### 2021-06-04 manuscript/motifs_vs_degree.tex
Initial Overleaf Import

\documentclass[12pt]{article} 
\usepackage{amsmath} 
\usepackage[dvips]{graphicx}
\usepackage{multirow} 
\usepackage{geometry} 
\usepackage{pdflscape}
\usepackage[labelfont=bf]{caption} 
\usepackage{setspace}
\usepackage[running]{lineno} 
% \usepackage[numbers,sort]{natbib}
\usepackage[round]{natbib} 
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{xr}

\newcommand{\methods}{\textit{Materials \& Methods}}
\newcommand{\SI}{\textit{Appendix}~}

\topmargin -1.5cm % 0.0cm 
\oddsidemargin 0.0cm % 0.2cm 
\textwidth 6.5in
\textheight 9.0in % 21cm
\footskip 1.0cm % 1.0cm

\usepackage{authblk}

\title{Do motif roles provide unique information about species risk of extinction?}

\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{3}$, Kate Wootton$^{2}$, Gyuri Barab\'{a}s$^{1}$, Anna Ekl\"{o}f$^{1}$} 
\date{\small$^1$Department of Theoretical Biology, Chemistry, and Physics\\ 
Link\"{o}ping University\\
Link\"{o}ping, Sweden\\
\medskip
\small$^2$ Swedish Agricultural University\\
Uppsala, Sweden\\
\medskip
\small$^3$Department of Agricultural Sciences\\
University of Helsinki\\
Helsinki, Finland\\
\medskip
$^\dagger$ Corresponding author:\\
}

\begin{document} 
\maketitle 
\raggedright

\setlength{\parindent}{15pt} 
\begin{spacing}{2.0}

% Maybe Proc B - approached Alyssa re: other motifs paper, flexible length limit.

\section*{Abstract}

    % In the Bayesian network framework, only the stable motifs are included.
    % We show that even in this case participation in different motifs is associated with different extinction risk.
    % Participation in different motifs is also correlated with in-degree and trophic level.
    % In a Bayesian network, trophic level has a very large effect on extinction risk, and likely explains a lot of what we observe for degree.
    % However, the trends we observe for motifs are not totally explained by the correlations of motifs with trophic level and degree, indicating that both simple and meso-scale descriptions of species roles provide unique information about extinction risk.

    % Omnivory may be stabilizing at high connectances because there is a low-TL "rescue prey" available to a high-risk, high-TL species. At lower connectances, perhaps higher-TL species encounter fewer disturbances and the rescue prey is less important?
    % In Lotka-Volterra systems, where top-down effects can also cause extinction, omnivory tends to be associated with less persistence. This suggests that what happens in real systems could depend on the dominant type of regulation.

    % - Add a panel showing in-degree at different TL's, like in Anna's fig. Or switch to persistence vs. in-deg for each TL in a panel.
\clearpage
\section*{Introduction}

    %\subsection*{1 Why are bottom-up effects important for understanding stability?}
     The plant community is the foundation upon which the myriad of depend for their survival \citep{}. Disturbances in the plant community have been shown to affect important ecosystem properties such as productivity \citep{}, nutrient cycling \citep{} and consumer diversity \citep{scherber2010bottom}.The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community promotes stability across all different levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}. Loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers, e.g. herbivores, predators, and parasitoids \citep{scherber2010bottom}.
     This has been shown in both theoretical work~\citep{} and empirical experiments ~\citep{} and results from the vertical propagation of disturbances through the network of interacting species.

     %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
    The study of how an initial decline or extinction leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, eklof2006species}. Traditionally, there are two main approaches for studying secondary extinctions. First, there are topological models, which are solely based on food web structure \citep{dunne2009cascading}. Here, initial extinctions only affect other species in the network once a consumer has lost all of its prey and therefore must go extinct. The second approach uses dynamical models, which explicitly simulate population dynamics using a system of differential equations \citep{binzer2011susceptibility}. Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network. These models therefore describe complex effects such as indirect interactions and are generally more realistic than topological models. However, dynamical models are parameter intensive, can only include a limited number of species, and simulations are time-consuming.

    A middle‐ground approach for simulating secondary extinctions are Bayesian networks \citep{Eklof2013}. In this framework, a consumer's probability of extinction depends on the fraction of resources lost and on a baseline probability of extinction which captures causes unrelated to the network structure (e.g., disease, stochastic extinction of small populations). Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
    As consumers do not affect the extinction probabilities of their resources, Bayesian networks can only capture bottom-up effects of disturbances within networks. However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
    This emphasizes the importance of bottom-up effects in food webs and makes Bayesian networks the ideal framework to use when studying them.

    %\subsection*{3 Global scale network properties are known to affect stability}
    The global structural properties of ecological networks have implications for their functioning \citep{petchey2002functional}, stability \citep{allesina2012stability} and robustness to disturbances \citep{dunne2002network, eklof2006species}. For example, more highly connected networks are more resistant to secondary extinctions \citep{dunne2002network, eklof2006species} and a modular organization where some groups of species are tightly interconnected with limited connections between groups can limit the propagation of disturbances \citep{}.
    %\subsection*{4 meso-scale nw propoerties also affect stability}
    Structural properties on the local scale, i.e. properties of single species, are also well-studied and there is consensus that species with high trophic levels, few prey (low in-degree), and low abundances are generally at higher risks of secondary extinction \citep{binzer2011susceptibility}. However, in-degree and trophic level are both relatively coarse descriptions of a species' role within its food-web~\citep{Cirtwill2018FoodWebs}. 
    For example, specialists may not necessarily have a high risk of extinction if it consumes a basal resource which is at low risk of being affected of disturbance. On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes few preys or prey which possess high risk of being affected by disturbance.

    %\subsection*{5 Motif participation may similarly affect species' particular extinction risk}
    One way to incorporate more nuance into descriptions of how species fit into their communities is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}. Importantly, these motif roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions.
    This means that two species with the same in-degrees and trophic levels can still have quite different motif roles if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}. This extended description of species structural roles may give additional information of extinction risk. 
    At the network level, the frequencies of different motifs is associated with community stability \citep{prill2005dynamic, bascompte2005simple}.
    Some motifs, such as the three-species chain, are likely to retain all species within them when modelled in isolation~\citep{Borrelli2015a} and over-represented in empirical food webs~\citep{Stouffer2007}.
    This suggests that some motifs may increase network stability by reducing the probability that species participating in them will go extinct.
    The plant community has been shown to be particularly relevant for how the motifs are distributed within a network. For example, a recent study in experimental grasslands showed that changes in the plant community, i.e. richness, induced shifts in the relative frequency of motifs in the food webs studied \cite{giling2019plant}.

    %\subsection{6. only 4 (stable) motifs can appear in Bayesian networks, so how do we expect them to affect stability in a bottom-up context?}
    Of the 13 three-species motifs that can occur in empirical food webs
    \citep{Milo2004}, only four do not include cycles (e.g., A eats B and B eats A or A eats B, B eats C, and C eats A). 
    As cycles must be removed from networks before calculating extinction probabilities in the Bayesian network framework, this study focuses on these four Bayesian-compatible motifs: the three-species chain, omnivory, apparent competition and direct competition motifs. 
    Despite this being a small subset of the motifs which can occur in empirical networks, our choice is justified by studies showing that these four motifs are over-represented in empirical networks and contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}.   
    When analyzing motifs within a Bayesian Network, it is important to keep in mind that only bottom-up effects are accounted for. Therefore, the implications of the four possible motifs are different than they would be in an empirical network.

    In particular, the lack of top-down effects removes the indirect effects from the two competition motifs. In the case of direct competition, two predators share a common resource. The predators will, within a Bayesian Network, not affect each other indirectly although both will be affected by any disturbance to their resource. Similarly, the two bottom-level species in the apparent competition motif will be independent of each other as there are no top-down effects of the shared consumer to link them.
    The difference between the direct competition and apparent competition motif is thus the number of bottom-level species, accounted for as prey items for a higher level species.
    A predator participating in many direct competition motifs is likely to have fewer prey items than a predator participating frequently in the apparent competition motif, and therefore a higher extinction risk.
    Note, however, that consumers can also appear in these motifs as prey to other consumers.  
    In that case, we would not expect the two motifs to have different effects on extinction risk.

    The structure of the omnivory motif and the three-species chain motif differ from the competition motifs by having an additional trophic level. Therefore, top-level species of these motifs might be more sensitive to disturbances since they can be affected by the loss of the basal resource in the focal motif or an increased probability of extinction of the middle species due to a disturbance to a basal resource not included in the focal motif.

       %\subsection*{Here we test...}
    Here we are interested in how disturbances at the basal level in food webs propagate in the network and whether a consumer's extinction risk depends on 1) the motif profile of the full food web and 2) the motifs in which a focal species participates. In particular, we want to understand whether the information provided by motifs can be explained by the classical global and local food web metrics such as network connectance and species trophic level, or whether motifs provide unique insight into bottom-up extinction.

   % \subsection*{Other, simpler measures of species' roles can also affect extinction risk, and may correlate with motif participation}

    	%Species with high trophic levels may be more likely to go extinct since they depend upon the stability of relatively long food chains~\citep{}. 
        %For the same reasons, the number of motifs a species participates in will also tend to be larger in larger and more-connected networks.
    	%Likewise, the range of possible trophic levels increases with network size, although higher connectance may keep trophic levels low as long as the shortest food chain carries more weight (as in prey-averaged and shortest path trophic levels). 
        %A species' trophic level in turn may affects a species' motif role by restricting the set of positions in which it can appear~\cite{Cirtwill2018EcolLett}. 
    	%To unravel to what extent species motif roles gives additional information on species' probabilities of going extinct, we therefore need to take network structure and other role definitions into account.

\section*{Methods}

	\subsection*{Initial network construction}

		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}. 
		To capture a range of plausible network architectures, we simulated networks with sizes ranging from 50 to 100 species (in steps of 10) and connectances ranging from 0.02 to 0.18 (in steps of 0.04). 
		For full details, see~\citet{Cirtwill2021_inprep}.
        We then filtered these simulated networks to remove any with unreasonable structures.
		Specifically, we removed any network containing disconnected components (species or groups of species not connected to the rest of the network) or any network where the shortest path between a consumer and the closest basal resource was \textgreater6 (as these high trophic levels are very uncommon in empirical food webs ~\citep{}).
		New networks were simulated to replace any removed networks, and this process was repeated until we obtained 100 suitable networks in each combination of size and connectance.

		A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}. 
		Therefore, all persistence calculations follow a strict bottom-up routine beginning by determining the status of the primary producers (who do not depend on other species) and continuing on to primary consumers (who depend only on primary producers), and so on up the network.

		While the niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}. We therefore removed any cycles within each network.
		To render networks acyclic, we followed~\citet{Allesina2009} and broke cycles by removing links which do not contribute to the robustness of the food web.
		This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009}.
		These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
		Next, we order the species from lowest to highest trophic level using a topological sorting routine following \citep{Tarjan1972, Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order. 
        After these steps, calculations of persistence probabilities can begin.

	\subsection*{Calculating persistences}

		First, we assign all species a baseline probability of extinction ($\pi_{base}$) -- the risk of going extinct given factors not related to the food web itself. 
		For consumer species, this also applies to extinction despite all resources being present. 
		Such events can for example reflect diseases or stochastic extinctions of small populations.

		Additionally, we analyze a set of scenarios where primary producers get a higher baseline probability of extinction ($\pi_{disturbed}$). 
		This reflects a threat to primary producers such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
		$\pi_{disturbed}$ ranged between $0.1-0.5$, in steps of $0.08$. The highest disturbance level, $\pi_{disturbed} = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
		Consumer species retained $\pi_{base}=0.1$ in all cases.

		As well as the intrinsic baseline probability of extinction, extinction risk for consumers is affected by access to resources. 
		When all resources are extant, the probability of extinction for a consumer $i$, $P(\lnot i)$, is equal to the baseline probability; $P(\lnot i|X_{1},...,X_{n}) = \pi_{base}$. 
		When all resources are extinct, the consumer will go extinct; $P(\lnot i|\lnot X_{1},...,\lnot X_{n})=1$.

		Calculating $P(i)$ or $P(\lnot i)$ is less straightforward when a consumer $i$ loses a fraction $k/n = f$ of its prey. 
		Plausible ecological responses to losing some prey include i) a topological response, where the consumer's probability of extinction is constant unless all resources are lost, ii) a linear response, in which the extinction probability increases linearly with resources loss, and iii) several nonlinear responses where the probability of extinction increases non-linearly with the fraction resource lost. 
		The extinction probability will differ depending on the response function. 
		We use a sigmoid non-linear functional response of consumers to the loss of resources ($\alpha >1, \beta >1$) which has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}. 
		We use a cumulative density function of a beta distribution for this purpose:
		\begin{equation}
		\label{betafunc}
		P(\lnot i|f) = \pi_{i} + (1 - \pi_{i}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}
		\end{equation}

		We calculate the fraction of resources lost ($f$) for each consumer $i$, to compute $P(\lnot i|f)$. 
		Persistence is determined by comparing $P(\lnot i|f)$ with a randomly drawn number from a uniform distribution between 0-1. 
		If this random number is lower than $P(\lnot i|f)$, the species is considered extinct. 
		The species' marginal probability of persistence is simply $P(i) = 1-P(\lnot i)$.
		In this manner we calculate persistence probabilities for all species in the network, starting from the bottom and working our way up. 
		Probabilities of extinction for resource species are used when calculating probabilities for higher level species. 
		Some species will be extinct due to primary extinctions (stochastic extinction with all resources left), while others will experience secondary extinctions; as a response to primary extinctions (partial or total loss of resources). 
		As this method is probabilistic, repeating the simulation will yield different results. 
		To minimize the effect of this variation we repeat these calculations 100 000 times per web, with unique random draws when calculating $P(i)$.
		The final persistence for each species is the mean over these runs.

	\subsection*{Defining motif profiles and species roles}

        As a first measure of the meso-scale structure, we calculated the `motif profile' of each network.
        This is, in our case, a four-dimensional vector of the number of each three-species motif in the network.
        To separate differences in motif structure from differences in network size, we normalized motif profiles by dividing the count of each motif by the total count of all motifs.
        The motif profile for each network therefore describes the proportions of each motif and has a sum of excatly one.

        We then describe meso-scale structure at the species level by defining each species' motif participation. [[Should stop calling this motif roles]]
		Finally, we defined species' motif participation following~\citet{Stouffer2012}. 
		For each network, we identified each group of three interacting species and assigned it to the corresponding motif. 
		A species' motif participation is a vector of the number of times it appears in each motif.
		We calculated species' motif participations using the Python package \emph{pymfinder}~\citep{pymfinder}.
        As with motif profiles of networks, we normalized participation vectors by dividing each count by the total number of motifs the species appears in.
        Note, however, that this normalization does \emph{not} control for differences due to degree \emph{per se} as high-degree species also tend to participate in different types of motifs than low-degree species~\citep{Cirtwill2021_inprep}.

		We began by calculating species' in-degrees (numbers of prey).
		As the Bayesian network approach only accounts  bottom-up effects, out-degree (number of predators) is irrelevant to extinction risk.
		Note, however, that out-degree may also be important in topological simulations which include both top-down and bottom-up effects.

        Next, we calculated species' trophic levels. 
		Among the many definitions of trophic level available, we chose shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}. 
		Basal resources have an STL of one, any species consuming a basal resource has an STL of two, etc.
        STL is therefore a measure of the minimum number of steps away a species is from any disturbance to the base of a food web.

	\subsection*{Statistical analysis}

        \subsubsection*{How does mean persistence vary depending on network properties?}

            Both species richness (S) and connectance (C) are known to influence persistence in both simulation studies and empirical food webs.
            In addition, increasing the probability of extinction for basal species ($\pi_{disturbed}$ above) is intended to decrease persistence of species at higher trophic levels.
            Although these effects are not the main focus of our manuscript, we must establish these background effects before measuring the effects of meso-scale structure on persistence.
            The persistence of basal species in our simulations is determined only by their baseline probability, being either $\pi_{base} = 0.1$ or higher, dependent on $\pi_{disturbed}$.
            We therefore considered only non-basal species (i.e., those with at least one prey) throughout this manuscript.

            To test for effects of global network properties, we fit a linear regression relating consumer species' probabilities of persistence to network size, connectance, the level of disturbance to basal resources, and all interactions between them. 
            All predictors were centered and scaled before fitting the model. 
            As the three-way interaction was significant, we did not simplify the model.
            All models were fit using the R~\citep{R} base function `lm'.

        \subsubsection*{How do network motif profiles vary with global structure?}

            As well as affecting persistence directly, global network structure (i.e., size and connectance) may affect the meso-scale structure of the network.
            This is especially likely since we include only networks which can retain all initial species after 1000 rounds of simulated Lotka-Volterra population dynamics~\citep{Cirtwill2021_inprep}:  some meso-scale structures might be more likely to allow all species to persist than others in networks with different size and connectance constraints.
            We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
            Significance was calculated based on 9999 permutations.
            PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
            To test whether this applies in our case, we fit an ANOVA of the dispersion of motif profiles about the centroid for each combination of network size and connectance. 
            We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
            To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif to network size, connectance, and their interaction.
            We fit each regression using the R~\citep{R} base function `lm'.

        \subsubsection*{How does the persistence vary with motif frequency in the web?}

            To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
            In case high or low mean persistences are associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable). 
            We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
            All statistical tests were conducted as described above.

            To supplement these overall tests, we also fit four regressions testing whether persistence had a linear relationship to the proportion of a network profile made up of each motif, and whether this varied with the level of disturbance.
            Each regression included the proportion of a motif, disturbance, and their interaction as predictors.
            All regressions were fit using the R~\citep{R} base function `lm'.

        \subsubsection*{How does species persistence vary with motif participation?}

            We are interested in whether having a greater proportion of a species' role be made up of a certain motif (e.g., direct competition) might increase persistence (see \emph{Appendix SXX} for relationships between persistence and counts of motifs).
            To test this, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
            These models included the proportion of the species' role made up by the focal motif, the level of disturbance to basal species, and the interaction between the two.
            We centered and scaled the predictors before fitting the LMMs.
            The models also included a random intercept for the interaction between species richness and connectance (equations.
            All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
            We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.

            The formula for these LMMs are:

            \begin{equation}
                \Psi_{ik} \approx \rho_{i} + \pi_{disturbed_k} + \rho_{i}\pi_{disturbed_k} +
                S:C_{i} ,
                \label{propreq}
            \end{equation}

            where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.

            We were also interested in whether these general trends were consistent across networks with different global properties and different levels of disturbance. 
            To test this, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources.
            These regressions were fit using the R~\citep{R} base function `lm'.
            We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.

        \subsubsection*{How does persistence vary with degree and trophic level?}

            Species' motif roles are often related to their degrees and trophic levels.
            To understand how this may have affected the results above, we first fit two linear mixed effects regressions relating persistence to degree or trophic level, $basal_p$, and their interaction.
            As above, we centered and scaled the predictors.
            We also included a random effect of the interaction between network size and connectance in order to account for the differences in persistence between different networks. 
            These regressions provide context about how persistence varies with simpler measures of species' roles within networks.

        \subsubsection*{How does motif participation vary with degree and trophic level (i.e., do simpler measures completely explain motif results)?}

            A further complication when trying to understand whether relationships between motifs and probability of extinction depend upon underlying relationships between degree or trophic level and extinction 
            We also fit linear mixed effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.

        \subsubsection*{How does persistence vary with motif positions?}

            As one more level of detail, we can consider the frequency with which a species appears in each \emph{position} in each motif. 
            Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.
            The full set of these positions, across all motifs, is the species' \emph{motif role}~\citep{Stouffer2010,Cirtwill2017}.
            As with motif participation, we normalized motif roles by dividing the number of times a species appeared in each position by the total number of times it appeared in all positions~\citep{Baker2015,Cirtwill2017}.

            [[Could add a PERMANOVA to show that species' roles overal affect their persistence. Haven't done so yet.]]

            To test whether participating in different motif positions was related to a species' probability of persistence, we fit similar models to those for motif participation.
            For each position, we fit a linear mixed effect model relating persistence to the frequency of the position in a species' motif role, disturbance, and their interaction.
            To account for differences in persistence across networks of different size and connectance, we included a random effect of the interaction between size and connectance.
            All regressions were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.

        \subsubsection*{How do motif roles vary with other network properties?}

            As with network motif profiles and motif participation, species' motif roles may vary with network size and connectance and with the species' in degree and trophic level.
            To identify these relationships, we fit two linear regressions for each motif position.
            The first related the frequency of the position in species' roles to network size, connectance, and their interaction.
            The second related the frequency of the position to in-degree, STL, and their interaction.
            All regressions were fit using the R~\citep{R} base function `lm'.

\section*{Results}

    \subsection*{How does mean persistence vary with global network properties?}

        Network average persistence of consumers naturally decrease with increasing level of disturbance of basal resources (Table S1, \emph{Appendix SA}). Network size and connectance also affect persistence, but to a much smaller extent (Figure~\ref{fig:lm_CS}). Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance: at low disturbance small and highly-connected webs show the lowest persistence, whereas at high disturbance level large and highly connected webs show the lowest persistence (Fig. S1, \emph{Appendix SA}).

        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.95\textwidth]{figures/persistence_vs_SC_lm.eps}
            \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while network size, connectance, and their interaction had smaller effects. \textbf{A)} Persistence decreased slightly with increasing network size at high levels of disturbance, increased slightly with increasing network size at low levels of disturbance, and did not vary noticeably with network size at moderate levels of disturbance. Networks with lower connectance (dotted line) had slightly higher persistence than networks with moderate (solid line) or high (dashed line) connectance.
            \textbf{B)} Persistence decreased with increasing connectance for all levels of disturbance and network size. Network size (line types) did not strongly affect these relationships.}
            \label{fig:lm_CS}
        \end{figure}

    \subsection*{How do motif profiles vary with global network properties?}

        A network's normalized motif profile (i.e., proportions of the total made up by each motif)
        % ; see \emph{Appendix SX} for non-normalized motif profiles) 
        is related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
        Unequal variation in motif profiles across levels of size, connectance, and their interaction may, however, mean that these PERMANOVA results are false positives (Fig. S2, \emph{Appendix SB}).

        The proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}). 
        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}). The proportion of direct competition, however, decreased with increasing connectance in small networks but showed very little change in large webs.

        \begin{figure}[h!]
            \centering
            \includegraphics[width=.75\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
            \caption{The proportions of a network's motif profile made up by the four motifs which can appear in acyclic networks vary with the size and connectance of the network. More-connected networks tend to contain relatively more omnivory motifs and fewer apparent competition motifs. The proportions of direct competition and three-species chain motifs were less variable.}
            \label{motif_proportion_lms}
        \end{figure}

    \subsection*{How does mean persistence vary with network motif profile?}

        The motif profile of a network does relate to the mean persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001). Mean persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}). Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif. These trends hold for all levels of disturbance (Table S3, \emph{Appendix SC}). Dispersion of motif profiles varies between networks with different levels of persistence ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001). For example, networks with higher persistence have more variable motif profiles ($\beta$=0.183, $p$\textless0.001).

        \begin{figure}
            \centering
            \includegraphics[height=.5\textheight]{figures/persistence_motif_profiles.eps}
            \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network. Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportions of the other three motifs. This was true for all levels of disturbance to basal resources; the interactions between motif profiles and disturbance were significant but small. Relationships are shown for the observed range of proportions for each motif.}      
            \label{fig:motif_profile_persistence}
        \end{figure}

    \subsection*{How does species persistence vary with motif participation?}

        % See \emph{Appendix SY} for relationships between persistence and non-normalized motif participation.

        \subsubsection*{Interactions between motif participation and disturbance}

            At low levels of disturbance, species' persistence increased with increasing proportions of the omnivory and direct competition motifs. Conversely, a species' persistence decreases with increasing proportion of the apparent competition motif. Finally, a species persistance is negatively but not significantly related to the proportion of the three-species chain (Figure~\ref{fig:prop_lmer_all}).

            At high levels of disturbance, a species' persistence decreases with increasing proportion of the three-species chain, but increases with increasing proportion of all other motifs (Figure~\ref{fig:prop_lmer_all}).
            This means that increased disturbance strengthen the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table S4, \emph{Appendix SD}).
            The effect of omnivory on persistence is still positive but weaker at higher levels of disturbance, and the effect of direct competition are largely unaffected by disturbance level.

            \begin{figure}[h!]
                \centering
                \includegraphics[width=0.85\textwidth]{figures/prop_lmer_allCS.pdf}
                \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis). The effect of participating in each motif is based on linear mixed-effect models following Equation~\ref{propreq}, except that models for each  disturbance level are fitted separately. The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance). Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
                \label{fig:prop_lmer_all}
            \end{figure}

        \clearpage

        \subsubsection*{Variability of motif effects changes with network properties}

            By fitting linear regressions of persistence against motif proportions for the species for each network separately, we can observe how the relationships between motifs and persistence change with the level of disturbance and global network structure. Network size had a much smaller effect than connectance (Fig. S3; \emph{Appendix SE}); we therefore focus on connectance here. 
            Networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}). 
            However, there was an interaction between disturbance and connectance for the direct competition and omnivory motifs.

            At low and medium connectances there is little change in the proportion of positive slopes (\textgreater80\%) for the different levels of disturbance. 
            At high connectances, however, the proportion of positive slopes is lower overall and decreases with increasing disturbance. 
            This means that at low levels of connectance, an increased proportion of the direct competition motif increases persistence in most networks, while the opposite is true with high connectance. 
            The trend is reversed for the omnivory motif.

        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.pdf}
            \caption{Density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. The slopes are derived from standard linear regression. Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right). Rows show various levels of connectance. The dotted, vertical line indicate zero on the x-axis. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence. The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend). }
            \label{fig:density_prop_C}
        \end{figure}

    \subsection*{How does persistence vary with degree and trophic level?}

        Persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}). Increased trophic level is associated with decreasing persistence, although the rate of this decline is slower at higher levels of disturbance (Table S5, \emph{Appendix SF}).
        The interaction between in-degree and disturbance is even stronger, such as having a large number of prey is associated with higher persistence in the no disturbance scenario, but lower persistence at high disturbance.

        % Anna notes (just for our (mine) understanding): 
        % When looking at the two figures of degree vs persistence and STL vs persistence - my conclusion is that species at lower trophic levels have the highest in-degree. Double check - yes, that’s true.
        % Lower TL are all well at low disturbances, and when disturbance increase, persistence decrease (looking at fig degree vs persistence, right end of the line).
        % Higher TL experience the same thing, although to a slightly smaller extent (looking at the left end of the lines in fig degree vs persistence and right end of lines in fig STL vs persistence). 
        % The slope of the line in fig degree vs persistence is not really relevant, since it's kind of based on the response of different TL level (which makes perfect sense in the BN framework)
        % weird thing though - at high levels of disturbance, lower TL seems more affected than higher TL in the persistence vs degree figure (since low degree belongs to higher TL, high degree to lower TL). This does not correspond with fig persistence vs STL. HOWEVER - not only high TL have a low degree, lower TL do as well, probably affecting the overall result. Only low TL have high degree - but the opposite is not true. Probably explains it

    \subsection*{Do simpler measures explain effects of motifs?}

        Motif participation varied with web size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
        Species in all networks appeared most frequently in the apparent competition motif, but the frequency of the apparent competition motif decreased with increasing network size and connectance.
        The frequency of the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size. 
        Species' motif roles also vary with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
        Species with higher in-degree tend to have higher proportions of their motif role made up by omnivory. Proportions of the other three motifs tend to decline with increasing degree.

        Species with higher trophic levels tend to have higher proportions of apparent competition and three-species chains.

            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
                \caption{As with motifs, persistence varied with simpler measures of species roles in networks. Moreover, motif frequencies correlated with in-degree and trophic level. \textbf{A-B)} Persistence increased with increasing in-degree for low levels of disturbance to basal resources but decreased with increasing degree for high levels of disturbance to basal resources.
                Persistence declined with inceasing STL for all levels of disturbance to basal resources.
                \textbf{C-D)} Counts of all motifs increased with increasing in-degree and decreased with increasing STL. The proportion of omnivory increased with increasing degree while all other proportions decreased. The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.}
                \label{fig:motifs_vs_TL_and_deg}
            \end{figure}

    \subsection*{How does persistence vary with motif positions?}

        A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance.
        \begin{table}[h!]
            \caption{Coefficients in a series of regressions relating probability of persistence to the frequency of a position in a species' motif role, disturbance, and their interaction. For ease of comparison between motifs, we did not center or scale predictors. Note, however, that some motifs and positions have higher frequencies than others.}
            \label{tab:persistence_vs_positions}
            \centering
            \begin{tabular}{l | c c c c}
            Position	& Intercept	& Disturbance	& Position	& Interaction \\
            \hline
            Chain: top	& 0.880 &	-0.985 & -0.00544 & -0.332 \\
            Chain: bottom	& 0.883 &	-1.07 &	-0.0510 &	0.647 \\
            Chain: middle	& 0.862 & -1.00 &	0.246 &	-0.244 \\
            \hline
            Dir. Comp.: top & 0.872 &	-0.995	& 0.0549 &	-0.178 \\
            Dir. Comp.: bottom & 0.886 &	-1.07 &	-0.113 &	0.939 \\
            \hline
            Omnivory: top	& 0.872 &	-0.984 &	0.180 &	-0.818 \\
            Omnivory: middle & 0.873 & -1.00 &	0.119 &	-0.324 \\
            Omnivory: bottom & 0.882 & -1.06 &	-0.0449 &	0.845 \\
            \hline
            App. Comp.: top & 0.865 &	-0.966 &	0.140 &	-0.520 \\
            App. Comp.: bottom & 0.910 &	-1.12 &	-0.100 &	0.323 \\
            \hline
            \end{tabular}
        \end{table}

        % Organizing results by motif - keep this OR by no. prey
        In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
        Peristence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
        Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.

        In the three-species chain, persistence decreased as the frequency of the top position increased at all levels of disturbance.
        Conversely, persistence increased with increasing frequency of the middle position at all levels of disturbance.
        Persistence decreased slightly with increasing frequency of the bottom position at low levels of disturbance but increased at high levels of disturbance.

        In the apparent competition motif, persistence decreased slightly with increasing frequencies of the bottom position (one predator, no prey within the motif) at low disturbance but increased slightly at high disturbance.
        Persistence increased with increasing frequency of the top position (two prey, no predator within the motif) at low disturbance but decreased at high disturbance.

        In the direct competition motif, persistence varied little with the frequency of the bottom position (two predators, no prey within the motif) at low disturbance but increased with the frequency of this position at high levels of disturbance.
        Persistence increased slightly with the frequency of the top position (one prey, no predators within the motif) at low levels of disturbance and decreased slightly at high levels of disturbance.

        \begin{figure}[h!]
        \includegraphics[width=.95\textwidth]{figures/persistence_positions_bymotif.eps}
        \label{figs:persistence_position_motifs}
        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role. Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common. Line colour indicates the probability of extinction for basal resources. Line style indicates the number of prey within the motif for each position. Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.}
        \end{figure}

        % Organizing by number of prey
        Across the four motifs we consider, positions with no prey within the motif tended to show weak relationships to persistence at low levels of disturbance and be associated with increased persistence at high levels of disturbance (Fig.~\ref{fig:persistence_positions_prey}).
        These trends were strongest for direct competition and omnivory and weakest for apparent competition.
        The two positions with stronger effects indicate a species with two predators within the motif, while the weaker effects are for motifs with one predator within the motif.

        Positions with one prey within the motif were more variable.
        The top position in the direct competition motif and middle position in the omnivory motif (both of which have one prey) were associated with increased persistence at low levels of disturbance and lower persistence at higher levels of disturbance, but these effects were relatively weak.
        The top and middle positions of the three-species chain both include a single prey; the middle position was associated with increased persistence at all levels of disturbance while the top position was associated with decreased persistence at all levels of disturbance.

        Positions with two prey in the apparent competition and omnivory motifs showed very similar patterns.
        At low levels of disturbance, persistence increased with the frequency of either position.
        At high levels of disturbance, persistence decreased as the frequency of either motif increased.
        This echoes the pattern for in-degree.

        \begin{figure}
            \centering
            \includegraphics[width=.95\textwidth]{figures/persistence_positions_byprey.eps}
            \caption{A species' probability of persistence varied with the frequency of different motif positions in its role. Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common. Line colour indicates the motif containing a position. Line style indicates the number of prey within the motif for each position. Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations. The top set of lines corresponds to a probability of extinction for basal resources of 0.1. The middle set of lines corresponds to a probability of extinction for basal resources of 0.26, and the lowest set of lines corresponds to a probability of extinction for basal resources of 0.5.}
            \label{fig:persistence_positions_prey}
        \end{figure}

    \subsection*{How do motif positions vary with other network properties?}

        \subsubsection*{Degree and trophic level}

            The top position in the omnivory motif increased with increasing degree across all trophic levels.
            The frequency of the middle position increased with the interaction of degree and trophic level.
            The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).

            The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
            The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.

            The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
            The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).

            The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
            The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
            The top position in the three-species chain increased with trophic level, degree, and their interaction.

        \subsubsection*{Network size and connectance}

            % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
            The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance.
            The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.

            The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
            The frequency of the bottom position in the direct competition motif was low in all networks.

            The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
            Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
            None of the positions in the three-species chain varied strongly with network size or connectance.

\section*{Discussion}
% * Apparent competition change from bad to good
% * Difference between 3-sp chain and omnivory, the importance of the rescue prey in omnivory (in dynamical sims omni was mostly bad, 3-sp chain good). Omni link destabilizing if strong, stabilizing if weak (other studies). In BN just a link as important as other links
% * Omnivory in fig 5 - high connectance, always good, not as good with lower levels of C
% * Extra prey in omnivory changes from being a security to a liability (with increasing disturbance?)
% * ADD REFS - high disturbance scenario is analogous to decreasing plant diversity

% Results summary - network level (following Anna)
We analysed how network properties at a variety of scales influence consumers' probabilities of extinction.
First, we considered how global-scale and meso-scale structures were related to the average persistence within a network.
Consistent with earlier work, networks with higher connectance or species richness tend to have lower persistences~\citep{}.
In our simulations, however, these effects were relatively small.
In contrast, and consistent with other studies of motif profiles, changes in the motif profiles of networks (i.e., the frequencies of different motifs across the whole food web) had relatively strong relationships to the mean persistence of species within the web~\citep{prill2005dynamic, bascompte2005simple}.
As changing connectance (and, to a lesser extent) network size 
affected network motif profiles, it appears that global-scale properties have stronger effects on persistence by changing meso-scale properties than directly.
In all cases, the effect of increasing disturbance to basal resources had a much larger effect on consumer persistences than did network structure.
This is a consequence of the bottom-up calculations done to receive the final persistence. The importance of the basal level for network robustness has been demonstrated in numerous studies \citep{scherber2010bottom}, as well as the sensitivity of high trophic levels (REF).

% Network-level motif profiles, persistence, and relationships to S and C.
In general, persistence was highest in networks with high frequencies of the apparent competition motif (two prey consumed by the same predator) and lowest in networks with high frequencies of the omnivory motif (a three-species food chain in which the top species also consumes the bottom species) while the direct competition and three-species chain motifs had smaller effects.
The negative relationship between omnivory and mean persistence is consistent with modelling work showing that an omnivory motif is less likely to retain all three species after a disturbance than the other motifs we consider~
\citep{Borrelli2015a}.
Motif profiles with high apparent competition and low omnivory were also more likely in less-connected food webs, which had higher mean persistence than highly-connected food webs.
Despite these general trends, networks with more variable motif profiles had higher mean persistence, suggesting that many different network structures support high probabilities of persistence. 
High connectance might reduce mean persistence by limiting the number of possible meso-scale network structures.
In particular, it is reasonable to expect higher frequencies of the omnivory motif at high connectances since this motif contains more links than the other three-species motifs we consider.
% The relationship between high connectance and frequency of omnivory motif should perhaps come earlier. 
% Close with something about how more links allow disturbance to propagate - borrow from connectance literature.

As well as the mean persistence of all consumers, we also considered the persistence of each consumer in a network.
While the persistence of a particular species is likely to be affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it will also depend on smaller-scale measures of the species' place in the network.
In particular, a species' motif participation (the frequency with which it appears in each of the four possible motifs) was significantly related to its probability of persistence.
Specifically, a high proportion of the direct competition or omnivory motifs increase the probability of persistence, while a high proportion of the three-species chain motif decreases probability of persistence.
%These trends held at any level of disturbance we considered, although at high levels of disturbance the positive effect of omnivory was reduced and the negative effect of three-species chains was increased.
%The effect of the apparent competition motif, however, varied with disturbance.
%A high proportion of the apparent competition motif reduces persistence at low levels of disturbance but increases persistence at high levels of disturbance.

% Two notable results: 1) omnivory bad at network level, good at species level, 2) apparent competition depends on disturbance
Combining these results with our network-level findings, we see that while being in a network with high frequencies of omnivory decreases persistence, participating in many of these omnivory motifs increases persistence at the species level.
At the network level, a higher frequency of the omnivory motif leads to more pathways by which disturbances to basal resources can propogate and cause extinctions of consumers.
If extinctions of basal resources are rare, these multiple pathways will tend to protect consumers by providing alternative channels of energy.
If extinctions of basal resources are common, however, these multiple pathways cause a negative impact as they will transmit more disturbances to the consumer.
To understand the species-level effect, we consider the relationships between stability and each position within the motif.
Appearing frequently in the top or middle positions of the omnivory motif is associated with higher probabilities of persistence at low levels of disturbance, while appearing frequently in the bottom position is associated with higher probabilities of persistence at high levels of disturbance.
This is consistent with the logic outlined for the network-level effect: when extinction of basal species is rare, consumers at the top or middle of omnivory motifs benefit from multiple energy channels but when extinction is common these multiple channels mean that the consumer is likely to be affected by multiple disturbances.
The bottom position shows the reverse effect: at low levels of disturbance, this position is nearly neutral. 
At high levels of disturbance, however, being low down in the motif (and obtaining energy by appearing as a predator in other motifs) means that the species is less likely to be affected by multiple lower-level extinctions.
The offsetting of effects of different positions within the omivory motif may explain why the motif as a whole is associated with greater probabilities of persistence across all levels of disturbance we considered.

% % Might be able to add some of this paragraph to the one above or below (especially the refs).
% The divergent effect on persistence of either having a high proportion of the three-species chain or omnivory motif can, within our Bayesian Network framework, be explained by the structure of the food web.
% When participating in a food chain motif---three-species chain or omnivory---the top species of this structure will have a higher susceptibility since disturbances amplify through the network. In our results, this is in particular evident for the three-species chain motif at higher levels of disturbance, but this also applies to species with a high proportion of the omnivory motif. While the relationship between omnivory and persistence is strongly positive at low bottom-level disturbance, the trend is reduced at high levels of disturbance. 
% This is because species participating to a high extent in omnivory motifs share the sensitivity of species participating to a high extent in three-species chain motif, and are similarly negatively affected by bottom-level disturbances. However, the top species in an omnivory motif also have the benefit of an additional prey species, located further down in the food web. This prey item can assist as a buffer, decreasing the negative effect of disturbance on the consumer species. Clearly, this additional prey is of importance, since it is more beneficial with a higher proportion of the omnivory motif for an individual species than a high proportion of the three-species chain motif. 
% Thus, the observed benefit of a high proportion of the omnivory motif can be compared to omnivores in a network feeding adaptively. Studies have demonstrated that adaptive omnivory can increase the permanence of three-species chains \citep{Fagan1997, KrivanDiehl2005, AbramsFung2010}.
% Although species in our model cannot swap prey dependent on bottom-level disturbances or lower resource densities, a species with a high proportion of omnivory will in a sense spread its risk between prey items at different trophic levels.
% Additionally, the effect of the omnivory motif on persistence was closely connected to connectance. In highly connected networks, a high proportion of the omnivory motif strongly contributed to higher persistence. This was not the case in less connected networks. 
% In a highly connected network, a disturbance could potentially spread to a higher extent than in a less connected network. 
% Therefore, in a highly connected web where disturbances have more ways to move, the lower level prey item in the omnivory motif may be of even higher importance assisting as a buffer, decreasing the negative effect of disturbance for the consumer species. In a less connected network, this additional prey species is of less importance.

% Broader context of omnivory is probably better after we introduce the species-level results. 
The relationship between omnivory and stability is subject to a longstanding debate in ecology. Omnivory has been shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, MonteiroFaria2016, Kratinaetal2012}. 
For example, omnivorous interactions may not be stable in isolation, but instead stabilized by interactions with other species when embedded in a larger food web \citep{Kratinaetal2012}. Additionally, the possibility of omnivores to feed adaptively from different trophic levels can spread the risks, for example when the prey level are disturbed \citep{Fagan1997}.
This might be why we have a divergent pattern between a full food web's motif profile versus each single species motif profile.

Our result contrasts with \citep{Cirtwill2021_inprep}, where a high proportion of the omnivory motif resulted in a shorter time to extinction, while the other three stable motifs were associated with longer times to extinction. However, the study by Cirtwill and Wootton used a different simulation approach, with body-mass structured vertebrate food webs and a dynamic simulation model accounting for top-down effects. 
    %In previous studies of the effect of omnivory motifs, the strength of the omnivorous link can potentially be important. If this link is strong, it is shown to be destabilizing (REF), while stabilizing if weaker (REF). In our Bayesian Network setting, this link is indistinguishable from any other link in the food web which partly may explain why omnivory results in increased persistence of consumer species.

    %These results can be explained by the structure of each motif. When a predator has a high proportion of the apparent competition motif, it tends to have more prey items. When the bottom-level is heavily disturbed, this becomes beneficial since if some prey items are lost, other might still exist. [but why is it bad to have a high proportion of app comp motif if disturbances are low?] 
    %Participating to a high extent in the direct competition motif means that a predator participates in many two-level food chains. It does not necessarily mean that the predator has few prey items, since it may participate in many direct competition motif. [why always good with high prop of direct comp motif? ]

    % % We can move this to the SI or delete entirely.
    % Although we performed analysis of both the effect of the absolute counts of motifs and the normalized motif role on persistence, we here focus on the normalized motif role.
    % This is because, the consistency of the relationships between persistence and the number of times a species appears in any of the possible motifs suggests that the observed effects of non-normalized motif roles are actually due to degree.
    % Species with higher degrees will with high probability appear in more motifs than species with low degrees.
    % It is possible for a species with low degree (few direct interaction partners) to participate in many motifs if it interacts with high-degree species (thereby obtaining many indirect interaction partners), but such a situation is fairly unlikely.
    % Because of this overlap between degree and absolute numbers of motifs, and because degree shows a similar trend to the motif counts [[to be confirmed]], we recommend that researchers consider degree and/or normalized motifs rather than absolute motifs.
    % When we consider the proportion of a species' role made up by each motif, the influence of degree is largely removed.
    % Given the similarly high $R^2$ values between the models using proportions of motifs and absolute counts, it does not appear that we lose much explanatory power by removing degree. 
    % Therefore, our main analysis focus on the effect of the normalized motif role on persistence.

[[Comparison to degree, tl?]]

    % This paragraph explains why motif trends aren't just due to TL
    Two of the focal motifs, omnivory and the three-species chain, include three trophic levels while apparent and direct competition only require two.
    However, the effects of motifs on persistence cannot be reduced to the number of trophic levels in each motif.
    The two predators in a direct competition motif or the two prey in an apparent competition motif are not necessarily at the same trophic level. This is demonstrated by the increasing participation in apparent competition and chain motifs and decreasing participation in direct competition and omnivory motifs with increasing trophic level (i.e., 2-TL and 3-TL motifs do not show consistent trends).
    In fact, a prey in a direct competition motif can have a higher trophic level than its predator.
    For example, blueberries and salmon consumed by a bear make an apparent competition motif where the salmon, a piscivorous fish, has a higher trophic level than the blueberries or the omnivorous bear.
    This means that motif participation and trophic level give complementary descriptions of a species' place in the network and explains why the simple decrease in participation with increasing trophic level does not explain the more complicated relationships between motif participation and persistence.

    In-degree (number of resources) may explain more of the relationships between motif participation and persistence.
    For example, higher proportions of the omnivory motif in a species' participation vector were associated with higher persistence for that species, and participation in omnivory significantly increased with increasing in-degree.
    Moreover, the relationship between degree and persistence  showed a strong interaction with disturbance, similar to the varying relationship between apparent competition and disturbance. 
    The shift from positive relationships between persistence and apparent competition at high levels of disturbance to negative relationships between persistence and apparent competition at low levels of disturbance could be partially explained by the negative relationship between apparent competition and in-degree.
    However, participation in the three-species chain also significantly decreased with increasing degree while participation in omnivory increased, and neither of these motifs shows a similar interaction with disturbance.
    As with trophic level, motifs provide some unique information not provided by in-degree (and vice versa). 
    Both ways of describing a species' position in the network show significant relationships to consumers' persistences after a disturbance and, because these relationships are not the same, likely capture different mechanisms leading to greater or lesser extinction risk.

% [[When can motifs still be useful predictors and add insight?]]
%     Our results suggest that some motifs are associated with higher persistence than others but that this depends on the level of bottom-up disturbance to the food web.
%     Detailed tests of how motifs affect network stability and species persistence ideally require the ability to manipulate the motif structure of a network directly (i.e., obtaining many replicate networks with set numbers of each motif). 
%     Current methods of generating simulated ecological networks (e.g., the niche and cascade models) do not incorporate information on motifs.
%     Researchers interested in relationships between motif structure and persistence must therefore generate large numbers of networks and hope that they contain the necessary variety of motifs.
%     This shotgun approach to simulating different distributions of motifs within networks makes it difficult to investigate the precise effects of different numbers of, for example, the three-species chain motif, whether it occurs high or low in the network, and similar questions.
%     Fortunately, software is currently in development to allow researchers to simulate networks with particular motif profiles.
%     We heartily encourage researchers interested in the effects of motifs on persistence to adopt these new methods as they become available.

\bibliographystyle{ecollett} 
\bibliography{anna_bib} % Abbreviate journal titles.

\end{spacing}

\end{document}

### 2021-06-04 manuscript/supplement_motifs.tex
Initial Overleaf Import

\documentclass[12pt]{article} 
\usepackage{amsmath} 
\usepackage[dvips]{graphicx}
\usepackage{multirow} 
\usepackage{geometry} 
\usepackage{pdflscape}
\usepackage[labelfont=bf]{caption} 
\usepackage{setspace}
\usepackage[running]{lineno} 
% \usepackage[numbers,sort]{natbib}
\usepackage[round]{natbib} 
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{xr}

\newcommand{\methods}{\textit{Materials \& Methods}}
\newcommand{\SI}{\textit{Appendix}~}

\topmargin -1.5cm % 0.0cm 
\oddsidemargin 0.0cm % 0.2cm 
\textwidth 6.5in
\textheight 9.0in % 21cm
\footskip 1.0cm % 1.0cm

\usepackage{authblk}

\begin{document}

\title{Do motif roles provide unique information about species risk of extinction?\\ \medskip Supplementary Information}

\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{2}$, Kate Wootton$^{3}$, Gyuri Barab\'{a}s$^{1}$, Anna Ekl\"{o}f$^{1}$} 
\date{\small$^1$Department of Theoretical Biology, Chemistry, and Physics\\ 
Link\"{o}ping University\\
Link\"{o}ping, Sweden\\
\medskip
\small$^2$Department of Agricultural Sciences\\
University of Helsinki\\
Helsinki, Finland\\
\medskip
\small$^3$ University of Colorado?\\
\medskip
$^\dagger$ Corresponding author:\\
}

\maketitle 
\raggedright

\setlength{\parindent}{15pt} 
\begin{spacing}{2.0}

\clearpage

\section*{Table of contents}
    \begin{spacing}{1.0}

    \begin{enumerate}

        \item How do counts of motifs in a network profile vary with S and C? \\

            \begin{itemize}
            \item Same methods as main text except using counts instead of proportions to define motif profiles
            \end{itemize}

        \item Motif profiles differed slightly between network sets

           \begin{itemize}
                \item Comparing proportions of motifs in un-filtered, filtered, and acyclic networks.
                \item AFAIK not referred to in text
            \end{itemize}

         \item Persistence and count-based motif participation

           \begin{itemize}
                \item Should be as main text section 4, but counts
            \end{itemize}

        \item Motif-persistence relationships and network size
            \begin{itemize}
                \item Detailed results for variability of slopes for persistence $\approx$ motif participation in networks of different sizes.
                \item Main text refers to C, moved figure to Main while we check that we've got everything.
            \end{itemize}

        \item How does persistence vary with degree and trophic level?

            \begin{itemize}
                \item Heat map showing persistence vs. TL with BP=0
                \item Heat map showing persistence vs. TL with BP=1
            \end{itemize}

    \end{enumerate}
\end{spacing}
\clearpage

\section{How do counts of motifs in a network profile vary with S and C?}

    As expected, all four motifs had higher counts in larger and more highly connected networks.
    The precise effects of size, connectance, and their interaction differed between motifs  (Table~\ref{network_motif_lms}).
    This suggests that the meso-scale structure of a network varies with size and connectance, not just the total number of motifs.
    Indeed, the motif profile of a network is significantly related to its size, connectance, and their interaction ($F_{1,2996}$=2279, $p$\textless0.001; $F_{1,2996}$=3067, $p$\textless0.001; and $F_{1,2996}$=1187, $p$\textless0.001, respectively).

    \begin{table}[h!]
        \centering
        \caption{Coefficients for linear regressions relating the total count of all motifs or the count of each three-species motif in a network to network size, connectance, and their interaction. Coefficients in \textbf{bold} are significant (\textless0.05). Note that, while there are 13 unique three-species motifs, only four are possible in acyclic networks such as those considered here.}
       \label{network_motif_lms}
       \begin{tabular}{c|c c c c c}
        Motif & Intercept & Size & Connectance & Interaction \\
        \hline
        Total count & \textbf{-2.85$\times10^3$} & \textbf{46.6} & \textbf{-2.12$\times10^{5}$} & \textbf{4.72$\times10^{3}$} \\
        \hline
        Thee-species Chain & \textbf{-1.40$\times10^{3}$} & \textbf{26.1} & \textbf{-3.74$\times10^{4}$} & \textbf{8.58$\times10^{2}$} \\
        Apparent Competition & \textbf{-2.23$\times10^3$} & \textbf{42.0} & 
        \textbf{-7.26$\times10^4$} & \textbf{1.63$\times10^3$} \\
        Direct Competition & 26.6 & 0.777 & \textbf{-4.85$\times10^4$} & \textbf{9.89$\times10^2$} \\
        Omnivory & \textbf{7.57} & \textbf{-22.2} & \textbf{-5.37$\times10^4$} & \textbf{1.24$\times10^3$} \\
        \hline
        \end{tabular}
    \end{table}

    \begin{figure}[h!]
        \centering
        \includegraphics[width=.75\textwidth]{figures/motif_profile_dispersion.pdf}
        \caption{Dispersion of motif profiles within a network about the centroid for that combination of network size and connectance. Each circle represents a single network; circle colour indicates network size. Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
        \label{dispersion_rawmotifs}
    \end{figure}

    Motif profiles were similarly variable across networks with different sizes, but were not homogeneously variable across connectances or size-connectance combinations ($F_{5,2994}$=0.365, $p$=0.873; $F_{4,2995}$=4.49, $p$=0.001; and $F_{29,2970}$=27.1, $p$\textless0.001 , respectively).
    These differences in variability of motif profiles can cause false positives in the PERMANOVA.
    Variability in motif profiles was lowest among large networks with moderate connectance and was highest among the least-connected and most-connected networks we considered (Fig.~\ref{dispersion_rawmotifs}).

    \begin{figure}[h!]
        \centering
        \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
        \caption{The effect of number of times a species appears (x-axis) in any of the various motifs (columns) on persistence (y-axis). The effect of each motif participation is fitted by linear mixed-effect models, following Equation~\ref{absreq}, except that the disturbance levels are fitted separately. The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
        \label{fig:abs_lmer_all}
    \end{figure}

\clearpage

\section{Motif profiles differed slightly in different network sets}

    Not all motifs were equally likely to occur in the networks we considered. 
    Across the initially stable, acyclic networks (i.e., those used in the main text), the omnivory motif tended to make up a smaller proportion of the motif profile than other motifs.
    The omnivory motif was also rarer than the three-species chain and apparent competition motifs before networks were rendered acyclic, but made up a similar proportion to the direct competition motif (Table~\ref{tab:acyclic_proportions}).
    The omnivory motif made up a larger proportion of the full set of initially stable networks, including those which were excluded from our final analysis due to overly high trophic levels.
    Nevertheless, in each group of networks the apparent competition and three-species chain motifs made up the largest proportion of the motif profile.
    This suggests that, while rendering networks acyclic does change proportions slightly, stable networks created by the niche model appear to be generally biased towards the chain and apparent competition motifs.

    \begin{table}[h!]
        \centering
        \caption{Proportions of the four motifs which appear in Bayesian networks (three-species chain: `Chain', apparent competition: `AC', direct competition `DC', and omnivory: `Omni') and other motifs (`Other') in the motif profiles of acyclic, filtered, and original networks. Acyclic networks are those used in the main text analyses. Filtered networks are those used in the main text analyses \emph{before} being rendered acyclic. Original networks are all simulated networks, including those removed because of overly high maximum trophic levels.}
        \label{tab:acyclic_proportions}            \footnotesize
        \begin{tabular}{c|l l l | l l l | l l l |}
        & \multicolumn{3}{c|}{Acyclic networks} & \multicolumn{3}{c|}{Filtered networks} & \multicolumn{3}{c|}{Original networks} \\
        Motif & Mean (SD) & Min & Max & Mean (SD) & Min & Max & Mean (SD) & Min & Max \\
        \hline
        Chain & 0.235 (0.042) & 0.109 & 0.412 & 0.225 (0.045) & 0.100 & 0.412 & 0.218 (0.043) & 0.096 & 0.413 \\
        AC & 0.422 (0.064) & 0.250 & 0.765 &
        0.417 (0.067) & 0.233 & 0.765 & 0.397 (0.064) & 0.231 & 0.765 \\
        DC & 0.179 (0.041) & 0.050 & 0.359 & 0.155 (0.046) & 0.040 & 0.359 & 0.140 (0.046) & 0.031 & 0.359 \\
        Omni & 0.164 (0.079) & 0.006 & 0.346 & 0.156 (0.075) & 0.006 & 0.319 & 0.184 (0.073) & 0.006 & 0.329 \\
        \hline
        Other & \multicolumn{3}{c|}{NA} & 0.047 (0.042) & 0.00 & 0.200 &
        0.062 (0.046) & 0.00 & 0.258 \\
        \end{tabular}
    \end{table}

\clearpage

\section{Persistence and count-based motif participation}

    \subsection*{Methods}

        We test whether participating in higher absolute numbers (i.e., counts) of a motif increases persistence using a set of models that parallels those used to test whether having a greater proportion of the role made up by a particular motif (see \emph{Main Text}).  
        Specifically, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
        In each case, we considered only non-basal species (i.e., those with at least one prey) as the persistence of basal species is determined only by their baseline probability.
        These models included the absolute number of times the focal species appeared in the focal motif, the increase in probability of extinction for basal species, and the interaction between the two as well as a random intercept for the interaction between species richness and connectance (equations.
        All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
        We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.

            The formula for these LMMs are:         
            \begin{equation}
                \Psi_{ik} \approx \chi_{i} + \pi_{disturbed_k} + \chi_{i}\pi_{disturbed_k} + S:C_{i} ,
                \label{absreq}
            \end{equation}

            \noindent where $\Psi_ik$ is the persistence of species $i$ during disturbance level $k$, $\chi_{i}$ is the number of times species $i$ appears in the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.

            % Additionally, we performed a similar linear regression of persistence against motif participation within each simulated network and for each level of disturbance to basal resources (XX regressions total).
            % These regressions were fit using the R~\citep{R} base function `lm'.
            % We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.

    \subsection*{Results}

        \begin{figure}[h!]
            \centering
            \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
            \caption{The effect of number of times a species appear (x-axis) in any of the various motifs (columns) on persistence (y-axis). The effect of each motif participation is fitted by linear mixed-effect models, following \emph{Equation 2; Main Text}, except that the disturbance levels are fitted separately. The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
            \label{fig:abs_lmer_all}
        \end{figure}

        Persistence increased significantly with the number of times a species appeared in any of the four possible motifs in the food webs (Table~\ref{tab:absolute_number} and Figure~\ref{fig:abs_lmer_all}). 
        These relationships were slightly stronger for the number of times a species appears in the three-species chain and direct competition motifs and somewhat weaker for apparent competition and omnivory. The absolute difference between the motifs was small, with a clear positive relationship between persistence and motif count present in all cases. 
        As expected, persistence decreased significantly with increasing disturbance of basal resources and the strength of this relationship was very similar across the four motifs.
        The level of disturbance to basal resources also affected the relationship between persistence and motif counts.
        For all four motifs, there was a negative interaction between motif count and disturbance such that, for high levels of disturbance, participating more often in any of the motifs would result in decreased persistence.

        \begin{table}[h!]
            \caption{Coefficients ($\beta$) for models relating persistence to the number of times a species appears in the focal motif, the level of disturbance to basal resources, and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{absreq}). $p$-values for all coefficients in all models were \textless0.001. We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
            \label{tab:absolute_number}        \centering
            \begin{tabular}{c|c  c c  c | c c |}
            Motif & Intercept & Motif count & Disturbance & Interaction & $R^2_M$ & $R^2_C$ \\
            \hline
            3-sp Chain & 0.762 & 1.14$\times10^{-4}$ & -0.388 & -1.36$\times10^{-4}$ & 0.897 & 0.913 \\
            Apparent Compet. & 0.770 & 3.41$\times10^{-5}$ & -0.397 & -4.36$\times10^{-5}$ & 0.898 & 0.911\\
            Direct Competition & 0.764 & 1.22$\times10^{-4}$ & -0.398 & -8.46$\times10^{-5}$ & 0.895 & 0.914 \\
            Omnivory & 0.770 & 6.01$\times10^{-5}$ & -0.400 & -5.91$\times10^{-5}$ & 0.897 & 0.912 \\                
            \end{tabular}
            \end{table}

\clearpage

\section{Motif-persistence relationships and network size}

    When comparing the outcomes from linear mixed-effect models (LMMs) relating persistence to motif roles across all networks (Fig. 3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately, the main trends are similar but different-sized networks showed slightly different results results for some motifs (Figure temporarily in main text).%(Figure~\ref{fig:density_prop_S}). 
    For the three-species chain and omnivory motifs, a significant interaction between disturbance and the effect of the focal motif means that having a larger proportion of the role be made up of the chain or omnivory motif is associated with lower persistence at high levels of disturbance (Fig. 3 and Table 1, \emph{Main Text}). 
    The same pattern holds overall for LRs fit within a network  %(Figure~\ref{fig:density_prop_S})
    : the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance. 
    This trend is most pronounced in large networks.% (bottom panel, Figure~\ref{fig:density_prop_S}). 
    Besides the decrease in fraction of positive slopes with increasing disturbance, the omnivory motif show a higher overall fraction of positive slopes in comparison to the three-species chain motif, corresponding to the positive main effect of omnivory (Fig. 3 and Table 1, \emph{Main Text}).

    % \begin{figure}[h!]
    %      \centering
    %      \includegraphics[width=\textwidth]{manuscript/figures/prop_dens_bp_vs_S_allC.pdf}

    %     \caption{Density (y-axis) of slopes (x-axis) for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. The slopes are derived from linear mixed-effect models, with a random intercept for the interaction between species richness and network size. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence. The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to motif (legend). Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right). Rows show various levels of network size. The vertical line indicate zero on the x-axis.}
    %     \label{fig:density_prop_S}
    % \end{figure}

    For the apparent and direct competition motifs, the LMMs include a positive interaction between disturbance and the proportion of the motif in a species' role (Fig. 3 and Table 1; \emph{Main Text}).
    This means that having more of the role made up of these two motifs is increasingly beneficial at higher levels of disturbance, a trend which is also evident from the increasing fraction of positive slopes with increasing disturbance based on LRs of each network at each level of disturbance. % (Figure~\ref{fig:density_prop_S}).
    As with the trends for omnivory and three-species chains, these increases are strongest in large networks. % (bottom row, Figure~\ref{fig:density_prop_S}).
    % For direct competition, the fraction of slopes are similar across disturbance levels. Although this interaction is not positive when accounting for network size, the fraction of positive slopes are throughout high.
    The increase in the proportion of positive slopes was much stronger for apparent competition than direct competition, consistent with the smaller (but still significant) interaction term for between direct competition in the LMMs.
    A high proportion of direct competition in a species' role tended to be associated with greater probability of persistence across levels of disturbance and network size.
    % In conclusion, species richness does not considerably alter the trends visible in the LMMs. High species richness more clearly display an interaction between motif proportion, persistences and disturbance levels, while this effect is smaller with low species richness.

\section{How does persistence vary with degree and trophic level?}

  \begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}

     \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$. The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis). Persistence decreases with darker color in the heat map.}
     \label{fig:heatmap_stl_BP0}
 \end{figure}

 \begin{figure}
     \centering
     \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}

     \caption{Primary producers experience a disturbance reducing their probability to go extinct to $\pi_{disturbed} = 0.5$ All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$. The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis). Persistence decreases with darker color in the heat map.}
     \label{fig:heatmap_stl_BP1}
 \end{figure}

\clearpage 
\bibliographystyle{ecollett} 
\bibliography{MyCollection} % Abbreviate journal titles.

\end{spacing}

\end{document}

### 2021-06-04 code/acyclic_sorted webs.R

@@ -1,0 +1,69 @@
+require(XRJulia) ## package allowing for Julia interfacing
+require(igraph)
+require(tidyverse) ## for efficient data manipulation and plotting
+source("BN-DFS.R") ## to make the adjacency matrix acyclic
+source("getTL.R") ## to get prey average TL, filter good webs
+
+## sort topological and remove cycles
+topo_sort_acyclic <- function(A) {
+    A<-as.matrix(A)
+    ## Make sure producers are the first rows; this is needed for the DFS
+    ## algorithm to work (which removes all cycles from the digraph)
+    A <- A[order(rowSums(A)),order(rowSums(A))]
+    ## Now make the digraph acyclic
+    A <- DFS(A)
+    
+    ## Find a sorting of A's rows and columns to make A lower triangular
+    o <- graph_from_adjacency_matrix(A) %>% topo_sort(mode="in")
+    ## Return sorted, lower triangular adjacency matrix A
+    return(A[o,o])
+}
+
+# 50sp webs done, 60sp webs done,
+sp_nos <- seq(50,100, by=10)
+c <- seq(0.02,0.18,by=0.04)
+web_nos <- c(0:649)
+for (i in 1:length(sp_nos)){
+    sp_no <- sp_nos[i]
+    print(c(sp_no,'Species'))
+    for (j in 1:length(c)){
+        connectance <- c[j]
+        print(c(connectance,'Connectance'))
+        goodfiles=as.character(list.files(path=paste0("../data/networks/acyclic_webs/",as.character(sp_no),"/",as.character(connectance)), full.names=FALSE, pattern="*.csv"))
+        if(length(goodfiles)<100){ # Keep addding files up to 100
+            print('incomplete')
+            n=length(goodfiles)
+            for (k in n:max(web_nos)){
+                if(!paste("initial_matrix_",k,".csv",sep='')%in%goodfiles && n<100){
+                    A <- read.table(paste("../data/networks/pre_disturbance/",sp_no, "/", connectance, "/", "initial_net_",
+                                          k,".csv",sep=''), sep=",", header=TRUE)
+                    rownames(A) <- colnames(A) <- 1:nrow(A) 
+                    image(t(apply(A,2,rev)))
+                    
+                    A <- topo_sort_acyclic(A)
+                    image(t(apply(A,2,rev)))
+
+                    # Calculate the PATL to sort good webs
+                    FW <- list()
+                    FW$M <- t(A)
+                    FW <- getTL(FW)
[…]

### 2021-06-04 code/calc_TL.R

@@ -1,0 +1,50 @@
+require(tidyverse) ## for efficient data manipulation and plotting
+require(igraph)
+source("getTL.R") ## to get prey average TL
+dir.create('../data/TL',showWarnings=FALSE) # Storing the degree/TL files in the same way as the motif files
+sp_nos <- seq(50,100, by=10)
+c <- seq(0.02,0.18,by=0.04)
+basal_proportion <- seq(0,1,by=0.1)
+
+for (i in 1:length(sp_nos)){
+    sp_no <- sp_nos[i]
+    print(sp_no)
+    dir.create(paste0('../data/TL/',sp_no,sep=''),showWarnings=FALSE)
+    for (j in 1:length(c)){
+        connectance <- c[j]
+        print(connectance)
+        dir.create(paste0('../data/TL/',sp_no,'/',connectance,sep=''),showWarnings=FALSE)
+        matrixfiles<-as.character(list.files(path=paste0("../data/networks/acyclic_webs/",as.character(sp_no),"/",as.character(connectance)), full.names=TRUE, pattern="*.csv"))
+        if(!length(matrixfiles)==100){
+            print(c(sp_no,connectance,length(matrixfiles)))
+            print('Error: Too many or too few webs')
+        }
+        for (k in 1:length(matrixfiles)){
+            infile=matrixfiles[k]
+            webnum=strsplit(strsplit(infile,'_')[[1]][4],'.csv')[[1]][1]
+            A <- read.table(infile, sep=",", header=TRUE)
+            S <- dim(A)[1]
+            A <- as.matrix(A)
+            dd <- distances(graph_from_adjacency_matrix(A), mode="out")
+            basals <- rownames(A[rowSums(A)==0,])
+            short_tl <- 1 + as.numeric(apply(dd[,as.numeric(basals)], 1, min))
+            # Getting a subscript out of bounds error from the dd[,basals] withtout the as.numeric()
+            FW <- list()
+            FW$M <- t(A)
+            FW <- getTL(FW)
+            
+            df <- data_frame(
+                sp_no = rownames(A), 
+                in_degree = rowSums(A),
+                out_degree = colSums(A),
+                shortest_tl = short_tl,
+                preyav_tl = round(FW$TL,2)
+            )
+            write.table(df,file=paste0("../data/TL/", sp_no, "/", connectance, "/web_",webnum, ".tsv"),sep='\t')
[…]

### 2021-06-04 code/calculate_motifs_and_roles.py

@@ -1,0 +1,77 @@
+import sys
+import os
+import pymfinder as py
+
+def main():
+
+    webdir='../data/edgelists/acyclic/'
+    motifdir='../data/motif_frequencies/acyclic/'
+    roledir='../data/roles/acyclic/'
+    pardir='../data/motif_participation/acyclic/'
+
+    for s in sorted(os.listdir(webdir)):
+      try:
+        os.mkdir(roledir+s)
+      except:
+        pass
+      try:
+        os.mkdir(motifdir+s)
+      except:
+        pass
+      try:
+        os.mkdir(pardir+s)
+      except:
+        pass
+      for c in sorted(os.listdir(webdir+s)):
+        print s,c
+        try:
+          os.mkdir(roledir+s+'/'+c)
+        except:
+          pass
+        try:
+          os.mkdir(motifdir+s+'/'+c)
+        except:
+          pass
+        try:
+          os.mkdir(pardir+s+'/'+c)
+        except:
+          pass
+        for web in os.listdir(webdir+s+'/'+c+'/'):
+          infile=webdir+s+'/'+c+'/'+web
+
+          size=3
+          rolefile=roledir+s+'/'+c+'/'+web.split('.tsv')[0]+'.roles_'+str(size)
+          motfile=motifdir+s+'/'+c+'/'+web.split('.tsv')[0]+'.motifs_'+str(size)
+          parfile=pardir+s+'/'+c+'/'+web.split('.tsv')[0]+'.participation_'+str(size)
+
+          result=py.motif_roles(infile,motifsize=size,stoufferIDs=False,allroles=True)
+          motifs=str(result).split('node')[0]
+          roles='node'+str(result).split('node')[1]
+
+
+          presult=py.motif_participation(infile,motifsize=size,stoufferIDs=False,allmotifs=True)
+          partic=str(presult).split('node')[1]
+
+          e=open(parfile,'w')
+          e.write(partic)
+          e.close()
+
+          f=open(rolefile,'w')
+          f.write(roles)
+          f.close()
+
+          g=open(motfile,'w')
+          g.write(motifs)
+          g.close()
+
+          size=4
+          parfile2=pardir+s+'/'+c+'/'+web.split('.tsv')[0]+'.participation_'+str(size)          
+          presult=py.motif_participation(infile,motifsize=size,stoufferIDs=False,allmotifs=True)
+          partic=str(presult).split('node')[1]
+
+          e=open(parfile2,'w')
[…]

### 2021-06-04 code/collect_motif_profiles.py

@@ -1,0 +1,61 @@
+import os
+import sys
+
+# Creates a dictionary with motif counts listed by species, connectance, network, motif.
+def read_role_file(roledir):
+	role_dict={}
+	for s in range(50,110,10):
+		role_dict[s]={}
+		for c in os.listdir(roledir+str(s)):
+			role_dict[s][c]={}
+			for netfile in os.listdir(roledir+str(s)+'/'+str(c)+'/'):
+				i=netfile.split('.motifs')[0].split('_')[2]
+				counts=[]
+				if '.motifs_3' in netfile:
+					role_dict[s][c][i]={}
+					net=roledir+'/'+str(s)+'/'+str(c)+'/'+netfile
+					f=open(net,'r')
+					for line in f:
+						if len(line.split())>1 and line.split()[0]!='motif':
+							motif=str(line.split()[0])
+							count=int(line.split()[1])
+							counts.append(count)
+							if motif in ['6','12','36','38']:
+								role_dict[s][c][i][motif]=count
+					f.close()
+				role_dict[s][c][i]['total']=sum(counts)
+	return role_dict
+
+
+def write_motif_file(role_dict,outfile):
+	f=open(outfile,'w')
+	f.write('Size\tConnectance\tNetwork\tm6\tm12\tm36\tm38\tTotal\n')
+	for s in range(50,110,10):
+		for c in ['0.02','0.06','0.1','0.14','0.18']:
+			for i in role_dict[s][c]:
+				f.write('\t'.join([str(s),c,str(i)]))
+				f.write('\t'+str(role_dict[s][c][i]['6']))
+				f.write('\t'+str(role_dict[s][c][i]['12']))
+				f.write('\t'+str(role_dict[s][c][i]['36']))
+				f.write('\t'+str(role_dict[s][c][i]['38']))
+				f.write('\t'+str(role_dict[s][c][i]['total'])+'\n')
+	f.close()
+
+	return
+
+
+def main():
+	# role_dict takes s, c, network, motif: count
+	# only motifs 6, 12, 36, 38 exist in the Bayesian networks
+	motif_dict=read_role_file('../data/motif_frequencies/acyclic/')
+
+	write_motif_file(motif_dict,'../data/3sp_motif_profiles.tsv')
+	print 'Motif roles read'
+
+	motif_dict=read_role_file('../data/motif_frequencies/original/')
+
[…]

### 2021-06-04 code/collect_roles.py

@@ -1,0 +1,245 @@
+import os
+import sys
+
+listdir='../data/edgelists/acyclic/'
+Annas_to_mine={}
+for S in os.listdir(listdir):
+	Annas_to_mine[S]={}
+	for C in os.listdir(listdir+S):
+		i=1
+		Annas_to_mine[S][C]={}
+		myfils=os.listdir(listdir+S+'/'+C)
+		for filname in sorted(myfils):
+			webnum=filname.split('_')[-1].split('.tsv')[0]
+			Annas_to_mine[S][C][str(i)]=webnum
+			i=i+1
+
+# Need to get Anna to redo "all_new" with web numbers from new webs, not sequentially.
+
+# Also have motif frequencies per web, may be interesting to compare to S, C?
+
+# Then bin the species according to ... things, like in Benno's paper.
+# Then do some stats and send to Anna.
+def read_persistence_file(persfile,role_dict):
+	persdict={}
+	TLdict={}
+	degdict={}
+	bpdict={}
+	for s in role_dict:
+		persdict[s]={}
+		TLdict[s]={}
+		degdict[s]={}
+		bpdict[s]={}
+		for c in role_dict[s]:
+			persdict[s][c]={}	
+			TLdict[s][c]={}
+			degdict[s][c]={}
+			for bp in [0.0,0.2,0.4,0.6,0.8,1.0]:
+				persdict[s][c][bp]={}	
+				TLdict[s][c][bp]={}
+				degdict[s][c][bp]={}
+				for i in role_dict[s][c]:
+					persdict[s][c][bp][i]={}
+					TLdict[s][c][bp][i]={}
+					degdict[s][c][bp][i]={}	
+
+	f=open(persfile,'r')
+	for line in f:
+		if line.split()[0]!='"sp_no"':
+			sp='"sp'+line.split()[1][1:-1]+'"'
+			indeg=int(line.split()[2])
+			outdeg=int(line.split()[3])
+			STL=int(line.split()[4])
+			PATL=float(line.split()[5])
+			per=float(line.split()[6])
+			basal_p=float(line.split()[7])
+			S=str(int(line.split()[8]))
+			C=str(float(line.split()[9]))
+			web=str(int(line.split()[10]))
+			Aweb=Annas_to_mine[S][C][web]
+			persdict[int(S)][C][basal_p][Aweb][sp]=(str(per))
[…]

### 2021-06-04 code/extract_edgelists.R

@@ -1,0 +1,27 @@
+library(igraph)
+# Convert graphs to edgelists
+
+for(s in seq(50,100,10)){
+	print(paste0(as.character(s),' species'))
+	dir.create(paste0('../data/edgelists/acyclic/',as.character(s)))
+	for(c in seq(0.02,0.18,0.04)){
+		print(paste0('connectance=',as.character(c)))
+		dir.create(paste0('../data/edgelists/acyclic/',as.character(s),'/',as.character(c)))
+        matrixfiles<-as.character(list.files(path=paste0("../data/networks/acyclic_webs/",as.character(s),"/",as.character(c)), full.names=TRUE, pattern="*.csv"))
+        if(!length(matrixfiles)==100){
+        	print(c(s,c,length(matrixfiles)))
+        	print('Error: Too many or too few webs')
+        }
+		for(i in 1:length(matrixfiles)){
+			# Select the json file to work with
+			infile=matrixfiles[i]
+			# Keeping the same number as the matrices
+			webnum=strsplit(strsplit(infile,'_')[[1]][4],'.csv')[[1]][1]
+			A=read.csv(infile,header=TRUE,sep=',')
+			# convert to edgelist and save
+			edges=get.edgelist(graph.adjacency(as.matrix(A)))
+			write.table(edges,file=paste0('../data/edgelists/acyclic/',as.character(s),'/',as.character(c),'/initial_edges_',as.character(webnum),'.tsv',sep=''),sep='\t',row.names=FALSE,col.names=FALSE)
+		}
+	}
+}
+

### 2021-06-04 code/getTL.R

@@ -1,0 +1,50 @@
+# ADD THIS IN MAIN CODE!
+# FW <- list()
+# FW$M <- t(A)
+# FW <- getTL(FW)
+
+normalizeM <- function(M){
+    colsum_M <- colSums(M);
+    colsum_M[colsum_M==0] <- 1;
+    return(t(t(M)/colsum_M));
+}
+
+getTL <- function(FW){
+    S <- dim(FW$M)[1] # number of species
+    ## normalize M such that a predator has no preference over
+    ## multiple prey.
+    ##Ref: A process-oriented approach to the multispecies functional response
+    ##  Koen-Alonso
+    ## From energetics to ecosystems: The dynamics and structure of ecological systems Springer (2007)
+    M <- normalizeM(FW$M)
+    
+    ## Since: 
+    ##
+    ##    	TL[j] = 1 + sum_over_i ( M[i, j] TL[i] )
+    ## 
+    ## which written in matrix notation is:
+    ##
+    ##		TL = 1 + t(M) %*% TL
+    ## =>  (I - t(M) ) %*% TL = 1    <- this is a vector
+    ## =>  TL = (I - t(M))^(-1) %*% rep(1, S);
+    ## if (I - t(M)) is not singular, 
+    ## then it is fine to do the matrix inversion
+    if(det( diag(S)-t(M) )!=0){
+        FW$TL <- solve(diag(S)-t(M), rep(1, S))
+    } else{
+        ## Otherwise, need to use
+        ## I + t(M) + t(M)%*% t(M) + t(M)%*%t(M) %*% t(M) + ... to approximate
+        ## (I - t(M))^(-1)
+        tmp <- diag(S)
+        for(i in 1:9){
+            tmp <- tmp  %*% t(M) + diag(S)
+        }
+        FW$TL <- tmp %*% rep(1, S)
+    }
+    FW$W <- M ## Save the normalized adjacency matrix as weight matrix
+    return(FW)
+}
+
+
+
+

### 2021-06-04 code/motif_SC_lms.py

@@ -1,0 +1,191 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Blues')  # The blue is very beautiful but maybe harder to see.
+colors.add_color(173,132,191,'Purple')
+colors.add_color(34,90,34,'Green')
+# colors.add_color(120,120,120,'grey')
+# colors.add_color(255,125,125,'lightish_red')
+# colors.add_color(200,200,200,'lightgrey')
+
+def read_dispfile(dispfile):
+  dispdict=[]
+
+  f=open(dispfile,'r')
+  for line in f:
+    if line.split()[0]!='"Distance"':
+      dist=float(line.split()[1][1:-1])
+      S=int(line.split()[3][1:-1])
+      C=float(line.split()[4][1:-1])
+      dispdict.append((S,C,dist))
+
+  f.close()
+
+  return dispdict
+
+def format_graph(graph,graphtype):
+  graph.yaxis.bar.linewidth=1
+  graph.xaxis.bar.linewidth=1
+  graph.frame.linewidth=1
+
+  graph.world.xmin=0
+  graph.world.xmax=0.21
+  graph.xaxis.tick.major=.05
+
+  graph.world.ymin=0
+  if graphtype=='total':
+    graph.world.ymax=50000
+    graph.yaxis.tick.major=10000
+    graph.yaxis.label.configure(text='Total motifs',char_size=1,just=2,place='normal')
+    graph.yaxis.ticklabel.configure(format='scientific',prec=1,char_size=.75)
+  else:
+    graph.world.ymax=.60000001
+    graph.yaxis.tick.major=.2
+    graph.yaxis.label.configure(text='Proportion',char_size=1,just=2,place='normal')
[…]

### 2021-06-04 code/plot_dispersion.py

@@ -1,0 +1,143 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Blues')  # The blue is very beautiful but maybe harder to see.
+colors.add_color(173,132,191,'Purple')
+colors.add_color(34,90,34,'Green')
+# colors.add_color(120,120,120,'grey')
+# colors.add_color(255,125,125,'lightish_red')
+# colors.add_color(200,200,200,'lightgrey')
+
+def read_dispfile(dispfile):
+  dispdict=[]
+
+  f=open(dispfile,'r')
+  for line in f:
+    if line.split()[0]!='"Distance"':
+      dist=float(line.split()[1][1:-1])
+      S=int(line.split()[3][1:-1])
+      C=float(line.split()[4][1:-1])
+      dispdict.append((S,C,dist))
+
+  f.close()
+
+  return dispdict
+
+def format_graph(graph,graphtype):
+  graph.yaxis.bar.linewidth=1
+  graph.xaxis.bar.linewidth=1
+  graph.frame.linewidth=1
+
+  graph.world.xmin=0
+  graph.world.xmax=0.21
+  graph.xaxis.tick.major=.05
+
+  graph.world.ymin=0
+  if graphtype=='prop':
+    graph.world.ymax=0.25
+    graph.yaxis.tick.major=.05
+    graph.yaxis.ticklabel.configure(format='decimal',prec=2,char_size=.75)
+  else:
+    graph.world.ymax=0.4
+    graph.yaxis.tick.major=0.1
+    graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
+
+  graph.xaxis.ticklabel.configure(format='decimal',prec=2,char_size=.75)
[…]

### 2021-06-04 code/plot_roles_vs_TL.py

@@ -1,0 +1,258 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Greys',n=10)  # The blue is very beautiful but maybe harder to see.
+# Match Anna's ggplot colours
+colors.add_color(94,79,162,'Rpurple')
+colors.add_color(50,136,189,'Rblue')
+colors.add_color(102,194,165,'Rgreen1')
+colors.add_color(171,221,164,'Rgreen2')
+colors.add_color(230,245,152,'Rgreen3')
+colors.add_color(255,255,191,'Ryellow1')
+colors.add_color(254,224,139,'Ryellow2')
+colors.add_color(253,174,97,'Rorange1')
+colors.add_color(244,109,67,'Rorange2')
+colors.add_color(213,62,79,'Rred1')
+colors.add_color(158,1,66,'Rred2')
+
+# colors.add_color(120,120,120,'grey')
+# colors.add_color(255,125,125,'lightish_red')
+# colors.add_color(200,200,200,'lightgrey')
+
+# Each point will be mean of one network
+def read_datafile(datafile):
+  speciesprops={}
+  f=open(datafile,'r')
+  for line in f:
+    if line.split()[0]!='Size':
+      S=int(line.split()[0])
+      C=float(line.split()[1])
+      basal_p=line.split()[2]
+      network=line.split()[3]
+      species=line.split()[4]
+      ID=str(S)+'_'+str(C)+'_'+network+'_'+species
+      pers=line.split()[5]
+      in_deg=float(line.split()[6])
+      out_deg=float(line.split()[7])
+      STL=int(line.split()[8])
+      PATL=float(line.split()[9])
[…]

### 2021-06-04 code/stat_analysis/fig_norm_motif_mailv.R

@@ -1,0 +1,70 @@
+#### TEST MOTIFS AND PERSISTENCE 
+require(ggpmisc)
+require(tidyverse)
+
+color_sc <- c("#CC79A7", "#E69F00", "#56B4E9")
+labs_bp <- c(
+    "0" = "0 %",
+    "0.4" = "40 %",
+    "0.8" = "80 %")
+
+data<-read.table(file = '3sp_roles_participation.tsv', sep = '\t', header = TRUE)
+
+cc <- c(0.02, 0.1, 0.18)
+bp <- c(0,0.4,0.8)
+
+tab2 <- data %>%
+    filter(in_Degree > 0, Size %in% c(50,100), Basal_p %in% bp, 
+           Connectance %in% cc) %>%
+    mutate(preyav_length=cut(PATL, breaks=c(1.9, 3, 4, 5, 6), 
+                             labels=c("2","3","4","5"))) %>% 
+    select(Network, Species, Persistence, m12, m36, m38, m6, 
+           preyav_length, Connectance, Basal_p, Size) %>%
+    mutate(omni=m38, chain=m12, dcomp=m36,acomp=m6) %>%
+    gather("motif","counts",4:7) %>%
+    group_by(Network, Species, Persistence, preyav_length, Basal_p,
+             omni, chain, dcomp, acomp) %>%
+    summarise(total=sum(counts)) %>%
+    ungroup() %>%
+    mutate(frac_omni = omni/total, frac_chain=chain/total,
+           frac_dcomp=dcomp/total, frac_acomp=acomp/total) %>%
+    gather("norm_motif", "frac", 11:14) 
+
+ggplot(tab2, aes(x=frac, y=Persistence, color=as.factor(Basal_p))) +
+    #geom_point(size=1, alpha=0.5) +
+    facet_grid(~norm_motif) +
+    theme(axis.text.x = element_text(angle = 90, size=7)) +
+    scale_color_manual(name = "Basal prop.", labels = labs_bp, 
+                       values=color_sc) +
+    stat_smooth(method="lm", se=FALSE) +
+    stat_fit_glance(method="lm", geom="text", label.x=c(0.7,0.7,0.7),
+                    label.y = c(0.58, 0.56, 0.54),
+                    aes(label=sprintf('italic(p)~"="~%.3f',
+                                      stat(..p.value..))),
+                    parse=TRUE,size = 3)
+
+### RAW COUNTS
+tab2 <- data %>%
[…]

### 2021-06-04 code/stat_analysis/persistence_vs_SC_lm.py

@@ -1,0 +1,173 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Spectral',n=13)  # The blue is very beautiful but maybe harder to see.
+colors.add_color(173,132,191,'Purple')
+colors.add_color(34,90,34,'Green')
+# colors.add_color(120,120,120,'grey')
+# colors.add_color(255,125,125,'lightish_red')
+# colors.add_color(200,200,200,'lightgrey')
+
+
+scales={
+  'Size':(79.1882,16.62225),
+  'Connectance':(0.1044983,0.05559927),
+  'Disturbance':(0.3,0.1366261),
+}
+
+lms={
+  'Intercept': 5.723e-01,
+  'Size':-5.148e-04,
+  'Connectance':-1.293e-02,
+  'Disturbance':-1.393e-01,
+  'Size:Connectance':-5.245e-04,
+  'Size:Disturbance':-3.485e-03,
+  'Connectance:Disturbance':-6.998e-05,
+  'Size:Connectance:Disturbance':-2.693e-04 }
+
+def format_graph(graph,simple):
+  graph.yaxis.bar.linewidth=1
+  graph.xaxis.bar.linewidth=1
+  graph.frame.linewidth=1
+
+
+  graph.world.ymin=0
+  graph.world.ymax=1
+  graph.yaxis.tick.major=.2
+  graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
+  graph.yaxis.label.text='Mean persistence'
+  if simple=='Size':
+    graph.xaxis.label.text='Network size'
+    graph.world.xmin=45
+    graph.world.xmax=105
[…]

### 2021-06-04 code/stat_analysis/persistence_vs_positions.py

@@ -1,0 +1,293 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Greys',n=10)  # The blue is very beautiful but maybe harder to see.
+# Match Anna's ggplot colours
+colors.add_color(68,1,84,'Rpurple')
+colors.add_color(65,68,135,'Rblue')
+colors.add_color(42,120,142,'Rgreen1')
+colors.add_color(34,168,132,'Rgreen2')
+colors.add_color(122,209,81,'Rgreen3')
+colors.add_color(253,231,37,'Ryellow')
+colors.add_color(254,224,139,'Ryellow2')
+colors.add_color(253,174,97,'Rorange1')
+colors.add_color(244,109,67,'Rorange2')
+colors.add_color(213,62,79,'Rred1')
+colors.add_color(158,1,66,'Rred2')
+
+codes={'Three-species chain':'p.12','Omnivory':'p.38','Direct competition':'p.36','Apparent competition':'p.6'}
+
+# Each point will be mean of one network
+def read_datafile(datafile):
+  netprops={'betas':{},'pvals':{}}
+  f=open(datafile,'r')
+  for line in f:
+    if line.split()[0]!='"Position"':
+      position=line.split()[0][1:-1]
+      betas=[float(b[1:-1]) for b in line.split()[1:5]]
+      pvals=[float(p[1:-1]) for p in line.split()[5:]]
+      netprops['betas'][position]=betas
+      netprops['pvals'][position]=pvals
+  f.close()
+  return netprops
+
+def format_graph(graph,simple):
+  graph.yaxis.bar.linewidth=1
+  graph.xaxis.bar.linewidth=1
+  graph.frame.linewidth=1
+
+  graph.world.xmin=0
[…]

### 2021-06-04 code/stat_analysis/persistence_vs_profiles.py

@@ -1,0 +1,225 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Greys',n=10)  # The blue is very beautiful but maybe harder to see.
+# Match Anna's ggplot colours
+colors.add_color(68,1,84,'Rpurple')
+colors.add_color(65,68,135,'Rblue')
+colors.add_color(42,120,142,'Rgreen1')
+colors.add_color(34,168,132,'Rgreen2')
+colors.add_color(122,209,81,'Rgreen3')
+colors.add_color(253,231,37,'Ryellow')
+colors.add_color(254,224,139,'Ryellow2')
+colors.add_color(253,174,97,'Rorange1')
+colors.add_color(244,109,67,'Rorange2')
+colors.add_color(213,62,79,'Rred1')
+colors.add_color(158,1,66,'Rred2')
+
+mot_ranges={
+  'omnivory':(0.005952381,0.345547537),
+  'apparent':(0.2499348,0.7647059),
+  'direct':(0.05010438,0.35878963),
+  'chain':(0.1088861,0.4125000),
+}
+
+scales={ # center, scale
+  'omnivory':(0.169822,0.07561787),
+  'apparent':(0.4169211,0.06134048),
+  'direct':(0.1805743,0.04025804),
+  'chain':(0.2326826,0.04136131),
+  'Disturbance':(0.3,0.1366261)
+}
+
+lms={ # Intercept, motif, disturbance, interaction
+  'omnivory':(5.723e-01,-1.328e-02,-1.393e-01,-4.886e-04),
[…]

### 2021-06-04 code/stat_analysis/roles_vs_degTL_collector.py

@@ -1,0 +1,166 @@
+import sys
+import os
+import decimal
+
+
+def permfile_reader(permfile):
+	temp={}
+	f=open(permfile,'r')
+	for line in f:
+		if line.split()[0]=='"1"':
+			output=float(line.split()[1])
+	f.close()
+	return output
+
+
+def pval_calculator(model_F,perms):
+	greater=[x for x in perms if x>=model_F]
+
+	p=float(len(greater))/float(len(perms))
+
+	return p
+
+
+def recreate_aovtab(datadict,outfile):
+
+	f=open(outfile,'w')
+	f.write('S\tC\text_corr\tdf\tSS\tMS\tModel_F\tR2\tpval\n')
+	for C in sorted(datadict):
+		dats=datadict[C]
+		f.write(str(dats['S'])+'\t'+str(dats['C'])+'\t'+str(dats['ext_corr'])+'\t')
+		f.write(str(dats['df'])+'\t'+str(dats['SS'])+'\t'+str(dats['MS'])+'\t')
+		f.write(str(dats['model_F'])+'\t'+str(dats['R2'])+'\t'+str(dats['pval'])+'\n')
+	f.close()
+
+
+def betadisper_reader(dispfile):
+	f=open(dispfile,'r')
+	for line in f:
+		if line.split()[0]=='"Groups"':
+			df=int(line.split()[1])
+			SSE=float(line.split()[2])
+			MSE=float(line.split()[3])
+			F=float(line.split()[4])
+			p=float(line.split()[5])
+	f.close()
+
+	result={'F':F,'p':p}
+	return result
+
+
+def tukey_reader(tukeyfile):
+	results={}
+	f=open(tukeyfile,'r')
+	for line in f:
+		if line.split()[0]!='"diff"':
+			groups=line.split()[0][1:-1]
+			diff=float(line.split()[1])
+			lower=float(line.split()[2])
+			upper=float(line.split()[3])
+			p=float(line.split()[4])
+			results[groups]=((diff,p))
+	f.close()
+
+	return results
+
+
+def tukey_outfile(dicto,style):
+	f=open('../../data/summaries/'+style+'_Tukey.tsv','w')
+	f.write('Species\tC\tGroup1\tGroup2\tDiff\tP\n')
+
+	for s in dicto:
+		for c in dicto[s]:
[…]

### 2021-06-04 code/stat_analysis/statistical_models.R

@@ -1,0 +1,252 @@
+#### TEST MOTIFS AND PERSISTENCE 
+require(ggpmisc)
+require(tidyverse)
+
+# Libraries to fit the non-plotted lms
+library(lme4)
+library(lmerTest)
+# For permanovas and betadisper
+library(vegan)
+# For R2 GLMM
+library(MuMIn)
+
+
+data<-read.table(file = '../../data/3sp_roles_participation.tsv', sep = '\t', header = TRUE)
+data$Disturbance<-0.1+0.4*data$Basal_p
+# ONLY FIT MODELS ON CONSUMERS!!!
+consumers<-data[which(data$in_Degree>0),]
+# Check that all predictors are numeric in case of R shenanigans
+# All should be numeric except for species ID
+for(c in c((1:4),(6:25))){
+	if(!is.numeric(consumers[,c])){
+		consumers[,c]=as.numeric(as.character(consumers[,c]))
+		print(paste0(colnames(consumers)[c],' made numeric'))
+	}
+}
+
+# # Proportions
+# Creating proportions. Note that because proportions always sum to 1, doesn't make sense to do the models with all motifs together. 
+# Have to remove one motif so that no predictor is a linear combination of the others.
+consumers$total_motifs=consumers$m12+consumers$m6+consumers$m36+consumers$m38
+consumers$prop_chain=consumers$m12/consumers$total_motifs
+consumers$prop_apparent=consumers$m6/consumers$total_motifs
+consumers$prop_direct=consumers$m36/consumers$total_motifs
+consumers$prop_omni=consumers$m38/consumers$total_motifs
+
+# Section 1: How does persistence vary with global network properties?
+
+    SCper=with(consumers,lm(Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)))
+
+
+# Section 2: How do motif profiles vary with S and C?
+    motdata=read.table('../../data/3sp_motif_profiles.tsv',header=TRUE)
+    motdata$mot_total=rowSums(motdata[,4:7])
+    motdata$prop_chain=motdata$m12/motdata$mot_total
+    motdata$prop_apparent=motdata$m6/motdata$mot_total
+    motdata$prop_direct=motdata$m36/motdata$mot_total
+    motdata$prop_omni=motdata$m38/motdata$mot_total
+
+    # Expanding motif profiles for each level of disturbance
[…]

### 2021-06-04 manuscript/motifs_vs_degree.tex
removing .RData since it's too large

@@
         In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
-        Peristence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
+        Persistence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
         Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.

### 2021-06-04 manuscript/motifs_vs_degree.tex
trying to sync git to overleaf

--- 2021-06-04 manuscript/motifs_vs_degree.tex
@@
 
-    \subsection*{How does mean persistence vary with global network properties?}
+    \subsection*{How does mean persistence vary with global network structure?}
 
-        Network average persistence of consumers naturally decrease with increasing level of disturbance of basal resources (Table S1, \emph{Appendix SA}).
+        Network average persistence of consumers naturally decrease with increasing level of disturbance of basal resources (Table~\ref{tab:per_vs_SC}).
 Network size and connectance also affect persistence, but to a much smaller extent (Figure~\ref{fig:lm_CS}).
-Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance: at low disturbance small and highly-connected webs show the lowest persistence, whereas at high disturbance level large and highly connected webs show the lowest persistence (Fig.
-S1, \emph{Appendix SA}).
+Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance: at low disturbance small and highly-connected webs show the lowest persistence, whereas at high disturbance level large and highly connected webs show the lowest persistence.
 
@@
 
-    \subsection*{How do motif profiles vary with global network properties?}
+        %\begin{figure}[h!]
+        %   \includegraphics[width=0.95\textwidth]{figures/heatmap_CS_BPcompare.eps}
+        %    \caption{Heat maps of average consumer persistence in networks with varying connectance (y-axis) and network size (x-axis) at two levels of disturbance to basal resources.
+Persistence decreases with darker color in the heat map.
+When basal species are not disturbed, all species experience a baseline extinction risk of $\pi = 0.1$.
+When basal species are highly disturbed, their baseline extinction risk rises to $\pi = 0.5$ while the baseline risk of consumers is not changed.}
+        %    \label{fig:heatmap_CS}
+        %\end{figure}
 
-        A network's normalized motif profile (i.e., proportions of the total made up by each motif)
-        % ; see \emph{Appendix SX} for non-normalized motif profiles) 
-        is related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
-        Unequal variation in motif profiles across levels of size, connectance, and their interaction may, however, mean that these PERMANOVA results are false positives (Fig. S2, \emph{Appendix SB}).
+        \begin{table}[hb!]
+            \caption{Persistence of consumers within a network depends on network size (S) and connectance (C), the level of disturbance to basal resources (D), and interactions between them.
+Here we show the effect sizes ($\beta$) and $p$-values for the regression described above.
+All predictors were centered and scaled before fitting the model.
+The level of disturbance corresponds to basal resources having a probability of extinction ranging between $0.1 - 0.5$ (see \emph{Methods}).
+            }
+            \label{tab:per_vs_SC}
+            \centering
+            \begin{tabular}{c|c c |}
+                Predictor & $\beta$ & $p$-value \\
+                \hline
+                Intercept & 0.572 & \textless0.001 \\
+                Size & -5.15$\times10^{-4}$  & \textless0.001 \\
+                Connectance & -0.0129 & \textless0.001 \\
+                Disturbance & -0.139 & \textless0.001 \\
+                S:C & -5.24$\times10^{-4}$ & \textless0.001 \\
+                S:D & -3.48$\times10^{-3}$ & \textless0.001 \\
+                C:D & -7.00$\times10^{-5}$ & 0.095 \\
+                S:C:D: & -2.69$\times10^{-4}$ & \textless0.001 \\
+            \end{tabular}
+        \end{table}
 
-        The proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}).
-        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).
-The proportion of direct competition, however, decreased with increasing connectance in small networks but showed very little change in large webs.
+    \clearpage
+
+    \subsection*{How do network motif profile vary with global network properties?}
+
+        A network's normalized motif profile (i.e., proportions of the total made up by each motif; see \emph{Appendix SX} for non-normalized motif profiles) is related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
+        The proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table~\ref{network_prop_lms}).
+        Network size and the interaction between network size and connectance did not strongly affect changes in normalized motif profiles with connectance, except for direct competition (Fig.~\ref{motif_proportion_lms}).
+The proportion of direct competition decreased with increasing connectance in small networks but showed very little change in large webs.
 
@@
 
+        \begin{table}[hb!]
+            \centering
+            \caption{Coefficients for linear regressions relating the proportion of each motif in the total motif profile of a network to network size, connectance, and their interaction.
+Coefficients in \textbf{bold} are significant (\textless0.05).}
+           \label{network_prop_lms}
+           \begin{tabular}{c|c c c c c}
+                Motif & Intercept & Size & Connectance & Interaction \\
+                \hline
+                Three-species Chain & \textbf{0.262} & \textbf{1.78$\times10^{-4}$} & \textbf{-0.157} & \textbf{-3.27$\times10^{-3}$} \\
+                Apparent Competition & \textbf{0.541} & \textbf{-5.06$\times10^{-4}$} & \textbf{-1.00} & \textbf{2.59$\times10^{-3}$} \\
+                Direct Competition & \textbf{0.193} & -3.56$\times10^{-5}$ & \textbf{-0.556} & \textbf{5.92$\times10^{-3}$} \\
+                Omnivory & 4.82$\times10^{-3}$ & \textbf{3.64$\times10^{-4}$} & \textbf{1.71} & \textbf{-5.24$\times10^{-3}$}\\
+                \hline
+                \end{tabular}
+        \end{table}
+
+        %However, normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).These differences in variability can cause false positives in the PERMANOVA test.
+Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).
+
+       % \begin{figure}
+        %    \centering
+        %    \includegraphics[width=.75\textwidth]{manuscript/figures/proportion_profile_dispersion.pdf}
+        %    \caption{Dispersion of normalized motif profiles (i.e., proportions of each motif) within a network about the centroid for that combination of network size and connectance.
+Each circle represents a single network; circle colour indicates network size.
+Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
+        %    \label{dispersion_normmotifs}
+        %\end{figure}
+
+    \clearpage
+
     \subsection*{How does mean persistence vary with network motif profile?}
@@
         The motif profile of a network does relate to the mean persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
-Mean persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
-Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif.
-These trends hold for all levels of disturbance (Table S3, \emph{Appendix SC}).
+Mean persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table~\ref{motif_profile_tab}).Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif.
+These trends hold for all levels of disturbance (Table~\ref{motif_profile_tab}).
 Dispersion of motif profiles varies between networks with different levels of persistence ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
@@
 
+        \begin{table}[h!]
+            \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
+Predictors were centered and scaled before model fitting.
+All effects were significant ($p$\textless0.001).
+We also provide the $R^2$.
+}
+            \label{motif_profile_tab}
+            \centering
+            \begin{tabular}{c|c c c c c | c}
+                Motif & Intercept & Proportion & Disturbance & Interaction & $R^2$\\
+                \hline
+                Omnivory & 0.572 & -0.0133 & -0.139 & -4.89$\times10^{-4}$ & 0.908 \\
+                Apparent competition & 0.572 & 0.0105 & -0.139 & 4.78$\times10^{-4}$ & 0.905 \\
+                Direct competition & 0.572 & 2.62$\times10^{-3}$ & -0.139 & 3.84$\times10^{-4}$ & 0.901 \\
+                Three-species chain & 0.572 & 6.12$\times10^{-3}$ & -0.139 & -1.90$\times10^{-4}$ & 0.901 \\ 
+            \end{tabular}
+        \end{table}
+
+    \clearpage
+
     \subsection*{How does species persistence vary with motif participation?}
 
-        % See \emph{Appendix SY} for relationships between persistence and non-normalized motif participation.
+        See \emph{Appendix SY} for relationships between persistence and non-normalized motif participation.
 
@@
 
-            At low levels of disturbance, species' persistence increased with increasing proportions of the omnivory and direct competition motifs.
+            Without disturbance, a species' persistence increases with increasing proportion of the omnivory and direct competition motifs.
 Conversely, a species' persistence decreases with increasing proportion of the apparent competition motif.
@@
             At high levels of disturbance, a species' persistence decreases with increasing proportion of the three-species chain, but increases with increasing proportion of all other motifs (Figure~\ref{fig:prop_lmer_all}).
-            This means that increased disturbance strengthen the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table S4, \emph{Appendix SD}).
+            This means that increased disturbance strengthen the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table~\ref{tab:proportion}).
             The effect of omnivory on persistence is still positive but weaker at higher levels of disturbance, and the effect of direct competition are largely unaffected by disturbance level.
+
+            \begin{table}[hb!]
+                \centering
+                \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{propreq}).
+Predictors were centered and scaled before model fitting.
+We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
+                \label{tab:proportion}                \footnotesize
+                \begin{tabular}{c|c c | c c | c c | c c | c c |}
+                & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c|}{$R^2$} \\
+                Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
+                \hline
+                3-sp Chain & 0.574 & \textless0.001 & -0.00508 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
+                App. Com. & 0.574 & \textless0.001 & -0.00312 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
+                Dir. Com. & 0.574 & \textless0.001 & 0.00564 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
+                Omnivory & 0.574 & \textless0.001 & 0.00709 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
+                \end{tabular}
+            \end{table}
 
@@
 
-            By fitting linear regressions of persistence against motif proportions for the species for each network separately, we can observe how the relationships between motifs and persistence change with the level of disturbance and global network structure.
-Network size had a much smaller effect than connectance (Fig.
-S3; \emph{Appendix SE}); we therefore focus on connectance here.
-            Networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
-            However, there was an interaction between disturbance and connectance for the direct competition and omnivory motifs.
+            By fitting linear regressions of persistence against motif proportions for the species for each network separately, we can extract the slopes and analyze their distribution.
+This reveals information on how the relationship change with the level of disturbance and global network structure.
+For the global network parameters the network size had a much smaller effect than connectance (Fig.~\ref{fig:density_prop_S}; \emph{Appendix SZ})and therefore we here focus on only connectance.
 
+            Networks with high connectance generally have more similar relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
+However, the interaction between a species' participation in the various motifs and disturbance changes with connectance  (Figure~\ref{fig:density_prop_C}).
+            This is in particular evident for the direct competition and omnivory motif.
 At low and medium connectances there is little change in the proportion of positive slopes (\textgreater80\%) for the different levels of disturbance.
-            At high connectances, however, the proportion of positive slopes is lower overall and decreases with increasing disturbance. 
-            This means that at low levels of connectance, an increased proportion of the direct competition motif increases persistence in most networks, while the opposite is true with high connectance.
+However, at high connectances the proportion of positive slopes is overall lower and also decreases with increasing disturbance.
+This means that at low levels of connectance, in a majority of cases an increased proportion of the direct competition motif increase persistence, while the opposite is true with high connectance.
             The trend is reversed for the omnivory motif.
+            Overall, the proportion of networks showing a positive effect of omnivory decreases with increasing disturbance.
+            This decrease is strongest at low and moderate connectances but very weak at high connectance.
+(Figure~\ref{fig:density_prop_C}).
+Notably, the distribution of slopes for the omnivory motif is wider (effect is more variable) the lower the connectance is.
+            Hence, the level of connectance influence whether a high proportion of the omnivory motif is associated with higher persistence or not.
 
@@
 
+        \begin{figure}[h!]
+             \centering
+             \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
+            \caption{Density (y-axis) of slopes (x-axis) for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
+The slopes are derived from standard linear regression.
+A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
+The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to motif (legend).
+Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
+Rows show various levels of network size.
+The vertical line indicate zero on the x-axis.}
+            \label{fig:density_prop_S}
+        \end{figure}
+
+    \clearpage
+
     \subsection*{How does persistence vary with degree and trophic level?}
 
-        Persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
-Increased trophic level is associated with decreasing persistence, although the rate of this decline is slower at higher levels of disturbance (Table S5, \emph{Appendix SF}).
+        Persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SZZ}).
+Increased trophic level is associated with decreasing persistence, although the rate of this decline is slower at higher levels of disturbance (Table~\ref{tab:per_vs_TLdeg}).
         The interaction between in-degree and disturbance is even stronger, such as having a large number of prey is associated with higher persistence in the no disturbance scenario, but lower persistence at high disturbance.
+
+        \begin{table}[hb!]
+            \caption{Coeffcients ($\beta$) for regressions of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction.
+Predictors were centered and scaled before model fitting.
+All $p$-values were \textless0.001.
+The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.}
+            \label{tab:per_vs_TLdeg}
+            \centering
+            \begin{tabular}{l|c  c  c |}
+                Simple role measure & Role & Disturbance & Interaction \\
+                \hline
+                STL & -0.0218 & -0.139 & 0.00508 \\
+                in-degree & -0.00130 & -0.139 & -0.00908 \\
+            \end{tabular}
+        \end{table}
 
@@
 
+    \clearpage
+
     \subsection*{Do simpler measures explain effects of motifs?}
 
-        Motif participation varied with web size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
-        Species in all networks appeared most frequently in the apparent competition motif, but the frequency of the apparent competition motif decreased with increasing network size and connectance.
-        The frequency of the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size. 
+        \subsubsection*{Motif participation and network size and connectance}
+
+            \begin{table}[hb!]
+            \centering
+            \caption{The meso-scale composition of a network varies with size and connectance.
+Here we give the range and mean ($\pm$ standard deviation) proportion of each motif in species' motif participation vectors.
+            We also show the effects ($\beta$) of network size, connectance, and their interaction in linear models predicting the proportion of each motif in a participation vector.
+            All effects were significant ($p$\textless0.001).
+            Predictors were not centred or standardized before model fitting. Motifs are ordered according to mean proportion.}
+            % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
+            \label{tab:partic_vs_SC}   
+            \footnotesize
+            \begin{tabular}{c|c c c | c c c}
+                Motif & Min & Max & Mean (SD) & Size & Connectance & Interaction \\
+                \hline
+                App. Comp & 0.250 & 0.765 & 0.422 (0.0645) & -5.06$\times10^{-4}$ & -1.00 & 2.59$\times10^{-3}$ \\
+                Chain & 0.109 & 0.413 & 0.235 (0.042) & 1.78$\times10^{-4}$ & -0.157 & -3.27$\times10^{-3}$ \\
+                Dir. Comp & 0.0501 & 0.359 & 0.179 (0.0407) & -3.56$\times10^{-5}$ & -0.556 & 5.92$\times10^{-3}$ \\
+                Omnivory & 0.006 & 0.346 & 0.164 (0.0790) & 3.64$\times10^{-4}$ & 1.71 & -5.24$\times10^{-3}$ \\   
+                \hline
+                \end{tabular}
+                \end{table}
+
+        The numbers of different motifs in a network may vary with web size, connectance, and their interaction.
+        To establish this context, we fit linear models relating the proportion of each motif in a species' role against network size, connectance, and their interaction.
+
+        \subsubsection*{Motif participation and simple role measures}
+
             Species' motif roles also vary with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
@@
             \end{figure}
+
+\clearpage
+
+    [[Big deal result: connectance is important for motif structure of network, size not so much]]
+
+    [[Persistence depends on S, C, interaction with $basal_p$ (lowest persistence swaps from highest S to lowest C in Figs.
+1-2, high C in both cases - but small absolute differences so maybe we don't want to make too big a deal out of this)]]
+
+    [[proportion of omnivory increases with increasing C, while the other doesn't. Connect to later findings involving omnivory]]
+
+\clearpage
 
@@
 The top set of lines corresponds to a probability of extinction for basal resources of 0.1.
-The middle set of lines corresponds to a probability of extinction for basal resources of 0.26, and the lowest set of lines corresponds to a probability of extinction for basal resources of 0.5.}
+The middle set of lines correspons to a probability of extinction for basal resources of 0.26, and the lowest set of lines corresponds to a probability of extinction for basal resources of 0.5.}
             \label{fig:persistence_positions_prey}
@@
 
-        \subsubsection*{Degree and trophic level}
-
-            The top position in the omnivory motif increased with increasing degree across all trophic levels.
-            The frequency of the middle position increased with the interaction of degree and trophic level.
-            The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).
-
-            The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
-            The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.
-
-            The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
-            The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).
-
-            The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
-            The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
-            The top position in the three-species chain increased with trophic level, degree, and their interaction.
-
-        \subsubsection*{Network size and connectance}
-
-            % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
-            The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance.
-            The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.
-
-            The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
-            The frequency of the bottom position in the direct competition motif was low in all networks.
-
-            The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
-            Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
-            None of the positions in the three-species chain varied strongly with network size or connectance.
+        [[have fit these models but no figure yet - seems like degree and the 2-prey positions show the same pattern so they're probably tightly linked.
+STL may explain the differences in top and middle positions for the chain.
+There's still a fair amount of differences among the 0 prey and 1 prey motifs though, so motifs still provide unique information that degree and STL don't.]]
 

### 2021-06-04 manuscript/supplement_motifs.tex
trying to sync git to overleaf

@@
 
-\section*{Table of contents}
-    \begin{spacing}{1.0}
+% \section*{Table of contents}
+%     \begin{spacing}{1.0}
 
-    \begin{enumerate}
+%     \begin{enumerate}
 
-        \item How do counts of motifs in a network profile vary with S and C? \\
+%         \item How do counts of motifs in a network profile vary with S and C? \\
 
-            \begin{itemize}
-            \item Same methods as main text except using counts instead of proportions to define motif profiles
-            \end{itemize}
+%             \begin{itemize}
+%             \item Same methods as main text except using counts instead of proportions to define motif profiles
+%             \end{itemize}
 
-        \item Motif profiles differed slightly between network sets
+%         \item Motif profiles differed slightly between network sets
 
-           \begin{itemize}
-                \item Comparing proportions of motifs in un-filtered, filtered, and acyclic networks.
-                \item AFAIK not referred to in text
-            \end{itemize}
+%            \begin{itemize}
+%                 \item Comparing proportions of motifs in un-filtered, filtered, and acyclic networks.
+%                 \item AFAIK not referred to in text
+%             \end{itemize}
 
-         \item Persistence and count-based motif participation
+%          \item Persistence and count-based motif participation
 
-           \begin{itemize}
-                \item Should be as main text section 4, but counts
-            \end{itemize}
+%            \begin{itemize}
+%                 \item Should be as main text section 4, but counts
+%             \end{itemize}
 
-        \item Motif-persistence relationships and network size
-            \begin{itemize}
-                \item Detailed results for variability of slopes for persistence $\approx$ motif participation in networks of different sizes.
-                \item Main text refers to C, moved figure to Main while we check that we've got everything.
-            \end{itemize}
+%         \item Motif-persistence relationships and network size
+%             \begin{itemize}
+%                 \item Detailed results for variability of slopes for persistence $\approx$ motif participation in networks of different sizes.
+%                 \item Main text refers to C, moved figure to Main while we check that we've got everything.
+%             \end{itemize}
 
-        \item How does persistence vary with degree and trophic level?
+%         \item How does persistence vary with degree and trophic level?
 
-            \begin{itemize}
-                \item Heat map showing persistence vs. TL with BP=0
-                \item Heat map showing persistence vs. TL with BP=1
-            \end{itemize}
+%             \begin{itemize}
+%                 \item Heat map showing persistence vs. TL with BP=0
+%                 \item Heat map showing persistence vs. TL with BP=1
+%             \end{itemize}
 
-    \end{enumerate}
-\end{spacing}
+%     \end{enumerate}
+% \end{spacing}
+% \clearpage
+
+\section{SA: Persistence and global network properties}
+
+    Connectance and species richness had smaller effects on consumer persistence than disturbance (Table~\ref{tab:per_vs_SC}, Fig.~\ref{fig:heatmap_CS}).
+
+    \begin{figure}[h!]
+      \includegraphics[width=0.95\textwidth]{figures/heatmap_CS_BPcompare.eps}
+       \caption{Heat maps of average consumer persistence in networks with varying connectance (y-axis) and network size (x-axis) at two levels of disturbance to basal resources.
+Persistence decreases with darker color in the heat map.
+When basal species are not disturbed, all species experience a baseline extinction risk of $\pi = 0.1$.
+When basal species are highly disturbed, their baseline extinction risk rises to $\pi = 0.5$ while the baseline risk of consumers is not changed.}
+       \label{fig:heatmap_CS}
+    \end{figure}
+
+    \begin{table}[hb!]
+        \caption{Persistence of consumers within a network depends on network size (S) and connectance (C), the level of disturbance to basal resources (D), and interactions between them.
+Here we show the effect sizes ($\beta$) and $p$-values for the regression described above.
+All predictors were centered and scaled before fitting the model.
+The level of disturbance corresponds to basal resources having a probability of extinction ranging between $0.1 - 0.5$ (see \emph{Methods}).
+        }
+        \label{tab:per_vs_SC}
+        \centering
+        \begin{tabular}{c|c c |}
+            Predictor & $\beta$ & $p$-value \\
+            \hline
+            Intercept & 0.572 & \textless0.001 \\
+            Size & -5.15$\times10^{-4}$  & \textless0.001 \\
+            Connectance & -0.0129 & \textless0.001 \\
+            Disturbance & -0.139 & \textless0.001 \\
+            S:C & -5.24$\times10^{-4}$ & \textless0.001 \\
+            S:D & -3.48$\times10^{-3}$ & \textless0.001 \\
+            C:D & -7.00$\times10^{-5}$ & 0.095 \\
+            S:C:D: & -2.69$\times10^{-4}$ & \textless0.001 \\
+        \end{tabular}
+    \end{table}
+
 \clearpage
 
-\section{How do counts of motifs in a network profile vary with S and C?}
+\section{SB: Motif profiles and global properties}
 
-    As expected, all four motifs had higher counts in larger and more highly connected networks.
-    The precise effects of size, connectance, and their interaction differed between motifs  (Table~\ref{network_motif_lms}).
-    This suggests that the meso-scale structure of a network varies with size and connectance, not just the total number of motifs.
-    Indeed, the motif profile of a network is significantly related to its size, connectance, and their interaction ($F_{1,2996}$=2279, $p$\textless0.001; $F_{1,2996}$=3067, $p$\textless0.001; and $F_{1,2996}$=1187, $p$\textless0.001, respectively).
+    Normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).These differences in variability can cause false positives in the PERMANOVA test.
+Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).
+
+   \begin{figure}[h!]
+       \centering
+       \includegraphics[width=.75\textwidth]{figures/proportion_profile_dispersion.eps}
+       \caption{Dispersion of normalized motif profiles (i.e., proportions of each motif) within a network about the centroid for that combination of network size and connectance.
+Each circle represents a single network; circle colour indicates network size.
+Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
+       \label{dispersion_normmotifs}
+    \end{figure}
+
+    The frequencies of the four possible motifs were related to global network properties.
+
+    \begin{table}[hb!]
+        \centering
+        \caption{Coefficients for linear regressions relating the proportion of each motif in the total motif profile of a network to network size, connectance, and their interaction.
+Coefficients in \textbf{bold} are significant (\textless0.05).}
+       \label{network_prop_lms}
+       \begin{tabular}{c|c c c c c}
+            Motif & Intercept & Size & Connectance & Interaction \\
+            \hline
+            Three-species Chain & \textbf{0.262} & \textbf{1.78$\times10^{-4}$} & \textbf{-0.157} & \textbf{-3.27$\times10^{-3}$} \\
+            Apparent Competition & \textbf{0.541} & \textbf{-5.06$\times10^{-4}$} & \textbf{-1.00} & \textbf{2.59$\times10^{-3}$} \\
+            Direct Competition & \textbf{0.193} & -3.56$\times10^{-5}$ & \textbf{-0.556} & \textbf{5.92$\times10^{-3}$} \\
+            Omnivory & 4.82$\times10^{-3}$ & \textbf{3.64$\times10^{-4}$} & \textbf{1.71} & \textbf{-5.24$\times10^{-3}$}\\
+            \hline
+            \end{tabular}
+    \end{table}        
+\clearpage
+
+\section{SC: Persistence and motif profiles}
+
+    Persistence varied with the frequencies of different motifs in a network profile.
+
+    \begin{table}[h!]
+        \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
+Predictors were centered and scaled before model fitting.
+All effects were significant ($p$\textless0.001).
+We also provide the $R^2$.
+}
+        \label{motif_profile_tab}
+        \centering
+        \begin{tabular}{c|c c c c c | c}
+            Motif & Intercept & Proportion & Disturbance & Interaction & $R^2$\\
+            \hline
+            Omnivory & 0.572 & -0.0133 & -0.139 & -4.89$\times10^{-4}$ & 0.908 \\
+            Apparent competition & 0.572 & 0.0105 & -0.139 & 4.78$\times10^{-4}$ & 0.905 \\
+            Direct competition & 0.572 & 2.62$\times10^{-3}$ & -0.139 & 3.84$\times10^{-4}$ & 0.901 \\
+            Three-species chain & 0.572 & 6.12$\times10^{-3}$ & -0.139 & -1.90$\times10^{-4}$ & 0.901 \\ 
+        \end{tabular}
+    \end{table}
+\clearpage
+
+\section{SD: Persistence and motif participation}
+
+    Persistence varied with motif participation.
 
@@
         \centering
-        \caption{Coefficients for linear regressions relating the total count of all motifs or the count of each three-species motif in a network to network size, connectance, and their interaction.
-Coefficients in \textbf{bold} are significant (\textless0.05).
-Note that, while there are 13 unique three-species motifs, only four are possible in acyclic networks such as those considered here.}
-       \label{network_motif_lms}
-       \begin{tabular}{c|c c c c c}
-        Motif & Intercept & Size & Connectance & Interaction \\
+        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{propreq}).
+Predictors were centered and scaled before model fitting.
+We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
+        \label{tab:proportion}                \footnotesize
+        \begin{tabular}{c|c c | c c | c c | c c | c c |}
+        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c|}{$R^2$} \\
+        Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
         \hline
-        Total count & \textbf{-2.85$\times10^3$} & \textbf{46.6} & \textbf{-2.12$\times10^{5}$} & \textbf{4.72$\times10^{3}$} \\
+        3-sp Chain & 0.574 & \textless0.001 & -0.00508 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
+        App. Com. & 0.574 & \textless0.001 & -0.00312 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
+        Dir. Com. & 0.574 & \textless0.001 & 0.00564 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
+        Omnivory & 0.574 & \textless0.001 & 0.00709 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
+        \end{tabular}
+    \end{table}        
+\clearpage
+
+\section{SE: Motif-persistence relationships \& network size}
+
+    When comparing the outcomes from linear mixed-effect models (LMMs) relating persistence to motif roles across all networks (Fig.
+3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately, the main trends are similar but different-sized networks showed slightly different results results for some motifs (Figure temporarily in main text).%(Figure~\ref{fig:density_prop_S}).
+    For the three-species chain and omnivory motifs, a significant interaction between disturbance and the effect of the focal motif means that having a larger proportion of the role be made up of the chain or omnivory motif is associated with lower persistence at high levels of disturbance (Fig.
+3 and Table 1, \emph{Main Text}).
+    The same pattern holds overall for LRs fit within a network  %(Figure~\ref{fig:density_prop_S})
+    : the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance. 
+    This trend is most pronounced in large networks.% (bottom panel, Figure~\ref{fig:density_prop_S}). 
+    Besides the decrease in fraction of positive slopes with increasing disturbance, the omnivory motif show a higher overall fraction of positive slopes in comparison to the three-species chain motif, corresponding to the positive main effect of omnivory (Fig.
+3 and Table 1, \emph{Main Text}).
+
+    \begin{figure}[h!]
+         \centering
+         % \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
+        \caption{Density (y-axis) of slopes (x-axis) for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
+The slopes are derived from linear mixed-effect models, with a random intercept for the interaction between species richness and network size.
+A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
+The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to motif (legend).
+Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
+Rows show various levels of network size.
+The vertical line indicate zero on the x-axis.}
+        \label{fig:density_prop_S}
+    \end{figure}
+
+    For the apparent and direct competition motifs, the LMMs include a positive interaction between disturbance and the proportion of the motif in a species' role (Fig.
+3 and Table 1; \emph{Main Text}).
+    This means that having more of the role made up of these two motifs is increasingly beneficial at higher levels of disturbance, a trend which is also evident from the increasing fraction of positive slopes with increasing disturbance based on LRs of each network at each level of disturbance.
+% (Figure~\ref{fig:density_prop_S}).
+    As with the trends for omnivory and three-species chains, these increases are strongest in large networks. % (bottom row, Figure~\ref{fig:density_prop_S}).
+    % For direct competition, the fraction of slopes are similar across disturbance levels.
+Although this interaction is not positive when accounting for network size, the fraction of positive slopes are throughout high.
+    The increase in the proportion of positive slopes was much stronger for apparent competition than direct competition, consistent with the smaller (but still significant) interaction term for between direct competition in the LMMs.
+    A high proportion of direct competition in a species' role tended to be associated with greater probability of persistence across levels of disturbance and network size.
+    % In conclusion, species richness does not considerably alter the trends visible in the LMMs.
+High species richness more clearly display an interaction between motif proportion, persistences and disturbance levels, while this effect is smaller with low species richness.
+\clearpage
+
+\section{SF: Persistence, degree, and trophic level}
+
+    Persistence varies with in-degree, trophic level, and their interaction.
+
+    \begin{table}[h!]
+        \caption{Coeffcients ($\beta$) for regressions of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction.
+Predictors were centered and scaled before model fitting.
+All $p$-values were \textless0.001.
+The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.}
+        \label{tab:per_vs_TLdeg}
+        \centering
+        \begin{tabular}{l|c  c  c |}
+            Simple role measure & Role & Disturbance & Interaction \\
             \hline
-        Thee-species Chain & \textbf{-1.40$\times10^{3}$} & \textbf{26.1} & \textbf{-3.74$\times10^{4}$} & \textbf{8.58$\times10^{2}$} \\
-        Apparent Competition & \textbf{-2.23$\times10^3$} & \textbf{42.0} & 
-        \textbf{-7.26$\times10^4$} & \textbf{1.63$\times10^3$} \\
-        Direct Competition & 26.6 & 0.777 & \textbf{-4.85$\times10^4$} & \textbf{9.89$\times10^2$} \\
-        Omnivory & \textbf{7.57} & \textbf{-22.2} & \textbf{-5.37$\times10^4$} & \textbf{1.24$\times10^3$} \\
+            STL & -0.0218 & -0.139 & 0.00508 \\
+            in-degree & -0.00130 & -0.139 & -0.00908 \\
+        \end{tabular}
+    \end{table}
+
+    \begin{figure}[h!]
+     \centering
+     % \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}
+     \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
+The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
+Persistence decreases with darker color in the heat map.}
+     \label{fig:heatmap_stl_BP0}
+    \end{figure}
+
+    \begin{figure}[h!]
+     \centering
+     % \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
+     \caption{Primary producers experience a disturbance reducing their probability to go extinct to $\pi_{disturbed} = 0.5$ All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
+The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
+Persistence decreases with darker color in the heat map.}
+     \label{fig:heatmap_stl_BP1}
+    \end{figure}
+\clearpage
+
+\section{SG: Motif participation and network properties}
+
+    Motif participation varies with global-scale network properies.
+
+    \begin{table}[hb!]
+        \centering
+        \caption{Motif participation varies with size and connectance. Here we give the range and mean ($\pm$ standard deviation) proportion of each motif in species' motif participation vectors. 
+        We also show the effects ($\beta$) of network size, connectance, and their interaction in linear models predicting the proportion of each motif in a participation vector.
+        All effects were significant ($p$\textless0.001).
+        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
+        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
+        \label{tab:partic_vs_SC}   
+        \footnotesize
+        \begin{tabular}{c|c c c | c c c}
+            Motif & Min & Max & Mean (SD) & Size & Connectance & Interaction \\
+            \hline
+            3-sp Chain & 0 & 1 & 0.249 (0.127) & 2.22$\times10^{-4}$ & -0.131 & -3.20$\times10^{-3}$ \\
+            App. Comp & 0.250 & 0.765 & 0.422 (0.0645) & -5.06$\times10^{-4}$ & -1.00 & 2.59$\times10^{-3}$ \\
+            Dir. Comp & 0.0501 & 0.359 & 0.179 (0.0407) & -3.56$\times10^{-5}$ & -0.556 & 5.92$\times10^{-3}$ \\
+            Omnivory & 0.006 & 0.346 & 0.164 (0.0790) & 3.64$\times10^{-4}$ & 1.71 & -5.24$\times10^{-3}$ \\   
             \hline
@@
         \centering
-        \includegraphics[width=.75\textwidth]{figures/motif_profile_dispersion.pdf}
-        \caption{Dispersion of motif profiles within a network about the centroid for that combination of network size and connectance.
-Each circle represents a single network; circle colour indicates network size.
-Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
-        \label{dispersion_rawmotifs}
+        \includegraphics[width=.95\textwidth]{figures/roles_vs_SC.eps}
+
+        \caption{The proportion of each motif in a species' participation vector varied with network size, connectance, and their interaction.
+These relationships were steepest for apparent competition and omnivory.}
+        \label{fig:roles_vs_SC}
     \end{figure}
+\clearpage
 
-    Motif profiles were similarly variable across networks with different sizes, but were not homogeneously variable across connectances or size-connectance combinations ($F_{5,2994}$=0.365, $p$=0.873; $F_{4,2995}$=4.49, $p$=0.001; and $F_{29,2970}$=27.1, $p$\textless0.001 , respectively).
-    These differences in variability of motif profiles can cause false positives in the PERMANOVA.
-    Variability in motif profiles was lowest among large networks with moderate connectance and was highest among the least-connected and most-connected networks we considered (Fig.~\ref{dispersion_rawmotifs}).
+\section*{SH: Motif roles and network properties}
+
+    \subsection*{Degree and trophic level}
+
+        \begin{table}[h!]
+            \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against in-degree, STL, and their interaction.}
+            \label{DT_tab}
+            \footnotesize
+            \begin{tabular}{c c | c c | cc | cc |}
+            Motif & Position & \multicolumn{2}{c}{Degree} & \multicolumn{2}{c}{STL} & \multicolumn{2}{c}{Interaction} \\
+            & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
+            \hline
+            Chain & bottom & -1.47 $\times10^{-5}$ & 0.769 & -0.0218 & \textless0.001 & -0.00160 & \textless0.001 \\
+            Chain & middle & 3.13$\times10^{-4}$ & \textless0.001 & -0.0353 & \textless0.001 & 1.85$\times10^{-5}$ & 0.275 \\
+            Chain & top & -0.0172 & \textless0.001 & 0.0259 & \textless0.001 & 0.00919 & \textless0.001 \\
+            \hline
+            Omnivory & bottom & 6.14$\times10^{-4}$ & \textless0.001 & -0.00270 & \textless0.001 & -0.00112 & \textless0.001 \\
+            Omnivory & middle & 3.13$\times10^{-4}$ & \textless0.001 & -0.0353 & \textless0.001 & 1.85$\times10^{-5}$ & \textless0.001 \\
+            Omnivory & top & 0.00413 & \textless0.001 & 0.00381 & \textless0.001 & 9.96$\times10^{-5}$ & \textless0.001 \\
+            \hline
+            App. comp. & bottom & -4.74$\times10^{-4}$ & \textless0.001 & -0.663 & \textless0.001 & 0.00224 & \textless0.001 \\
+            App. comp. & top & 2.12$\times10^{-5}$ & 0.183 & -0.112 & \textless0.001 & -4.47$\times10^{-4}$ & \textless0.001 \\
+            \hline
+            Dir. comp. & bottom & 8.12$\times10^{-4}$ & \textless0.001 & -5.57$\times10^{-3}$ & \textless0.001 & -1.37$\times10^{-3}$ & \textless0.001 \\
+            Dir. comp. & top & -0.00405 & \textless0.001 & -0.0435 & \textless0.001 & 0.00260 & \textless0.001 \\
+            \hline
+            \end{tabular}
+            \end{table}
 
@@
             \centering
-        \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
-        \caption{The effect of number of times a species appears (x-axis) in any of the various motifs (columns) on persistence (y-axis).
-The effect of each motif participation is fitted by linear mixed-effect models, following Equation~\ref{absreq}, except that the disturbance levels are fitted separately.
-The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
-        \label{fig:abs_lmer_all}
+            \includegraphics[width=.75\textwidth]{figures/positions_byTL_Omnivory.eps}
+            \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
+            \label{omnivory_DT}
+        \end{figure}
+
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=.75\textwidth]{figures/positions_byTL_Apparentcompetition.eps}
+            \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
+            \label{appcomp_DT}
+        \end{figure}
+
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=.75\textwidth]{figures/positions_byTL_Directcompetition.eps}
+            \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
+            \label{dircomp_DT}
+        \end{figure}
+
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=.75\textwidth]{figures/positions_byTL_Three-specieschain.eps}
+            \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
+            \label{appcomp_DT}
+        \end{figure}
+
+    \subsection*{Network size and connectance}
+
+        \begin{table}[h!]
+            \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against network size, connectance, and their interaction.}
+            \label{DT_tab}
+            \footnotesize
+            \begin{tabular}{c c | c c | cc | cc |}
+            Motif & Position & \multicolumn{2}{c}{Size} & \multicolumn{2}{c}{Connectance} & \multicolumn{2}{c}{Interaction} \\
+            & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
+            \hline
+            Chain & bottom & -4.63$\times10^{-5}$ & \textless0.001 & -0.108 & \textless0.001 & -1.94$\times10^{-4}$ & 0.035 \\
+            Chain & middle & 1.68$\times10^{-4}$ & \textless0.001 & -0.0530 & \textless0.001 & -0.00212 & \textless0.001 \\
+            Chain & top & 1.00$\times10^{-4}$ & \textless0.001 & 0.0304 & 0.015 & -8.79$\times10^{-4}$ & \textless0.001 \\
+            \hline
+            Omnivory & bottom & 2.00$\times10^{-4}$ & \textless0.001 & 0.575 & \textless0.001 & -0.00240 & \textless0.001 \\
+            Omnivory & middle & 1.32$\times10^{-4}$ & \textless0.001 & 0.551 & \textless0.001 & -0.00215 & \textless0.001 \\
+            Omnivory & top & 6.42$\times10^{-5}$ & \textless0.001 & 0.448 & \textless0.001 & -0.00115 & \textless0.001 \\
+            \hline
+            App. comp. & bottom & -4.74$\times10^{-4}$ & \textless0.001 & -0.663 & \textless0.001 & 0.00224 & \textless0.001 \\
+            App. comp. & top & 2.12$\times10^{-5}$ & 0.183 & -0.112 & \textless0.001 & -4.47$\times10^{-4}$ & \textless0.001 \\
+            \hline
+            Dir. comp. & bottom & 7.94$\times10^{-5}$ & \textless0.001 & -0.0663 & \textless0.001 & 0.00101 & \textless0.001 \\
+            Dir. comp. & top & -2.45$\times10^{-4}$ & \textless0.001 & -0.602 & \textless0.001 & 0.00610 & \textless0.001 \\
+            \hline
+            \end{tabular}
+            \end{table}
+
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=.75\textwidth]{figures/positions_bySC_Omnivory.eps}
+            \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
+            \label{omnivory_DT}
+        \end{figure}
+
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=.75\textwidth]{figures/positions_bySC_Apparentcompetition.eps}
+            \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
+            \label{appcomp_DT}
+        \end{figure}
+
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=.75\textwidth]{figures/positions_bySC_Directcompetition.eps}
+            \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
+            \label{dircomp_DT}
+        \end{figure}
+
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=.75\textwidth]{figures/positions_bySC_Three-specieschain.eps}
+            \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
+            \label{appcomp_DT}
         \end{figure}
@@
 
-\section{Motif profiles differed slightly in different network sets}
+% \section{How do counts of motifs in a network profile vary with S and C?}
+
+%     As expected, all four motifs had higher counts in larger and more highly connected networks.
+%     The precise effects of size, connectance, and their interaction differed between motifs  (Table~\ref{network_motif_lms}).
+%     This suggests that the meso-scale structure of a network varies with size and connectance, not just the total number of motifs.
+%     Indeed, the motif profile of a network is significantly related to its size, connectance, and their interaction ($F_{1,2996}$=2279, $p$\textless0.001; $F_{1,2996}$=3067, $p$\textless0.001; and $F_{1,2996}$=1187, $p$\textless0.001, respectively).
+
+%     \begin{table}[h!]
+%         \centering
+%         \caption{Coefficients for linear regressions relating the total count of all motifs or the count of each three-species motif in a network to network size, connectance, and their interaction.
+Coefficients in \textbf{bold} are significant (\textless0.05).
+Note that, while there are 13 unique three-species motifs, only four are possible in acyclic networks such as those considered here.}
+%        \label{network_motif_lms}
+%        \begin{tabular}{c|c c c c c}
+%         Motif & Intercept & Size & Connectance & Interaction \\
+%         \hline
+%         Total count & \textbf{-2.85$\times10^3$} & \textbf{46.6} & \textbf{-2.12$\times10^{5}$} & \textbf{4.72$\times10^{3}$} \\
+%         \hline
+%         Thee-species Chain & \textbf{-1.40$\times10^{3}$} & \textbf{26.1} & \textbf{-3.74$\times10^{4}$} & \textbf{8.58$\times10^{2}$} \\
+%         Apparent Competition & \textbf{-2.23$\times10^3$} & \textbf{42.0} & 
+%         \textbf{-7.26$\times10^4$} & \textbf{1.63$\times10^3$} \\
+%         Direct Competition & 26.6 & 0.777 & \textbf{-4.85$\times10^4$} & \textbf{9.89$\times10^2$} \\
+%         Omnivory & \textbf{7.57} & \textbf{-22.2} & \textbf{-5.37$\times10^4$} & \textbf{1.24$\times10^3$} \\
+%         \hline
+%         \end{tabular}
+%     \end{table}
+
+%     \begin{figure}[h!]
+%         \centering
+%         \includegraphics[width=.75\textwidth]{figures/motif_profile_dispersion.pdf}
+%         \caption{Dispersion of motif profiles within a network about the centroid for that combination of network size and connectance.
+Each circle represents a single network; circle colour indicates network size.
+Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
+%         \label{dispersion_rawmotifs}
+%     \end{figure}
+
+%     Motif profiles were similarly variable across networks with different sizes, but were not homogeneously variable across connectances or size-connectance combinations ($F_{5,2994}$=0.365, $p$=0.873; $F_{4,2995}$=4.49, $p$=0.001; and $F_{29,2970}$=27.1, $p$\textless0.001 , respectively).
+%     These differences in variability of motif profiles can cause false positives in the PERMANOVA.
+%     Variability in motif profiles was lowest among large networks with moderate connectance and was highest among the least-connected and most-connected networks we considered (Fig.~\ref{dispersion_rawmotifs}).
+
+%     \begin{figure}[h!]
+%         \centering
+%         \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
+%         \caption{The effect of number of times a species appears (x-axis) in any of the various motifs (columns) on persistence (y-axis).
+The effect of each motif participation is fitted by linear mixed-effect models, following Equation~\ref{absreq}, except that the disturbance levels are fitted separately.
+The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
+%         \label{fig:abs_lmer_all}
+%     \end{figure}
+
+\section{Motif profiles in different network sets}
 
@@
 
+    \begin{table}[h!]
+        \centering
+        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{propreq}).
+Predictors were centered and scaled before model fitting.
+We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
+        \label{tab:proportion}                \footnotesize
+        \begin{tabular}{c|c c | c c | c c | c c | c c |}
+        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c|}{$R^2$} \\
+        Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
+        \hline
+        3-sp Chain & 0.574 & \textless0.001 & -0.00508 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
+        App. Com. & 0.574 & \textless0.001 & -0.00312 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
+        Dir. Com. & 0.574 & \textless0.001 & 0.00564 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
+        Omnivory & 0.574 & \textless0.001 & 0.00709 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
+        \end{tabular}
+    \end{table}        
 \clearpage
 
-\section{Persistence and count-based motif participation}
+% \section{Persistence and count-based motif participation}
 
-    \subsection*{Methods}
+%     \subsection*{Methods}
 
-        We test whether participating in higher absolute numbers (i.e., counts) of a motif increases persistence using a set of models that parallels those used to test whether having a greater proportion of the role made up by a particular motif (see \emph{Main Text}).
-        Specifically, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
-        In each case, we considered only non-basal species (i.e., those with at least one prey) as the persistence of basal species is determined only by their baseline probability.
-        These models included the absolute number of times the focal species appeared in the focal motif, the increase in probability of extinction for basal species, and the interaction between the two as well as a random intercept for the interaction between species richness and connectance (equations.
-        All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-        We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.
+%         We test whether participating in higher absolute numbers (i.e., counts) of a motif increases persistence using a set of models that parallels those used to test whether having a greater proportion of the role made up by a particular motif (see \emph{Main Text}).
+%         Specifically, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
+%         In each case, we considered only non-basal species (i.e., those with at least one prey) as the persistence of basal species is determined only by their baseline probability.
+%         These models included the absolute number of times the focal species appeared in the focal motif, the increase in probability of extinction for basal species, and the interaction between the two as well as a random intercept for the interaction between species richness and connectance (equations.
+%         All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+%         We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.
 
-            The formula for these LMMs are:         
-            \begin{equation}
-                \Psi_{ik} \approx \chi_{i} + \pi_{disturbed_k} + \chi_{i}\pi_{disturbed_k} + S:C_{i} ,
-                \label{absreq}
-            \end{equation}
+%             The formula for these LMMs are:         
+%             \begin{equation}
+%                 \Psi_{ik} \approx \chi_{i} + \pi_{disturbed_k} + \chi_{i}\pi_{disturbed_k} + S:C_{i} ,
+%                 \label{absreq}
+%             \end{equation}
 
-            \noindent where $\Psi_ik$ is the persistence of species $i$ during disturbance level $k$, $\chi_{i}$ is the number of times species $i$ appears in the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
+%             \noindent where $\Psi_ik$ is the persistence of species $i$ during disturbance level $k$, $\chi_{i}$ is the number of times species $i$ appears in the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
 
-            % Additionally, we performed a similar linear regression of persistence against motif participation within each simulated network and for each level of disturbance to basal resources (XX regressions total).
-            % These regressions were fit using the R~\citep{R} base function `lm'.
-            % We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.
+%             % Additionally, we performed a similar linear regression of persistence against motif participation within each simulated network and for each level of disturbance to basal resources (XX regressions total).
+%             % These regressions were fit using the R~\citep{R} base function `lm'.
+%             % We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.
 
-    \subsection*{Results}
-
-        \begin{figure}[h!]
-            \centering
-            \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
-            \caption{The effect of number of times a species appear (x-axis) in any of the various motifs (columns) on persistence (y-axis).
-The effect of each motif participation is fitted by linear mixed-effect models, following \emph{Equation 2; Main Text}, except that the disturbance levels are fitted separately.
-The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
-            \label{fig:abs_lmer_all}
-        \end{figure}
-
-        Persistence increased significantly with the number of times a species appeared in any of the four possible motifs in the food webs (Table~\ref{tab:absolute_number} and Figure~\ref{fig:abs_lmer_all}).
-        These relationships were slightly stronger for the number of times a species appears in the three-species chain and direct competition motifs and somewhat weaker for apparent competition and omnivory.
-The absolute difference between the motifs was small, with a clear positive relationship between persistence and motif count present in all cases.
-        As expected, persistence decreased significantly with increasing disturbance of basal resources and the strength of this relationship was very similar across the four motifs.
-        The level of disturbance to basal resources also affected the relationship between persistence and motif counts.
-        For all four motifs, there was a negative interaction between motif count and disturbance such that, for high levels of disturbance, participating more often in any of the motifs would result in decreased persistence.
-
-        \begin{table}[h!]
-            \caption{Coefficients ($\beta$) for models relating persistence to the number of times a species appears in the focal motif, the level of disturbance to basal resources, and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{absreq}).
-$p$-values for all coefficients in all models were \textless0.001.
-We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
-            \label{tab:absolute_number}        \centering
-            \begin{tabular}{c|c  c c  c | c c |}
-            Motif & Intercept & Motif count & Disturbance & Interaction & $R^2_M$ & $R^2_C$ \\
-            \hline
-            3-sp Chain & 0.762 & 1.14$\times10^{-4}$ & -0.388 & -1.36$\times10^{-4}$ & 0.897 & 0.913 \\
-            Apparent Compet. & 0.770 & 3.41$\times10^{-5}$ & -0.397 & -4.36$\times10^{-5}$ & 0.898 & 0.911\\
-            Direct Competition & 0.764 & 1.22$\times10^{-4}$ & -0.398 & -8.46$\times10^{-5}$ & 0.895 & 0.914 \\
-            Omnivory & 0.770 & 6.01$\times10^{-5}$ & -0.400 & -5.91$\times10^{-5}$ & 0.897 & 0.912 \\                
-            \end{tabular}
-            \end{table}
-
-\clearpage
-
-\section{Motif-persistence relationships and network size}
-
-    When comparing the outcomes from linear mixed-effect models (LMMs) relating persistence to motif roles across all networks (Fig.
-3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately, the main trends are similar but different-sized networks showed slightly different results results for some motifs (Figure temporarily in main text).%(Figure~\ref{fig:density_prop_S}).
-    For the three-species chain and omnivory motifs, a significant interaction between disturbance and the effect of the focal motif means that having a larger proportion of the role be made up of the chain or omnivory motif is associated with lower persistence at high levels of disturbance (Fig.
-3 and Table 1, \emph{Main Text}).
-    The same pattern holds overall for LRs fit within a network  %(Figure~\ref{fig:density_prop_S})
-    : the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance. 
-    This trend is most pronounced in large networks.% (bottom panel, Figure~\ref{fig:density_prop_S}). 
-    Besides the decrease in fraction of positive slopes with increasing disturbance, the omnivory motif show a higher overall fraction of positive slopes in comparison to the three-species chain motif, corresponding to the positive main effect of omnivory (Fig.
-3 and Table 1, \emph{Main Text}).
+%     \subsection*{Results}
 
@@
 %             \centering
-    %      \includegraphics[width=\textwidth]{manuscript/figures/prop_dens_bp_vs_S_allC.pdf}
-
-    %     \caption{Density (y-axis) of slopes (x-axis) for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
-The slopes are derived from linear mixed-effect models, with a random intercept for the interaction between species richness and network size.
-A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
-The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to motif (legend).
-Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
-Rows show various levels of network size.
-The vertical line indicate zero on the x-axis.}
-    %     \label{fig:density_prop_S}
+%             \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
+%             \caption{The effect of number of times a species appear (x-axis) in any of the various motifs (columns) on persistence (y-axis).
+The effect of each motif participation is fitted by linear mixed-effect models, following \emph{Equation 2; Main Text}, except that the disturbance levels are fitted separately.
+The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
+%             \label{fig:abs_lmer_all}
 %         \end{figure}
 
-    For the apparent and direct competition motifs, the LMMs include a positive interaction between disturbance and the proportion of the motif in a species' role (Fig.
-3 and Table 1; \emph{Main Text}).
-    This means that having more of the role made up of these two motifs is increasingly beneficial at higher levels of disturbance, a trend which is also evident from the increasing fraction of positive slopes with increasing disturbance based on LRs of each network at each level of disturbance.
-% (Figure~\ref{fig:density_prop_S}).
-    As with the trends for omnivory and three-species chains, these increases are strongest in large networks. % (bottom row, Figure~\ref{fig:density_prop_S}).
-    % For direct competition, the fraction of slopes are similar across disturbance levels.
-Although this interaction is not positive when accounting for network size, the fraction of positive slopes are throughout high.
-    The increase in the proportion of positive slopes was much stronger for apparent competition than direct competition, consistent with the smaller (but still significant) interaction term for between direct competition in the LMMs.
-    A high proportion of direct competition in a species' role tended to be associated with greater probability of persistence across levels of disturbance and network size.
-    % In conclusion, species richness does not considerably alter the trends visible in the LMMs.
-High species richness more clearly display an interaction between motif proportion, persistences and disturbance levels, while this effect is smaller with low species richness.
+%         Persistence increased significantly with the number of times a species appeared in any of the four possible motifs in the food webs (Table~\ref{tab:absolute_number} and Figure~\ref{fig:abs_lmer_all}).
+%         These relationships were slightly stronger for the number of times a species appears in the three-species chain and direct competition motifs and somewhat weaker for apparent competition and omnivory.
+The absolute difference between the motifs was small, with a clear positive relationship between persistence and motif count present in all cases.
+%         As expected, persistence decreased significantly with increasing disturbance of basal resources and the strength of this relationship was very similar across the four motifs.
+%         The level of disturbance to basal resources also affected the relationship between persistence and motif counts.
+%         For all four motifs, there was a negative interaction between motif count and disturbance such that, for high levels of disturbance, participating more often in any of the motifs would result in decreased persistence.
 
-\section{How does persistence vary with degree and trophic level?}
+%         \begin{table}[h!]
+%             \caption{Coefficients ($\beta$) for models relating persistence to the number of times a species appears in the focal motif, the level of disturbance to basal resources, and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{absreq}).
+$p$-values for all coefficients in all models were \textless0.001.
+We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
+%             \label{tab:absolute_number}        \centering
+%             \begin{tabular}{c|c  c c  c | c c |}
+%             Motif & Intercept & Motif count & Disturbance & Interaction & $R^2_M$ & $R^2_C$ \\
+%             \hline
+%             3-sp Chain & 0.762 & 1.14$\times10^{-4}$ & -0.388 & -1.36$\times10^{-4}$ & 0.897 & 0.913 \\
+%             Apparent Compet. & 0.770 & 3.41$\times10^{-5}$ & -0.397 & -4.36$\times10^{-5}$ & 0.898 & 0.911\\
+%             Direct Competition & 0.764 & 1.22$\times10^{-4}$ & -0.398 & -8.46$\times10^{-5}$ & 0.895 & 0.914 \\
+%             Omnivory & 0.770 & 6.01$\times10^{-5}$ & -0.400 & -5.91$\times10^{-5}$ & 0.897 & 0.912 \\                
+%             \end{tabular}
+%             \end{table}
 
-  \begin{figure}
-     \centering
-     \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}
+% \clearpage
 
-     \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
-The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
-Persistence decreases with darker color in the heat map.}
-     \label{fig:heatmap_stl_BP0}
- \end{figure}
-
- \begin{figure}
-     \centering
-     \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
-
-     \caption{Primary producers experience a disturbance reducing their probability to go extinct to $\pi_{disturbed} = 0.5$ All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
-The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
-Persistence decreases with darker color in the heat map.}
-     \label{fig:heatmap_stl_BP1}
- \end{figure}
-
-\clearpage 
+\bibliography{anna_bib} % Abbreviate journal titles.
 \bibliographystyle{ecollett}
-\bibliography{MyCollection} % Abbreviate journal titles.
 

### 2021-06-04 code/stat_analysis/statistical_models.R

@@ -35,7 +35,7 @@ consumers$prop_omni=consumers$m38/consumers$total_motifs
 
 # Section 1: How does persistence vary with global network properties?
 
-    SCper=with(consumers,lm(Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)))
+    SCper=with(consumers,lm(            Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)))
 
 
 # Section 2: How do motif profiles vary with S and C?
@@ -155,6 +155,13 @@ consumers$prop_omni=consumers$m38/consumers$total_motifs
     chain_prop=with(subdata,lm(prop_chain~Size*Connectance))
     apparent_prop=with(subdata,lm(prop_apparent~Size*Connectance))
     direct_prop=with(subdata,lm(prop_direct~Size*Connectance))
+    results=matrix(nrow=4,ncol=5)
+    results[1,]=c("Omnivory",summary(omni_prop)$coefficients[,1])
+    results[2,]=c("Chain",summary(chain_prop)$coefficients[,1])
+    results[3,]=c("Apparent",summary(apparent_prop)$coefficients[,1])
+    results[4,]=c("Direct",summary(direct_prop)$coefficients[,1])
+    write.table(results,file='roles_vs_SC.tsv',sep='\t')
+
 
     # Deg and TL
     # relating counts and proportions to degree, TL (repeats analysis in Cirwill & Wootton, in prep.)
@@ -208,6 +215,7 @@ consumers$prop_omni=consumers$m38/consumers$total_motifs
     }
 
 # Other, supplemental tests:\
+save.image('all_tests.Rdata')
 
 
 # Do proportions of motifs vary due to network processing?

### 2021-06-04 manuscript/motifs_vs_degree.tex
Updates from Overleaf

@@
      %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
-    The study of how an initial decline or extinction leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, eklof2006species}.
+    The study of how an initial decline or extinction leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
 Traditionally, there are two main approaches for studying secondary extinctions.
@@
     %\subsection*{3 Global scale network properties are known to affect stability}
-    The global structural properties of ecological networks have implications for their functioning \citep{petchey2002functional}, stability \citep{allesina2012stability} and robustness to disturbances \citep{dunne2002network, eklof2006species}.
-For example, more highly connected networks are more resistant to secondary extinctions \citep{dunne2002network, eklof2006species} and a modular organization where some groups of species are tightly interconnected with limited connections between groups can limit the propagation of disturbances \citep{}.
+    The global structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002d, Eklof2006}.
+For example, more highly connected networks are more resistant to secondary extinctions \citep{Dunne2002d, Eklof2006} and a modular organization where some groups of species are tightly interconnected with limited connections between groups can limit the propagation of disturbances \citep{}.
     %\subsection*{4 meso-scale nw propoerties also affect stability}
@@
 
-		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}.
+		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.
 		To capture a range of plausible network architectures, we simulated networks with sizes ranging from 50 to 100 species (in steps of 10) and connectances ranging from 0.02 to 0.18 (in steps of 0.04).

### 2021-06-04 manuscript/motifs_vs_degree.tex
Updates from Overleaf

@@
      The plant community is the foundation upon which the myriad of depend for their survival \citep{}.
-Disturbances in the plant community have been shown to affect important ecosystem properties such as productivity \citep{}, nutrient cycling \citep{} and consumer diversity \citep{scherber2010bottom}.The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community promotes stability across all different levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}.
+Disturbances in the plant community have been shown to affect important ecosystem properties such as nutrient cycling \citep{} and consumer diversity \citep{scherber2010bottom,Baiser2013}.The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community promotes stability across all different levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}.
 Loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers, e.g.
@@
 
-        Network average persistence of consumers naturally decrease with increasing level of disturbance of basal resources (Table~\ref{tab:per_vs_SC}).
+        Network average persistence of consumers naturally decrease with increasing level of disturbance of basal resources (Table S1, \emph{Appendix SA}).
 Network size and connectance also affect persistence, but to a much smaller extent (Figure~\ref{fig:lm_CS}).
-Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance: at low disturbance small and highly-connected webs show the lowest persistence, whereas at high disturbance level large and highly connected webs show the lowest persistence.
+Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance: at low disturbance, small and highly-connected webs show the lowest persistence, whereas at high disturbance, large and highly connected webs show the lowest persistence (Fig.
+S1, \emph{Appendix SA}).
 
@@
 
-        %\begin{figure}[h!]
-        %   \includegraphics[width=0.95\textwidth]{figures/heatmap_CS_BPcompare.eps}
-        %    \caption{Heat maps of average consumer persistence in networks with varying connectance (y-axis) and network size (x-axis) at two levels of disturbance to basal resources.
-Persistence decreases with darker color in the heat map.
-When basal species are not disturbed, all species experience a baseline extinction risk of $\pi = 0.1$.
-When basal species are highly disturbed, their baseline extinction risk rises to $\pi = 0.5$ while the baseline risk of consumers is not changed.}
-        %    \label{fig:heatmap_CS}
-        %\end{figure}
-
-        \begin{table}[hb!]
-            \caption{Persistence of consumers within a network depends on network size (S) and connectance (C), the level of disturbance to basal resources (D), and interactions between them.
-Here we show the effect sizes ($\beta$) and $p$-values for the regression described above.
-All predictors were centered and scaled before fitting the model.
-The level of disturbance corresponds to basal resources having a probability of extinction ranging between $0.1 - 0.5$ (see \emph{Methods}).
-            }
-            \label{tab:per_vs_SC}
-            \centering
-            \begin{tabular}{c|c c |}
-                Predictor & $\beta$ & $p$-value \\
-                \hline
-                Intercept & 0.572 & \textless0.001 \\
-                Size & -5.15$\times10^{-4}$  & \textless0.001 \\
-                Connectance & -0.0129 & \textless0.001 \\
-                Disturbance & -0.139 & \textless0.001 \\
-                S:C & -5.24$\times10^{-4}$ & \textless0.001 \\
-                S:D & -3.48$\times10^{-3}$ & \textless0.001 \\
-                C:D & -7.00$\times10^{-5}$ & 0.095 \\
-                S:C:D: & -2.69$\times10^{-4}$ & \textless0.001 \\
-            \end{tabular}
-        \end{table}
-
     \clearpage
@@
 
-        A network's normalized motif profile (i.e., proportions of the total made up by each motif; see \emph{Appendix SX} for non-normalized motif profiles) is related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
-        The proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table~\ref{network_prop_lms}).
-        Network size and the interaction between network size and connectance did not strongly affect changes in normalized motif profiles with connectance, except for direct competition (Fig.~\ref{motif_proportion_lms}).
-The proportion of direct competition decreased with increasing connectance in small networks but showed very little change in large webs.
+        A network's normalized motif profile (i.e., proportions of the total made up by each motif)
+        % ; see \emph{Appendix SX} for non-normalized motif profiles) 
+        is related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
+        Unequal variation in motif profiles across levels of size, connectance, and their interaction may, however, mean that these PERMANOVA results are false positives (Fig. S2, \emph{Appendix SB}).
+
+        The proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}).
+        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).
+The proportion of direct competition, however, decreased with increasing connectance in small networks but showed very little change in large webs.
 
@@
 
-        \begin{table}[hb!]
-            \centering
-            \caption{Coefficients for linear regressions relating the proportion of each motif in the total motif profile of a network to network size, connectance, and their interaction.
-Coefficients in \textbf{bold} are significant (\textless0.05).}
-           \label{network_prop_lms}
-           \begin{tabular}{c|c c c c c}
-                Motif & Intercept & Size & Connectance & Interaction \\
-                \hline
-                Three-species Chain & \textbf{0.262} & \textbf{1.78$\times10^{-4}$} & \textbf{-0.157} & \textbf{-3.27$\times10^{-3}$} \\
-                Apparent Competition & \textbf{0.541} & \textbf{-5.06$\times10^{-4}$} & \textbf{-1.00} & \textbf{2.59$\times10^{-3}$} \\
-                Direct Competition & \textbf{0.193} & -3.56$\times10^{-5}$ & \textbf{-0.556} & \textbf{5.92$\times10^{-3}$} \\
-                Omnivory & 4.82$\times10^{-3}$ & \textbf{3.64$\times10^{-4}$} & \textbf{1.71} & \textbf{-5.24$\times10^{-3}$}\\
-                \hline
-                \end{tabular}
-        \end{table}
-
-        %However, normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).These differences in variability can cause false positives in the PERMANOVA test.
-Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).
-
-       % \begin{figure}
-        %    \centering
-        %    \includegraphics[width=.75\textwidth]{manuscript/figures/proportion_profile_dispersion.pdf}
-        %    \caption{Dispersion of normalized motif profiles (i.e., proportions of each motif) within a network about the centroid for that combination of network size and connectance.
-Each circle represents a single network; circle colour indicates network size.
-Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
-        %    \label{dispersion_normmotifs}
-        %\end{figure}
-
     \clearpage
@@
 
-        The motif profile of a network does relate to the mean persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
-Mean persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table~\ref{motif_profile_tab}).Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif.
-These trends hold for all levels of disturbance (Table~\ref{motif_profile_tab}).
+        The motif profile of a network was related to the mean persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
+Mean persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif.
+These trends hold for all levels of disturbance (Table S3, \emph{Appendix SC}).
 Dispersion of motif profiles varies between networks with different levels of persistence ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
@@
 
-        \begin{table}[h!]
-            \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
-Predictors were centered and scaled before model fitting.
-All effects were significant ($p$\textless0.001).
-We also provide the $R^2$.
-}
-            \label{motif_profile_tab}
-            \centering
-            \begin{tabular}{c|c c c c c | c}
-                Motif & Intercept & Proportion & Disturbance & Interaction & $R^2$\\
-                \hline
-                Omnivory & 0.572 & -0.0133 & -0.139 & -4.89$\times10^{-4}$ & 0.908 \\
-                Apparent competition & 0.572 & 0.0105 & -0.139 & 4.78$\times10^{-4}$ & 0.905 \\
-                Direct competition & 0.572 & 2.62$\times10^{-3}$ & -0.139 & 3.84$\times10^{-4}$ & 0.901 \\
-                Three-species chain & 0.572 & 6.12$\times10^{-3}$ & -0.139 & -1.90$\times10^{-4}$ & 0.901 \\ 
-            \end{tabular}
-        \end{table}
-
     \clearpage
@@
 
-        See \emph{Appendix SY} for relationships between persistence and non-normalized motif participation.
+        % See \emph{Appendix SY} for relationships between persistence and non-normalized motif participation.
 
@@
             At high levels of disturbance, a species' persistence decreases with increasing proportion of the three-species chain, but increases with increasing proportion of all other motifs (Figure~\ref{fig:prop_lmer_all}).
-            This means that increased disturbance strengthen the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table~\ref{tab:proportion}).
+            This means that increased disturbance strengthen the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table S4, \emph{Appendix SD}).
             The effect of omnivory on persistence is still positive but weaker at higher levels of disturbance, and the effect of direct competition are largely unaffected by disturbance level.
-
-            \begin{table}[hb!]
-                \centering
-                \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{propreq}).
-Predictors were centered and scaled before model fitting.
-We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
-                \label{tab:proportion}                \footnotesize
-                \begin{tabular}{c|c c | c c | c c | c c | c c |}
-                & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c|}{$R^2$} \\
-                Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
-                \hline
-                3-sp Chain & 0.574 & \textless0.001 & -0.00508 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
-                App. Com. & 0.574 & \textless0.001 & -0.00312 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
-                Dir. Com. & 0.574 & \textless0.001 & 0.00564 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
-                Omnivory & 0.574 & \textless0.001 & 0.00709 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
-                \end{tabular}
-            \end{table}
 
@@
 
-            Networks with high connectance generally have more similar relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
+            Networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
 However, the interaction between a species' participation in the various motifs and disturbance changes with connectance  (Figure~\ref{fig:density_prop_C}).
@@
 
-        Persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SZZ}).
-Increased trophic level is associated with decreasing persistence, although the rate of this decline is slower at higher levels of disturbance (Table~\ref{tab:per_vs_TLdeg}).
+        Persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
+Increased trophic level is associated with decreasing persistence, although the rate of this decline is slower at higher levels of disturbance (Table S5, \emph{Appendix SF}).
         The interaction between in-degree and disturbance is even stronger, such as having a large number of prey is associated with higher persistence in the no disturbance scenario, but lower persistence at high disturbance.
-
-        \begin{table}[hb!]
-            \caption{Coeffcients ($\beta$) for regressions of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction.
-Predictors were centered and scaled before model fitting.
-All $p$-values were \textless0.001.
-The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.}
-            \label{tab:per_vs_TLdeg}
-            \centering
-            \begin{tabular}{l|c  c  c |}
-                Simple role measure & Role & Disturbance & Interaction \\
-                \hline
-                STL & -0.0218 & -0.139 & 0.00508 \\
-                in-degree & -0.00130 & -0.139 & -0.00908 \\
-            \end{tabular}
-        \end{table}
 

### 2021-06-04 manuscript/supplement_motifs.tex
moved one more table to SI

@@
     \end{figure}
+
 \clearpage
@@
 
-\section*{SH: Motif roles and network properties}
+\section{SH: Persistence and motif roles}
+
+    \begin{table}[h!]
+        \caption{Coefficients in a series of regressions relating probability of persistence to the frequency of a position in a species' motif role, disturbance, and their interaction.
+For ease of comparison between motifs, we did not center or scale predictors.
+Note, however, that some motifs and positions have higher frequencies than others.}
+        \label{tab:persistence_vs_positions}
+        \centering
+        \begin{tabular}{l | c c c c}
+        Position    & Intercept & Disturbance   & Position  & Interaction \\
+        \hline
+        Chain: top  & 0.880 &   -0.985 & -0.00544 & -0.332 \\
+        Chain: bottom   & 0.883 &   -1.07 & -0.0510 &   0.647 \\
+        Chain: middle   & 0.862 & -1.00 &   0.246 & -0.244 \\
+        \hline
+        Dir. Comp.: top & 0.872 &   -0.995  & 0.0549 &  -0.178 \\
+        Dir. Comp.: bottom & 0.886 &    -1.07 & -0.113 &    0.939 \\
+        \hline
+        Omnivory: top   & 0.872 &   -0.984 &    0.180 & -0.818 \\
+        Omnivory: middle & 0.873 & -1.00 &  0.119 & -0.324 \\
+        Omnivory: bottom & 0.882 & -1.06 &  -0.0449 &   0.845 \\
+        \hline
+        App. Comp.: top & 0.865 &   -0.966 &    0.140 & -0.520 \\
+        App. Comp.: bottom & 0.910 &    -1.12 & -0.100 &    0.323 \\
+        \hline
+        \end{tabular}
+    \end{table}
+\clearpage
+
+\section{SI: Motif roles and network properties}
 

### 2021-06-04 manuscript/motifs_vs_degree.tex

@@
 
-	\subsection*{Calculating persistences}
+	\subsection*{Calculating persistence}
 
-		First, we assign all species a baseline probability of extinction ($\pi_{base}$) -- the risk of going extinct given factors not related to the food web itself. 
+		First, we assigned all species a baseline probability of extinction ($\pi_{base}$) -- the risk of going extinct given factors not related to the food web itself. 
 		For consumer species, this also applies to extinction despite all resources being present. 
-		Such events can for example reflect diseases or stochastic extinctions of small populations.
+		Such events can, for example, reflect diseases or stochastic extinctions of small populations.
 
-		Additionally, we analyze a set of scenarios where primary producers get a higher baseline probability of extinction ($\pi_{disturbed}$). 
-		This reflects a threat to primary producers such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
+		As well as this baseline scenario, we simulate a set of scenarios where primary producers get a higher baseline probability of extinction ($\pi_{disturbed}$). 
+		These scenarios reflect threats to primary producers such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
 		$\pi_{disturbed}$ ranged between $0.1-0.5$, in steps of $0.08$. The highest disturbance level, $\pi_{disturbed} = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
@@
 
-        We then describe meso-scale structure at the species level by defining each species' motif participation. [[Should stop calling this motif roles]]
-		Finally, we defined species' motif participation following~\citet{Stouffer2012}. 
-		For each network, we identified each group of three interacting species and assigned it to the corresponding motif. 
-		A species' motif participation is a vector of the number of times it appears in each motif.
-		We calculated species' motif participations using the Python package \emph{pymfinder}~\citep{pymfinder}.
-        As with motif profiles of networks, we normalized participation vectors by dividing each count by the total number of motifs the species appears in.
+        We then describe meso-scale structure at the species level by defining each species' motif participation and species' roles.
+        Motif participation describes the frequency with which a focal species appears in each of the four motifs, analogous to the network motif profile~\citep{Stouffer2012}.
+        A species' role describes the frequency with which a focal species appears in each \emph{unique position} (e.g., top, middle, or bottom species in the three-species chain, top or bottom species in the apparent competition motif) within each motif ~\citep{Stouffer2012,Cirtwill2017}.
+		We calculated species' motif participation and roles using the Python package \emph{pymfinder}~\citep{pymfinder}.
+        As with motif profiles of networks, we normalized participation (and role) vectors by dividing each count by the total number of motifs (motif positions) the species appears in.
         Note, however, that this normalization does \emph{not} control for differences due to degree \emph{per se} as high-degree species also tend to participate in different types of motifs than low-degree species~\citep{Cirtwill2021_inprep}.
 
-		We began by calculating species' in-degrees (numbers of prey).
+        Finally, we calculated simpler measures of a species' role within its community: degree and trophic level.
 		As the Bayesian network approach only accounts  bottom-up effects, out-degree (number of predators) is irrelevant to extinction risk.
+		We therefore focused on species' in-degrees (number of prey).
 		Note, however, that out-degree may also be important in topological simulations which include both top-down and bottom-up effects.
 
-        Next, we calculated species' trophic levels. 
-		Among the many definitions of trophic level available, we chose shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}.
+		There are many definitions of trophic level available, most of which tend to be very similar~\citep{Carscallen2012}.
+		Here we use shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}. 
 		Basal resources have an STL of one, any species consuming a basal resource has an STL of two, etc.
@@
 
-            We are interested in whether having a greater proportion of a species' role be made up of a certain motif (e.g., direct competition) might increase persistence (see \emph{Appendix SXX} for relationships between persistence and counts of motifs).
+            We are interested in whether having a greater proportion of a species' role be made up of a certain motif (e.g., direct competition) might increase persistence.
+            % (see \emph{Appendix SXX} for relationships between persistence and counts of motifs).
             To test this, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
@@
 
-            Without disturbance, a species' persistence increases with increasing proportion of the omnivory and direct competition motifs.
-Conversely, a species' persistence decreases with increasing proportion of the apparent competition motif.
-Finally, a species persistance is  negatively but not significantly related to the proportion of the three-species chain (Figure~\ref{fig:prop_lmer_all}).
+            At low levels of disturbance, species' persistences increased with increasing proportions of the omnivory and direct competition motifs.
+Conversely, persistence decreased with as the  proportion of the apparent competition motif increased.
+Persistence was not significantly related to the proportion of the three-species chain (Figure~\ref{fig:prop_lmer_all}).
 
-            At high levels of disturbance, a species' persistence decreases with increasing proportion of the three-species chain, but increases with increasing proportion of all other motifs (Figure~\ref{fig:prop_lmer_all}).
-            This means that increased disturbance strengthen the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table S4, \emph{Appendix SD}).
-            The effect of omnivory on persistence is still positive but weaker at higher levels of disturbance, and the effect of direct competition are largely unaffected by disturbance level.
+            At high levels of disturbance, persistence decreased as the proportion of the three-species chain increased, but increased as the proportion of all other motifs increased (Figure~\ref{fig:prop_lmer_all}).
+            % This means that increased disturbance strengthened the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table S4, \emph{Appendix SD}).
+            The effect of omnivory on persistence was  positive but weaker at higher levels of disturbance, and the effect of direct competition was largely unaffected by disturbance level.
 
@@
 
-            By fitting linear regressions of persistence against motif proportions for the species for each network separately, we can extract the slopes and analyze their distribution.
-This reveals information on how the relationship change with the level of disturbance and global network structure.
-For the global network parameters the network size had a much smaller effect than connectance (Fig.~\ref{fig:density_prop_S}; \emph{Appendix SZ})and therefore we here focus on only connectance.
+            By fitting linear regressions of persistence against motif proportions for the species for each network separately, we can observe how the relationships between motifs and persistence change with the level of disturbance and global network structure.
+Network size had a much smaller effect than connectance (Fig.
+S3; \emph{Appendix SE}); we therefore focus on connectance here.
+            Networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
+            However, there was an interaction disturbance and connectance for the direct competition and omnivory motifs.
 
-            Networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
-However, the interaction between a species' participation in the various motifs and disturbance changes with connectance  (Figure~\ref{fig:density_prop_C}).
-            This is in particular evident for the direct competition and omnivory motif.
             At low and medium connectances there is little change in the proportion of positive slopes (\textgreater80\%) for the different levels of disturbance. 
-However, at high connectances the proportion of positive slopes is overall lower and also decreases with increasing disturbance.
-This means that at low levels of connectance, in a majority of cases an increased proportion of the direct competition motif increase persistence, while the opposite is true with high connectance.
+            At high connectances, however, the proportion of positive slopes is lower overall and decreases with increasing disturbance. 
+            This means that at low levels of connectance, an increased proportion of the direct competition motif increases persistence in most networks, while the opposite is true with high connectance.
             The trend is reversed for the omnivory motif.
-            Overall, the proportion of networks showing a positive effect of omnivory decreases with increasing disturbance.
-            This decrease is strongest at low and moderate connectances but very weak at high connectance.
-(Figure~\ref{fig:density_prop_C}).
-Notably, the distribution of slopes for the omnivory motif is wider (effect is more variable) the lower the connectance is.
-            Hence, the level of connectance influence whether a high proportion of the omnivory motif is associated with higher persistence or not.
 
@@
 
-            \begin{table}[hb!]
-            \centering
-            \caption{The meso-scale composition of a network varies with size and connectance.
-Here we give the range and mean ($\pm$ standard deviation) proportion of each motif in species' motif participation vectors.
-            We also show the effects ($\beta$) of network size, connectance, and their interaction in linear models predicting the proportion of each motif in a participation vector.
-            All effects were significant ($p$\textless0.001).
-            Predictors were not centred or standardized before model fitting. Motifs are ordered according to mean proportion.}
-            % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
-            \label{tab:partic_vs_SC}   
-            \footnotesize
-            \begin{tabular}{c|c c c | c c c}
-                Motif & Min & Max & Mean (SD) & Size & Connectance & Interaction \\
-                \hline
-                App. Comp & 0.250 & 0.765 & 0.422 (0.0645) & -5.06$\times10^{-4}$ & -1.00 & 2.59$\times10^{-3}$ \\
-                Chain & 0.109 & 0.413 & 0.235 (0.042) & 1.78$\times10^{-4}$ & -0.157 & -3.27$\times10^{-3}$ \\
-                Dir. Comp & 0.0501 & 0.359 & 0.179 (0.0407) & -3.56$\times10^{-5}$ & -0.556 & 5.92$\times10^{-3}$ \\
-                Omnivory & 0.006 & 0.346 & 0.164 (0.0790) & 3.64$\times10^{-4}$ & 1.71 & -5.24$\times10^{-3}$ \\   
-                \hline
-                \end{tabular}
-                \end{table}
-
-        The numbers of different motifs in a network may vary with web size, connectance, and their interaction.
-        To establish this context, we fit linear models relating the proportion of each motif in a species' role against network size, connectance, and their interaction.
-
-        \subsubsection*{Motif participation and simple role measures}
-
+        Motif participation varied with web size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
+        Species in all networks appeared most frequently in the apparent competition motif, but the frequency of the apparent competition motif decreased with increasing network size and connectance.
+        The frequency of the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size. 
         Species' motif roles also vary with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
@@
             \end{figure}
-
-\clearpage
-
-    [[Big deal result: connectance is important for motif structure of network, size not so much]]
-
-    [[Persistence depends on S, C, interaction with $basal_p$ (lowest persistence swaps from highest S to lowest C in Figs.
-1-2, high C in both cases - but small absolute differences so maybe we don't want to make too big a deal out of this)]]
-
-    [[proportion of omnivory increases with increasing C, while the other doesn't. Connect to later findings involving omnivory]]
-
-\clearpage
 
@@
 The top set of lines corresponds to a probability of extinction for basal resources of 0.1.
-The middle set of lines correspons to a probability of extinction for basal resources of 0.26, and the lowest set of lines corresponds to a probability of extinction for basal resources of 0.5.}
+The middle set of lines corresponds to a probability of extinction for basal resources of 0.26, and the lowest set of lines corresponds to a probability of extinction for basal resources of 0.5.}
             \label{fig:persistence_positions_prey}
@@
 
-        [[have fit these models but no figure yet - seems like degree and the 2-prey positions show the same pattern so they're probably tightly linked.
-STL may explain the differences in top and middle positions for the chain.
-There's still a fair amount of differences among the 0 prey and 1 prey motifs though, so motifs still provide unique information that degree and STL don't.]]
+        \subsubsection*{Degree and trophic level}
+
+            The top position in the omnivory motif increased with increasing degree across all trophic levels.
+            The frequency of the middle position increased with the interaction of degree and trophic level.
+            The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).
+
+            The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
+            The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.
+
+            The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
+            The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).
+
+            The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
+            The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
+            The top position in the three-species chain increased with trophic level, degree, and their interaction.
+
+        \subsubsection*{Network size and connectance}
+
+            % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
+            The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance.
+            The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.
+
+            The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
+            The frequency of the bottom position in the direct competition motif was low in all networks.
+
+            The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
+            Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
+            None of the positions in the three-species chain varied strongly with network size or connectance.
+
+\clearpage
+
+    [[Big deal result: connectance is important for motif structure of network, size not so much]]
+
+    [[Persistence depends on S, C, interaction with $basal_p$ (lowest persistence swaps from highest S to lowest C in Figs.
+1-2, high C in both cases - but small absolute differences so maybe we don't want to make too big a deal out of this)]]
+
+    [[proportion of omnivory increases with increasing C, while the other doesn't. Connect to later findings involving omnivory]]
+
+\clearpage
+
+    \subsection*{How does persistence vary with motif positions?}
+
+        A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance (Table S7, \emph{Appendix SH}).
+
+        % Organizing results by motif - keep this OR by no. prey
+        In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
+        Peristence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
+        Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.
+
+        In the three-species chain, persistence decreased as the frequency of the top position increased at all levels of disturbance.
+        Conversely, persistence increased with increasing frequency of the middle position at all levels of disturbance.
+        Persistence decreased slightly with increasing frequency of the bottom position at low levels of disturbance but increased at high levels of disturbance.
+
+        In the apparent competition motif, persistence decreased slightly with increasing frequencies of the bottom position (one predator, no prey within the motif) at low disturbance but increased slightly at high disturbance.
+        Persistence increased with increasing frequency of the top position (two prey, no predator within the motif) at low disturbance but decreased at high disturbance.
+
+        In the direct competition motif, persistence varied little with the frequency of the bottom position (two predators, no prey within the motif) at low disturbance but increased with the frequency of this position at high levels of disturbance.
+        Persistence increased slightly with the frequency of the top position (one prey, no predators within the motif) at low levels of disturbance and decreased slightly at high levels of disturbance.
+
+        \begin{figure}[h!]
+        \includegraphics[width=.95\textwidth]{figures/persistence_positions_bymotif.eps}
+        \label{figs:persistence_position_motifs}
+        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
+Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common.
+Line colour indicates the probability of extinction for basal resources.
+Line style indicates the number of prey within the motif for each position.
+Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.}
+        \end{figure}
+
+        % Organizing by number of prey
+        Across the four motifs we consider, positions with no prey within the motif tended to show weak relationships to persistence at low levels of disturbance and be associated with increased persistence at high levels of disturbance (Fig.~\ref{fig:persistence_positions_prey}).
+        These trends were strongest for direct competition and omnivory and weakest for apparent competition.
+        The two positions with stronger effects indicate a species with two predators within the motif, while the weaker effects are for motifs with one predator within the motif.
+
+        Positions with one prey within the motif were more variable.
+        The top position in the direct competition motif and middle position in the omnivory motif (both of which have one prey) were associated with increased persistence at low levels of disturbance and lower persistence at higher levels of disturbance, but these effects were relatively weak.
+        The top and middle positions of the three-species chain both include a single prey; the middle position was associated with increased persistence at all levels of disturbance while the top position was associated with decreased persistence at all levels of disturbance.
+
+        Positions with two prey in the apparent competition and omnivory motifs showed very similar patterns.
+        At low levels of disturbance, persistence increased with the frequency of either position.
+        At high levels of disturbance, persistence decreased as the frequency of either motif increased.
+        This echoes the pattern for in-degree.
+
+        \begin{figure}
+            \centering
+            \includegraphics[width=.95\textwidth]{figures/persistence_positions_byprey.eps}
+            \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
+Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common.
+Line colour indicates the motif containing a position.
+Line style indicates the number of prey within the motif for each position.
+Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.
+The top set of lines corresponds to a probability of extinction for basal resources of 0.1.
+The middle set of lines corresponds to a probability of extinction for basal resources of 0.26, and the lowest set of lines corresponds to a probability of extinction for basal resources of 0.5.}
+            \label{fig:persistence_positions_prey}
+        \end{figure}
+
+    \subsection*{How do motif positions vary with other network properties?}
+
+        The frequencies of motif positions in a species' role varied strongly and significantly with degree and trophic level, but only weakly with network size and connectance (\emph{Appendix SI}).
+
+           \subsubsection*{Degree and trophic level}
+
+            The top position in the omnivory motif increased with increasing degree across all trophic levels.
+            The frequency of the middle position increased with the interaction of degree and trophic level.
+            The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).
+
+            The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
+            The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.
+
+            The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
+            The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).
+
+            The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
+            The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
+            The top position in the three-species chain increased with trophic level, degree, and their interaction.
+
+        \subsubsection*{Network size and connectance}
+
+            % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
+            The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance.
+            The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.
+
+            The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
+            The frequency of the bottom position in the direct competition motif was low in all networks.
+
+            The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
+            Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
+            None of the positions in the three-species chain varied strongly with network size or connectance.
 

### 2021-06-04 manuscript/motifs_vs_degree.tex
All old changes have been recovered as far as I can tell.

@@
     For example, specialists may not necessarily have a high risk of extinction if it consumes a basal resource which is at low risk of being affected of disturbance.
-On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes few preys or prey which possess high risk of being affected by disturbance.
+On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes few prey species or prey which possess high risk of being affected by disturbance.
 

### 2021-06-16 manuscript/supplement_motifs.tex
moved a few figures over

@@
 
-    Connectance and species richness had smaller effects on consumer persistence than disturbance (Table~\ref{tab:per_vs_SC}, Fig.~\ref{fig:heatmap_CS}).
+    Connectance and species richness had smaller effects on consumer persistence than disturbance (Figs.~\ref{fig:lm_CS}, \ref{fig:heatmap_CS}, Table~\ref{tab:per_vs_SC}).
+
+    \begin{figure}[h!]
+        \centering
+        \includegraphics[width=0.95\textwidth]{figures/persistence_vs_SC_lm.eps}
+        \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while network size, connectance, and their interaction had smaller effects.
+\textbf{A)} Persistence decreased slightly with increasing network size at high levels of disturbance, increased slightly with increasing network size at low levels of disturbance, and did not vary noticeably with network size at moderate levels of disturbance.
+Networks with lower connectance (dotted line) had slightly higher persistence than networks with moderate (solid line) or high (dashed line) connectance.
+        \textbf{B)} Persistence decreased with increasing connectance for all levels of disturbance and network size. Network size (line types) did not strongly affect these relationships.}
+        \label{fig:lm_CS}
+    \end{figure}
 
@@
 
-    Normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).These differences in variability can cause false positives in the PERMANOVA test.
+    A network's normalized motif profile (i.e., proportions of the total made up by each motif) was related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
+    However, normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).These differences in variability can cause false positives in the PERMANOVA test.
 Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).
@@
     \end{table}
+
 \clearpage
@@
     \subsection*{Degree and trophic level}
+
+        The top position in the omnivory motif increased with increasing degree across all trophic levels.
+        The frequency of the middle position increased with the interaction of degree and trophic level.
+        The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).
+
+        The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
+        The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.
+
+        The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
+        The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).
+
+        The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
+        The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
+        The top position in the three-species chain increased with trophic level, degree, and their interaction.
 
@@
     \subsection*{Network size and connectance}
+
+        % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
+        The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance.
+        The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.
+
+        The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
+        The frequency of the bottom position in the direct competition motif was low in all networks.
+
+        The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
+        Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
+        None of the positions in the three-species chain varied strongly with network size or connectance.
 

### 2021-06-28 code/motif_SC_lms.py
redoing figs

@@ -24,21 +24,6 @@ colors.add_color(34,90,34,'Green')
 # colors.add_color(255,125,125,'lightish_red')
 # colors.add_color(200,200,200,'lightgrey')
 
-def read_dispfile(dispfile):
-  dispdict=[]
-
-  f=open(dispfile,'r')
-  for line in f:
-    if line.split()[0]!='"Distance"':
-      dist=float(line.split()[1][1:-1])
-      S=int(line.split()[3][1:-1])
-      C=float(line.split()[4][1:-1])
-      dispdict.append((S,C,dist))
-
-  f.close()
-
-  return dispdict
-
 def format_graph(graph,graphtype):
   graph.yaxis.bar.linewidth=1
   graph.xaxis.bar.linewidth=1
@@ -145,8 +130,6 @@ def populate_graph(graph,nettype):
 ###############################################################################################
 ###############################################################################################
 
-dispfile='stat_analysis/motif_variability_SC.tsv'
-dispdict=read_dispfile(dispfile) # S, C, dict
 
 grace=MultiPanelGrace(colors=colors)
 # grace.add_label_scheme('dummy',['Start of flowering','Mass flowering','Inferred peak','End of flowering','Season length','F','G','H'])

### 2021-06-28 code/plot_dispersion.py

@@ -31,12 +31,11 @@ def read_dispfile(dispfile):
   for line in f:
     if line.split()[0]!='"Distance"':
       dist=float(line.split()[1][1:-1])
-      S=int(line.split()[3][1:-1])
-      C=float(line.split()[4][1:-1])
+      S=int(line.split()[3].split('.')[0][1:-1])
+      C=float('0.'+line.split()[3].split('.')[1][:-1])
       dispdict.append((S,C,dist))
 
   f.close()
-
   return dispdict
 
 def format_graph(graph,graphtype):
@@ -100,26 +99,26 @@ def populate_graph(graph,dispdict,graphtype):
 ###############################################################################################
 ###############################################################################################
 
-dispfile='stat_analysis/motif_variability_SC.tsv'
-dispdict=read_dispfile(dispfile) # S, C, dict
+# dispfile='stat_analysis/motif_variability_SC.tsv'
+# dispdict=read_dispfile(dispfile) # S, C, dict
 
-grace=MultiPanelGrace(colors=colors)
-# grace.add_label_scheme('dummy',['Start of flowering','Mass flowering','Inferred peak','End of flowering','Season length','F','G','H'])
-# grace.set_label_scheme('dummy')
+# grace=MultiPanelGrace(colors=colors)
+# # grace.add_label_scheme('dummy',['Start of flowering','Mass flowering','Inferred peak','End of flowering','Season length','F','G','H'])
+# # grace.set_label_scheme('dummy')
 
-graph=grace.add_graph(Panel)
-graph=format_graph(graph,'heatmap')
-graph=populate_graph(graph,dispdict,'heatmap')
-graph.panel_label.configure(placement='iul',char_size=1,dx=.03,dy=.03)
+# graph=grace.add_graph(Panel)
+# graph=format_graph(graph,'heatmap')
+# graph=populate_graph(graph,dispdict,'heatmap')
+# graph.panel_label.configure(placement='iul',char_size=1,dx=.03,dy=.03)
 
-# grace.multi(rows=3,cols=2,vgap=.09,hgap=.04)
-# grace.hide_redundant_labels()
-# grace.set_row_xaxislabel(label='Degree (Number of interaction partners per species)',row=1,colspan=(None,None),char_size=1,perpendicular_offset=.05)
[…]

### 2021-06-28 code/plot_roles_vs_TL.py

@@ -35,37 +35,6 @@ colors.add_color(158,1,66,'Rred2')
 # colors.add_color(255,125,125,'lightish_red')
 # colors.add_color(200,200,200,'lightgrey')
 
-# Each point will be mean of one network
-def read_datafile(datafile):
-  speciesprops={}
-  f=open(datafile,'r')
-  for line in f:
-    if line.split()[0]!='Size':
-      S=int(line.split()[0])
-      C=float(line.split()[1])
-      basal_p=line.split()[2]
-      network=line.split()[3]
-      species=line.split()[4]
-      ID=str(S)+'_'+str(C)+'_'+network+'_'+species
-      pers=line.split()[5]
-      in_deg=float(line.split()[6])
-      out_deg=float(line.split()[7])
-      STL=int(line.split()[8])
-      PATL=float(line.split()[9])
-      chain=int(line.split()[10])
-      dircomp=int(line.split()[11])
-      omnivory=int(line.split()[12])
-      appcomp=int(line.split()[13])
-      total=chain+dircomp+omnivory+appcomp
-      speciesprops[ID]={'TL':STL,'Deg':in_deg,
-      'Raw':{'chain':chain,'dircomp':dircomp,'omnivory':omnivory,'appcomp':appcomp},
-      'Prop':{'chain':float(chain)/float(total),'dircomp':float(dircomp)/float(total),
-         'omnivory':float(omnivory)/float(total),'appcomp':float(appcomp)/float(total)}}
-
-  f.close()
-
-  return speciesprops
-
 def read_lmfile(lmfile):
   lmdict={'Deg':{'Count':{},'Prop':{}},'TL':{'Count':{},'Prop':{}}}
 
@@ -183,20 +152,27 @@ def populate_graph(graph,minidict,simple,roletype):
 
 def populate_persgraph(graph,persdict,simple):
   if simple=='Deg':
-    key='in_Degree'
+    key='scale(in_Degree)'
+    scal=12.51637
+    cent=11.95232
   else:
-    key='STL'
+    key='scale(STL)'
+    scal=0.4803683
+    cent=2.274325
+
+  print persdict.keys()
 
   j=12
-  for basal_p in [0,0.2,0.4,0.6,0.8,1]:
-    p=0.1+0.4*basal_p
+  for basal_p in [0,0.18,0.26,0.34,0.42,0.5]:
+    p=(basal_p-0.3)/0.1366261
 
     dats=[]
     # lower=[]
     # upper=[]
[…]

### 2021-06-28 code/stat_analysis/persistence_vs_positions.py

@@ -291,3 +291,17 @@ grace.hide_redundant_labels()
 
 grace.write_file('../../manuscript/figures/persistence_positions_byprey.eps')
 
+
+# Omnivory only, with a clearer legend. May also want to indicate mean proportion for each position.
+# Do the same for the other 3 motifs?
+# Needs to wait for re-run ...
+grace=Grace(colors=colors)
+
+mot='Omnivory'
+graph2=grace.add_graph()
+graph2=format_graph(graph2,'persistence')
+graph2=populate_persgraph(graph2,mot,predictors)
+graph2.panel_label.configure(placement='ouc',char_size=1,dx=.0,dy=.01)
+
+grace.write_file('../../manuscript/figures/persistence_positions_omnivory.eps')
+

### 2021-06-28 code/stat_analysis/statistical_models.R

@@ -35,7 +35,7 @@ consumers$prop_omni=consumers$m38/consumers$total_motifs
 
 # Section 1: How does persistence vary with global network properties?
 
-    SCper=with(consumers,lm(            Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)))
+    SCper=with(consumers,lm(Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)))
 
 
 # Section 2: How do motif profiles vary with S and C?

### 2021-06-28 manuscript/motifs_vs_degree.tex

--- 2021-06-04 manuscript/motifs_vs_degree.tex
@@
     %\subsection*{1 Why are bottom-up effects important for understanding stability?}
-     The plant community is the foundation upon which the myriad of depend for their survival \citep{}.
-Disturbances in the plant community have been shown to affect important ecosystem properties such as nutrient cycling \citep{} and consumer diversity \citep{scherber2010bottom,Baiser2013}.The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community promotes stability across all different levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}.
+     The plant community is the foundation upon which the myriad of species in food webs depend for their survival \citep{}.
+Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{} and secondary production \citep{}, soil respiration and carbon cycling \citep{} and consumer diversity \citep{scherber2010bottom,Baiser2013}.
+The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community promotes stability across all different levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}.
 Loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers, e.g.
 herbivores, predators, and parasitoids \citep{scherber2010bottom}.
-     This has been shown in both theoretical work~\citep{} and empirical experiments ~\citep{} and results from the vertical propagation of disturbances through the network of interacting species.
+     %This has been shown in both theoretical work~\citep{} and empirical experiments ~\citep{} and results from the vertical propagation of disturbances through the network of interacting species.
 
      %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
-    The study of how an initial decline or extinction leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    The study of how an initial population decline or extinction leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
 Traditionally, there are two main approaches for studying secondary extinctions.
@@
     The global structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002d, Eklof2006}.
-For example, more highly connected networks are more resistant to secondary extinctions \citep{Dunne2002d, Eklof2006} and a modular organization where some groups of species are tightly interconnected with limited connections between groups can limit the propagation of disturbances \citep{}.
+For example, more highly connected networks are more resistant to secondary extinctions \citep{Dunne2002d, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{}.
     %\subsection*{4 meso-scale nw propoerties also affect stability}
@@
     For example, specialists may not necessarily have a high risk of extinction if it consumes a basal resource which is at low risk of being affected of disturbance.
-On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes few preys or prey which possess high risk of being affected by disturbance.
+On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes preys which possess high risk of being affected by disturbance.
 
@@
     %\subsection{6. only 4 (stable) motifs can appear in Bayesian networks, so how do we expect them to affect stability in a bottom-up context?}
-    Of the 13 three-species motifs that can occur in empirical food webs
-    \citep{Milo2004}, only four do not include cycles (e.g., A eats B and B eats A or A eats B, B eats C, and C eats A). 
-    As cycles must be removed from networks before calculating extinction probabilities in the Bayesian network framework, this study focuses on these four Bayesian-compatible motifs: the three-species chain, omnivory, apparent competition and direct competition motifs.
-    Despite this being a small subset of the motifs which can occur in empirical networks, our choice is justified by studies showing that these four motifs are over-represented in empirical networks and contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
-    When analyzing motifs within a Bayesian Network, it is important to keep in mind that only bottom-up effects are accounted for.
+    Of the 13 three-species motifs that can occur in empirical food webs there are four motifs that are over-represented in empirical networks \citep{Stouffer2007} and contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}: the three-species chain, omnivory, apparent competition and direct competition motifs.
+    These four motifs are also the only ones which can occur in the acyclic networks required for Bayesian network simulations~\citep{Eklof2013}.
+Because of this restriction, we only consider these four particularly important motifs.
+    % only four do not include cycles (e.g., A eats B and B eats A or A eats B, B eats C, and C eats A). 
+    %As cycles must be removed from networks before calculating extinction probabilities in the Bayesian network framework, this study focuses on these four Bayesian-compatible motifs:
+
+    %When analyzing motifs within a Bayesian Network, it is important to keep in mind that only bottom-up effects are accounted for.
 Therefore, the implications of the four possible motifs are different than they would be in an empirical network.
 
-    In particular, the lack of top-down effects removes the indirect effects from the two competition motifs.
+    %In particular, the lack of top-down effects removes the indirect effects from the two competition motifs.
 In the case of direct competition, two predators share a common resource.
@@
 Similarly, the two bottom-level species in the apparent competition motif will be independent of each other as there are no top-down effects of the shared consumer to link them.
-    The difference between the direct competition and apparent competition motif is thus the number of bottom-level species, accounted for as prey items for a higher level species.
-    A predator participating in many direct competition motifs is likely to have fewer prey items than a predator participating frequently in the apparent competition motif, and therefore a higher extinction risk.
-    Note, however, that consumers can also appear in these motifs as prey to other consumers.  
+    %The difference between the direct competition and apparent competition motif is thus the number of bottom-level species, accounted for as prey items for a higher level species.
+   % A predator participating in many direct competition motifs is likely to have fewer prey items than a predator participating frequently in the apparent competition motif, and therefore a higher extinction risk.Note, however, that consumers can also appear in these motifs as prey to other consumers.
 In that case, we would not expect the two motifs to have different effects on extinction risk.
-
-    The structure of the omnivory motif and the three-species chain motif differ from the competition motifs by having an additional trophic level.
+    %The structure of the omnivory motif and the three-species chain motif differ from the competition motifs by having an additional trophic level.
 Therefore, top-level species of these motifs might be more sensitive to disturbances since they can be affected by the loss of the basal resource in the focal motif or an increased probability of extinction of the middle species due to a disturbance to a basal resource not included in the focal motif.
@@
        %\subsection*{Here we test...}
-    Here we are interested in how disturbances at the basal level in food webs propagate in the network and whether a consumer's extinction risk depends on 1) the motif profile of the full food web and 2) the motifs in which a focal species participates.
-In particular, we want to understand whether the information provided by motifs can be explained by the classical global and local food web metrics such as network connectance and species trophic level, or whether motifs provide unique insight into bottom-up extinction.
+    Here we are interested in how disturbances at the basal level in food webs propagate up the network and whether a consumer's extinction risk depends on 1) the motif profile of the full food web and 2) the motifs in which a focal species participates.
+In particular, we want to understand whether the information provided by motifs can be explained by the classical global and local food web properties such as network connectance and species trophic level, or whether motifs provide unique insight into bottom-up extinction cascades.
 
@@
 
-    \subsection*{How does mean persistence vary with global network structure?}
+    \subsection*{Persistence did not vary strongly with global network structure}
 
-        Network average persistence of consumers naturally decrease with increasing level of disturbance of basal resources (Table S1, \emph{Appendix SA}).
-Network size and connectance also affect persistence, but to a much smaller extent (Figure~\ref{fig:lm_CS}).
+        Network average persistence of consumers naturally decreases with increasing level of disturbance of basal resources (Table S1, \emph{Appendix SA}).
+Network size and connectance also affect persistence, but to a much smaller extent (Figure S1, \emph{Appendix SA}).
 Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance: at low disturbance, small and highly-connected webs show the lowest persistence, whereas at high disturbance, large and highly connected webs show the lowest persistence (Fig.
-S1, \emph{Appendix SA}).
+S2, \emph{Appendix SA}).
 
-        \begin{figure}[h!]
-            \centering
-            \includegraphics[width=0.95\textwidth]{figures/persistence_vs_SC_lm.eps}
-            \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while network size, connectance, and their interaction had smaller effects.
-\textbf{A)} Persistence decreased slightly with increasing network size at high levels of disturbance, increased slightly with increasing network size at low levels of disturbance, and did not vary noticeably with network size at moderate levels of disturbance.
-Networks with lower connectance (dotted line) had slightly higher persistence than networks with moderate (solid line) or high (dashed line) connectance.
-            \textbf{B)} Persistence decreased with increasing connectance for all levels of disturbance and network size. Network size (line types) did not strongly affect these relationships.}
-            \label{fig:lm_CS}
-        \end{figure}
+    \subsection*{Network motif profiles varied with global network properties}
 
-    \clearpage
-
-    \subsection*{How do network motif profile vary with global network properties?}
-
-        A network's normalized motif profile (i.e., proportions of the total made up by each motif)
-        % ; see \emph{Appendix SX} for non-normalized motif profiles) 
-        is related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
-        Unequal variation in motif profiles across levels of size, connectance, and their interaction may, however, mean that these PERMANOVA results are false positives (Fig. S2, \emph{Appendix SB}).
-
-        The proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}).
+        A PERMANOVA test suggests that a network's normalized motif profile as a whole (i.e., proportions of the total made up by each motif) is related to the network's size, connectance, and the interaction between them, but unequal variation in motif profiles across global network properties may, however, mean that these PERMANOVA results are false positives (Fig.
+S2, \emph{Appendix SB}).
+        Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}).
         Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).
@@
 
-    \subsection*{How does mean persistence vary with network motif profile?}
+    \subsection*{Persistence decreases with greater omnivory in a network}
 
         The motif profile of a network was related to the mean persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
-Mean persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif.
+Mean persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
+Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif.
 These trends hold for all levels of disturbance (Table S3, \emph{Appendix SC}).
 Dispersion of motif profiles varies between networks with different levels of persistence ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
-For example, networks with higher persistence have more variable motif profiles ($\beta$=0.183, $p$\textless0.001).
+For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).
 
@@
 
-    \clearpage
-
-    \subsection*{How does species persistence vary with motif participation?}
+    \subsection*{Species persistence varies with motif participation}
 
@@
 
-        \subsubsection*{Interactions between motif participation and disturbance}
+        \subsubsection*{Effects of motif participation depend on disturbance}
 
@@
 
-        \subsubsection*{Variability of motif effects changes with network properties}
+        \subsubsection*{Effects of motif participation are more consistent in highly-connected networks}
 
@@
 
-        \begin{figure}
+        \begin{figure}[h!]
             \centering
@@
 
-        \begin{figure}[h!]
-             \centering
-             \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
-            \caption{Density (y-axis) of slopes (x-axis) for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
-The slopes are derived from standard linear regression.
-A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
-The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to motif (legend).
-Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
-Rows show various levels of network size.
-The vertical line indicate zero on the x-axis.}
-            \label{fig:density_prop_S}
-        \end{figure}
-
-    \clearpage
-
-    \subsection*{How does persistence vary with degree and trophic level?}
+    \subsection*{Persistence increases with degree but decreases with STL}
 
@@
 
-    \subsection*{Do simpler measures explain effects of motifs?}
-
-        \subsubsection*{Motif participation and network size and connectance}
+    \subsection*{Simpler measures do not fully explain effects of motifs}
 
@@
         The frequency of the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size.
+
         Species' motif roles also vary with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
         Species with higher in-degree tend to have higher proportions of their motif role made up by omnivory. Proportions of the other three motifs tend to decline with increasing degree.
-
         Species with higher trophic levels tend to have higher proportions of apparent competition and three-species chains.
@@
 
-    \subsection*{How does persistence vary with motif positions?}
+    \subsection*{Persistence varies with a species' motif role}
 
-        A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance.
-        \begin{table}[h!]
-            \caption{Coefficients in a series of regressions relating probability of persistence to the frequency of a position in a species' motif role, disturbance, and their interaction.
-For ease of comparison between motifs, we did not center or scale predictors.
-Note, however, that some motifs and positions have higher frequencies than others.}
-            \label{tab:persistence_vs_positions}
-            \centering
-            \begin{tabular}{l | c c c c}
-            Position	& Intercept	& Disturbance	& Position	& Interaction \\
-            \hline
-            Chain: top	& 0.880 &	-0.985 & -0.00544 & -0.332 \\
-            Chain: bottom	& 0.883 &	-1.07 &	-0.0510 &	0.647 \\
-            Chain: middle	& 0.862 & -1.00 &	0.246 &	-0.244 \\
-            \hline
-            Dir. Comp.: top & 0.872 &	-0.995	& 0.0549 &	-0.178 \\
-            Dir. Comp.: bottom & 0.886 &	-1.07 &	-0.113 &	0.939 \\
-            \hline
-            Omnivory: top	& 0.872 &	-0.984 &	0.180 &	-0.818 \\
-            Omnivory: middle & 0.873 & -1.00 &	0.119 &	-0.324 \\
-            Omnivory: bottom & 0.882 & -1.06 &	-0.0449 &	0.845 \\
-            \hline
-            App. Comp.: top & 0.865 &	-0.966 &	0.140 &	-0.520 \\
-            App. Comp.: bottom & 0.910 &	-1.12 &	-0.100 &	0.323 \\
-            \hline
-            \end{tabular}
-        \end{table}
-
-        % Organizing results by motif - keep this OR by no. prey
+        A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance (Table SX, \emph{Appendix SH}).
         In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
@@
         \includegraphics[width=.95\textwidth]{figures/persistence_positions_bymotif.eps}
-        \label{figs:persistence_position_motifs}
         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
@@
 Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.}
+        \label{figs:persistence_position_motifs}
         \end{figure}
 
-        % Organizing by number of prey
-        Across the four motifs we consider, positions with no prey within the motif tended to show weak relationships to persistence at low levels of disturbance and be associated with increased persistence at high levels of disturbance (Fig.~\ref{fig:persistence_positions_prey}).
-        These trends were strongest for direct competition and omnivory and weakest for apparent competition.
-        The two positions with stronger effects indicate a species with two predators within the motif, while the weaker effects are for motifs with one predator within the motif.
-
-        Positions with one prey within the motif were more variable.
-        The top position in the direct competition motif and middle position in the omnivory motif (both of which have one prey) were associated with increased persistence at low levels of disturbance and lower persistence at higher levels of disturbance, but these effects were relatively weak.
-        The top and middle positions of the three-species chain both include a single prey; the middle position was associated with increased persistence at all levels of disturbance while the top position was associated with decreased persistence at all levels of disturbance.
-
-        Positions with two prey in the apparent competition and omnivory motifs showed very similar patterns.
-        At low levels of disturbance, persistence increased with the frequency of either position.
-        At high levels of disturbance, persistence decreased as the frequency of either motif increased.
-        This echoes the pattern for in-degree.
-
-        \begin{figure}
-            \centering
-            \includegraphics[width=.95\textwidth]{figures/persistence_positions_byprey.eps}
-            \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
-Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common.
-Line colour indicates the motif containing a position.
-Line style indicates the number of prey within the motif for each position.
-Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.
-The top set of lines corresponds to a probability of extinction for basal resources of 0.1.
-The middle set of lines corresponds to a probability of extinction for basal resources of 0.26, and the lowest set of lines corresponds to a probability of extinction for basal resources of 0.5.}
-            \label{fig:persistence_positions_prey}
-        \end{figure}
-
-    \subsection*{How do motif positions vary with other network properties?}
-
-        \subsubsection*{Degree and trophic level}
-
-            The top position in the omnivory motif increased with increasing degree across all trophic levels.
-            The frequency of the middle position increased with the interaction of degree and trophic level.
-            The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).
-
-            The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
-            The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.
-
-            The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
-            The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).
-
-            The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
-            The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
-            The top position in the three-species chain increased with trophic level, degree, and their interaction.
-
-        \subsubsection*{Network size and connectance}
-
-            % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
-            The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance.
-            The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.
-
-            The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
-            The frequency of the bottom position in the direct competition motif was low in all networks.
-
-            The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
-            Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
-            None of the positions in the three-species chain varied strongly with network size or connectance.
+        The frequencies of motif positions in a species' role varied with network properties (\emph{Appendix SI}).
+        Motif roles had stronger relationships with degree and STL than with network size or connectance.
 
@@
 
-    [[Big deal result: connectance is important for motif structure of network, size not so much]]
+Somewhere early in discusssion (needs tidying up):
 
-    [[Persistence depends on S, C, interaction with $basal_p$ (lowest persistence swaps from highest S to lowest C in Figs.
-1-2, high C in both cases - but small absolute differences so maybe we don't want to make too big a deal out of this)]]
+Apparent comp and direct comp have no indirect effects in BN, so we should be able to explain them with simpler properties.
+AC is strongly negatively correlated with in-degree.
+Higher in degree is associated with both high persistence at low disturbance (Fig.
+7A) and low participation in AC (Fig.
+7C).
+Low participation in AC is in turn associated with high persistence at low disturbance (Fig.
+4).
+Note that trends for degree and AC reverse at high disturbance in the same way!
+DC is not strongly correlated with degree but does decrease strongly at high trophic levels.
+High trophic level and low DC are both associated with low persistence at all levels of disturbance, making it likely that the effects we observe for DC are trophic level in disguise.
 
-    [[proportion of omnivory increases with increasing C, while the other doesn't. Connect to later findings involving omnivory]]
-
-\clearpage
-
-    \subsection*{How does persistence vary with motif positions?}
-
-        A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance (Table S7, \emph{Appendix SH}).
-
-        % Organizing results by motif - keep this OR by no. prey
-        In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
-        Peristence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
-        Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.
-
-        In the three-species chain, persistence decreased as the frequency of the top position increased at all levels of disturbance.
-        Conversely, persistence increased with increasing frequency of the middle position at all levels of disturbance.
-        Persistence decreased slightly with increasing frequency of the bottom position at low levels of disturbance but increased at high levels of disturbance.
-
-        In the apparent competition motif, persistence decreased slightly with increasing frequencies of the bottom position (one predator, no prey within the motif) at low disturbance but increased slightly at high disturbance.
-        Persistence increased with increasing frequency of the top position (two prey, no predator within the motif) at low disturbance but decreased at high disturbance.
-
-        In the direct competition motif, persistence varied little with the frequency of the bottom position (two predators, no prey within the motif) at low disturbance but increased with the frequency of this position at high levels of disturbance.
-        Persistence increased slightly with the frequency of the top position (one prey, no predators within the motif) at low levels of disturbance and decreased slightly at high levels of disturbance.
-
-        \begin{figure}[h!]
-        \includegraphics[width=.95\textwidth]{figures/persistence_positions_bymotif.eps}
-        \label{figs:persistence_position_motifs}
-        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
-Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common.
-Line colour indicates the probability of extinction for basal resources.
-Line style indicates the number of prey within the motif for each position.
-Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.}
-        \end{figure}
-
-        % Organizing by number of prey
-        Across the four motifs we consider, positions with no prey within the motif tended to show weak relationships to persistence at low levels of disturbance and be associated with increased persistence at high levels of disturbance (Fig.~\ref{fig:persistence_positions_prey}).
-        These trends were strongest for direct competition and omnivory and weakest for apparent competition.
-        The two positions with stronger effects indicate a species with two predators within the motif, while the weaker effects are for motifs with one predator within the motif.
-
-        Positions with one prey within the motif were more variable.
-        The top position in the direct competition motif and middle position in the omnivory motif (both of which have one prey) were associated with increased persistence at low levels of disturbance and lower persistence at higher levels of disturbance, but these effects were relatively weak.
-        The top and middle positions of the three-species chain both include a single prey; the middle position was associated with increased persistence at all levels of disturbance while the top position was associated with decreased persistence at all levels of disturbance.
-
-        Positions with two prey in the apparent competition and omnivory motifs showed very similar patterns.
-        At low levels of disturbance, persistence increased with the frequency of either position.
-        At high levels of disturbance, persistence decreased as the frequency of either motif increased.
-        This echoes the pattern for in-degree.
-
-        \begin{figure}
-            \centering
-            \includegraphics[width=.95\textwidth]{figures/persistence_positions_byprey.eps}
-            \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
-Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common.
-Line colour indicates the motif containing a position.
-Line style indicates the number of prey within the motif for each position.
-Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.
-The top set of lines corresponds to a probability of extinction for basal resources of 0.1.
-The middle set of lines corresponds to a probability of extinction for basal resources of 0.26, and the lowest set of lines corresponds to a probability of extinction for basal resources of 0.5.}
-            \label{fig:persistence_positions_prey}
-        \end{figure}
-
-    \subsection*{How do motif positions vary with other network properties?}
-
-        The frequencies of motif positions in a species' role varied strongly and significantly with degree and trophic level, but only weakly with network size and connectance (\emph{Appendix SI}).
-
-           \subsubsection*{Degree and trophic level}
-
-            The top position in the omnivory motif increased with increasing degree across all trophic levels.
-            The frequency of the middle position increased with the interaction of degree and trophic level.
-            The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).
-
-            The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
-            The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.
-
-            The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
-            The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).
-
-            The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
-            The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
-            The top position in the three-species chain increased with trophic level, degree, and their interaction.
-
-        \subsubsection*{Network size and connectance}
-
-            % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
-            The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance.
-            The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.
-
-            The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
-            The frequency of the bottom position in the direct competition motif was low in all networks.
-
-            The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
-            Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
-            None of the positions in the three-species chain varied strongly with network size or connectance.
+Three-species chain and omnivory do have indirect effects in a BN setting, and they can't be fully explained by degree or trophic level (since the result doesn't follow the result for degree).
+We can see this because despite the strong correlations with degree and trophic level these motifs don't show the same patterns with persistence.
+We therefore focus on them here.
 
 \section*{Discussion}
-% * Apparent competition change from bad to good
+% Connect back to intro -- importance of the plant community for a healthy consumer community.
+% Focus on what additional information is revealed from analyzing motif compared to global network properties or species characteristics.
+% Connect to traits -- Cirtwill & Eklöf, 2018, Ecol Lett. 
+% Bayesian network approach - what extra information does motifs provide when we are interested in bottom- up effects. 
 % * Difference between 3-sp chain and omnivory, the importance of the rescue prey in omnivory (in dynamical sims omni was mostly bad, 3-sp chain good).
@@
 % Results summary - network level (following Anna)
-We analysed how network properties at a variety of scales influence consumers' probabilities of extinction.
-First, we considered how global-scale and meso-scale structures were related to the average persistence within a network.
-Consistent with earlier work, networks with higher connectance or species richness tend to have lower persistences~\citep{}.
-In our simulations, however, these effects were relatively small.
-In contrast, and consistent with other studies of motif profiles, changes in the motif profiles of networks (i.e., the frequencies of different motifs across the whole food web) had relatively strong relationships to the mean persistence of species within the web~\citep{prill2005dynamic, bascompte2005simple}.
-As changing connectance (and, to a lesser extent) network size 
-affected network motif profiles, it appears that global-scale properties have stronger effects on persistence by changing meso-scale properties than directly.
-In all cases, the effect of increasing disturbance to basal resources had a much larger effect on consumer persistences than did network structure.
+Changes in the primary producer level in food webs have profound effects on all other species in the network \citep{}.
+In particular, the disturbances can percolate across trophic levels as all consumers depend directly or indirectly on primary producers for their survival \citep{}.
+Understanding which characteristics of a network and its species influence the severity of outcomes after a disturbance is crucial, e.g.
+for conservation actions \citep{}.
+Here we were particularly interested in how the distribution of meso-scale structures in ecological networks are related to average network persistence when the basal level is disturbed by species loss.
+In order to disentangle the effects caused by the motif distribution we also analyzed global (network-wide) and local (species) properties.
+%Consistent with earlier work, networks with higher connectance or species richness tend to have lower persistences~\citep{}.
+A network's motif profile had relatively strong relationships to the network mean persistence, while effects from global scale structures, such as network size and connectance, were relatively small. 
+However, connectance (and, to a lesser extent network size) do affect network motif profiles.
+Thus, global-scale properties seem to have stronger effects on network persistence by changing meso-scale properties than directly.
+~\citep{prill2005dynamic, bascompte2005simple}.
+
+%In all cases, the effect of increasing disturbance to basal resources had a much larger effect on consumer persistences than did network structure.
 This is a consequence of the bottom-up calculations done to receive the final persistence.
@@
 % Network-level motif profiles, persistence, and relationships to S and C.
-In general, persistence was highest in networks with high frequencies of the apparent competition motif (two prey consumed by the same predator) and lowest in networks with high frequencies of the omnivory motif (a three-species food chain in which the top species also consumes the bottom species) while the direct competition and three-species chain motifs had smaller effects.
-The negative relationship between omnivory and mean persistence is consistent with modelling work showing that an omnivory motif is less likely to retain all three species after a disturbance than the other motifs we consider~
-\citep{Borrelli2015a}.
+A network's mean persistence depended on which motifs were most common in the network profile.
+In general, persistence was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif.
+Direct competition and three-species chain motifs had smaller effects.
+The effects were clearly pronounced when disturbance was low but attenuated with increased disturbance.
+The negative relationship between omnivory and mean persistence is consistent with other studies showing that an omnivory motif is less likely to be stable than the other motifs considered~\citep{Borrelli2015a}.
+
+High frequency of the apparent competition motif was correlated with an on average higher in-degree for consumers. In-degree in turn is often related to higher persistence \citep{}.
+
 Motif profiles with high apparent competition and low omnivory were also more likely in less-connected food webs, which had higher mean persistence than highly-connected food webs.
@@
 High connectance might reduce mean persistence by limiting the number of possible meso-scale network structures.
-In particular, it is reasonable to expect higher frequencies of the omnivory motif at high connectances since this motif contains more links than the other three-species motifs we consider.
+In particular, it is reasonable to expect higher frequencies of the omnivory motif at high connectances since this motif contains more links than the other three-species motifs considered here.
+% Top species directly connected to the disturbance, disturbance not mitigated via herbivores. Duality in the role of omnivory.
 % The relationship between high connectance and frequency of omnivory motif should perhaps come earlier. 
-% Close with something about how more links allow disturbance to propagate - borrow from connectance literature.
+% Close with something about how more links allow disturbance to propagate.
 
 As well as the mean persistence of all consumers, we also considered the persistence of each consumer in a network.
-While the persistence of a particular species is likely to be affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it will also depend on smaller-scale measures of the species' place in the network.
-In particular, a species' motif participation (the frequency with which it appears in each of the four possible motifs) was significantly related to its probability of persistence.
+While the persistence of a particular species is likely to be affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it will also depend on species-specific characteristics.
+In particular, a species' motif participation was significantly related to its probability of persistence.
 Specifically, a high proportion of the direct competition or omnivory motifs increase the probability of persistence, while a high proportion of the three-species chain motif decreases probability of persistence.
@@
 Combining these results with our network-level findings, we see that while being in a network with high frequencies of omnivory decreases persistence, participating in many of these omnivory motifs increases persistence at the species level.
-At the network level, a higher frequency of the omnivory motif leads to more pathways by which disturbances to basal resources can propogate and cause extinctions of consumers.
-If extinctions of basal resources are rare, these multiple pathways will tend to protect consumers by providing alternative channels of energy.
+At the network level, a higher frequency of the omnivory motif leads to more pathways by which disturbances to basal resources can propagate and cause extinctions of consumers.
+At a species level, if extinctions of basal resources are rare, these multiple pathways will tend to protect consumers participating in many omnivory motifs by providing alternative channels of energy.
 If extinctions of basal resources are common, however, these multiple pathways cause a negative impact as they will transmit more disturbances to the consumer.
@@
 The offsetting of effects of different positions within the omivory motif may explain why the motif as a whole is associated with greater probabilities of persistence across all levels of disturbance we considered.
+
+%The different positions in the three-species chain motif - connect to network stability?
+3-sp chain always good in fig 3 - when a full food web has a high proportion of this motif.
+In fig 4, a species with a high proportion of the 3-sp chain always negative for persistence, most so with increasing disturbance.
+In fig 8, mainly the top position that reduces persistence, more so with high disturbance.
+Bottom and middle position largely unchanged.
+
+% in app comp - bottom species drive the trend visible in fig 4? the top species shows opposite patterns anyway.
 
@@
 In our Bayesian Network setting, this link is indistinguishable from any other link in the food web which partly may explain why omnivory results in increased persistence of consumer species.
-
-    %These results can be explained by the structure of each motif.
-When a predator has a high proportion of the apparent competition motif, it tends to have more prey items.
-When the bottom-level is heavily disturbed, this becomes beneficial since if some prey items are lost, other might still exist.
-[but why is it bad to have a high proportion of app comp motif if disturbances are low?]
-    %Participating to a high extent in the direct competition motif means that a predator participates in many two-level food chains.
-It does not necessarily mean that the predator has few prey items, since it may participate in many direct competition motif.
-[why always good with high prop of direct comp motif?
-]
 

### 2021-06-28 code/stat_analysis/persistence_vs_positions.py
one motif at a time

@@ -47,7 +47,7 @@ def read_datafile(datafile):
   f.close()
   return netprops
 
-def format_graph(graph,simple):
+def format_graph(graph,simple,motif):
   graph.yaxis.bar.linewidth=1
   graph.xaxis.bar.linewidth=1
   graph.frame.linewidth=1
@@ -63,7 +63,7 @@ def format_graph(graph,simple):
     graph.yaxis.tick.major=.1
     graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
     graph.yaxis.label.text='Mean persistence'
-    graph.xaxis.label.text='Proportion of network profile'
+    graph.xaxis.label.text='Proportion of '+motif+' in network profile'
 
     graph.xaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
     graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
@@ -93,19 +93,26 @@ def populate_persgraph(graph,motif,datadict):
     npred=int(pos.split('..')[1])
     nprey=int(pos.split('..')[2].split('.')[0])
     j=12
-    if nprey==0:
-      linestyle=2
-    elif nprey==1:
-      linestyle=3
-    else:
-      linestyle=1
+    if motif=='Omnivory':
+      if nprey==0:
+        linestyle=2
+      elif nprey==1:
+        linestyle=3
+      else:
+        linestyle=1
+    elif motif=='Three-species chain':
+      if nprey==0:
+        linestyle=2
+      if npred==0:
+        linestyle=1
+      if nprey==1 and npred==1:
+        linestyle=3
+    else: # both competition motifs can use same code
+      if npred==0:
+        linestyle=1
+      else:
+        linestyle=2
 
-    if npred==0:
-      shap=3
-    elif npred==1:
-      shap=1
-    else:
-      shap=2
 
     # betas go intercept, disturbance, position, interaction
     lms=datadict['betas'][pos]
@@ -118,130 +125,55 @@ def populate_persgraph(graph,motif,datadict):
         dats.append((x,y))
 
       data=graph.add_dataset(dats)
-      data.symbol.configure(shape=shap,color=j,fill_color=j)
+      data.symbol.configure(shape=0,color=j,fill_color=j)
       data.line.configure(linestyle=linestyle,linewdith=1.5,color=j)
 
[…]

### 2021-06-28 code/stat_analysis/roles_vs_degTL_collector.py

@@ -1,166 +1,0 @@
-import sys
-import os
-import decimal
-
-
-def permfile_reader(permfile):
-	temp={}
-	f=open(permfile,'r')
-	for line in f:
-		if line.split()[0]=='"1"':
-			output=float(line.split()[1])
-	f.close()
-	return output
-
-
-def pval_calculator(model_F,perms):
-	greater=[x for x in perms if x>=model_F]
-
-	p=float(len(greater))/float(len(perms))
-
-	return p
-
-
-def recreate_aovtab(datadict,outfile):
-
-	f=open(outfile,'w')
-	f.write('S\tC\text_corr\tdf\tSS\tMS\tModel_F\tR2\tpval\n')
-	for C in sorted(datadict):
-		dats=datadict[C]
-		f.write(str(dats['S'])+'\t'+str(dats['C'])+'\t'+str(dats['ext_corr'])+'\t')
-		f.write(str(dats['df'])+'\t'+str(dats['SS'])+'\t'+str(dats['MS'])+'\t')
-		f.write(str(dats['model_F'])+'\t'+str(dats['R2'])+'\t'+str(dats['pval'])+'\n')
-	f.close()
-
-
-def betadisper_reader(dispfile):
-	f=open(dispfile,'r')
-	for line in f:
-		if line.split()[0]=='"Groups"':
-			df=int(line.split()[1])
-			SSE=float(line.split()[2])
-			MSE=float(line.split()[3])
-			F=float(line.split()[4])
-			p=float(line.split()[5])
-	f.close()
-
-	result={'F':F,'p':p}
-	return result
-
-
-def tukey_reader(tukeyfile):
-	results={}
-	f=open(tukeyfile,'r')
-	for line in f:
-		if line.split()[0]!='"diff"':
-			groups=line.split()[0][1:-1]
-			diff=float(line.split()[1])
-			lower=float(line.split()[2])
-			upper=float(line.split()[3])
-			p=float(line.split()[4])
-			results[groups]=((diff,p))
-	f.close()
-
-	return results
-
-
-def tukey_outfile(dicto,style):
-	f=open('../../data/summaries/'+style+'_Tukey.tsv','w')
-	f.write('Species\tC\tGroup1\tGroup2\tDiff\tP\n')
-
-	for s in dicto:
-		for c in dicto[s]:
[…]

### 2021-06-28 code/stat_analysis/statistical_models.R
cleaning up last changes

@@ -218,19 +218,19 @@ consumers$prop_omni=consumers$m38/consumers$total_motifs
 save.image('all_tests.Rdata')
 
 
-# Do proportions of motifs vary due to network processing?
+# # Do proportions of motifs vary due to network processing?
 
-original_data=read.table('../../data/3sp_motif_profiles_cyclicwebs.tsv',header=TRUE)
-original_data$prop_chain=original_data$m12/original_data$Total
-original_data$prop_apparent=original_data$m6/original_data$Total
-original_data$prop_direct=original_data$m36/original_data$Total
-original_data$prop_omni=original_data$m38/original_data$Total
-original_data$prop_other=(original_data$Total-rowSums(original_data[,4:7]))/original_data$Total
+# original_data=read.table('../../data/3sp_motif_profiles_cyclicwebs.tsv',header=TRUE)
+# original_data$prop_chain=original_data$m12/original_data$Total
+# original_data$prop_apparent=original_data$m6/original_data$Total
+# original_data$prop_direct=original_data$m36/original_data$Total
+# original_data$prop_omni=original_data$m38/original_data$Total
+# original_data$prop_other=(original_data$Total-rowSums(original_data[,4:7]))/original_data$Total
 
-data$label=paste(data$Size,data$Connectance,data$Network,sep=':')
-original_data$label=paste(original_data$Size,original_data$Connectance,original_data$Network,sep=':')
+# data$label=paste(data$Size,data$Connectance,data$Network,sep=':')
+# original_data$label=paste(original_data$Size,original_data$Connectance,original_data$Network,sep=':')
 
-used=original_data[which(original_data$label%in%data$label),]
+# used=original_data[which(original_data$label%in%data$label),]
 
 
 

### 2021-06-28 code/stat_analysis/persistence_vs_positions.py
fixing x axis

@@ -63,7 +63,7 @@ def format_graph(graph,simple,motif):
     graph.yaxis.tick.major=.1
     graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
     graph.yaxis.label.text='Mean persistence'
-    graph.xaxis.label.text='Proportion of '+motif+' in network profile'
+    graph.xaxis.label.text='Proportion of '+motif+' positions in network profile'
 
     graph.xaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
     graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)

### 2021-06-28 manuscript/motifs_vs_degree.tex

@@
      The plant community is the foundation upon which the myriad of species in food webs depend for their survival \citep{}.
-Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{} and secondary production \citep{}, soil respiration and carbon cycling \citep{} and consumer diversity \citep{scherber2010bottom,Baiser2013}.
+Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{} and secondary production \citep{}, soil respiration and carbon cycling \citep{} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
 The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community promotes stability across all different levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}.
@@
 Because of this restriction, we only consider these four particularly important motifs.
-    % only four do not include cycles (e.g., A eats B and B eats A or A eats B, B eats C, and C eats A). 
-    %As cycles must be removed from networks before calculating extinction probabilities in the Bayesian network framework, this study focuses on these four Bayesian-compatible motifs:
-
-    %When analyzing motifs within a Bayesian Network, it is important to keep in mind that only bottom-up effects are accounted for.
-Therefore, the implications of the four possible motifs are different than they would be in an empirical network.
-
-    %In particular, the lack of top-down effects removes the indirect effects from the two competition motifs.
-In the case of direct competition, two predators share a common resource.
-The predators will, within a Bayesian Network, not affect each other indirectly although both will be affected by any disturbance to their resource.
-Similarly, the two bottom-level species in the apparent competition motif will be independent of each other as there are no top-down effects of the shared consumer to link them.
-    %The difference between the direct competition and apparent competition motif is thus the number of bottom-level species, accounted for as prey items for a higher level species.
-   % A predator participating in many direct competition motifs is likely to have fewer prey items than a predator participating frequently in the apparent competition motif, and therefore a higher extinction risk.Note, however, that consumers can also appear in these motifs as prey to other consumers.
-In that case, we would not expect the two motifs to have different effects on extinction risk.
-    %The structure of the omnivory motif and the three-species chain motif differ from the competition motifs by having an additional trophic level.
-Therefore, top-level species of these motifs might be more sensitive to disturbances since they can be affected by the loss of the basal resource in the focal motif or an increased probability of extinction of the middle species due to a disturbance to a basal resource not included in the focal motif.
 
@@
             Networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
-            However, there was an interaction disturbance and connectance for the direct competition and omnivory motifs.
+            However, there was an interaction between disturbance and connectance for the direct competition and omnivory motifs.
 
-            At low and medium connectances there is little change in the proportion of positive slopes (\textgreater80\%) for the different levels of disturbance. 
+            At low and medium connectances with a high proportion of the direct competition motif, there is little change in the proportion of positive slopes (\textgreater80\%) for the different levels of disturbance.
             At high connectances, however, the proportion of positive slopes is lower overall and decreases with increasing disturbance. 
@@
                 Persistence declined with inceasing STL for all levels of disturbance to basal resources.
-                \textbf{C-D)} Counts of all motifs increased with increasing in-degree and decreased with increasing STL.
-The proportion of omnivory increased with increasing degree while all other proportions decreased.
+                \textbf{C-D)} The proportion of omnivory increased with increasing degree while all other proportions decreased.
 The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.}
@@
 
-Somewhere early in discusssion (needs tidying up):
-
-Apparent comp and direct comp have no indirect effects in BN, so we should be able to explain them with simpler properties.
-AC is strongly negatively correlated with in-degree.
+%Somewhere early in discussion (needs tidying up):
+%Apparent comp and direct comp have no indirect effects in BN, so we should be able to explain them with simpler properties.
+%AC is strongly negatively correlated with in-degree.
 Higher in degree is associated with both high persistence at low disturbance (Fig.
@@
 Note that trends for degree and AC reverse at high disturbance in the same way!
-DC is not strongly correlated with degree but does decrease strongly at high trophic levels.
+%DC is not strongly correlated with degree but does decrease strongly at high trophic levels.
 High trophic level and low DC are both associated with low persistence at all levels of disturbance, making it likely that the effects we observe for DC are trophic level in disguise.
-
-Three-species chain and omnivory do have indirect effects in a BN setting, and they can't be fully explained by degree or trophic level (since the result doesn't follow the result for degree).
+%Three-species chain and omnivory do have indirect effects in a BN setting, and they can't be fully explained by degree or trophic level (since the result doesn't follow the result for degree).
 We can see this because despite the strong correlations with degree and trophic level these motifs don't show the same patterns with persistence.
@@
 
-%In all cases, the effect of increasing disturbance to basal resources had a much larger effect on consumer persistences than did network structure.
-This is a consequence of the bottom-up calculations done to receive the final persistence.
-The importance of the basal level for network robustness has been demonstrated in numerous studies \citep{scherber2010bottom}, as well as the sensitivity of high trophic levels (REF).
-
 % Network-level motif profiles, persistence, and relationships to S and C.
@@
 
-As well as the mean persistence of all consumers, we also considered the persistence of each consumer in a network.
-While the persistence of a particular species is likely to be affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it will also depend on species-specific characteristics.
-In particular, a species' motif participation was significantly related to its probability of persistence.
-Specifically, a high proportion of the direct competition or omnivory motifs increase the probability of persistence, while a high proportion of the three-species chain motif decreases probability of persistence.
-%These trends held at any level of disturbance we considered, although at high levels of disturbance the positive effect of omnivory was reduced and the negative effect of three-species chains was increased.
-%The effect of the apparent competition motif, however, varied with disturbance.
-%A high proportion of the apparent competition motif reduces persistence at low levels of disturbance but increases persistence at high levels of disturbance.
+% HERE ENDS MEAN PERSISTENCE AND SPECIES WISE START
+As well as the mean persistence of all consumers, we also considered the species-specific persistence of each consumer in a network.
+While the persistence of a particular species is likely to be affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it also depends on a species' position in the network.
+There are many ways of measuring species' positions, and in some cases multiple measures may capture the same trend.
+In our results, motif participation is significantly related to persistence, but the effects of two motifs (\emph{apparent competition} and \emph{direct competition}) might be explained more simply by in-degree and trophic level.
+This is especially likely for these two motifs, since Bayesian Network modeling lacks top-down effects and therefore removes the indirect effects from these motifs.
+
+% Getting rid of AC/DC "quickly" here so we can focus on omnivory
+Participation in the apparent competition motif is strongly negatively correlated with in-degree---a high proportion of the apparent competition motif implies a low in-degree. 
+Both higher in-degree and lower participation in the apparent competition motif are associated with greater persistence at low disturbance but decreased persistence at high disturbance; because the trends are so similar, it is likely that a common mechanism underlies both.
+% The effect on persistence is consistent between species with a low in-degree and a high proportion of the apparent competition motif. 
+%Low in-degree decreases persistence at low disturbance, and increases persistence at high disturbance, similar to the trend visible for a high participation in the apparent competition motif. 
+% Similarly, the effect on species persistence is reversed for a low proportion of the apparent competition motif in which species have a high in-degree.
+
+While the direct competition motif is not correlated with in-degree, it is strongly negatively correlated with trophic level---species with a high proportion of the direct competition motif have a low trophic levels.
+Low trophic levels are always associated with higher persistence because these species are less affected by bottom-up disturbances~\citep{}.
+It then follows that species which participate in many direct competition motifs may have higher probabilities of persistence because they are also likely to have low trophic levels.
+
+In contrast to the competition motifs, the three-species chain and omnivory motifs do have indirect effects in a Bayesian Network setting and the effects of these motifs cannot be fully explained by degree or trophic level.
+For example, the three-species chain motif is negatively correlated with in-degree and positively correlated with trophic level. 
+Despite this correlation, participation in the three-species chain does not interact the same way with disturbance as either simpler measure and so has a unique relationship to species persistence.
+This is likely due to indirect effects of the bottom species in the chain on the top species.
+
+The most interesting motif in which to consider indirect effects is omnivory.
+While participation in the omnivory motif is positively correlated with in-degree, a high proportion of the omnivory motif is always associated with increased persistence, while the effect of in-degree varies with disturbance.
+Thus, the effect of a high proportion of the three-species chain motif and the omnivory motif does not follow the corresponding effect of in-degree and trophic level.
 
 % Two notable results: 1) omnivory bad at network level, good at species level, 2) apparent competition depends on disturbance
-Combining these results with our network-level findings, we see that while being in a network with high frequencies of omnivory decreases persistence, participating in many of these omnivory motifs increases persistence at the species level.
-At the network level, a higher frequency of the omnivory motif leads to more pathways by which disturbances to basal resources can propagate and cause extinctions of consumers.
-At a species level, if extinctions of basal resources are rare, these multiple pathways will tend to protect consumers participating in many omnivory motifs by providing alternative channels of energy.
+Combining the results from our species wise exploration with our network-level findings, we see that while being in a network with high frequencies of omnivory decreases persistence, participating in many of these omnivory motifs increases persistence at the species level.
+At the network level, a higher frequency of the omnivory motif leads to more pathways by which the disturbances to basal resources can propagate and cause extinctions of consumers.
+At the species level on the other hand, if extinctions of basal resources are rare, these multiple pathways will tend to protect consumers participating in many omnivory motifs by providing alternative channels of energy.
 If extinctions of basal resources are common, however, these multiple pathways cause a negative impact as they will transmit more disturbances to the consumer.
+
+Additionally, the effect of the omnivory motif on persistence was closely connected to connectance.
+In highly connected networks, a high proportion of the omnivory motif strongly contributed to higher persistence.
+This was not the case in less connected networks.
+In a highly connected network, a disturbance could potentially spread to a higher extent than in a less connected network. 
+Therefore, in a highly connected web where disturbances have more ways to move, the lower level prey item in the omnivory motif may be of even higher importance assisting as a buffer, decreasing the negative effect of disturbance for the consumer species.
+In a less connected network, this additional prey species is of less importance.
+
 To understand the species-level effect, we consider the relationships between stability and each position within the motif.
@@
 The offsetting of effects of different positions within the omivory motif may explain why the motif as a whole is associated with greater probabilities of persistence across all levels of disturbance we considered.
-
-%The different positions in the three-species chain motif - connect to network stability?
-3-sp chain always good in fig 3 - when a full food web has a high proportion of this motif.
-In fig 4, a species with a high proportion of the 3-sp chain always negative for persistence, most so with increasing disturbance.
-In fig 8, mainly the top position that reduces persistence, more so with high disturbance.
-Bottom and middle position largely unchanged.
-
-% in app comp - bottom species drive the trend visible in fig 4? the top species shows opposite patterns anyway.
 
@@
 % Although species in our model cannot swap prey dependent on bottom-level disturbances or lower resource densities, a species with a high proportion of omnivory will in a sense spread its risk between prey items at different trophic levels.
-% Additionally, the effect of the omnivory motif on persistence was closely connected to connectance.
-In highly connected networks, a high proportion of the omnivory motif strongly contributed to higher persistence.
-This was not the case in less connected networks.
-% In a highly connected network, a disturbance could potentially spread to a higher extent than in a less connected network. 
-% Therefore, in a highly connected web where disturbances have more ways to move, the lower level prey item in the omnivory motif may be of even higher importance assisting as a buffer, decreasing the negative effect of disturbance for the consumer species.
-In a less connected network, this additional prey species is of less importance.
 
@@
 
-Our result contrasts with \citep{Cirtwill2021_inprep}, where a high proportion of the omnivory motif resulted in a shorter time to extinction, while the other three stable motifs were associated with longer times to extinction.
-However, the study by Cirtwill and Wootton used a different simulation approach, with body-mass structured vertebrate food webs and a dynamic simulation model accounting for top-down effects.
+ \cite{Cirtwill2021_inprep} analysed how species time to extinction varied depending on their network profile.
+They found that a high proportion of the omnivory motif resulted in a shorter time to extinction, while the other three stable motifs were associated with longer times to extinction.
+However, the study by Cirtwill and Wootton used a different approach using a dynamic simulation model also including top-down effects.
     %In previous studies of the effect of omnivory motifs, the strength of the omnivorous link can potentially be important.
@@
 
+ %When analyzing motifs within a Bayesian Network, it is important to keep in mind that only bottom-up effects are accounted for.
+Therefore, the implications of the four possible motifs are different than they would be in an empirical network.
+
+    %In particular, the lack of top-down effects removes the indirect effects from the two competition motifs.
+In the case of direct competition, two predators share a common resource.
+The predators will, within a Bayesian Network, not affect each other indirectly although both will be affected by any disturbance to their resource.
+Similarly, the two bottom-level species in the apparent competition motif will be independent of each other as there are no top-down effects of the shared consumer to link them.
+    %The difference between the direct competition and apparent competition motif is thus the number of bottom-level species, accounted for as prey items for a higher level species.
+   % A predator participating in many direct competition motifs is likely to have fewer prey items than a predator participating frequently in the apparent competition motif, and therefore a higher extinction risk.Note, however, that consumers can also appear in these motifs as prey to other consumers.
+In that case, we would not expect the two motifs to have different effects on extinction risk.
+    %The structure of the omnivory motif and the three-species chain motif differ from the competition motifs by having an additional trophic level.
+Therefore, top-level species of these motifs might be more sensitive to disturbances since they can be affected by the loss of the basal resource in the focal motif or an increased probability of extinction of the middle species due to a disturbance to a basal resource not included in the focal motif.
+
 % [[When can motifs still be useful predictors and add insight?]]
@@
 \bibliographystyle{ecollett} 
-\bibliography{anna_bib} % Abbreviate journal titles.
+\bibliography{anna_bib_new} % Abbreviate journal titles.
 

### 2021-06-28 manuscript/motifs_vs_degree.tex

@@
     \begin{figure}[h!]
-        \includegraphics[width=.95\textwidth]{figures/persistence_positions_bymotif.eps}
+        \includegraphics[width=.95\textwidth]{figures/persistence_positions_Omnivory.eps}
         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
-Relationships are shown for the range 0-0.8 for all positions, but some positions (especially in the omnivory motif) are less common.
-Line colour indicates the probability of extinction for basal resources.
-Line style indicates the number of prey within the motif for each position.
-Symbol shape indicates the number of predators within the motif for each position; note that only bottom-up extinctions were included in our simulations.}
+Here we show the relationships between persistence and the proportions of the three positions in the omnivory motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
+Line style indicates the motif  position.
+}
         \label{figs:persistence_position_motifs}

### 2021-06-29 code/stat_analysis/persistence_vs_profiles.py
persistence vs motif profiles updated

@@ -154,9 +154,21 @@ def populate_graph(graph,bindict):
 
 def populate_persgraph(graph,motif):
 
-  j=12
-  for dist in [0.1,0.18,0.26,0.34,0.42,0.5]:
+  for dist in [0.1,0.5]:
+    if motif=='apparent':
+      j=13
+      sty=2
+    elif motif=='chain':
+      j=14
+      sty=3
+    elif motif=='direct':
+      j=16
+      sty=5
+    else:
+      j=18
+      sty=1
 
+    # for dist in [0.1,0.18,0.26,0.34,0.42,0.5]:
     scaldist=(dist-scales['Disturbance'][0])/scales['Disturbance'][1]
 
     dats=[]
@@ -171,18 +183,11 @@ def populate_persgraph(graph,motif):
 
     data=graph.add_dataset(dats)
     data.symbol.shape=0
-    data.line.configure(linestyle=1,linewdith=1.5,color=j)
+    data.line.configure(linestyle=sty,linewdith=1.5,color=j)
 
+    if dist==0.1:
+      data.legend=motif_names[motif]
 
-    if motif=='omnivory':
-      data.legend=str(dist)
-    j+=1
-
-  if motif=='omnivory':
-    graph.add_drawing_object(DrawText,text='Basal species',x=0.5,y=0.755,loctype='world',char_size=.5)
-    graph.add_drawing_object(DrawText,text='extinction',x=0.5,y=0.725,loctype='world',char_size=.5)
-    graph.add_drawing_object(DrawText,text='probability',x=0.5,y=0.695,loctype='world',char_size=.5)
-    graph.legend.configure(char_size=.5,loc=(0.5,0.685),loctype='world',box_linestyle=0,fill_pattern=0)
   return graph
 
 ###############################################################################################
@@ -197,24 +202,26 @@ def populate_persgraph(graph,motif):
 # datafile='motif_proportions_persistence.tsv'
 # netprops=read_datafile(datafile)
 
-grace=MultiPanelGrace(colors=colors)
-grace.add_label_scheme('dummy',['Omnivory','Three-species chain','Apparent competition','Direct competition'])
-grace.set_label_scheme('dummy')
+grace=Grace(colors=colors)
+# grace.add_label_scheme('dummy',['Omnivory','Three-species chain','Apparent competition','Direct competition'])
+# grace.set_label_scheme('dummy')
 
[…]

### 2021-06-29 manuscript/supplement_motifs.tex

@@
 
+    \begin{figure}
+        \centering
+        \includegraphics{figures/persistence_positions_Apparent.eps}
+        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
+Here we show the relationships between persistence and the proportions of the three positions in the apparent competition motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
+Line style indicates the motif  position.
+}
+        \label{fig:apparent_positions}
+    \end{figure}
+
+    \begin{figure}
+        \centering
+        \includegraphics{figures/persistence_positions_Direct.eps}
+        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
+Here we show the relationships between persistence and the proportions of the three positions in the direct competition motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
+Line style indicates the motif  position.
+}
+        \label{fig:direct_positions}
+    \end{figure}
+
+    \begin{figure}
+        \centering
+        \includegraphics{figures/persistence_positions_Three-species.eps}
+        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
+Here we show the relationships between persistence and the proportions of the three positions in the three-species chain motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
+Line style indicates the motif  position.
+}
+        \label{fig:chain_positions}
+    \end{figure}
+
 \clearpage

### 2021-06-29 manuscript/motifs_vs_degree.tex
Updates from Overleaf

@@
 
-        Network average persistence of consumers naturally decreases with increasing level of disturbance of basal resources (Table S1, \emph{Appendix SA}).
-Network size and connectance also affect persistence, but to a much smaller extent (Figure S1, \emph{Appendix SA}).
-Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance: at low disturbance, small and highly-connected webs show the lowest persistence, whereas at high disturbance, large and highly connected webs show the lowest persistence (Fig.
+         Network size and connectance affected persistence, but to a much smaller extent than increasing disturbance to basal resources (Figure S1, Table S1, \emph{Appendix SA}).
+Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance.
+At low disturbance, small and highly-connected webs show the lowest persistence, whereas at high disturbance, large and highly connected webs show the lowest persistence (Fig.
 S2, \emph{Appendix SA}).

### 2021-06-29 manuscript/motifs_vs_degree.tex
Updates from Overleaf

@@
 Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportions of the other three motifs.
-This was true for all levels of disturbance to basal resources; the interactions between motif profiles and disturbance were significant but small.
+Interactions between motif profiles and disturbance were significant but small.
+Here we show the relationships of probabilities of extinction of basal resources of 0.1 (top set of lines) and 0.5 (bottom set of lines).
 Relationships are shown for the observed range of proportions for each motif.}

### 2021-06-29 code/plot_roles_vs_TL.py
got some better colours for TL

@@ -114,10 +114,19 @@ def format_graph(graph,simple,roletype):
   return graph
 
 def populate_graph(graph,minidict,simple,roletype):
-  j=19
   for motif in ['Apparent','Chain','Omnivory','Direct']:
-    if j==23:
-      j=1
+    if motif=='Apparent':
+      j=13
+      sty=2
+    elif motif=='Chain':
+      j=14
+      sty=3
+    elif motif=='Direct':
+      j=16
+      sty=5
+    else:
+      j=19
+      sty=1
     dats=[]
     lower=[]
     upper=[]
@@ -128,21 +137,18 @@ def populate_graph(graph,minidict,simple,roletype):
 
     upply=graph.add_dataset(upper)
     upply.symbol.shape=0
-    upply.line.configure(linestyle=1,linewidth=1,color=j)
-    # upply.fill.configure(pattern=4,type=2,color=j-1,)
+    upply.line.configure(linestyle=sty,linewidth=1,color=j)
 
     lowly=graph.add_dataset(lower)
     lowly.symbol.shape=0
-    lowly.line.configure(linestyle=1,linewidth=1,color=j)
-    # lowly.fill.configure(type=2,pattern=0,color=0)
+    lowly.line.configure(linestyle=sty,linewidth=1,color=j)
 
     data=graph.add_dataset(dats)
     data.symbol.shape=0
-    data.line.configure(linestyle=1,linewdith=1.5,color=j)
+    data.line.configure(linestyle=sty,linewdith=1.5,color=j)
 
     if simple=='TL':
       data.legend=motif
-    j+=1
 
   if simple=='TL':
     graph.add_drawing_object(DrawText,text='Motif',x=5.25,y=0.47,loctype='world',char_size=.75) 

### 2021-07-08 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 
+    \subsection*{Methods}
+
+        As one more level of detail, we can consider the frequency with which a species appears in each \emph{position} in each motif. 
+        Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.
+        The full set of these positions, across all motifs, is the species' \emph{motif role}~\citep{Stouffer2010,Cirtwill2017}.
+        As with motif participation, we normalized motif roles by dividing the number of times a species appeared in each position by the total number of times it appeared in all positions~\citep{Baker2015,Cirtwill2017}.
+
+        [[Could add a PERMANOVA to show that species' roles overal affect their persistence. Haven't done so yet.]]
+
+        To test whether participating in different motif positions was related to a species' probability of persistence, we fit similar models to those for motif participation.
+        For each position, we fit a linear mixed-effect model relating persistence to the frequency of the position in a species' motif role, disturbance, and their interaction.
+        To account for differences in persistence across networks of different size and connectance, we included a random effect of the interaction between size and connectance.
+        All regressions were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+
+    \subsection*{Results}
+
     \begin{table}[h!]
@@
 
+    \subsection*{Methods}
+
+        As with network motif profiles and motif participation, species' motif roles may vary with network size and connectance and with the species' in degree and trophic level.
+        To identify these relationships, we fit two linear regressions for each motif position.
+        The first related the frequency of the position in species' roles to network size, connectance, and their interaction.
+        The second related the frequency of the position to in-degree, STL, and their interaction.
+        All regressions were fit using the R~\citep{R} base function `lm'.
+
+    \subsection*{Results}
         \subsection*{Degree and trophic level}

### 2021-07-20 code/stat_analysis/persistence_vs_SC_lm.py
starting to explore effects of prop basal

@@ -29,6 +29,7 @@ scales={
   'Size':(79.1882,16.62225),
   'Connectance':(0.1044983,0.05559927),
   'Disturbance':(0.3,0.1366261),
+  'Basal':(0.1601632,0.07986507)
 }
 
 lms={
@@ -40,6 +41,25 @@ lms={
   'Size:Disturbance':-3.485e-03,
   'Connectance:Disturbance':-6.998e-05,
   'Size:Connectance:Disturbance':-2.693e-04 }
+
+Blms={'Intercept':5.777e-01,
+      'Basal':2.841e-02,
+      'Size':5.309e-03,
+      'Connectance':6.482e-03,
+      'Disturbance':-1.452e-01,
+      'Basal:Size':1.704e-03,
+      'Basal:Connectance':6.680e-03,
+      'Size:Connectance':2.302e-03,
+      'Basal:Disturbance':-1.116e-02,
+      'Size:Disturbance':-6.699e-03,
+      'Connectance:Disturbance':-6.235e-03,
+      'Basal:Size:Connectance':6.934e-04,
+      'Basal:Size:Disturbance':-1.587e-03,
+      'Basal:Connectance:Disturbance':-7.291e-03,
+      'Size:Connectance:Disturbance':-2.105e-03,
+      'Basal:Size:Connectance:Disturbance':-1.445e-03
+}
+
 
 def format_graph(graph,simple):
   graph.yaxis.bar.linewidth=1
@@ -58,12 +78,18 @@ def format_graph(graph,simple):
     graph.world.xmax=105
     graph.xaxis.tick.major=10
     graph.xaxis.ticklabel.configure(format='decimal',prec=0,char_size=.75)
-  else:
+  elif simple=='Connectance':
     graph.xaxis.label.text='Connectance'
     graph.world.xmin=0
     graph.world.xmax=.2
     graph.xaxis.tick.major=.05
     graph.xaxis.ticklabel.configure(format='decimal',prec=2,char_size=.75)
+  else:
+    graph.xaxis.label.text='Proportion basal'
+    graph.world.xmin=0
+    graph.world.xmax=.5
+    graph.xaxis.tick.major=.1
+    graph.xaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
 
   graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
   graph.xaxis.label.configure(char_size=1,just=2,place='normal')
@@ -71,6 +97,18 @@ def format_graph(graph,simple):
[…]

### 2021-07-20 code/stat_analysis/prop_basal_models.R

@@ -1,0 +1,113 @@
+load.image('all_tests.Rdata')
+
+# Libraries to fit the non-plotted lms
+library(lme4)
+library(lmerTest)
+# For permanovas and betadisper
+library(vegan)
+# For R2 GLMM
+library(MuMIn)
+
+# data is full data, including basals
+data$network_ID=paste(data$Size,data$Connectance,data$Network,sep='_')
+for(r in 1:nrow(data)){
+    netdat=data[which(data$network_ID==data$network_ID[r]),]
+    nBasal=length(which(netdat$in_Degree==0))
+    n=data$Size[r]
+    data$prop_Basal[r]=nBasal/n
+}
+
+# consumers is heterotrophs only
+
+
+# A: Does persistence vary with proportion of basal resources?
+SCper=with(consumers,lm(Persistence~scale(prop_Basal)*scale(Size)*scale(Connectance)*scale(Disturbance)))
+
+
+# B: How do motif profiles vary with prop_Basal and C?
+    motdata=read.table('../../data/3sp_motif_profiles.tsv',header=TRUE)
+    motdata$mot_total=rowSums(motdata[,4:7])
+    motdata$prop_chain=motdata$m12/motdata$mot_total
+    motdata$prop_apparent=motdata$m6/motdata$mot_total
+    motdata$prop_direct=motdata$m36/motdata$mot_total
+    motdata$prop_omni=motdata$m38/motdata$mot_total
+
+    # Expanding motif profiles for each level of disturbance
+    motdata$netID=paste(motdata$Size,motdata$Connectance,motdata$Network,sep=':')
+    consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep=':')
+    consumers$net_chain=0
+    consumers$net_apparent=0
+    consumers$net_direct=0
+    consumers$net_omni=0
+    for(r in 1:nrow(consumers)){
+        net=consumers$netID[r]
+        consumers$net_chain[r]=motdata$prop_chain[which(motdata$netID==net)]
+        consumers$net_apparent[r]=motdata$prop_apparent[which(motdata$netID==net)]
+        consumers$net_direct[r]=motdata$prop_direct[which(motdata$netID==net)]
+        consumers$net_omni[r]=motdata$prop_omni[which(motdata$netID==net)]
+    }
+
+    save.image('analysis.Rdata')
+
[…]

### 2021-07-20 code/stat_analysis/statistical_models.R

@@ -69,7 +69,7 @@ consumers$prop_omni=consumers$m38/consumers$total_motifs
     # PERMANOVA and variability of motif profiles
     propdist=vegdist(motdata[,10:13],method="bray")
     propperm=adonis(propdist~motdata$Size*motdata$Connectance,perm=9999)
-    motdata$interact=paste0(motdata$Size,motdata$Connectance,sep=':')
+    motdata$interact=paste0(motdata$Size,':',motdata$Connectance)
     propinterdisp=betadisper(propdist,motdata$interact,type="centroid")
     Spropdisp=betadisper(propdist,motdata$Size,type="centroid")
     Cpropdisp=betadisper(propdist,motdata$Connectance,type="centroid")
@@ -125,7 +125,7 @@ consumers$prop_omni=consumers$m38/consumers$total_motifs
 
 # Section 4: How does persistence vary with proportions?
     # lmer with random intercepts for each level of S:C
-    consumers$Global=paste0(consumers$Size,consumers$Connectance,':')
+    consumers$Global=paste0(consumers$Size,':',consumers$Connectance)
     propchain_lmer1<-with(consumers,lmer(Persistence~scale(prop_chain)*scale(Disturbance)+(1|Global)))
     propapparent_lmer1<-with(consumers,lmer(Persistence~scale(prop_apparent)*scale(Disturbance)+(1|Global)))
     propdirect_lmer1<-with(consumers,lmer(Persistence~scale(prop_direct)*scale(Disturbance)+(1|Global)))

### 2021-07-20 manuscript/temp_motif_persistence.tex

--- 2021-06-29 manuscript/motifs_vs_degree.tex
@@
             A further complication when trying to understand whether relationships between motifs and probability of extinction depend upon underlying relationships between degree or trophic level and extinction
-            We also fit linear mixed effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+            We also fit linear mixed effect models relating the proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
 
@@
 
-    \subsection*{Persistence did not vary strongly with global network structure}
+We are interested in i) how a the distribution of motifs in a network affect network persistence, and ii) how a species motif profile affects its persistence.
+We analyse this for different levels of disturbance to the basal level.
 
-         Network size and connectance affected persistence, but to a much smaller extent than increasing disturbance to basal resources (Figure S1, Table S1, \emph{Appendix SA}).
+The results are organized as follows: 1) How network persistence varies with network motif distribution.
+2) How network persistence varies with global structural properties.
+3) How network motif distribution varies with global structural properties.
+4) How species probability of persistence varies with its motif profile.
+5) How species persistence varies with its characteristics such as TL and in-degree.
+6) How a species motif profile depend on its characteristics.
+7) How a species persistence  varies with its motif role (position in respective motif).
+
+    \subsection*{Network persistence and level of disturbance and global network properties}
+    The level of disturbance naturally had clear and profound effects on the average network persistence, such as a increased probability of extinction of basal species always decreased network persistence.
+
+    Network size and connectance affect persistence, but to a much smaller extent than increasing disturbance (Figure S1, Table S1, \emph{Appendix SA}).
 Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance.
@@
 
-    \subsection*{Network motif profiles varied with global network properties}
+    \subsection*{Network motif profile and global network properties}
 
-        A PERMANOVA test suggests that a network's normalized motif profile as a whole (i.e., proportions of the total made up by each motif) is related to the network's size, connectance, and the interaction between them, but unequal variation in motif profiles across global network properties may, however, mean that these PERMANOVA results are false positives (Fig.
-S2, \emph{Appendix SB}).
         Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}).
 Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).
-The proportion of direct competition, however, decreased with increasing connectance in small networks but showed very little change in large webs.
 
@@
 
-    \subsection*{Persistence decreases with greater omnivory in a network}
+    \subsection*{Network motif profile and network persistence}
 
-        The motif profile of a network was related to the mean persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
-Mean persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
-Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif.
-These trends hold for all levels of disturbance (Table S3, \emph{Appendix SC}).
-Dispersion of motif profiles varies between networks with different levels of persistence ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
+        The motif profile of a network is related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
+Average persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
+Conversely, average persistence is lower in networks with higher proportions of the omnivory motif.
+These trends hold for all levels of disturbance (Fig.~\ref{motif_profile_persistence}, Table S3), \emph{Appendix SC}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
 For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).
@@
 
-    \subsection*{Species persistence varies with motif participation}
+    \subsection*{Species motif participation and species persistence}
 
-        % See \emph{Appendix SY} for relationships between persistence and non-normalized motif participation.
+       The effects of motif participation on a species persistence depend on level of disturbance.
+When disturbance is low species' persistence increase with increasing proportions of the omnivory and direct competition motifs.
+Conversely, a species' persistence decrease with increasing proportion of the apparent competition motif.
+Persistence is not significantly related to the proportion of the three-species chain motif (Figure~\ref{fig:prop_lmer_all}).
 
-        \subsubsection*{Effects of motif participation depend on disturbance}
-
-            At low levels of disturbance, species' persistences increased with increasing proportions of the omnivory and direct competition motifs.
-Conversely, persistence decreased with as the  proportion of the apparent competition motif increased.
-Persistence was not significantly related to the proportion of the three-species chain (Figure~\ref{fig:prop_lmer_all}).
-
-            At high levels of disturbance, persistence decreased as the proportion of the three-species chain increased, but increased as the proportion of all other motifs increased (Figure~\ref{fig:prop_lmer_all}).
-            % This means that increased disturbance strengthened the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table S4, \emph{Appendix SD}).
-            The effect of omnivory on persistence was  positive but weaker at higher levels of disturbance, and the effect of direct competition was largely unaffected by disturbance level.
+        At high levels of disturbance, species persistence decrease as the proportion of the three-species chain increase, but increase as the proportion of all other motifs increase (Figure~\ref{fig:prop_lmer_all}).
+This means that increased disturbance strengthened the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table S4, \emph{Appendix SD}).
+The effect of omnivory on persistence is still positive but weaker at higher levels of disturbance, and the effect of direct competition was largely unaffected by disturbance level.
 
@@
 
-        \subsubsection*{Effects of motif participation are more consistent in highly-connected networks}
+        \subsubsection*{Consistency in species persistence versus motif participation}
 
-            By fitting linear regressions of persistence against motif proportions for the species for each network separately, we can observe how the relationships between motifs and persistence change with the level of disturbance and global network structure.
+            In order to evaluate how consistent the relationships between a species persistence and motif participation is, and how it varies with  level of disturbance and global network structure, we fit linear regressions of species persistence against motif proportions for the species for each network separately.
+We then analyze the density distribution of the values of the slopes.
 Network size had a much smaller effect than connectance (Fig.
 S3; \emph{Appendix SE}); we therefore focus on connectance here.
+
             Networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
-            However, there was an interaction between disturbance and connectance for the direct competition and omnivory motifs.
-
-            At low and medium connectances with a high proportion of the direct competition motif, there is little change in the proportion of positive slopes (\textgreater80\%) for the different levels of disturbance.
-            At high connectances, however, the proportion of positive slopes is lower overall and decreases with increasing disturbance. 
-            This means that at low levels of connectance, an increased proportion of the direct competition motif increases persistence in most networks, while the opposite is true with high connectance.
-            The trend is reversed for the omnivory motif.
+However, there is an interaction between disturbance and connectance for the direct competition and omnivory motifs; At low and medium connectances there is little change in the proportion of positive slopes when the proportion of  direct competition motif is high, independent of level of disturbance.
+At high connectances, however, the proportion of positive slopes is lower and also decreases with increasing disturbance.
+For the omnivory motif the trend is reversed.
+Additionally, the omnivory motif stands out having a wider curve when network connectance is low.
 
@@
 
-    \subsection*{Persistence increases with degree but decreases with STL}
+    \subsection*{Species characteristics and persistence}
 
-        Persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
-Increased trophic level is associated with decreasing persistence, although the rate of this decline is slower at higher levels of disturbance (Table S5, \emph{Appendix SF}).
-        The interaction between in-degree and disturbance is even stronger, such as having a large number of prey is associated with higher persistence in the no disturbance scenario, but lower persistence at high disturbance.
+        A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
+Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S5, \emph{Appendix SF}).
+        In-degree shows an even stronger interaction with disturbance, as a having more  prey species is associated with higher persistence in the no disturbance scenario, but lower persistence at high disturbance scenario.
 
@@
 
-    \subsection*{Simpler measures do not fully explain effects of motifs}
+    \subsection*{Species motif participation and network properties and species characteristics}
 
-        Motif participation varied with web size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
-        Species in all networks appeared most frequently in the apparent competition motif, but the frequency of the apparent competition motif decreased with increasing network size and connectance.
-        The frequency of the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size.
+       Species' motif roles varied depending on the properties of the network; size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
+        Species appeared most often in the apparent competition motif in all types of networks, but the frequency decreased with increasing network size and connectance.
+        The frequency with which species participated in the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size.
 
-        Species' motif roles also vary with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
+        Species' motif roles also varied with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
         Species with higher in-degree tend to have higher proportions of their motif role made up by omnivory. Proportions of the other three motifs tend to decline with increasing degree.
@@
 
-    \subsection*{Persistence varies with a species' motif role}
+    \subsection*{Persistence and species' positioning in motifs}
 
@@
 % Results summary - network level (following Anna)
-Changes in the primary producer level in food webs have profound effects on all other species in the network \citep{}.
-In particular, the disturbances can percolate across trophic levels as all consumers depend directly or indirectly on primary producers for their survival \citep{}.
-Understanding which characteristics of a network and its species influence the severity of outcomes after a disturbance is crucial, e.g.
-for conservation actions \citep{}.
-Here we were particularly interested in how the distribution of meso-scale structures in ecological networks are related to average network persistence when the basal level is disturbed by species loss.
-In order to disentangle the effects caused by the motif distribution we also analyzed global (network-wide) and local (species) properties.
-%Consistent with earlier work, networks with higher connectance or species richness tend to have lower persistences~\citep{}.
-A network's motif profile had relatively strong relationships to the network mean persistence, while effects from global scale structures, such as network size and connectance, were relatively small. 
-However, connectance (and, to a lesser extent network size) do affect network motif profiles.
-Thus, global-scale properties seem to have stronger effects on network persistence by changing meso-scale properties than directly.
-~\citep{prill2005dynamic, bascompte2005simple}.
+Changes in the basal level in food webs have profound effects on all other species in the network \citep{}, as all consumers depend directly or indirectly on primary producers for their survival \citep{}.It has repeatedly been shown that global network properties are important for how well a network can handle disturbances, such as loss of species \citep{Eklof2006, Dunne2002}.
+At the same time, from a conservation perspective it is often more relevant to look at specific species to which actions can be targeted \cite{}.
+Nevertheless, since species do exist in a network their interactions, direct as well as indirect, will affect how they respond to a disturbance.
+Motifs are the basic building blocks in networks and as such represents a bridge between network-wide (global) and species-specific (local) properties.
+Here we were particularly interested in how the distribution of these meso-scale structures in ecological networks are related to network and species persistence when the plant community is affected by species loss.
 
-% Network-level motif profiles, persistence, and relationships to S and C.
-A network's mean persistence depended on which motifs were most common in the network profile.
-In general, persistence was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif.
-Direct competition and three-species chain motifs had smaller effects.
-The effects were clearly pronounced when disturbance was low but attenuated with increased disturbance.
-The negative relationship between omnivory and mean persistence is consistent with other studies showing that an omnivory motif is less likely to be stable than the other motifs considered~\citep{Borrelli2015a}.
+A network's motif profile had relatively strong relationships to the network average persistence, compared to global structures, such as network size and connectance, whose effects were smaller.
+However, connectance (and, to a lesser extent network size) do affect network which motifs that are most prominent.
+Thus, global-scale properties have stronger effects on network persistence by changing meso-scale properties than directly which indicates that network motifs do provide additional information.
 
-High frequency of the apparent competition motif was correlated with an on average higher in-degree for consumers. In-degree in turn is often related to higher persistence \citep{}.
+% Network-level persistence
+ In general, average network persistence was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif.
+Direct competition and three-species chain motifs both showed positive but weaker relationships.
+The level of disturbance naturally affected the level of persistence, such as increased disturbance decreased average persistence, but trends were largely unaffected.
+
+% About fraction of omnivory and apparent competition in the full network 
+The negative relationship between omnivory and network persistence is consistent with other studies showing that an omnivory motif is less likely to be stable than the other motifs considered~\citep{Borrelli2015a}.
+
+High frequency of the apparent competition in a network motif was correlated with an on average higher in-degree for consumers.
+In-degree has repeatedly been suggested to be beneficial for a species persistence  \citep{}.
 
@@
 
-% HERE ENDS MEAN PERSISTENCE AND SPECIES WISE START
-As well as the mean persistence of all consumers, we also considered the species-specific persistence of each consumer in a network.
-While the persistence of a particular species is likely to be affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it also depends on a species' position in the network.
-There are many ways of measuring species' positions, and in some cases multiple measures may capture the same trend.
-In our results, motif participation is significantly related to persistence, but the effects of two motifs (\emph{apparent competition} and \emph{direct competition}) might be explained more simply by in-degree and trophic level.
-This is especially likely for these two motifs, since Bayesian Network modeling lacks top-down effects and therefore removes the indirect effects from these motifs.
+% Species level persistence
+% What we want to discuss here: 1) Apparent competition change direction depending on disturbance.
+2) High disturbance strengthen the effect of chain.
+3) Omnivory most positive in low disturbance scenarios.
 
-% Getting rid of AC/DC "quickly" here so we can focus on omnivory
-Participation in the apparent competition motif is strongly negatively correlated with in-degree---a high proportion of the apparent competition motif implies a low in-degree. 
-Both higher in-degree and lower participation in the apparent competition motif are associated with greater persistence at low disturbance but decreased persistence at high disturbance; because the trends are so similar, it is likely that a common mechanism underlies both.
-% The effect on persistence is consistent between species with a low in-degree and a high proportion of the apparent competition motif. 
-%Low in-degree decreases persistence at low disturbance, and increases persistence at high disturbance, similar to the trend visible for a high participation in the apparent competition motif. 
-% Similarly, the effect on species persistence is reversed for a low proportion of the apparent competition motif in which species have a high in-degree.
+While the persistence of a particular species is likely to be affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it also depends on what characterize a species position in the network., such as trophic level, number of interactions and which motifs the species participate in.
 
-While the direct competition motif is not correlated with in-degree, it is strongly negatively correlated with trophic level---species with a high proportion of the direct competition motif have a low trophic levels.
-Low trophic levels are always associated with higher persistence because these species are less affected by bottom-up disturbances~\citep{}.
-It then follows that species which participate in many direct competition motifs may have higher probabilities of persistence because they are also likely to have low trophic levels.
+In our results, motif participation is significantly related to persistence, but some of the effects can to some extent be explained by local properties such as in-degree and trophic level.
+This is especially likely for the apparent competition and direct competition motifs, since in Bayesian Network framework lacks top-down effects and therefore the indirect effects in these motifs are not relevant.
 
-In contrast to the competition motifs, the three-species chain and omnivory motifs do have indirect effects in a Bayesian Network setting and the effects of these motifs cannot be fully explained by degree or trophic level.
+Participation in the apparent competition motif is strongly negatively correlated with in-degree --a high proportion of the apparent competition motif implies a low in-degree.
+Both higher in-degree and lower participation in the apparent competition motif are associated with greater persistence at low disturbance but decreased persistence at high disturbance.
+This suggests that in-degree is in fact the factor driving the positive relationship between proportion of apparent competition motif and species persistence.
+Interestingly though, the relationship changes from positive to negative with increasing level of disturbance.
+When disturbance is high having many prey species acts as an insurance, while at high disturbance connections to many prey species is in fact negative as the consumer receives the negative impact from several sources.
+
+However, the coupling between a species proportion of motif participation and degree is neither evident or always true.
+For example, it is possible for a species with low degree to participate in many motifs if it interacts with high-degree species and thereby obtain many indirect interaction partners.
+This is one of the strengths with analyses of motifs that it considers local structures additionally include indirect connections one link away.
+
+A species participation in the direct competition motif is only weakly correlated with in-degree, but is instead strongly negatively correlated with trophic level---species with a high proportion of the direct competition motif are more likely to have low trophic levels.
+Low trophic levels are naturally associated with higher persistence in our analyzes  as they receive less negative impact from the bottom-up disturbances ~\citep{Eklof2013}.
+The clear positive relationship between proportion of direct competition motifs and species persistence is therefore likely due to the tendency for species at low trophic levels to have a higher proportion of that motif.
+Therefore, contrary to the apparent comptition motif, we do not observe a change in the relationship with changed disturbance.
+
+The three species chain motif have a nearly neutral relationship with persistence when disturbance is low, but a clearly negative relationship when disturbance is high.
+This can partly be explained with the positive relationship between trophic level and a species proportion on the chain motif.
+In the three species chain motif the species are on average more often dependent on two trophic levels (if the focal species possesses the top-position in the motif), and as such affected negatively from two trophic levels.
+A high level of disturbance (high probability for a basal species to go extinct) will then sharply decrease persistence for the top species.
+
+%In contrast to the two competition motifs, the three-species chain and omnivory motifs have a less evident relationship to species in-degree or trophic level.
 For example, the three-species chain motif is negatively correlated with in-degree and positively correlated with trophic level.
-Despite this correlation, participation in the three-species chain does not interact the same way with disturbance as either simpler measure and so has a unique relationship to species persistence.
+Despite this correlation, participation in the three-species chain does not interact the same way with disturbance as either of the simpler measure and so has a unique relationship to species persistence.
 This is likely due to indirect effects of the bottom species in the chain on the top species.
 
-The most interesting motif in which to consider indirect effects is omnivory.
+The perhaps most interesting motif is omnivory.
+Omnivory shows a reversed relationship with disturbance level compared to the three species chain; at low disturbance a high proportion of omnivory is positive for persistence whereas the effect is nearly neural for high levels of disturbance.
+
 While participation in the omnivory motif is positively correlated with in-degree, a high proportion of the omnivory motif is always associated with increased persistence, while the effect of in-degree varies with disturbance.
@@
 
-% % Might be able to add some of this paragraph to the one above or below (especially the refs).
 % The divergent effect on persistence of either having a high proportion of the three-species chain or omnivory motif can, within our Bayesian Network framework, be explained by the structure of the food web.
@@
 However, the study by Cirtwill and Wootton used a different approach using a dynamic simulation model also including top-down effects.
-    %In previous studies of the effect of omnivory motifs, the strength of the omnivorous link can potentially be important.
-If this link is strong, it is shown to be destabilizing (REF), while stabilizing if weaker (REF).
-In our Bayesian Network setting, this link is indistinguishable from any other link in the food web which partly may explain why omnivory results in increased persistence of consumer species.
 
@@
     % Species with higher degrees will with high probability appear in more motifs than species with low degrees.
-    % It is possible for a species with low degree (few direct interaction partners) to participate in many motifs if it interacts with high-degree species (thereby obtaining many indirect interaction partners), but such a situation is fairly unlikely.
-    % Because of this overlap between degree and absolute numbers of motifs, and because degree shows a similar trend to the motif counts [[to be confirmed]], we recommend that researchers consider degree and/or normalized motifs rather than absolute motifs.
+    % It is possible for a species with low degree  to participate in many motifs if it interacts with high-degree species (thereby obtaining many indirect interaction partners).
+
     % When we consider the proportion of a species' role made up by each motif, the influence of degree is largely removed.

### 2021-07-20 manuscript/motifs_vs_degree.tex

@@
 
-% Maybe Proc B - approached Alyssa re: other motifs paper, flexible length limit.
+% Maybe Proc B - approached Alyssa re: other motifs paper, flexible length limit (but pay extra over X pages).
 
@@
 
-    % In the Bayesian network framework, only the stable motifs are included.
-    % We show that even in this case participation in different motifs is associated with different extinction risk.
-    % Participation in different motifs is also correlated with in-degree and trophic level.
-    % In a Bayesian network, trophic level has a very large effect on extinction risk, and likely explains a lot of what we observe for degree.
-    % However, the trends we observe for motifs are not totally explained by the correlations of motifs with trophic level and degree, indicating that both simple and meso-scale descriptions of species roles provide unique information about extinction risk.
+    - can we condense omnivory, in-degree any further?
+    - - at least omnivory
+    - can we shorten any more? Nice to save some money?
+    - glossary
+    - refs
+    - ending
 
-    % Omnivory may be stabilizing at high connectances because there is a low-TL "rescue prey" available to a high-risk, high-TL species.
-At lower connectances, perhaps higher-TL species encounter fewer disturbances and the rescue prey is less important?
-    % In Lotka-Volterra systems, where top-down effects can also cause extinction, omnivory tends to be associated with less persistence.
-This suggests that what happens in real systems could depend on the dominant type of regulation.
+    - repeat network size analyses with \# basal resources - maybe adjust Fig. 1 to reflect n(basal) and C instead of C:S?
+    - Fig. 4 not much in discussion, could be a candidate for the SI? Or add more discussion about apparent competition, etc. in discussion sections per motif
 
-    % - Add a panel showing in-degree at different TL's, like in Anna's fig. Or switch to persistence vs. in-deg for each TL in a panel.
 \clearpage
@@
      The plant community is the foundation upon which the myriad of species in food webs depend for their survival \citep{}.
-Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{} and secondary production \citep{}, soil respiration and carbon cycling \citep{} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
+Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
 The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community promotes stability across all different levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}.
@@
     The global structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002d, Eklof2006}.
-For example, more highly connected networks are more resistant to secondary extinctions \citep{Dunne2002d, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{}.
+For example, more highly-connected networks are more resistant to secondary extinctions \citep{Dunne2002d, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{}.
     %\subsection*{4 meso-scale nw propoerties also affect stability}
@@
     For example, specialists may not necessarily have a high risk of extinction if it consumes a basal resource which is at low risk of being affected of disturbance.
-On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes preys which possess high risk of being affected by disturbance.
+On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes preys which themselves possess high risk of being affected by disturbance.
 
@@
 Importantly, these motif roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions.
-    This means that two species with the same in-degrees and trophic levels can still have quite different motif roles if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
-This extended description of species structural roles may give additional information of extinction risk.
-    At the network level, the frequencies of different motifs is associated with community stability \citep{prill2005dynamic, bascompte2005simple}.
-    Some motifs, such as the three-species chain, are likely to retain all species within them when modelled in isolation~\citep{Borrelli2015a} and over-represented in empirical food webs~\citep{Stouffer2007}.
-    This suggests that some motifs may increase network stability by reducing the probability that species participating in them will go extinct.
-    The plant community has been shown to be particularly relevant for how the motifs are distributed within a network.
-For example, a recent study in experimental grasslands showed that changes in the plant community, i.e.
-richness, induced shifts in the relative frequency of motifs in the food webs studied \cite{giling2019plant}.
+    This means that two species with the same in-degrees and trophic levels can still participate in different sets of motifs if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
+    This extended description of species structural roles may give additional information of extinction risk or synthesize the information provided by multiple other measures of a species' place in its community.
+    At the network level, the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities~\cite{giling2019plant}.
+    We therefore expect that the motif structure of a network and species' participation in that structure should be related to risks due to disturbance to basal resources.
 
-    %\subsection{6. only 4 (stable) motifs can appear in Bayesian networks, so how do we expect them to affect stability in a bottom-up context?}
-    Of the 13 three-species motifs that can occur in empirical food webs there are four motifs that are over-represented in empirical networks \citep{Stouffer2007} and contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}: the three-species chain, omnivory, apparent competition and direct competition motifs.
-    These four motifs are also the only ones which can occur in the acyclic networks required for Bayesian network simulations~\citep{Eklof2013}.
-Because of this restriction, we only consider these four particularly important motifs.
+    %\subsection{6. only 4 (stable) motifs can appear in Bayesian networks}
+    Of the 13 three-species motifs that can occur in empirical food webs there are four motifs that are over-represented in empirical networks and are thought to contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}: the three-species chain, omnivory, apparent competition and direct competition motifs.
+    These four motifs are also the ones which can occur in the acyclic networks required for Bayesian network simulations~\citep{Eklof2013}.
+Therefore, we focus on these four particularly important motifs.
 
@@
     Here we are interested in how disturbances at the basal level in food webs propagate up the network and whether a consumer's extinction risk depends on 1) the motif profile of the full food web and 2) the motifs in which a focal species participates.
-In particular, we want to understand whether the information provided by motifs can be explained by the classical global and local food web properties such as network connectance and species trophic level, or whether motifs provide unique insight into bottom-up extinction cascades.
+In particular, we want to understand whether the information provided by motifs is related to classical global and local food web properties such as network connectance and species trophic level, or whether motifs provide unique insight into bottom-up extinction cascades.
 
@@
 
-            We are interested in whether having a greater proportion of a species' role be made up of a certain motif (e.g., direct competition) might increase persistence.
+            We are interested in whether participating more often in a certain motif (e.g., direct competition) might increase persistence.
             % (see \emph{Appendix SXX} for relationships between persistence and counts of motifs).
@@
             We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance. 
+            We performed similar analyses to test whether appearing in different positions in each motif (i.e., the motif role) is related to a species' persistence (\emph{Appendix SH}).
 
@@
 
-            Species' motif roles are often related to their degrees and trophic levels.
-            To understand how this may have affected the results above, we first fit two linear mixed effects regressions relating persistence to degree or trophic level, $basal_p$, and their interaction.
+            A species' motif participation is likely related to its degrees and trophic level since all three are measures of network position.
+            To understand how this may have affected the results above, we first fit two linear mixed-effects regressions relating persistence to degree or trophic level, $\pi_{disturbed}$, and their interaction.
             As above, we centered and scaled the predictors.
@@
             A further complication when trying to understand whether relationships between motifs and probability of extinction depend upon underlying relationships between degree or trophic level and extinction
-            We also fit linear mixed effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+            We also fit linear mixed-effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+            We performed similar analyses to test whether a species' motif role is related it in-degree or trophic level (\emph{Appendix SI}).
 
-        \subsubsection*{How does persistence vary with motif positions?}
+        % \subsubsection*{How does persistence vary with motif positions?}
 
-            As one more level of detail, we can consider the frequency with which a species appears in each \emph{position} in each motif. 
-            Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.
-            The full set of these positions, across all motifs, is the species' \emph{motif role}~\citep{Stouffer2010,Cirtwill2017}.
-            As with motif participation, we normalized motif roles by dividing the number of times a species appeared in each position by the total number of times it appeared in all positions~\citep{Baker2015,Cirtwill2017}.
+        %     As one more level of detail, we can consider the frequency with which a species appears in each \emph{position} in each motif. 
+        %     Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.
+        %     The full set of these positions, across all motifs, is the species' \emph{motif role}~\citep{Stouffer2010,Cirtwill2017}.
+        %     As with motif participation, we normalized motif roles by dividing the number of times a species appeared in each position by the total number of times it appeared in all positions~\citep{Baker2015,Cirtwill2017}.
 
-            [[Could add a PERMANOVA to show that species' roles overal affect their persistence. Haven't done so yet.]]
+        %     [[Could add a PERMANOVA to show that species' roles overal affect their persistence. Haven't done so yet.]]
 
-            To test whether participating in different motif positions was related to a species' probability of persistence, we fit similar models to those for motif participation.
-            For each position, we fit a linear mixed effect model relating persistence to the frequency of the position in a species' motif role, disturbance, and their interaction.
-            To account for differences in persistence across networks of different size and connectance, we included a random effect of the interaction between size and connectance.
-            All regressions were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+        %     To test whether participating in different motif positions was related to a species' probability of persistence, we fit similar models to those for motif participation.
+        %     For each position, we fit a linear mixed-effect model relating persistence to the frequency of the position in a species' motif role, disturbance, and their interaction.
+        %     To account for differences in persistence across networks of different size and connectance, we included a random effect of the interaction between size and connectance.
+        %     All regressions were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
 
-        \subsubsection*{How do motif roles vary with other network properties?}
+        % \subsubsection*{How do motif roles vary with other network properties?}
 
-            As with network motif profiles and motif participation, species' motif roles may vary with network size and connectance and with the species' in degree and trophic level.
-            To identify these relationships, we fit two linear regressions for each motif position.
-            The first related the frequency of the position in species' roles to network size, connectance, and their interaction.
-            The second related the frequency of the position to in-degree, STL, and their interaction.
-            All regressions were fit using the R~\citep{R} base function `lm'.
+        %     As with network motif profiles and motif participation, species' motif roles may vary with network size and connectance and with the species' in degree and trophic level.
+        %     To identify these relationships, we fit two linear regressions for each motif position.
+        %     The first related the frequency of the position in species' roles to network size, connectance, and their interaction.
+        %     The second related the frequency of the position to in-degree, STL, and their interaction.
+        %     All regressions were fit using the R~\citep{R} base function `lm'.
 
@@
 
-    \subsection*{Persistence did not vary strongly with global network structure}
+    We are interested in i) how the distribution of motifs in a network (i.e., the network's \emph{motif profile}) affect network persistence, and ii) how a species motif participation affects its persistence.
+We analyse this for different levels of disturbance to the basal level.
+To provide context, we also investigate relationships between persistence and classic global (size and connectance) and local (in-degree and trophic level) network properties.
+    % The results are organized as follows: 1) How network persistence varies with network motif distribution.
+2) How network persistence varies with global structural properties.
+3) How network motif distribution varies with global structural properties.
+4) How species probability of persistence varies with its motif profile.
+5) How species persistence varies with its characteristics such as TL and in-degree.
+6) How a species motif profile depend on its characteristics.
+7) How a species persistence varies with its motif role (position in respective motif).
 
-         Network size and connectance affected persistence, but to a much smaller extent than increasing disturbance to basal resources (Figure S1, Table S1, \emph{Appendix SA}).
+    \subsection*{Mean persistence, disturbance, and global network properties}
+
+        The mean probability of persistence for all consumers in a network naturally decreased with increasing probability of extinction of basal species.
+These effects were clear across the ranges of network size and connectance we considered.
+        Network size and connectance also affect network mean persistence, but to a much smaller extent than increasing disturbance (Figure S1, Table S1, \emph{Appendix SA}).
 Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance.
@@
 
-    \subsection*{Network motif profiles varied with global network properties}
+    \subsection*{Network motif profiles and global network properties}
 
-        A PERMANOVA test suggests that a network's normalized motif profile as a whole (i.e., proportions of the total made up by each motif) is related to the network's size, connectance, and the interaction between them, but unequal variation in motif profiles across global network properties may, however, mean that these PERMANOVA results are false positives (Fig.
-S2, \emph{Appendix SB}).
+        Connectance was significantly related to the frequencies of different motifs in a network.
+        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).
         Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}).
-        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).
-The proportion of direct competition, however, decreased with increasing connectance in small networks but showed very little change in large webs.
 
@@
 
-    \subsection*{Persistence decreases with greater omnivory in a network}
+    \subsection*{Network motif profiles and network persistence}
 
-        The motif profile of a network was related to the mean persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
-Mean persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
-Conversely, mean persistence is lower in networks with higher proportions of the omnivory motif.
-These trends hold for all levels of disturbance (Table S3, \emph{Appendix SC}).
-Dispersion of motif profiles varies between networks with different levels of persistence ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
+        The motif profile of a network is related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
+Average persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
+Conversely, average persistence is lower in networks with higher proportions of the omnivory motif.
+These trends hold for all levels of disturbance (Fig.~\ref{motif_profile_persistence}, Table S3), \emph{Appendix SC}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
 For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).
@@
 
-    \subsection*{Species persistence varies with motif participation}
+    \subsection*{Species motif participation and species persistence}
 
-        % See \emph{Appendix SY} for relationships between persistence and non-normalized motif participation.
+       The effects of motif participation on a species' probability of persistence depends on the level of disturbance.
+When disturbance is low, species' persistence increases with increasing proportions of the omnivory and direct competition motifs.
+Conversely, a species' persistence decreases with an increasing proportion of the apparent competition motif.
+Persistence is not significantly related to the proportion of the three-species chain motif (Figure~\ref{fig:prop_lmer_all}).
 
-        \subsubsection*{Effects of motif participation depend on disturbance}
-
-            At low levels of disturbance, species' persistences increased with increasing proportions of the omnivory and direct competition motifs.
-Conversely, persistence decreased with as the  proportion of the apparent competition motif increased.
-Persistence was not significantly related to the proportion of the three-species chain (Figure~\ref{fig:prop_lmer_all}).
-
-            At high levels of disturbance, persistence decreased as the proportion of the three-species chain increased, but increased as the proportion of all other motifs increased (Figure~\ref{fig:prop_lmer_all}).
-            % This means that increased disturbance strengthened the effects of the three-species chain motif and changes direction of the effect from the apparent competition motif  (Table S4, \emph{Appendix SD}).
-            The effect of omnivory on persistence was  positive but weaker at higher levels of disturbance, and the effect of direct competition was largely unaffected by disturbance level.
+        At high levels of disturbance, a species' persistence decreases as the proportion of the three-species chain increases, but persistence increases as the proportion of any other motif increases (Figure~\ref{fig:prop_lmer_all}).
+This means that increased disturbance strengthens the effects of the three-species chain motif and changes the direction of the effect of the apparent competition motif  (Table S4, \emph{Appendix SD}).
+The effect of omnivory on persistence is still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.
 
@@
 
-        \subsubsection*{Effects of motif participation are more consistent in highly-connected networks}
+        \subsubsection*{Consistency in effects of motif participation on species persistence}
 
-            By fitting linear regressions of persistence against motif proportions for the species for each network separately, we can observe how the relationships between motifs and persistence change with the level of disturbance and global network structure.
+            % In order to evaluate how consistent the relationships between a species persistence and motif participation is, and how it varies with  level of disturbance and global network structure, we fit linear regressions of species persistence against motif proportions for the species for each network separately.
+We then analyze the density distribution of the values of the slopes.
 Network size had a much smaller effect than connectance (Fig.
 S3; \emph{Appendix SE}); we therefore focus on connectance here.
-            Networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
-            However, there was an interaction between disturbance and connectance for the direct competition and omnivory motifs.
 
-            At low and medium connectances with a high proportion of the direct competition motif, there is little change in the proportion of positive slopes (\textgreater80\%) for the different levels of disturbance.
-            At high connectances, however, the proportion of positive slopes is lower overall and decreases with increasing disturbance. 
-            This means that at low levels of connectance, an increased proportion of the direct competition motif increases persistence in most networks, while the opposite is true with high connectance.
-            The trend is reversed for the omnivory motif.
+            When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
+However, there is an interaction between disturbance and connectance for the direct competition and omnivory motifs; At low and medium connectances there is little change in the proportion of positive slopes when the proportion of  direct competition motif is high, independent of level of disturbance.
+At high connectances, however, the proportion of positive slopes is lower and also decreases with increasing disturbance.
+For the omnivory motif the trend is reversed.
+Additionally, the omnivory motif stands out having a wider curve when network connectance is low.
+            Network size had a much smaller effect than connectance (Fig. S3; \emph{Appendix SE}).
 
@@
             \caption{Density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
-The slopes are derived from standard linear regression.
 Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
@@
 
-    \subsection*{Persistence increases with degree but decreases with STL}
+    \subsection*{Species characteristics and persistence}
 
-        Persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
-Increased trophic level is associated with decreasing persistence, although the rate of this decline is slower at higher levels of disturbance (Table S5, \emph{Appendix SF}).
-        The interaction between in-degree and disturbance is even stronger, such as having a large number of prey is associated with higher persistence in the no disturbance scenario, but lower persistence at high disturbance.
+        A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
+Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S5, \emph{Appendix SF}).
+        In-degree shows an even stronger interaction with disturbance; having more prey species is associated with higher persistence without disturbance, but lower persistence at high levels of disturbance.
 
@@
 
-    \subsection*{Simpler measures do not fully explain effects of motifs}
+    \subsection*{Motif participation, network properties, and species characteristics}
 
-        Motif participation varied with web size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
-        Species in all networks appeared most frequently in the apparent competition motif, but the frequency of the apparent competition motif decreased with increasing network size and connectance.
-        The frequency of the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size.
+       Species' motif participation varied depending on the properties of the network; size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
+        Species appeared most often in the apparent competition motif in all types of networks, but the frequency decreased with increasing network size and connectance.
+        The frequency with which species participated in the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size.
 
-        Species' motif roles also vary with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
-        Species with higher in-degree tend to have higher proportions of their motif role made up by omnivory. Proportions of the other three motifs tend to decline with increasing degree.
+        Species' motif participation also varied with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
+        Species with higher in-degree tend to have higher frequencies of participation in the omnivory motif. Proportions of the other three motifs tend to decline with increasing degree.
         Species with higher trophic levels tend to have higher proportions of apparent competition and three-species chains.
@@
 
-    \subsection*{Persistence varies with a species' motif role}
+    % \subsection*{Persistence and species' positioning in motifs}
 
-        A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance (Table SX, \emph{Appendix SH}).
-        In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
-        Persistence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
-        Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.
+    %     A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance (Table SX, \emph{Appendix SH}).
+    %     In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
+    %     Persistence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
+    %     Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.
 
-        In the three-species chain, persistence decreased as the frequency of the top position increased at all levels of disturbance.
-        Conversely, persistence increased with increasing frequency of the middle position at all levels of disturbance.
-        Persistence decreased slightly with increasing frequency of the bottom position at low levels of disturbance but increased at high levels of disturbance.
+    %     In the three-species chain, persistence decreased as the frequency of the top position increased at all levels of disturbance.
+    %     Conversely, persistence increased with increasing frequency of the middle position at all levels of disturbance.
+    %     Persistence decreased slightly with increasing frequency of the bottom position at low levels of disturbance but increased at high levels of disturbance.
 
-        In the apparent competition motif, persistence decreased slightly with increasing frequencies of the bottom position (one predator, no prey within the motif) at low disturbance but increased slightly at high disturbance.
-        Persistence increased with increasing frequency of the top position (two prey, no predator within the motif) at low disturbance but decreased at high disturbance.
+    %     In the apparent competition motif, persistence decreased slightly with increasing frequencies of the bottom position (one predator, no prey within the motif) at low disturbance but increased slightly at high disturbance.
+    %     Persistence increased with increasing frequency of the top position (two prey, no predator within the motif) at low disturbance but decreased at high disturbance.
 
-        In the direct competition motif, persistence varied little with the frequency of the bottom position (two predators, no prey within the motif) at low disturbance but increased with the frequency of this position at high levels of disturbance.
-        Persistence increased slightly with the frequency of the top position (one prey, no predators within the motif) at low levels of disturbance and decreased slightly at high levels of disturbance.
+    %     In the direct competition motif, persistence varied little with the frequency of the bottom position (two predators, no prey within the motif) at low disturbance but increased with the frequency of this position at high levels of disturbance.
+    %     Persistence increased slightly with the frequency of the top position (one prey, no predators within the motif) at low levels of disturbance and decreased slightly at high levels of disturbance.
 
-    \begin{figure}[h!]
-        \includegraphics[width=.95\textwidth]{figures/persistence_positions_Omnivory.eps}
-        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
+    % \begin{figure}[h!]
+    %     \includegraphics[width=.95\textwidth]{figures/persistence_positions_Omnivory.eps}
+    %     \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
 Here we show the relationships between persistence and the proportions of the three positions in the omnivory motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
@@
 }
-        \label{figs:persistence_position_motifs}
-        \end{figure}
+    %     \label{figs:persistence_position_motifs}
+    %     \end{figure}
 
-        The frequencies of motif positions in a species' role varied with network properties (\emph{Appendix SI}).
-        Motif roles had stronger relationships with degree and STL than with network size or connectance.
+    %     The frequencies of motif positions in a species' role varied with network properties (\emph{Appendix SI}).Motif roles had stronger relationships with degree and STL than with network size or connectance.
 
@@
 %Apparent comp and direct comp have no indirect effects in BN, so we should be able to explain them with simpler properties.
-%AC is strongly negatively correlated with in-degree.
-Higher in degree is associated with both high persistence at low disturbance (Fig.
-7A) and low participation in AC (Fig.
-7C).
-Low participation in AC is in turn associated with high persistence at low disturbance (Fig.
-4).
-Note that trends for degree and AC reverse at high disturbance in the same way!
-%DC is not strongly correlated with degree but does decrease strongly at high trophic levels.
-High trophic level and low DC are both associated with low persistence at all levels of disturbance, making it likely that the effects we observe for DC are trophic level in disguise.
-%Three-species chain and omnivory do have indirect effects in a BN setting, and they can't be fully explained by degree or trophic level (since the result doesn't follow the result for degree).
-We can see this because despite the strong correlations with degree and trophic level these motifs don't show the same patterns with persistence.
-We therefore focus on them here.
 
 \section*{Discussion}
-% Connect back to intro -- importance of the plant community for a healthy consumer community.
+
 % Focus on what additional information is revealed from analyzing motif compared to global network properties or species characteristics.
@@
 % Results summary - network level (following Anna)
-Changes in the primary producer level in food webs have profound effects on all other species in the network \citep{}.
-In particular, the disturbances can percolate across trophic levels as all consumers depend directly or indirectly on primary producers for their survival \citep{}.
-Understanding which characteristics of a network and its species influence the severity of outcomes after a disturbance is crucial, e.g.
-for conservation actions \citep{}.
-Here we were particularly interested in how the distribution of meso-scale structures in ecological networks are related to average network persistence when the basal level is disturbed by species loss.
-In order to disentangle the effects caused by the motif distribution we also analyzed global (network-wide) and local (species) properties.
-%Consistent with earlier work, networks with higher connectance or species richness tend to have lower persistences~\citep{}.
-A network's motif profile had relatively strong relationships to the network mean persistence, while effects from global scale structures, such as network size and connectance, were relatively small. 
-However, connectance (and, to a lesser extent network size) do affect network motif profiles.
-Thus, global-scale properties seem to have stronger effects on network persistence by changing meso-scale properties than directly.
-~\citep{prill2005dynamic, bascompte2005simple}.
+Changes in the basal level in food webs have profound effects on all other species in the network \citep{}, as all consumers depend directly or indirectly on primary producers for their survival \citep{}.
+It has repeatedly been shown that global network properties are important for how well a network can handle disturbances, such as loss of species \citep{Eklof2006, Dunne2002}.
+At the same time, from a conservation perspective it is often more relevant to focus interventions on particular target species~\citep{}.
+Nevertheless, the process of identifying target species should take the network context into account since both direct and indirect interactions can affect how species respond to a disturbance~\citep{}.
 
-% Network-level motif profiles, persistence, and relationships to S and C.
-A network's mean persistence depended on which motifs were most common in the network profile.
-In general, persistence was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif.
-Direct competition and three-species chain motifs had smaller effects.
-The effects were clearly pronounced when disturbance was low but attenuated with increased disturbance.
-The negative relationship between omnivory and mean persistence is consistent with other studies showing that an omnivory motif is less likely to be stable than the other motifs considered~\citep{Borrelli2015a}.
+%Since motifs, basic building blocks of networks~\citep{Milo2002} and bridges between network-wide (global) and species-specific (local) properties, describe both direct and indirect effects, the distribution of these meso-scale structures in ecological networks is likely related to network and species persistence when the plant community is affected by species loss.
+%Since motifs describe direct and indirect effects~\citep{Milo2002}, the distribution of these meso-scale structures in ecological networks is likely related to network and species persistence after a disturbance~\citep{}.
 
-High frequency of the apparent competition motif was correlated with an on average higher in-degree for consumers. In-degree in turn is often related to higher persistence \citep{}.
+As basic building blocks of networks~\citep{Milo2002}, motifs connect network-wide (global) and species-specific (local) properties, describing both direct and indirect effects.
+As such, the distribution of these meso-scale structures can be of importance to network and species persistence~\citep{}.
+Indeed, we find that both the distribution of motifs at the network level and the set of motifs a target species participates in are related to persistence after disturbance to the plant community, and that the nature of these relationships synthesizes the effects of other local and global-scale properties.
 
-Motif profiles with high apparent competition and low omnivory were also more likely in less-connected food webs, which had higher mean persistence than highly-connected food webs.
-Despite these general trends, networks with more variable motif profiles had higher mean persistence, suggesting that many different network structures support high probabilities of persistence. 
-High connectance might reduce mean persistence by limiting the number of possible meso-scale network structures.
-In particular, it is reasonable to expect higher frequencies of the omnivory motif at high connectances since this motif contains more links than the other three-species motifs considered here.
-% Top species directly connected to the disturbance, disturbance not mitigated via herbivores. Duality in the role of omnivory.
-% The relationship between high connectance and frequency of omnivory motif should perhaps come earlier. 
-% Close with something about how more links allow disturbance to propagate.
+A network's motif profile had a relatively strong relationship to the network mean persistence compared to global structures.
+However, connectance did affect which network motifs were most common.
+For example, highly-connected networks tended to contain higher frequencies of the omnivory motif which includes more links than the other three-species motifs considered here.
+Thus, global-scale properties may have stronger effects on network persistence by changing meso-scale properties than directly.
+This in turn suggests that network motif profiles may be more useful than global network properties when predicting network-level extinction risk.
 
-% HERE ENDS MEAN PERSISTENCE AND SPECIES WISE START
-As well as the mean persistence of all consumers, we also considered the species-specific persistence of each consumer in a network.
-While the persistence of a particular species is likely to be affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it also depends on a species' position in the network.
-There are many ways of measuring species' positions, and in some cases multiple measures may capture the same trend.
-In our results, motif participation is significantly related to persistence, but the effects of two motifs (\emph{apparent competition} and \emph{direct competition}) might be explained more simply by in-degree and trophic level.
-This is especially likely for these two motifs, since Bayesian Network modeling lacks top-down effects and therefore removes the indirect effects from these motifs.
+% Network-level persistence
+In general, average network persistence was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif. 
+%The negative relationship between omnivory and network persistence is consistent with other studies showing that an omnivory motif is less likely to be stable than the other motifs considered~\citep{Borrelli2015a}.
+Motif profiles with high apparent competition and low omnivory were more likely in less-connected food webs, which generally had higher mean persistence than highly-connected food webs.
+Despite these general trends, networks with more variable motif profiles had higher mean persistence, suggesting that presence of a variety of different network structures support persistence.
 
-% Getting rid of AC/DC "quickly" here so we can focus on omnivory
-Participation in the apparent competition motif is strongly negatively correlated with in-degree---a high proportion of the apparent competition motif implies a low in-degree. 
-Both higher in-degree and lower participation in the apparent competition motif are associated with greater persistence at low disturbance but decreased persistence at high disturbance; because the trends are so similar, it is likely that a common mechanism underlies both.
-% The effect on persistence is consistent between species with a low in-degree and a high proportion of the apparent competition motif. 
-%Low in-degree decreases persistence at low disturbance, and increases persistence at high disturbance, similar to the trend visible for a high participation in the apparent competition motif. 
-% Similarly, the effect on species persistence is reversed for a low proportion of the apparent competition motif in which species have a high in-degree.
+A network's motif profile ...
+Both omnivory and apparent competition have been shown to have consequences for the stability of ecological networks.
+However, particularly the role of omnivory is debated and it has been shown to act both stabilizing and destabilizing.
+[Add examples of both ...
+]
 
-While the direct competition motif is not correlated with in-degree, it is strongly negatively correlated with trophic level---species with a high proportion of the direct competition motif have a low trophic levels.
-Low trophic levels are always associated with higher persistence because these species are less affected by bottom-up disturbances~\citep{}.
-It then follows that species which participate in many direct competition motifs may have higher probabilities of persistence because they are also likely to have low trophic levels.
+Moreover, species within a network can have substantially different probabilities of persistence than the mean across all species.
+For targeted conservation actions this type of information is often more relevant.
 
-In contrast to the competition motifs, the three-species chain and omnivory motifs do have indirect effects in a Bayesian Network setting and the effects of these motifs cannot be fully explained by degree or trophic level.
-For example, the three-species chain motif is negatively correlated with in-degree and positively correlated with trophic level. 
-Despite this correlation, participation in the three-species chain does not interact the same way with disturbance as either simpler measure and so has a unique relationship to species persistence.
-This is likely due to indirect effects of the bottom species in the chain on the top species.
+% Species level persistence
+While the persistence of a particular species is affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it also depends on local-scale (trophic level and in-degree) network properties.
+Local properties may be especially important for the apparent competition and direct competition motifs.
+The indirect effects in these two motifs are not fully expressed since the Bayesian Network framework does not include top-down effects.
+The direct effects they do describe echo the effects of in-degree and trophic level.
 
-The most interesting motif in which to consider indirect effects is omnivory.
-While participation in the omnivory motif is positively correlated with in-degree, a high proportion of the omnivory motif is always associated with increased persistence, while the effect of in-degree varies with disturbance.
-Thus, the effect of a high proportion of the three-species chain motif and the omnivory motif does not follow the corresponding effect of in-degree and trophic level.
+However, the effects of in-degree and trophic level differed.
+A high trophic level was associated with lower persistence across all levels of disturbance.
+This was expected, as species at higher trophic levels here by definition have equal or higher probability of extinction compared to their prey ~\citep{Eklof2013}.
+The effect of in-degree is less obvious and depended on the level of disturbance; a high in-degree had a positive effect on persistence at low levels of disturbance but not at high levels of disturbance.
+In addition to these different relationships with persistence, in-degree and trophic level had different relationships to the four motifs we investigated.
 
-% Two notable results: 1) omnivory bad at network level, good at species level, 2) apparent competition depends on disturbance
-Combining the results from our species wise exploration with our network-level findings, we see that while being in a network with high frequencies of omnivory decreases persistence, participating in many of these omnivory motifs increases persistence at the species level.
+Participation in the apparent competition motif was strongly negatively correlated with in-degree --a high proportion of the apparent competition motif implies a low in-degree, and both are associated with greater persistence at low disturbance but decreased persistence at high disturbance.
+This suggests that in-degree is in fact the factor driving the positive relationship between proportion of apparent competition motif and species persistence.
+Interestingly though, the relationship changes from positive to negative with increasing level of disturbance.
+When disturbance is low many prey species acts as an insurance, while at high disturbance connections to many prey species is in fact negative as the consumer receives the negative impact from several sources.
+In-degree does not, however, explain all of the trends we observe for motifs as degree and motif participation are not always tightly coupled.
+For example, it is possible for a species with low degree to participate in many motifs if it interacts with high-degree species and thereby obtain many indirect interaction partners.
+This is one of the strengths with analyses of motifs that it considers local structures additionally include indirect connections one link away.
+
+Trophic level, like motifs, includes some information about the risk of indirect effects; species at high trophic levels can be at increased risk of extinction due to disturbances at lower trophic levels~\citep{Eklof2013}.
+Species with high trophic levels also tend to have low participation in the direct competition motif, which is in turn associated with lower probability of persistence at all levels of disturbance.
+The clear positive relationship between proportion of direct competition motifs and species persistence is therefore likely due to the tendency for species at low trophic levels to have a higher proportion of that motif.
+
+The remaining two motifs are strongly related to both in-degree and trophic level and must be explained using a combination of the two simpler network properties.
+Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between them.
+The three-species chain motif, for example, is moderately negatively correlated with in-degree and moderately positively correlated with trophic level. 
+The relationship between persistence and the chain motif therefore combines the overall negative effect of high trophic level with the interaction between in-degree and disturbance; leading the chain motif to have little relationship to persistence at low disturbance and an increasingly negative relationship at higher disturbance.
+
+The omnivory motif likewise integrates effects of in-degree and trophic level.
+Species participation in the omnivory motif was associated with greater persistence, especially at low levels of disturbance.
+This is consistent with a strong negative correlation between omnivory and trophic level (leading to the overall positive effect of omnivory) and a strong positive correlation with in-degree (leading to the reduced benefit of omnivory at high levels of disturbance).
+
+Despite participating more often in the omnivory motif being associated with greater persistence for a focal species, having a greater proportion of omnivory motifs in the network was associated with lower network persistence.
 At the network level, a higher frequency of the omnivory motif leads to more pathways by which the disturbances to basal resources can propagate and cause extinctions of consumers.
@@
 This was not the case in less connected networks.
-In a highly connected network, a disturbance could potentially spread to a higher extent than in a less connected network. 
+Following the same reasoning as above, in a highly connected network,the numerous links pave the way for a disturbance to spread to a higher extent than in a less connected network. 
 Therefore, in a highly connected web where disturbances have more ways to move, the lower level prey item in the omnivory motif may be of even higher importance assisting as a buffer, decreasing the negative effect of disturbance for the consumer species.
-In a less connected network, this additional prey species is of less importance.
-
-To understand the species-level effect, we consider the relationships between stability and each position within the motif.
-Appearing frequently in the top or middle positions of the omnivory motif is associated with higher probabilities of persistence at low levels of disturbance, while appearing frequently in the bottom position is associated with higher probabilities of persistence at high levels of disturbance.
-This is consistent with the logic outlined for the network-level effect: when extinction of basal species is rare, consumers at the top or middle of omnivory motifs benefit from multiple energy channels but when extinction is common these multiple channels mean that the consumer is likely to be affected by multiple disturbances.
-The bottom position shows the reverse effect: at low levels of disturbance, this position is nearly neutral.
-At high levels of disturbance, however, being low down in the motif (and obtaining energy by appearing as a predator in other motifs) means that the species is less likely to be affected by multiple lower-level extinctions.
-The offsetting of effects of different positions within the omivory motif may explain why the motif as a whole is associated with greater probabilities of persistence across all levels of disturbance we considered.
 
@@
 
-% Broader context of omnivory is probably better after we introduce the species-level results. 
 The relationship between omnivory and stability is subject to a longstanding debate in ecology.
-Omnivory has been shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, MonteiroFaria2016, Kratinaetal2012}.
+Omnivory has been shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, Monteiro2016, Kratina2012}.
 For example, omnivorous interactions may not be stable in isolation, but instead stabilized by interactions with other species when embedded in a larger food web \citep{Kratinaetal2012}.
@@
 They found that a high proportion of the omnivory motif resulted in a shorter time to extinction, while the other three stable motifs were associated with longer times to extinction.
-However, the study by Cirtwill and Wootton used a different approach using a dynamic simulation model also including top-down effects.
+However, the study by Cirtwill and Wootton used a different approach using a fully dynamic simulation model also including top-down effects.
+
+Our study of network and species motif profiles show that information on and synthesis of the global and local network structures can provide important information on which conservation actions that will most successfully preserve biodiversity in ecological networks.
+When affected by disturbance of plant community...
+
     %In previous studies of the effect of omnivory motifs, the strength of the omnivorous link can potentially be important.
@@
 
-[[Comparison to degree, tl?]]
-
-    % This paragraph explains why motif trends aren't just due to TL
-    Two of the focal motifs, omnivory and the three-species chain, include three trophic levels while apparent and direct competition only require two.
-    However, the effects of motifs on persistence cannot be reduced to the number of trophic levels in each motif.
-    The two predators in a direct competition motif or the two prey in an apparent competition motif are not necessarily at the same trophic level.
-This is demonstrated by the increasing participation in apparent competition and chain motifs and decreasing participation in direct competition and omnivory motifs with increasing trophic level (i.e., 2-TL and 3-TL motifs do not show consistent trends).
-    In fact, a prey in a direct competition motif can have a higher trophic level than its predator.
-    For example, blueberries and salmon consumed by a bear make an apparent competition motif where the salmon, a piscivorous fish, has a higher trophic level than the blueberries or the omnivorous bear.
-    This means that motif participation and trophic level give complementary descriptions of a species' place in the network and explains why the simple decrease in participation with increasing trophic level does not explain the more complicated relationships between motif participation and persistence.
-
-    In-degree (number of resources) may explain more of the relationships between motif participation and persistence.
-    For example, higher proportions of the omnivory motif in a species' participation vector were associated with higher persistence for that species, and participation in omnivory significantly increased with increasing in-degree.
-    Moreover, the relationship between degree and persistence  showed a strong interaction with disturbance, similar to the varying relationship between apparent competition and disturbance. 
-    The shift from positive relationships between persistence and apparent competition at high levels of disturbance to negative relationships between persistence and apparent competition at low levels of disturbance could be partially explained by the negative relationship between apparent competition and in-degree.
-    However, participation in the three-species chain also significantly decreased with increasing degree while participation in omnivory increased, and neither of these motifs shows a similar interaction with disturbance.
-    As with trophic level, motifs provide some unique information not provided by in-degree (and vice versa). 
-    Both ways of describing a species' position in the network show significant relationships to consumers' persistences after a disturbance and, because these relationships are not the same, likely capture different mechanisms leading to greater or lesser extinction risk.
-
- %When analyzing motifs within a Bayesian Network, it is important to keep in mind that only bottom-up effects are accounted for.
-Therefore, the implications of the four possible motifs are different than they would be in an empirical network.
-
-    %In particular, the lack of top-down effects removes the indirect effects from the two competition motifs.
-In the case of direct competition, two predators share a common resource.
-The predators will, within a Bayesian Network, not affect each other indirectly although both will be affected by any disturbance to their resource.
-Similarly, the two bottom-level species in the apparent competition motif will be independent of each other as there are no top-down effects of the shared consumer to link them.
-    %The difference between the direct competition and apparent competition motif is thus the number of bottom-level species, accounted for as prey items for a higher level species.
-   % A predator participating in many direct competition motifs is likely to have fewer prey items than a predator participating frequently in the apparent competition motif, and therefore a higher extinction risk.Note, however, that consumers can also appear in these motifs as prey to other consumers.
-In that case, we would not expect the two motifs to have different effects on extinction risk.
-    %The structure of the omnivory motif and the three-species chain motif differ from the competition motifs by having an additional trophic level.
-Therefore, top-level species of these motifs might be more sensitive to disturbances since they can be affected by the loss of the basal resource in the focal motif or an increased probability of extinction of the middle species due to a disturbance to a basal resource not included in the focal motif.
-
 % [[When can motifs still be useful predictors and add insight?]]
@@
 
+\section*{Glossary}
+
+\begin{table}[h!]
+\label{glossary}
+\caption{Glossary of terms relating to motifs and Bayesian networks}
+
+\begin{tabular}{l|l}
+    Term & Definition \\
+    \hline
+    Motif &  \\
+    Network profile & \\
+    Motif participation & \\
+    Motif role & \\
+    Bayesian network & \\
+    Network persistence & \\
+    Species persistence & \\
+    In-degree & \\
+    Trophic level (STL) & \\
+    Disturbance ($\pi_{disturbed}$) & Probability that a basal resource will go extinct when extra disturbance is added. \\
+    Baseline extinction&  When no disturbance is applied, $\pi_{base}$=0.1.\\
+\end{tabular}
+\end{table}
+
+% Notes on what expressions and words to use
+% "probability of extinction" or "level of disturbance" for the different disturbace scenarios?
+
 \bibliographystyle{ecollett} 

### 2021-07-20 manuscript/supplement_motifs.tex

@@
         \label{fig:lm_CS}
+    \end{figure}
+
+    \begin{figure}[h!]
+        \centering
+        \includegraphics[height=0.75\textheight]{figures/persistence_vs_BSC_lm.eps}
+        \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while the proportion of the network made up of basal resources, network size, connectance, and their interactions had smaller effects.
+We show relationships with persistence for low disturbance ($\pi_{Basal}$=0.1) and high disturbance($\pi_{Basal}$=0.5).
+For each predictor other than disturbance, we show the relationship between the focal predictor and persistence for combinations of the lowest and highest observed values of the other predictors.
+For proportion of basal resources, these values were 0.02 and 0.5; for network size, 50 and 100; and for connectance, 0.02 and 0.18.}
+        \label{fig:lm_BCS}
     \end{figure}

### 2021-07-20 code/motif_BC_lms.py
proportions vs basal resources

@@ -1,0 +1,134 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Blues', reverse=True)  # The blue is very beautiful but maybe harder to see.
+colors.add_color(173,132,191,'Purple')
+colors.add_color(34,90,34,'Green')
+# colors.add_color(120,120,120,'grey')
+# colors.add_color(255,125,125,'lightish_red')
+# colors.add_color(200,200,200,'lightgrey')
+
+def format_graph(graph,graphtype):
+  graph.yaxis.bar.linewidth=1
+  graph.xaxis.bar.linewidth=1
+  graph.frame.linewidth=1
+
+  graph.world.xmin=0
+  graph.world.xmax=0.51
+  graph.xaxis.tick.major=.1
+
+  graph.world.ymin=0
+  graph.world.ymax=.6000001
+  graph.yaxis.tick.major=.2
+  graph.yaxis.label.configure(text='Proportion',char_size=1,just=2,place='normal')
+  graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
+
+
+  graph.xaxis.ticklabel.configure(format='decimal',prec=2,char_size=.75)
+  graph.xaxis.label.configure(text='Proportion basal resources',char_size=1,just=2,place='normal')
+
+  graph.xaxis.tick.configure(onoff='on',minor_ticks=0,major_size=.4,minor_size=.5,place='both',major_linewidth=.5,minor_linewidth=1)
+  graph.yaxis.tick.configure(onoff='on',minor_ticks=0,major_size=.4,minor_size=.5,place='both',major_linewidth=.5,minor_linewidth=1)
+
[…]

### 2021-07-20 code/motif_BC_lms.py
fixed x axis

@@ -40,7 +40,7 @@ def format_graph(graph,graphtype):
   graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
 
 
-  graph.xaxis.ticklabel.configure(format='decimal',prec=2,char_size=.75)
+  graph.xaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
   graph.xaxis.label.configure(text='Proportion basal resources',char_size=1,just=2,place='normal')
 
   graph.xaxis.tick.configure(onoff='on',minor_ticks=0,major_size=.4,minor_size=.5,place='both',major_linewidth=.5,minor_linewidth=1)
@@ -121,7 +121,7 @@ for lm in ['apparent','direct','omnivory','chain']:
 
 grace.multi(rows=2,cols=2,vgap=.03,hgap=.04)
 grace.hide_redundant_labels()
-grace.set_row_xaxislabel(label='Connectance',row=1,colspan=(None,None),char_size=1,perpendicular_offset=.05,just=2)
+grace.set_row_xaxislabel(label='Proportion basal resources',row=1,colspan=(None,None),char_size=1,perpendicular_offset=.05,just=2)
 grace.set_col_yaxislabel(label='Proportion of total motifs',col=0,rowspan=(None,None),char_size=1,perpendicular_offset=.05,just=2)
 # grace.set_row_xaxislabel(label='Day of year',row=2,colspan=(None,None),char_size=1,perpendicular_offset=.05)
 for graph in grace.graphs:

### 2021-07-20 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
         \label{fig:lm_CS}
-    \end{figure}
-
-    \begin{figure}[h!]
-        \centering
-        \includegraphics[height=0.75\textheight]{figures/persistence_vs_BSC_lm.eps}
-        \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while the proportion of the network made up of basal resources, network size, connectance, and their interactions had smaller effects.
-We show relationships with persistence for low disturbance ($\pi_{Basal}$=0.1) and high disturbance($\pi_{Basal}$=0.5).
-For each predictor other than disturbance, we show the relationship between the focal predictor and persistence for combinations of the lowest and highest observed values of the other predictors.
-For proportion of basal resources, these values were 0.02 and 0.5; for network size, 50 and 100; and for connectance, 0.02 and 0.18.}
-        \label{fig:lm_BCS}
     \end{figure}
@@
         \centering
-        \includegraphics{figures/persistence_positions_Apparent.eps}
+        \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Apparent.eps}
         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
@@
         \centering
-        \includegraphics{figures/persistence_positions_Direct.eps}
+        \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Direct.eps}
         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
@@
         \centering
-        \includegraphics{figures/persistence_positions_Three-species.eps}
+        \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Three-species.eps}
         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
@@
 
+            Positions in all three motifs were strongly related to in-degree and/or trophic level (Table~\ref{DT_tab}).
             The top position in the omnivory motif increased with increasing degree across all trophic levels.
@@
                 \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
-                \label{appcomp_DT}
+                \label{chain_DT}
             \end{figure}
@@
             % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
-            The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance.
+            The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance (Table~\ref{SC_tab}).
             The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.
@@
                 \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against network size, connectance, and their interaction.}
-                \label{DT_tab}
+                \label{SC_tab}
                 \footnotesize
@@
                 \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
-                \label{omnivory_DT}
+                \label{omnivory_SC}
             \end{figure}
@@
                 \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
-                \label{appcomp_DT}
+                \label{appcomp_SC}
             \end{figure}
@@
                 \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
-                \label{dircomp_DT}
+                \label{dircomp_SC}
             \end{figure}
@@
                 \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
-                \label{appcomp_DT}
+                \label{chain_SC}
             \end{figure}
@@
     \clearpage
+
+\clearpage
+
+\section*{SJ: Proportion of basal resources in the network}
+
+    \subsection*{Basal resources and mean persistence}
+
+        We fit a linear regression relating consumer persistence to network size, connectance, disturbance, the proportion of the network made up of basal resources (pBasal), and all interactions between them.
+All terms were significant, indicating that the pBasal affects a consumer's probability of persistence but that this effect depends upon other aspects of network structure.
+        As with the effects of network size and connectance, however, the effects of pBasal were small in the range of networks we considered here (Fig.~\ref{fig:lm_BCS}).
+        % Table is from basal_models.Rdata: BCper, if we want it.
+
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[height=0.75\textheight]{figures/persistence_vs_BSC_lm.eps}
+            \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while the proportion of the network made up of basal resources, network size, connectance, and their interactions had smaller effects.
+We show relationships with persistence for low disturbance ($\pi_{Basal}$=0.1) and high disturbance($\pi_{Basal}$=0.5).
+For each predictor other than disturbance, we show the relationship between the focal predictor and persistence for combinations of the lowest and highest observed values of the other predictors.
+For proportion of basal resources, these values were 0.02 and 0.5; for network size, 50 and 100; and for connectance, 0.02 and 0.18.}
+            \label{fig:lm_BCS}
+        \end{figure}
 
@@
 
+    \subsection*{Motif profiles and proportion of basal resources}
+
+        We repeated our analyses testing whether motif profiles were related to network size and connectance, replacing network size with the proportion of the network made up by basal resources (pBasal).
+        A network's overall motif profile was significantly related to pBasal ($F_{1,2996}$=196, $p$\textless0.001, connecance ($F_{1,2996}$=10.0, $p$\textless0.001), and their interaction ($F_{1,2996}$=27.3, $p$\textless0.001).
+        However, this result could be a false positive as the dispersion of motif profiles was non-homogeneous across connectance ($F_{4,2995}$=5.17, $p$\textless0.001), pBasal ($F_{128,2871}$=19.0, $p$\textless0.001), and their interaction ($F_{291,2708}$=13.4, $p$\textless0.001).
+        Distance from group centroids was low for networks with extremely low or extremely high pBasal, perhaps because there were few such networks.
+        The ranges and median distances from group centroids was similar across levels of connectance.
+
+        Considered separately, the proportions of all four motifs in a network's motif profile were related to the proportion of basal resources in a network (Fig.~\ref{fig:Blms}).
+        The proportion of the three-species chain decreased in networks with more basal resources ($\beta_{pBasal}$=-0.080, $p$\textless0.001) and networks with higher connectance ($\beta_{C}$=0.498, $p$\textless0.001).
+        The interaction between pBasal and connectance was not significant ($\beta_{pBasal:C}$=0.0554, $p$=0.766).
+
+        The proportion of the omnivory motif decreased in networks with more basal resources ($\beta_{pBasal}$=-0.198, $p$\textless0.001), increased in networks with higher connectance ($\beta_{C}$=0.895, $p$\textless0.001), and increased most strongly in networks with high connectance and many basal resources ($\beta_{pBasal:C}$=2.53, $p$\textless0.001).
+
+        The proportion of the apparent competition motif increased with in networks with more basal resources ($\beta_{pBasal}$=0.202, $p$\textless0.001) and decreased in more connected networks ($\beta_C$=-0.596, $p$\textless0.001).
+        The interaction between the proportion of basal resources and connectance was not significant ($\beta_{pBasal:C}$=0.258, $p$\textless0.001).
+
+        The proportion of the direct competition motif increased in networks with more basal resources ($\beta_{pBasal}$=0.0759, $p$\textless0.001) and more-connected networks ($\beta_C$=0.199, $p$\textless0.001), but decreased with the interaction between them ($\beta_{pBasal:C}$=-2.85, $p$\textless0.001).
+
+        \begin{figure}
+            \centering
+            \includegraphics[width=.75\textwidth]{figures/motif_proportion_basal_lms.eps}
+            \caption{Mean proportion of each motif in a network's motif profile varied with connectance and the proportion of basal resources in the network.
+For direct competition and omnivory, the frequency of the motif also varied with the interaction between connectance and basal resources.}
+            \label{fig:Blms}
+        \end{figure}
+
+\clearpage
+
+    \subsection*{}
+
 \section{Motif profiles in different network sets}
@@
 
-    \begin{table}[h!]
-        \centering
-        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{propreq}).
-Predictors were centered and scaled before model fitting.
-We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
-        \label{tab:proportion}                \footnotesize
-        \begin{tabular}{c|c c | c c | c c | c c | c c |}
-        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c|}{$R^2$} \\
-        Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
-        \hline
-        3-sp Chain & 0.574 & \textless0.001 & -0.00508 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
-        App. Com. & 0.574 & \textless0.001 & -0.00312 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
-        Dir. Com. & 0.574 & \textless0.001 & 0.00564 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
-        Omnivory & 0.574 & \textless0.001 & 0.00709 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
-        \end{tabular}
-    \end{table}        
 \clearpage

### 2021-07-21 code/stat_analysis/participation_vs_Basal.py
same figure for participation vs BC

@@ -1,0 +1,149 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Greys',n=10)  # The blue is very beautiful but maybe harder to see.
+# Match Anna's ggplot colours
+colors.add_color(68,1,84,'Rpurple')
+colors.add_color(65,68,135,'Rblue')
+colors.add_color(42,120,142,'Rgreen1')
+colors.add_color(34,168,132,'Rgreen2')
+colors.add_color(122,209,81,'Rgreen3')
+colors.add_color(253,231,37,'Ryellow')
+colors.add_color(254,224,139,'Ryellow2')
+colors.add_color(253,174,97,'Rorange1')
+colors.add_color(244,109,67,'Rorange2')
+colors.add_color(213,62,79,'Rred1')
+colors.add_color(158,1,66,'Rred2')
+
+mot_ranges={
+  'omnivory':(0,0.54917),
+  'apparent':(0,1),
+  'direct':(0,1),
+  'chain':(0,1),
+}
+
+motif_names={'Omnivory':'Omnivory','Apparent':'Apparent competition','Direct':'Direct competition','Chain':'Three-species chain'}
+
+# Each point will be mean of one network
+def read_datafile(datafile):
+  netprops={}
+  f=open(datafile,'r')
+  for line in f:
+    if line.split()[0]!='"V1"':
+      motif=line.split()[1][1:-1]
+      intercept=float(line.split()[2][1:-1])
+      pBasal=float(line.split()[3][1:-1])
+      C=float(line.split()[4][1:-1])
+      interaction=float(line.split()[5][1:-1])
+
+      netprops[motif]=((intercept,pBasal,C,interaction))
+
+  f.close()
+  return netprops
+
[…]

### 2021-07-21 manuscript/supplement_motifs.tex

@@
         \centering
-        \includegraphics[width=.95\textwidth]{figures/roles_vs_SC.eps}
-
-        \caption{The proportion of each motif in a species' participation vector varied with network size, connectance, and their interaction.
+        \includegraphics[width=.95\textwidth]{figures/participation_vs_SC.eps}
+        \caption{The proportion of each motif in a consumer's participation vector varied with network size, connectance, and their interaction.
 These relationships were steepest for apparent competition and omnivory.}
@@
 
-    \subsection*{}
+    \subsection*{Motif participation and proportion of basal resources}
+
+        We fit regressions relating the proportion of each motif in a consumer's motif participation vector to the proportion of basal resources in the network (pBasal), connectance, and the interaction between them.
+        Participation in all four motifs was significantly related to pBasal, connectance, and the interaction between them (Fig.~\ref{fig:basal_participation}).
+
+        \begin{figure}
+            \centering
+            \includegraphics{}
+            \caption{Caption}
+            \label{fig:basal_participation}
+        \end{figure}
+
+        Participation in the three-species chain increased in networks with more basal resources ($\beta_{pBasal}$=0.0447, $p$\textless0.001), decreased in more-connected networks ($\beta_C$=-0.296, $p$\textless0.001), and decreased with the interaction between them ($\beta_{pBasal:C}$=-0.525, $p$\textless0.001).
+        Participation in the omnivory motif decreased in networks with more basal resources ($\beta_{pBasal}$=-0.198, $p$\textless0.001), increased in more-connected networks ($\beta_C$=0.701, $p$\textless0.001), and increased strongly with the interaction between them ($\beta_{pBasal:C}$=2.81, $p$\textless0.001).
+        Participation in the apparent competition motif decreased in networks with more basal resources and more-connected networks ($\beta_{pBasal}$=-0.0825, $p$\textless0.001 and $\beta_C$=-0.846, $p$\textless0.001) but increased strongly with the interaction between them ($\beta_{pBasal:C}$=1.66, $p$\textless0.001).
+        Participation in the direct competition motif increased in networks with more basal resources and more-connected networks ($\beta_{pBasal}$=0.235, $p$\textless0.001 and $\beta_C$=0.442, $p$\textless0.002), but decreased strongly with the interaction between them ($\beta_{pBasal:C}$=-3.94, $p$\textless0.001).
 

### 2021-07-21 code/stat_analysis/prop_basal_models.R
fit models

@@ -10,22 +10,34 @@ library(MuMIn)
 
 # data is full data, including basals
 data$network_ID=paste(data$Size,data$Connectance,data$Network,sep='_')
-for(r in 1:nrow(data)){
-    netdat=data[which(data$network_ID==data$network_ID[r]),]
+subdata=data[which(data$Disturbance==0.1),]
+for(r in 1:nrow(subdata)){
+    netdat=subdata[which(subdata$network_ID==subdata$network_ID[r]),]
     nBasal=length(which(netdat$in_Degree==0))
-    n=data$Size[r]
-    data$prop_Basal[r]=nBasal/n
+    n=subdata$Size[r]
+    subdata$prop_Basal[r]=nBasal/n
 }
 
 # consumers is heterotrophs only
-
+for(r in 1:nrow(consumers)){
+    netID=consumers$network_ID[r]
+    netter=subdata[which(subdata$network_ID==netID),]
+    consumers$prop_Basal[r]=netter$prop_Basal[1]
+}
+save.image('basal_models.Rdata')
 
 # A: Does persistence vary with proportion of basal resources?
-SCper=with(consumers,lm(Persistence~scale(prop_Basal)*scale(Size)*scale(Connectance)*scale(Disturbance)))
+BCper=with(consumers,lm(Persistence~scale(prop_Basal)*scale(Size)*scale(Connectance)*scale(Disturbance)))
 
 
 # B: How do motif profiles vary with prop_Basal and C?
     motdata=read.table('../../data/3sp_motif_profiles.tsv',header=TRUE)
+    motdata$network_ID=paste0(motdata$Size,'_',motdata$Connectance,'_',motdata$Network)
+    for(r in 1:nrow(motdata)){
+        netdat=subdata[which(subdata$network_ID==motdata$network_ID[r]),]
+        motdata$prop_Basal[r]=netdat$prop_Basal[1]
+    }
+
     motdata$mot_total=rowSums(motdata[,4:7])
     motdata$prop_chain=motdata$m12/motdata$mot_total
     motdata$prop_apparent=motdata$m6/motdata$mot_total
@@ -47,15 +59,17 @@ SCper=with(consumers,lm(Persistence~scale(prop_Basal)*scale(Size)*scale(Connecta
         consumers$net_omni[r]=motdata$prop_omni[which(motdata$netID==net)]
     }
 
-    save.image('analysis.Rdata')
+    save.image('basal_models.Rdata')
 
[…]

### 2021-08-12 code/stat_analysis/persistence_vs_profiles.py
colours closer, lines bolder|

@@ -19,8 +19,8 @@ from PyGrace.Styles.el import ElGraph, ElLogColorBar
 
 colors=ColorBrewerScheme('Greys',n=10)  # The blue is very beautiful but maybe harder to see.
 # Match Anna's ggplot colours
-colors.add_color(68,1,84,'Rpurple')
-colors.add_color(65,68,135,'Rblue')
+colors.add_color(158,101,184,'Rpurple')
+colors.add_color(145,148,215,'Rblue')
 colors.add_color(42,120,142,'Rgreen1')
 colors.add_color(34,168,132,'Rgreen2')
 colors.add_color(122,209,81,'Rgreen3')
@@ -133,39 +133,20 @@ def format_graph(graph,simple):
 
   return graph
 
-def populate_graph(graph,bindict):
-  j=2
-  motnames={'appcomp':'Apparent competition','chain':'3-species chain','dircomp':'Direct competition'}
-  for motif in ['appcomp','chain','dircomp']:
-    data=graph.add_dataset(bindict[motif],type='xydy')
-    data.symbol.shape=0
-      
-    data.line.configure(linestyle=1,linewdith=1.5,color=j)
-    data.errorbar.configure(riser_color=j,color=j,riser_linewidth=1,size=.5,linewidth=1)
-
-    # data.legend=motnames[motif]
-    # graph.legend.configure(char_size=.75,loc=(0.2,0.58),loctype='world',box_linestyle=0,fill_pattern=0)
-    if j==11:
-      j=13
-    elif j==2:
-      j=11
-
-  return graph
-
 def populate_persgraph(graph,motif):
 
   for dist in [0.1,0.5]:
     if motif=='apparent':
-      j=13
+      j='Rpurple'
       sty=2
     elif motif=='chain':
-      j=14
+      j='Rblue'
       sty=3
     elif motif=='direct':
-      j=16
-      sty=5
+      j='Rgreen2'
+      sty=8
     else:
-      j=18
+      j='Rorange1'
       sty=1
 
     # for dist in [0.1,0.18,0.26,0.34,0.42,0.5]:
@@ -183,7 +164,7 @@ def populate_persgraph(graph,motif):
 
     data=graph.add_dataset(dats)
     data.symbol.shape=0
-    data.line.configure(linestyle=sty,linewdith=1.5,color=j)
+    data.line.configure(linestyle=sty,linewidth=4,color=j)
 
     if dist==0.1:
       data.legend=motif_names[motif]
[…]

### 2021-08-12 manuscript/motifs_vs_degree.tex

@@
 
-%Since motifs, basic building blocks of networks~\citep{Milo2002} and bridges between network-wide (global) and species-specific (local) properties, describe both direct and indirect effects, the distribution of these meso-scale structures in ecological networks is likely related to network and species persistence when the plant community is affected by species loss.
-%Since motifs describe direct and indirect effects~\citep{Milo2002}, the distribution of these meso-scale structures in ecological networks is likely related to network and species persistence after a disturbance~\citep{}.
-
-As basic building blocks of networks~\citep{Milo2002}, motifs connect network-wide (global) and species-specific (local) properties, describing both direct and indirect effects.
+Motifs are the basic building blocks of networks~\citep{Milo2002}, connecting network-wide (global) and species-specific (local) properties, describing both direct and indirect effects.
 As such, the distribution of these meso-scale structures can be of importance to network and species persistence~\citep{}.
-Indeed, we find that both the distribution of motifs at the network level and the set of motifs a target species participates in are related to persistence after disturbance to the plant community, and that the nature of these relationships synthesizes the effects of other local and global-scale properties.
+We find that both the distribution of motifs at the network level and the set of motifs a target species participates in are related to persistence after disturbance to the plant community.
+Also, the nature of these relationships synthesizes the effects of other local and global-scale properties.
 
@@
 While the persistence of a particular species is affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it also depends on local-scale (trophic level and in-degree) network properties.
-Local properties may be especially important for the apparent competition and direct competition motifs.
-The indirect effects in these two motifs are not fully expressed since the Bayesian Network framework does not include top-down effects.
-The direct effects they do describe echo the effects of in-degree and trophic level.
+Local properties may be especially important for the apparent competition and direct competition motifs, due to the indirect effects not being fully expressed in these two motifs since the Bayesian Network framework does not include top-down effects.
 
-However, the effects of in-degree and trophic level differed.
+The effect of the local-scale network properties in-degree and trophic level differed.
 A high trophic level was associated with lower persistence across all levels of disturbance.
@@
 
-Participation in the apparent competition motif was strongly negatively correlated with in-degree --a high proportion of the apparent competition motif implies a low in-degree, and both are associated with greater persistence at low disturbance but decreased persistence at high disturbance.
-This suggests that in-degree is in fact the factor driving the positive relationship between proportion of apparent competition motif and species persistence.
-Interestingly though, the relationship changes from positive to negative with increasing level of disturbance.
-When disturbance is low many prey species acts as an insurance, while at high disturbance connections to many prey species is in fact negative as the consumer receives the negative impact from several sources.
+Participation in the apparent competition motif was strongly negatively correlated with in-degree --a high proportion of the apparent competition motif implies a low in-degree, and both are associated with lower persistence at low disturbance but increased persistence at high disturbance.
+This suggests that in-degree is in fact the factor driving the relationship between proportion of apparent competition motif and species persistence.
+The turnover from a negative to a positive relationship with an increasing level of disturbance indicate that when disturbance is high, connections to many prey species is negative as the consumer receives the negative impact from several sources.
+A species with a high proportion of the apparent competition motif, which also has a low in-degree, thus gets an advantage when plant level disturbance is high since fewer prey items transmit the disturbance.
 In-degree does not, however, explain all of the trends we observe for motifs as degree and motif participation are not always tightly coupled.
@@
 
-Trophic level, like motifs, includes some information about the risk of indirect effects; species at high trophic levels can be at increased risk of extinction due to disturbances at lower trophic levels~\citep{Eklof2013}.
-Species with high trophic levels also tend to have low participation in the direct competition motif, which is in turn associated with lower probability of persistence at all levels of disturbance.
+The direct competition motif was not strongly correlated to in-degree, but had a strong negative relationship with trophic level.
+Thus, species with a high proportion of the direct competition motif has commonly a low trophic level, which is associated with higher probability of persistence at all levels of disturbance.
 The clear positive relationship between proportion of direct competition motifs and species persistence is therefore likely due to the tendency for species at low trophic levels to have a higher proportion of that motif.
 
-The remaining two motifs are strongly related to both in-degree and trophic level and must be explained using a combination of the two simpler network properties.
-Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between them.
-The three-species chain motif, for example, is moderately negatively correlated with in-degree and moderately positively correlated with trophic level. 
-The relationship between persistence and the chain motif therefore combines the overall negative effect of high trophic level with the interaction between in-degree and disturbance; leading the chain motif to have little relationship to persistence at low disturbance and an increasingly negative relationship at higher disturbance.
+%The remaining two motifs are strongly related to both in-degree and trophic level and must be explained using a combination of the two simpler network properties.
+%Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between them.
+%The three-species chain motif, for example, is moderately negatively correlated with in-degree and moderately positively correlated with trophic level. 
+%The relationship between persistence and the chain motif therefore combines the overall negative effect of high trophic level with the interaction between in-degree and disturbance; leading the chain motif to have little relationship to persistence at low disturbance and an increasingly negative relationship at higher disturbance.
+Likewise, the three-species chain motif is correlated with trophic level, but the relationship is positive and a high proportion of this motif indicates a high trophic level.
+Hence, the negative effect on persistence of having a high proportion of the three-species chain motif is likely connected to the high trophic level.
 
-The omnivory motif likewise integrates effects of in-degree and trophic level.
+%The omnivory motif likewise integrates effects of in-degree and trophic level.
+The omnivory motif is strongly related to both in-degree and trophic level and must be explained using a combination of the two simpler network properties.
+Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between them
 Species participation in the omnivory motif was associated with greater persistence, especially at low levels of disturbance.
@@
 
-Additionally, the effect of the omnivory motif on persistence was closely connected to connectance.
+Additionally, the effect of the omnivory motif on a focal species' persistence was closely connected to connectance.
 In highly connected networks, a high proportion of the omnivory motif strongly contributed to higher persistence.
 This was not the case in less connected networks.
-Following the same reasoning as above, in a highly connected network,the numerous links pave the way for a disturbance to spread to a higher extent than in a less connected network. 
+Following the same reasoning as above, in a highly connected network, numerous links pave the way for a disturbance to spread to a higher extent than in a less connected network. 
 Therefore, in a highly connected web where disturbances have more ways to move, the lower level prey item in the omnivory motif may be of even higher importance assisting as a buffer, decreasing the negative effect of disturbance for the consumer species.

### 2021-08-12 manuscript/supplement_motifs.tex

@@
         \centering
-        \includegraphics[width=.95\textwidth]{figures/participation_vs_SC.eps}
+        \includegraphics[height=.65\textheight]{figures/participation_vs_SC.eps}
         \caption{The proportion of each motif in a consumer's participation vector varied with network size, connectance, and their interaction.
@@
         Participation in all four motifs was significantly related to pBasal, connectance, and the interaction between them (Fig.~\ref{fig:basal_participation}).
+        These interactions were strongest for apparent and direct competition.
 
@@
             \centering
-            \includegraphics{}
-            \caption{Caption}
+            \includegraphics[height=.65\textheight]{figures/participation_vs_BC.eps}
+            \caption{The proportion of each motif in a consumer's participation vector varied with the proportion of the network made up by basal resources, connectance, and their interaction.}
             \label{fig:basal_participation}
@@
         Participation in the direct competition motif increased in networks with more basal resources and more-connected networks ($\beta_{pBasal}$=0.235, $p$\textless0.001 and $\beta_C$=0.442, $p$\textless0.002), but decreased strongly with the interaction between them ($\beta_{pBasal:C}$=-3.94, $p$\textless0.001).
+
+    \subsection*{Frequencies of positions and proportion of basal resources}
+
+        As with the frequencies of each motif in a species' motif participation vector, the frequencies of positions in a species' motif role varied with the proportion of basal resources in the network, connectance, and their interaction (Table~\ref{BC_tab}).
+
+        \begin{table}[h!]
+            \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against the proportion of the network made up of basal resources (pBasal), connectance, and their interaction.}
+            \label{BC_tab}
+            \footnotesize
+            \begin{tabular}{c c | c c | cc | cc |}
+            Motif & Position & \multicolumn{2}{c}{pBasal} & \multicolumn{2}{c}{Connectance} & \multicolumn{2}{c}{Interaction} \\
+            & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
+            \hline
+            Chain & bottom & -0.116 & \textless0.001 & -0.266 & \textless0.001 & 0.217 & \textless0.001 \\
+            Chain & middle & 0.0779 & \textless0.001 & -0.107 & \textless0.001 & -0.386 & \textless0.001 \\
+            Chain & top & 0.0829 & \textless0.001 & 0.0776 & \textless0.001 & -0.356 & \textless0.001 \\
+            \hline
+            Omnivory & bottom & -0.104 & \textless0.001 & 0.203 & \textless0.001 & 0.937 & \textless0.001 \\
+            Omnivory & middle & -0.0668 & \textless0.001 & 0.217 & \textless0.001 & 1.24 & \textless0.001 \\
+            Omnivory & top & -0.0269 & \textless0.001 & 0.281 & \textless0.001 & 0.638 & \textless0.001 \\
+            \hline
+            App. comp. & bottom & -0.237 & \textless0.001 & -0.874 & \textless0.001 & 1.76 & \textless0.001 \\
+            App. comp. & top & 0.154 & \textless0.001 & 0.0279 & \textless0.001 & -0.0947 & 0.013 \\
+            \hline
+            Dir. comp. & bottom & -0.0631 & \textless0.001 & -0.0247 & \textless0.001 & -0.418 & \textless0.001 \\
+            Dir. comp. & top & 0.298 & \textless0.001 & 0.466 & \textless0.001 & =3.53 & \textless0.001 \\
+            \hline
+            \end{tabular}
+            \end{table}
+
+\clearpage
 

### 2021-08-12 manuscript/motifs_vs_degree.tex

@@
     %\subsection*{1 Why are bottom-up effects important for understanding stability?}
-     The plant community is the foundation upon which the myriad of species in food webs depend for their survival \citep{}.
+     The plant community is the foundation upon which the myriad of species in terrestrial food webs depend for their survival \citep{}.
 Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
@@
     One way to incorporate more nuance into descriptions of how species fit into their communities is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
-Importantly, these motif roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions.
+Importantly, these `motif roles' capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions.
     This means that two species with the same in-degrees and trophic levels can still participate in different sets of motifs if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
@@
 
-    %\subsection{6. only 4 (stable) motifs can appear in Bayesian networks}
+           %\subsection*{Here we test...}
+    Specifically, we are interested in whether a consumer's extinction risk depends on 1) the motif profile of the full food web and 2) the motifs in which a focal species participates. 
+    In addition, we want to understand how the information provided by motifs is related to classical global and local food web properties in order to better integrate different aspects of network theory.
     Of the 13 three-species motifs that can occur in empirical food webs there are four motifs that are over-represented in empirical networks and are thought to contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}: the three-species chain, omnivory, apparent competition and direct competition motifs.
@@
     Therefore, we focus on these four particularly important motifs.
-
-       %\subsection*{Here we test...}
-    Here we are interested in how disturbances at the basal level in food webs propagate up the network and whether a consumer's extinction risk depends on 1) the motif profile of the full food web and 2) the motifs in which a focal species participates.
-In particular, we want to understand whether the information provided by motifs is related to classical global and local food web properties such as network connectance and species trophic level, or whether motifs provide unique insight into bottom-up extinction cascades.
 
@@
 
-	\subsection*{Defining motif profiles and species roles}
+	\subsection*{Defining motif profiles and participation [species roles]}
 
@@
         To separate differences in motif structure from differences in network size, we normalized motif profiles by dividing the count of each motif by the total count of all motifs.
-        The motif profile for each network therefore describes the proportions of each motif and has a sum of excatly one.
+        The motif profile for each network therefore describes the proportions of each motif and has a sum of exactly one.
 
-        We then describe meso-scale structure at the species level by defining each species' motif participation and species' roles.
+        We then describe meso-scale structure at the species level by defining each species' motif participation and motif roles.
         Motif participation describes the frequency with which a focal species appears in each of the four motifs, analogous to the network motif profile~\citep{Stouffer2012}.
-        A species' role describes the frequency with which a focal species appears in each \emph{unique position} (e.g., top, middle, or bottom species in the three-species chain, top or bottom species in the apparent competition motif) within each motif ~\citep{Stouffer2012,Cirtwill2017}.
+        A species' motif role describes the frequency with which a focal species appears in each \emph{unique position} (e.g., top, middle, or bottom species in the three-species chain, top or bottom species in the apparent competition motif) within each motif ~\citep{Stouffer2012,Cirtwill2017}.
 		We calculated species' motif participation and roles using the Python package \emph{pymfinder}~\citep{pymfinder}.
@@
 
-        \subsubsection*{How does mean persistence vary depending on network properties?}
+        \subsubsection*{How does network mean persistence vary depending on network properties?}
 
@@
 
-        \subsubsection*{How does the persistence vary with motif frequency in the web?}
+        \subsubsection*{How does network mean persistence vary with motif profiles?}
 
             To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
+            We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
             In case high or low mean persistences are associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable).
-            We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
-            All statistical tests were conducted as described above.
+            All statistical tests were conducted as when relating motif profiles to global network structure.
 
@@
 
-        \subsubsection*{How does motif participation vary with degree and trophic level (i.e., do simpler measures completely explain motif results)?}
+        \subsubsection*{How does motif participation vary with degree and trophic level?}
 
@@
 
-    \subsection*{Mean persistence, disturbance, and global network properties}
+    \subsection*{Network mean persistence and global network properties}
 
-        The mean probability of persistence for all consumers in a network naturally decreased with increasing probability of extinction of basal species.
-These effects were clear across the ranges of network size and connectance we considered.
-        Network size and connectance also affect network mean persistence, but to a much smaller extent than increasing disturbance (Figure S1, Table S1, \emph{Appendix SA}).
-Persistences generally decrease when networks are more connected, but the effect of network size varies depending on level of disturbance.
-At low disturbance, small and highly-connected webs show the lowest persistence, whereas at high disturbance, large and highly connected webs show the lowest persistence (Fig.
+        Increasing the probability of extinction for basal species necessarily decreased the mean probability of persistence for consumers. 
+        This effect was clear across the ranges of network size and connectance we considered. 
+        Network size and connectance also affected network mean persistence, but to a much smaller extent than increasing disturbance (Figure S1, Table S1, \emph{Appendix SA}). 
+        Persistences generally decreased in more-connected networks, but the effect of network size varied depending on level of disturbance. 
+        At low disturbance, small and highly-connected webs showed the lowest mean persistence, whereas at high disturbance, large and highly connected webs showed the lowest mean persistence (Fig.
 S2, \emph{Appendix SA}).
@@
 
-        Connectance was significantly related to the frequencies of different motifs in a network.
+        Although the PERMANOVA test relating overall motif profiles to network properties was not reliable (Fig. S3, \emph{Appendix SB}), 
+        connectance was significantly related to the frequencies of different motifs in a network when considering motifs individually.
         Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).
@@
 
-    \subsection*{Network motif profiles and network persistence}
+    \subsection*{Network motif profiles and network mean persistence}
 
-        The motif profile of a network is related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
+        The overall motif profile of a network is related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
 Average persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
@@
             \centering
-            \includegraphics[height=.5\textheight]{figures/persistence_motif_profiles.eps}
+            \includegraphics[width=.5\textwidth]{figures/persistence_motif_profiles.eps}
             \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.

### 2021-08-12 code/stat_analysis/persistence_vs_profiles.py
further legibility tweaks

@@ -74,28 +74,6 @@ def read_datafile(datafile):
   f.close()
   return netprops
 
-def binner(netprops,binby):
-  if binby=='omnivory':
-    bindict={'chain':[],'dircomp':[],'appcomp':[]}
-    step=.01
-    minni=0
-    for i in range(0,100):
-      binmin=minni+step*i
-      binmed=minni+step*i+step/2
-      binmax=minni+step*(i+1)
-
-      for motif in bindict:
-        motpoints=[]
-        for ID in netprops:
-          if netprops[ID]['omnivory']>=binmin and netprops[ID]['omnivory']<binmax:
-            motpoints.append(netprops[ID][motif])
-        if motpoints!=[]:
-          bindict[motif].append((binmed,np.mean(motpoints),np.std(motpoints)/len(motpoints)))
-  else:
-    bindict={}
-
-  return bindict
-
 def format_graph(graph,simple):
   graph.yaxis.bar.linewidth=1
   graph.xaxis.bar.linewidth=1
@@ -122,13 +100,13 @@ def format_graph(graph,simple):
     graph.xaxis.label.text='Proportion of network profile'
 
 
-  graph.xaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
-  graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
-  graph.xaxis.label.configure(char_size=1,just=2,place='normal')
+  graph.xaxis.ticklabel.configure(format='decimal',prec=1,char_size=1)
+  graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=1)
+  graph.xaxis.label.configure(char_size=1.5,just=2,place='normal')
 
-  graph.yaxis.label.configure(char_size=1,just=2,place='normal')
+  graph.yaxis.label.configure(char_size=1.5,just=2,place='normal')
   graph.xaxis.tick.configure(onoff='on',minor_ticks=0,major_size=.4,place='both',major_linewidth=.5,minor_linewidth=1)
-  graph.yaxis.tick.configure(onoff='on',minor_ticks=1,major_size=.4,minor_size=.5,place='both',major_linewidth=.5,minor_linewidth=1)
+  graph.yaxis.tick.configure(onoff='on',minor_ticks=0,major_size=.4,minor_size=.5,place='both',major_linewidth=.5,minor_linewidth=1)
 
 
   return graph
[…]

### 2021-08-12 code/stat_analysis/persistence_vs_profiles.py
further legibility tweaks

@@ -94,7 +94,7 @@ def format_graph(graph,simple):
     graph.world.ymax=.9
     graph.world.xmax=.8
     graph.xaxis.tick.major=.2
-    graph.yaxis.tick.major=.1
+    graph.yaxis.tick.major=.2
     graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
     graph.yaxis.label.text='Mean persistence'
     graph.xaxis.label.text='Proportion of network profile'
@@ -106,7 +106,7 @@ def format_graph(graph,simple):
 
   graph.yaxis.label.configure(char_size=1.5,just=2,place='normal')
   graph.xaxis.tick.configure(onoff='on',minor_ticks=0,major_size=.4,place='both',major_linewidth=.5,minor_linewidth=1)
-  graph.yaxis.tick.configure(onoff='on',minor_ticks=0,major_size=.4,minor_size=.5,place='both',major_linewidth=.5,minor_linewidth=1)
+  graph.yaxis.tick.configure(onoff='on',minor_ticks=0,major_size=.4,place='both',major_linewidth=.5,minor_linewidth=1)
 
 
   return graph

### 2021-08-13 manuscript/motifs_vs_degree.tex

@@
 
-            A species' motif participation is likely related to its degrees and trophic level since all three are measures of network position.
-            To understand how this may have affected the results above, we first fit two linear mixed-effects regressions relating persistence to degree or trophic level, $\pi_{disturbed}$, and their interaction.
+            A species' motif participation is likely related to its in-degree and trophic level since all three properties are measures of network position.
+            To provide context for the results above, we therefore fit two linear mixed-effects regressions relating persistence to degree or trophic level, $\pi_{disturbed}$, and their interaction.
             As above, we centered and scaled the predictors.
             We also included a random effect of the interaction between network size and connectance in order to account for the differences in persistence between different networks.
-            These regressions provide context about how persistence varies with simpler measures of species' roles within networks.
 
@@
 
-        % \subsubsection*{How does persistence vary with motif positions?}
-
-        %     As one more level of detail, we can consider the frequency with which a species appears in each \emph{position} in each motif. 
-        %     Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.
-        %     The full set of these positions, across all motifs, is the species' \emph{motif role}~\citep{Stouffer2010,Cirtwill2017}.
-        %     As with motif participation, we normalized motif roles by dividing the number of times a species appeared in each position by the total number of times it appeared in all positions~\citep{Baker2015,Cirtwill2017}.
-
-        %     [[Could add a PERMANOVA to show that species' roles overal affect their persistence. Haven't done so yet.]]
-
-        %     To test whether participating in different motif positions was related to a species' probability of persistence, we fit similar models to those for motif participation.
-        %     For each position, we fit a linear mixed-effect model relating persistence to the frequency of the position in a species' motif role, disturbance, and their interaction.
-        %     To account for differences in persistence across networks of different size and connectance, we included a random effect of the interaction between size and connectance.
-        %     All regressions were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-
-        % \subsubsection*{How do motif roles vary with other network properties?}
-
-        %     As with network motif profiles and motif participation, species' motif roles may vary with network size and connectance and with the species' in degree and trophic level.
-        %     To identify these relationships, we fit two linear regressions for each motif position.
-        %     The first related the frequency of the position in species' roles to network size, connectance, and their interaction.
-        %     The second related the frequency of the position to in-degree, STL, and their interaction.
-        %     All regressions were fit using the R~\citep{R} base function `lm'.
+    % % % Things in SI that are not referred to:
+    % persistence vs. motif positions
+    % motif roles (positions) vs. other network properties
+    % persistence vs. % Basal
+    % motif profiles vs. % Basal
 
@@
 
-    We are interested in i) how the distribution of motifs in a network (i.e., the network's \emph{motif profile}) affect network persistence, and ii) how a species motif participation affects its persistence.
-We analyse this for different levels of disturbance to the basal level.
-To provide context, we also investigate relationships between persistence and classic global (size and connectance) and local (in-degree and trophic level) network properties.
-    % The results are organized as follows: 1) How network persistence varies with network motif distribution.
-2) How network persistence varies with global structural properties.
-3) How network motif distribution varies with global structural properties.
-4) How species probability of persistence varies with its motif profile.
-5) How species persistence varies with its characteristics such as TL and in-degree.
-6) How a species motif profile depend on its characteristics.
-7) How a species persistence varies with its motif role (position in respective motif).
+    We were interested in i) how the distribution of motifs in a network (i.e., the network's \emph{motif profile}) affect network persistence, and ii) how a species motif participation affects its persistence.
+    We analysed this for different levels of disturbance to the basal level. 
+    To provide context, we also investigated relationships between persistence and classic global (size and connectance) and local (in-degree and trophic level) network properties.
 
@@
             \includegraphics[width=.75\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
-            \caption{The proportions of a network's motif profile made up by the four motifs which can appear in acyclic networks vary with the size and connectance of the network.
-More-connected networks tend to contain relatively more omnivory motifs and fewer apparent competition motifs.
+            \caption{The proportions of a network's motif profile made up by the four motifs which can appear in acyclic networks varied with the size and connectance of the network.
+More-connected networks tended to contain relatively more omnivory motifs and fewer apparent competition motifs.
 The proportions of direct competition and three-species chain motifs were less variable.}
@@
 
-    \clearpage
-
     \subsection*{Network motif profiles and network mean persistence}
 
-        The overall motif profile of a network is related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
-Average persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
-Conversely, average persistence is lower in networks with higher proportions of the omnivory motif.
-These trends hold for all levels of disturbance (Fig.~\ref{motif_profile_persistence}, Table S3), \emph{Appendix SC}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
+        The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001). 
+        Average persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}). 
+        Conversely, average persistence was lower in networks with higher proportions of the omnivory motif. 
+        These trends were consistent across  levels of disturbance (Fig.~\ref{motif_profile_persistence}, Table S3), \emph{Appendix SC}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
 For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).
@@
             \includegraphics[width=.5\textwidth]{figures/persistence_motif_profiles.eps}
-            \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.
-Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportions of the other three motifs.
+            \caption{The proportions of the four motifs in a network's motif profile were related to the mean persistence of species in the network.
+Specifically, persistence decreased as the proportion of omnivory in the network's motif profile increased while persistence increased with increasing proportions of the other three motifs.
 Interactions between motif profiles and disturbance were significant but small.
@@
 
-       The effects of motif participation on a species' probability of persistence depends on the level of disturbance.
-When disturbance is low, species' persistence increases with increasing proportions of the omnivory and direct competition motifs.
-Conversely, a species' persistence decreases with an increasing proportion of the apparent competition motif.
-Persistence is not significantly related to the proportion of the three-species chain motif (Figure~\ref{fig:prop_lmer_all}).
+       The relationship between frequency of participation in a motif and a species' probability of persistence depended on the level of disturbance. 
+       When disturbance was low, species' persistence increased with increasing proportions of the omnivory and direct competition motifs. 
+       Conversely, persistence decreased with an increasing proportion of the apparent competition motif. 
+       Persistence was not significantly related to the proportion of the three-species chain motif (Figure~\ref{fig:prop_lmer_all}).
 
-        At high levels of disturbance, a species' persistence decreases as the proportion of the three-species chain increases, but persistence increases as the proportion of any other motif increases (Figure~\ref{fig:prop_lmer_all}).
-This means that increased disturbance strengthens the effects of the three-species chain motif and changes the direction of the effect of the apparent competition motif  (Table S4, \emph{Appendix SD}).
-The effect of omnivory on persistence is still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.
+        At high levels of disturbance, a species' persistence decreased as the proportion of the three-species chain increased, but persistence increased as the proportion of any other motif increased (Figure~\ref{fig:prop_lmer_all}).
+        This means that increased disturbance strengthened the effects of the three-species chain motif and changeed the direction of the effect of the apparent competition motif  (Table S4, \emph{Appendix SD}).
+        The relationship between omnivory and persistence was still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.
 
@@
                 \centering
-                \includegraphics[width=0.85\textwidth]{figures/prop_lmer_allCS.pdf}
+                \includegraphics[width=\textwidth]{figures/prop_lmer_allCS.pdf}
                 \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).
-The effect of participating in each motif is based on linear mixed-effect models following Equation~\ref{propreq}, except that models for each  disturbance level are fitted separately.
+The effect of participating in each motif was based on linear mixed-effect models following Equation~\ref{propreq}, except that models for each  disturbance level are fitted separately.
 The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance).
@@
 
-        \subsubsection*{Consistency in effects of motif participation on species persistence}
+        \subsubsection*{Consistency of  relationships across networks}
 
-            % In order to evaluate how consistent the relationships between a species persistence and motif participation is, and how it varies with  level of disturbance and global network structure, we fit linear regressions of species persistence against motif proportions for the species for each network separately.
-We then analyze the density distribution of the values of the slopes.
-Network size had a much smaller effect than connectance (Fig.
-S3; \emph{Appendix SE}); we therefore focus on connectance here.
-
-            When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
-However, there is an interaction between disturbance and connectance for the direct competition and omnivory motifs; At low and medium connectances there is little change in the proportion of positive slopes when the proportion of  direct competition motif is high, independent of level of disturbance.
-At high connectances, however, the proportion of positive slopes is lower and also decreases with increasing disturbance.
-For the omnivory motif the trend is reversed.
-Additionally, the omnivory motif stands out having a wider curve when network connectance is low.
+            When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally had more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
+            However, this trend varied at different levels of disturbance, especially for the direct competition and omnivory motifs.
+            At low and medium connectances there was little change in the distribution of slopes for direct competition across all levels of disturbance. 
+            At high connectances, however, the proportion of positive slopes was both lower overall and decreased with increasing disturbance. 
+            For the omnivory motif these trends were reversed. 
+            Moreover, the omnivory motif has a particularly broad distribution of slopes (i.e., an inconsistent relationship to persistence) when network connectance was low. 
             Network size had a much smaller effect than connectance (Fig. S3; \emph{Appendix SE}).
@@
             \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.pdf}
-            \caption{Density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
+            \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
 Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
@@
 
-        A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
-Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S5, \emph{Appendix SF}).
-        In-degree shows an even stronger interaction with disturbance; having more prey species is associated with higher persistence without disturbance, but lower persistence at high levels of disturbance.
-
-        % Anna notes (just for our (mine) understanding): 
-        % When looking at the two figures of degree vs persistence and STL vs persistence - my conclusion is that species at lower trophic levels have the highest in-degree.
-Double check - yes, that’s true.
-        % Lower TL are all well at low disturbances, and when disturbance increase, persistence decrease (looking at fig degree vs persistence, right end of the line).
-        % Higher TL experience the same thing, although to a slightly smaller extent (looking at the left end of the lines in fig degree vs persistence and right end of lines in fig STL vs persistence).
-        % The slope of the line in fig degree vs persistence is not really relevant, since it's kind of based on the response of different TL level (which makes perfect sense in the BN framework)
-        % weird thing though - at high levels of disturbance, lower TL seems more affected than higher TL in the persistence vs degree figure (since low degree belongs to higher TL, high degree to lower TL).
-This does not correspond with fig persistence vs STL.
-HOWEVER - not only high TL have a low degree, lower TL do as well, probably affecting the overall result.
-Only low TL have high degree - but the opposite is not true.
-Probably explains it
-
-    \clearpage
+        Species persistence varied with both in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}). 
+        Increased trophic level was generally associated with decreasing species persistence, although the rate of decline was slower at high disturbance (Table S5, \emph{Appendix SF}).
+        In-degree showed an even stronger interaction with disturbance; having more prey species was associated with higher persistence without disturbance, but lower persistence at high levels of disturbance.
 
@@
 
-            \begin{figure}
+            \begin{figure}[h!]
                 \centering
-                \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
+                \includegraphics[width=.75\textwidth]{figures/roles_vs_TL.eps}
                 \caption{As with motifs, persistence varied with simpler measures of species roles in networks.
@@
             \end{figure}
-
-    % \subsection*{Persistence and species' positioning in motifs}
-
-    %     A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance (Table SX, \emph{Appendix SH}).
-    %     In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
-    %     Persistence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
-    %     Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.
-
-    %     In the three-species chain, persistence decreased as the frequency of the top position increased at all levels of disturbance.
-    %     Conversely, persistence increased with increasing frequency of the middle position at all levels of disturbance.
-    %     Persistence decreased slightly with increasing frequency of the bottom position at low levels of disturbance but increased at high levels of disturbance.
-
-    %     In the apparent competition motif, persistence decreased slightly with increasing frequencies of the bottom position (one predator, no prey within the motif) at low disturbance but increased slightly at high disturbance.
-    %     Persistence increased with increasing frequency of the top position (two prey, no predator within the motif) at low disturbance but decreased at high disturbance.
-
-    %     In the direct competition motif, persistence varied little with the frequency of the bottom position (two predators, no prey within the motif) at low disturbance but increased with the frequency of this position at high levels of disturbance.
-    %     Persistence increased slightly with the frequency of the top position (one prey, no predators within the motif) at low levels of disturbance and decreased slightly at high levels of disturbance.
-
-    % \begin{figure}[h!]
-    %     \includegraphics[width=.95\textwidth]{figures/persistence_positions_Omnivory.eps}
-    %     \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
-Here we show the relationships between persistence and the proportions of the three positions in the omnivory motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
-Line style indicates the motif  position.
-}
-    %     \label{figs:persistence_position_motifs}
-    %     \end{figure}
-
-    %     The frequencies of motif positions in a species' role varied with network properties (\emph{Appendix SI}).Motif roles had stronger relationships with degree and STL than with network size or connectance.
 

### 2021-08-18 manuscript/reorder_main_text.tex
a revised order for the main text

--- 2021-08-12 manuscript/motifs_vs_degree.tex
@@
 
-        \subsubsection*{How does network mean persistence vary depending on network properties?}
-
-            Both species richness (S) and connectance (C) are known to influence persistence in both simulation studies and empirical food webs.
-            In addition, increasing the probability of extinction for basal species ($\pi_{disturbed}$ above) is intended to decrease persistence of species at higher trophic levels.
-            Although these effects are not the main focus of our manuscript, we must establish these background effects before measuring the effects of meso-scale structure on persistence.
-            The persistence of basal species in our simulations is determined only by their baseline probability, being either $\pi_{base} = 0.1$ or higher, dependent on $\pi_{disturbed}$.
-            We therefore considered only non-basal species (i.e., those with at least one prey) throughout this manuscript.
-
-            To test for effects of global network properties, we fit a linear regression relating consumer species' probabilities of persistence to network size, connectance, the level of disturbance to basal resources, and all interactions between them.
-            All predictors were centered and scaled before fitting the model. 
-            As the three-way interaction was significant, we did not simplify the model.
-            All models were fit using the R~\citep{R} base function `lm'.
-
-        \subsubsection*{How do network motif profiles vary with global structure?}
-
-            As well as affecting persistence directly, global network structure (i.e., size and connectance) may affect the meso-scale structure of the network.
-            This is especially likely since we include only networks which can retain all initial species after 1000 rounds of simulated Lotka-Volterra population dynamics~\citep{Cirtwill2021_inprep}:  some meso-scale structures might be more likely to allow all species to persist than others in networks with different size and connectance constraints.
-            We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
-            Significance was calculated based on 9999 permutations.
-            PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
-            To test whether this applies in our case, we fit an ANOVA of the dispersion of motif profiles about the centroid for each combination of network size and connectance. 
-            We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
-            To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif to network size, connectance, and their interaction.
-            We fit each regression using the R~\citep{R} base function `lm'.
-
-        \subsubsection*{How does network mean persistence vary with motif profiles?}
-
-            To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
-            We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
-            In case high or low mean persistences are associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable).
-            All statistical tests were conducted as when relating motif profiles to global network structure.
-
-            To supplement these overall tests, we also fit four regressions testing whether persistence had a linear relationship to the proportion of a network profile made up of each motif, and whether this varied with the level of disturbance.
-            Each regression included the proportion of a motif, disturbance, and their interaction as predictors.
-            All regressions were fit using the R~\citep{R} base function `lm'.
-
         \subsubsection*{How does species persistence vary with motif participation?}
@@
 
-            We were also interested in whether these general trends were consistent across networks with different global properties and different levels of disturbance. 
+        \subsubsection*{How does motif participation vary with degree and trophic level?}
+
+            A further complication when trying to understand whether relationships between motifs and probability of extinction depend upon underlying relationships between degree or trophic level and extinction
+            We also fit linear mixed-effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+            We performed similar analyses to test whether a species' motif role is related it in-degree or trophic level (\emph{Appendix SI}).
+
+    	\subsubsection*{How consistent are motif-persistence relationships?}
+
+    		As well as identifying general relationships between motif participation and persistence, we were also interested in whether these general trends were consistent across networks with different global properties and different levels of disturbance.
             To test this, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources.
@@
             We performed similar analyses to test whether appearing in different positions in each motif (i.e., the motif role) is related to a species' persistence (\emph{Appendix SH}).
+
+        \subsubsection*{How does network mean persistence vary with motif profiles?}
+
+            To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
+            We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
+            In case high or low mean persistences are associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable).
+            All statistical tests were conducted as when relating motif profiles to global network structure.
+
+            To supplement these overall tests, we also fit four regressions testing whether persistence had a linear relationship to the proportion of a network profile made up of each motif, and whether this varied with the level of disturbance.
+            Each regression included the proportion of a motif, disturbance, and their interaction as predictors.
+            All regressions were fit using the R~\citep{R} base function `lm'.
+
+        \subsubsection*{How do network motif profiles vary with global structure?}
+
+            As well as affecting persistence directly, global network structure (i.e., size and connectance) may affect the meso-scale structure of the network.
+            This is especially likely since we include only networks which can retain all initial species after 1000 rounds of simulated Lotka-Volterra population dynamics~\citep{Cirtwill2021_inprep}:  some meso-scale structures might be more likely to allow all species to persist than others in networks with different size and connectance constraints.
+            We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
+            Significance was calculated based on 9999 permutations.
+            PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
+            To test whether this applies in our case, we fit an ANOVA of the dispersion of motif profiles about the centroid for each combination of network size and connectance. 
+            We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
+            To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif to network size, connectance, and their interaction.
+            We fit each regression using the R~\citep{R} base function `lm'.
+
+        \subsubsection*{How does network mean persistence vary depending on network properties?}
+
+            Both species richness (S) and connectance (C) are known to influence persistence in both simulation studies and empirical food webs.
+            In addition, increasing the probability of extinction for basal species ($\pi_{disturbed}$ above) is intended to decrease persistence of species at higher trophic levels.
+            Although these effects are not the main focus of our manuscript, we must establish these background effects before measuring the effects of meso-scale structure on persistence.
+            The persistence of basal species in our simulations is determined only by their baseline probability, being either $\pi_{base} = 0.1$ or higher, dependent on $\pi_{disturbed}$.
+            We therefore considered only non-basal species (i.e., those with at least one prey) throughout this manuscript.
+
+            To test for effects of global network properties, we fit a linear regression relating consumer species' probabilities of persistence to network size, connectance, the level of disturbance to basal resources, and all interactions between them.
+            All predictors were centered and scaled before fitting the model. 
+            As the three-way interaction was significant, we did not simplify the model.
+            All models were fit using the R~\citep{R} base function `lm'.
 
@@
 
-        \subsubsection*{How does motif participation vary with degree and trophic level?}
-
-            A further complication when trying to understand whether relationships between motifs and probability of extinction depend upon underlying relationships between degree or trophic level and extinction
-            We also fit linear mixed-effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-            We performed similar analyses to test whether a species' motif role is related it in-degree or trophic level (\emph{Appendix SI}).
-
-        % \subsubsection*{How does persistence vary with motif positions?}
-
-        %     As one more level of detail, we can consider the frequency with which a species appears in each \emph{position} in each motif. 
-        %     Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.
-        %     The full set of these positions, across all motifs, is the species' \emph{motif role}~\citep{Stouffer2010,Cirtwill2017}.
-        %     As with motif participation, we normalized motif roles by dividing the number of times a species appeared in each position by the total number of times it appeared in all positions~\citep{Baker2015,Cirtwill2017}.
-
-        %     [[Could add a PERMANOVA to show that species' roles overal affect their persistence. Haven't done so yet.]]
-
-        %     To test whether participating in different motif positions was related to a species' probability of persistence, we fit similar models to those for motif participation.
-        %     For each position, we fit a linear mixed-effect model relating persistence to the frequency of the position in a species' motif role, disturbance, and their interaction.
-        %     To account for differences in persistence across networks of different size and connectance, we included a random effect of the interaction between size and connectance.
-        %     All regressions were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-
-        % \subsubsection*{How do motif roles vary with other network properties?}
-
-        %     As with network motif profiles and motif participation, species' motif roles may vary with network size and connectance and with the species' in degree and trophic level.
-        %     To identify these relationships, we fit two linear regressions for each motif position.
-        %     The first related the frequency of the position in species' roles to network size, connectance, and their interaction.
-        %     The second related the frequency of the position to in-degree, STL, and their interaction.
-        %     All regressions were fit using the R~\citep{R} base function `lm'.
-
 \section*{Results}
 
-    We are interested in i) how the distribution of motifs in a network (i.e., the network's \emph{motif profile}) affect network persistence, and ii) how a species motif participation affects its persistence.
+    We are interested in how a species' participation in different motifs affects its persistence, and how these relationships relate to other measures of network structure. 
     We analyse this for different levels of disturbance to the basal level. 
@@
 7) How a species persistence varies with its motif role (position in respective motif).
-
-    \subsection*{Network mean persistence and global network properties}
-
-        Increasing the probability of extinction for basal species necessarily decreased the mean probability of persistence for consumers. 
-        This effect was clear across the ranges of network size and connectance we considered. 
-        Network size and connectance also affected network mean persistence, but to a much smaller extent than increasing disturbance (Figure S1, Table S1, \emph{Appendix SA}). 
-        Persistences generally decreased in more-connected networks, but the effect of network size varied depending on level of disturbance. 
-        At low disturbance, small and highly-connected webs showed the lowest mean persistence, whereas at high disturbance, large and highly connected webs showed the lowest mean persistence (Fig.
-S2, \emph{Appendix SA}).
-
-    \subsection*{Network motif profiles and global network properties}
-
-        Although the PERMANOVA test relating overall motif profiles to network properties was not reliable (Fig. S3, \emph{Appendix SB}), 
-        connectance was significantly related to the frequencies of different motifs in a network when considering motifs individually.
-        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).
-        Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}).
-
-        \begin{figure}[h!]
-            \centering
-            \includegraphics[width=.75\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
-            \caption{The proportions of a network's motif profile made up by the four motifs which can appear in acyclic networks vary with the size and connectance of the network.
-More-connected networks tend to contain relatively more omnivory motifs and fewer apparent competition motifs.
-The proportions of direct competition and three-species chain motifs were less variable.}
-            \label{motif_proportion_lms}
-        \end{figure}
-
-    \clearpage
-
-    \subsection*{Network motif profiles and network mean persistence}
-
-        The overall motif profile of a network is related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
-Average persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
-Conversely, average persistence is lower in networks with higher proportions of the omnivory motif.
-These trends hold for all levels of disturbance (Fig.~\ref{motif_profile_persistence}, Table S3), \emph{Appendix SC}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
-For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).
-
-        \begin{figure}
-            \centering
-            \includegraphics[width=.5\textwidth]{figures/persistence_motif_profiles.eps}
-            \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.
-Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportions of the other three motifs.
-Interactions between motif profiles and disturbance were significant but small.
-Here we show the relationships of probabilities of extinction of basal resources of 0.1 (top set of lines) and 0.5 (bottom set of lines).
-Relationships are shown for the observed range of proportions for each motif.}
-            \label{fig:motif_profile_persistence}
-        \end{figure}
 
@@
 
-        \clearpage
-
-        \subsubsection*{Consistency in effects of motif participation on species persistence}
+    \subsection*{Consistency in effects of motif participation on species persistence}
 
@@
 
-    \subsection*{Species characteristics and persistence}
-
-        A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
-Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S5, \emph{Appendix SF}).
-        In-degree shows an even stronger interaction with disturbance; having more prey species is associated with higher persistence without disturbance, but lower persistence at high levels of disturbance.
-
-        % Anna notes (just for our (mine) understanding): 
-        % When looking at the two figures of degree vs persistence and STL vs persistence - my conclusion is that species at lower trophic levels have the highest in-degree.
-Double check - yes, that’s true.
-        % Lower TL are all well at low disturbances, and when disturbance increase, persistence decrease (looking at fig degree vs persistence, right end of the line).
-        % Higher TL experience the same thing, although to a slightly smaller extent (looking at the left end of the lines in fig degree vs persistence and right end of lines in fig STL vs persistence).
-        % The slope of the line in fig degree vs persistence is not really relevant, since it's kind of based on the response of different TL level (which makes perfect sense in the BN framework)
-        % weird thing though - at high levels of disturbance, lower TL seems more affected than higher TL in the persistence vs degree figure (since low degree belongs to higher TL, high degree to lower TL).
-This does not correspond with fig persistence vs STL.
-HOWEVER - not only high TL have a low degree, lower TL do as well, probably affecting the overall result.
-Only low TL have high degree - but the opposite is not true.
-Probably explains it
-
-    \clearpage
-
     \subsection*{Motif participation, network properties, and species characteristics}
@@
             \end{figure}
+
+    \subsection*{Network motif profiles and network mean persistence}
+
+        The overall motif profile of a network is related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
+Average persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
+Conversely, average persistence is lower in networks with higher proportions of the omnivory motif.
+These trends hold for all levels of disturbance (Fig.~\ref{motif_profile_persistence}, Table S3), \emph{Appendix SC}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
+For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).
+
+        \begin{figure}
+            \centering
+            \includegraphics[width=.5\textwidth]{figures/persistence_motif_profiles.eps}
+            \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.
+Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportions of the other three motifs.
+Interactions between motif profiles and disturbance were significant but small.
+Here we show the relationships of probabilities of extinction of basal resources of 0.1 (top set of lines) and 0.5 (bottom set of lines).
+Relationships are shown for the observed range of proportions for each motif.}
+            \label{fig:motif_profile_persistence}
+        \end{figure}
+
+    \subsection*{Network motif profiles and global network properties}
+
+        Although the PERMANOVA test relating overall motif profiles to network properties was not reliable (Fig. S3, \emph{Appendix SB}), 
+        connectance was significantly related to the frequencies of different motifs in a network when considering motifs individually.
+        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).
+        Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}).
+
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=.75\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
+            \caption{The proportions of a network's motif profile made up by the four motifs which can appear in acyclic networks vary with the size and connectance of the network.
+More-connected networks tend to contain relatively more omnivory motifs and fewer apparent competition motifs.
+The proportions of direct competition and three-species chain motifs were less variable.}
+            \label{motif_proportion_lms}
+        \end{figure}
+
+     \clearpage
+
+\section*{Results to SI entirely?}
+
+    \subsection*{Network mean persistence and global network properties}
+
+        Increasing the probability of extinction for basal species necessarily decreased the mean probability of persistence for consumers. 
+        This effect was clear across the ranges of network size and connectance we considered. 
+        Network size and connectance also affected network mean persistence, but to a much smaller extent than increasing disturbance (Figure S1, Table S1, \emph{Appendix SA}). 
+        Persistences generally decreased in more-connected networks, but the effect of network size varied depending on level of disturbance. 
+        At low disturbance, small and highly-connected webs showed the lowest mean persistence, whereas at high disturbance, large and highly connected webs showed the lowest mean persistence (Fig.
+S2, \emph{Appendix SA}).
+
+    \subsection*{Species characteristics and persistence}
+
+        A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
+Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S5, \emph{Appendix SF}).
+        In-degree shows an even stronger interaction with disturbance; having more prey species is associated with higher persistence without disturbance, but lower persistence at high levels of disturbance.
+
+        % Anna notes (just for our (mine) understanding): 
+        % When looking at the two figures of degree vs persistence and STL vs persistence - my conclusion is that species at lower trophic levels have the highest in-degree.
+Double check - yes, that’s true.
+        % Lower TL are all well at low disturbances, and when disturbance increase, persistence decrease (looking at fig degree vs persistence, right end of the line).
+        % Higher TL experience the same thing, although to a slightly smaller extent (looking at the left end of the lines in fig degree vs persistence and right end of lines in fig STL vs persistence).
+        % The slope of the line in fig degree vs persistence is not really relevant, since it's kind of based on the response of different TL level (which makes perfect sense in the BN framework)
+        % weird thing though - at high levels of disturbance, lower TL seems more affected than higher TL in the persistence vs degree figure (since low degree belongs to higher TL, high degree to lower TL).
+This does not correspond with fig persistence vs STL.
+HOWEVER - not only high TL have a low degree, lower TL do as well, probably affecting the overall result.
+Only low TL have high degree - but the opposite is not true.
+Probably explains it
+
+    \clearpage
 
@@
 
-% Results summary - network level (following Anna)
+% Framing: network structure is important for predicting extinction risk
 Changes in the basal level in food webs have profound effects on all other species in the network \citep{}, as all consumers depend directly or indirectly on primary producers for their survival \citep{}.
@@
 
+% Framing: motifs are related to network stability, species persistence (our results)
 Motifs are the basic building blocks of networks~\citep{Milo2002}, connecting network-wide (global) and species-specific (local) properties, describing both direct and indirect effects. 
-As such, the distribution of these meso-scale structures can be of importance to network and species persistence~\citep{}.
-We find that both the distribution of motifs at the network level and the set of motifs a target species participates in are related to persistence after disturbance to the plant community.
-Also, the nature of these relationships synthesizes the effects of other local and global-scale properties.
+As such, the distribution of these meso-scale structures within a network can be of importance to network and species persistence~\citep{}.
+Moreover, we find that a species' frequency of participation in different motifs is related to its participation in different motifs.
 
-A network's motif profile had a relatively strong relationship to the network mean persistence compared to global structures.
+% Species-level results summary
+Specifically, a species' probability of persistence increased with its frequency of participation in the direct competition and omnivory motifs and decreased with its frequency of participation in the apparent competition motif.
+The relationship between persistence and the frequency of participation in the apparent competition motif depended strongly on the level of disturbance.
+These trends can be partially explained by underlying relationships between motif participation and in-degree and/or trophic level.
+For example, the frequency of participation in the direct competition motif was strongly and negatively correlated with trophic level.
+Accordingly, while a high trophic level is associated with lower persistence at all levels of disturbance, the opposite is true for participation in the direct competition motif.
+
+% Degree and TL are both important for omnivory
+The relationship between participation in the omnivory motif and persistence sythesizes relationships between participation and both in-degree and trophic level.
+Species participation in the omnivory motif was associated with greater persistence, especially at low levels of disturbance.
+This is consistent with a strong negative correlation between omnivory and trophic level (leading to the overall positive effect of omnivory) and a strong positive correlation with in-degree (leading to the reduced benefit of omnivory at high levels of disturbance).
+Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between them.
+
+% Degree and TL don't explain everything
+The relationships between persistence and participation in the apparent competition and three-species chain motifs also demonstrate motif roles' ability to syntesize different aspects of network structure.
+In these cases, in-degree and trophic level are not sufficient to explain the trends we observe in the motifs.
+Participation in the apparent competition motif was strongly negatively correlated with in-degree and shows the opposite relationship to persistence.
+However, participation in the apparent competition motif was also strongly and positively correlated with trophic level (which is negatively associated with persistence since predators must have lower probability of persistence than their prey~\citep{Eklof2013}) and this correlation is not reflected in the relationship between participation in the apparent competition motif and persistence.
+Similarly, participation in the three-species chain motif was significantly correlated with both in-degree and trophic level, but the relationship between the motif and persistence did not reflect the relationship between in-degree and persistence.
+This suggests the influence of some other aspect of network structure on species' persistence.
+As motifs capture information about direct and indirect interactions and include a larger section of the network than degree or trophic level,~\citep{} it is likely that effects of these indirect interactions may be included in the relationships we observe between motifs and persistence.
+
+% Moving to network context
+As well as clarifying the relationships between network structure and the effect of bottom-up disturbance on particular species, motifs were also related to the mean probability of persistence across all species in a network.
+In fact, a network's motif profile had a stronger relationship to mean persistence than network size or connectance. 
 However, connectance did affect which network motifs were most common. 
@@
 
-% Network-level persistence
-In general, average network persistence was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif. 
-%The negative relationship between omnivory and network persistence is consistent with other studies showing that an omnivory motif is less likely to be stable than the other motifs considered~\citep{Borrelli2015a}.
-Motif profiles with high apparent competition and low omnivory were more likely in less-connected food webs, which generally had higher mean persistence than highly-connected food webs.
-Despite these general trends, networks with more variable motif profiles had higher mean persistence, suggesting that presence of a variety of different network structures support persistence.
+% Details of network-level results
+In general, average network persistence was highest in networks with low frequencies of the omnivory motif and high frequencies of the apparent competition motif.
+These networks also tended to have lower connectance, but the relationship between connectance and mean persistence was weaker than that for motifs.
+Both omnivory and apparent competition have been shown to have consequences for the stability of ecological networks.
 
-A network's motif profile ...
-Both omnivory and apparent competition have been shown to have consequences for the stability of ecological networks.
-However, particularly the role of omnivory is debated and it has been shown to act both stabilizing and destabilizing.
-[Add examples of both ...
-]
-
-Moreover, species within a network can have substantially different probabilities of persistence than the mean across all species.
-For targeted conservation actions this type of information is often more relevant.
-
-% Species level persistence
-While the persistence of a particular species is affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it also depends on local-scale (trophic level and in-degree) network properties.
-Local properties may be especially important for the apparent competition and direct competition motifs, due to the indirect effects not being fully expressed in these two motifs since the Bayesian Network framework does not include top-down effects.
-
-The effect of the local-scale network properties in-degree and trophic level differed.
-A high trophic level was associated with lower persistence across all levels of disturbance.
-This was expected, as species at higher trophic levels here by definition have equal or higher probability of extinction compared to their prey ~\citep{Eklof2013}.
-The effect of in-degree is less obvious and depended on the level of disturbance; a high in-degree had a positive effect on persistence at low levels of disturbance but not at high levels of disturbance.
-In addition to these different relationships with persistence, in-degree and trophic level had different relationships to the four motifs we investigated.
-
-Participation in the apparent competition motif was strongly negatively correlated with in-degree --a high proportion of the apparent competition motif implies a low in-degree, and both are associated with lower persistence at low disturbance but increased persistence at high disturbance.
-This suggests that in-degree is in fact the factor driving the relationship between proportion of apparent competition motif and species persistence.
-The turnover from a negative to a positive relationship with an increasing level of disturbance indicate that when disturbance is high, connections to many prey species is negative as the consumer receives the negative impact from several sources.
-A species with a high proportion of the apparent competition motif, which also has a low in-degree, thus gets an advantage when plant level disturbance is high since fewer prey items transmit the disturbance.
-In-degree does not, however, explain all of the trends we observe for motifs as degree and motif participation are not always tightly coupled.
-For example, it is possible for a species with low degree to participate in many motifs if it interacts with high-degree species and thereby obtain many indirect interaction partners.
-This is one of the strengths with analyses of motifs that it considers local structures additionally include indirect connections one link away.
-
-The direct competition motif was not strongly correlated to in-degree, but had a strong negative relationship with trophic level.
-Thus, species with a high proportion of the direct competition motif has commonly a low trophic level, which is associated with higher probability of persistence at all levels of disturbance.
-The clear positive relationship between proportion of direct competition motifs and species persistence is therefore likely due to the tendency for species at low trophic levels to have a higher proportion of that motif.
-
-%The remaining two motifs are strongly related to both in-degree and trophic level and must be explained using a combination of the two simpler network properties.
-%Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between them.
-%The three-species chain motif, for example, is moderately negatively correlated with in-degree and moderately positively correlated with trophic level. 
-%The relationship between persistence and the chain motif therefore combines the overall negative effect of high trophic level with the interaction between in-degree and disturbance; leading the chain motif to have little relationship to persistence at low disturbance and an increasingly negative relationship at higher disturbance.
-Likewise, the three-species chain motif is correlated with trophic level, but the relationship is positive and a high proportion of this motif indicates a high trophic level.
-Hence, the negative effect on persistence of having a high proportion of the three-species chain motif is likely connected to the high trophic level.
-
-%The omnivory motif likewise integrates effects of in-degree and trophic level.
-The omnivory motif is strongly related to both in-degree and trophic level and must be explained using a combination of the two simpler network properties.
-Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between them
-Species participation in the omnivory motif was associated with greater persistence, especially at low levels of disturbance.
-This is consistent with a strong negative correlation between omnivory and trophic level (leading to the overall positive effect of omnivory) and a strong positive correlation with in-degree (leading to the reduced benefit of omnivory at high levels of disturbance).
-
-Despite participating more often in the omnivory motif being associated with greater persistence for a focal species, having a greater proportion of omnivory motifs in the network was associated with lower network persistence.
-At the network level, a higher frequency of the omnivory motif leads to more pathways by which the disturbances to basal resources can propagate and cause extinctions of consumers.
-At the species level on the other hand, if extinctions of basal resources are rare, these multiple pathways will tend to protect consumers participating in many omnivory motifs by providing alternative channels of energy.
-If extinctions of basal resources are common, however, these multiple pathways cause a negative impact as they will transmit more disturbances to the consumer.
-
-Additionally, the effect of the omnivory motif on a focal species' persistence was closely connected to connectance.
-In highly connected networks, a high proportion of the omnivory motif strongly contributed to higher persistence.
-This was not the case in less connected networks.
-Following the same reasoning as above, in a highly connected network, numerous links pave the way for a disturbance to spread to a higher extent than in a less connected network. 
-Therefore, in a highly connected web where disturbances have more ways to move, the lower level prey item in the omnivory motif may be of even higher importance assisting as a buffer, decreasing the negative effect of disturbance for the consumer species.
-
-% % Might be able to add some of this paragraph to the one above or below (especially the refs).
-% The divergent effect on persistence of either having a high proportion of the three-species chain or omnivory motif can, within our Bayesian Network framework, be explained by the structure of the food web.
-% When participating in a food chain motif---three-species chain or omnivory---the top species of this structure will have a higher susceptibility since disturbances amplify through the network.
-In our results, this is in particular evident for the three-species chain motif at higher levels of disturbance, but this also applies to species with a high proportion of the omnivory motif.
-While the relationship between omnivory and persistence is strongly positive at low bottom-level disturbance, the trend is reduced at high levels of disturbance.
-% This is because species participating to a high extent in omnivory motifs share the sensitivity of species participating to a high extent in three-species chain motif, and are similarly negatively affected by bottom-level disturbances.
-However, the top species in an omnivory motif also have the benefit of an additional prey species, located further down in the food web.
-This prey item can assist as a buffer, decreasing the negative effect of disturbance on the consumer species.
-Clearly, this additional prey is of importance, since it is more beneficial with a higher proportion of the omnivory motif for an individual species than a high proportion of the three-species chain motif.
-% Thus, the observed benefit of a high proportion of the omnivory motif can be compared to omnivores in a network feeding adaptively.
-Studies have demonstrated that adaptive omnivory can increase the permanence of three-species chains \citep{Fagan1997, KrivanDiehl2005, AbramsFung2010}.
-% Although species in our model cannot swap prey dependent on bottom-level disturbances or lower resource densities, a species with a high proportion of omnivory will in a sense spread its risk between prey items at different trophic levels.
-
-The relationship between omnivory and stability is subject to a longstanding debate in ecology.
+% Heading into omnivory
+The relationship between omnivory and stability in particular has long been a subject of debate~\citep{}. 
 Omnivory has been shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, Monteiro2016, Kratina2012}. 
@@
 
- \cite{Cirtwill2021_inprep} analysed how species time to extinction varied depending on their network profile.
+%  % Should go in one of these two paragraphs...
+ % \cite{Cirtwill2021_inprep} analysed how species time to extinction varied depending on their network profile.
 They found that a high proportion of the omnivory motif resulted in a shorter time to extinction, while the other three stable motifs were associated with longer times to extinction.
 However, the study by Cirtwill and Wootton used a different approach using a fully dynamic simulation model also including top-down effects.
-
-Our study of network and species motif profiles show that information on and synthesis of the global and local network structures can provide important information on which conservation actions that will most successfully preserve biodiversity in ecological networks.
-When affected by disturbance of plant community...
-
     %In previous studies of the effect of omnivory motifs, the strength of the omnivorous link can potentially be important.
@@
 
-    % % We can move this to the SI or delete entirely.
-    % Although we performed analysis of both the effect of the absolute counts of motifs and the normalized motif role on persistence, we here focus on the normalized motif role.
-    % This is because, the consistency of the relationships between persistence and the number of times a species appears in any of the possible motifs suggests that the observed effects of non-normalized motif roles are actually due to degree.
-    % Species with higher degrees will with high probability appear in more motifs than species with low degrees.
-    % It is possible for a species with low degree (few direct interaction partners) to participate in many motifs if it interacts with high-degree species (thereby obtaining many indirect interaction partners), but such a situation is fairly unlikely.
-    % Because of this overlap between degree and absolute numbers of motifs, and because degree shows a similar trend to the motif counts [[to be confirmed]], we recommend that researchers consider degree and/or normalized motifs rather than absolute motifs.
-    % When we consider the proportion of a species' role made up by each motif, the influence of degree is largely removed.
-    % Given the similarly high $R^2$ values between the models using proportions of motifs and absolute counts, it does not appear that we lose much explanatory power by removing degree. 
-    % Therefore, our main analysis focus on the effect of the normalized motif role on persistence.
+% Contrasting effects of omnivory at network and species levels
+High proportions of omnivory in the network profile being associated with lower mean persistence while higher participation in the omnivory motif being associated with higher probability of a focal species persisting suggests that the ideals situation would be for a species to be highly omnivorous in a low-omnivory setting.
+Both the frequency of omnivory in a network profile and the frequency of participation in a network tended to increase with connectance.
+As the omnivory motif contains more links than the other three-species motifs, it is not surprising that it increases with increasing connectance.
+In a high-connectance network, there are many pathways by which a bottom-up disturbance can propagate through the network.
+However, as species which participate in many omnivory motifs also tend to have many prey and low trophic levels, they may be less-severely affected by these propogating disturbances than other species.
+In empirical studies, adaptive omnivory can increase the permanence of three-species chains~\citep{Fagan1997, KrivanDiehl2005, AbramsFung2010}.
+Although species in our model cannot swap prey dependent on bottom-level disturbances or lower resource densities, a species with a high proportion of omnivory will in a sense spread its risk between prey items at different trophic levels.
 
-% [[When can motifs still be useful predictors and add insight?]]
-%     Our results suggest that some motifs are associated with higher persistence than others but that this depends on the level of bottom-up disturbance to the food web.
-%     Detailed tests of how motifs affect network stability and species persistence ideally require the ability to manipulate the motif structure of a network directly (i.e., obtaining many replicate networks with set numbers of each motif).
-%     Current methods of generating simulated ecological networks (e.g., the niche and cascade models) do not incorporate information on motifs.
-%     Researchers interested in relationships between motif structure and persistence must therefore generate large numbers of networks and hope that they contain the necessary variety of motifs.
-%     This shotgun approach to simulating different distributions of motifs within networks makes it difficult to investigate the precise effects of different numbers of, for example, the three-species chain motif, whether it occurs high or low in the network, and similar questions.
-%     Fortunately, software is currently in development to allow researchers to simulate networks with particular motif profiles.
-%     We heartily encourage researchers interested in the effects of motifs on persistence to adopt these new methods as they become available.
+In summary, our results show that participating in different three-species motifs is related to a species' risk of extinction following a bottom-up disturbance, and that these relationships syntesize information provided by other, simpler measures of species roles.
+Our results also underscore the importance of the level of disturbance when identifying which species are most at risk of extinction.
+There were strong interactions between the level of disturbance and the relationships between persistence and measures of species roles (motifs and in-degree).
+These simulated results echo empirical studies which report `tipping points' where previously beneficial traits can become risky~\citep{}.
+These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.
 

### 2021-08-18 manuscript/motifs_vs_degree.tex

@@
 
+% Try to reorder following a sensible discussion
+% Mention chain, direct competition briefly and then get rid of them.
+
 \section*{Abstract}
 
+    Bottom-up effects of disturbance to basal resources can strongly affect the persistence of consumers.
+    However, predicting which species will be most affected by increasing levels of disturbance is not straightforward.
+    In-degree and trophic level are known to affect a consumer's vulnerability to bottom-up disturbance, but these are relatively coarse measures of a species' place in its food web.
+    Here we explore how roles based on participation in three-species motifs are related to persistence, and how these motif-based roles relate to simpler measures.
+    We found that networks containing a higher proportion of the `omnivory' motif, where a predator consumes a prey and the prey of that prey, tend to have higher mean persistence following disturbance to basal species, and that more frequent participation in the omnivory motif was associated with higher probability of persistence.
+    However, the relationship between omnivory and persistence was highly variable across networks; much more so than for the other motifs we consider.
+    % % Something to wrap all of this up, including degree and TL.
+
     - can we condense omnivory, in-degree any further?
-    - - at least omnivory
+    - - at least omnivoryhttps://www.overleaf.com/project/602a809ce82587324badb832
     - can we shorten any more? Nice to save some money?
-    - glossary
+    - glossaryhttps://www.overleaf.com/project/602a809ce82587324badb832
     - refs
@@
 
-    - repeat network size analyses with \# basal resources - maybe adjust Fig. 1 to reflect n(basal) and C instead of C:S?
     - Fig. 4 not much in discussion, could be a candidate for the SI? Or add more discussion about apparent competition, etc. in discussion sections per motif
@@
 
-        \subsubsection*{How does motif participation vary with degree and trophic level?}
+        \subsubsection*{How does motif participation vary with species and network properties?}
 
-            A further complication when trying to understand whether relationships between motifs and probability of extinction depend upon underlying relationships between degree or trophic level and extinction
+            As with network profiles, species' motif participation may covary with other measures of network structure and species' roles within the network.
+            To establish this context for motif participation, we fit a linear mixed-effect model relating the frequency of participation in each motif to network size, connectance, and their interaction, as for network motif profiles (\emph{Appendix SG}).
             We also fit linear mixed-effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
@@
 
-    % % % Things in SI that are not referred to:
-    % persistence vs. motif positions
-    % motif roles (positions) vs. other network properties
+% % % Things in SI only
+% persistence vs. motif positions - SH
+% motif roles (positions) vs. other network properties - SI
 % persistence vs. % Basal
@@
         At high levels of disturbance, a species' persistence decreased as the proportion of the three-species chain increased, but persistence increased as the proportion of any other motif increased (Figure~\ref{fig:prop_lmer_all}).
-        This means that increased disturbance strengthened the effects of the three-species chain motif and changeed the direction of the effect of the apparent competition motif  (Table S4, \emph{Appendix SD}).
+        This means that increased disturbance strengthened the effects of the three-species chain motif and changed the direction of the effect of the apparent competition motif  (Table S4, \emph{Appendix SD}).
         The relationship between omnivory and persistence was still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.
@@
         Species' motif participation also varied with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
-        Species with higher in-degree tend to have higher frequencies of participation in the omnivory motif. Proportions of the other three motifs tend to decline with increasing degree.
-        Species with higher trophic levels tend to have higher proportions of apparent competition and three-species chains.
+        Species with higher in-degree tended to have higher frequencies of participation in the omnivory motif. Proportions of the other three motifs tended to decline with increasing degree.
+        Species with higher trophic levels tended to have higher proportions of apparent competition and three-species chains.
 
@@
 % Results summary - network level (following Anna)
-Changes in the basal level in food webs have profound effects on all other species in the network \citep{}, as all consumers depend directly or indirectly on primary producers for their survival \citep{}.
-It has repeatedly been shown that global network properties are important for how well a network can handle disturbances, such as loss of species \citep{Eklof2006, Dunne2002}.
+    It has repeatedly been shown that global network properties are important for how well a network can handle disturbances such as loss of species \citep{Eklof2006, Dunne2002}. 
     At the same time, from a conservation perspective it is often more relevant to focus interventions on particular target species~\citep{}. 
-Nevertheless, the process of identifying target species should take the network context into account since both direct and indirect interactions can affect how species respond to a disturbance~\citep{}.
+    If target species are not selected \emph{a priori} based on economic or cultural concerns~\citep{}, then it is logical to target conservation efforts towards species that are most at risk of extinction.
+    Identifying characteristics of a species' position within a network that are associated with increased or decreased persistence in the face of disturbance can assist with such efforts to direct scarce resources to the species which need them most.
 
-Motifs are the basic building blocks of networks~\citep{Milo2002}, connecting network-wide (global) and species-specific (local) properties, describing both direct and indirect effects.
-As such, the distribution of these meso-scale structures can be of importance to network and species persistence~\citep{}.
-We find that both the distribution of motifs at the network level and the set of motifs a target species participates in are related to persistence after disturbance to the plant community.
-Also, the nature of these relationships synthesizes the effects of other local and global-scale properties.
+    Motifs, the basic building blocks of networks~\citep{Milo2002}, are promising candidates for network structures that may be relevant to persistence.
+    Motifs connect network-wide (global) and species-specific (local) properties, describing direct and indirect effects which can both affect network stability~\citep{} and species persistence~\citep{}.
+    As such, the distribution of three-species motifs is also related to network stability~\citep{}.
+    We extend this previous research and show that both the distribution of motifs at the network level and the set of motifs a target species participates in are related to that species' persistence after disturbance to the plant community.
 
-A network's motif profile had a relatively strong relationship to the network mean persistence compared to global structures.
+    A network's motif profile had a relatively strong relationship to the network mean persistence compared to global structures (i.e., size and connectance). 
     However, connectance did affect which network motifs were most common. 
     For example, highly-connected networks tended to contain higher frequencies of the omnivory motif which includes more links than the other three-species motifs considered here.
-Thus, global-scale properties may have stronger effects on network persistence by changing meso-scale properties than directly.
-This in turn suggests that network motif profiles may be more useful than global network properties when predicting network-level extinction risk.
+    Thus, global-scale properties seem to have stronger effects on network persistence by changing meso-scale properties than directly.
+    This in turn suggests that network motif profiles may be more useful than global network properties when predicting network-level mean extinction risk.
 
     % Network-level persistence
-In general, average network persistence was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif. 
-%The negative relationship between omnivory and network persistence is consistent with other studies showing that an omnivory motif is less likely to be stable than the other motifs considered~\citep{Borrelli2015a}.
-Motif profiles with high apparent competition and low omnivory were more likely in less-connected food webs, which generally had higher mean persistence than highly-connected food webs.
-Despite these general trends, networks with more variable motif profiles had higher mean persistence, suggesting that presence of a variety of different network structures support persistence.
+    In general, average persistence across a network was lowest in networks with high frequencies of the omnivory motif and highest in networks with high frequencies of the apparent competition motif.
+    Both of these motifs have been linked to stability of ecological networks.
+    Apparent competition ...
+    Omnivory has been proposed as both a stabilizing an destabilizing structure.
+    When modelled in isolation, the omnivory motif is moderately likely to be stable while the other motifs we consider are highly likely to be stable~\citep{Borrelli2015a}.
+    In empirical systems, higher levels of omnivory ...
 
-A network's motif profile ...
-Both omnivory and apparent competition have been shown to have consequences for the stability of ecological networks.
-However, particularly the role of omnivory is debated and it has been shown to act both stabilizing and destabilizing.
-[Add examples of both ...
-]
+    As expected based on the network-level effects, participation in the omnivory and apparent competition motifs was also strongly related to each species' probability of persistence.
+    In contrast to the network-level results, participating more frequently in the omnivory motif \emph{increased} a species' probability of persistence with participating more frequently in the apparent competition motif \emph{decreased} persistence, except at very high levels of disturbance.
+    These species-level trends may be explained by relationships between motif participation, in-degree, and trophic level.
 
-Moreover, species within a network can have substantially different probabilities of persistence than the mean across all species.
-For targeted conservation actions this type of information is often more relevant.
+    Participation in the apparent competition motif was strongly negatively correlated with in-degree and positively correlated with trophic level --a high proportion of the apparent competition motif implies a low in-degree and high trophic level, both of which are associated with lower probabilities of persistence.
+    Moreover, the relationships between persistence and in-degree and apparent competition have a similar interaction with disturbance.
+    Both are associated with lower persistence at low disturbance but increased persistence at high disturbance. 
+    This may occur because, when disturbance is high, connections to many prey species mean many sources of bottom-up disturbance.
 
-% Species level persistence
-While the persistence of a particular species is affected by global-scale (size and connectance) and meso-scale (motif profiles) network structure, it also depends on local-scale (trophic level and in-degree) network properties.
-Local properties may be especially important for the apparent competition and direct competition motifs, due to the indirect effects not being fully expressed in these two motifs since the Bayesian Network framework does not include top-down effects.
+    The omnivory motif showed opposite trends, being strongly and positively correlated with in-degree and negatively correlated with trophic level.
+    A species with a high frequency of participation in the omnivory motif is therefore likely to have many prey and a low trophic level, both of which are associated with higher probabilities of persistence.
+    The combination of low trophic level and high in-degree means that participating in the omnivory motif was positive at all levels of disturbance, while a high in-degree alone was not.
+    This also demonstrates how motif participation can synthesize complementary information provided by other measures of species' positions within networks.
 
-The effect of the local-scale network properties in-degree and trophic level differed.
-A high trophic level was associated with lower persistence across all levels of disturbance.
+    Discussion of the distribution results could go here...
+
+% The effect of the local-scale network properties in-degree and trophic level differed.
+% A high trophic level was associated with lower persistence across all levels of disturbance.
 This was expected, as species at higher trophic levels here by definition have equal or higher probability of extinction compared to their prey ~\citep{Eklof2013}.
-The effect of in-degree is less obvious and depended on the level of disturbance; a high in-degree had a positive effect on persistence at low levels of disturbance but not at high levels of disturbance.
-In addition to these different relationships with persistence, in-degree and trophic level had different relationships to the four motifs we investigated.
-
-Participation in the apparent competition motif was strongly negatively correlated with in-degree --a high proportion of the apparent competition motif implies a low in-degree, and both are associated with lower persistence at low disturbance but increased persistence at high disturbance.
-This suggests that in-degree is in fact the factor driving the relationship between proportion of apparent competition motif and species persistence.
-The turnover from a negative to a positive relationship with an increasing level of disturbance indicate that when disturbance is high, connections to many prey species is negative as the consumer receives the negative impact from several sources.
-A species with a high proportion of the apparent competition motif, which also has a low in-degree, thus gets an advantage when plant level disturbance is high since fewer prey items transmit the disturbance.
-In-degree does not, however, explain all of the trends we observe for motifs as degree and motif participation are not always tightly coupled.
-For example, it is possible for a species with low degree to participate in many motifs if it interacts with high-degree species and thereby obtain many indirect interaction partners.
-This is one of the strengths with analyses of motifs that it considers local structures additionally include indirect connections one link away.
+% The effect of in-degree is less obvious and depended on the level of disturbance; a high in-degree had a positive effect on persistence at low levels of disturbance but not at high levels of disturbance.
+% In addition to these different relationships with persistence, in-degree and trophic level had different relationships to the four motifs we investigated.
 
 The direct competition motif was not strongly correlated to in-degree, but had a strong negative relationship with trophic level.
-Thus, species with a high proportion of the direct competition motif has commonly a low trophic level, which is associated with higher probability of persistence at all levels of disturbance.
-The clear positive relationship between proportion of direct competition motifs and species persistence is therefore likely due to the tendency for species at low trophic levels to have a higher proportion of that motif.
+Thus, species with a high proportion of the direct competition motif has in general a low trophic level, which is associated with higher probability of persistence at all levels of disturbance.
+The clear positive relationship between proportion of the direct competition motif and species persistence is therefore likely due to the tendency for species at low trophic levels to have a higher proportion of that motif.
 
@@
 Hence, the negative effect on persistence of having a high proportion of the three-species chain motif is likely connected to the high trophic level.
-
-%The omnivory motif likewise integrates effects of in-degree and trophic level.
-The omnivory motif is strongly related to both in-degree and trophic level and must be explained using a combination of the two simpler network properties.
-Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between them
-Species participation in the omnivory motif was associated with greater persistence, especially at low levels of disturbance.
-This is consistent with a strong negative correlation between omnivory and trophic level (leading to the overall positive effect of omnivory) and a strong positive correlation with in-degree (leading to the reduced benefit of omnivory at high levels of disturbance).
-
-Despite participating more often in the omnivory motif being associated with greater persistence for a focal species, having a greater proportion of omnivory motifs in the network was associated with lower network persistence.
-At the network level, a higher frequency of the omnivory motif leads to more pathways by which the disturbances to basal resources can propagate and cause extinctions of consumers.
-At the species level on the other hand, if extinctions of basal resources are rare, these multiple pathways will tend to protect consumers participating in many omnivory motifs by providing alternative channels of energy.
-If extinctions of basal resources are common, however, these multiple pathways cause a negative impact as they will transmit more disturbances to the consumer.
-
-Additionally, the effect of the omnivory motif on a focal species' persistence was closely connected to connectance.
-In highly connected networks, a high proportion of the omnivory motif strongly contributed to higher persistence.
-This was not the case in less connected networks.
-Following the same reasoning as above, in a highly connected network, numerous links pave the way for a disturbance to spread to a higher extent than in a less connected network. 
-Therefore, in a highly connected web where disturbances have more ways to move, the lower level prey item in the omnivory motif may be of even higher importance assisting as a buffer, decreasing the negative effect of disturbance for the consumer species.
+- sig. correlated with in-degree (I think???) but does not show up in the relationship to persistence
+- chain has to be capturing effect of something else that acts like positive correlation with in-degree and is stronger than effect of in-degree.
 
@@
 
-The relationship between omnivory and stability is subject to a longstanding debate in ecology.
+%The relationship between omnivory and stability is subject to a longstanding debate in ecology.
 Omnivory has been shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, Monteiro2016, Kratina2012}.
-For example, omnivorous interactions may not be stable in isolation, but instead stabilized by interactions with other species when embedded in a larger food web \citep{Kratinaetal2012}.
+%For example, omnivorous interactions may not be stable in isolation, but instead stabilized by interactions with other species when embedded in a larger food web \citep{Kratinaetal2012}.
 Additionally, the possibility of omnivores to feed adaptively from different trophic levels can spread the risks, for example when the prey level are disturbed \citep{Fagan1997}.
-This might be why we have a divergent pattern between a full food web's motif profile versus each single species motif profile.
 

### 2021-08-18 manuscript/motifs_vs_degree.tex

@@
     This may occur because, when disturbance is high, connections to many prey species mean many sources of bottom-up disturbance.
+    - also strongly associated with trophic level but effects don't match. So there must be something else affecting the affect of apparent competition
 
@@
 - sig. correlated with in-degree (I think???) but does not show up in the relationship to persistence
-- chain has to be capturing effect of something else that acts like positive correlation with in-degree and is stronger than effect of in-degree.
+- something else is related to chain that acts like a positive association with in-degree that is strong enough to cancel out the negative relationship with in-degree
 
@@
 When affected by disturbance of plant community...
+
+- maybe end with a brief summary/discussion of the effect of disturbance: seem to see a "tipping point" in both degree and motifs.
 

### 2021-09-28 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-\title{Do motif roles provide unique information about species risk of extinction?}
+\title{Does motif participation provide unique information about species risk of extinction?}
 
@@
 \section*{Abstract}
+    Bottom-up effects of disturbance to basal resources can strongly affect the persistence of consumers.
+    However, predicting which species will be most affected by increasing disturbance is not straightforward.
+    In-degree and trophic level are known to affect a consumer's vulnerability to bottom-up disturbance, but these are relatively coarse measures of a species' place in a food web.
+    Here we explore how roles based on participation in three-species motifs are related to persistence, and how these motif-based roles relate to simpler measures.
+    We show that a consumer species' frequency of participation in different motifs is related to its probability of persistence.
+Similarly, the overall motif profile of a network was related to the average persistence of consumer species.
+    Notably, we found that networks containing a higher proportion of the `omnivory' motif tend to have lower mean persistence following disturbance to basal species, while more frequent participation in the omnivory motif was associated with higher probability of persistence.
+    Additionally, we find that relationships between motif participation and persistence can synthesize the information provided by in-degree and trophic level.
+    Motif participation therefore provides a relatively rich description of a species' structural role in its community and captures important trends in vulnerability to extinction.
 
-    - can we condense omnivory, in-degree any further?
-    - - at least omnivory
-    - can we shorten any more? Nice to save some money?
+    Anna AA: The whole article is now approximately 4400 words (not counting fig texts); intro 850, methods 1600, results 620, discussion 1300.
+See discussion on section statistical analysis in the methods, I think we can make it shorter
     - glossary
@@
     %\subsection*{1 Why are bottom-up effects important for understanding stability?}
-     The plant community is the foundation upon which the myriad of species in terrestrial food webs depend for their survival \citep{}.
-Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community promotes stability across all different levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}.
-Loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers, e.g.
-herbivores, predators, and parasitoids \citep{scherber2010bottom}.
-     %This has been shown in both theoretical work~\citep{} and empirical experiments ~\citep{} and results from the vertical propagation of disturbances through the network of interacting species.
+     The plant community is the foundation upon which the myriad species in terrestrial food webs depend for their survival.
+Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
+The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community affects stability across all levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}.
+Loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators,  parasitoids, etc.
+\citep{scherber2010bottom}.
 
      %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
-    The study of how an initial population decline or extinction leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    How population decline or extinction of one species leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
 Traditionally, there are two main approaches for studying secondary extinctions.
-First, there are topological models, which are solely based on food web structure \citep{dunne2009cascading}.
-Here, initial extinctions only affect other species in the network once a consumer has lost all of its prey and therefore must go extinct.
-The second approach uses dynamical models, which explicitly simulate population dynamics using a system of differential equations \citep{binzer2011susceptibility}.
+First, there are topological models, based only on food web structure \citep{dunne2009cascading}.
+Here,  extinctions only affect other species in the network once a consumer has lost all of its prey and therefore must go extinct.
+The second approach uses dynamical models which explicitly simulate population dynamics using a system of differential equations \citep{binzer2011susceptibility}.
 Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network.
-These models therefore describe complex effects such as indirect interactions and are generally more realistic than topological models.
-However, dynamical models are parameter intensive, can only include a limited number of species, and simulations are time-consuming.
+    This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are more parameter intensive, and simulations are more time-consuming, than topological models.
 
-    A middle‐ground approach for simulating secondary extinctions are Bayesian networks \citep{Eklof2013}.
+    A middle‐ground approach for simulating secondary extinctions is to use  Bayesian networks \citep{Eklof2013}. 
     In this framework, a consumer's probability of extinction depends on the fraction of resources lost and on a baseline probability of extinction which captures causes unrelated to the network structure (e.g., disease, stochastic extinction of small populations).
@@
     %\subsection*{3 Global scale network properties are known to affect stability}
-    The global structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002d, Eklof2006}.
-For example, more highly-connected networks are more resistant to secondary extinctions \citep{Dunne2002d, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{}.
+    Using Bayesian networks, researchers can explore how the structures of realistically-large networks shape their responses to disturbance.
+    The global (whole-network) structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002, Eklof2006}.
+    For example, more highly-connected networks are more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{Krause2003, Teng2004}.
+
     %\subsection*{4 meso-scale nw propoerties also affect stability}
-    Structural properties on the local scale, i.e.
-properties of single species, are also well-studied and there is consensus that species with high trophic levels, few prey (low in-degree), and low abundances are generally at higher risks of secondary extinction \citep{binzer2011susceptibility}.
+    Structural properties on the local (single-species) scale are also known to affect response to disturbance. 
+    Species with high trophic levels (long paths to basal resources), low in-degree (few prey), and low abundances are generally at higher risks of secondary extinction \citep{binzer2011susceptibility}.
     However, in-degree and trophic level are both relatively coarse descriptions of a species' role within its food-web~\citep{Cirtwill2018FoodWebs}. 
-    For example, specialists may not necessarily have a high risk of extinction if it consumes a basal resource which is at low risk of being affected of disturbance.
-On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes preys which themselves possess high risk of being affected by disturbance.
+    For example, a specialist may not necessarily have a high risk of extinction if it consumes a basal resource which is unlikely to be affected by disturbance. 
+    On the other hand, a consumer with low trophic level may be at high risk of extinction if it consumes prey which are themselves very likely to be affected by disturbance.
 
@@
     One way to incorporate more nuance into descriptions of how species fit into their communities is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
-Importantly, these `motif roles' capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions.
+Importantly, these `motif participation roles' capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015JAE}.
     This means that two species with the same in-degrees and trophic levels can still participate in different sets of motifs if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
-    This extended description of species structural roles may give additional information of extinction risk or synthesize the information provided by multiple other measures of a species' place in its community.
-    At the network level, the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities~\cite{giling2019plant}.
+    This extra detail  may be useful when predicting  extinction risk since, at the network level, the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities~\cite{giling2019plant}.
     We therefore expect that the motif structure of a network and species' participation in that structure should be related to risks due to disturbance to basal resources.
@@
     Specifically, we are interested in whether a consumer's extinction risk depends on 1) the motif profile of the full food web and 2) the motifs in which a focal species participates. 
-    In addition, we want to understand how the information provided by motifs is related to classical global and local food web properties in order to better integrate different aspects of network theory.
-    Of the 13 three-species motifs that can occur in empirical food webs there are four motifs that are over-represented in empirical networks and are thought to contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}: the three-species chain, omnivory, apparent competition and direct competition motifs.
-    These four motifs are also the ones which can occur in the acyclic networks required for Bayesian network simulations~\citep{Eklof2013}. 
+    In addition, we want to understand how the information provided by motifs is related to classical global and local food web properties in order to better synthesize information provided by different scales of network structure.
+    Of the 13 three-species motifs that can occur in empirical food webs, there are four motifs that are over-represented in empirical networks and are thought to contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
+These four motifs, the three-species chain, omnivory, apparent competition, and direct competition, are also the only ones which can occur in the acyclic networks required for Bayesian network simulations~\citep{Eklof2013}.
     Therefore, we focus on these four particularly important motifs.
-
-   % \subsection*{Other, simpler measures of species' roles can also affect extinction risk, and may correlate with motif participation}
-
-    	%Species with high trophic levels may be more likely to go extinct since they depend upon the stability of relatively long food chains~\citep{}. 
-        %For the same reasons, the number of motifs a species participates in will also tend to be larger in larger and more-connected networks.
-    	%Likewise, the range of possible trophic levels increases with network size, although higher connectance may keep trophic levels low as long as the shortest food chain carries more weight (as in prey-averaged and shortest path trophic levels).
-        %A species' trophic level in turn may affects a species' motif role by restricting the set of positions in which it can appear~\cite{Cirtwill2018EcolLett}. 
-    	%To unravel to what extent species motif roles gives additional information on species' probabilities of going extinct, we therefore need to take network structure and other role definitions into account.
 
@@
 
-		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.
+		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}.
 		To capture a range of plausible network architectures, we simulated networks with sizes ranging from 50 to 100 species (in steps of 10) and connectances ranging from 0.02 to 0.18 (in steps of 0.04).
 		For full details, see~\citet{Cirtwill2021_inprep}.
-        We then filtered these simulated networks to remove any with unreasonable structures.
-		Specifically, we removed any network containing disconnected components (species or groups of species not connected to the rest of the network) or any network where the shortest path between a consumer and the closest basal resource was \textgreater6 (as these high trophic levels are very uncommon in empirical food webs ~\citep{}).
+        We then filtered these simulated networks and removed any network containing disconnected components
+        % (species or groups of species not connected to the rest of the network) 
+        or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
 		New networks were simulated to replace any removed networks, and this process was repeated until we obtained 100 suitable networks in each combination of size and connectance.
+		All networks were rendered acyclic following~\citet{Allesina2009} and topologically sorted from lowest to highest trophic level (Appendix\emph{S1}).
 
 		A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}.
-		Therefore, all persistence calculations follow a strict bottom-up routine beginning by determining the status of the primary producers (who do not depend on other species) and continuing on to primary consumers (who depend only on primary producers), and so on up the network.
+		All persistence calculations therefore begin  by determining the status of the primary producers (who do not depend on other species).
+		Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on primary producers), and so on up the network.
 
-		While the niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
-We therefore removed any cycles within each network.
-		To render networks acyclic, we followed~\citet{Allesina2009} and broke cycles by removing links which do not contribute to the robustness of the food web.
-		This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009}.
-		These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
-		Next, we order the species from lowest to highest trophic level using a topological sorting routine following \citep{Tarjan1972, Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
+		While niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
+		We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009}.
+		%This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009}.
+		%These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
+		Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citep{Tarjan1972, Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
         After these steps, calculations of persistence probabilities can begin.
@@
 		Calculating $P(i)$ or $P(\lnot i)$ is less straightforward when a consumer $i$ loses a fraction $k/n = f$ of its prey. 
-		Plausible ecological responses to losing some prey include i) a topological response, where the consumer's probability of extinction is constant unless all resources are lost, ii) a linear response, in which the extinction probability increases linearly with resources loss, and iii) several nonlinear responses where the probability of extinction increases non-linearly with the fraction resource lost.
-		The extinction probability will differ depending on the response function. 
-		We use a sigmoid non-linear functional response of consumers to the loss of resources ($\alpha >1, \beta >1$) which has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
-		We use a cumulative density function of a beta distribution for this purpose:
-		\begin{equation}
-		\label{betafunc}
-		P(\lnot i|f) = \pi_{i} + (1 - \pi_{i}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}
-		\end{equation}
-
+		We use a sigmoid non-linear functional response of consumers to the loss of resources, which has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
+For details see SI.
 		We calculate the fraction of resources lost ($f$) for each consumer $i$, to compute $P(\lnot i|f)$. 
@@
 		Some species will be extinct due to primary extinctions (stochastic extinction with all resources left), while others will experience secondary extinctions; as a response to primary extinctions (partial or total loss of resources).
+		Although it is possible to solve the Bayesian network exactly \citep{Eklof2013}, this is cumbersome for larger networks.
+As numerical simulations are highly efficient and produce the same result \citep{Haussler2020} we use that method here.
 		As this method is probabilistic, repeating the simulation will yield different results. 
@@
 
-	\subsection*{Defining motif profiles and participation [species roles]}
+	\subsection*{Defining motif profiles and species roles}
 
-        As a first measure of the meso-scale structure, we calculated the `motif profile' of each network.
+        We calculated the `motif profile' of each network.
 This is, in our case, a four-dimensional vector of the number of each three-species motif in the network.
@@
 
-        We then describe meso-scale structure at the species level by defining each species' motif participation and motif roles.
+        We then defined each individual species' motif participation.
 Motif participation describes the frequency with which a focal species appears in each of the four motifs, analogous to the network motif profile~\citep{Stouffer2012}.
-        A species' motif role describes the frequency with which a focal species appears in each \emph{unique position} (e.g., top, middle, or bottom species in the three-species chain, top or bottom species in the apparent competition motif) within each motif ~\citep{Stouffer2012,Cirtwill2017}.
-		We calculated species' motif participation and roles using the Python package \emph{pymfinder}~\citep{pymfinder}.
+
+		We calculated species' motif participation using the Python package \emph{pymfinder}~\citep{pymfinder}.
         As with motif profiles of networks, we normalized participation (and role) vectors by dividing each count by the total number of motifs (motif positions) the species appears in.
@@
 
-        Finally, we calculated simpler measures of a species' role within its community: degree and trophic level.
-		As the Bayesian network approach only accounts  bottom-up effects, out-degree (number of predators) is irrelevant to extinction risk.
-		We therefore focused on species' in-degrees (number of prey).
-		Note, however, that out-degree may also be important in topological simulations which include both top-down and bottom-up effects.
-
-		There are many definitions of trophic level available, most of which tend to be very similar~\citep{Carscallen2012}.
-		Here we use shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}. 
-		Basal resources have an STL of one, any species consuming a basal resource has an STL of two, etc.
-        STL is therefore a measure of the minimum number of steps away a species is from any disturbance to the base of a food web.
+        Finally, we calculated simpler measures of a species' role within its community: in-degree and trophic level (shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}).
 
 	\subsection*{Statistical analysis} 
+	We performed multiple statistical analysis, investigating 1) how species persistence vary with motif participation (Appendix XX), 2) how motif participation vary with degree and trophic level (Appendix XX), 3) how network mean persistence vary with network motif profiles (Appendix XX), 4) how network motif profiles vary with global structural properties (Appendix XX), 5) how network mean persistence vary depending on network properties (Appendix XX), and finally 6) how persistence varies with degree and trophic level (Appendix XX).
+The analyses are briefly described below.
+For detailed descriptions see Appendix XX.
 
-        \subsubsection*{How does species persistence vary with motif participation?}
+    \subsection*{AE - Statistical analysis}
 
-            We are interested in whether participating more often in a certain motif (e.g., direct competition) might increase persistence.
-            % (see \emph{Appendix SXX} for relationships between persistence and counts of motifs).
-            To test this, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
+            Effects on persistence from global network properties (i.e.
+$S$ and $C$) must be established before measuring the effects of meso-scale structure on persistence.
+Therefore, we fit a linear regression relating consumer species' probabilities of persistence to network size, connectance, the level of disturbance to basal resources, and all interactions between them.
+As the three-way interaction was significant, we did not simplify the model.
+
+            Global network properties may affect the meso-scale structure of the network.
+Therfore we test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
+For details see SI(XX).
+
+            To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
+For details see SI(XX).
+
+            To supplement these overall tests, we also fit four regressions testing whether persistence had a linear relationship to the proportion of a network profile made up of each motif, and whether this varied with the level of disturbance.
+            Each regression included the proportion of a motif, disturbance, and their interaction as predictors.
+
+            To test if participating more often in a certain motif affect a species persistence, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
             These models included the proportion of the species' role made up by the focal motif, the level of disturbance to basal species, and the interaction between the two.
-            We centered and scaled the predictors before fitting the LMMs.
-            The models also included a random intercept for the interaction between species richness and connectance (equations.
-            All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-            We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.
+The models also included a random intercept for the interaction between species richness and connectance.
 
@@
 
-        \subsubsection*{How does motif participation vary with degree and trophic level?}
+            To test whether these general trends were consistent across networks with different global properties and different levels of disturbance, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources.
+We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.
 
-            A further complication when trying to understand whether relationships between motifs and probability of extinction depend upon underlying relationships between degree or trophic level and extinction
-            We also fit linear mixed-effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-            We performed similar analyses to test whether a species' motif role is related it in-degree or trophic level (\emph{Appendix SI}).
+            A species' motif participation can be related to its in-degree and trophic level.
+Therefore we fit two linear mixed-effects regressions relating persistence to degree or trophic level, $\pi_{disturbed}$, and their interaction.
+We also included a random effect of the interaction between network size and connectance in order to account for the differences in persistence between different networks.
 
-    	\subsubsection*{How consistent are motif-persistence relationships?}
+            Species' motif participation may covary with measures of network structure and species' position within the network.To establish this context for motif participation, we fit a linear mixed-effect model relating the frequency of participation in each motif to network size, connectance, and their interaction, as for network motif profiles (\emph{Appendix SG}).We performed similar analyses to test whether a species' motif role is related it in-degree or trophic level (\emph{Appendix SI}).
 
-    		As well as identifying general relationships between motif participation and persistence, we were also interested in whether these general trends were consistent across networks with different global properties and different levels of disturbance.
-            To test this, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources.
-            These regressions were fit using the R~\citep{R} base function `lm'.
-            We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance. 
-            We performed similar analyses to test whether appearing in different positions in each motif (i.e., the motif role) is related to a species' persistence (\emph{Appendix SH}).
-
-        \subsubsection*{How does network mean persistence vary with motif profiles?}
-
-            To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
-            We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
-            In case high or low mean persistences are associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable).
-            All statistical tests were conducted as when relating motif profiles to global network structure.
-
-            To supplement these overall tests, we also fit four regressions testing whether persistence had a linear relationship to the proportion of a network profile made up of each motif, and whether this varied with the level of disturbance.
-            Each regression included the proportion of a motif, disturbance, and their interaction as predictors.
-            All regressions were fit using the R~\citep{R} base function `lm'.
-
-        \subsubsection*{How do network motif profiles vary with global structure?}
-
-            As well as affecting persistence directly, global network structure (i.e., size and connectance) may affect the meso-scale structure of the network.
-            This is especially likely since we include only networks which can retain all initial species after 1000 rounds of simulated Lotka-Volterra population dynamics~\citep{Cirtwill2021_inprep}:  some meso-scale structures might be more likely to allow all species to persist than others in networks with different size and connectance constraints.
-            We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
-            Significance was calculated based on 9999 permutations.
-            PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
-            To test whether this applies in our case, we fit an ANOVA of the dispersion of motif profiles about the centroid for each combination of network size and connectance. 
-            We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
-            To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif to network size, connectance, and their interaction.
-            We fit each regression using the R~\citep{R} base function `lm'.
-
-        \subsubsection*{How does network mean persistence vary depending on network properties?}
-
-            Both species richness (S) and connectance (C) are known to influence persistence in both simulation studies and empirical food webs.
-            In addition, increasing the probability of extinction for basal species ($\pi_{disturbed}$ above) is intended to decrease persistence of species at higher trophic levels.
-            Although these effects are not the main focus of our manuscript, we must establish these background effects before measuring the effects of meso-scale structure on persistence.
-            The persistence of basal species in our simulations is determined only by their baseline probability, being either $\pi_{base} = 0.1$ or higher, dependent on $\pi_{disturbed}$.
-            We therefore considered only non-basal species (i.e., those with at least one prey) throughout this manuscript.
-
-            To test for effects of global network properties, we fit a linear regression relating consumer species' probabilities of persistence to network size, connectance, the level of disturbance to basal resources, and all interactions between them.
-            All predictors were centered and scaled before fitting the model. 
-            As the three-way interaction was significant, we did not simplify the model.
-            All models were fit using the R~\citep{R} base function `lm'.
-
-        \subsubsection*{How does persistence vary with degree and trophic level?}
-
-            A species' motif participation is likely related to its degrees and trophic level since all three are measures of network position.
-            To understand how this may have affected the results above, we first fit two linear mixed-effects regressions relating persistence to degree or trophic level, $\pi_{disturbed}$, and their interaction.
-            As above, we centered and scaled the predictors.
-            We also included a random effect of the interaction between network size and connectance in order to account for the differences in persistence between different networks. 
-            These regressions provide context about how persistence varies with simpler measures of species' roles within networks.
-
+            All predictors were centered and scaled before fitting the models. 
 \section*{Results}
 
-    We are interested in how a species' participation in different motifs affects its persistence, and how these relationships relate to other measures of network structure. 
+    %We are interested in how a species' participation in different motifs affects its persistence, and how the overall motif profile of a network relate to the average persistence of consumer species.
 We analyse this for different levels of disturbance to the basal level.
 To provide context, we also investigate relationships between persistence and classic global (size and connectance) and local (in-degree and trophic level) network properties.
-    % The results are organized as follows: 1) How network persistence varies with network motif distribution.
-2) How network persistence varies with global structural properties.
-3) How network motif distribution varies with global structural properties.
-4) How species probability of persistence varies with its motif profile.
-5) How species persistence varies with its characteristics such as TL and in-degree.
-6) How a species motif profile depend on its characteristics.
-7) How a species persistence varies with its motif role (position in respective motif).
 
@@
 
-       The effects of motif participation on a species' probability of persistence depends on the level of disturbance.
-When disturbance is low, species' persistence increases with increasing proportions of the omnivory and direct competition motifs.
-Conversely, a species' persistence decreases with an increasing proportion of the apparent competition motif.
-Persistence is not significantly related to the proportion of the three-species chain motif (Figure~\ref{fig:prop_lmer_all}).
+       The relationship between frequency of participation in a motif and a species' probability of persistence depended on the level of disturbance. 
+       When disturbance was low, species' persistence increased with increasing proportions of the omnivory and direct competition motifs. 
+       Conversely, persistence decreased with an increasing proportion of the apparent competition motif. 
+       Persistence was not significantly related to the proportion of the three-species chain motif (Figure~\ref{fig:prop_lmer_all}).
 
-        At high levels of disturbance, a species' persistence decreases as the proportion of the three-species chain increases, but persistence increases as the proportion of any other motif increases (Figure~\ref{fig:prop_lmer_all}).
-This means that increased disturbance strengthens the effects of the three-species chain motif and changes the direction of the effect of the apparent competition motif  (Table S4, \emph{Appendix SD}).
-The effect of omnivory on persistence is still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.
+        At high levels of disturbance, a species' persistence decreased as the proportion of the three-species chain increased, but persistence increased as the proportion of any other motif increased (Figure~\ref{fig:prop_lmer_all}).
+        This means that increased disturbance strengthened the effects of the three-species chain motif and changed the direction of the effect of the apparent competition motif  (Table S4, \emph{Appendix SD}).
+        The relationship between omnivory and persistence was still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.
 
@@
 
-    \subsection*{Consistency in effects of motif participation on species persistence}
-
-        % In order to evaluate how consistent the relationships between a species persistence and motif participation is, and how it varies with  level of disturbance and global network structure, we fit linear regressions of species persistence against motif proportions for the species for each network separately.
-We then analyze the density distribution of the values of the slopes.
-Network size had a much smaller effect than connectance (Fig.
-S3; \emph{Appendix SE}); we therefore focus on connectance here.
-
-        When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally have more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
-However, there is an interaction between disturbance and connectance for the direct competition and omnivory motifs; At low and medium connectances there is little change in the proportion of positive slopes when the proportion of  direct competition motif is high, independent of level of disturbance.
-At high connectances, however, the proportion of positive slopes is lower and also decreases with increasing disturbance.
-For the omnivory motif the trend is reversed.
-Additionally, the omnivory motif stands out having a wider curve when network connectance is low.
-        Network size had a much smaller effect than connectance (Fig. S3; \emph{Appendix SE}).
-
-        \begin{figure}[h!]
-            \centering
-            \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.pdf}
-            \caption{Density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
-Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
-Rows show various levels of connectance.
-The dotted, vertical line indicate zero on the x-axis.
-A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
-The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend).
-}
-            \label{fig:density_prop_C}
-        \end{figure}
-
-    \subsection*{Motif participation, network properties, and species characteristics}
+    \subsection*{Motif participation and other network properties}
 
@@
        Species' motif participation also varied with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
-        Species with higher in-degree tend to have higher frequencies of participation in the omnivory motif. Proportions of the other three motifs tend to decline with increasing degree.
-        Species with higher trophic levels tend to have higher proportions of apparent competition and three-species chains.
+       Species with higher in-degree tended to have higher frequencies of participation in the omnivory motif. Proportions of the other three motifs tended to decline with increasing degree.
+       Species with higher trophic levels tended to have higher proportions of apparent competition and three-species chains.
+
+       A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
+Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S5, \emph{Appendix SF}).
+       In-degree shows an even stronger interaction with disturbance; having more prey species is associated with higher persistence without disturbance, but lower persistence at high levels of disturbance.
 
@@
 
+    \subsection*{Consistency in effects of motif participation on species persistence}
+
+        % In order to evaluate how consistent the relationships between a species persistence and motif participation is, and how it varies with  level of disturbance and global network structure, we fit linear regressions of species persistence against motif proportions for the species for each network separately.
+We then analyze the density distribution of the values of the slopes.
+Network size had a much smaller effect than connectance (Fig.
+S3; \emph{Appendix SE}); we therefore focus on connectance here.
+
+        When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally had more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
+        However, this trend varied at different levels of disturbance, especially for the direct competition and omnivory motifs.
+        At low and medium connectances there was little change in the distribution of slopes for direct competition across all levels of disturbance. 
+        At high connectances, however, the proportion of positive slopes was both lower overall and decreased with increasing disturbance. 
+        For the omnivory motif these trends were reversed. 
+        Moreover, the omnivory motif has a particularly broad distribution of slopes (i.e., an inconsistent relationship to persistence) when network connectance was low. 
+        Network size had a much smaller effect than connectance (Fig. S3; \emph{Appendix SE}).
+
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.pdf}
+            \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
+Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
+Rows show various levels of connectance.
+The dotted, vertical line indicate zero on the x-axis.
+A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
+The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend).
+}
+            \label{fig:density_prop_C}
+        \end{figure}
+
     \subsection*{Network motif profiles and network mean persistence}
 
-        The overall motif profile of a network is related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001).
-Average persistence is higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}).
-Conversely, average persistence is lower in networks with higher proportions of the omnivory motif.
-These trends hold for all levels of disturbance (Fig.~\ref{motif_profile_persistence}, Table S3), \emph{Appendix SC}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
+        The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001). 
+        Average persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}). 
+        Conversely, average persistence was lower in networks with higher proportions of the omnivory motif. 
+        These trends were consistent across levels of disturbance (Fig.~\ref{fig:motif_profile_persistence}, Table S3), \emph{Appendix SC}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
 For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).
@@
             \includegraphics[width=.75\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
-            \caption{The proportions of a network's motif profile made up by the four motifs which can appear in acyclic networks vary with the size and connectance of the network.
-More-connected networks tend to contain relatively more omnivory motifs and fewer apparent competition motifs.
-The proportions of direct competition and three-species chain motifs were less variable.}
+            \caption{The proportions of the four motifs in a network's motif profile were related to the mean persistence of species in the network.
+Specifically, persistence decreased as the proportion of omnivory in the network's motif profile increased while persistence increased with increasing proportions of the other three motifs.
+Interactions between motif profiles and disturbance were significant but small.
+Here we show the relationships of probabilities of extinction of basal resources of 0.1 (top set of lines) and 0.5 (bottom set of lines).
+Relationships are shown for the observed range of proportions for each motif.}
             \label{motif_proportion_lms}
@@
 
-        Increasing the probability of extinction for basal species necessarily decreased the mean probability of persistence for consumers. 
-        This effect was clear across the ranges of network size and connectance we considered. 
-        Network size and connectance also affected network mean persistence, but to a much smaller extent than increasing disturbance (Figure S1, Table S1, \emph{Appendix SA}). 
-        Persistences generally decreased in more-connected networks, but the effect of network size varied depending on level of disturbance. 
-        At low disturbance, small and highly-connected webs showed the lowest mean persistence, whereas at high disturbance, large and highly connected webs showed the lowest mean persistence (Fig.
+        %Increasing the probability of extinction for basal species necessarily decreased the mean probability of persistence for consumers. 
+        %This effect was clear across the ranges of network size and connectance we considered. 
+        %Network size and connectance also affected network mean persistence, but to a much smaller extent than increasing disturbance (Figure S1, Table S1, \emph{Appendix SA}). 
+        %Persistences generally decreased in more-connected networks, but the effect of network size varied depending on level of disturbance. 
+        %At low disturbance, small and highly-connected webs showed the lowest mean persistence, whereas at high disturbance, large and highly connected webs showed the lowest mean persistence (Fig.
 S2, \emph{Appendix SA}).
@@
     \subsection*{Species characteristics and persistence}
-
-        A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
-Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S5, \emph{Appendix SF}).
-        In-degree shows an even stronger interaction with disturbance; having more prey species is associated with higher persistence without disturbance, but lower persistence at high levels of disturbance.
 
@@
 % Focus on what additional information is revealed from analyzing motif compared to global network properties or species characteristics.
-% Connect to traits -- Cirtwill & Eklöf, 2018, Ecol Lett. 
-% Bayesian network approach - what extra information does motifs provide when we are interested in bottom- up effects. 
-% * Difference between 3-sp chain and omnivory, the importance of the rescue prey in omnivory (in dynamical sims omni was mostly bad, 3-sp chain good).
-Omni link destabilizing if strong, stabilizing if weak (other studies).
-In BN just a link as important as other links
-% * Omnivory in fig 5 - high connectance, always good, not as good with lower levels of C
-% * Extra prey in omnivory changes from being a security to a liability (with increasing disturbance?)
 % * ADD REFS - high disturbance scenario is analogous to decreasing plant diversity
@@
 % Framing: network structure is important for predicting extinction risk
-Changes in the basal level in food webs have profound effects on all other species in the network \citep{}, as all consumers depend directly or indirectly on primary producers for their survival \citep{}.
-It has repeatedly been shown that global network properties are important for how well a network can handle disturbances, such as loss of species \citep{Eklof2006, Dunne2002}. 
-At the same time, from a conservation perspective it is often more relevant to focus interventions on particular target species~\citep{}. 
-Nevertheless, the process of identifying target species should take the network context into account since both direct and indirect interactions can affect how species respond to a disturbance~\citep{}.
+
+It is well known that bottom-up effects are strong in most ecosystems and that global network properties are important for how well a network can handle disturbances to those energy flows, such as loss of species \citep{Dunne2002, Eklof2006, PascualDunne2006}.
+At the same time, from a conservation perspective it is often more relevant to focus interventions on particular target species~\citep{Bottrilletal2008}. 
+Nevertheless, the process of identifying target species should take the network context into account since both direct and indirect interactions can affect how species respond to a disturbance~\citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
 
 % Framing: motifs are related to network stability, species persistence (our results)
-Motifs are the basic building blocks of networks~\citep{Milo2002}, connecting network-wide (global) and species-specific (local) properties, describing both direct and indirect effects. 
-As such, the distribution of these meso-scale structures within a network can be of importance to network and species persistence~\citep{}.
-Moreover, we find that a species' frequency of participation in different motifs is related to its participation in different motifs.
+Global network properties are often too coarse grained and identification of vulnerable species in that context becomes to broad (e.g.
+primary producers).
+Motifs represent a middle-ground approach as they are the basic building blocks of networks~\citep{Milo2002}, connecting network-wide (global) and species-specific (local) properties, describing both direct and indirect effects.
+As such, the distribution of these meso-scale structures within a network is of importance to network stability ~\citep{prill2005dynamic, bascompte2005simple}.
+We here show that a species' frequency of participation in different motifs is related to its probability of persistence.
 
 % Species-level results summary
-Specifically, a species' probability of persistence increased with its frequency of participation in the direct competition and omnivory motifs and decreased with its frequency of participation in the apparent competition motif.
+Specifically, we found that a species' probability of persistence increased with its frequency of participation in the direct competition and omnivory motifs and decreased with its frequency of participation in the three-species chain motif.
 The relationship between persistence and the frequency of participation in the apparent competition motif depended strongly on the level of disturbance.
@@
 % Degree and TL are both important for omnivory
-The relationship between participation in the omnivory motif and persistence sythesizes relationships between participation and both in-degree and trophic level.
+The relationship between participation in the omnivory motif and persistence connects motif participation with both in-degree and trophic level.
 Species participation in the omnivory motif was associated with greater persistence, especially at low levels of disturbance.
 This is consistent with a strong negative correlation between omnivory and trophic level (leading to the overall positive effect of omnivory) and a strong positive correlation with in-degree (leading to the reduced benefit of omnivory at high levels of disturbance).
-Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between them.
+Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between the measurements.
 
 % Degree and TL don't explain everything
-The relationships between persistence and participation in the apparent competition and three-species chain motifs also demonstrate motif roles' ability to syntesize different aspects of network structure.
+Also the relationships between persistence and participation in the apparent competition and three-species chain motifs demonstrate motif participation roles' ability to syntesize different aspects of network structure.
 In these cases, in-degree and trophic level are not sufficient to explain the trends we observe in the motifs.
-Participation in the apparent competition motif was strongly negatively correlated with in-degree and shows the opposite relationship to persistence.
-However, participation in the apparent competition motif was also strongly and positively correlated with trophic level (which is negatively associated with persistence since predators must have lower probability of persistence than their prey~\citep{Eklof2013}) and this correlation is not reflected in the relationship between participation in the apparent competition motif and persistence.
+Participation in the apparent competition motif was strongly negatively correlated with in-degree and shows a similar relationship to persistence.
+However, participation in the apparent competition motif was also strongly and positively correlated with trophic level (which is negatively associated with persistence since predators by necessity have lower probability of persistence than their prey~\citep{Eklof2013}) and this correlation is not reflected in the relationship between participation in the apparent competition motif and persistence.
 Similarly, participation in the three-species chain motif was significantly correlated with both in-degree and trophic level, but the relationship between the motif and persistence did not reflect the relationship between in-degree and persistence.
 This suggests the influence of some other aspect of network structure on species' persistence.
-As motifs capture information about direct and indirect interactions and include a larger section of the network than degree or trophic level,~\citep{} it is likely that effects of these indirect interactions may be included in the relationships we observe between motifs and persistence.
+As motifs capture information about direct and indirect interactions and include a larger section of the network than degree or trophic level~\citep{Cirtwill2018FoodWebs}, it is likely that effects of these indirect interactions may be included in the relationships we observe between motifs and persistence.
 
@@
 As well as clarifying the relationships between network structure and the effect of bottom-up disturbance on particular species, motifs were also related to the mean probability of persistence across all species in a network.
-In fact, a network's motif profile had a stronger relationship to mean persistence than network size or connectance. 
-However, connectance did affect which network motifs were most common. 
+In fact, a network's motif profile had a stronger relationship to mean persistence compared to network size or connectance. 
+However, connectance naturally did affect which network motifs were most common. 
 For example, highly-connected networks tended to contain higher frequencies of the omnivory motif which includes more links than the other three-species motifs considered here.
-Thus, global-scale properties may have stronger effects on network persistence by changing meso-scale properties than directly.
-This in turn suggests that network motif profiles may be more useful than global network properties when predicting network-level extinction risk.
-
-% Details of network-level results
-In general, average network persistence was highest in networks with low frequencies of the omnivory motif and high frequencies of the apparent competition motif.
-These networks also tended to have lower connectance, but the relationship between connectance and mean persistence was weaker than that for motifs.
-Both omnivory and apparent competition have been shown to have consequences for the stability of ecological networks.
+Thus, global-scale properties seems to have stronger effects on network persistence by changing meso-scale properties than directly.
+This in turn suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.
+In general, we found that average network persistence was highest in networks with low frequencies of the omnivory motif and high frequencies of the apparent competition motif.
+These two motifs that have been shown to have consequences for the stability of ecological networks \citep{Borrelli2015a}.
 
 % Heading into omnivory
-The relationship between omnivory and stability in particular has long been a subject of debate~\citep{}. 
-Omnivory has been shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, Monteiro2016, Kratina2012}. 
-For example, omnivorous interactions may not be stable in isolation, but instead stabilized by interactions with other species when embedded in a larger food web \citep{Kratinaetal2012}.
-Additionally, the possibility of omnivores to feed adaptively from different trophic levels can spread the risks, for example when the prey level are disturbed \citep{Fagan1997}.
-This might be why we have a divergent pattern between a full food web's motif profile versus each single species motif profile.
+The effect of omnivory on stability has long been a subject of debate \citep{Kratina2012}. 
+Omnivory has been shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, Monteiro2016}. 
+For example, omnivorous interactions may not be stable in isolation but instead stabilized by interactions with other species when embedded in a larger food web \citep{Kratina2012}. 
+This might be why we here find a divergent pattern between a full food web's motif profile (where a high proportion of omnivory indicates lower average network persistence) versus each single species motif profile (where a high proportion of the omnivory motif indicates a higher focal species persistence).
+Thus, the ideal situation for a species could potentially be a highly omnivorous species in a low-omnivory setting.
 
-%  % Should go in one of these two paragraphs...
- % \cite{Cirtwill2021_inprep} analysed how species time to extinction varied depending on their network profile.
-They found that a high proportion of the omnivory motif resulted in a shorter time to extinction, while the other three stable motifs were associated with longer times to extinction.
-However, the study by Cirtwill and Wootton used a different approach using a fully dynamic simulation model also including top-down effects.
-    %In previous studies of the effect of omnivory motifs, the strength of the omnivorous link can potentially be important.
-If this link is strong, it is shown to be destabilizing (REF), while stabilizing if weaker (REF).
-In our Bayesian Network setting, this link is indistinguishable from any other link in the food web which partly may explain why omnivory results in increased persistence of consumer species.
+% High C webs and omnivory
+Additionally, we show that both the presence as well as the persistence of omnivorous species are higher in high-connected networks.
+The many pathways present in a high-connectance network, by which a bottom-up disturbance can propagate through the network, might explain the positive effect on persistence for species that feeds on multiple trophic levels.
+%Species that participate in many omnivory motifs have feeding interactions to species at lower trophic levels in addition to the directly adjacent one.
+Those species will therefore be less-severely affected by these propagating disturbances enforced in high-connectance networks, in comparison to other species.
+In a sense, a species with a high proportion of omnivory will spread its risk between prey items at different trophic levels.
+Although species in our model cannot swap prey dependent on bottom-level disturbances or lower resource densities, this mechanism resembles empirical studies showing that adaptive omnivory can increase the permanence of three-species chains~\citep{Fagan1997, Kvrivan2005, AbramsFung2010}.
 
-% Contrasting effects of omnivory at network and species levels
-High proportions of omnivory in the network profile being associated with lower mean persistence while higher participation in the omnivory motif being associated with higher probability of a focal species persisting suggests that the ideals situation would be for a species to be highly omnivorous in a low-omnivory setting.
-Both the frequency of omnivory in a network profile and the frequency of participation in a network tended to increase with connectance.
-As the omnivory motif contains more links than the other three-species motifs, it is not surprising that it increases with increasing connectance.
-In a high-connectance network, there are many pathways by which a bottom-up disturbance can propagate through the network.
-However, as species which participate in many omnivory motifs also tend to have many prey and low trophic levels, they may be less-severely affected by these propogating disturbances than other species.
-In empirical studies, adaptive omnivory can increase the permanence of three-species chains~\citep{Fagan1997, KrivanDiehl2005, AbramsFung2010}.
-Although species in our model cannot swap prey dependent on bottom-level disturbances or lower resource densities, a species with a high proportion of omnivory will in a sense spread its risk between prey items at different trophic levels.
+An additional important predictor for a species persistence is its traits \citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000}.
+In turn, a species' set of traits also influence the likelihood of it participating in different motifs \citep{cirtwill2018feeding}.
+This association between motifs and traits further emphasise the role of motifs for understanding how species' roles in a network have implications for its persistence.
 
-In summary, our results show that participating in different three-species motifs is related to a species' risk of extinction following a bottom-up disturbance, and that these relationships syntesize information provided by other, simpler measures of species roles.
+There are some limitations to the approach presented here.
+Most important is the lack of top-down effects in the extinction risk calculations.
+This is because Bayesian networks operate on a strict bottom-up principle whereby prey influence their predators, but not vice versa \citep{Eklof2013}.
+Here our focus was to analyze the importance of disturbances to the primary producers and how effects propagate up in the food web.
+Earlier work indeed showed that the Bayesian network approach capture a majority of the secondary extinctions when compared to a fully-fledged dynamical model \citep{Eklof2013}, but if top-down effects are assumed important other methods should be used.
+Ongoing work using a dynamical model show that time to extinction for species following disturbance is tightly coupled to the species motif profile \citep{Cirtwill2021_inprep}.
+
+In summary, our results show that participating in different three-species motifs is related to a species' risk of extinction following a bottom-up disturbance, and that these relationships synthesize information provided by other, simpler measures of species roles.
 Our results also underscore the importance of the level of disturbance when identifying which species are most at risk of extinction.
@@
 These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.
+
+\clearpage
 
@@
     Motif role & \\
-    Bayesian network & \\
-    Network persistence & \\
-    Species persistence & \\
-    In-degree & \\
-    Trophic level (STL) & \\
-    Disturbance ($\pi_{disturbed}$) & Probability that a basal resource will go extinct when extra disturbance is added. \\
-    Baseline extinction&  When no disturbance is applied, $\pi_{base}$=0.1.\\
+    Bayesian network & A directed acyclic graph, used to predict species' likelihood of persistence. \\
+    Network persistence & The mean likelihood of consumers in a network not going extinct.\\
+    Species persistence & The individual likelihood of a species not going extinct.\\
+    In-degree & Number of prey species to a consumer species.\\
+    Trophic level (STL) & The shortest food chain between the focal species and any basal species.\\
+    Disturbance ($\pi_{disturbed}$) & Probability of extinction of a basal resource when extra disturbance is added. \\
+    Baseline extinction &  When no disturbance is applied, $\pi_{base}$=0.1.\\
 \end{tabular}

### 2021-09-28 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 % \clearpage
+
+\section{Functional response of consumers}
+        As well as the intrinsic baseline probability of extinction, extinction risk for consumers is affected by access to resources. 
+		When all resources are extant, the probability of extinction for a consumer $i$, $P(\lnot i)$, is equal to the baseline probability; $P(\lnot i|X_{1},...,X_{n}) = \pi_{base}$. 
+		When all resources are extinct, the consumer will go extinct; $P(\lnot i|\lnot X_{1},...,\lnot X_{n})=1$.
+
+		Calculating $P(i)$ or $P(\lnot i)$ is less straightforward when a consumer $i$ loses a fraction $k/n = f$ of its prey. 
+		Plausible ecological responses to losing some prey include i) a topological response, where the consumer's probability of extinction is constant unless all resources are lost, ii) a linear response, in which the extinction probability increases linearly with resources loss, and iii) several nonlinear responses where the probability of extinction increases non-linearly with the fraction resource lost.
+		The extinction probability will differ depending on the response function. 
+		We use a sigmoid non-linear functional response of consumers to the loss of resources ($\alpha >1, \beta >1$) which has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
+		We use a cumulative density function of a beta distribution for this purpose:
+		\begin{equation}
+		\label{betafunc}
+		P(\lnot i|f) = \pi_{i} + (1 - \pi_{i}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}
+		\end{equation}
+
+\section{Statistical analysis}
+ \subsection*{How does species persistence vary with motif participation?}
+
+            We are interested in whether participating more often in a certain motif (e.g., direct competition) might increase persistence.
+            % (see \emph{Appendix SXX} for relationships between persistence and counts of motifs).
+            To test this, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
+            These models included the proportion of the species' role made up by the focal motif, the level of disturbance to basal species, and the interaction between the two.
+            We centered and scaled the predictors before fitting the LMMs.
+            The models also included a random intercept for the interaction between species richness and connectance (equations.
+            All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+            We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.
+
+            The formula for these LMMs are:
+
+            \begin{equation}
+                \Psi_{ik} \approx \rho_{i} + \pi_{disturbed_k} + \rho_{i}\pi_{disturbed_k} +
+                S:C_{i} ,
+                \label{propreq}
+            \end{equation}
+
+            where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
+
+            We were also interested in whether these general trends were consistent across networks with different global properties and different levels of disturbance. 
+            To test this, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources.
+            These regressions were fit using the R~\citep{R} base function `lm'.
+            We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance. 
+            We performed similar analyses to test whether appearing in different positions in each motif (i.e., the motif role) is related to a species' persistence (\emph{Appendix SH}).
+
+        \subsection*{How does motif participation vary with species and network properties?}
+
+            As with network profiles, species' motif participation may covary with other measures of network structure and species' roles within the network.
+            To establish this context for motif participation, we fit a linear mixed-effect model relating the frequency of participation in each motif to network size, connectance, and their interaction, as for network motif profiles (\emph{Appendix SG}).
+            We also fit linear mixed-effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+            We performed similar analyses to test whether a species' motif role is related it in-degree or trophic level (\emph{Appendix SI}).
+
+        \subsection*{How does network mean persistence vary with motif profiles?}
+
+            To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
+            We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
+            In case high or low mean persistences are associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable).
+            All statistical tests were conducted as when relating motif profiles to global network structure.
+
+            To supplement these overall tests, we also fit four regressions testing whether persistence had a linear relationship to the proportion of a network profile made up of each motif, and whether this varied with the level of disturbance.
+            Each regression included the proportion of a motif, disturbance, and their interaction as predictors.
+            All regressions were fit using the R~\citep{R} base function `lm'.
+
+        \subsection*{How do network motif profiles vary with global structure?}
+
+            As well as affecting persistence directly, global network structure (i.e., size and connectance) may affect the meso-scale structure of the network.
+            This is especially likely since we include only networks which can retain all initial species after 1000 rounds of simulated Lotka-Volterra population dynamics~\citep{Cirtwill2021_inprep}:  some meso-scale structures might be more likely to allow all species to persist than others in networks with different size and connectance constraints.
+            We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
+            Significance was calculated based on 9999 permutations.
+            PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
+            To test whether this applies in our case, we fit an ANOVA of the dispersion of motif profiles about the centroid for each combination of network size and connectance. 
+            We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
+            To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif to network size, connectance, and their interaction.
+            We fit each regression using the R~\citep{R} base function `lm'.
+
+        \subsection*{How does network mean persistence vary depending on network properties?}
+
+            Both species richness (S) and connectance (C) are known to influence persistence in both simulation studies and empirical food webs.
+            In addition, increasing the probability of extinction for basal species ($\pi_{disturbed}$ above) is intended to decrease persistence of species at higher trophic levels.
+            Although these effects are not the main focus of our manuscript, we must establish these background effects before measuring the effects of meso-scale structure on persistence.
+            The persistence of basal species in our simulations is determined only by their baseline probability, being either $\pi_{base} = 0.1$ or higher, dependent on $\pi_{disturbed}$.
+            We therefore considered only non-basal species (i.e., those with at least one prey) throughout this manuscript.
+
+            To test for effects of global network properties, we fit a linear regression relating consumer species' probabilities of persistence to network size, connectance, the level of disturbance to basal resources, and all interactions between them.
+            All predictors were centered and scaled before fitting the model. 
+            As the three-way interaction was significant, we did not simplify the model.
+            All models were fit using the R~\citep{R} base function `lm'.
+
+        \subsection*{How does persistence vary with degree and trophic level?}
+
+            A species’ motif participation is likely related to its in-degree and trophic level since all three properties are measures of network position.
+To provide context for the results above, we therefore fit two linear mixed-effects regressions relating persistence to degree or trophic level, $\pi_{disturbed}$, and their interaction.
+As above, we centered and scaled the predictors.
+We also included a random effect of the interaction between network size and connectance in order to account for the differences in persistence between different networks.
 

### 2021-09-28 manuscript/supplement_motifs.tex
Added some persistence stuff to SI

@@
 
-\section{Functional response of consumers}
+\section{Network creation and processing}
+
+    A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}.
+    All persistence calculations therefore begin  by determining the status of the primary producers (who do not depend on other species).
+    Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on primary producers), and so on up the network.
+
+    While niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
+    We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009}.
+    This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009}.
+    These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
+    Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citep{Tarjan1972, Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
+    After these steps, calculations of persistence probabilities can begin.
+
+\section{Calculating persistence}
+
+    \subsection{Stochastic extinction}
+
+    \subsection{Functional response of consumers}
+
         As well as the intrinsic baseline probability of extinction, extinction risk for consumers is affected by access to resources. 
@@
 		\end{equation}
+
+    \subsection{Numerical simulation}
 

### 2021-09-28 manuscript/reorder_main_text.tex

@@
 		All networks were rendered acyclic following~\citet{Allesina2009} and topologically sorted from lowest to highest trophic level (Appendix\emph{S1}).
+		These processing steps allow the strictly bottom-up calculation of probability of persistence used in the Bayesian network framework.
 
-		A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}.
-		All persistence calculations therefore begin  by determining the status of the primary producers (who do not depend on other species).
-		Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on primary producers), and so on up the network.
+% 		A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}.
+% 		All persistence calculations therefore begin  by determining the status of the primary producers (who do not depend on other species).
+% 		Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on primary producers), and so on up the network.
 
-		While niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
-		We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009}.
-		%This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009}.
-		%These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
-		Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citep{Tarjan1972, Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
-        After these steps, calculations of persistence probabilities can begin.
+% 		While niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
+% 		We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009}.
+% 		%This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009}.
+% 		%These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
+% 		Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citep{Tarjan1972, Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
+%         After these steps, calculations of persistence probabilities can begin.
 
@@
 		First, we assigned all species a baseline probability of extinction ($\pi_{base}$) -- the risk of going extinct given factors not related to the food web itself. 
-		For consumer species, this also applies to extinction despite all resources being present. 
+		For consumer species, this also represents extinction despite all resources being present. 
 		Such events can, for example, reflect diseases or stochastic extinctions of small populations.
 
-		As well as this baseline scenario, we simulate a set of scenarios where primary producers get a higher baseline probability of extinction ($\pi_{disturbed}$). 
+		As well as this baseline scenario, we simulated a set of scenarios where primary producers had a higher baseline probability of extinction ($\pi_{disturbed}$). 
 		These scenarios reflect threats to primary producers such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
-		$\pi_{disturbed}$ ranged between $0.1-0.5$, in steps of $0.08$. The highest disturbance level, $\pi_{disturbed} = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
+		$\pi_{disturbed}$ ranged between $0.1-0.5$, in steps of $0.08$. 
+		The highest disturbance level, $\pi_{disturbed} = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
 		Consumer species retained $\pi_{base}=0.1$ in all cases.
@@
 		Calculating $P(i)$ or $P(\lnot i)$ is less straightforward when a consumer $i$ loses a fraction $k/n = f$ of its prey. 
-		We use a sigmoid non-linear functional response of consumers to the loss of resources, which has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
-For details see SI.
-		We calculate the fraction of resources lost ($f$) for each consumer $i$, to compute $P(\lnot i|f)$. 
+		We used a sigmoid non-linear functional response of consumers to the loss of resources (Appendix \emph{S2}), which has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
+		Probabilities of extinction for resource species are used when calculating probabilities for higher level species.
+
+		Although it is possible to solve the Bayesian network exactly \citep{Eklof2013}, this is cumbersome for larger networks. 
+		As numerical simulations are highly efficient and produce the same result \citep{Haussler2020}, we also use simulations here.
+		To compute $P(\lnot i|f)$, we calculated the fraction of resources lost ($f$) for each consumer $i$.
 		Persistence is determined by comparing $P(\lnot i|f)$ with a randomly drawn number from a uniform distribution between 0-1. 
@@
 		The species' marginal probability of persistence is simply $P(i) = 1-P(\lnot i)$.
-		In this manner we calculate persistence probabilities for all species in the network, starting from the bottom and working our way up. 
-		Probabilities of extinction for resource species are used when calculating probabilities for higher level species. 
-		Some species will be extinct due to primary extinctions (stochastic extinction with all resources left), while others will experience secondary extinctions; as a response to primary extinctions (partial or total loss of resources).
-		Although it is possible to solve the Bayesian network exactly \citep{Eklof2013}, this is cumbersome for larger networks.
-As numerical simulations are highly efficient and produce the same result \citep{Haussler2020} we use that method here.
-		As this method is probabilistic, repeating the simulation will yield different results. 
-		To minimize the effect of this variation we repeat these calculations 100 000 times per web, with unique random draws when calculating $P(i)$.
+		To minimize noise due to the random draws, we repeat these calculations 100 000 times per web, with unique random draws when calculating $P(i)$.
 		The final persistence for each species is the mean over these runs.
 
-	\subsection*{Defining motif profiles and species roles}
+	\subsection*{Defining species roles}
 
-        We calculated the `motif profile' of each network.
-This is, in our case, a four-dimensional vector of the number of each three-species motif in the network.
-To separate differences in motif structure from differences in network size, we normalized motif profiles by dividing the count of each motif by the total count of all motifs.
-The motif profile for each network therefore describes the proportions of each motif and has a sum of exactly one.
-
-        We then defined each individual species' motif participation.
-Motif participation describes the frequency with which a focal species appears in each of the four motifs, analogous to the network motif profile~\citep{Stouffer2012}.
-
-		We calculated species' motif participation using the Python package \emph{pymfinder}~\citep{pymfinder}.
-        As with motif profiles of networks, we normalized participation (and role) vectors by dividing each count by the total number of motifs (motif positions) the species appears in.
+        A species' ``motif participation role'' is the frequency with which a focal species appears in each of the motifs present in a network~\citep{Stouffer2012}.
+        In our case, these are four-dimensional vectors since only four three-species motifs can appear in acyclic networks.
+        As we are interested in relationships between the type of motif participated in rather than the total number of motifs, we normalized participation vectors by dividing each count by the total number of motifs the species appears in.
         Note, however, that this normalization does \emph{not} control for differences due to degree \emph{per se} as high-degree species also tend to participate in different types of motifs than low-degree species~\citep{Cirtwill2021_inprep}.
 
-        Finally, we calculated simpler measures of a species' role within its community: in-degree and trophic level (shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}).
+        To provide context for species' motif participation, we also calculated network ``motif profiles'': four-dimensional vectors of the number of each three-species motif in each network~\citep{Stouffer2012}.
+        To separate differences in motif structure from differences in network size (larger networks contain more motifs), we normalized motif profiles by dividing the count of each motif by the total count of all motifs.
+        % The motif profile for each network therefore describes the proportions of each motif and has a sum of exactly one.
+		We calculated species' motif participation using the Python package \emph{pymfinder}~\citep{pymfinder}.
+
+        Finally, we calculated simpler measures of a species' role within its community.
+        These were in-degree (number of prey) and trophic level (shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}).
+        Both simple role measures were calculated in R~\citep{R}.
 
 	\subsection*{Statistical analysis} 
+	% Appendix 3, with numbered subheadings.
+
 	We performed multiple statistical analysis, investigating 1) how species persistence vary with motif participation (Appendix XX), 2) how motif participation vary with degree and trophic level (Appendix XX), 3) how network mean persistence vary with network motif profiles (Appendix XX), 4) how network motif profiles vary with global structural properties (Appendix XX), 5) how network mean persistence vary depending on network properties (Appendix XX), and finally 6) how persistence varies with degree and trophic level (Appendix XX).

### 2021-09-28 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-	We performed multiple statistical analysis, investigating 1) how species persistence vary with motif participation (Appendix XX), 2) how motif participation vary with degree and trophic level (Appendix XX), 3) how network mean persistence vary with network motif profiles (Appendix XX), 4) how network motif profiles vary with global structural properties (Appendix XX), 5) how network mean persistence vary depending on network properties (Appendix XX), and finally 6) how persistence varies with degree and trophic level (Appendix XX).
+	We tested 1) how species persistence varied with motif participation (Appendix \emph{3.1}), 2) how motif participation varied with other network properties (Appendix \emph{3.2}), 3) how network mean persistence varied with network motif profiles (Appendix \emph{3.3}), 4) how network motif profiles varied with global structural properties (Appendix \emph{3.4}), and 5) how persistence varied with other network properties (Appendix \emph{3.5}).
 The analyses are briefly described below.
-For detailed descriptions see Appendix XX.
 
-    \subsection*{AE - Statistical analysis}
+    \subsubsection*{1) Persistence and motif participation}
 
-            Effects on persistence from global network properties (i.e.
-$S$ and $C$) must be established before measuring the effects of meso-scale structure on persistence.
-Therefore, we fit a linear regression relating consumer species' probabilities of persistence to network size, connectance, the level of disturbance to basal resources, and all interactions between them.
-As the three-way interaction was significant, we did not simplify the model.
-
-            Global network properties may affect the meso-scale structure of the network.
-Therfore we test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
-For details see SI(XX).
-
-            To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
-For details see SI(XX).
-
-            To supplement these overall tests, we also fit four regressions testing whether persistence had a linear relationship to the proportion of a network profile made up of each motif, and whether this varied with the level of disturbance.
-            Each regression included the proportion of a motif, disturbance, and their interaction as predictors.
-
-            To test if participating more often in a certain motif affect a species persistence, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
-            These models included the proportion of the species' role made up by the focal motif, the level of disturbance to basal species, and the interaction between the two.
-The models also included a random intercept for the interaction between species richness and connectance.
-
-            The formula for these LMMs are:
+        To test if participating more often in a certain motif affected a species' probability of persistence, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks):
 
@@
                 \Psi_{ik} \approx \rho_{i} + \pi_{disturbed_k} + \rho_{i}\pi_{disturbed_k} +
-                S:C_{i} ,
+                S_{i}C_{i} ,
                 \label{propreq}
@@
 
-            where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
-
-            To test whether these general trends were consistent across networks with different global properties and different levels of disturbance, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources.
+        where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
+        All predictors were centered and scaled before fitting the models.         
+        To test whether these general trends were consistent across networks, we performed additional linear regressions of persistence against motif participation within each simulated network, for each level of disturbance to basal resources (600 regressions).
         We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.
 
-            A species' motif participation can be related to its in-degree and trophic level.
-Therefore we fit two linear mixed-effects regressions relating persistence to degree or trophic level, $\pi_{disturbed}$, and their interaction.
-We also included a random effect of the interaction between network size and connectance in order to account for the differences in persistence between different networks.
+    \subsubsection*{2) Participation and other network properties}
 
-            Species' motif participation may covary with measures of network structure and species' position within the network.To establish this context for motif participation, we fit a linear mixed-effect model relating the frequency of participation in each motif to network size, connectance, and their interaction, as for network motif profiles (\emph{Appendix SG}).We performed similar analyses to test whether a species' motif role is related it in-degree or trophic level (\emph{Appendix SI}).
+        To test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
+        In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMMs:
 
-            All predictors were centered and scaled before fitting the models. 
+        \begin{equation}
+            \rho_{i} \approx S_{i} + C_{i} + S_{i}C_{i}
+        \end{equation}
+        where all symbols are as in equation~\ref{propreq}.
+
+    \subsubsection*{3) Persistence and motif profiles}
+
+        As well as species-specific probabilities of persistence, motifs may provide information about the overall probability of any species going extinct (since they incorporate information on indirect effects).
+        To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
+        To identify the relationships between particular motifs and network mean persistence, we fit a set of four LMMs (one per motif):
+
+                \begin{equation}
+                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{disturbed_k} + \Bar{\rho}_{i}\pi_{disturbed_k} +
+                S_{i}C_{i} ,
+                \label{netpropeq}
+            \end{equation}
+
+        where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and all other symbols are as in equation~\ref{propreq}.
+
+    \subsubsection*{4) Motif profiles and network properties}
+
+        Different scales of network structure (e.g., local and meso-scale) are not independent. 
+        To test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
+        In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMMs:
+
+        \begin{equation}
+            \rho_{i} \approx S_{i} + C_{i} + S_{i}C_{i}
+        \end{equation}
+        where all symbols are as in equation~\ref{propreq}.
+
+    \subsubsection*{5) Persistence and network properties}
+
+        Persistence is known to vary with global-scale and local-scale network properties.
+        To correctly interpret any relationships between motifs and persistence, we must consider the context of these other relationships.
+        To identify the relationship between global network structure and mean probability of persistence within a network, we fit a linear regression:
+            \begin{equation}
+                \Psi_{ik} \approx S_{i} + C_{i} + \pi_{disturbed_k} + S_{i}:C_{i} + S_{i}\pi_{disturbed_k} + C_{i}\pi_{disturbed_k} + S_{i}C_{i}\pi_{disturbed_k},
+                \label{SCeq}
+            \end{equation}
+        where all symbols are as in equation~\ref{propreq}.
+        As the three-way interaction was significant, we did not simplify the model.
+
+        Persistence is also known to vary with local-scale measures of network structure. 
+        To identify the relationships betwen local-scale properties and species' probabilities of persistence, we fit two more LMMs:
+
+        \begin{equation}
+                \Psi_{ik} \approx \delta_{i} + \pi_{disturbed_k} + \delta_{i}\pi_{disturbed_k} +
+                S_{i}C_{i} ,
+                \label{degeq}
+            \end{equation}
+
+        \begin{equation}
+                \Psi_{ik} \approx \tau_{i} + \pi_{disturbed_k} + \tau_{i}\pi_{disturbed_k} +
+                S_{i}C_{i} ,
+                \label{TLeq}
+            \end{equation}
+
+        where $\delta_{i}$ and $\tau_i$ are the in-degree and trophic level of species $i$, respectively, and all other terms are as in equation~\ref{propreq}.
+
 \section*{Results}

### 2021-09-29 manuscript/supplement_motifs.tex
minor changes

@@
 \section{Statistical analysis}
- \subsection*{How does species persistence vary with motif participation?}
 
-            We are interested in whether participating more often in a certain motif (e.g., direct competition) might increase persistence.
-            % (see \emph{Appendix SXX} for relationships between persistence and counts of motifs).
-            To test this, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
+    \subsection{How does species persistence vary with motif participation?}
+
+        We are first interested in whether participating more often in a certain motif (e.g., direct competition) might increase a species' probability of persistence.
+        To test this, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks):
+
+        \begin{equation}
+            \Psi_{ik} \approx \rho_{i} + \pi_{disturbed_k} + \rho_{i}\pi_{disturbed_k} +
+            S_{i}:C_{i} ,
+            \label{propreq}
+        \end{equation}
+
         These models included the proportion of the species' role made up by the focal motif, the level of disturbance to basal species, and the interaction between the two.
@@
         The formula for these LMMs are:
-
-            \begin{equation}
-                \Psi_{ik} \approx \rho_{i} + \pi_{disturbed_k} + \rho_{i}\pi_{disturbed_k} +
-                S:C_{i} ,
-                \label{propreq}
-            \end{equation}
 
@@
 
-        \subsection*{How does motif participation vary with species and network properties?}
+    \subsection{How does motif participation vary with species and network properties?}
 
@@
         We also fit linear mixed-effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-            We performed similar analyses to test whether a species' motif role is related it in-degree or trophic level (\emph{Appendix SI}).
+        We performed similar analyses to test whether a species' motif role is related it in-degree or trophic lev,el (\emph{Appendix SI}).
 
-        \subsection*{How does network mean persistence vary with motif profiles?}
+    \subsection{How does network mean persistence vary with motif profiles?}
 
@@
 
-        \subsection*{How do network motif profiles vary with global structure?}
+    \subsection{How do network motif profiles vary with global structure?}
 
@@
 
-        \subsection*{How does network mean persistence vary depending on network properties?}
+    \subsection{How does persistence vary with species and network properties?}
+
+        \subsubsection{How does network mean persistence vary depending on network properties?}
 
@@
 
-        \subsection*{How does persistence vary with degree and trophic level?}
+        \subsubsection{How does persistence vary with degree and trophic level?}
 

### 2021-09-29 manuscript/supplement_motifs.tex
working on SI methods still

@@
 
-        These models included the proportion of the species' role made up by the focal motif, the level of disturbance to basal species, and the interaction between the two.
+        \noindent where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
         We centered and scaled the predictors before fitting the LMMs.
-        The models also included a random intercept for the interaction between species richness and connectance (equations.
         All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
@@
 
-        The formula for these LMMs are:
+        We were also interested in whether these general trends were consistent across networks with different global properties and different levels of disturbance. 
+        To test this, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources:
 
-        where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
+        \begin{equation}
+            \Psi_{ik} \approx \rho_{i} ,
+            \label{mineq}
+        \end{equation}
 
-        We were also interested in whether these general trends were consistent across networks with different global properties and different levels of disturbance. 
-        To test this, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources.
+        where all symbols are as in equation~\ref{propreq}.
         These regressions were fit using the R~\citep{R} base function `lm'.
-        We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance. 
-        We performed similar analyses to test whether appearing in different positions in each motif (i.e., the motif role) is related to a species' persistence (\emph{Appendix SH}).
+        We then visually examine the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.
 
-    \subsection{How does motif participation vary with species and network properties?}
+    \subsection{How does motif participation vary with network properties?}
 
         As with network profiles, species' motif participation may covary with other measures of network structure and species' roles within the network.
-        To establish this context for motif participation, we fit a linear mixed-effect model relating the frequency of participation in each motif to network size, connectance, and their interaction, as for network motif profiles (\emph{Appendix SG}).
-        We also fit linear mixed-effect models relating the count or proportion of each motif to degree or trophic level and a random effect of the interaction between network size and connectance using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-        We performed similar analyses to test whether a species' motif role is related it in-degree or trophic lev,el (\emph{Appendix SI}).
+        To establish this context for motif participation, we first fit a PERMANOVA~\citep{Anderson2001} to test whether a species' overall motif participation role is related to network size, connectance, or the interaction between them.
+        To avoid inflating similarity between species which do not appear in the same motifs (i.e., which have ``shared zeros''), we use Bray-Curtis dissimilarity between motif participation roles.
+        Significance was calculated using 9999 permutations.
+        The PERMANOVA was fit using the R~\citep{R} function `adonis' from the package \emph{vegan}~\citep{vegan}.
+
+        To obtain a more detailed picture of the relationships between motif participation and global network structure, we fit a set of four linear models (one per motif):
+
+        \begin{equation}
+            \rho_{i} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i} ,
+            \label{partic_SC}
+        \end{equation}
+
+        where $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif,
+        $\Sigma_{i}$ is the size of the network containing species $i$, and $\zeta_{i}$ is the connectance of the network containing species $i$.
+        Predictors were not centered or scaled before fitting.
+        All models were fit using the R~\citep{R} base function `lm'.
+
+        We fit a similar set of eight linear mixed-effect models relating motif participation to in-degree or trophic level:
+
+        \begin{equation}
+            \rho_{i} \approx \delta_{i} + S_{i}C_{i} ,
+            \label{partic_deg}
+        \end{equation}
+
+        \begin{equation}
+            \rho_{i} \approx \tau_{i} + S_{i}C_{i} ,
+            \label{partic_deg}
+        \end{equation}
+
+        where $\rho_{i}$ is the frequency with which species $i$ participtes in the focal motif, $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, $S_{i}C_{i}$ is a random effect of the combination of size and connectance for the network containing species $i$.
+        These models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
 

### 2021-09-29 manuscript/reorder_main_text.tex

@@
 
-	We tested 1) how species persistence varied with motif participation (Appendix \emph{3.1}), 2) how motif participation varied with other network properties (Appendix \emph{3.2}), 3) how network mean persistence varied with network motif profiles (Appendix \emph{3.3}), 4) how network motif profiles varied with global structural properties (Appendix \emph{3.4}), and 5) how persistence varied with other network properties (Appendix \emph{3.5}).
-The analyses are briefly described below.
+	We tested 1) how species persistence varied with motif participation (Appendix \emph{3.1}), 2) how motif participation varied with other network properties (Appendix \emph{3.2}), 3) how network mean persistence varied with network motif profiles (Appendix \emph{3.3}), 4) how network motif profiles varied with global structural properties (Appendix \emph{3.4}), and 5) how species persistence and network mean persistence varied with other network properties (Appendix \emph{3.5}).
+The analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan} as detailed in the appendices.
 
@@
 
-        where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
+        \noindent where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
         All predictors were centered and scaled before fitting the models.         
-        To test whether these general trends were consistent across networks, we performed additional linear regressions of persistence against motif participation within each simulated network, for each level of disturbance to basal resources (600 regressions).
-        We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.
+        To test whether these general trends were consistent across networks, we examined the distribution of slopes of simplified regressions fit to a single network and level of disturbance (Appendix \emph{3.1}).
 
@@
 
-        To test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
-        In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMMs:
+        We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} (Appendix \emph{3.2}). 
+        To identify correlations between size, connectance, and participation in particular motifs, we fit an additional set of four LMMs:
 
         \begin{equation}
-            \rho_{i} \approx S_{i} + C_{i} + S_{i}C_{i}
+            \rho_{i} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i} ,
+            \label{partic_SC}
         \end{equation}
-        where all symbols are as in equation~\ref{propreq}.
+
+        where $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif,
+        $\Sigma_{i}$ is the size of the network containing species $i$, and $\zeta_{i}$ is the connectance of the network containing species $i$.
+
+        We fit a similar set of eight LMMs relating motif participation to in-degree or trophic level:
+
+        \begin{equation}
+            \rho_{i} \approx \delta_{i} + S_{i}C_{i} ,
+            \label{partic_deg}
+        \end{equation}
+
+        \begin{equation}
+            \rho_{i} \approx \tau_{i} + S_{i}C_{i} ,
+            \label{partic_deg}
+        \end{equation}
+
+        where $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, and all other symbols are as in equation~\ref{propreq}.
 

### 2021-09-29 manuscript/supplement_motifs.tex
still syncing SI and main

@@
         Significance was calculated using 9999 permutations.
-        The PERMANOVA was fit using the R~\citep{R} function `adonis' from the package \emph{vegan}~\citep{vegan}.
+
+        PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
+        To test whether this applies in our case, we fit an ANOVA of the dispersion of motif participation roles about the centroid for each combination of network size and connectance. 
+        We then regressed dispersion against persistence to test whether networks with higher or lower persistence had more consistent motif profiles.
+        We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
+        We fit the regression using the R~\citep{R} base function `lm'.
 
@@
         To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
+        As with the PERMANOVA relating motif participation to network properties, we fit the PERMANOVA using the R~\citep{R} function `vegan' from the package \emph{adonis}~\citep{adonis} and calculated significance based on 9999 permutations.
+
         We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
+        Dispersion was calculated using the R~\citep{R} function `betadisper' from the package \emph{vegan}~\citep{vegan}.
         In case high or low mean persistences are associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable).
-        All statistical tests were conducted as when relating motif profiles to global network structure.
 
-        To supplement these overall tests, we also fit four regressions testing whether persistence had a linear relationship to the proportion of a network profile made up of each motif, and whether this varied with the level of disturbance.
-        Each regression included the proportion of a motif, disturbance, and their interaction as predictors.
+        To supplement these overall tests, we also fit four linear models relating mean persistence to network motif profiles, disturbance, and the interaction between them:
+
+            \begin{equation}
+                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{disturbed_k} + \Bar{\rho}_{i}\pi_{disturbed_k} ,
+                \label{netpropeq}
+            \end{equation}
+
+        where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and all other symbols are as in equation~\ref{propreq}. 
         All regressions were fit using the R~\citep{R} base function `lm'.
@@
         We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
-        To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif to network size, connectance, and their interaction.
+
+        To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif to network size, connectance, and their interaction:
+
+        \begin{equation}
+            \Bar{\rho_{i}} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i}
+        \end{equation}
+        where all symbols are as in equations~\ref{propreq} and~\ref{netpropeq}.
         We fit each regression using the R~\citep{R} base function `lm'.
@@
     \subsection{How does persistence vary with species and network properties?}
+
+        S, C for random effects. $\Sigma$ and $\zeta$ for fixed effects.
 

### 2021-09-29 manuscript/reorder_main_text.tex

@@
         We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} (Appendix \emph{3.2}). 
-        To identify correlations between size, connectance, and participation in particular motifs, we fit an additional set of four LMMs:
+        To identify correlations between size, connectance, and participation in particular motifs, we fit a set of four linear models (LMs):
 
@@
         As well as species-specific probabilities of persistence, motifs may provide information about the overall probability of any species going extinct (since they incorporate information on indirect effects).
-        To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
-        To identify the relationships between particular motifs and network mean persistence, we fit a set of four LMMs (one per motif):
+        We tested whether networks with similar motif profiles tend to have similar mean persistences among consumers using a PERMANOVA (Appendix \emph{S3.3}.
+        To identify the relationships between particular motifs and network mean persistence, we fit a set of four LMs (one per motif):
 
             \begin{equation}
-                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{disturbed_k} + \Bar{\rho}_{i}\pi_{disturbed_k} +
-                S_{i}C_{i} ,
+                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{disturbed_k} + \Bar{\rho}_{i}\pi_{disturbed_k},
                 \label{netpropeq}
@@
         Different scales of network structure (e.g., local and meso-scale) are not independent. 
-        To test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
-        In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMMs:
+        To test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
+        In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMs:
 
         \begin{equation}
-            \rho_{i} \approx S_{i} + C_{i} + S_{i}C_{i}
+            \Bar{\rho_{i}} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i}
         \end{equation}
-        where all symbols are as in equation~\ref{propreq}.
+        where all symbols are as in equations~\ref{propreq} and~\ref{netpropeq}.
 
     \subsubsection*{5) Persistence and network properties}
+        $\Sigma$ and $\zeta$ for fixed effects, SC for random.
 

### 2021-09-29 manuscript/supplement_motifs.tex
methods synced

@@
 
-        S, C for random effects. $\Sigma$ and $\zeta$ for fixed effects.
-
         \subsubsection{How does network mean persistence vary depending on network properties?}
@@
 
-        \subsubsection{How does persistence vary with degree and trophic level?}
+            \begin{equation}
+                \Psi_{ik} \approx \Sigma_{i} + \zeta_{i} + \pi_{disturbed_k} + \Sigma_{i}:\zeta_{i} + \Sigma_{i}\pi_{disturbed_k} + \zeta_{i}\pi_{disturbed_k} + \Sigma_{i}\zeta_{i}\pi_{disturbed_k},
+                \label{SCeq}
+            \end{equation}
+            where all symbols are as in equation~\ref{propreq} and~\ref{partic_SC}.
 
-            A species’ motif participation is likely related to its in-degree and trophic level since all three properties are measures of network position.
-To provide context for the results above, we therefore fit two linear mixed-effects regressions relating persistence to degree or trophic level, $\pi_{disturbed}$, and their interaction.
-As above, we centered and scaled the predictors.
-We also included a random effect of the interaction between network size and connectance in order to account for the differences in persistence between different networks.
+        \subsubsection{How does persistence vary with in-degree and trophic level?}
+
+            Likewise, in-degree and trophic level are known to influence persistence in many cases.
+            These effects, in combination with differences in mean persistence across networks with different global properties, should also be taken into account when considering relationships between motif participation and persistence.
+            To establish this necessary context, we therefore fit two linear mixed-effects regressions:
+
+            \begin{equation}
+                    \Psi_{ik} \approx \delta_{i} + \pi_{disturbed_k} + \delta_{i}\pi_{disturbed_k} +
+                    S_{i}C_{i} ,
+                    \label{degeq}
+                \end{equation}
+
+            \begin{equation}
+                    \Psi_{ik} \approx \tau_{i} + \pi_{disturbed_k} + \tau_{i}\pi_{disturbed_k} +
+                    S_{i}C_{i} ,
+                    \label{TLeq}
+                \end{equation}
+
+            where $\delta_{i}$ and $\tau_i$ are the in-degree and trophic level of species $i$, respectively, and all other terms are as in equation~\ref{propreq}.
+            We centered and scaled the predictors before fitting the models.
 

### 2021-09-29 manuscript/reorder_main_text.tex

@@
     \subsubsection*{5) Persistence and network properties}
-        $\Sigma$ and $\zeta$ for fixed effects, SC for random.
 
         Persistence is known to vary with global-scale and local-scale network properties.
-        To correctly interpret any relationships between motifs and persistence, we must consider the context of these other relationships.
+        To correctly interpret any relationships between motifs and persistence, we must consider the context of these other relationships (Appendix \emph{S3.5}).
         To identify the relationship between global network structure and mean probability of persistence within a network, we fit a linear regression:
+
             \begin{equation}
-                \Psi_{ik} \approx S_{i} + C_{i} + \pi_{disturbed_k} + S_{i}:C_{i} + S_{i}\pi_{disturbed_k} + C_{i}\pi_{disturbed_k} + S_{i}C_{i}\pi_{disturbed_k},
+                \Psi_{ik} \approx \Sigma_{i} + \zeta_{i} + \pi_{disturbed_k} + \Sigma_{i}:\zeta_{i} + \Sigma_{i}\pi_{disturbed_k} + \zeta_{i}\pi_{disturbed_k} + \Sigma_{i}\zeta_{i}\pi_{disturbed_k},
                 \label{SCeq}
             \end{equation}
-        where all symbols are as in equation~\ref{propreq}.
+
+        where all symbols are as in equation~\ref{propreq} and~\ref{partic_SC}.
         As the three-way interaction was significant, we did not simplify the model.
@@
         Persistence is also known to vary with local-scale measures of network structure. 
-        To identify the relationships betwen local-scale properties and species' probabilities of persistence, we fit two more LMMs:
+        To identify the relationships betwen local-scale properties and species' probabilities of persistence, we fit a final two LMMs:
 

### 2021-10-01 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 \usepackage{pdflscape}
+\usepackage{amsmath}
 \usepackage[labelfont=bf]{caption} 
@@
            %\subsection*{Here we test...}
-    Specifically, we are interested in whether a consumer's extinction risk depends on 1) the motif profile of the full food web and 2) the motifs in which a focal species participates. 
+    Specifically, we are interested in whether a consumer's extinction risk depends on 1) the motifs in which a focal species participates and 2) the motif profile of the full food web. 
     In addition, we want to understand how the information provided by motifs is related to classical global and local food web properties in order to better synthesize information provided by different scales of network structure.
@@
 	\subsection*{Calculating persistence}	
+    % basic idea behind BN modeling
+		We use Bayesian Network modeling to calculate the persistence of each species in our simulated networks. 
+		A Bayesian Network represent a graphical model for probabilistic relationships among a set of variables.
+The nodes of the network represents the variables, and the links represent conditional dependencies among the variables.
+Each node has a set of local probability dependencies, conditional on the state of the parent nodes.
+		In our ecological setting, species act as nodes and is a Bernoulli random variable, being either extant or extinct. The links act as feeding interactions between the nodes.
+		The extinction probability of each species is a function of the state of its resources.
 
-		First, we assigned all species a baseline probability of extinction ($\pi_{base}$) -- the risk of going extinct given factors not related to the food web itself. 
-		For consumer species, this also represents extinction despite all resources being present. 
-		Such events can, for example, reflect diseases or stochastic extinctions of small populations.
+		% probability of extinction equation
+		Consider a consumer species with several resource species $n$.
+The extinction probability of this consumer will depend on both the total state of its resource species $n$, and on a baseline extinction probability of extinction, $\pi_{base}$.
+$\pi_{base}$ represent the risk of going extinct given factors not related to the food web itself, and could reflect diseases or stochastic extinctions of small populations.
+		All the species in the network retain a baseline probability of extinction, which can be different to reflect sensitivity of some groups of species.
+		If a consumer looses some resources $k$, the fraction of resources lost $f$ will be $f = (k/n)$.
+Note that we here make a strong assumption, where we only care about the fraction of resources lost and not the identity of the resources species.
 
-		As well as this baseline scenario, we simulated a set of scenarios where primary producers had a higher baseline probability of extinction ($\pi_{disturbed}$). 
+		The response of a consumer to the loss of a fraction of its resources can be modelled in several biologically plausible ways. For example, the response could be linear or non-linear.
+		A sigmoid non-linear functional response of consumers to the loss of resources ($\alpha >1, \beta >1$) has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
+		Therefore, we use a cumulative density function of a beta distribution to get each species \textit{i} probability of extinction:
+		\begin{equation}
+		\label{betafunc}
+		P(\lnot i|f) = \pi_{base} + (1 - \pi_{base}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}
+		\end{equation}
+
+		If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability will be $P(\lnot C|f) = \pi_{base}$. 
+		Similarly, if all resources are extinct ($f = 1$), the extinction probability $P(\lnot C|f)$ will be equal to 1 and the consumer will go extinct.
+		If the $f$ is not 1 nor 0, the non-linear sigmoid curve will determine the response of the consumer, which will be more severe if the fraction is high and less severe if the fraction is small.
+This will in turn affect the extinction probability $P(\lnot C|f)$.
+		For a basal species, we assume that abiotic resources are always available. Therefore, a primary producer is independent of resource species and will always have $f = 0$.
+
+		% how we calculate persistences 
+		As consumer species depends on its resource species, we start determining the status of each primary producer in the network.
+	    Systematically, we use eqn \ref{betafunc} to receive each species probability of extinction $P(\lnot i|f)$, and next we perform Bernoulli trials. 
+	    Imagine a basal species \textit{A} with extinction probability $P(\lnot A) = \pi_{base}$ where $\pi_{base} = 0.1$.
+We draw a random number $r$ from a uniform distribution between $0-1$.
+The basal species would be considered extinct if $r < 0.1$, and extant otherwise.
+		We use this information when performing a Bernoulli trial on a consumer \textit{C} connected to this basal species, inserting our information of the $\pi_{base}$ and fraction of resources lost $f$ of this consumer, into eqn \ref{betafunc} to
+	    receive $P(\lnot C|f)$.
+In the same manner as for the basal species above, we compare $P(\lnot C|f)$ with a random number $r$, and if $r < P(\lnot C|f)$ the consumer species is considered extinct.
+	    In this way, we evaluate each species existence, starting from the bottom and working our way up.
+	    As this method is probabilistic, we repeat these calculations 100 000 times per network, with unique random draws.
+	    For each species, we then estimate the fraction of simulations in which the species persist, and this value is the approximated probability of persistence.
+	    To approximate persistence probabilities through numerical simulation are shown to be highly efficient and produce the same result as exact methods of solving Bayesian networks \citep{Haussler2020}.
+
+    % various disturbance scenarios
+        In our baseline scenario, primary producers had a baseline extinction probability of $\pi_{base} = 0.1$. 
+		We also simulated a set of scenarios where primary producers had a higher baseline probability of extinction ($\pi_{disturbed}$). 
 		These scenarios reflect threats to primary producers such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
@@
 
-		As well as the intrinsic baseline probability of extinction, extinction risk for consumers is affected by access to resources. 
-		When all resources are extant, the probability of extinction for a consumer $i$, $P(\lnot i)$, is equal to the baseline probability; $P(\lnot i|X_{1},...,X_{n}) = \pi_{base}$. 
-		When all resources are extinct, the consumer will go extinct; $P(\lnot i|\lnot X_{1},...,\lnot X_{n})=1$.
+		%As well as the intrinsic baseline probability of extinction, extinction risk for consumers is affected by access to resources. 
+		%When all resources are extant, the probability of extinction for a consumer $i$, $P(\lnot i)$, is equal to the baseline probability; $P(\lnot i|X_{1},...,X_{n}) = \pi_{base}$. 
+		%When all resources are extinct, the consumer will go extinct; $P(\lnot i|\lnot X_{1},...,\lnot X_{n})=1$.
 
-		Calculating $P(i)$ or $P(\lnot i)$ is less straightforward when a consumer $i$ loses a fraction $k/n = f$ of its prey. 
-		We used a sigmoid non-linear functional response of consumers to the loss of resources (Appendix \emph{S2}), which has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
-		Probabilities of extinction for resource species are used when calculating probabilities for higher level species.
+		%Calculating $P(i)$ or $P(\lnot i)$ is less straightforward when a consumer $i$ loses a fraction $k/n = f$ of its prey. 
+		%We used a sigmoid non-linear functional response of consumers to the loss of resources (Appendix \emph{S2}), which has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
+		%Probabilities of extinction for resource species are used when calculating probabilities for higher level species.
 
-		Although it is possible to solve the Bayesian network exactly \citep{Eklof2013}, this is cumbersome for larger networks. 
-		As numerical simulations are highly efficient and produce the same result \citep{Haussler2020}, we also use simulations here.
-		To compute $P(\lnot i|f)$, we calculated the fraction of resources lost ($f$) for each consumer $i$.
-		Persistence is determined by comparing $P(\lnot i|f)$ with a randomly drawn number from a uniform distribution between 0-1. 
-		If this random number is lower than $P(\lnot i|f)$, the species is considered extinct. 
-		The species' marginal probability of persistence is simply $P(i) = 1-P(\lnot i)$.
-		To minimize noise due to the random draws, we repeat these calculations 100 000 times per web, with unique random draws when calculating $P(i)$.
-		The final persistence for each species is the mean over these runs.
+		%Although it is possible to solve the Bayesian network exactly \citep{Eklof2013}, this is cumbersome for larger networks. 
+		%As numerical simulations are highly efficient and produce the same result \citep{Haussler2020}, we also use simulations here.
+		%To compute $P(\lnot i|f)$, we calculated the fraction of resources lost ($f$) for each consumer $i$.
+		%Persistence is determined by comparing $P(\lnot i|f)$ with a randomly drawn number from a uniform distribution between 0-1. 
+		%If this random number is lower than $P(\lnot i|f)$, the species is considered extinct. 
+		%The species' marginal probability of persistence is simply $P(i) = 1-P(\lnot i)$.
+		%To minimize noise due to the random draws, we repeat these calculations 100 000 times per web, with unique random draws when calculating $P(i)$.
+		%The final persistence for each species is the mean over these runs.
 
@@
         To separate differences in motif structure from differences in network size (larger networks contain more motifs), we normalized motif profiles by dividing the count of each motif by the total count of all motifs.
-        % The motif profile for each network therefore describes the proportions of each motif and has a sum of exactly one.
 		We calculated species' motif participation using the Python package \emph{pymfinder}~\citep{pymfinder}.

### 2021-10-01 manuscript/supplement_motifs.tex
added Annas raw text for persistence calc

@@
 
-    \subsection{Stochastic extinction}
+        We use Bayesian Network modelling to calculate the persistence of each species in our simulated networks. 
+        A Bayesian Network represents a graphical model for probabilistic relationships among a set of variables.
+The nodes of the network represents the variables, and the links represent conditional dependencies among the variables.
+Each node has a set of local probability dependencies, conditional on the state of the parent nodes.
+        In our ecological setting, species act as nodes and are Bernoulli random variables, being either extant or extinct. The links act as feeding interactions between the nodes.
+        The extinction probability of each species is a function of the state of its resources.
+
+    \subsection{Baseline probability of extinction}
+
+        Consider a consumer species with several resource species $n$. 
+        The extinction probability
+        of this consumer will depend on both the total state of its resource species $n$, and on a
+        baseline extinction probability of extinction, $\pi_{base}$. 
+        $\pi_{base}$ represents the risk of going extinct
+        given factors not related to the food web itself, and could reflect diseases or stochastic
+        extinctions of small populations. 
+        All the species in the network retain a baseline
+        probability of extinction, which can be different to reflect greater sensitivity in some groups of species.
+
+    \subsection{Extinction due to food-web effects}
+
+        If a consumer with $n$ resources loses $k$ of them, the fraction of resources lost $f$ will be
+        $f = (k/n)$. 
+        Note that we here make a strong assumption, where we only care about the fraction of resources lost and not the identity of the resources species.
 
@@
 
-        As well as the intrinsic baseline probability of extinction, extinction risk for consumers is affected by access to resources. 
-		When all resources are extant, the probability of extinction for a consumer $i$, $P(\lnot i)$, is equal to the baseline probability; $P(\lnot i|X_{1},...,X_{n}) = \pi_{base}$. 
-		When all resources are extinct, the consumer will go extinct; $P(\lnot i|\lnot X_{1},...,\lnot X_{n})=1$.
-
-		Calculating $P(i)$ or $P(\lnot i)$ is less straightforward when a consumer $i$ loses a fraction $k/n = f$ of its prey. 
-		Plausible ecological responses to losing some prey include i) a topological response, where the consumer's probability of extinction is constant unless all resources are lost, ii) a linear response, in which the extinction probability increases linearly with resources loss, and iii) several nonlinear responses where the probability of extinction increases non-linearly with the fraction resource lost.
-		The extinction probability will differ depending on the response function. 
-		We use a sigmoid non-linear functional response of consumers to the loss of resources ($\alpha >1, \beta >1$) which has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
-		We use a cumulative density function of a beta distribution for this purpose:
+        The response of a consumer to the loss of a fraction of its resources can be modelled in several biologically plausible ways. For example, the response could be linear or non-linear.
+        A sigmoid non-linear functional response of consumers to the loss of resources ($\alpha >1, \beta >1$) has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
+        Therefore, we use a cumulative density function of a beta distribution to get each species \textit{i} probability of extinction:
         \begin{equation}
         \label{betafunc}
-		P(\lnot i|f) = \pi_{i} + (1 - \pi_{i}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}
+        P(\lnot i|f) = \pi_{base} + (1 - \pi_{base}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}
         \end{equation}
 
-    \subsection{Numerical simulation}
+        If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability will be $P(\lnot C|f) = \pi_{base}$. 
+        Similarly, if all resources are extinct ($f = 1$), the extinction probability $P(\lnot C|f)$ will be equal to 1 and the consumer will go extinct.
+        If the fraction $f$ is not 1 nor 0, the non-linear sigmoid curve will determine the response of the consumer, which will be more severe if the fraction resources lost is high and less severe if the fraction is small.
+This will in turn affect the extinction probability $P(\lnot C|f)$.
+        For a basal species, we assume that abiotic resources are always available. Therefore, a primary producer is independent of resource species and will always have $f = 0$.
+
+        % how we calculate persistences 
+        As consumer species depends on its resource species, we start determining the status of each primary producer in the network.
+        Systematically, we use eqn \ref{betafunc} to receive each species probability of extinction $P(\lnot i|f)$, and next we perform Bernoulli trials to determine whether the species goes extinct or not.
+        Imagine a basal species \textit{A} with extinction probability $P(\lnot A) = \pi_{base}$ where $\pi_{base} = 0.1$.
+We draw a random number $r$ from a uniform distribution between $0-1$.
+The basal species would be considered extinct if $r < 0.1$, and extant otherwise.
+        We use this information when performing a Bernoulli trial on a consumer \textit{C} connected to this basal species, inserting our information of the $\pi_{base}$ and fraction of resources lost $f$ of this consumer, into eqn \ref{betafunc} to
+        receive $P(\lnot C|f)$.
+In the same manner as for the basal species above, we compare $P(\lnot C|f)$ with a random number $r$, and if $r < P(\lnot C|f)$ the consumer species is considered extinct.
+        In this way, we evaluate each species existence, starting from the bottom and working our way up.
+        As this method is probabilistic, we repeat these calculations 100 000 times per network, with unique random draws.
+        For each species, we then estimate the fraction of simulations in which the species persist, and this value is the approximated probability of persistence.
+        To approximate persistence probabilities through numerical simulation are shown to be highly efficient and produce the same result as exact methods of solving Bayesian networks \citep{Haussler2020}.
+
+    % various disturbance scenarios
+        In our baseline scenario, primary producers had a baseline extinction probability of $\pi_{base} = 0.1$. 
+        We also simulated a set of scenarios where primary producers had a higher baseline probability of extinction ($\pi_{disturbed}$). 
+        These scenarios reflect threats to primary producers such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
+        $\pi_{disturbed}$ ranged between $0.1-0.5$, in steps of $0.08$. 
+        The highest disturbance level, $\pi_{disturbed} = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
+        Consumer species retained $\pi_{base}=0.1$ in all cases.
 

### 2021-10-01 manuscript/reorder_main_text.tex

--- 2021-09-29 manuscript/reorder_main_text.tex
@@
 		New networks were simulated to replace any removed networks, and this process was repeated until we obtained 100 suitable networks in each combination of size and connectance.
-		All networks were rendered acyclic following~\citet{Allesina2009} and topologically sorted from lowest to highest trophic level (Appendix\emph{S1}).
+		All networks were rendered acyclic following~\citet{Allesina2009} and topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).
 		These processing steps allow the strictly bottom-up calculation of probability of persistence used in the Bayesian network framework.
@@
 	\subsection*{Calculating persistence}	
+    % basic idea behind BN modeling
+        Bayesian Network modelling represents species and feeding interactions as the nodes and links of a network (see Appendix \emph{S2} for full details).
+        Each species (node) is a Bernoulli random variable, being either extant or extinct. 
+        The probability that a particular species is extinct is a function of the states of its resources.
 
-		First, we assigned all species a baseline probability of extinction ($\pi_{base}$) -- the risk of going extinct given factors not related to the food web itself. 
-		For consumer species, this also represents extinction despite all resources being present. 
-		Such events can, for example, reflect diseases or stochastic extinctions of small populations.
+		% probability of extinction equation
+		Consider a species with $n$ resource species. 
+		Before taking food-web structure into account, this species already has a baseline extinction  probability of extinction, $\pi_{base}$. 
+		$\pi_{base}$ represents the risk of going extinct due to diseases, stochastic extinction of small populations, or other factors not related to the food web.
+		$\pi_{base}$ may be constant, or it can vary to reflect greater sensitivity in some groups of species.
+		In our simulations, we vary $\pi_{base}$ for basal resources between 0.1 and 0.5 but keep constant $\pi_{base}=0.1$ for consumers.
 
-		As well as this baseline scenario, we simulated a set of scenarios where primary producers had a higher baseline probability of extinction ($\pi_{disturbed}$). 
+        If the focal species is a consumer with $n$ resources, its probability of extinction depends on $\pi_{base}$ and on the fraction $f$ of its resources that have gone extinct.
+        Note that we assume that only the fraction of resources lost is important and not the identities of the missing resources.
+        We then model the response of the consumer to changes in $f$ using a sigmoid non-linear function ($\alpha > 1, \beta > 1$; Appendix \emph{S2}.
+		This type of response most accurately captures the secondary extinctions produced by dynamical models~\citep{Eklof2013}.
+		The probability of extinction for each species $i$ can therefore be represented using the cumulative density function of a beta distribution:
+		\begin{equation}
+		\label{betafunc}
+		P(\lnot i|f) = \pi_{base} + (1 - \pi_{base}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}.
+		\end{equation}
+
+		If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability will be $P(\lnot C|f) = \pi_{base}$. 
+		Similarly, if all resources are extinct ($f = 1$), the extinction probability $P(\lnot C|f)$ will be equal to 1 and the consumer will go extinct.
+		If the fraction $f$ is not 1 nor 0, the non-linear sigmoid curve will determine the response of the consumer, which will be more severe if the fraction resources lost is high and less severe if the fraction is small.
+This will in turn affect the extinction probability $P(\lnot C|f)$.
+		For basal species, we assume that abiotic resources are always available and $f=0$, such that extinction depends only on $\pi_{base}$.
+
+		% how we calculate persistences 
+        Beginning with primary producers and working systematically upward through the food web, we use Equation \ref{betafunc} to calculate each species' probability of extinction $P(\lnot i|f)$.
+        Using this probability, we then perform Bernoulli trials to determine whether the species goes extinct or not. 
+        Species are considered extinct if a random number drawn from a uniform distribution $[0-1]$ is less than $P(\lnot i|f)$.
+
+        These simulated extinctions are used when calculating the fraction of resources lost $f$ for each consumer.
+        This allows us to update Equation~\ref{betafunc} with information on extinctions of a focal consumer's resources and simulate bottom-up effects.
+As this method is probabilistic, we repeat these calculations 100 000 times per network, with unique random draws.
+	    For each species, we then define probability of persistence as the fraction of simulations in which the species persists.
+	    This process is highly efficient and produces the same result as exact methods of solving Bayesian networks \citep{Haussler2020}.
+
+    % various disturbance scenarios
+        In our baseline scenario, primary producers had a baseline extinction probability of $\pi_{base} = 0.1$. 
+		We also simulated a set of scenarios where primary producers had a higher baseline probability of extinction ($\pi_{disturbed}$). 
 		These scenarios reflect threats to primary producers such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
-		$\pi_{disturbed}$ ranged between $0.1-0.5$, in steps of $0.08$. 
+		$\pi_{disturbed}$ ranged between $0.18-0.5$, in steps of $0.08$. 
 		The highest disturbance level, $\pi_{disturbed} = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
 		Consumer species retained $\pi_{base}=0.1$ in all cases.
-
-		As well as the intrinsic baseline probability of extinction, extinction risk for consumers is affected by access to resources. 
-		When all resources are extant, the probability of extinction for a consumer $i$, $P(\lnot i)$, is equal to the baseline probability; $P(\lnot i|X_{1},...,X_{n}) = \pi_{base}$. 
-		When all resources are extinct, the consumer will go extinct; $P(\lnot i|\lnot X_{1},...,\lnot X_{n})=1$.
-
-		Calculating $P(i)$ or $P(\lnot i)$ is less straightforward when a consumer $i$ loses a fraction $k/n = f$ of its prey. 
-		We used a sigmoid non-linear functional response of consumers to the loss of resources (Appendix \emph{S2}), which has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
-		Probabilities of extinction for resource species are used when calculating probabilities for higher level species.
-
-		Although it is possible to solve the Bayesian network exactly \citep{Eklof2013}, this is cumbersome for larger networks. 
-		As numerical simulations are highly efficient and produce the same result \citep{Haussler2020}, we also use simulations here.
-		To compute $P(\lnot i|f)$, we calculated the fraction of resources lost ($f$) for each consumer $i$.
-		Persistence is determined by comparing $P(\lnot i|f)$ with a randomly drawn number from a uniform distribution between 0-1. 
-		If this random number is lower than $P(\lnot i|f)$, the species is considered extinct. 
-		The species' marginal probability of persistence is simply $P(i) = 1-P(\lnot i)$.
-		To minimize noise due to the random draws, we repeat these calculations 100 000 times per web, with unique random draws when calculating $P(i)$.
-		The final persistence for each species is the mean over these runs.
 

### 2021-10-01 manuscript/supplement_motifs.tex
synced order of appendices with main text

@@
 
-\section{Network creation and processing}
+\section{Supplemental methods: Network creation}
 
@@
 
-\section{Calculating persistence}
+\clearpage
+
+\section{Supplemental methods: Calculating persistence}
 
@@
 
-\section{Statistical analysis}
+\clearpage
+
+\section{Supplemental methods: Statistical analysis}
 
@@
 
-        As with network profiles, species' motif participation may covary with other measures of network structure and species' roles within the network.
-        To establish this context for motif participation, we first fit a PERMANOVA~\citep{Anderson2001} to test whether a species' overall motif participation role is related to network size, connectance, or the interaction between them.
-        To avoid inflating similarity between species which do not appear in the same motifs (i.e., which have ``shared zeros''), we use Bray-Curtis dissimilarity between motif participation roles.
-        Significance was calculated using 9999 permutations.
-
-        PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
-        To test whether this applies in our case, we fit an ANOVA of the dispersion of motif participation roles about the centroid for each combination of network size and connectance. 
-        We then regressed dispersion against persistence to test whether networks with higher or lower persistence had more consistent motif profiles.
-        We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
-        We fit the regression using the R~\citep{R} base function `lm'.
-
-        To obtain a more detailed picture of the relationships between motif participation and global network structure, we fit a set of four linear models (one per motif):
+        Species' motif participation may covary with other measures of network structure and species' roles within the network.
+        To determine the relationships between motif participation and global network structure, we fit a set of four linear models (one per motif):
 
@@
 
-    \subsection{How does persistence vary with species and network properties?}
+    \subsection{How does persistence vary with other properties?}
 
-        \subsubsection{How does network mean persistence vary depending on network properties?}
+        \subsubsection{Network mean persistence and network properties}
 
@@
 
-        \subsubsection{How does persistence vary with in-degree and trophic level?}
+        \subsubsection{Species persistence and species properties}
 
@@
             We centered and scaled the predictors before fitting the models.
+
+\clearpage
+
+\section{Results: Persistence and motif participation}
+
+    Persistence varied with motif participation.
+
+    \begin{table}[h!]
+        \centering
+        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{propreq}).
+Predictors were centered and scaled before model fitting.
+We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
+        \label{tab:proportion}                \footnotesize
+        \begin{tabular}{c|c c | c c | c c | c c | c c |}
+        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c|}{$R^2$} \\
+        Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
+        \hline
+        3-sp Chain & 0.574 & \textless0.001 & -0.00508 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
+        App. Com. & 0.574 & \textless0.001 & -0.00312 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
+        Dir. Com. & 0.574 & \textless0.001 & 0.00564 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
+        Omnivory & 0.574 & \textless0.001 & 0.00709 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
+        \end{tabular}
+    \end{table}        
+\clearpage
+
+\section{Results: Motif participation and network properties}
+
+    Motif participation varies with global-scale network properies.
+
+    \begin{table}[hb!]
+        \centering
+        \caption{Motif participation varies with size and connectance. Here we give the range and mean ($\pm$ standard deviation) proportion of each motif in species' motif participation vectors. 
+        We also show the effects ($\beta$) of network size, connectance, and their interaction in linear models predicting the proportion of each motif in a participation vector.
+        All effects were significant ($p$\textless0.001).
+        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
+        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
+        \label{tab:partic_vs_SC}   
+        \footnotesize
+        \begin{tabular}{c|c c c | c c c}
+            Motif & Min & Max & Mean (SD) & Size & Connectance & Interaction \\
+            \hline
+            3-sp Chain & 0 & 1 & 0.249 (0.127) & 2.22$\times10^{-4}$ & -0.131 & -3.20$\times10^{-3}$ \\
+            App. Comp & 0.250 & 0.765 & 0.422 (0.0645) & -5.06$\times10^{-4}$ & -1.00 & 2.59$\times10^{-3}$ \\
+            Dir. Comp & 0.0501 & 0.359 & 0.179 (0.0407) & -3.56$\times10^{-5}$ & -0.556 & 5.92$\times10^{-3}$ \\
+            Omnivory & 0.006 & 0.346 & 0.164 (0.0790) & 3.64$\times10^{-4}$ & 1.71 & -5.24$\times10^{-3}$ \\   
+            \hline
+            \end{tabular}
+            \end{table}
+
+    \begin{figure}[h!]
+        \centering
+        \includegraphics[height=.5\textheight]{figures/participation_vs_SC.eps}
+        \caption{The proportion of each motif in a consumer's participation vector varied with network size, connectance, and their interaction.
+These relationships were steepest for apparent competition and omnivory.}
+        \label{fig:roles_vs_SC}
+    \end{figure}
+\clearpage
+
+\section{Results: Persistence, degree, and trophic level}
+
+    Persistence varies with in-degree, trophic level, and their interaction.
+
+    \begin{table}[h!]
+        \caption{Coeffcients ($\beta$) for regressions of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction.
+Predictors were centered and scaled before model fitting.
+All $p$-values were \textless0.001.
+The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.}
+        \label{tab:per_vs_TLdeg}
+        \centering
+        \begin{tabular}{l|c  c  c |}
+            Simple role measure & Role & Disturbance & Interaction \\
+            \hline
+            STL & -0.0218 & -0.139 & 0.00508 \\
+            in-degree & -0.00130 & -0.139 & -0.00908 \\
+        \end{tabular}
+    \end{table}
+
+    \begin{figure}[h!]
+     \centering
+     % \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}
+     \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
+The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
+Persistence decreases with darker color in the heat map.}
+     \label{fig:heatmap_stl_BP0}
+    \end{figure}
+
+    \begin{figure}[h!]
+     \centering
+     % \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
+     \caption{Primary producers experience a disturbance reducing their probability to go extinct to $\pi_{disturbed} = 0.5$ All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
+The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
+Persistence decreases with darker color in the heat map.}
+     \label{fig:heatmap_stl_BP1}
+    \end{figure}
+
+\clearpage
+
+\section{Results: Motif-persistence relationships}
+
+    When comparing the outcomes from linear mixed-effect models (LMMs) relating persistence to motif roles across all networks (Fig.
+3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately, the main trends are similar but different-sized networks showed slightly different results results for some motifs (Figure temporarily in main text).%(Figure~\ref{fig:density_prop_S}).
+    For the three-species chain and omnivory motifs, a significant interaction between disturbance and the effect of the focal motif means that having a larger proportion of the role be made up of the chain or omnivory motif is associated with lower persistence at high levels of disturbance (Fig.
+3 and Table 1, \emph{Main Text}).
+    The same pattern holds overall for LRs fit within a network  %(Figure~\ref{fig:density_prop_S})
+    : the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance. 
+    This trend is most pronounced in large networks.% (bottom panel, Figure~\ref{fig:density_prop_S}). 
+    Besides the decrease in fraction of positive slopes with increasing disturbance, the omnivory motif show a higher overall fraction of positive slopes in comparison to the three-species chain motif, corresponding to the positive main effect of omnivory (Fig.
+3 and Table 1, \emph{Main Text}).
+
+    \begin{figure}[h!]
+         \centering
+         % \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
+        \caption{Density (y-axis) of slopes (x-axis) for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
+The slopes are derived from linear mixed-effect models, with a random intercept for the interaction between species richness and network size.
+A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
+The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to motif (legend).
+Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
+Rows show various levels of network size.
+The vertical line indicate zero on the x-axis.}
+        \label{fig:density_prop_S}
+    \end{figure}
+
+    For the apparent and direct competition motifs, the LMMs include a positive interaction between disturbance and the proportion of the motif in a species' role (Fig.
+3 and Table 1; \emph{Main Text}).
+    This means that having more of the role made up of these two motifs is increasingly beneficial at higher levels of disturbance, a trend which is also evident from the increasing fraction of positive slopes with increasing disturbance based on LRs of each network at each level of disturbance.
+% (Figure~\ref{fig:density_prop_S}).
+    As with the trends for omnivory and three-species chains, these increases are strongest in large networks. % (bottom row, Figure~\ref{fig:density_prop_S}).
+    % For direct competition, the fraction of slopes are similar across disturbance levels.
+Although this interaction is not positive when accounting for network size, the fraction of positive slopes are throughout high.
+    The increase in the proportion of positive slopes was much stronger for apparent competition than direct competition, consistent with the smaller (but still significant) interaction term for between direct competition in the LMMs.
+    A high proportion of direct competition in a species' role tended to be associated with greater probability of persistence across levels of disturbance and network size.
+    % In conclusion, species richness does not considerably alter the trends visible in the LMMs.
+High species richness more clearly display an interaction between motif proportion, persistences and disturbance levels, while this effect is smaller with low species richness.
+\clearpage
+
+\section{Results: Persistence and motif profiles}
+
+    Persistence varied with the frequencies of different motifs in a network profile.
+
+    \begin{table}[h!]
+        \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
+Predictors were centered and scaled before model fitting.
+All effects were significant ($p$\textless0.001).
+We also provide the $R^2$.
+}
+        \label{motif_profile_tab}
+        \centering
+        \begin{tabular}{c|c c c c c | c}
+            Motif & Intercept & Proportion & Disturbance & Interaction & $R^2$\\
+            \hline
+            Omnivory & 0.572 & -0.0133 & -0.139 & -4.89$\times10^{-4}$ & 0.908 \\
+            Apparent competition & 0.572 & 0.0105 & -0.139 & 4.78$\times10^{-4}$ & 0.905 \\
+            Direct competition & 0.572 & 2.62$\times10^{-3}$ & -0.139 & 3.84$\times10^{-4}$ & 0.901 \\
+            Three-species chain & 0.572 & 6.12$\times10^{-3}$ & -0.139 & -1.90$\times10^{-4}$ & 0.901 \\ 
+        \end{tabular}
+    \end{table}
+\clearpage
+
+\section{Results: Motif profiles and global properties}
+
+    A network's normalized motif profile (i.e., proportions of the total made up by each motif) was related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
+    However, normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).These differences in variability can cause false positives in the PERMANOVA test.
+Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).
+
+   \begin{figure}[h!]
+       \centering
+       \includegraphics[width=.75\textwidth]{figures/proportion_profile_dispersion.eps}
+       \caption{Dispersion of normalized motif profiles (i.e., proportions of each motif) within a network about the centroid for that combination of network size and connectance.
+Each circle represents a single network; circle colour indicates network size.
+Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
+       \label{dispersion_normmotifs}
+    \end{figure}
+
+    The frequencies of the four possible motifs were related to global network properties.
+
+    \begin{table}[hb!]
+        \centering
+        \caption{Coefficients for linear regressions relating the proportion of each motif in the total motif profile of a network to network size, connectance, and their interaction.
+Coefficients in \textbf{bold} are significant (\textless0.05).}
+       \label{network_prop_lms}
+       \begin{tabular}{c|c c c c c}
+            Motif & Intercept & Size & Connectance & Interaction \\
+            \hline
+            Three-species Chain & \textbf{0.262} & \textbf{1.78$\times10^{-4}$} & \textbf{-0.157} & \textbf{-3.27$\times10^{-3}$} \\
+            Apparent Competition & \textbf{0.541} & \textbf{-5.06$\times10^{-4}$} & \textbf{-1.00} & \textbf{2.59$\times10^{-3}$} \\
+            Direct Competition & \textbf{0.193} & -3.56$\times10^{-5}$ & \textbf{-0.556} & \textbf{5.92$\times10^{-3}$} \\
+            Omnivory & 4.82$\times10^{-3}$ & \textbf{3.64$\times10^{-4}$} & \textbf{1.71} & \textbf{-5.24$\times10^{-3}$}\\
+            \hline
+            \end{tabular}
+    \end{table}        
+\clearpage
 
@@
 
-\clearpage
-
-\section{SB: Motif profiles and global properties}
-
-    A network's normalized motif profile (i.e., proportions of the total made up by each motif) was related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
-    However, normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).These differences in variability can cause false positives in the PERMANOVA test.
-Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).
-
-   \begin{figure}[h!]
-       \centering
-       \includegraphics[width=.75\textwidth]{figures/proportion_profile_dispersion.eps}
-       \caption{Dispersion of normalized motif profiles (i.e., proportions of each motif) within a network about the centroid for that combination of network size and connectance.
-Each circle represents a single network; circle colour indicates network size.
-Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
-       \label{dispersion_normmotifs}
-    \end{figure}
-
-    The frequencies of the four possible motifs were related to global network properties.
-
-    \begin{table}[hb!]
-        \centering
-        \caption{Coefficients for linear regressions relating the proportion of each motif in the total motif profile of a network to network size, connectance, and their interaction.
-Coefficients in \textbf{bold} are significant (\textless0.05).}
-       \label{network_prop_lms}
-       \begin{tabular}{c|c c c c c}
-            Motif & Intercept & Size & Connectance & Interaction \\
-            \hline
-            Three-species Chain & \textbf{0.262} & \textbf{1.78$\times10^{-4}$} & \textbf{-0.157} & \textbf{-3.27$\times10^{-3}$} \\
-            Apparent Competition & \textbf{0.541} & \textbf{-5.06$\times10^{-4}$} & \textbf{-1.00} & \textbf{2.59$\times10^{-3}$} \\
-            Direct Competition & \textbf{0.193} & -3.56$\times10^{-5}$ & \textbf{-0.556} & \textbf{5.92$\times10^{-3}$} \\
-            Omnivory & 4.82$\times10^{-3}$ & \textbf{3.64$\times10^{-4}$} & \textbf{1.71} & \textbf{-5.24$\times10^{-3}$}\\
-            \hline
-            \end{tabular}
-    \end{table}        
-\clearpage
-
-\section{SC: Persistence and motif profiles}
-
-    Persistence varied with the frequencies of different motifs in a network profile.
-
-    \begin{table}[h!]
-        \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
-Predictors were centered and scaled before model fitting.
-All effects were significant ($p$\textless0.001).
-We also provide the $R^2$.
-}
-        \label{motif_profile_tab}
-        \centering
-        \begin{tabular}{c|c c c c c | c}
-            Motif & Intercept & Proportion & Disturbance & Interaction & $R^2$\\
-            \hline
-            Omnivory & 0.572 & -0.0133 & -0.139 & -4.89$\times10^{-4}$ & 0.908 \\
-            Apparent competition & 0.572 & 0.0105 & -0.139 & 4.78$\times10^{-4}$ & 0.905 \\
-            Direct competition & 0.572 & 2.62$\times10^{-3}$ & -0.139 & 3.84$\times10^{-4}$ & 0.901 \\
-            Three-species chain & 0.572 & 6.12$\times10^{-3}$ & -0.139 & -1.90$\times10^{-4}$ & 0.901 \\ 
-        \end{tabular}
-    \end{table}
-
-\clearpage
-
-\section{SD: Persistence and motif participation}
-
-    Persistence varied with motif participation.
-
-    \begin{table}[h!]
-        \centering
-        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{propreq}).
-Predictors were centered and scaled before model fitting.
-We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
-        \label{tab:proportion}                \footnotesize
-        \begin{tabular}{c|c c | c c | c c | c c | c c |}
-        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c|}{$R^2$} \\
-        Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
-        \hline
-        3-sp Chain & 0.574 & \textless0.001 & -0.00508 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
-        App. Com. & 0.574 & \textless0.001 & -0.00312 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
-        Dir. Com. & 0.574 & \textless0.001 & 0.00564 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
-        Omnivory & 0.574 & \textless0.001 & 0.00709 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
-        \end{tabular}
-    \end{table}        
-\clearpage
-
-\section{SE: Motif-persistence relationships \& network size}
-
-    When comparing the outcomes from linear mixed-effect models (LMMs) relating persistence to motif roles across all networks (Fig.
-3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately, the main trends are similar but different-sized networks showed slightly different results results for some motifs (Figure temporarily in main text).%(Figure~\ref{fig:density_prop_S}).
-    For the three-species chain and omnivory motifs, a significant interaction between disturbance and the effect of the focal motif means that having a larger proportion of the role be made up of the chain or omnivory motif is associated with lower persistence at high levels of disturbance (Fig.
-3 and Table 1, \emph{Main Text}).
-    The same pattern holds overall for LRs fit within a network  %(Figure~\ref{fig:density_prop_S})
-    : the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance. 
-    This trend is most pronounced in large networks.% (bottom panel, Figure~\ref{fig:density_prop_S}). 
-    Besides the decrease in fraction of positive slopes with increasing disturbance, the omnivory motif show a higher overall fraction of positive slopes in comparison to the three-species chain motif, corresponding to the positive main effect of omnivory (Fig.
-3 and Table 1, \emph{Main Text}).
-
-    \begin{figure}[h!]
-         \centering
-         % \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
-        \caption{Density (y-axis) of slopes (x-axis) for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
-The slopes are derived from linear mixed-effect models, with a random intercept for the interaction between species richness and network size.
-A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
-The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to motif (legend).
-Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
-Rows show various levels of network size.
-The vertical line indicate zero on the x-axis.}
-        \label{fig:density_prop_S}
-    \end{figure}
-
-    For the apparent and direct competition motifs, the LMMs include a positive interaction between disturbance and the proportion of the motif in a species' role (Fig.
-3 and Table 1; \emph{Main Text}).
-    This means that having more of the role made up of these two motifs is increasingly beneficial at higher levels of disturbance, a trend which is also evident from the increasing fraction of positive slopes with increasing disturbance based on LRs of each network at each level of disturbance.
-% (Figure~\ref{fig:density_prop_S}).
-    As with the trends for omnivory and three-species chains, these increases are strongest in large networks. % (bottom row, Figure~\ref{fig:density_prop_S}).
-    % For direct competition, the fraction of slopes are similar across disturbance levels.
-Although this interaction is not positive when accounting for network size, the fraction of positive slopes are throughout high.
-    The increase in the proportion of positive slopes was much stronger for apparent competition than direct competition, consistent with the smaller (but still significant) interaction term for between direct competition in the LMMs.
-    A high proportion of direct competition in a species' role tended to be associated with greater probability of persistence across levels of disturbance and network size.
-    % In conclusion, species richness does not considerably alter the trends visible in the LMMs.
-High species richness more clearly display an interaction between motif proportion, persistences and disturbance levels, while this effect is smaller with low species richness.
-\clearpage
-
-\section{SF: Persistence, degree, and trophic level}
-
-    Persistence varies with in-degree, trophic level, and their interaction.
-
-    \begin{table}[h!]
-        \caption{Coeffcients ($\beta$) for regressions of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction.
-Predictors were centered and scaled before model fitting.
-All $p$-values were \textless0.001.
-The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.}
-        \label{tab:per_vs_TLdeg}
-        \centering
-        \begin{tabular}{l|c  c  c |}
-            Simple role measure & Role & Disturbance & Interaction \\
-            \hline
-            STL & -0.0218 & -0.139 & 0.00508 \\
-            in-degree & -0.00130 & -0.139 & -0.00908 \\
-        \end{tabular}
-    \end{table}
-
-    \begin{figure}[h!]
-     \centering
-     % \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}
-     \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
-The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
-Persistence decreases with darker color in the heat map.}
-     \label{fig:heatmap_stl_BP0}
-    \end{figure}
-
-    \begin{figure}[h!]
-     \centering
-     % \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
-     \caption{Primary producers experience a disturbance reducing their probability to go extinct to $\pi_{disturbed} = 0.5$ All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
-The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
-Persistence decreases with darker color in the heat map.}
-     \label{fig:heatmap_stl_BP1}
-    \end{figure}
-
-\clearpage
-
-\section{SG: Motif participation and network properties}
-
-    Motif participation varies with global-scale network properies.
-
-    \begin{table}[hb!]
-        \centering
-        \caption{Motif participation varies with size and connectance. Here we give the range and mean ($\pm$ standard deviation) proportion of each motif in species' motif participation vectors. 
-        We also show the effects ($\beta$) of network size, connectance, and their interaction in linear models predicting the proportion of each motif in a participation vector.
-        All effects were significant ($p$\textless0.001).
-        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
-        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
-        \label{tab:partic_vs_SC}   
-        \footnotesize
-        \begin{tabular}{c|c c c | c c c}
-            Motif & Min & Max & Mean (SD) & Size & Connectance & Interaction \\
-            \hline
-            3-sp Chain & 0 & 1 & 0.249 (0.127) & 2.22$\times10^{-4}$ & -0.131 & -3.20$\times10^{-3}$ \\
-            App. Comp & 0.250 & 0.765 & 0.422 (0.0645) & -5.06$\times10^{-4}$ & -1.00 & 2.59$\times10^{-3}$ \\
-            Dir. Comp & 0.0501 & 0.359 & 0.179 (0.0407) & -3.56$\times10^{-5}$ & -0.556 & 5.92$\times10^{-3}$ \\
-            Omnivory & 0.006 & 0.346 & 0.164 (0.0790) & 3.64$\times10^{-4}$ & 1.71 & -5.24$\times10^{-3}$ \\   
-            \hline
-            \end{tabular}
-            \end{table}
-
-    \begin{figure}[h!]
-        \centering
-        \includegraphics[height=.65\textheight]{figures/participation_vs_SC.eps}
-        \caption{The proportion of each motif in a consumer's participation vector varied with network size, connectance, and their interaction.
-These relationships were steepest for apparent competition and omnivory.}
-        \label{fig:roles_vs_SC}
-    \end{figure}
 \clearpage
@@
 
-    % \section{How do counts of motifs in a network profile vary with S and C?}
-
-    %     As expected, all four motifs had higher counts in larger and more highly connected networks.
-    %     The precise effects of size, connectance, and their interaction differed between motifs  (Table~\ref{network_motif_lms}).
-    %     This suggests that the meso-scale structure of a network varies with size and connectance, not just the total number of motifs.
-    %     Indeed, the motif profile of a network is significantly related to its size, connectance, and their interaction ($F_{1,2996}$=2279, $p$\textless0.001; $F_{1,2996}$=3067, $p$\textless0.001; and $F_{1,2996}$=1187, $p$\textless0.001, respectively).
-
-    %     \begin{table}[h!]
-    %         \centering
-    %         \caption{Coefficients for linear regressions relating the total count of all motifs or the count of each three-species motif in a network to network size, connectance, and their interaction.
-Coefficients in \textbf{bold} are significant (\textless0.05).
-Note that, while there are 13 unique three-species motifs, only four are possible in acyclic networks such as those considered here.}
-    %        \label{network_motif_lms}
-    %        \begin{tabular}{c|c c c c c}
-    %         Motif & Intercept & Size & Connectance & Interaction \\
-    %         \hline
-    %         Total count & \textbf{-2.85$\times10^3$} & \textbf{46.6} & \textbf{-2.12$\times10^{5}$} & \textbf{4.72$\times10^{3}$} \\
-    %         \hline
-    %         Thee-species Chain & \textbf{-1.40$\times10^{3}$} & \textbf{26.1} & \textbf{-3.74$\times10^{4}$} & \textbf{8.58$\times10^{2}$} \\
-    %         Apparent Competition & \textbf{-2.23$\times10^3$} & \textbf{42.0} & 
-    %         \textbf{-7.26$\times10^4$} & \textbf{1.63$\times10^3$} \\
-    %         Direct Competition & 26.6 & 0.777 & \textbf{-4.85$\times10^4$} & \textbf{9.89$\times10^2$} \\
-    %         Omnivory & \textbf{7.57} & \textbf{-22.2} & \textbf{-5.37$\times10^4$} & \textbf{1.24$\times10^3$} \\
-    %         \hline
-    %         \end{tabular}
-    %     \end{table}
-
-    %     \begin{figure}[h!]
-    %         \centering
-    %         \includegraphics[width=.75\textwidth]{figures/motif_profile_dispersion.pdf}
-    %         \caption{Dispersion of motif profiles within a network about the centroid for that combination of network size and connectance.
-Each circle represents a single network; circle colour indicates network size.
-Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
-    %         \label{dispersion_rawmotifs}
-    %     \end{figure}
-
-    %     Motif profiles were similarly variable across networks with different sizes, but were not homogeneously variable across connectances or size-connectance combinations ($F_{5,2994}$=0.365, $p$=0.873; $F_{4,2995}$=4.49, $p$=0.001; and $F_{29,2970}$=27.1, $p$\textless0.001 , respectively).
-    %     These differences in variability of motif profiles can cause false positives in the PERMANOVA.
-    %     Variability in motif profiles was lowest among large networks with moderate connectance and was highest among the least-connected and most-connected networks we considered (Fig.~\ref{dispersion_rawmotifs}).
-
-    %     \begin{figure}[h!]
-    %         \centering
-    %         \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
-    %         \caption{The effect of number of times a species appears (x-axis) in any of the various motifs (columns) on persistence (y-axis).
-The effect of each motif participation is fitted by linear mixed-effect models, following Equation~\ref{absreq}, except that the disturbance levels are fitted separately.
-The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
-    %         \label{fig:abs_lmer_all}
-    %     \end{figure}
-
     \subsection*{Motif profiles and proportion of basal resources}
@@
         \end{figure}
-
-\clearpage
 
@@
             \end{table}
-
 \clearpage
 
-\section{Motif profiles in different network sets}
+\section*{SK: Motif profiles in different network sets}
 
@@
     \end{table}
-
 \clearpage
-
-% \section{Persistence and count-based motif participation}
-
-%     \subsection*{Methods}
-
-%         We test whether participating in higher absolute numbers (i.e., counts) of a motif increases persistence using a set of models that parallels those used to test whether having a greater proportion of the role made up by a particular motif (see \emph{Main Text}).
-%         Specifically, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks).
-%         In each case, we considered only non-basal species (i.e., those with at least one prey) as the persistence of basal species is determined only by their baseline probability.
-%         These models included the absolute number of times the focal species appeared in the focal motif, the increase in probability of extinction for basal species, and the interaction between the two as well as a random intercept for the interaction between species richness and connectance (equations.
-%         All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-%         We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.
-
-%             The formula for these LMMs are:         
-%             \begin{equation}
-%                 \Psi_{ik} \approx \chi_{i} + \pi_{disturbed_k} + \chi_{i}\pi_{disturbed_k} + S:C_{i} ,
-%                 \label{absreq}
-%             \end{equation}
-
-%             \noindent where $\Psi_ik$ is the persistence of species $i$ during disturbance level $k$, $\chi_{i}$ is the number of times species $i$ appears in the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S:C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
-
-%             % Additionally, we performed a similar linear regression of persistence against motif participation within each simulated network and for each level of disturbance to basal resources (XX regressions total).
-%             % These regressions were fit using the R~\citep{R} base function `lm'.
-%             % We then analyse the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.
-
-%     \subsection*{Results}
-
-%         \begin{figure}[h!]
-%             \centering
-%             \includegraphics[width=\textwidth]{figures/absolute_lmer_allCS.pdf}
-%             \caption{The effect of number of times a species appear (x-axis) in any of the various motifs (columns) on persistence (y-axis).
-The effect of each motif participation is fitted by linear mixed-effect models, following \emph{Equation 2; Main Text}, except that the disturbance levels are fitted separately.
-The different colored lines indicate various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (in the top) to $\pi_{disturbed} = 0.5$ (in the bottom).}
-%             \label{fig:abs_lmer_all}
-%         \end{figure}
-
-%         Persistence increased significantly with the number of times a species appeared in any of the four possible motifs in the food webs (Table~\ref{tab:absolute_number} and Figure~\ref{fig:abs_lmer_all}).
-%         These relationships were slightly stronger for the number of times a species appears in the three-species chain and direct competition motifs and somewhat weaker for apparent competition and omnivory.
-The absolute difference between the motifs was small, with a clear positive relationship between persistence and motif count present in all cases.
-%         As expected, persistence decreased significantly with increasing disturbance of basal resources and the strength of this relationship was very similar across the four motifs.
-%         The level of disturbance to basal resources also affected the relationship between persistence and motif counts.
-%         For all four motifs, there was a negative interaction between motif count and disturbance such that, for high levels of disturbance, participating more often in any of the motifs would result in decreased persistence.
-
-%         \begin{table}[h!]
-%             \caption{Coefficients ($\beta$) for models relating persistence to the number of times a species appears in the focal motif, the level of disturbance to basal resources, and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{absreq}).
-$p$-values for all coefficients in all models were \textless0.001.
-We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
-%             \label{tab:absolute_number}        \centering
-%             \begin{tabular}{c|c  c c  c | c c |}
-%             Motif & Intercept & Motif count & Disturbance & Interaction & $R^2_M$ & $R^2_C$ \\
-%             \hline
-%             3-sp Chain & 0.762 & 1.14$\times10^{-4}$ & -0.388 & -1.36$\times10^{-4}$ & 0.897 & 0.913 \\
-%             Apparent Compet. & 0.770 & 3.41$\times10^{-5}$ & -0.397 & -4.36$\times10^{-5}$ & 0.898 & 0.911\\
-%             Direct Competition & 0.764 & 1.22$\times10^{-4}$ & -0.398 & -8.46$\times10^{-5}$ & 0.895 & 0.914 \\
-%             Omnivory & 0.770 & 6.01$\times10^{-5}$ & -0.400 & -5.91$\times10^{-5}$ & 0.897 & 0.912 \\                
-%             \end{tabular}
-%             \end{table}
-
-% \clearpage
 

### 2021-10-01 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 \newcommand{\SI}{\textit{Appendix}~}
+
+\setlength{\abovedisplayskip}{0pt}
+\setlength{\belowdisplayskip}{0pt}
 
@@
     % basic idea behind BN modeling
-        Bayesian Network modelling represents species and feeding interactions as the nodes and links of a network (see Appendix \emph{S2} for full details).
-        Each species (node) is a Bernoulli random variable, being either extant or extinct. 
+        Bayesian Network modelling can represent species and feeding interactions as the nodes and links of a network (see Appendix \emph{S2} for full details).
+        In our setting, each species (node) is a Bernoulli random variable, being either extant or extinct. 
         The probability that a particular species is extinct is a function of the states of its resources.
@@
 		% probability of extinction equation
-		Consider a species with $n$ resource species. 
-		Before taking food-web structure into account, this species already has a baseline extinction  probability of extinction, $\pi_{base}$. 
+		%Consider a species with $n$ resource species. 
+		Before taking food-web structure into account, all species already have a baseline probability of extinction, $\pi_{base}$. 
 		$\pi_{base}$ represents the risk of going extinct due to diseases, stochastic extinction of small populations, or other factors not related to the food web.
 		$\pi_{base}$ may be constant, or it can vary to reflect greater sensitivity in some groups of species.
-		In our simulations, we vary $\pi_{base}$ for basal resources between 0.1 and 0.5 but keep constant $\pi_{base}=0.1$ for consumers.
+		%In our simulations, we vary $\pi_{base}$ for basal resources between 0.1 and 0.5 but keep constant $\pi_{base}=0.1$ for consumers.
 
-        If the focal species is a consumer with $n$ resources, its probability of extinction depends on $\pi_{base}$ and on the fraction $f$ of its resources that have gone extinct.
+        If the focal species is a consumer with $n$ resources, its probability of extinction depends on $\pi_{base}$ and on the fraction $f = k/n$ of its resources that have gone extinct ($k$).
         Note that we assume that only the fraction of resources lost is important and not the identities of the missing resources.
@@
 This will in turn affect the extinction probability $P(\lnot C|f)$.
-		For basal species, we assume that abiotic resources are always available and $f=0$, such that extinction depends only on $\pi_{base}$.
+		For primary producers, we assume that abiotic resources are always available and $f=0$. Following Equation \ref{betafunc}, the extinction will only depend on $\pi_{base}$.
 
 		% how we calculate persistences 
-        Beginning with primary producers and working systematically upward through the food web, we use Equation \ref{betafunc} to calculate each species' probability of extinction $P(\lnot i|f)$.
+        Beginning with primary producers and working systematically upward through the food web, we use Equation \ref{betafunc} to calculate each species' $i$ probability of extinction $P(\lnot i|f)$.
         Using this probability, we then perform Bernoulli trials to determine whether the species goes extinct or not. 
@@
 
-	We tested 1) how species persistence varied with motif participation (Appendix \emph{3.1}), 2) how motif participation varied with other network properties (Appendix \emph{3.2}), 3) how network mean persistence varied with network motif profiles (Appendix \emph{3.3}), 4) how network motif profiles varied with global structural properties (Appendix \emph{3.4}), and 5) how species persistence and network mean persistence varied with other network properties (Appendix \emph{3.5}).
-The analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan} as detailed in the appendices.
+	We tested:
+	\noindent1) how species persistence varied with motif participation (Appendix \emph{3.1}),
+
+	\noindent2) how motif participation varied with other network properties (Appendix \emph{3.2}),
+
+	\noindent3) how network mean persistence varied with network motif profiles (Appendix \emph{3.3}),
+
+	\noindent4) how network motif profiles varied with global structural properties (Appendix \emph{3.4}), and
+
+	\noindent5) how species persistence and network mean persistence varied with other network properties (Appendix \emph{3.5}). 
+	All analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan} as detailed in the appendices.
 
@@
         To test if participating more often in a certain motif affected a species' probability of persistence, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks):
-
             \begin{equation}
@@
             \end{equation}
-
         \noindent where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
@@
 
-        We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} (Appendix \emph{3.2}). 
-        To identify correlations between size, connectance, and participation in particular motifs, we fit a set of four linear models (LMs):
-
+        Different scales of network structure (e.g., local and meso-scale) are not independent. 
+        To identify correlations between network size, connectance, and participation in particular motifs, we fit a set of four linear models (LMs):
         \begin{equation}
@@
         \end{equation}
-
-        where $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif,
+        \noindent where $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif,
         $\Sigma_{i}$ is the size of the network containing species $i$, and $\zeta_{i}$ is the connectance of the network containing species $i$.
-
-        We fit a similar set of eight LMMs relating motif participation to in-degree or trophic level:
-
+        We fit a similar set of eight LMMs (two per motif) relating motif participation to in-degree or trophic level:
         \begin{equation}
-            \rho_{i} \approx \delta_{i} + S_{i}C_{i} ,
-            \label{partic_deg}
+            \rho_{i} \approx \delta_{i} + S_{i}C_{i} , \hspace{2cm} 
+            \rho_{i} \approx \tau_{i} + S_{i}C_{i} ,
+            \label{partic_degTL}
         \end{equation}
-
-        \begin{equation}
-            \rho_{i} \approx \tau_{i} + S_{i}C_{i} ,
-            \label{partic_deg}
-        \end{equation}
-
-        where $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, and all other symbols are as in equation~\ref{propreq}.
+        \noindent where $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, and all other symbols are as in equation~\ref{propreq}.
 
@@
 
-        As well as species-specific probabilities of persistence, motifs may provide information about the overall probability of any species going extinct (since they incorporate information on indirect effects).
+        As well as species-specific probabilities of persistence, the overall frequency of motifs in a network may provide information about the overall probability of any species going extinct (since they incorporate information on indirect effects).
         We tested whether networks with similar motif profiles tend to have similar mean persistences among consumers using a PERMANOVA (Appendix \emph{S3.3}.
-        To identify the relationships between particular motifs and network mean persistence, we fit a set of four LMs (one per motif):
-
+        To identify the relationships between particular motifs and network mean persistence in more detail, we fit a set of four LMs (one per motif):
             \begin{equation}
@@
             \end{equation}
-
-        where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and all other symbols are as in equation~\ref{propreq}.
+        \noindent where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and all other symbols are as in equation~\ref{propreq}.
 
@@
 
-        Different scales of network structure (e.g., local and meso-scale) are not independent. 
         To test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
         In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMs:
-
         \begin{equation}
@@
         To identify the relationship between global network structure and mean probability of persistence within a network, we fit a linear regression:
-
             \begin{equation}
@@
             \end{equation}
-
-        where all symbols are as in equation~\ref{propreq} and~\ref{partic_SC}.
+        \noindent where all symbols are as in equation~\ref{propreq} and~\ref{partic_SC}.
         As the three-way interaction was significant, we did not simplify the model.
@@
         To identify the relationships betwen local-scale properties and species' probabilities of persistence, we fit a final two LMMs:
-
         \begin{equation}
@@
             \end{equation}
-
         \begin{equation}
@@
             \end{equation}
-
-        where $\delta_{i}$ and $\tau_i$ are the in-degree and trophic level of species $i$, respectively, and all other terms are as in equation~\ref{propreq}.
+        \noindent where $\delta_{i}$ and $\tau_i$ are the in-degree and trophic level of species $i$, respectively, and all other terms are as in equation~\ref{propreq}.
 
@@
        Conversely, persistence decreased with an increasing proportion of the apparent competition motif. 
-       Persistence was not significantly related to the proportion of the three-species chain motif (Figure~\ref{fig:prop_lmer_all}).
+       Persistence was not significantly related to the proportion of the three-species chain motif (Figure~\ref{fig:prop_lmer_all}; upper lines in each panel).
 
-        At high levels of disturbance, a species' persistence decreased as the proportion of the three-species chain increased, but persistence increased as the proportion of any other motif increased (Figure~\ref{fig:prop_lmer_all}).
-        This means that increased disturbance strengthened the effects of the three-species chain motif and changed the direction of the effect of the apparent competition motif  (Table S4, \emph{Appendix SD}).
+        At high levels of disturbance, a species' persistence decreased as the proportion of the three-species chain increased, but persistence increased as the proportion of any other motif increased (Figure~\ref{fig:prop_lmer_all}; lower lines in each panel).
+        This means that increased disturbance strengthened the effects of the three-species chain motif and changed the direction of the effect of the apparent competition motif  (Table S1, \emph{Appendix S4}).
         The relationship between omnivory and persistence was still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.
-
-            \begin{figure}[h!]
-                \centering
-                \includegraphics[width=0.85\textwidth]{figures/prop_lmer_allCS.pdf}
-                \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).
-The effect of participating in each motif is based on linear mixed-effect models following Equation~\ref{propreq}, except that models for each  disturbance level are fitted separately.
-The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance).
-Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
-                \label{fig:prop_lmer_all}
-            \end{figure}
 
@@
 
-       Species' motif participation varied depending on the properties of the network; size, connectance, and their interaction (Fig. S4, Table S6; \emph{Appendix SG}).
+       Species' motif participation varied depending on the global properties of the network: size, connectance, and their interaction (Fig. S1, Table S2; \emph{Appendix S5}).
        Species appeared most often in the apparent competition motif in all types of networks, but the frequency decreased with increasing network size and connectance.
@@
 
-       A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix SF}).
-Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S5, \emph{Appendix SF}).
+       A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix S6}).
+Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S3, \emph{Appendix S6}).
        In-degree shows an even stronger interaction with disturbance; having more prey species is associated with higher persistence without disturbance, but lower persistence at high levels of disturbance.
-
-            \begin{figure}
-                \centering
-                \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
-                \caption{As with motifs, persistence varied with simpler measures of species roles in networks.
-Moreover, motif frequencies correlated with in-degree and trophic level.
-\textbf{A-B)} Persistence increased with increasing in-degree for low levels of disturbance to basal resources but decreased with increasing degree for high levels of disturbance to basal resources.
-                Persistence declined with inceasing STL for all levels of disturbance to basal resources.
-                \textbf{C-D)} The proportion of omnivory increased with increasing degree while all other proportions decreased.
-The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.}
-                \label{fig:motifs_vs_TL_and_deg}
-            \end{figure}
 
@@
 
-        % In order to evaluate how consistent the relationships between a species persistence and motif participation is, and how it varies with  level of disturbance and global network structure, we fit linear regressions of species persistence against motif proportions for the species for each network separately.
-We then analyze the density distribution of the values of the slopes.
-Network size had a much smaller effect than connectance (Fig.
-S3; \emph{Appendix SE}); we therefore focus on connectance here.
-
-        When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally had more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
+        When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally showed more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
         However, this trend varied at different levels of disturbance, especially for the direct competition and omnivory motifs.
@@
         Moreover, the omnivory motif has a particularly broad distribution of slopes (i.e., an inconsistent relationship to persistence) when network connectance was low. 
-        Network size had a much smaller effect than connectance (Fig. S3; \emph{Appendix SE}).
-
-        \begin{figure}[h!]
-            \centering
-            \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.pdf}
-            \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
-Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
-Rows show various levels of connectance.
-The dotted, vertical line indicate zero on the x-axis.
-A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
-The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend).
-}
-            \label{fig:density_prop_C}
-        \end{figure}
+        Network size had a much smaller effect than connectance (Fig. S4; \emph{Appendix S7}).
 
@@
         The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001). 
-        Average persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S3, \emph{Appendix SC}). 
+        Average persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S4, \emph{Appendix S8}). 
         Conversely, average persistence was lower in networks with higher proportions of the omnivory motif. 
-        These trends were consistent across levels of disturbance (Fig.~\ref{fig:motif_profile_persistence}, Table S3), \emph{Appendix SC}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
+        These trends were consistent across levels of disturbance (Fig.~\ref{fig:motif_profile_persistence}, Table S4), \emph{Appendix S8}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
 For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).
-
-        \begin{figure}
-            \centering
-            \includegraphics[width=.5\textwidth]{figures/persistence_motif_profiles.eps}
-            \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.
-Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportions of the other three motifs.
-Interactions between motif profiles and disturbance were significant but small.
-Here we show the relationships of probabilities of extinction of basal resources of 0.1 (top set of lines) and 0.5 (bottom set of lines).
-Relationships are shown for the observed range of proportions for each motif.}
-            \label{fig:motif_profile_persistence}
-        \end{figure}
 
@@
 
-        Although the PERMANOVA test relating overall motif profiles to network properties was not reliable (Fig. S3, \emph{Appendix SB}), 
-        connectance was significantly related to the frequencies of different motifs in a network when considering motifs individually.
-        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs (Fig.~\ref{motif_proportion_lms}).
-        Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S2, \emph{Appendix SB}).
+        Although the PERMANOVA test relating overall motif profiles to network properties was not reliable (Fig. S5, \emph{Appendix S9}), 
+        connectance was significantly related to the frequencies of different motifs in a network when considering motifs individually (Fig.~\ref{motif_proportion_lms}).
+        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs .
+        Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S5, \emph{Appendix S9}).
 
-        \begin{figure}[h!]
-            \centering
-            \includegraphics[width=.75\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
-            \caption{The proportions of the four motifs in a network's motif profile were related to the mean persistence of species in the network.
-Specifically, persistence decreased as the proportion of omnivory in the network's motif profile increased while persistence increased with increasing proportions of the other three motifs.
-Interactions between motif profiles and disturbance were significant but small.
-Here we show the relationships of probabilities of extinction of basal resources of 0.1 (top set of lines) and 0.5 (bottom set of lines).
-Relationships are shown for the observed range of proportions for each motif.}
-            \label{motif_proportion_lms}
-        \end{figure}
-
-     \clearpage
-
-\section*{Results to SI entirely?}
-
-    \subsection*{Network mean persistence and global network properties}
-
-        %Increasing the probability of extinction for basal species necessarily decreased the mean probability of persistence for consumers. 
-        %This effect was clear across the ranges of network size and connectance we considered. 
-        %Network size and connectance also affected network mean persistence, but to a much smaller extent than increasing disturbance (Figure S1, Table S1, \emph{Appendix SA}). 
-        %Persistences generally decreased in more-connected networks, but the effect of network size varied depending on level of disturbance. 
-        %At low disturbance, small and highly-connected webs showed the lowest mean persistence, whereas at high disturbance, large and highly connected webs showed the lowest mean persistence (Fig.
-S2, \emph{Appendix SA}).
-
-    \subsection*{Species characteristics and persistence}
-
-        % Anna notes (just for our (mine) understanding): 
-        % When looking at the two figures of degree vs persistence and STL vs persistence - my conclusion is that species at lower trophic levels have the highest in-degree.
-Double check - yes, that’s true.
-        % Lower TL are all well at low disturbances, and when disturbance increase, persistence decrease (looking at fig degree vs persistence, right end of the line).
-        % Higher TL experience the same thing, although to a slightly smaller extent (looking at the left end of the lines in fig degree vs persistence and right end of lines in fig STL vs persistence).
-        % The slope of the line in fig degree vs persistence is not really relevant, since it's kind of based on the response of different TL level (which makes perfect sense in the BN framework)
-        % weird thing though - at high levels of disturbance, lower TL seems more affected than higher TL in the persistence vs degree figure (since low degree belongs to higher TL, high degree to lower TL).
-This does not correspond with fig persistence vs STL.
-HOWEVER - not only high TL have a low degree, lower TL do as well, probably affecting the overall result.
-Only low TL have high degree - but the opposite is not true.
-Probably explains it
-
-    \clearpage
-
-    % \subsection*{Persistence and species' positioning in motifs}
-
-    %     A species probability of persisting is significantly related to the frequency of all positions, disturbance, and the interaction between frequency of each position and disturbance (Table SX, \emph{Appendix SH}).
-    %     In the omnivory motif, persistence increased with the frequency of the bottom position (two predators, no prey within the motif) across all levels of disturbance but most steeply at high levels of disturbance (Fig.~\ref{figs:persistence_position_motifs}).
-    %     Persistence increased slightly with the frequency of the middle position (one prey, one predator) at low levels of disturbance but showed little relationship at higher levels of disturbance.
-    %     Persistence increased with the frequency of the top position (two prey, no predators within the motif) at low disturbance but decreased at higher levels of disturbance.
-
-    %     In the three-species chain, persistence decreased as the frequency of the top position increased at all levels of disturbance.
-    %     Conversely, persistence increased with increasing frequency of the middle position at all levels of disturbance.
-    %     Persistence decreased slightly with increasing frequency of the bottom position at low levels of disturbance but increased at high levels of disturbance.
-
-    %     In the apparent competition motif, persistence decreased slightly with increasing frequencies of the bottom position (one predator, no prey within the motif) at low disturbance but increased slightly at high disturbance.
-    %     Persistence increased with increasing frequency of the top position (two prey, no predator within the motif) at low disturbance but decreased at high disturbance.
-
-    %     In the direct competition motif, persistence varied little with the frequency of the bottom position (two predators, no prey within the motif) at low disturbance but increased with the frequency of this position at high levels of disturbance.
-    %     Persistence increased slightly with the frequency of the top position (one prey, no predators within the motif) at low levels of disturbance and decreased slightly at high levels of disturbance.
-
-    % \begin{figure}[h!]
-    %     \includegraphics[width=.95\textwidth]{figures/persistence_positions_Omnivory.eps}
-    %     \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
-Here we show the relationships between persistence and the proportions of the three positions in the omnivory motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
-Line style indicates the motif  position.
-}
-    %     \label{figs:persistence_position_motifs}
-    %     \end{figure}
-
-    %     The frequencies of motif positions in a species' role varied with network properties (\emph{Appendix SI}).Motif roles had stronger relationships with degree and STL than with network size or connectance.
-
-\clearpage
-
-%Somewhere early in discussion (needs tidying up):
-%Apparent comp and direct comp have no indirect effects in BN, so we should be able to explain them with simpler properties.
-
+% Appendices SA (persistence vs.
+S,C), SH-I (analyses with motif positions), SJ-K (repeated analyses with number of basal resources instead of disturbance) not referred to in text.
+Keeping SH-K in commented-out form in case reviewers ask.
+We should refer to SA somewhere in Discussion (will be Appendix 10).
 \section*{Discussion}
@@
 
-\section*{Glossary}
+\section*{Figure captions}
 
-\begin{table}[h!]
-\label{glossary}
-\caption{Glossary of terms relating to motifs and Bayesian networks}
+\section*{Figures}
 
-\begin{tabular}{l|l}
-    Term & Definition \\
-    \hline
-    Motif &  \\
-    Network profile & \\
-    Motif participation & \\
-    Motif role & \\
-    Bayesian network & A directed acyclic graph, used to predict species' likelihood of persistence. \\
-    Network persistence & The mean likelihood of consumers in a network not going extinct.\\
-    Species persistence & The individual likelihood of a species not going extinct.\\
-    In-degree & Number of prey species to a consumer species.\\
-    Trophic level (STL) & The shortest food chain between the focal species and any basal species.\\
-    Disturbance ($\pi_{disturbed}$) & Probability of extinction of a basal resource when extra disturbance is added. \\
-    Baseline extinction &  When no disturbance is applied, $\pi_{base}$=0.1.\\
-\end{tabular}
-\end{table}
+    \begin{figure}[h!]
+        \centering
+        \includegraphics[width=0.85\textwidth]{figures/prop_lmer_allCS.pdf}
+        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).
+The effect of participating in each motif is based on linear mixed-effect models following Equation~\ref{propreq}, except that models for each  disturbance level are fitted separately.
+The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance).
+Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
+    \label{fig:prop_lmer_all}
+    \end{figure}
+
+    \begin{figure}
+        \centering
+        \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
+        \caption{As with motifs, persistence varied with simpler measures of species roles in networks.
+Moreover, motif frequencies correlated with in-degree and trophic level.
+\textbf{A-B)} Persistence increased with increasing in-degree for low levels of disturbance to basal resources but decreased with increasing degree for high levels of disturbance to basal resources.
+        Persistence declined with inceasing STL for all levels of disturbance to basal resources.
+        \textbf{C-D)} The proportion of omnivory increased with increasing degree while all other proportions decreased.
+The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.}
+        \label{fig:motifs_vs_TL_and_deg}
+    \end{figure}
+
+    \begin{figure}[h!]
+    \centering
+        \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.pdf}
+        \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
+Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
+Rows show various levels of connectance.
+The dotted, vertical line indicate zero on the x-axis.
+A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
+The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend).
+}
+        \label{fig:density_prop_C}
+    \end{figure}
+
+    \begin{figure}
+        \centering
+        \includegraphics[width=.5\textwidth]{figures/persistence_motif_profiles.eps}
+        \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.
+Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportions of the other three motifs.
+Interactions between motif profiles and disturbance were significant but small.
+Here we show the relationships of probabilities of extinction of basal resources of 0.1 (top set of lines) and 0.5 (bottom set of lines).
+Relationships are shown for the observed range of proportions for each motif.}
+        \label{fig:motif_profile_persistence}
+    \end{figure}    
+    \begin{figure}[h!]
+        \centering
+        \includegraphics[width=.75\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
+        \caption{The proportions of the four motifs in a network's motif profile were related to the mean persistence of species in the network.
+Specifically, persistence decreased as the proportion of omnivory in the network's motif profile increased while persistence increased with increasing proportions of the other three motifs.
+Interactions between motif profiles and disturbance were significant but small.
+Here we show the relationships of probabilities of extinction of basal resources of 0.1 (top set of lines) and 0.5 (bottom set of lines).
+Relationships are shown for the observed range of proportions for each motif.}
+        \label{motif_proportion_lms}
+    \end{figure}
+
+\clearpage
+
+% \section*{Glossary}
+% \begin{table}[h!]
+% \label{glossary}
+% \caption{Glossary of terms relating to motifs and Bayesian networks}
+%     \footnotesize{
+% \begin{tabular}{l|l}
+%     Term & Definition \\
+%     \hline
+%     Motif & Set of $n$ interacting species. In this case, $n=3$ \\
+%     Network motif profile & Vector describing the frequency of each motif in the network. \\
+%     & Normalised by dividing counts of each motif by the total across all motifs.\\
+%     Motif participation role & Vector describing the frequency with which a focal species appears in each motif.\\
+%     & Normalised by dividing counts for each motif by the total across all motifs. \\
+%     Bayesian network & A directed acyclic graph, used to predict species' likelihood of persistence. \\
+%     Network persistence & The mean likelihood of consumers in a network not going extinct.\\
+%     Species persistence & The individual likelihood of a species not going extinct.\\
+%     In-degree & Number of prey species to a consumer species.\\
+%     Trophic level (STL) & The shortest food chain between the focal species and any basal species.\\
+%     Disturbance ($\pi_{disturbed}$) & Probability of extinction of a basal resource when extra disturbance is added. \\
+%     Baseline extinction &  When no disturbance is applied, $\pi_{base}$=0.1.\\
+% \end{tabular}}
+% \end{table}
 

### 2021-10-01 manuscript/supplement_motifs.tex
commented out appendices not referred to yet

@@
 
-\section{SH: Persistence and motif roles}
+% \section{SH: Persistence and motif roles}
 
-    \subsection*{Methods}
+%     \subsection*{Methods}
 
-        As one more level of detail, we can consider the frequency with which a species appears in each \emph{position} in each motif. 
-        Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.
-        The full set of these positions, across all motifs, is the species' \emph{motif role}~\citep{Stouffer2010,Cirtwill2017}.
-        As with motif participation, we normalized motif roles by dividing the number of times a species appeared in each position by the total number of times it appeared in all positions~\citep{Baker2015,Cirtwill2017}.
+%         As one more level of detail, we can consider the frequency with which a species appears in each \emph{position} in each motif. 
+%         Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.
+%         The full set of these positions, across all motifs, is the species' \emph{motif role}~\citep{Stouffer2010,Cirtwill2017}.
+%         As with motif participation, we normalized motif roles by dividing the number of times a species appeared in each position by the total number of times it appeared in all positions~\citep{Baker2015,Cirtwill2017}.
 
-        [[Could add a PERMANOVA to show that species' roles overal affect their persistence. Haven't done so yet.]]
+%         [[Could add a PERMANOVA to show that species' roles overal affect their persistence. Haven't done so yet.]]
 
-        To test whether participating in different motif positions was related to a species' probability of persistence, we fit similar models to those for motif participation.
-        For each position, we fit a linear mixed-effect model relating persistence to the frequency of the position in a species' motif role, disturbance, and their interaction.
-        To account for differences in persistence across networks of different size and connectance, we included a random effect of the interaction between size and connectance.
-        All regressions were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+%         To test whether participating in different motif positions was related to a species' probability of persistence, we fit similar models to those for motif participation.
+%         For each position, we fit a linear mixed-effect model relating persistence to the frequency of the position in a species' motif role, disturbance, and their interaction.
+%         To account for differences in persistence across networks of different size and connectance, we included a random effect of the interaction between size and connectance.
+%         All regressions were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
 
-    \subsection*{Results}
+%     \subsection*{Results}
 
-    \begin{table}[h!]
-        \caption{Coefficients in a series of regressions relating probability of persistence to the frequency of a position in a species' motif role, disturbance, and their interaction.
+%     \begin{table}[h!]
+%         \caption{Coefficients in a series of regressions relating probability of persistence to the frequency of a position in a species' motif role, disturbance, and their interaction.
 For ease of comparison between motifs, we did not center or scale predictors.
 Note, however, that some motifs and positions have higher frequencies than others.}
-        \label{tab:persistence_vs_positions}
-        \centering
-        \begin{tabular}{l | c c c c}
-        Position    & Intercept & Disturbance   & Position  & Interaction \\
-        \hline
-        Chain: top  & 0.880 &   -0.985 & -0.00544 & -0.332 \\
-        Chain: bottom   & 0.883 &   -1.07 & -0.0510 &   0.647 \\
-        Chain: middle   & 0.862 & -1.00 &   0.246 & -0.244 \\
-        \hline
-        Dir. Comp.: top & 0.872 &   -0.995  & 0.0549 &  -0.178 \\
-        Dir. Comp.: bottom & 0.886 &    -1.07 & -0.113 &    0.939 \\
-        \hline
-        Omnivory: top   & 0.872 &   -0.984 &    0.180 & -0.818 \\
-        Omnivory: middle & 0.873 & -1.00 &  0.119 & -0.324 \\
-        Omnivory: bottom & 0.882 & -1.06 &  -0.0449 &   0.845 \\
-        \hline
-        App. Comp.: top & 0.865 &   -0.966 &    0.140 & -0.520 \\
-        App. Comp.: bottom & 0.910 &    -1.12 & -0.100 &    0.323 \\
-        \hline
-        \end{tabular}
-    \end{table}
+%         \label{tab:persistence_vs_positions}
+%         \centering
+%         \begin{tabular}{l | c c c c}
+%         Position    & Intercept & Disturbance   & Position  & Interaction \\
+%         \hline
+%         Chain: top  & 0.880 &   -0.985 & -0.00544 & -0.332 \\
+%         Chain: bottom   & 0.883 &   -1.07 & -0.0510 &   0.647 \\
+%         Chain: middle   & 0.862 & -1.00 &   0.246 & -0.244 \\
+%         \hline
+%         Dir. Comp.: top & 0.872 &   -0.995  & 0.0549 &  -0.178 \\
+%         Dir. Comp.: bottom & 0.886 &    -1.07 & -0.113 &    0.939 \\
+%         \hline
+%         Omnivory: top   & 0.872 &   -0.984 &    0.180 & -0.818 \\
+%         Omnivory: middle & 0.873 & -1.00 &  0.119 & -0.324 \\
+%         Omnivory: bottom & 0.882 & -1.06 &  -0.0449 &   0.845 \\
+%         \hline
+%         App. Comp.: top & 0.865 &   -0.966 &    0.140 & -0.520 \\
+%         App. Comp.: bottom & 0.910 &    -1.12 & -0.100 &    0.323 \\
+%         \hline
+%         \end{tabular}
+%     \end{table}
 
-    \begin{figure}
-        \centering
-        \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Apparent.eps}
-        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
+%     \begin{figure}
+%         \centering
+%         \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Apparent.eps}
+%         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
 Here we show the relationships between persistence and the proportions of the three positions in the apparent competition motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
@@
 }
-        \label{fig:apparent_positions}
-    \end{figure}
+%         \label{fig:apparent_positions}
+%     \end{figure}
 
-    \begin{figure}
-        \centering
-        \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Direct.eps}
-        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
+%     \begin{figure}
+%         \centering
+%         \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Direct.eps}
+%         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
 Here we show the relationships between persistence and the proportions of the three positions in the direct competition motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
@@
 }
-        \label{fig:direct_positions}
-    \end{figure}
+%         \label{fig:direct_positions}
+%     \end{figure}
 
-    \begin{figure}
-        \centering
-        \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Three-species.eps}
-        \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
+%     \begin{figure}
+%         \centering
+%         \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Three-species.eps}
+%         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
 Here we show the relationships between persistence and the proportions of the three positions in the three-species chain motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
@@
 }
-        \label{fig:chain_positions}
-    \end{figure}
+%         \label{fig:chain_positions}
+%     \end{figure}
 
-\clearpage
+% \clearpage
 
-\section{SI: Motif roles and network properties}
+% \section{SI: Motif roles and network properties}
 
-    \subsection*{Methods}
+%     \subsection*{Methods}
 
-        As with network motif profiles and motif participation, species' motif roles may vary with network size and connectance and with the species' in degree and trophic level.
-        To identify these relationships, we fit two linear regressions for each motif position.
-        The first related the frequency of the position in species' roles to network size, connectance, and their interaction.
-        The second related the frequency of the position to in-degree, STL, and their interaction.
-        All regressions were fit using the R~\citep{R} base function `lm'.
+%         As with network motif profiles and motif participation, species' motif roles may vary with network size and connectance and with the species' in degree and trophic level.
+%         To identify these relationships, we fit two linear regressions for each motif position.
+%         The first related the frequency of the position in species' roles to network size, connectance, and their interaction.
+%         The second related the frequency of the position to in-degree, STL, and their interaction.
+%         All regressions were fit using the R~\citep{R} base function `lm'.
 
-    \subsection*{Results}
-        \subsection*{Degree and trophic level}
+%     \subsection*{Results}
+%         \subsection*{Degree and trophic level}
 
-            Positions in all three motifs were strongly related to in-degree and/or trophic level (Table~\ref{DT_tab}).
-            The top position in the omnivory motif increased with increasing degree across all trophic levels.
-            The frequency of the middle position increased with the interaction of degree and trophic level.
-            The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).
+%             Positions in all three motifs were strongly related to in-degree and/or trophic level (Table~\ref{DT_tab}).
+%             The top position in the omnivory motif increased with increasing degree across all trophic levels.
+%             The frequency of the middle position increased with the interaction of degree and trophic level.
+%             The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).
 
-            The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
-            The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.
+%             The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
+%             The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.
 
-            The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
-            The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).
+%             The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
+%             The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).
 
-            The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
-            The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
-            The top position in the three-species chain increased with trophic level, degree, and their interaction.
+%             The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
+%             The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
+%             The top position in the three-species chain increased with trophic level, degree, and their interaction.
 
-            \begin{table}[h!]
-                \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against in-degree, STL, and their interaction.}
-                \label{DT_tab}
-                \footnotesize
-                \begin{tabular}{c c | c c | cc | cc |}
-                Motif & Position & \multicolumn{2}{c}{Degree} & \multicolumn{2}{c}{STL} & \multicolumn{2}{c}{Interaction} \\
-                & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
-                \hline
-                Chain & bottom & -1.47 $\times10^{-5}$ & 0.769 & -0.0218 & \textless0.001 & -0.00160 & \textless0.001 \\
-                Chain & middle & 3.13$\times10^{-4}$ & \textless0.001 & -0.0353 & \textless0.001 & 1.85$\times10^{-5}$ & 0.275 \\
-                Chain & top & -0.0172 & \textless0.001 & 0.0259 & \textless0.001 & 0.00919 & \textless0.001 \\
-                \hline
-                Omnivory & bottom & 6.14$\times10^{-4}$ & \textless0.001 & -0.00270 & \textless0.001 & -0.00112 & \textless0.001 \\
-                Omnivory & middle & 3.13$\times10^{-4}$ & \textless0.001 & -0.0353 & \textless0.001 & 1.85$\times10^{-5}$ & \textless0.001 \\
-                Omnivory & top & 0.00413 & \textless0.001 & 0.00381 & \textless0.001 & 9.96$\times10^{-5}$ & \textless0.001 \\
-                \hline
-                App. comp. & bottom & -4.74$\times10^{-4}$ & \textless0.001 & -0.663 & \textless0.001 & 0.00224 & \textless0.001 \\
-                App. comp. & top & 2.12$\times10^{-5}$ & 0.183 & -0.112 & \textless0.001 & -4.47$\times10^{-4}$ & \textless0.001 \\
-                \hline
-                Dir. comp. & bottom & 8.12$\times10^{-4}$ & \textless0.001 & -5.57$\times10^{-3}$ & \textless0.001 & -1.37$\times10^{-3}$ & \textless0.001 \\
-                Dir. comp. & top & -0.00405 & \textless0.001 & -0.0435 & \textless0.001 & 0.00260 & \textless0.001 \\
-                \hline
-                \end{tabular}
-                \end{table}
+%             \begin{table}[h!]
+%                 \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against in-degree, STL, and their interaction.}
+%                 \label{DT_tab}
+%                 \footnotesize
+%                 \begin{tabular}{c c | c c | cc | cc |}
+%                 Motif & Position & \multicolumn{2}{c}{Degree} & \multicolumn{2}{c}{STL} & \multicolumn{2}{c}{Interaction} \\
+%                 & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
+%                 \hline
+%                 Chain & bottom & -1.47 $\times10^{-5}$ & 0.769 & -0.0218 & \textless0.001 & -0.00160 & \textless0.001 \\
+%                 Chain & middle & 3.13$\times10^{-4}$ & \textless0.001 & -0.0353 & \textless0.001 & 1.85$\times10^{-5}$ & 0.275 \\
+%                 Chain & top & -0.0172 & \textless0.001 & 0.0259 & \textless0.001 & 0.00919 & \textless0.001 \\
+%                 \hline
+%                 Omnivory & bottom & 6.14$\times10^{-4}$ & \textless0.001 & -0.00270 & \textless0.001 & -0.00112 & \textless0.001 \\
+%                 Omnivory & middle & 3.13$\times10^{-4}$ & \textless0.001 & -0.0353 & \textless0.001 & 1.85$\times10^{-5}$ & \textless0.001 \\
+%                 Omnivory & top & 0.00413 & \textless0.001 & 0.00381 & \textless0.001 & 9.96$\times10^{-5}$ & \textless0.001 \\
+%                 \hline
+%                 App. comp. & bottom & -4.74$\times10^{-4}$ & \textless0.001 & -0.663 & \textless0.001 & 0.00224 & \textless0.001 \\
+%                 App. comp. & top & 2.12$\times10^{-5}$ & 0.183 & -0.112 & \textless0.001 & -4.47$\times10^{-4}$ & \textless0.001 \\
+%                 \hline
+%                 Dir. comp. & bottom & 8.12$\times10^{-4}$ & \textless0.001 & -5.57$\times10^{-3}$ & \textless0.001 & -1.37$\times10^{-3}$ & \textless0.001 \\
+%                 Dir. comp. & top & -0.00405 & \textless0.001 & -0.0435 & \textless0.001 & 0.00260 & \textless0.001 \\
+%                 \hline
+%                 \end{tabular}
+%                 \end{table}
 
-            \begin{figure}[h!]
-                \centering
-                \includegraphics[width=.75\textwidth]{figures/positions_byTL_Omnivory.eps}
-                \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
-                \label{omnivory_DT}
-            \end{figure}
+%             \begin{figure}[h!]
+%                 \centering
+%                 \includegraphics[width=.75\textwidth]{figures/positions_byTL_Omnivory.eps}
+%                 \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
+%                 \label{omnivory_DT}
+%             \end{figure}
 
-            \begin{figure}[h!]
-                \centering
-                \includegraphics[width=.75\textwidth]{figures/positions_byTL_Apparentcompetition.eps}
-                \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
-                \label{appcomp_DT}
-            \end{figure}
+%             \begin{figure}[h!]
+%                 \centering
+%                 \includegraphics[width=.75\textwidth]{figures/positions_byTL_Apparentcompetition.eps}
+%                 \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
+%                 \label{appcomp_DT}
+%             \end{figure}
 
-            \begin{figure}[h!]
-                \centering
-                \includegraphics[width=.75\textwidth]{figures/positions_byTL_Directcompetition.eps}
-                \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
-                \label{dircomp_DT}
-            \end{figure}
+%             \begin{figure}[h!]
+%                 \centering
+%                 \includegraphics[width=.75\textwidth]{figures/positions_byTL_Directcompetition.eps}
+%                 \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
+%                 \label{dircomp_DT}
+%             \end{figure}
 
-            \begin{figure}[h!]
-                \centering
-                \includegraphics[width=.75\textwidth]{figures/positions_byTL_Three-specieschain.eps}
-                \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
-                \label{chain_DT}
-            \end{figure}
+%             \begin{figure}[h!]
+%                 \centering
+%                 \includegraphics[width=.75\textwidth]{figures/positions_byTL_Three-specieschain.eps}
+%                 \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
+%                 \label{chain_DT}
+%             \end{figure}
 
-        \subsection*{Network size and connectance}
+%         \subsection*{Network size and connectance}
 
-            % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
-            The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance (Table~\ref{SC_tab}).
-            The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.
+%             % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
+%             The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance (Table~\ref{SC_tab}).
+%             The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.
 
-            The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
-            The frequency of the bottom position in the direct competition motif was low in all networks.
+%             The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
+%             The frequency of the bottom position in the direct competition motif was low in all networks.
 
-            The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
-            Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
-            None of the positions in the three-species chain varied strongly with network size or connectance.
+%             The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
+%             Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
+%             None of the positions in the three-species chain varied strongly with network size or connectance.
 
-            \begin{table}[h!]
-                \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against network size, connectance, and their interaction.}
-                \label{SC_tab}
-                \footnotesize
-                \begin{tabular}{c c | c c | cc | cc |}
-                Motif & Position & \multicolumn{2}{c}{Size} & \multicolumn{2}{c}{Connectance} & \multicolumn{2}{c}{Interaction} \\
-                & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
-                \hline
-                Chain & bottom & -4.63$\times10^{-5}$ & \textless0.001 & -0.108 & \textless0.001 & -1.94$\times10^{-4}$ & 0.035 \\
-                Chain & middle & 1.68$\times10^{-4}$ & \textless0.001 & -0.0530 & \textless0.001 & -0.00212 & \textless0.001 \\
-                Chain & top & 1.00$\times10^{-4}$ & \textless0.001 & 0.0304 & 0.015 & -8.79$\times10^{-4}$ & \textless0.001 \\
-                \hline
-                Omnivory & bottom & 2.00$\times10^{-4}$ & \textless0.001 & 0.575 & \textless0.001 & -0.00240 & \textless0.001 \\
-                Omnivory & middle & 1.32$\times10^{-4}$ & \textless0.001 & 0.551 & \textless0.001 & -0.00215 & \textless0.001 \\
-                Omnivory & top & 6.42$\times10^{-5}$ & \textless0.001 & 0.448 & \textless0.001 & -0.00115 & \textless0.001 \\
-                \hline
-                App. comp. & bottom & -4.74$\times10^{-4}$ & \textless0.001 & -0.663 & \textless0.001 & 0.00224 & \textless0.001 \\
-                App. comp. & top & 2.12$\times10^{-5}$ & 0.183 & -0.112 & \textless0.001 & -4.47$\times10^{-4}$ & \textless0.001 \\
-                \hline
-                Dir. comp. & bottom & 7.94$\times10^{-5}$ & \textless0.001 & -0.0663 & \textless0.001 & 0.00101 & \textless0.001 \\
-                Dir. comp. & top & -2.45$\times10^{-4}$ & \textless0.001 & -0.602 & \textless0.001 & 0.00610 & \textless0.001 \\
-                \hline
-                \end{tabular}
-                \end{table}
+%             \begin{table}[h!]
+%                 \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against network size, connectance, and their interaction.}
+%                 \label{SC_tab}
+%                 \footnotesize
+%                 \begin{tabular}{c c | c c | cc | cc |}
+%                 Motif & Position & \multicolumn{2}{c}{Size} & \multicolumn{2}{c}{Connectance} & \multicolumn{2}{c}{Interaction} \\
+%                 & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
+%                 \hline
+%                 Chain & bottom & -4.63$\times10^{-5}$ & \textless0.001 & -0.108 & \textless0.001 & -1.94$\times10^{-4}$ & 0.035 \\
+%                 Chain & middle & 1.68$\times10^{-4}$ & \textless0.001 & -0.0530 & \textless0.001 & -0.00212 & \textless0.001 \\
+%                 Chain & top & 1.00$\times10^{-4}$ & \textless0.001 & 0.0304 & 0.015 & -8.79$\times10^{-4}$ & \textless0.001 \\
+%                 \hline
+%                 Omnivory & bottom & 2.00$\times10^{-4}$ & \textless0.001 & 0.575 & \textless0.001 & -0.00240 & \textless0.001 \\
+%                 Omnivory & middle & 1.32$\times10^{-4}$ & \textless0.001 & 0.551 & \textless0.001 & -0.00215 & \textless0.001 \\
+%                 Omnivory & top & 6.42$\times10^{-5}$ & \textless0.001 & 0.448 & \textless0.001 & -0.00115 & \textless0.001 \\
+%                 \hline
+%                 App. comp. & bottom & -4.74$\times10^{-4}$ & \textless0.001 & -0.663 & \textless0.001 & 0.00224 & \textless0.001 \\
+%                 App. comp. & top & 2.12$\times10^{-5}$ & 0.183 & -0.112 & \textless0.001 & -4.47$\times10^{-4}$ & \textless0.001 \\
+%                 \hline
+%                 Dir. comp. & bottom & 7.94$\times10^{-5}$ & \textless0.001 & -0.0663 & \textless0.001 & 0.00101 & \textless0.001 \\
+%                 Dir. comp. & top & -2.45$\times10^{-4}$ & \textless0.001 & -0.602 & \textless0.001 & 0.00610 & \textless0.001 \\
+%                 \hline
+%                 \end{tabular}
+%                 \end{table}
 
-            \begin{figure}[h!]
-                \centering
-                \includegraphics[width=.75\textwidth]{figures/positions_bySC_Omnivory.eps}
-                \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
-                \label{omnivory_SC}
-            \end{figure}
+%             \begin{figure}[h!]
+%                 \centering
+%                 \includegraphics[width=.75\textwidth]{figures/positions_bySC_Omnivory.eps}
+%                 \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
+%                 \label{omnivory_SC}
+%             \end{figure}
 
-            \begin{figure}[h!]
-                \centering
-                \includegraphics[width=.75\textwidth]{figures/positions_bySC_Apparentcompetition.eps}
-                \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
-                \label{appcomp_SC}
-            \end{figure}
+%             \begin{figure}[h!]
+%                 \centering
+%                 \includegraphics[width=.75\textwidth]{figures/positions_bySC_Apparentcompetition.eps}
+%                 \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
+%                 \label{appcomp_SC}
+%             \end{figure}
 
-            \begin{figure}[h!]
-                \centering
-                \includegraphics[width=.75\textwidth]{figures/positions_bySC_Directcompetition.eps}
-                \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
-                \label{dircomp_SC}
-            \end{figure}
+%             \begin{figure}[h!]
+%                 \centering
+%                 \includegraphics[width=.75\textwidth]{figures/positions_bySC_Directcompetition.eps}
+%                 \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
+%                 \label{dircomp_SC}
+%             \end{figure}
 
-            \begin{figure}[h!]
-                \centering
-                \includegraphics[width=.75\textwidth]{figures/positions_bySC_Three-specieschain.eps}
-                \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
-                \label{chain_SC}
-            \end{figure}
+%             \begin{figure}[h!]
+%                 \centering
+%                 \includegraphics[width=.75\textwidth]{figures/positions_bySC_Three-specieschain.eps}
+%                 \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
+%                 \label{chain_SC}
+%             \end{figure}
 
-    \clearpage
+%     \clearpage
 
-\clearpage
+% \clearpage
 
-\section*{SJ: Proportion of basal resources in the network}
+% \section*{SJ: Proportion of basal resources in the network}
 
-    \subsection*{Basal resources and mean persistence}
+%     \subsection*{Basal resources and mean persistence}
 
-        We fit a linear regression relating consumer persistence to network size, connectance, disturbance, the proportion of the network made up of basal resources (pBasal), and all interactions between them.
+%         We fit a linear regression relating consumer persistence to network size, connectance, disturbance, the proportion of the network made up of basal resources (pBasal), and all interactions between them.
 All terms were significant, indicating that the pBasal affects a consumer's probability of persistence but that this effect depends upon other aspects of network structure.
-        As with the effects of network size and connectance, however, the effects of pBasal were small in the range of networks we considered here (Fig.~\ref{fig:lm_BCS}).
-        % Table is from basal_models.Rdata: BCper, if we want it.
+%         As with the effects of network size and connectance, however, the effects of pBasal were small in the range of networks we considered here (Fig.~\ref{fig:lm_BCS}).
+%         % Table is from basal_models.Rdata: BCper, if we want it.
 
-        \begin{figure}[h!]
-            \centering
-            \includegraphics[height=0.75\textheight]{figures/persistence_vs_BSC_lm.eps}
-            \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while the proportion of the network made up of basal resources, network size, connectance, and their interactions had smaller effects.
+%         \begin{figure}[h!]
+%             \centering
+%             \includegraphics[height=0.75\textheight]{figures/persistence_vs_BSC_lm.eps}
+%             \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while the proportion of the network made up of basal resources, network size, connectance, and their interactions had smaller effects.
 We show relationships with persistence for low disturbance ($\pi_{Basal}$=0.1) and high disturbance($\pi_{Basal}$=0.5).
@@
 For proportion of basal resources, these values were 0.02 and 0.5; for network size, 50 and 100; and for connectance, 0.02 and 0.18.}
-            \label{fig:lm_BCS}
-        \end{figure}
+%             \label{fig:lm_BCS}
+%         \end{figure}
 
-    \subsection*{Motif profiles and proportion of basal resources}
+%     \subsection*{Motif profiles and proportion of basal resources}
 
-        We repeated our analyses testing whether motif profiles were related to network size and connectance, replacing network size with the proportion of the network made up by basal resources (pBasal).
-        A network's overall motif profile was significantly related to pBasal ($F_{1,2996}$=196, $p$\textless0.001, connecance ($F_{1,2996}$=10.0, $p$\textless0.001), and their interaction ($F_{1,2996}$=27.3, $p$\textless0.001).
-        However, this result could be a false positive as the dispersion of motif profiles was non-homogeneous across connectance ($F_{4,2995}$=5.17, $p$\textless0.001), pBasal ($F_{128,2871}$=19.0, $p$\textless0.001), and their interaction ($F_{291,2708}$=13.4, $p$\textless0.001).
-        Distance from group centroids was low for networks with extremely low or extremely high pBasal, perhaps because there were few such networks.
-        The ranges and median distances from group centroids was similar across levels of connectance.
+%         We repeated our analyses testing whether motif profiles were related to network size and connectance, replacing network size with the proportion of the network made up by basal resources (pBasal).
+%         A network's overall motif profile was significantly related to pBasal ($F_{1,2996}$=196, $p$\textless0.001, connecance ($F_{1,2996}$=10.0, $p$\textless0.001), and their interaction ($F_{1,2996}$=27.3, $p$\textless0.001).
+%         However, this result could be a false positive as the dispersion of motif profiles was non-homogeneous across connectance ($F_{4,2995}$=5.17, $p$\textless0.001), pBasal ($F_{128,2871}$=19.0, $p$\textless0.001), and their interaction ($F_{291,2708}$=13.4, $p$\textless0.001).
+%         Distance from group centroids was low for networks with extremely low or extremely high pBasal, perhaps because there were few such networks.
+%         The ranges and median distances from group centroids was similar across levels of connectance.
 
-        Considered separately, the proportions of all four motifs in a network's motif profile were related to the proportion of basal resources in a network (Fig.~\ref{fig:Blms}).
-        The proportion of the three-species chain decreased in networks with more basal resources ($\beta_{pBasal}$=-0.080, $p$\textless0.001) and networks with higher connectance ($\beta_{C}$=0.498, $p$\textless0.001).
-        The interaction between pBasal and connectance was not significant ($\beta_{pBasal:C}$=0.0554, $p$=0.766).
+%         Considered separately, the proportions of all four motifs in a network's motif profile were related to the proportion of basal resources in a network (Fig.~\ref{fig:Blms}).
+%         The proportion of the three-species chain decreased in networks with more basal resources ($\beta_{pBasal}$=-0.080, $p$\textless0.001) and networks with higher connectance ($\beta_{C}$=0.498, $p$\textless0.001).
+%         The interaction between pBasal and connectance was not significant ($\beta_{pBasal:C}$=0.0554, $p$=0.766).
 
-        The proportion of the omnivory motif decreased in networks with more basal resources ($\beta_{pBasal}$=-0.198, $p$\textless0.001), increased in networks with higher connectance ($\beta_{C}$=0.895, $p$\textless0.001), and increased most strongly in networks with high connectance and many basal resources ($\beta_{pBasal:C}$=2.53, $p$\textless0.001).
+%         The proportion of the omnivory motif decreased in networks with more basal resources ($\beta_{pBasal}$=-0.198, $p$\textless0.001), increased in networks with higher connectance ($\beta_{C}$=0.895, $p$\textless0.001), and increased most strongly in networks with high connectance and many basal resources ($\beta_{pBasal:C}$=2.53, $p$\textless0.001).
 
-        The proportion of the apparent competition motif increased with in networks with more basal resources ($\beta_{pBasal}$=0.202, $p$\textless0.001) and decreased in more connected networks ($\beta_C$=-0.596, $p$\textless0.001).
-        The interaction between the proportion of basal resources and connectance was not significant ($\beta_{pBasal:C}$=0.258, $p$\textless0.001).
+%         The proportion of the apparent competition motif increased with in networks with more basal resources ($\beta_{pBasal}$=0.202, $p$\textless0.001) and decreased in more connected networks ($\beta_C$=-0.596, $p$\textless0.001).
+%         The interaction between the proportion of basal resources and connectance was not significant ($\beta_{pBasal:C}$=0.258, $p$\textless0.001).
 
-        The proportion of the direct competition motif increased in networks with more basal resources ($\beta_{pBasal}$=0.0759, $p$\textless0.001) and more-connected networks ($\beta_C$=0.199, $p$\textless0.001), but decreased with the interaction between them ($\beta_{pBasal:C}$=-2.85, $p$\textless0.001).
+%         The proportion of the direct competition motif increased in networks with more basal resources ($\beta_{pBasal}$=0.0759, $p$\textless0.001) and more-connected networks ($\beta_C$=0.199, $p$\textless0.001), but decreased with the interaction between them ($\beta_{pBasal:C}$=-2.85, $p$\textless0.001).
 
-        \begin{figure}
-            \centering
-            \includegraphics[width=.75\textwidth]{figures/motif_proportion_basal_lms.eps}
-            \caption{Mean proportion of each motif in a network's motif profile varied with connectance and the proportion of basal resources in the network.
+%         \begin{figure}
+%             \centering
+%             \includegraphics[width=.75\textwidth]{figures/motif_proportion_basal_lms.eps}
+%             \caption{Mean proportion of each motif in a network's motif profile varied with connectance and the proportion of basal resources in the network.
 For direct competition and omnivory, the frequency of the motif also varied with the interaction between connectance and basal resources.}
-            \label{fig:Blms}
-        \end{figure}
+%             \label{fig:Blms}
+%         \end{figure}
 
-    \subsection*{Motif participation and proportion of basal resources}
+%     \subsection*{Motif participation and proportion of basal resources}
 
-        We fit regressions relating the proportion of each motif in a consumer's motif participation vector to the proportion of basal resources in the network (pBasal), connectance, and the interaction between them.
-        Participation in all four motifs was significantly related to pBasal, connectance, and the interaction between them (Fig.~\ref{fig:basal_participation}).
-        These interactions were strongest for apparent and direct competition.
+%         We fit regressions relating the proportion of each motif in a consumer's motif participation vector to the proportion of basal resources in the network (pBasal), connectance, and the interaction between them.
+%         Participation in all four motifs was significantly related to pBasal, connectance, and the interaction between them (Fig.~\ref{fig:basal_participation}).
+%         These interactions were strongest for apparent and direct competition.
 
-        \begin{figure}
-            \centering
-            \includegraphics[height=.65\textheight]{figures/participation_vs_BC.eps}
-            \caption{The proportion of each motif in a consumer's participation vector varied with the proportion of the network made up by basal resources, connectance, and their interaction.}
-            \label{fig:basal_participation}
-        \end{figure}
+%         \begin{figure}
+%             \centering
+%             \includegraphics[height=.65\textheight]{figures/participation_vs_BC.eps}
+%             \caption{The proportion of each motif in a consumer's participation vector varied with the proportion of the network made up by basal resources, connectance, and their interaction.}
+%             \label{fig:basal_participation}
+%         \end{figure}
 
-        Participation in the three-species chain increased in networks with more basal resources ($\beta_{pBasal}$=0.0447, $p$\textless0.001), decreased in more-connected networks ($\beta_C$=-0.296, $p$\textless0.001), and decreased with the interaction between them ($\beta_{pBasal:C}$=-0.525, $p$\textless0.001).
-        Participation in the omnivory motif decreased in networks with more basal resources ($\beta_{pBasal}$=-0.198, $p$\textless0.001), increased in more-connected networks ($\beta_C$=0.701, $p$\textless0.001), and increased strongly with the interaction between them ($\beta_{pBasal:C}$=2.81, $p$\textless0.001).
-        Participation in the apparent competition motif decreased in networks with more basal resources and more-connected networks ($\beta_{pBasal}$=-0.0825, $p$\textless0.001 and $\beta_C$=-0.846, $p$\textless0.001) but increased strongly with the interaction between them ($\beta_{pBasal:C}$=1.66, $p$\textless0.001).
-        Participation in the direct competition motif increased in networks with more basal resources and more-connected networks ($\beta_{pBasal}$=0.235, $p$\textless0.001 and $\beta_C$=0.442, $p$\textless0.002), but decreased strongly with the interaction between them ($\beta_{pBasal:C}$=-3.94, $p$\textless0.001).
+%         Participation in the three-species chain increased in networks with more basal resources ($\beta_{pBasal}$=0.0447, $p$\textless0.001), decreased in more-connected networks ($\beta_C$=-0.296, $p$\textless0.001), and decreased with the interaction between them ($\beta_{pBasal:C}$=-0.525, $p$\textless0.001).
+%         Participation in the omnivory motif decreased in networks with more basal resources ($\beta_{pBasal}$=-0.198, $p$\textless0.001), increased in more-connected networks ($\beta_C$=0.701, $p$\textless0.001), and increased strongly with the interaction between them ($\beta_{pBasal:C}$=2.81, $p$\textless0.001).
+%         Participation in the apparent competition motif decreased in networks with more basal resources and more-connected networks ($\beta_{pBasal}$=-0.0825, $p$\textless0.001 and $\beta_C$=-0.846, $p$\textless0.001) but increased strongly with the interaction between them ($\beta_{pBasal:C}$=1.66, $p$\textless0.001).
+%         Participation in the direct competition motif increased in networks with more basal resources and more-connected networks ($\beta_{pBasal}$=0.235, $p$\textless0.001 and $\beta_C$=0.442, $p$\textless0.002), but decreased strongly with the interaction between them ($\beta_{pBasal:C}$=-3.94, $p$\textless0.001).
 
-    \subsection*{Frequencies of positions and proportion of basal resources}
+%     \subsection*{Frequencies of positions and proportion of basal resources}
 
-        As with the frequencies of each motif in a species' motif participation vector, the frequencies of positions in a species' motif role varied with the proportion of basal resources in the network, connectance, and their interaction (Table~\ref{BC_tab}).
+%         As with the frequencies of each motif in a species' motif participation vector, the frequencies of positions in a species' motif role varied with the proportion of basal resources in the network, connectance, and their interaction (Table~\ref{BC_tab}).
 
-        \begin{table}[h!]
-            \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against the proportion of the network made up of basal resources (pBasal), connectance, and their interaction.}
-            \label{BC_tab}
-            \footnotesize
-            \begin{tabular}{c c | c c | cc | cc |}
-            Motif & Position & \multicolumn{2}{c}{pBasal} & \multicolumn{2}{c}{Connectance} & \multicolumn{2}{c}{Interaction} \\
-            & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
-            \hline
-            Chain & bottom & -0.116 & \textless0.001 & -0.266 & \textless0.001 & 0.217 & \textless0.001 \\
-            Chain & middle & 0.0779 & \textless0.001 & -0.107 & \textless0.001 & -0.386 & \textless0.001 \\
-            Chain & top & 0.0829 & \textless0.001 & 0.0776 & \textless0.001 & -0.356 & \textless0.001 \\
-            \hline
-            Omnivory & bottom & -0.104 & \textless0.001 & 0.203 & \textless0.001 & 0.937 & \textless0.001 \\
-            Omnivory & middle & -0.0668 & \textless0.001 & 0.217 & \textless0.001 & 1.24 & \textless0.001 \\
-            Omnivory & top & -0.0269 & \textless0.001 & 0.281 & \textless0.001 & 0.638 & \textless0.001 \\
-            \hline
-            App. comp. & bottom & -0.237 & \textless0.001 & -0.874 & \textless0.001 & 1.76 & \textless0.001 \\
-            App. comp. & top & 0.154 & \textless0.001 & 0.0279 & \textless0.001 & -0.0947 & 0.013 \\
-            \hline
-            Dir. comp. & bottom & -0.0631 & \textless0.001 & -0.0247 & \textless0.001 & -0.418 & \textless0.001 \\
-            Dir. comp. & top & 0.298 & \textless0.001 & 0.466 & \textless0.001 & =3.53 & \textless0.001 \\
-            \hline
-            \end{tabular}
-            \end{table}
-\clearpage
+%         \begin{table}[h!]
+%             \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against the proportion of the network made up of basal resources (pBasal), connectance, and their interaction.}
+%             \label{BC_tab}
+%             \footnotesize
+%             \begin{tabular}{c c | c c | cc | cc |}
+%             Motif & Position & \multicolumn{2}{c}{pBasal} & \multicolumn{2}{c}{Connectance} & \multicolumn{2}{c}{Interaction} \\
+%             & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
+%             \hline
+%             Chain & bottom & -0.116 & \textless0.001 & -0.266 & \textless0.001 & 0.217 & \textless0.001 \\
+%             Chain & middle & 0.0779 & \textless0.001 & -0.107 & \textless0.001 & -0.386 & \textless0.001 \\
+%             Chain & top & 0.0829 & \textless0.001 & 0.0776 & \textless0.001 & -0.356 & \textless0.001 \\
+%             \hline
+%             Omnivory & bottom & -0.104 & \textless0.001 & 0.203 & \textless0.001 & 0.937 & \textless0.001 \\
+%             Omnivory & middle & -0.0668 & \textless0.001 & 0.217 & \textless0.001 & 1.24 & \textless0.001 \\
+%             Omnivory & top & -0.0269 & \textless0.001 & 0.281 & \textless0.001 & 0.638 & \textless0.001 \\
+%             \hline
+%             App. comp. & bottom & -0.237 & \textless0.001 & -0.874 & \textless0.001 & 1.76 & \textless0.001 \\
+%             App. comp. & top & 0.154 & \textless0.001 & 0.0279 & \textless0.001 & -0.0947 & 0.013 \\
+%             \hline
+%             Dir. comp. & bottom & -0.0631 & \textless0.001 & -0.0247 & \textless0.001 & -0.418 & \textless0.001 \\
+%             Dir. comp. & top & 0.298 & \textless0.001 & 0.466 & \textless0.001 & =3.53 & \textless0.001 \\
+%             \hline
+%             \end{tabular}
+%             \end{table}
+% \clearpage
 
-\section*{SK: Motif profiles in different network sets}
+% \section*{SK: Motif profiles in different network sets}
 
-    Not all motifs were equally likely to occur in the networks we considered. 
-    Across the initially stable, acyclic networks (i.e., those used in the main text), the omnivory motif tended to make up a smaller proportion of the motif profile than other motifs.
-    The omnivory motif was also rarer than the three-species chain and apparent competition motifs before networks were rendered acyclic, but made up a similar proportion to the direct competition motif (Table~\ref{tab:acyclic_proportions}).
-    The omnivory motif made up a larger proportion of the full set of initially stable networks, including those which were excluded from our final analysis due to overly high trophic levels.
-    Nevertheless, in each group of networks the apparent competition and three-species chain motifs made up the largest proportion of the motif profile.
-    This suggests that, while rendering networks acyclic does change proportions slightly, stable networks created by the niche model appear to be generally biased towards the chain and apparent competition motifs.
+%     Not all motifs were equally likely to occur in the networks we considered. 
+%     Across the initially stable, acyclic networks (i.e., those used in the main text), the omnivory motif tended to make up a smaller proportion of the motif profile than other motifs.
+%     The omnivory motif was also rarer than the three-species chain and apparent competition motifs before networks were rendered acyclic, but made up a similar proportion to the direct competition motif (Table~\ref{tab:acyclic_proportions}).
+%     The omnivory motif made up a larger proportion of the full set of initially stable networks, including those which were excluded from our final analysis due to overly high trophic levels.
+%     Nevertheless, in each group of networks the apparent competition and three-species chain motifs made up the largest proportion of the motif profile.
+%     This suggests that, while rendering networks acyclic does change proportions slightly, stable networks created by the niche model appear to be generally biased towards the chain and apparent competition motifs.
 
-    \begin{table}[h!]
-        \centering
-        \caption{Proportions of the four motifs which appear in Bayesian networks (three-species chain: `Chain', apparent competition: `AC', direct competition `DC', and omnivory: `Omni') and other motifs (`Other') in the motif profiles of acyclic, filtered, and original networks.
+%     \begin{table}[h!]
+%         \centering
+%         \caption{Proportions of the four motifs which appear in Bayesian networks (three-species chain: `Chain', apparent competition: `AC', direct competition `DC', and omnivory: `Omni') and other motifs (`Other') in the motif profiles of acyclic, filtered, and original networks.
 Acyclic networks are those used in the main text analyses.
@@
 Original networks are all simulated networks, including those removed because of overly high maximum trophic levels.}
-        \label{tab:acyclic_proportions}            \footnotesize
-        \begin{tabular}{c|l l l | l l l | l l l |}
-        & \multicolumn{3}{c|}{Acyclic networks} & \multicolumn{3}{c|}{Filtered networks} & \multicolumn{3}{c|}{Original networks} \\
-        Motif & Mean (SD) & Min & Max & Mean (SD) & Min & Max & Mean (SD) & Min & Max \\
-        \hline
-        Chain & 0.235 (0.042) & 0.109 & 0.412 & 0.225 (0.045) & 0.100 & 0.412 & 0.218 (0.043) & 0.096 & 0.413 \\
-        AC & 0.422 (0.064) & 0.250 & 0.765 &
-        0.417 (0.067) & 0.233 & 0.765 & 0.397 (0.064) & 0.231 & 0.765 \\
-        DC & 0.179 (0.041) & 0.050 & 0.359 & 0.155 (0.046) & 0.040 & 0.359 & 0.140 (0.046) & 0.031 & 0.359 \\
-        Omni & 0.164 (0.079) & 0.006 & 0.346 & 0.156 (0.075) & 0.006 & 0.319 & 0.184 (0.073) & 0.006 & 0.329 \\
-        \hline
-        Other & \multicolumn{3}{c|}{NA} & 0.047 (0.042) & 0.00 & 0.200 &
-        0.062 (0.046) & 0.00 & 0.258 \\
-        \end{tabular}
-    \end{table}
-\clearpage
+%         \label{tab:acyclic_proportions}            \footnotesize
+%         \begin{tabular}{c|l l l | l l l | l l l |}
+%         & \multicolumn{3}{c|}{Acyclic networks} & \multicolumn{3}{c|}{Filtered networks} & \multicolumn{3}{c|}{Original networks} \\
+%         Motif & Mean (SD) & Min & Max & Mean (SD) & Min & Max & Mean (SD) & Min & Max \\
+%         \hline
+%         Chain & 0.235 (0.042) & 0.109 & 0.412 & 0.225 (0.045) & 0.100 & 0.412 & 0.218 (0.043) & 0.096 & 0.413 \\
+%         AC & 0.422 (0.064) & 0.250 & 0.765 &
+%         0.417 (0.067) & 0.233 & 0.765 & 0.397 (0.064) & 0.231 & 0.765 \\
+%         DC & 0.179 (0.041) & 0.050 & 0.359 & 0.155 (0.046) & 0.040 & 0.359 & 0.140 (0.046) & 0.031 & 0.359 \\
+%         Omni & 0.164 (0.079) & 0.006 & 0.346 & 0.156 (0.075) & 0.006 & 0.319 & 0.184 (0.073) & 0.006 & 0.329 \\
+%         \hline
+%         Other & \multicolumn{3}{c|}{NA} & 0.047 (0.042) & 0.00 & 0.200 &
+%         0.062 (0.046) & 0.00 & 0.258 \\
+%         \end{tabular}
+%     \end{table}
+% \clearpage
 

### 2021-10-04 manuscript/reorder_main_text.tex

@@
 \newcommand{\SI}{\textit{Appendix}~}
-
-\setlength{\abovedisplayskip}{0pt}
-\setlength{\belowdisplayskip}{0pt}
 
@@
 \setlength{\parindent}{15pt}
-\begin{spacing}{2.0}
 
@@
 \section*{Abstract}
+
     Bottom-up effects of disturbance to basal resources can strongly affect the persistence of consumers.
@@
 
-    Anna AA: The whole article is now approximately 4400 words (not counting fig texts); intro 850, methods 1600, results 620, discussion 1300.
+    % Anna AA: The whole article is now approximately 4400 words (not counting fig texts); intro 850, methods 1600, results 620, discussion 1300.
 See discussion on section statistical analysis in the methods, I think we can make it shorter
-    - glossary
-    - refs
-    - ending
+    % - glossary
+    % - refs
+    % - ending
 
-    - Fig. 4 not much in discussion, could be a candidate for the SI? Or add more discussion about apparent competition, etc. in discussion sections per motif
+    % - Fig. 4 not much in discussion, could be a candidate for the SI? Or add more discussion about apparent competition, etc. in discussion sections per motif
+
+\begin{spacing}{2.0}
 
@@
     % basic idea behind BN modeling
-        Bayesian Network modelling can represent species and feeding interactions as the nodes and links of a network (see Appendix \emph{S2} for full details).
+        Bayesian Networks can be used to represent species and their feeding interactions as the nodes and links of a network (see Appendix \emph{S2} for full details).
         In our setting, each species (node) is a Bernoulli random variable, being either extant or extinct. 
@@
 		$\pi_{base}$ represents the risk of going extinct due to diseases, stochastic extinction of small populations, or other factors not related to the food web.
-		$\pi_{base}$ may be constant, or it can vary to reflect greater sensitivity in some groups of species.
-		%In our simulations, we vary $\pi_{base}$ for basal resources between 0.1 and 0.5 but keep constant $\pi_{base}=0.1$ for consumers.
+		$\pi_{base}$ may be constant, or it can vary between species to reflect greater sensitivity in some groups of species.
 
-        If the focal species is a consumer with $n$ resources, its probability of extinction depends on $\pi_{base}$ and on the fraction $f = k/n$ of its resources that have gone extinct ($k$).
-        Note that we assume that only the fraction of resources lost is important and not the identities of the missing resources.
-        We then model the response of the consumer to changes in $f$ using a sigmoid non-linear function ($\alpha > 1, \beta > 1$; Appendix \emph{S2}.
+        A species probability of extinction also depends on the state of its resources.
+If the focal species is a consumer with $n$ resources, of which $k$ have gone extinct, its probability of extinction depends on $\pi_{base}$ and on the fraction $f = k/n$.
+
+        The response of the consumer to changes in $f$ can take any form. Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$; Appendix \emph{S2}.
+
 		This type of response most accurately captures the secondary extinctions produced by dynamical models~\citep{Eklof2013}.
+
 		The probability of extinction for each species $i$ can therefore be represented using the cumulative density function of a beta distribution:
@@
 This will in turn affect the extinction probability $P(\lnot C|f)$.
-		For primary producers, we assume that abiotic resources are always available and $f=0$. Following Equation \ref{betafunc}, the extinction will only depend on $\pi_{base}$.
+		For basal resources, we assume that abiotic resources are always available and $f=0$. Following Equation \ref{betafunc}, the extinction will only depend on $\pi_{base}$.
+
+		Note that we here assume that only the fraction, and not the identity, of the resources lost is important.
+However, the model can easily be modified to incorporate weighted resource interactions \citep[see][]{Eklof2013}.
 
 		% how we calculate persistences 
-        Beginning with primary producers and working systematically upward through the food web, we use Equation \ref{betafunc} to calculate each species' $i$ probability of extinction $P(\lnot i|f)$.
+        Beginning with basal resources and working systematically up through the food web, we use Equation \ref{betafunc} to calculate each species' $i$ probability of extinction $P(\lnot i|f)$.
         Using this probability, we then perform Bernoulli trials to determine whether the species goes extinct or not. 
@@
 	    For each species, we then define probability of persistence as the fraction of simulations in which the species persists.
-	    This process is highly efficient and produces the same result as exact methods of solving Bayesian networks \citep{Haussler2020}.
+There are methods for solving Bayesian networks exactly \citep{Eklof2013}.
+	    However, for larger networks the described numerical evaluation is highly efficient and produces close to identical results \citep{Haussler2020}.
+
+		Modelling species extinctions in food webs using Bayesian networks have two main limitations.
+First, due to the strict bottom-up approach the food webs must be acyclic.
+Second, as consumers are influenced by the state of their resources, but resources are not affected by their consumers, the method does not capture top-down effects.
 
     % various disturbance scenarios
-        In our baseline scenario, primary producers had a baseline extinction probability of $\pi_{base} = 0.1$. 
-		We also simulated a set of scenarios where primary producers had a higher baseline probability of extinction ($\pi_{disturbed}$). 
-		These scenarios reflect threats to primary producers such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
-		$\pi_{disturbed}$ ranged between $0.18-0.5$, in steps of $0.08$. 
+        In our baseline scenario, primary producers had an extinction probability of $\pi_{base} = 0.1$. 
+		To simulate various threats to primary producers, making some species more vulnerable to extinction, we increase the affected species' probability of extinction ($\pi_{disturbed}$). 
+		We varied $\pi_{disturbed}$ between $0.18-0.5$, in steps of $0.08$. 
 		The highest disturbance level, $\pi_{disturbed} = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
@@
 	We tested:
-	\noindent1) how species persistence varied with motif participation (Appendix \emph{3.1}),
+	\noindent1) how species persistence varied with motif participation.
 
-	\noindent2) how motif participation varied with other network properties (Appendix \emph{3.2}),
+	\noindent2) how motif participation varied with other network properties,
 
-	\noindent3) how network mean persistence varied with network motif profiles (Appendix \emph{3.3}),
+	\noindent3) how network mean persistence varied with network motif profiles,
 
-	\noindent4) how network motif profiles varied with global structural properties (Appendix \emph{3.4}), and
+	\noindent4) how network motif profiles varied with global structural properties, and
 
-	\noindent5) how species persistence and network mean persistence varied with other network properties (Appendix \emph{3.5}). 
+	\noindent5) how species persistence and network mean persistence varied with other network properties (see Appendix \emph{3.1 - 3.5}). 
 	All analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan} as detailed in the appendices.
@@
 
-    %We are interested in how a species' participation in different motifs affects its persistence, and how the overall motif profile of a network relate to the average persistence of consumer species.
-We analyse this for different levels of disturbance to the basal level.
-To provide context, we also investigate relationships between persistence and classic global (size and connectance) and local (in-degree and trophic level) network properties.
-
     \subsection*{Species motif participation and species persistence}
@@
 We should refer to SA somewhere in Discussion (will be Appendix 10).
+
 \section*{Discussion}
 
-% Focus on what additional information is revealed from analyzing motif compared to global network properties or species characteristics.
-% * ADD REFS - high disturbance scenario is analogous to decreasing plant diversity
+    The effects of bottom-up disturbances on consumers depend on the global structure of the affected food web~\citep{Dunne2002, Eklof2006, PascualDunne2006}.
+    Likewise, the distribution of meso-scale structures within a network affects the network's stability~\citep{prill2005dynamic, bascompte2005simple}.
+    In addition, both direct and indirect interactions (i.e., the local and meso-scale structure of a food web  [\citealp[]{Cirtwill2018FoodWebs}]) can affect which species are most strongly affected~\citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    Identifying these species is a high priority for conservation efforts~\citep{Bottrilletal2008}.
+    To advance this goal, we systematically explore the relationships between meso-scale structures and species' extinction risk, as well as how these trends overlap with the effects of global- and local-scale network properties.
 
-% Framing: network structure is important for predicting extinction risk
-
-It is well known that bottom-up effects are strong in most ecosystems and that global network properties are important for how well a network can handle disturbances to those energy flows, such as loss of species \citep{Dunne2002, Eklof2006, PascualDunne2006}.
-At the same time, from a conservation perspective it is often more relevant to focus interventions on particular target species~\citep{Bottrilletal2008}. 
-Nevertheless, the process of identifying target species should take the network context into account since both direct and indirect interactions can affect how species respond to a disturbance~\citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
-
-% Framing: motifs are related to network stability, species persistence (our results)
-Global network properties are often too coarse grained and identification of vulnerable species in that context becomes to broad (e.g.
+    % % Framing: motifs are related to network stability, species persistence (our results)
+    % Global network properties are often too coarse grained and identification of vulnerable species in that context becomes to broad (e.g.
 primary producers).
 Motifs represent a middle-ground approach as they are the basic building blocks of networks~\citep{Milo2002}, connecting network-wide (global) and species-specific (local) properties, describing both direct and indirect effects.
-As such, the distribution of these meso-scale structures within a network is of importance to network stability ~\citep{prill2005dynamic, bascompte2005simple}.
-We here show that a species' frequency of participation in different motifs is related to its probability of persistence.
 
     % Species-level results summary
-Specifically, we found that a species' probability of persistence increased with its frequency of participation in the direct competition and omnivory motifs and decreased with its frequency of participation in the three-species chain motif.
-The relationship between persistence and the frequency of participation in the apparent competition motif depended strongly on the level of disturbance.
-These trends can be partially explained by underlying relationships between motif participation and in-degree and/or trophic level.
-For example, the frequency of participation in the direct competition motif was strongly and negatively correlated with trophic level.
-Accordingly, while a high trophic level is associated with lower persistence at all levels of disturbance, the opposite is true for participation in the direct competition motif.
+    A species' frequent participation in the direct competition and omnivory motifs was associated with a higher probability of persistence while participation in the three-species chain motif was associated with lower persistence, regardless of the strength of disturbance.
+    The strength of disturbance strongly affected the relationship between persistence and participation in the apparent competition motif.
+    These trends can be partially explained by underlying relationships between motif participation and local-scale network properties.
+    For example, species with higher trophic levels tended to participate less often in the direct competition motif.
+    A high trophic level is always associated with lower persistence since predators necessarily have lower probabilities of persistence than their prey~\citep{Eklof2013}
+    Accordingly, high participation in the direct competition motif, by indicating a low trophic level, is associated with high probability of persistence.
 
@@
     This is consistent with a strong negative correlation between omnivory and trophic level (leading to the overall positive effect of omnivory) and a strong positive correlation with in-degree (leading to the reduced benefit of omnivory at high levels of disturbance).
-Motifs can, in this sense, act as a tool for synthesizing the information provided by multiple other measures of a species' place in its community, as well as interactions between the measurements.
+    Motifs can thus act as a tool for synthesizing the information provided by, and interactions among, other measures of a species' place in its community.
 
     % Degree and TL don't explain everything
-Also the relationships between persistence and participation in the apparent competition and three-species chain motifs demonstrate motif participation roles' ability to syntesize different aspects of network structure.
+    The relationships between persistence and participation in the apparent competition and three-species chain motifs likewise demonstrate the potential for motif participation to synthesize different aspects of network structure.
     In these cases, in-degree and trophic level are not sufficient to explain the trends we observe in the motifs.
     Participation in the apparent competition motif was strongly negatively correlated with in-degree and shows a similar relationship to persistence.
-However, participation in the apparent competition motif was also strongly and positively correlated with trophic level (which is negatively associated with persistence since predators by necessity have lower probability of persistence than their prey~\citep{Eklof2013}) and this correlation is not reflected in the relationship between participation in the apparent competition motif and persistence.
-Similarly, participation in the three-species chain motif was significantly correlated with both in-degree and trophic level, but the relationship between the motif and persistence did not reflect the relationship between in-degree and persistence.
-This suggests the influence of some other aspect of network structure on species' persistence.
-As motifs capture information about direct and indirect interactions and include a larger section of the network than degree or trophic level~\citep{Cirtwill2018FoodWebs}, it is likely that effects of these indirect interactions may be included in the relationships we observe between motifs and persistence.
+    However, participation in the apparent competition motif was also strongly and positively correlated with trophic level, and the negative effect of high trophic levels on persistence is not reflected in the relationship between persistence and participation in the apparent competition motif.
+
+    Similarly, participation in the three-species chain motif was significantly correlated with both in-degree and trophic level, but the relationship between motif participation and persistence did not reflect the disturbance-dependent relationship between in-degree and persistence.
+    This suggests that the relationship between persistence and participation in the chain motif synthesizes the effects of in-degree, trophic level, and some additional effect which counterbalances the effect of in-degree.
+    This additional effect may come from indirect interactions, which are described by motifs but not in-degree~\citep{Cirtwill2018FoodWebs}. 
+    Although we did not include species traits in our simulations except when initially creating the food webs, in empirical webs motif participation can also capture information about species traits~\citep{cirtwill2018feeding} which could also affect extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000}.
 
@@
     In fact, a network's motif profile had a stronger relationship to mean persistence compared to network size or connectance. 
-However, connectance naturally did affect which network motifs were most common. 
-For example, highly-connected networks tended to contain higher frequencies of the omnivory motif which includes more links than the other three-species motifs considered here.
-Thus, global-scale properties seems to have stronger effects on network persistence by changing meso-scale properties than directly.
-This in turn suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.
-In general, we found that average network persistence was highest in networks with low frequencies of the omnivory motif and high frequencies of the apparent competition motif.
-These two motifs that have been shown to have consequences for the stability of ecological networks \citep{Borrelli2015a}.
+    This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.
+    Connectance did, however, affect which motifs were most common within a network. 
+    Highly-connected networks tended to contain higher frequencies of the omnivory motif, which includes more links than the other three-species motifs considered here.
+    Global-scale properties may thus  have stronger effects on network persistence by changing meso-scale properties than directly.
 
-% Heading into omnivory
-The effect of omnivory on stability has long been a subject of debate \citep{Kratina2012}. 
-Omnivory has been shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, Monteiro2016}. 
-For example, omnivorous interactions may not be stable in isolation but instead stabilized by interactions with other species when embedded in a larger food web \citep{Kratina2012}. 
-This might be why we here find a divergent pattern between a full food web's motif profile (where a high proportion of omnivory indicates lower average network persistence) versus each single species motif profile (where a high proportion of the omnivory motif indicates a higher focal species persistence).
-Thus, the ideal situation for a species could potentially be a highly omnivorous species in a low-omnivory setting.
+    In general, average persistence was highest in networks with low frequencies of the omnivory motif and high frequencies of the apparent competition motif. 
+    These two motifs that have been shown to have consequences for the stability of ecological networks \citep{Borrelli2015a,XX}.
+    Omnivory has been previously shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, Monteiro2016}. 
+    Omnivorous interactions may not be stable in isolation but instead stabilized by interactions with other species when embedded in a larger food web \citep{Kratina2012}. 
+    This might be why we find different patterns between a network's motif profile (where a high proportion of omnivory indicates lower average network persistence) versus a single species motif participation (where a high proportion of the omnivory motif indicates a higher focal species persistence).
+    Thus, the ideal situation appears to be being a highly omnivorous species in a low-omnivory setting.
 
     % High C webs and omnivory
-Additionally, we show that both the presence as well as the persistence of omnivorous species are higher in high-connected networks.
-The many pathways present in a high-connectance network, by which a bottom-up disturbance can propagate through the network, might explain the positive effect on persistence for species that feeds on multiple trophic levels.
-%Species that participate in many omnivory motifs have feeding interactions to species at lower trophic levels in addition to the directly adjacent one.
-Those species will therefore be less-severely affected by these propagating disturbances enforced in high-connectance networks, in comparison to other species.
-In a sense, a species with a high proportion of omnivory will spread its risk between prey items at different trophic levels.
-Although species in our model cannot swap prey dependent on bottom-level disturbances or lower resource densities, this mechanism resembles empirical studies showing that adaptive omnivory can increase the permanence of three-species chains~\citep{Fagan1997, Kvrivan2005, AbramsFung2010}.
-
-An additional important predictor for a species persistence is its traits \citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000}.
-In turn, a species' set of traits also influence the likelihood of it participating in different motifs \citep{cirtwill2018feeding}.
-This association between motifs and traits further emphasise the role of motifs for understanding how species' roles in a network have implications for its persistence.
+    Moreover, the positive relationship between participation in omnivory and species persistence is stronger in highly-connected networks.
+    These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}.
+    In such a network, it may be especially beneficial for a species to feed on multiple trophic levels and, in a sense, spread its risk of losing prey to bottom-up effects.
+    Although species in our model cannot swap prey, the reduced extinction risk for omnivorous species resembles empirical studies showing that adaptive omnivory can increase the permanence of three-species chains~\citep{Fagan1997, Kvrivan2005, AbramsFung2010}.
 
     There are some limitations to the approach presented here.
-Most important is the lack of top-down effects in the extinction risk calculations.
-This is because Bayesian networks operate on a strict bottom-up principle whereby prey influence their predators, but not vice versa \citep{Eklof2013}.
-Here our focus was to analyze the importance of disturbances to the primary producers and how effects propagate up in the food web.
-Earlier work indeed showed that the Bayesian network approach capture a majority of the secondary extinctions when compared to a fully-fledged dynamical model \citep{Eklof2013}, but if top-down effects are assumed important other methods should be used.
+    Most importantly, the Bayesian Network framework operates on a  strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
+    Here our focus was to analyze the importance of disturbances to the primary producers, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
+    If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
+    % other methods should be used. 
     Ongoing work using a dynamical model show that time to extinction for species following disturbance is tightly coupled to the species motif profile \citep{Cirtwill2021_inprep}.
@@
     In summary, our results show that participating in different three-species motifs is related to a species' risk of extinction following a bottom-up disturbance, and that these relationships synthesize information provided by other, simpler measures of species roles.
-Our results also underscore the importance of the level of disturbance when identifying which species are most at risk of extinction.
-There were strong interactions between the level of disturbance and the relationships between persistence and measures of species roles (motifs and in-degree).
+    Our results also underscore the importance of considering the level of disturbance when identifying which species are most at risk of extinction.
+    There were strong interactions between the level of disturbance and the relationships between persistence and multiple measures of species roles (motifs and in-degree).
     These simulated results echo empirical studies which report `tipping points' where previously beneficial traits can become risky~\citep{}.

### 2021-10-06 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{3}$, Kate Wootton$^{2}$, Gyuri Barab\'{a}s$^{1}$, Anna Ekl\"{o}f$^{1}$} 
+\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{2}$, Kate Wootton$^{3}$, Anna Ekl\"{o}f$^{1}$} 
 \date{\small$^1$Department of Theoretical Biology, Chemistry, and Physics\\ 
@@
 \medskip
-\small$^2$ Swedish Agricultural University\\
-Uppsala, Sweden\\
-\medskip
-\small$^3$Department of Agricultural Sciences\\
+\small$^2$Department of Agricultural Sciences\\
 University of Helsinki\\
 Helsinki, Finland\\
+\medskip
+\small$^3$ Swedish Agricultural University\\
+Uppsala, Sweden\\
+Colorado?
 \medskip
@@
     One way to incorporate more nuance into descriptions of how species fit into their communities is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
-Importantly, these `motif participation roles' capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015JAE}.
+Importantly, these `motif participation' roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015JAE}.
     This means that two species with the same in-degrees and trophic levels can still participate in different sets of motifs if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
     This extra detail may be useful when predicting  extinction risk since, at the network level, the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities~\cite{giling2019plant}.
-    We therefore expect that the motif structure of a network and species' participation in that structure should be related to risks due to disturbance to basal resources.
+    We therefore expect that both the motif structure of a network and species' participation in that structure should be related to risks due to disturbance to basal resources.
 
@@
 		For full details, see~\citet{Cirtwill2021_inprep}.
-        We then filtered these simulated networks and removed any network containing disconnected components
-        % (species or groups of species not connected to the rest of the network) 
-        or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
-		New networks were simulated to replace any removed networks, and this process was repeated until we obtained 100 suitable networks in each combination of size and connectance.
+		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
 		All networks were rendered acyclic following~\citet{Allesina2009} and topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).
 		These processing steps allow the strictly bottom-up calculation of probability of persistence used in the Bayesian network framework.
-
-% 		A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}.
-% 		All persistence calculations therefore begin  by determining the status of the primary producers (who do not depend on other species).
-% 		Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on primary producers), and so on up the network.
-
-% 		While niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
-% 		We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009}.
-% 		%This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009}.
-% 		%These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
-% 		Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citep{Tarjan1972, Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
-%         After these steps, calculations of persistence probabilities can begin.
 
@@
 		%Consider a species with $n$ resource species. 
-		Before taking food-web structure into account, all species already have a baseline probability of extinction, $\pi_{base}$. 
-		$\pi_{base}$ represents the risk of going extinct due to diseases, stochastic extinction of small populations, or other factors not related to the food web.
-		$\pi_{base}$ may be constant, or it can vary between species to reflect greater sensitivity in some groups of species.
+		Before taking food-web structure into account, all species already have a baseline probability of extinction, $\pi$. 
+		$\pi$ represents the risk of going extinct due to diseases, stochastic extinction of small populations, or other factors not related to the food web.
+		A species probability of extinction also depends on the state of its resources. 
+        If the focal species is a consumer with $n$ resources, of which $k$ have gone extinct, its probability of extinction depends on $\pi$ and on the fraction $f = k/n$.
 
-        A species probability of extinction also depends on the state of its resources.
-If the focal species is a consumer with $n$ resources, of which $k$ have gone extinct, its probability of extinction depends on $\pi_{base}$ and on the fraction $f = k/n$.
-
-        The response of the consumer to changes in $f$ can take any form. Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$; Appendix \emph{S2}.
-
-		This type of response most accurately captures the secondary extinctions produced by dynamical models~\citep{Eklof2013}.
-
+        The response of the consumer to changes in $f$ can take any form. Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$) 
+		This type of response most accurately captures the secondary extinctions produced by dynamical models (\citealp[]{Eklof2013}; Appendix \emph{S2}).
 		The probability of extinction for each species $i$ can therefore be represented using the cumulative density function of a beta distribution:
 		\begin{equation}
+		P(\lnot i|f) = \pi_{base} + (1 - \pi_{base}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}.
 				\label{betafunc}
-		P(\lnot i|f) = \pi_{base} + (1 - \pi_{base}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}.
         \end{equation}
@@
 This will in turn affect the extinction probability $P(\lnot C|f)$.
-		For basal resources, we assume that abiotic resources are always available and $f=0$. Following Equation \ref{betafunc}, the extinction will only depend on $\pi_{base}$.
-
+		For basal resources, we assume that abiotic resources are always available and $f=0$. Following Equation \ref{betafunc}, the extinction will only depend on $\pi$.
 		Note that we here assume that only the fraction, and not the identity, of the resources lost is important. 
@@
 
-		Modelling species extinctions in food webs using Bayesian networks have two main limitations.
-First, due to the strict bottom-up approach the food webs must be acyclic.
-Second, as consumers are influenced by the state of their resources, but resources are not affected by their consumers, the method does not capture top-down effects.
+		Modelling species extinctions in food webs using Bayesian networks have two main limitations. First, due to the strict bottom-up approach the food webs must be acyclic. 
+		Second, as consumers are influenced by the state of their resources, but not vice versa, the method does not capture top-down effects.
 
     % various disturbance scenarios
-        In our baseline scenario, primary producers had an extinction probability of $\pi_{base} = 0.1$. 
-		To simulate various threats to primary producers, making some species more vulnerable to extinction, we increase the affected species' probability of extinction ($\pi_{disturbed}$). 
-		We varied $\pi_{disturbed}$ between $0.18-0.5$, in steps of $0.08$. 
-		The highest disturbance level, $\pi_{disturbed} = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
-		Consumer species retained $\pi_{base}=0.1$ in all cases.
+        In our baseline scenario, basal resources had an extinction probability of $\pi = 0.1$. 
+		To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase the affected species' probability of extinction ($\pi$). 
+		We varied $\pi$ between $0.18-0.5$, in steps of $0.08$. 
+		The highest disturbance level, $\pi = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
+		Consumer species retained $\pi=0.1$ in all cases.
 
@@
         These were in-degree (number of prey) and trophic level (shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}).
-        Both simple role measures were calculated in R~\citep{R}.
+        Both simple role measures were calculated in R~\citep{R} using the same algorithm as~\citet{Eklof2013}.
 
@@
 
-	We tested:
-	\noindent1) how species persistence varied with motif participation.
-
-	\noindent2) how motif participation varied with other network properties,
-
-	\noindent3) how network mean persistence varied with network motif profiles,
-
-	\noindent4) how network motif profiles varied with global structural properties, and
-
-	\noindent5) how species persistence and network mean persistence varied with other network properties (see Appendix \emph{3.1 - 3.5}). 
-	All analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan} as detailed in the appendices.
+	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties, 3) how network mean persistence varied with network motif profiles, 4) how network motif profiles varied with global structural properties, and 5) how species persistence and network mean persistence varied with other network properties.
+	See Appendix \emph{3.1 - 3.5} for details. 
+	All analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.
 
@@
             \begin{equation}
-                \Psi_{ik} \approx \rho_{i} + \pi_{disturbed_k} + \rho_{i}\pi_{disturbed_k} +
+                \Psi_{ik} \approx \rho_{i} + \pi_{k} + \rho_{i}\pi_{k} +
                 S_{i}C_{i} ,
@@
 
-    % % Framing: motifs are related to network stability, species persistence (our results)
-    % Global network properties are often too coarse grained and identification of vulnerable species in that context becomes to broad (e.g.
-primary producers).
-Motifs represent a middle-ground approach as they are the basic building blocks of networks~\citep{Milo2002}, connecting network-wide (global) and species-specific (local) properties, describing both direct and indirect effects.
-
     % Species-level results summary
@@
     This additional effect may come from indirect interactions, which are described by motifs but not in-degree~\citep{Cirtwill2018FoodWebs}.
-    Although we did not include species traits in our simulations except when initially creating the food webs, in empirical webs motif participation can also capture information about species traits~\citep{cirtwill2018feeding} which could also affect extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000}.
 
@@
     % High C webs and omnivory
-    Moreover, the positive relationship between participation in omnivory and species persistence is stronger in highly-connected networks.
+    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks \emph{and} the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}).
     These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}.
@@
 
-    There are some limitations to the approach presented here.
+    There are limitations to the approach presented here.
     Most importantly, the Bayesian Network framework operates on a  strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
-    Here our focus was to analyze the importance of disturbances to the primary producers, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
+    Here our focus was to analyze the importance of disturbances to the basal resources, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
     If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
-    % other methods should be used. 
-    Ongoing work using a dynamical model show that time to extinction for species following disturbance is tightly coupled to the species motif profile \citep{Cirtwill2021_inprep}.
+    Ongoing work using a dynamical model shows that time to extinction for species following disturbance is tightly coupled to the motifs a species participates in, and that adding top-down effects can lead omnivory to be negatively correlated with persistence \citep{Cirtwill2021_inprep}.
+    If top-down effects are not crucial, the Bayesian Network approach could be extended by explicitly incorporating information on species traits.
+    In empirical webs, species traits are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000},
+    adding an extra dimension of information which we do not consider here.
 

### 2021-10-06 manuscript/supplement_motifs.tex
any last changes
Updates from Overleaf

@@
 
-\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{2}$, Kate Wootton$^{3}$, Gyuri Barab\'{a}s$^{1}$, Anna Ekl\"{o}f$^{1}$} 
+\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{2}$, Kate Wootton$^{3}$, Anna Ekl\"{o}f$^{1}$} 
 \date{\small$^1$Department of Theoretical Biology, Chemistry, and Physics\\ 
@@
 
+    After simulating our initial niche-model networks, we then filtered networks for biological plausibility.
+    Specifically, we removed any network containing disconnected components
+    (species or groups of species not connected to the rest of the network) 
+    or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
+    Any removed network was replaced with a fresh simulation and re-checked until 100 realistic networks for each size-connectance combination had been obtained.
+
     A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}.
-    All persistence calculations therefore begin  by determining the status of the primary producers (who do not depend on other species).
-    Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on primary producers), and so on up the network.
+    All persistence calculations therefore begin  by determining the status of the basal resources (who do not depend on other species).
+    Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on basal resources), and so on up the network.
 
@@
         of this consumer will depend on both the total state of its resource species $n$, and on a
-        baseline extinction probability of extinction, $\pi_{base}$. 
-        $\pi_{base}$ represents the risk of going extinct
+        baseline extinction probability of extinction, $\pi$. 
+        $\pi$ represents the risk of going extinct
         given factors not related to the food web itself, and could reflect diseases or stochastic
@@
         \label{betafunc}
-        P(\lnot i|f) = \pi_{base} + (1 - \pi_{base}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}
+        P(\lnot i|f) = \pi + (1 - \pi) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}
         \end{equation}
 
-        If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability will be $P(\lnot C|f) = \pi_{base}$. 
+        If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability will be $P(\lnot C|f) = \pi$. 
         Similarly, if all resources are extinct ($f = 1$), the extinction probability $P(\lnot C|f)$ will be equal to 1 and the consumer will go extinct.
@@
         Systematically, we use eqn \ref{betafunc} to receive each species probability of extinction $P(\lnot i|f)$, and next we perform Bernoulli trials to determine whether the species goes extinct or not.
-        Imagine a basal species \textit{A} with extinction probability $P(\lnot A) = \pi_{base}$ where $\pi_{base} = 0.1$.
+        Imagine a basal species \textit{A} with extinction probability $P(\lnot A) = \pi$ where $\pi = 0.1$.
 We draw a random number $r$ from a uniform distribution between $0-1$.
 The basal species would be considered extinct if $r < 0.1$, and extant otherwise.
-        We use this information when performing a Bernoulli trial on a consumer \textit{C} connected to this basal species, inserting our information of the $\pi_{base}$ and fraction of resources lost $f$ of this consumer, into eqn \ref{betafunc} to
+        We use this information when performing a Bernoulli trial on a consumer \textit{C} connected to this basal species, inserting our information of the $\pi$ and fraction of resources lost $f$ of this consumer, into eqn \ref{betafunc} to
         receive $P(\lnot C|f)$.
@@
     % various disturbance scenarios
-        In our baseline scenario, primary producers had a baseline extinction probability of $\pi_{base} = 0.1$. 
-        We also simulated a set of scenarios where primary producers had a higher baseline probability of extinction ($\pi_{disturbed}$). 
-        These scenarios reflect threats to primary producers such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
-        $\pi_{disturbed}$ ranged between $0.1-0.5$, in steps of $0.08$. 
-        The highest disturbance level, $\pi_{disturbed} = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
-        Consumer species retained $\pi_{base}=0.1$ in all cases.
+        In our baseline scenario, basal resources had a baseline extinction probability of $\pi = 0.1$. 
+        We also simulated a set of scenarios where basal resources had a higher baseline probability of extinction ($\pi$). 
+        These scenarios reflect threats to basal resources such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
+        $\pi$ ranged between $0.1-0.5$, in steps of $0.08$. 
+        The highest disturbance level, $\pi = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
+        Consumer species retained $\pi=0.1$ in all cases.
 
@@
         \begin{equation}
-            \Psi_{ik} \approx \rho_{i} + \pi_{disturbed_k} + \rho_{i}\pi_{disturbed_k} +
+            \Psi_{ik} \approx \rho_{i} + \pi_{k} + \rho_{i}\pi_{k} +
             S_{i}:C_{i} ,
@@
 
-        \noindent where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
+        \noindent where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
         We centered and scaled the predictors before fitting the LMMs.
@@
             \begin{equation}
-                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{disturbed_k} + \Bar{\rho}_{i}\pi_{disturbed_k} ,
+                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{k} + \Bar{\rho}_{i}\pi_{k} ,
                 \label{netpropeq}
@@
             Both species richness (S) and connectance (C) are known to influence persistence in both simulation studies and empirical food webs.
-            In addition, increasing the probability of extinction for basal species ($\pi_{disturbed}$ above) is intended to decrease persistence of species at higher trophic levels.
+            In addition, increasing the probability of extinction for basal species ($\pi$ above) is intended to decrease persistence of species at higher trophic levels.
             Although these effects are not the main focus of our manuscript, we must establish these background effects before measuring the effects of meso-scale structure on persistence.
-            The persistence of basal species in our simulations is determined only by their baseline probability, being either $\pi_{base} = 0.1$ or higher, dependent on $\pi_{disturbed}$.
+            The persistence of basal species in our simulations is determined only by their baseline probability, being either $\pi = 0.1$ or higher, dependent on $\pi$.
             We therefore considered only non-basal species (i.e., those with at least one prey) throughout this manuscript.
@@
             \begin{equation}
-                \Psi_{ik} \approx \Sigma_{i} + \zeta_{i} + \pi_{disturbed_k} + \Sigma_{i}:\zeta_{i} + \Sigma_{i}\pi_{disturbed_k} + \zeta_{i}\pi_{disturbed_k} + \Sigma_{i}\zeta_{i}\pi_{disturbed_k},
+                \Psi_{ik} \approx \Sigma_{i} + \zeta_{i} + \pi_{k} + \Sigma_{i}:\zeta_{i} + \Sigma_{i}\pi_{k} + \zeta_{i}\pi_{k} + \Sigma_{i}\zeta_{i}\pi_{k},
                 \label{SCeq}
@@
             \begin{equation}
-                    \Psi_{ik} \approx \delta_{i} + \pi_{disturbed_k} + \delta_{i}\pi_{disturbed_k} +
+                    \Psi_{ik} \approx \delta_{i} + \pi_{k} + \delta_{i}\pi_{k} +
                     S_{i}C_{i} ,
@@
             \begin{equation}
-                    \Psi_{ik} \approx \tau_{i} + \pi_{disturbed_k} + \tau_{i}\pi_{disturbed_k} +
+                    \Psi_{ik} \approx \tau_{i} + \pi_{k} + \tau_{i}\pi_{k} +
                     S_{i}C_{i} ,
@@
      \centering
-     % \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}
+     \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}
      \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
@@
      \centering
-     % \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
-     \caption{Primary producers experience a disturbance reducing their probability to go extinct to $\pi_{disturbed} = 0.5$ All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
+     \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
+     \caption{Basal resources experience a disturbance reducing their probability to go extinct to $\pi = 0.5$ All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
 The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
@@
     When comparing the outcomes from linear mixed-effect models (LMMs) relating persistence to motif roles across all networks (Fig.
-3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately, the main trends are similar but different-sized networks showed slightly different results results for some motifs (Figure temporarily in main text).%(Figure~\ref{fig:density_prop_S}).
+3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately, the main trends are similar but different-sized networks showed slightly different results results for some motifs (Figure~\ref{fig:density_prop_S}).
     For the three-species chain and omnivory motifs, a significant interaction between disturbance and the effect of the focal motif means that having a larger proportion of the role be made up of the chain or omnivory motif is associated with lower persistence at high levels of disturbance (Fig.
 3 and Table 1, \emph{Main Text}).
-    The same pattern holds overall for LRs fit within a network  %(Figure~\ref{fig:density_prop_S})
+    The same pattern holds overall for LRs fit within a network  (Figure~\ref{fig:density_prop_S})
     : the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance. 
-    This trend is most pronounced in large networks.% (bottom panel, Figure~\ref{fig:density_prop_S}). 
+    This trend is most pronounced in large networks. (bottom panel, Figure~\ref{fig:density_prop_S}). 
     Besides the decrease in fraction of positive slopes with increasing disturbance, the omnivory motif show a higher overall fraction of positive slopes in comparison to the three-species chain motif, corresponding to the positive main effect of omnivory (Fig.
@@
     \centering
-         % \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
-        \caption{Density (y-axis) of slopes (x-axis) for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
-The slopes are derived from linear mixed-effect models, with a random intercept for the interaction between species richness and network size.
+        \includegraphics[width=0.75\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
+        \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all connectances - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
+Columns show the result for various disturbances on the basal level, from $\pi = 0.1$ (left) to $\pi = 0.5$ (right).
+Rows show various network sizes.
+The dotted, vertical line indicate zero on the x-axis.
 A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
-The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to motif (legend).
-Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
-Rows show various levels of network size.
-The vertical line indicate zero on the x-axis.}
+The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend).
+}
         \label{fig:density_prop_S}
@@
 
-\section{SA: Persistence and global network properties}
+\section{Results: Persistence and global network properties}
 
@@
 %             \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while the proportion of the network made up of basal resources, network size, connectance, and their interactions had smaller effects.
-We show relationships with persistence for low disturbance ($\pi_{Basal}$=0.1) and high disturbance($\pi_{Basal}$=0.5).
+We show relationships with persistence for low disturbance ($\pi$=0.1) and high disturbance($\pi$=0.5).
 For each predictor other than disturbance, we show the relationship between the focal predictor and persistence for combinations of the lowest and highest observed values of the other predictors.

### 2021-10-06 code/plot_roles_vs_TL.py
flipped fig

@@ -212,18 +212,18 @@ lmdict=read_lmfile(lmfile)
 persdict=read_persfiles('stat_analysis/persistence_vs_Deg.tsv','stat_analysis/persistence_vs_TL.tsv')
 
 grace=MultiPanelGrace(colors=colors)
-for graphtype in ['Persistence','Proportion']:
+for graphtype in ['Proportion','Persistence']:
   for simple in ['Deg','TL']:
     if graphtype=='Proportion':
       graph=grace.add_graph(Panel)
       graph=format_graph(graph,simple,'Prop')
       graph=populate_graph(graph,lmdict[simple]['Prop'],simple,'Prop')
-      graph.panel_label.configure(placement='iul',char_size=1,dx=.03,dy=.03)
+      graph.panel_label.configure(placement='iul',char_size=1,dx=.02,dy=.02)
     elif graphtype=='Persistence':
       graph=grace.add_graph(Panel)
       graph=format_graph(graph,simple,'persistence')
       graph=populate_persgraph(graph,persdict[simple],simple)
-      graph.panel_label.configure(placement='iul',char_size=1,dx=.03,dy=.03)
+      graph.panel_label.configure(placement='iul',char_size=1,dx=.02,dy=.02)
       
     
 grace.multi(rows=2,cols=2,vgap=.05,hgap=.05)

### 2021-10-06 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
     Motif participation therefore provides a relatively rich description of a species' structural role in its community and captures important trends in vulnerability to extinction.
-
-    % Anna AA: The whole article is now approximately 4400 words (not counting fig texts); intro 850, methods 1600, results 620, discussion 1300.
-See discussion on section statistical analysis in the methods, I think we can make it shorter
-    % - glossary
-    % - refs
-    % - ending
-
-    % - Fig. 4 not much in discussion, could be a candidate for the SI? Or add more discussion about apparent competition, etc. in discussion sections per motif
 
@@
     % basic idea behind BN modeling
-        Bayesian Networks can be used to represent species and their feeding interactions as the nodes and links of a network (see Appendix \emph{S2} for full details).
+        Bayesian networks can be used to represent species and their feeding interactions as the nodes and links of a network (see Appendix \emph{S2} for full details).
         In our setting, each species (node) is a Bernoulli random variable, being either extant or extinct. 
@@
     There are limitations to the approach presented here.
-    Most importantly, the Bayesian Network framework operates on a  strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
+    Most importantly, the Bayesian network framework operates on a  strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
     Here our focus was to analyze the importance of disturbances to the basal resources, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
@@
     Ongoing work using a dynamical model shows that time to extinction for species following disturbance is tightly coupled to the motifs a species participates in, and that adding top-down effects can lead omnivory to be negatively correlated with persistence \citep{Cirtwill2021_inprep}.
-    If top-down effects are not crucial, the Bayesian Network approach could be extended by explicitly incorporating information on species traits.
+    If top-down effects are not crucial, the Bayesian network approach could be extended by explicitly incorporating information on species traits.
     In empirical webs, species traits are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000},
@@
         \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
-        \caption{As with motifs, persistence varied with simpler measures of species roles in networks.
-Moreover, motif frequencies correlated with in-degree and trophic level.
-\textbf{A-B)} Persistence increased with increasing in-degree for low levels of disturbance to basal resources but decreased with increasing degree for high levels of disturbance to basal resources.
-        Persistence declined with inceasing STL for all levels of disturbance to basal resources.
-        \textbf{C-D)} The proportion of omnivory increased with increasing degree while all other proportions decreased.
-The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.}
+        \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence.
+\textbf{A-B)} The proportion of omnivory increased with increasing degree while all other proportions decreased.
+The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.
+\textbf{C-D)} Persistence increased with increasing in-degree for low levels of disturbance to basal resources but decreased with increasing degree for high levels of disturbance to basal resources.
+        Persistence declined with inceasing STL for all levels of disturbance to basal resources.}
         \label{fig:motifs_vs_TL_and_deg}

### 2021-10-06 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 
-        We use Bayesian Network modelling to calculate the persistence of each species in our simulated networks. 
-        A Bayesian Network represents a graphical model for probabilistic relationships among a set of variables.
+        We use Bayesian network modelling to calculate the persistence of each species in our simulated networks. 
+        A Bayesian network represents a graphical model for probabilistic relationships among a set of variables.
 The nodes of the network represents the variables, and the links represent conditional dependencies among the variables.

### 2021-10-08 manuscript/cover_letter.tex
Updates from Overleaf

\documentclass[11pt]{letter}

\usepackage[bitstream-charter]{mathdesign}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage[letterpaper, margin=0.9in, top=0.4in, bottom=0.4in]{geometry}

\longindentation=30pt

\address{Anna {\AA}kesson \\ Link\"oping University \\ Division of Theoretical Biology, Dept.~IFM \\ SE-58183 Link\"oping, Sweden}

\date{October XX, 2021}

\begin{document}

\begin{letter}{
\vspace{-2.5cm}

Dr.~XX\\
Editor-in-Chief\\
Proceedings of the Royal Society B} % Name/title of the addressee

\opening{Dear Dr.~XX,}

Please find enclosed the manuscript ``Does motif participation provide unique information about species risk of extinction?'' by Anna {\AA}kesson et al., which we would like to be considered for publication in \emph{Proceedings of the Royal Society B} as an Article.

Alyssa Cirtwill has earlier been approached by the Proceedings B Preprint Editorial Team, considering another paper regarding motifs.

While there is pressing need to understand ...
Here we study the effects of these processes ... 
Several key results emerge ...

First ...

Second ...

Third ...

We reach our conclusions by ...

Due to the relevance of our results to understanding ..., for which \emph{ Proceedings of the Royal Society B} would be an ideal platform.

We thank you for your time and attention dedicated to our work.

\vspace{2\parskip}

\hspace{4\parskip} Sincerely,

\hspace{4\parskip} Anna {\AA}kesson, Alyssa R. Cirtwill, Kate Wootton, and Anna Ekl\"of

\end{letter} 
\end{document}

### 2021-10-08 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
     One way to incorporate more nuance into descriptions of how species fit into their communities is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
-Importantly, these `motif participation' roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015JAE}.
+Importantly, these `motif participation' roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a}.
     This means that two species with the same in-degrees and trophic levels can still participate in different sets of motifs if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
-    This extra detail may be useful when predicting  extinction risk since, at the network level, the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities~\cite{giling2019plant}.
+    This extra detail may be useful when predicting  extinction risk since, at the network level, the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
     We therefore expect that both the motif structure of a network and species' participation in that structure should be related to risks due to disturbance to basal resources.
@@
 		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
-		All networks were rendered acyclic following~\citet{Allesina2009} and topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).
+		All networks were rendered acyclic following~\citet{Allesina2009functional} and topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).
 		These processing steps allow the strictly bottom-up calculation of probability of persistence used in the Bayesian network framework.
@@
 
-        The response of the consumer to changes in $f$ can take any form. Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$) 
-		This type of response most accurately captures the secondary extinctions produced by dynamical models (\citealp[]{Eklof2013}; Appendix \emph{S2}).
+        The response of the consumer to changes in $f$ can take any form (Appendix \emph{S2}). Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$).
+		This type of response most accurately captures the secondary extinctions produced by dynamical models (\citealp[]{Eklof2013}).
 		The probability of extinction for each species $i$ can therefore be represented using the cumulative density function of a beta distribution:
 		\begin{equation}
-		P(\lnot i|f) = \pi_{base} + (1 - \pi_{base}) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}.
+		P(\lnot i|f) = \pi + (1 - \pi) \frac{B(f;\alpha,\beta)}{B(\alpha,\beta)}.
 				\label{betafunc}
@@
 
-		If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability will be $P(\lnot C|f) = \pi_{base}$. 
+		If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability will be $P(\lnot C|f) = \pi$. 
 		Similarly, if all resources are extinct ($f = 1$), the extinction probability $P(\lnot C|f)$ will be equal to 1 and the consumer will go extinct.
@@
             \end{equation}
-        \noindent where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{disturbed_k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
+        \noindent where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
         All predictors were centered and scaled before fitting the models.         
@@
         Different scales of network structure (e.g., local and meso-scale) are not independent. 
-        To identify correlations between network size, connectance, and participation in particular motifs, we fit a set of four linear models (LMs):
-        \begin{equation}
-            \rho_{i} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i} ,
-            \label{partic_SC}
-        \end{equation}
-        \noindent where $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif,
-        $\Sigma_{i}$ is the size of the network containing species $i$, and $\zeta_{i}$ is the connectance of the network containing species $i$.
-        We fit a similar set of eight LMMs (two per motif) relating motif participation to in-degree or trophic level:
-        \begin{equation}
-            \rho_{i} \approx \delta_{i} + S_{i}C_{i} , \hspace{2cm} 
-            \rho_{i} \approx \tau_{i} + S_{i}C_{i} ,
-            \label{partic_degTL}
-        \end{equation}
-        \noindent where $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, and all other symbols are as in equation~\ref{propreq}.
+        To identify correlations between global network structure and participation in particular motifs, we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
+        We fit a similar set of  LMMs relating motif participation to in-degree or trophic level (Appendix \emph{S3.2}).
+        Taken together, these supplemental models demonstrate how motif participation relates to other measures of network structure.
 
@@
         As well as species-specific probabilities of persistence, the overall frequency of motifs in a network may provide information about the overall probability of any species going extinct (since they incorporate information on indirect effects).
-        We tested whether networks with similar motif profiles tend to have similar mean persistences among consumers using a PERMANOVA (Appendix \emph{S3.3}.
+        We tested whether networks with similar motif profiles tend to have similar mean persistences among consumers using a PERMANOVA (Appendix \emph{S3.3}).
         To identify the relationships between particular motifs and network mean persistence in more detail, we fit a set of four LMs (one per motif):
             \begin{equation}
-                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{disturbed_k} + \Bar{\rho}_{i}\pi_{disturbed_k},
+                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{k} + \Bar{\rho}_{i}\pi_{k},
                 \label{netpropeq}
@@
             \Bar{\rho_{i}} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i}
+        \label{profile_global}
         \end{equation}
-        where all symbols are as in equations~\ref{propreq} and~\ref{netpropeq}.
+        \noindent where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and $\Sigma_i$ and $\zeta_{i}$ are the size and connectance of the network to which $i$ belongs.
 
@@
             \begin{equation}
-                \Psi_{ik} \approx \Sigma_{i} + \zeta_{i} + \pi_{disturbed_k} + \Sigma_{i}:\zeta_{i} + \Sigma_{i}\pi_{disturbed_k} + \zeta_{i}\pi_{disturbed_k} + \Sigma_{i}\zeta_{i}\pi_{disturbed_k},
+                \Psi_{ik} \approx \Sigma_{i} + \zeta_{i} + \pi_k + \Sigma_{i}:\zeta_{i} + \Sigma_{i}\pi_k + \zeta_{i}\pi_k + \Sigma_{i}\zeta_{i}\pi_k,
                 \label{SCeq}
             \end{equation}
-        \noindent where all symbols are as in equation~\ref{propreq} and~\ref{partic_SC}.
+        \noindent where all symbols are as in equation~\ref{propreq} and~\ref{profile_global}.
         As the three-way interaction was significant, we did not simplify the model.
@@
         \begin{equation}
-                \Psi_{ik} \approx \delta_{i} + \pi_{disturbed_k} + \delta_{i}\pi_{disturbed_k} +
+                \Psi_{ik} \approx \delta_{i} + \pi_k + \delta_{i}\pi_k +
                 S_{i}C_{i} ,
@@
         \begin{equation}
-                \Psi_{ik} \approx \tau_{i} + \pi_{disturbed_k} + \tau_{i}\pi_{disturbed_k} +
+                \Psi_{ik} \approx \tau_{i} + \pi_k + \tau_{i}\pi_k +
                 S_{i}C_{i} ,
@@
     Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks \emph{and} the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}).
-    These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}.
+    These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
     In such a network, it may be especially beneficial for a species to feed on multiple trophic levels and, in a sense, spread its risk of losing prey to bottom-up effects.

### 2021-10-08 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 \newcommand{\methods}{\textit{Materials \& Methods}}
-\newcommand{\SI}{\textit{Appendix}~}
+\setcounter{section}{0}
+\renewcommand{\thesection}{S\arabic{section}}
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}%
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}%
+\setcounter{equation}{0}
+\renewcommand{\theequation}{S\arabic{equation}}
 
@@
 
-% \section*{Table of contents}
-%     \begin{spacing}{1.0}
+\section*{Table of contents}
+    \begin{spacing}{1.0}
 
-%     \begin{enumerate}
+    \begin{enumerate}
 
-%         \item How do counts of motifs in a network profile vary with S and C? \\
+        \item Supplemental methods: Network creation \\
 
-%             \begin{itemize}
-%             \item Same methods as main text except using counts instead of proportions to define motif profiles
-%             \end{itemize}
+            \begin{itemize}
+            \item Construction of simulated networks
+            \item Filtering of biologically unrealistic networks 
+            \item Conversion of networks to acyclic form
+            \end{itemize}
 
-%         \item Motif profiles differed slightly between network sets
+        \item Supplemental methods: Calculating persistence
 
-%            \begin{itemize}
-%                 \item Comparing proportions of motifs in un-filtered, filtered, and acyclic networks.
-%                 \item AFAIK not referred to in text
-%             \end{itemize}
+            \begin{itemize}
+                \item Details on functional response of consumers
+                \item Contains Eqn. S1
+            \end{itemize}
 
-%          \item Persistence and count-based motif participation
+         \item Supplemental methods: Statistical analysis
 
-%            \begin{itemize}
-%                 \item Should be as main text section 4, but counts
-%             \end{itemize}
+            \begin{itemize}
+                \item Species persistence and motif participation (Eqn. S2-3)
+                \item Motif participation and network properties (Eqn. S4-6)
+                \item Network mean persistence and motif profiles (Eqn. S7)
+                \item Network motif profiles and global network structure (Eqn. S8)
+                \item Persistence and other network properties (Eqn. S9-11)
+            \end{itemize}
 
-%         \item Motif-persistence relationships and network size
-%             \begin{itemize}
-%                 \item Detailed results for variability of slopes for persistence $\approx$ motif participation in networks of different sizes.
-%                 \item Main text refers to C, moved figure to Main while we check that we've got everything.
-%             \end{itemize}
+        \item Results: Persistence and motif participation
+            \begin{itemize}
+                \item Contains Table S1
+            \end{itemize}
 
-%         \item How does persistence vary with degree and trophic level?
+        \item Results: Participation and network properties
 
-%             \begin{itemize}
-%                 \item Heat map showing persistence vs. TL with BP=0
-%                 \item Heat map showing persistence vs. TL with BP=1
-%             \end{itemize}
+            \begin{itemize}
+                \item Contains Table S2
+                \item Contains Fig. S1
+            \end{itemize}
 
-%     \end{enumerate}
-% \end{spacing}
-% \clearpage
+        \item Results: Persistence, degree, and trophic level
+
+            \begin{itemize}
+                \item Contains Table S3
+                \item Contains Fig. S2-3
+            \end{itemize}
+
+        \item Results: Motif-persistence relationships
+
+            \begin{itemize}
+                \item Contains Fig. S4
+            \end{itemize}
+
+        \item Results: Persistence and network motif profiles
+
+            \begin{itemize}
+                \item Contains Table S4
+            \end{itemize}    
+    \end{enumerate}    
+\end{spacing}
+\clearpage
 
@@
     We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009}.
-    This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009}.
+    This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009functional}.
     These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
-    Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citep{Tarjan1972, Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
+    Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citet{Tarjan1972} and \citet{Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
     After these steps, calculations of persistence probabilities can begin.
@@
         A Bayesian network represents a graphical model for probabilistic relationships among a set of variables. 
-The nodes of the network represents the variables, and the links represent conditional dependencies among the variables.
+        In our ecological setting, the nodes of the network represent species and links represent feeding interactions between species.
+        In network-theory terms, the species are Bernoulli random variables, and the links represent conditional dependencies among these variables. 
         Each node has a set of local probability dependencies, conditional on the state of the parent nodes. 
-        In our ecological setting, species act as nodes and are Bernoulli random variables, being either extant or extinct. The links act as feeding interactions between the nodes.
-        The extinction probability of each species is a function of the state of its resources.
-
-    \subsection{Baseline probability of extinction}
-
-        Consider a consumer species with several resource species $n$. 
-        The extinction probability
-        of this consumer will depend on both the total state of its resource species $n$, and on a
-        baseline extinction probability of extinction, $\pi$. 
-        $\pi$ represents the risk of going extinct
-        given factors not related to the food web itself, and could reflect diseases or stochastic
-        extinctions of small populations. 
-        All the species in the network retain a baseline
-        probability of extinction, which can be different to reflect greater sensitivity in some groups of species.
-
-    \subsection{Extinction due to food-web effects}
-
-        If a consumer with $n$ resources loses $k$ of them, the fraction of resources lost $f$ will be
-        $f = (k/n)$. 
-        Note that we here make a strong assumption, where we only care about the fraction of resources lost and not the identity of the resources species.
+        Since the nodes are species, their state can be either extant or extinct.
+        The extinction probability of each species is a function of the state of its resources and also depends on a baseline probability of extinction ($\pi$) representing risk of extinction due to factors unrelated to network structure.
+        In our baseline scenario, basal resources had a baseline extinction probability of $\pi = 0.1$. 
+        We also simulated a set of scenarios where basal resources had a higher baseline probability of extinction ($\pi$). 
+        These scenarios reflect threats to basal resources such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
+        $\pi$ ranged between $0.1-0.5$, in steps of $0.08$. 
+        The highest disturbance level, $\pi = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
+        Consumer species retained $\pi=0.1$ in all cases.
 
@@
 
-        The response of a consumer to the loss of a fraction of its resources can be modelled in several biologically plausible ways. For example, the response could be linear or non-linear.
+        The response of a consumer to the loss of a fraction of its resources can be modelled in several biologically plausible ways: i) a topological response, where the consumer's probability of extinction is constant unless all resources are lost, ii) a linear response, in which the extinction probability increases linearly with resources loss, and iii) several nonlinear responses where the probability of extinction increases non-linearly with the fraction resource lost.
+		The extinction probability will differ depending on the response function.
+
         A sigmoid non-linear functional response of consumers to the loss of resources ($\alpha >1, \beta >1$) has been shown to most accurately capture the secondary extinctions produced by a dynamical model~\citep{Eklof2013}.
@@
 
-        If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability will be $P(\lnot C|f) = \pi$. 
+        \noindent If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability will be $P(\lnot C|f) = \pi$. 
         Similarly, if all resources are extinct ($f = 1$), the extinction probability $P(\lnot C|f)$ will be equal to 1 and the consumer will go extinct.
@@
 
-        % how we calculate persistences 
-        As consumer species depends on its resource species, we start determining the status of each primary producer in the network.
-        Systematically, we use eqn \ref{betafunc} to receive each species probability of extinction $P(\lnot i|f)$, and next we perform Bernoulli trials to determine whether the species goes extinct or not.
+        As consumer species depend on their resources, we start by determining the status of each primary producer in the network.
+        Systematically, we use Equation \ref{betafunc} to calculate each species' probability of extinction $P(\lnot i|f)$, and then perform Bernoulli trials to determine whether the species goes extinct or not.
         Imagine a basal species \textit{A} with extinction probability $P(\lnot A) = \pi$ where $\pi = 0.1$.
@@
 The basal species would be considered extinct if $r < 0.1$, and extant otherwise.
-        We use this information when performing a Bernoulli trial on a consumer \textit{C} connected to this basal species, inserting our information of the $\pi$ and fraction of resources lost $f$ of this consumer, into eqn \ref{betafunc} to
-        receive $P(\lnot C|f)$.
+
+        We then insert the simulated fraction of species lost into Equation \ref{betafunc} to calculate $P(\lnot C|f)$ for the consumer species \textit{C}. 
         In the same manner as for the basal species above, we compare $P(\lnot C|f)$ with a random number $r$, and if $r < P(\lnot C|f)$ the consumer species is considered extinct. 
         In this way, we evaluate each species existence, starting from the bottom and working our way up.
+
         As this method is probabilistic, we repeat these calculations 100 000 times per network, with unique random draws.
@@
         To approximate persistence probabilities through numerical simulation are shown to be highly efficient and produce the same result as exact methods of solving Bayesian networks \citep{Haussler2020}.
-
-    % various disturbance scenarios
-        In our baseline scenario, basal resources had a baseline extinction probability of $\pi = 0.1$. 
-        We also simulated a set of scenarios where basal resources had a higher baseline probability of extinction ($\pi$). 
-        These scenarios reflect threats to basal resources such as increased temperatures resulting in droughts, making some species more vulnerable to extinction.
-        $\pi$ ranged between $0.1-0.5$, in steps of $0.08$. 
-        The highest disturbance level, $\pi = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
-        Consumer species retained $\pi=0.1$ in all cases.
 
@@
 
-        where all symbols are as in equation~\ref{propreq}.
+        \noindent where all symbols are as in equation~\ref{propreq}.
         These regressions were fit using the R~\citep{R} base function `lm'.
@@
 
-        where $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif,
+        \noindent where $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif,
         $\Sigma_{i}$ is the size of the network containing species $i$, and $\zeta_{i}$ is the connectance of the network containing species $i$.
@@
 
-        where $\rho_{i}$ is the frequency with which species $i$ participtes in the focal motif, $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, $S_{i}C_{i}$ is a random effect of the combination of size and connectance for the network containing species $i$.
+        \noindent where $\rho_{i}$ is the frequency with which species $i$ participtes in the focal motif, $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, $S_{i}C_{i}$ is a random effect of the combination of size and connectance for the network containing species $i$.
         These models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
 
-    \subsection{How does network mean persistence vary with motif profiles?}
+    \subsection{How does mean persistence vary with motif profiles?}
 
@@
 
-        where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and all other symbols are as in equation~\ref{propreq}. 
+        \noindent where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and all other symbols are as in equation~\ref{propreq}. 
         All regressions were fit using the R~\citep{R} base function `lm'.
@@
             Both species richness (S) and connectance (C) are known to influence persistence in both simulation studies and empirical food webs.
-            In addition, increasing the probability of extinction for basal species ($\pi$ above) is intended to decrease persistence of species at higher trophic levels.
+            In addition, increasing the probability of extinction for basal species ($\pi$) is intended to decrease persistence of species at higher trophic levels.
             Although these effects are not the main focus of our manuscript, we must establish these background effects before measuring the effects of meso-scale structure on persistence.
@@
 
-\section{Results: Motif participation and network properties}
+\section{Results: Participation and network properties}
 
@@
 
-    When comparing the outcomes from linear mixed-effect models (LMMs) relating persistence to motif roles across all networks (Fig.
-3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately, the main trends are similar but different-sized networks showed slightly different results results for some motifs (Figure~\ref{fig:density_prop_S}).
+    We compared the outcomes from linear mixed-effect models (LMMs) relating persistence to motif roles across all networks (Fig.
+3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately.
+    The main trends are similar in both analyses, but different-sized networks showed slightly different results results for some motifs when considered individually (Figure~\ref{fig:density_prop_S}).
     For the three-species chain and omnivory motifs, a significant interaction between disturbance and the effect of the focal motif means that having a larger proportion of the role be made up of the chain or omnivory motif is associated with lower persistence at high levels of disturbance (Fig.
@@
     This trend is most pronounced in large networks. (bottom panel, Figure~\ref{fig:density_prop_S}). 
-    Besides the decrease in fraction of positive slopes with increasing disturbance, the omnivory motif show a higher overall fraction of positive slopes in comparison to the three-species chain motif, corresponding to the positive main effect of omnivory (Fig.
+    % Besides the decrease in fraction of positive slopes with increasing disturbance, the omnivory motif showed a higher overall fraction of positive slopes in comparison to the three-species chain motif, corresponding to the positive main effect of omnivory (Fig.
 3 and Table 1, \emph{Main Text}).

### 2021-10-11 code/stat_analysis/persistence_vs_profiles.py
adding disturbance labels in fig

@@ -168,6 +168,9 @@ graph2=format_graph(graph2,'persistence')
 for mot in ['omnivory','chain','apparent','direct']:
   graph2=populate_persgraph(graph2,mot)
 
+graph2.add_drawing_object(DrawText,x=0.5,y=0.75,loctype='world',text='Lowest disturbance',just=0,char_size=1)
+graph2.add_drawing_object(DrawText,x=0.5,y=0.34,loctype='world',text='Highest disturbance',just=0,char_size=1)
+
 graph2.legend.configure(char_size=1,loc=(0.05,0.685),loctype='world',box_linestyle=0,fill_pattern=0)
 
 grace.write_file('../../manuscript/figures/persistence_motif_profiles.eps')

### 2021-10-11 manuscript/reorder_main_text.tex

@@
 
-\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{2}$, Kate Wootton$^{3}$, Anna Ekl\"{o}f$^{1}$} 
+\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{2}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{1}$} 
 \date{\small$^1$Department of Theoretical Biology, Chemistry, and Physics\\ 
@@
 \medskip
-\small$^3$ Swedish Agricultural University\\
-Uppsala, Sweden\\
-Colorado?
+\small$^3$ BioFrontiers Institute\\
+University of Colorado, Boulder\\
+U.S.A.
 \medskip
@@
     We show that a consumer species' frequency of participation in different motifs is related to its probability of persistence.
-Similarly, the overall motif profile of a network was related to the average persistence of consumer species.
+Similarly, the overall motif profile of a network was related to the average persistence of its consumer species.
     Notably, we found that networks containing a higher proportion of the `omnivory' motif tend to have lower mean persistence following disturbance to basal species, while more frequent participation in the omnivory motif was associated with higher probability of persistence.
@@
      %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
-    How population decline or extinction of one species leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    How population decline or extinction of one species leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
 Traditionally, there are two main approaches for studying secondary extinctions.
@@
 		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}.
-		To capture a range of plausible network architectures, we simulated networks with sizes ranging from 50 to 100 species (in steps of 10) and connectances ranging from 0.02 to 0.18 (in steps of 0.04).
+		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
 		For full details, see~\citet{Cirtwill2021_inprep}.
@@
 
-		Modelling species extinctions in food webs using Bayesian networks have two main limitations. First, due to the strict bottom-up approach the food webs must be acyclic. 
+		Modelling species extinctions in food webs using Bayesian networks has two main limitations. First, due to the strict bottom-up approach the food webs must be acyclic. 
 		Second, as consumers are influenced by the state of their resources, but not vice versa, the method does not capture top-down effects.
@@
 		To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase the affected species' probability of extinction ($\pi$). 
-		We varied $\pi$ between $0.18-0.5$, in steps of $0.08$. 
+		We increased $\pi$ for basal resources from $0.10$ to $0.5$, in steps of $0.08$. 
 		The highest disturbance level, $\pi = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
@@
 
-	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties, 3) how network mean persistence varied with network motif profiles, 4) how network motif profiles varied with global structural properties, and 5) how species persistence and network mean persistence varied with other network properties.
+	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties.
 	See Appendix \emph{3.1 - 3.5} for details. 
@@
 
-        To test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
+        To test whether networks with the the same size and connectance have more similar motif profiles, we used a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
         In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMs:
@@
         Persistence is also known to vary with local-scale measures of network structure. 
-        To identify the relationships betwen local-scale properties and species' probabilities of persistence, we fit a final two LMMs:
+        To identify the relationships betwen local-scale properties and species' probabilities of persistence, we fit two final LMMs:
         \begin{equation}
@@
        The relationship between frequency of participation in a motif and a species' probability of persistence depended on the level of disturbance. 
-       When disturbance was low, species' persistence increased with increasing proportions of the omnivory and direct competition motifs. 
+       When disturbance ($\pi$) was low, species' persistence increased with increasing proportions of the omnivory and direct competition motifs. 
        Conversely, persistence decreased with an increasing proportion of the apparent competition motif. 
@@
         connectance was significantly related to the frequencies of different motifs in a network when considering motifs individually (Fig.~\ref{motif_proportion_lms}).
-        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs .
+        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs.
         Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S5, \emph{Appendix S9}).
@@
     For example, species with higher trophic levels tended to participate less often in the direct competition motif.
-    A high trophic level is always associated with lower persistence since predators necessarily have lower probabilities of persistence than their prey~\citep{Eklof2013}
+    A high trophic level is always associated with lower persistence since predators necessarily have lower probabilities of persistence than their prey~\citep{Eklof2013}.
     Accordingly, high participation in the direct competition motif, by indicating a low trophic level, is associated with high probability of persistence.
@@
     In general, average persistence was highest in networks with low frequencies of the omnivory motif and high frequencies of the apparent competition motif. 
-    These two motifs that have been shown to have consequences for the stability of ecological networks \citep{Borrelli2015a,XX}.
+    These two motifs that have been shown to have consequences for the stability of ecological networks \citep{McLeod2019,Picot2018}.
     Omnivory has been previously shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, Monteiro2016}. 
@@
     There were strong interactions between the level of disturbance and the relationships between persistence and multiple measures of species roles (motifs and in-degree).
-    These simulated results echo empirical studies which report `tipping points' where previously beneficial traits can become risky~\citep{}.
+    These simulated results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
     These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.
+
+\section*{Acknowledgements}
+
+    The authors thank György Barabás for excellent discussion of the ideas in this manuscript. ARC is supported by a Finnish Academy postdoctoral research grant (\#332999).
 
@@
 
-    \begin{figure}[h!]
+    \begin{figure}[hb!]
         \centering
@@
 
-    \begin{figure}
+    \begin{figure}[hb!]
         \centering
@@
 
-    \begin{figure}[h!]
+    \begin{figure}[hb!]
     \centering
@@
 Rows show various levels of connectance.
-The dotted, vertical line indicate zero on the x-axis.
-A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
+The dotted, vertical lines indicate zero on the x-axis.
+A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increases persistence.
 The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend).
@@
 
-    \begin{figure}
+    \begin{figure}[hb!]
         \centering
-        \includegraphics[width=.5\textwidth]{figures/persistence_motif_profiles.eps}
+        \includegraphics[width=\textwidth]{figures/persistence_motif_profiles.eps}
         \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.
@@
 Interactions between motif profiles and disturbance were significant but small.
-Here we show the relationships of probabilities of extinction of basal resources of 0.1 (top set of lines) and 0.5 (bottom set of lines).
+Here we show the relationships of probabilities of extinction of basal resources of $\pi=0.1$ (top set of lines) and $\pi=0.5$ (bottom set of lines).
 Relationships are shown for the observed range of proportions for each motif.}
@@
     \end{figure}
-    \begin{figure}[h!]
+
+    \begin{figure}[hb!]
         \centering

### 2021-10-11 manuscript/reorder_main_text.tex

@@
        Species' motif participation varied depending on the global properties of the network: size, connectance, and their interaction (Fig. S1, Table S2; \emph{Appendix S5}).
-       Species appeared most often in the apparent competition motif in all types of networks, but the frequency decreased with increasing network size and connectance.
-       The frequency with which species participated in the omnivory motif increased sharply with increasing connectance, while the frequency of the direct competition motif increased slightly with increasing network size.
+       Species appeared most often in the apparent competition motif in all types of networks, but the proportion of apparent competition decreased with increasing network size and connectance.
+       The proportion of the omnivory motif in a species' participation role increased sharply with increasing connectance, while the proportion of the direct competition motif increased slightly with increasing network size.
 
@@
 
-    In general, average persistence was highest in networks with low frequencies of the omnivory motif and high frequencies of the apparent competition motif. 
-    These two motifs that have been shown to have consequences for the stability of ecological networks \citep{McLeod2019,Picot2018}.
-    Omnivory has been previously shown to both stabilize and destabilize communities, and the impact is highly dependent on network context \citep{bascompte2005simple, Monteiro2016}. 
-    Omnivorous interactions may not be stable in isolation but instead stabilized by interactions with other species when embedded in a larger food web \citep{Kratina2012}. 
+    In general, average persistence was highest in networks with low proportions of the omnivory motif and high proportions of the apparent competition motif. 
+    These results contrast with earlier work showing that higher numbers of omnivory motifs and lower numbers of apparent competition motifs are associated with greater food-web persistence~\citep{Stouffer2010b}.
+    By using counts of motifs rather than proportions,~\citet{Stouffer2010b} may be combining effects of total numbers of motifs with effects of motif proportions.
+    Moreover, the effect of omnivory on stability is known to be highly context-dependent \citep{bascompte2005simple, Monteiro2016,McLeod2021}. 
+    The stability of omnivory motifs in isolation depends on the strength of the omnivore-resource interaction~\citep{McLeod2021}, but at the whole-network level interactions with other species can stabilize even intrinsically unstable omnivory \citep{Kratina2012}.
     This might be why we find different patterns between a network's motif profile (where a high proportion of omnivory indicates lower average network persistence) versus a single species motif participation (where a high proportion of the omnivory motif indicates a higher focal species persistence).
@@
     % High C webs and omnivory
-    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks \emph{and} the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}).
+    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks \emph{and} the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}) as in~\citep{McLeod2021}.
     These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
@@
         \centering
-        \includegraphics[width=.75\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
-        \caption{The proportions of the four motifs in a network's motif profile were related to the mean persistence of species in the network.
-Specifically, persistence decreased as the proportion of omnivory in the network's motif profile increased while persistence increased with increasing proportions of the other three motifs.
-Interactions between motif profiles and disturbance were significant but small.
-Here we show the relationships of probabilities of extinction of basal resources of 0.1 (top set of lines) and 0.5 (bottom set of lines).
-Relationships are shown for the observed range of proportions for each motif.}
+        \includegraphics[width=\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
+        \caption{The proportions of the four motifs in a network's motif profile were network size, connectance, and their interaction.
+The effect of connectance was much stronger than the main effect or interaction with network size.
+The proportion of apparent competition in a network's profile decreased with increasing connectance, while the proportion of omnivory increased.
+In small networks, the proportion of direct competition also decreased with increasing connectance.}
         \label{motif_proportion_lms}

### 2021-10-12 code/stat_analysis/prop_basal_models.R
starting to update figures with ranefs for network ID

@@ -1,4 +1,4 @@
-load.image('all_tests.Rdata')
+load('all_tests.Rdata')
 
 # Libraries to fit the non-plotted lms
 library(lme4)
@@ -119,8 +119,8 @@ BCper=with(consumers,lm(Persistence~scale(prop_Basal)*scale(Size)*scale(Connecta
 
     for(c in 1:ncol(posprops)){
         posname=colnames(posprops)[c]
-        SClm=lm(posprops[,c]~consumers$prop_Basal*consumers$Connectance)
-        write.table(summary(SClm)$coefficients,file=paste0('positions_vs_other/BC_',posname,'.tsv',sep=''),sep='\t')
+        BClm=lm(posprops[,c]~consumers$prop_Basal*consumers$Connectance)
+        write.table(summary(BClm)$coefficients,file=paste0('positions_vs_other/BC_',posname,'.tsv',sep=''),sep='\t')
     }
 
 save.image('basal_tests.Rdata')

### 2021-10-12 code/stat_analysis/statistical_models.R

@@ -32,10 +32,12 @@ consumers$prop_chain=consumers$m12/consumers$total_motifs
 consumers$prop_apparent=consumers$m6/consumers$total_motifs
 consumers$prop_direct=consumers$m36/consumers$total_motifs
 consumers$prop_omni=consumers$m38/consumers$total_motifs
+consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep=':')
 
 # Section 1: How does persistence vary with global network properties?
 
     SCper=with(consumers,lm(Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)))
+    SCper_randnet=with(consumers,lmer(Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)+(1|netID)))
 
 
 # Section 2: How do motif profiles vary with S and C?
@@ -48,7 +50,6 @@ consumers$prop_omni=consumers$m38/consumers$total_motifs
 
     # Expanding motif profiles for each level of disturbance
     motdata$netID=paste(motdata$Size,motdata$Connectance,motdata$Network,sep=':')
-    consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep=':')
     consumers$net_chain=0
     consumers$net_apparent=0
     consumers$net_direct=0
@@ -61,14 +62,12 @@ consumers$prop_omni=consumers$m38/consumers$total_motifs
         consumers$net_omni[r]=motdata$prop_omni[which(motdata$netID==net)]
     }
 
-    save.image('analysis.Rdata')
-
     total_lm=with(motdata,lm(mot_total~Size*Connectance))
     summary(total_lm)
 
     # PERMANOVA and variability of motif profiles
     propdist=vegdist(motdata[,10:13],method="bray")
-    propperm=adonis(propdist~motdata$Size*motdata$Connectance,perm=9999)
+    propperm=adonis(propdist~motdata$Size*motdata$Connectance,perm=9999) # Should I stratify within a network? Probably doesn't matter so much since dispersion not normal
     motdata$interact=paste0(motdata$Size,':',motdata$Connectance)
     propinterdisp=betadisper(propdist,motdata$interact,type="centroid")
     Spropdisp=betadisper(propdist,motdata$Size,type="centroid")
[…]

### 2021-10-12 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 \renewcommand{\theequation}{S\arabic{equation}}
+\renewcommand{\labelenumi}{S\arabic{enumi}}
 
@@
 \medskip
-\small$^3$ University of Colorado?\\
+\small$^3$ BioFrontiers Institute\\
+University of Colorado, Boulder\\
+U.S.A.
 \medskip
@@
 
+    Appendices S1-S3 provide supplemental methods, while appendices S4-S8 provide supplemental results.
+
     \begin{enumerate}
 
-        \item Supplemental methods: Network creation \\
+        \item Network creation \\
 
@@
 
-        \item Supplemental methods: Calculating persistence
+        \item Calculating persistence
 
@@
 
-         \item Supplemental methods: Statistical analysis
+         \item Statistical analysis
 
@@
 
-        \item Results: Persistence and motif participation
+        \item Persistence and motif participation
             \begin{itemize}
@@
 
-        \item Results: Participation and network properties
+        \item Participation and network properties
 
@@
 
-        \item Results: Persistence, degree, and trophic level
+        \item Persistence, degree, and trophic level
 
@@
 
-        \item Results: Motif-persistence relationships
+        \item Within-network motif-persistence relationships
 
@@
 
-        \item Results: Persistence and network motif profiles
+        \item Persistence and network motif profiles
 
@@
 
-\section{Supplemental methods: Network creation}
+\section{Network creation}
 
-    After simulating our initial niche-model networks, we then filtered networks for biological plausibility.
+    After simulating our initial niche-model networks, we filtered networks for biological plausibility.
     Specifically, we removed any network containing disconnected components
@@
 
-\section{Supplemental methods: Calculating persistence}
+\section{Calculating persistence}
 
@@
 
-\section{Supplemental methods: Statistical analysis}
+\section{Statistical analysis}
 
@@
         \begin{equation}
-            \Psi_{ik} \approx \rho_{i} + \pi_{k} + \rho_{i}\pi_{k} +
-            S_{i}:C_{i} ,
+            \Psi_{ink} \approx \rho_{i} + \pi_{k} + \rho_{i}\pi_{k} +
+            S_{n}:C_{n} + N_n,
             \label{propreq}
@@
 
-        \noindent where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
+        \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{n}C_{n}$ is a random intercept for the species richness and connectance of the network containing species $i$, and $N_n$ is a random effect of network ID.
+        The network-level random effect accounts for the fact that species in the same network are not independent.
         We centered and scaled the predictors before fitting the LMMs.
@@
         Species' motif participation may covary with other measures of network structure and species' roles within the network.
-        To determine the relationships between motif participation and global network structure, we fit a set of four linear models (one per motif):
+        To determine the relationships between motif participation and global network structure, we fit a set of four linear mixed-effect models (one per motif):
 
         \begin{equation}
-            \rho_{i} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i} ,
+            \rho_{in} \approx \Sigma_{n} + \zeta_{n} + \Sigma_{n}\zeta_{n} + N_n ,
             \label{partic_SC}
@@
 
-        \noindent where $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif,
-        $\Sigma_{i}$ is the size of the network containing species $i$, and $\zeta_{i}$ is the connectance of the network containing species $i$.
+        \noindent where $\rho_{in}$ is the proportion of the role of species $i$, belonging to network $n$, that is made up by the focal motif,
+        $\Sigma_{n}$ is the size of network $n$, $\zeta_{n}$ is the connectance of the network $n$, and $N_n$ is a random effect of network ID.
         Predictors were not centered or scaled before fitting.
@@
         \begin{equation}
-            \rho_{i} \approx \delta_{i} + S_{i}C_{i} ,
+            \rho_{in} \approx \delta_{i} + S_{n}C_{n} + N_n,
             \label{partic_deg}
@@
         \begin{equation}
-            \rho_{i} \approx \tau_{i} + S_{i}C_{i} ,
+            \rho_{in} \approx \tau_{i} + S_{n}C_{n} + N_n,
             \label{partic_deg}
@@
 
-        \noindent where $\rho_{i}$ is the frequency with which species $i$ participtes in the focal motif, $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, $S_{i}C_{i}$ is a random effect of the combination of size and connectance for the network containing species $i$.
+        \noindent where $\rho_{in}$ is the frequency with which species $i$, belonging to network $n$, participates in the focal motif, $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, $S_{n}C_{n}$ is a random effect of the combination of size and connectance for network $n$, and $N_n$ is a random effect of network ID.
         These models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
@@
         Dispersion was calculated using the R~\citep{R} function `betadisper' from the package \emph{vegan}~\citep{vegan}.
-        In case high or low mean persistences are associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable).
+        In case high or low mean persistence is associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable).
 
@@
             \begin{equation}
-                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{k} + \Bar{\rho}_{i}\pi_{k} ,
+                \overline{\Psi_{nk}} \approx \Bar{\rho}_{n} + \pi_{k} + \Bar{\rho}_{n}\pi_{k} ,
                 \label{netpropeq}
@@
 
-        \noindent where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and all other symbols are as in equation~\ref{propreq}. 
+        \noindent where $\overline{\Psi_{nk}}$ is the mean persistence of all consumers in web $n$ at disturbance level $k$, $\Bar{\rho}_{n}$ is the proportion of the focal motif in the network's motif profile, and $\pi_k$ is the level of disturbance (probability of extinction for basal resources) as in equation~\ref{propreq}.
         All regressions were fit using the R~\citep{R} base function `lm'.
@@
 
-        To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif to network size, connectance, and their interaction:
+        To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif in a network's motif profile to network size, connectance, and their interaction:
 
         \begin{equation}
-            \Bar{\rho_{i}} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i}
+            \Bar{\rho_{n}} \approx \Sigma_{n} + \zeta_{n} + \Sigma_{n}\zeta_{n}
         \end{equation}
@@
             \begin{equation}
-                \Psi_{ik} \approx \Sigma_{i} + \zeta_{i} + \pi_{k} + \Sigma_{i}:\zeta_{i} + \Sigma_{i}\pi_{k} + \zeta_{i}\pi_{k} + \Sigma_{i}\zeta_{i}\pi_{k},
+                \Psi_{ink} \approx \Sigma_{n} + \zeta_{n} + \pi_{k} + \Sigma_{n}:\zeta_{n} + \Sigma_{n}\pi_{k} + \zeta_{n}\pi_{k} + \Sigma_{n}\zeta_{n}\pi_{k},
                 \label{SCeq}
@@
             \begin{equation}
-                    \Psi_{ik} \approx \delta_{i} + \pi_{k} + \delta_{i}\pi_{k} +
-                    S_{i}C_{i} ,
+                    \Psi_{ink} \approx \delta_{i} + \pi_{k} + \delta_{i}\pi_{k} +
+                    S_{n}C_{n} ,
                     \label{degeq}
@@
             \begin{equation}
-                    \Psi_{ik} \approx \tau_{i} + \pi_{k} + \tau_{i}\pi_{k} +
-                    S_{i}C_{i} ,
+                    \Psi_{ink} \approx \tau_{i} + \pi_{k} + \tau_{i}\pi_{k} +
+                    S_{n}C_{n} ,
                     \label{TLeq}
@@
 
-\section{Results: Persistence and motif participation}
+\section{Persistence and motif participation}
 
@@
 
+    % Updated with new random effect
     \begin{table}[h!]
         \centering
-        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as a random intercept to account for differences in persistence across levels of network size and connectance (Equation~\ref{propreq}).
+        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as random effects to account for differences in persistence across levels of network size and connectance and connectance (Equation~\ref{propreq}).
 Predictors were centered and scaled before model fitting.
@@
         \hline
-        3-sp Chain & 0.574 & \textless0.001 & -0.00508 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
-        App. Com. & 0.574 & \textless0.001 & -0.00312 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
-        Dir. Com. & 0.574 & \textless0.001 & 0.00564 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
-        Omnivory & 0.574 & \textless0.001 & 0.00709 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
+        3-sp Chain & 0.574 & \textless0.001 & -0.00492 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
+        App. Com. & 0.574 & \textless0.001 & -0.00390 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
+        Dir. Com. & 0.574 & \textless0.001 & 0.00607 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
+        Omnivory & 0.575 & \textless0.001 & 0.00806 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
         \end{tabular}
@@
 
-\section{Results: Participation and network properties}
+\section{Participation and network properties}
 
     Motif participation varies with global-scale network properies.
-
+    % Updated with network-level random effect
     \begin{table}[hb!]
@@
         \caption{Motif participation varies with size and connectance. Here we give the range and mean ($\pm$ standard deviation) proportion of each motif in species' motif participation vectors. 
-        We also show the effects ($\beta$) of network size, connectance, and their interaction in linear models predicting the proportion of each motif in a participation vector.
-        All effects were significant ($p$\textless0.001).
+        We also show the effects ($\beta$) and $p$-values for network size, connectance, and their interaction in linear mixed-effect models predicting the proportion of each motif in a participation vector.
+        These models also included a random effect of network ID, as species within the same network are not independent.
         Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
@@
         \footnotesize
-        \begin{tabular}{c|c c c | c c c}
-            Motif & Min & Max & Mean (SD) & Size & Connectance & Interaction \\
+        \begin{tabular}{c|c c c | c c | c c | c c}
+             &  &  &  & \multicolumn{2}{c}{Size} & \multicolumn{2}{|c}{Connectance} & \multicolumn{2}{|c}{Interaction} \\
+            Motif & Min & Max & Mean (SD) & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
             \hline
-            3-sp Chain & 0 & 1 & 0.249 (0.127) & 2.22$\times10^{-4}$ & -0.131 & -3.20$\times10^{-3}$ \\
-            App. Comp & 0.250 & 0.765 & 0.422 (0.0645) & -5.06$\times10^{-4}$ & -1.00 & 2.59$\times10^{-3}$ \\
-            Dir. Comp & 0.0501 & 0.359 & 0.179 (0.0407) & -3.56$\times10^{-5}$ & -0.556 & 5.92$\times10^{-3}$ \\
-            Omnivory & 0.006 & 0.346 & 0.164 (0.0790) & 3.64$\times10^{-4}$ & 1.71 & -5.24$\times10^{-3}$ \\   
+            3-sp Chain & 0 & 1 & 0.249 (0.127) & 2.83$\times10^{-4}$ & \textless0.001 & -0.0982 & 0.053 & -3.66$\times10^{-3}$ & \textless0.001 \\
+            App. Comp & 0.250 & 0.765 & 0.422 (0.0645) & -5.85$\times10^{-4}$ & \textless0.001 & -0.840 & \textless0.001 & 2.69$\times10^{-3}$ & 0.003 \\
+            Dir. Comp & 0.0501 & 0.359 & 0.179 (0.0407) & -6.99$\times10^{-5}$ & 0.450 & -0.636 & \textless0.001 & 6.46$\times10^{-3}$ & \textless0.001 \\
+            Omnivory & 0.006 & 0.346 & 0.164 (0.0790) & 3.70$\times10^{-4}$ & \textless0.001 & 1.57 & \textless0.001 & -5.48$\times10^{-3}$ & \textless0.001\\   
             \hline
@@
 
+    % Updated with network-level random effect
     \begin{figure}[h!]
@@
 
-\section{Results: Persistence, degree, and trophic level}
+\section{Persistence, degree, and trophic level}
 
@@
      \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
-     \caption{Basal resources experience a disturbance reducing their probability to go extinct to $\pi = 0.5$ All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
+     \caption{Basal resources experience a disturbance increasing their probability to go extinct to $\pi = 0.5$.
+All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
 The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
@@
 
-\section{Results: Motif-persistence relationships}
+\section{Within-network motif-persistence relationships}
 
@@
 3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately.
-    The main trends are similar in both analyses, but different-sized networks showed slightly different results results for some motifs when considered individually (Figure~\ref{fig:density_prop_S}).
+    The main trends are similar in both analyses, but different-sized networks showed slightly different results for some motifs when considered individually (Figure~\ref{fig:density_prop_S}).
     For the three-species chain and omnivory motifs, a significant interaction between disturbance and the effect of the focal motif means that having a larger proportion of the role be made up of the chain or omnivory motif is associated with lower persistence at high levels of disturbance (Fig.
@@
 Rows show various network sizes.
-The dotted, vertical line indicate zero on the x-axis.
+The dotted, vertical lines indicate zero on the x-axis.
 A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
@@
 
-\section{Results: Persistence and motif profiles}
+\section{Persistence and motif profiles}
 
@@
 
-\section{Results: Motif profiles and global properties}
+\section{Motif profiles and global properties}
 
@@
 
-\section{Results: Persistence and global network properties}
+\section{Persistence and global network properties}
 

### 2021-10-12 manuscript/reorder_main_text.tex

@@
             \begin{equation}
-                \Psi_{ik} \approx \rho_{i} + \pi_{k} + \rho_{i}\pi_{k} +
-                S_{i}C_{i} ,
+                \Psi_{ink} \approx \rho_{i} + \pi_{k} + \rho_{i}\pi_{k} +
+                S_{n}C_{n} ,
                 \label{propreq}
             \end{equation}
-        \noindent where $\Psi_{ik}$ is the persistence of species $i$ during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{i}C_{i}$ is a random intercept for the species richness and connectance of the network containing species $i$.
+        \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{n}C_{n}$ is a random intercept for the species richness and connectance of network $n$.
         All predictors were centered and scaled before fitting the models.         
@@
             \begin{equation}
-                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{k} + \Bar{\rho}_{i}\pi_{k},
+                \overline{\Psi_{nk}} \approx \Bar{\rho}_{n} + \pi_{k} + \Bar{\rho}_{n}\pi_{k},
                 \label{netpropeq}
             \end{equation}
-        \noindent where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and all other symbols are as in equation~\ref{propreq}.
+        \noindent where $\overline{\Psi_{nk}}$ is the mean persistence of all consumers in network $n$ at disturbance level $k$, $\Bar{\rho}_{n}$ is the proportion of the focal motif in the network's motif profile, and $\pi_k$ is the level of disturbance (probability of extinction for basal resources) as in equation~\ref{propreq}.
 
@@
         \begin{equation}
-            \Bar{\rho_{i}} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i}
+            \Bar{\rho_{n}} \approx \Sigma_{n} + \zeta_{n} + \Sigma_{n}\zeta_{n}
         \label{profile_global}
         \end{equation}
-        \noindent where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and $\Sigma_i$ and $\zeta_{i}$ are the size and connectance of the network to which $i$ belongs.
+        \noindent where $\Bar{\rho}_{n}$ is the proportion of the focal motif in the motif profile of network $n$ and $\Sigma_n$ and $\zeta_{n}$ are the size and connectance of network $n$.
 
@@
             \begin{equation}
-                \Psi_{ik} \approx \Sigma_{i} + \zeta_{i} + \pi_k + \Sigma_{i}:\zeta_{i} + \Sigma_{i}\pi_k + \zeta_{i}\pi_k + \Sigma_{i}\zeta_{i}\pi_k,
+                \overline{\Psi_{nk}} \approx \Sigma_{n} + \zeta_{n} + \pi_k + \Sigma_{n}:\zeta_{n} + \Sigma_{n}\pi_k + \zeta_{n}\pi_k + \Sigma_{n}\zeta_{n}\pi_k,
                 \label{SCeq}
@@
         Persistence is also known to vary with local-scale measures of network structure. 
-        To identify the relationships betwen local-scale properties and species' probabilities of persistence, we fit two final LMMs:
+        To identify the relationships between local-scale properties and species' probabilities of persistence, we fit two final LMMs:
         \begin{equation}
-                \Psi_{ik} \approx \delta_{i} + \pi_k + \delta_{i}\pi_k +
-                S_{i}C_{i} ,
+                \Psi_{ink} \approx \delta_{i} + \pi_k + \delta_{i}\pi_k +
+                S_{n}C_{n} ,
                 \label{degeq}
@@
         \begin{equation}
-                \Psi_{ik} \approx \tau_{i} + \pi_k + \tau_{i}\pi_k +
-                S_{i}C_{i} ,
+                \Psi_{ink} \approx \tau_{i} + \pi_k + \tau_{i}\pi_k +
+                S_{n}C_{n} ,
                 \label{TLeq}
@@
     \subsection*{Species motif participation and species persistence} 
-
+    % Coefficients changed slightly with network-level random effect, but conclusions similar
        The relationship between frequency of participation in a motif and a species' probability of persistence depended on the level of disturbance. 
@@
     \subsection*{Motif participation and other network properties}
-
+        % Updated with network-level random effect. No visible change.
        Species' motif participation varied depending on the global properties of the network: size, connectance, and their interaction (Fig. S1, Table S2; \emph{Appendix S5}).
@@
 
+        % Updated with network-level random effect. No visible change.    
        Species' motif participation also varied with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).

### 2021-10-13 manuscript/supplement_motifs.tex
updated persistence vs SCg

@@
 
-            To test for effects of global network properties, we fit a linear regression relating consumer species' probabilities of persistence to network size, connectance, the level of disturbance to basal resources, and all interactions between them.
+            To test for effects of global network properties, we fit a linear regression relating the mean of consumer species' probabilities of persistence in a network to network size, connectance, the level of disturbance to basal resources, and all interactions between them.
             All predictors were centered and scaled before fitting the model. 
@@
             \begin{equation}
-                \Psi_{ink} \approx \Sigma_{n} + \zeta_{n} + \pi_{k} + \Sigma_{n}:\zeta_{n} + \Sigma_{n}\pi_{k} + \zeta_{n}\pi_{k} + \Sigma_{n}\zeta_{n}\pi_{k},
+                \overline{\Psi_{nk}} \approx \Sigma_{n} + \zeta_{n} + \pi_k + \Sigma_{n}:\zeta_{n} + \Sigma_{n}\pi_k + \zeta_{n}\pi_k + \Sigma_{n}\zeta_{n}\pi_k,
                 \label{SCeq}
@@
                     \Psi_{ink} \approx \delta_{i} + \pi_{k} + \delta_{i}\pi_{k} +
-                    S_{n}C_{n} ,
+                    S_{n}C_{n} + N_n,
                     \label{degeq}
@@
                     \Psi_{ink} \approx \tau_{i} + \pi_{k} + \tau_{i}\pi_{k} +
-                    S_{n}C_{n} ,
+                    S_{n}C_{n} N_n,
                     \label{TLeq}
@@
         \centering
-        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as random effects to account for differences in persistence across levels of network size and connectance and connectance (Equation~\ref{propreq}).
+        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as random effects to account for differences in persistence across levels of network size, connectance, and network ID  (Equation~\ref{propreq}).
 Predictors were centered and scaled before model fitting.
@@
 
+    % Updated with network-level random effect
     \begin{table}[h!]
-        \caption{Coeffcients ($\beta$) for regressions of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction.
+        \caption{Coeffcients ($\beta$) for linear mixed-effect models of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction, as well as a random effect of network ID.
 Predictors were centered and scaled before model fitting.
@@
             \hline
-            STL & -0.0218 & -0.139 & 0.00508 \\
-            in-degree & -0.00130 & -0.139 & -0.00908 \\
+            STL & -0.0210 & -0.139 & 0.00508 \\
+            in-degree & -0.00188 & -0.139 & -0.00908 \\
         \end{tabular}
@@
 
+    % Updated to use mean rather than all species.
     \begin{table}[hb!]
-        \caption{Persistence of consumers within a network depends on network size (S) and connectance (C), the level of disturbance to basal resources (D), and interactions between them.
+        \caption{Mean persistence of consumers within a network depends on network size (S) and connectance (C), the level of disturbance to basal resources (D), and interactions between them.
 Here we show the effect sizes ($\beta$) and $p$-values for the regression described above.
@@
             \hline
-            Intercept & 0.572 & \textless0.001 \\
-            Size & -5.15$\times10^{-4}$  & \textless0.001 \\
-            Connectance & -0.0129 & \textless0.001 \\
+            Intercept & 0.574 & \textless0.001 \\
+            Size & -4.82$\times10^{-4}$  & 0.011 \\
+            Connectance & -0.0137 & \textless0.001 \\
             Disturbance & -0.139 & \textless0.001 \\
-            S:C & -5.24$\times10^{-4}$ & \textless0.001 \\
-            S:D & -3.48$\times10^{-3}$ & \textless0.001 \\
-            C:D & -7.00$\times10^{-5}$ & 0.095 \\
-            S:C:D: & -2.69$\times10^{-4}$ & \textless0.001 \\
+            S:C & -9.77$\times10^{-4}$ & \textless0.001 \\
+            S:D & -3.46$\times10^{-3}$ & \textless0.001 \\
+            C:D & -2.25$\times10^{-4}$ & 0.233 \\
+            S:C:D: & -3.97$\times10^{-4}$ & 0.036 \\
         \end{tabular}

### 2021-10-13 code/stat_analysis/persistence_vs_SC_lm.py

@@ -33,14 +33,14 @@ scales={
 }
 
 lms={
-  'Intercept': 5.723e-01,
-  'Size':-5.148e-04,
+  'Intercept': 0.5743430,
+  'Size':-0.0004818,
   'Connectance':-1.293e-02,
-  'Disturbance':-1.393e-01,
-  'Size:Connectance':-5.245e-04,
-  'Size:Disturbance':-3.485e-03,
-  'Connectance:Disturbance':-6.998e-05,
-  'Size:Connectance:Disturbance':-2.693e-04 }
+  'Disturbance':-0.0136768,
+  'Size:Connectance':-0.0009770,
+  'Size:Disturbance':-0.0034566,
+  'Connectance:Disturbance':-0.0002251,
+  'Size:Connectance:Disturbance':-0.0003969 }
 
 Blms={'Intercept':5.777e-01,
       'Basal':2.841e-02,

### 2021-10-13 manuscript/supplement_motifs.tex
updated all with Kate's random effect

@@
 
-    Persistence varied with the frequencies of different motifs in a network profile.
+    % Updated and expanded to include permanova tests
+    Persistence varied with the frequencies of different motifs in a network profile ($F_{1,2998}$=645, $p$\textless0.001 for a PERMANOVA of dissimilarity in motif profiles against mean persistence).
+    However, motif profiles were not evenly distributed across mean persistences ($F_{134,2865}$=1.77, $p$\textless0.001 for an ANOVA of dissimilarity in motif profiles against mean persistence, rounded to 3 decimal places).
+    Specifically, motif profiles were more variable (higher dissimilarities) in networks with higher mean persistence.
+    As non-homogeneous variability can cause a PERMANOVA to return a false positive~\citep{Anderson2001}, this means that we must treat our PERMANOVA result as tentative and place more weight on the single-motif regressions.
 
+    % Updated to use mean instead of all species
     \begin{table}[h!]
@@
 Predictors were centered and scaled before model fitting.
-All effects were significant ($p$\textless0.001).
-We also provide the $R^2$.
+All effects were significant ($p$\textless0.05).
+We also provide the $R^2$ (multiple $R^2$).
 }
@@
             \hline
-            Omnivory & 0.572 & -0.0133 & -0.139 & -4.89$\times10^{-4}$ & 0.908 \\
-            Apparent competition & 0.572 & 0.0105 & -0.139 & 4.78$\times10^{-4}$ & 0.905 \\
-            Direct competition & 0.572 & 2.62$\times10^{-3}$ & -0.139 & 3.84$\times10^{-4}$ & 0.901 \\
-            Three-species chain & 0.572 & 6.12$\times10^{-3}$ & -0.139 & -1.90$\times10^{-4}$ & 0.901 \\ 
+            Omnivory & 0.574 & -0.0140 & -0.139 & -5.59$\times10^{-4}$ & 0.968 \\
+            Apparent competition & 0.574 & 0.0112 & -0.139 & 5.90$\times10^{-4}$ & 0.964 \\
+            Direct competition & 0.574 & 3.38$\times10^{-3}$ & -0.139 & 1.81$\times10^{-4}$ & 0.958 \\
+            Three-species chain & 0.574 & 5.80$\times10^{-3}$ & -0.138 & -3.07$\times10^{-5}$ & 0.960 \\ 
         \end{tabular}
@@
 \section{Motif profiles and global properties}
+% No update needed
 
@@
 
+    % updated to use network means rather than all species
     \begin{figure}[h!]

### 2021-10-13 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
                 \Psi_{ink} \approx \rho_{i} + \pi_{k} + \rho_{i}\pi_{k} +
-                S_{n}C_{n} ,
+                S_{n}C_{n} + N_n,
                 \label{propreq}
             \end{equation}
-        \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{n}C_{n}$ is a random intercept for the species richness and connectance of network $n$.
+        \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$,  $S_{n}C_{n}$ is a random intercept for the species richness and connectance of network $n$, and $N_n$ is a random effect of network $n$, intended to account for the non-independence of motif participations for species in the same network.
         All predictors were centered and scaled before fitting the models.         
@@
     \subsubsection*{2) Participation and other network properties}    
-
         Different scales of network structure (e.g., local and meso-scale) are not independent. 
@@
                 \Psi_{ink} \approx \delta_{i} + \pi_k + \delta_{i}\pi_k +
-                S_{n}C_{n} ,
+                S_{n}C_{n} + N_n,
                 \label{degeq}
@@
                 \Psi_{ink} \approx \tau_{i} + \pi_k + \tau_{i}\pi_k +
-                S_{n}C_{n} ,
+                S_{n}C_{n} + N_n,
                 \label{TLeq}
@@
        Species with higher trophic levels tended to have higher proportions of apparent competition and three-species chains. 
-
-       A species persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix S6}).
+        % Updated with network-level random effect. No visible change.   
+       A species' persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix S6}).
 Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S3, \emph{Appendix S6}).
@@
 
+        %Update not needed
         When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally showed more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
@@
     \subsection*{Network motif profiles and network mean persistence}
-
-        The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=571, $p$\textless0.001). 
+        % Updated to use means instead of all species
+        The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=645, $p$\textless0.001), although this result must be considered tentative due to non-homogeneous distribution of roles (\emph{Appendix S8}).
         Average persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S4, \emph{Appendix S8}). 

### 2021-10-13 code/stat_analysis/persistence_vs_profiles.py
updated Fig. 4, no visible change

@@ -39,18 +39,18 @@ mot_ranges={
 }
 
 scales={ # center, scale
-  'omnivory':(0.169822,0.07561787),
-  'apparent':(0.4169211,0.06134048),
-  'direct':(0.1805743,0.04025804),
-  'chain':(0.2326826,0.04136131),
-  'Disturbance':(0.3,0.1366261)
+  'omnivory':(0.1641884,0.07895101),
+  'apparent':(0.4222605,0.06446181),
+  'direct':(0.1787041,0.0407462),
+  'chain':(0.234847,0.04233961),
+  'Disturbance':(0.3,0.1366298)
 }
 
 lms={ # Intercept, motif, disturbance, interaction
-  'omnivory':(5.723e-01,-1.328e-02,-1.393e-01,-4.886e-04),
-  'apparent':(5.723e-01,1.053e-02,-1.393e-01,4.781e-04),
-  'direct':(5.723e-01,2.618e-03,-1.393e-01,3.844e-04),
-  'chain':(0.5723309,0.0061244,-0.1393351,-0.0001900)  
+  'omnivory':(0.5743430145,-0.0140077658,-0.1385163991,-0.0005586564),
+  'apparent':(0.574343015,0.011206293,-0.138516399,0.000589999),
+  'direct':(0.5743430145,0.0033850297,-0.1385163991,0.0001809503),
+  'chain':(5.743430e-01,5.801241e-03,-1.385164e-01,-3.067919e-05)  
 }
 
 motif_names={'omnivory':'Omnivory','apparent':'Apparent competition','direct':'Direct competition','chain':'Three-species chain'}

### 2021-10-13 code/stat_analysis/statistical_models.R
updating code file with added random effects

@@ -34,11 +34,6 @@ consumers$prop_direct=consumers$m36/consumers$total_motifs
 consumers$prop_omni=consumers$m38/consumers$total_motifs
 consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep=':')
 
-# Section 1: How does persistence vary with global network properties?
-
-    SCper=with(consumers,lm(Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)))
-    SCper_randnet=with(consumers,lmer(Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)+(1|netID)))
-
 
 # Section 2: How do motif profiles vary with S and C?
     motdata=read.table('../../data/3sp_motif_profiles.tsv',header=TRUE)
@@ -137,6 +132,10 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     # LM dispersion ~ mean persistence
     persist_lm=lm(persist_disp$distances~round(motdata$persistence,3))
 
+# Section 1: How does persistence vary with global network properties?
+
+    # SCper=with(consumers,lm(Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)))
+    SCper_mean=with(simdata,lm(per~scale(Size)*scale(Connectance)*scale(dist)))
 
 # Section 4: How does persistence vary with proportions?
     # # lmer with random intercepts for each level of S:C
@@ -164,14 +163,14 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
 
 # Section 5. How does persistence vary with degree and TL?
     # Persistence vs. STL and Degree
-    TLper=with(consumers,lmer(Persistence~scale(STL)*scale(Disturbance)+(1|Global)))
-    Degper=with(consumers,lmer(Persistence~scale(in_Degree)*scale(Disturbance)+(1|Global)))
+    # TLper=with(consumers,lmer(Persistence~scale(STL)*scale(Disturbance)+(1|Global)))
+    # Degper=with(consumers,lmer(Persistence~scale(in_Degree)*scale(Disturbance)+(1|Global)))
 
     TLper_randnet=with(consumers,lmer(Persistence~scale(STL)*scale(Disturbance)+(1|Global)+(1|netID)))
     Degper_randnet=with(consumers,lmer(Persistence~scale(in_Degree)*scale(Disturbance)+(1|Global)+(1|netID)))
 
[…]

### 2021-10-14 manuscript/cover_letter.tex
Updates from Overleaf

@@
 
-Alyssa Cirtwill has earlier been approached by the Proceedings B Preprint Editorial Team, considering another paper regarding motifs.
+Alyssa Cirtwill was asked by the Proceedings B Preprint Editorial Team to consider publishing a thematically similar paper in this journal.
+That manuscript is currently in revisions elsewhere, but we consider that this manuscript will also be a good topical fit.
+Both manuscripts deal with the relationship between species' motif roles and risk of extinction, although in different modelling frameworks and using different types of disturbance.
 
@@
 
+%The cover letter is used to tell the handling editor why you have submitted to this journal, and provide any additional information that might not be available within the manuscript, for example setting the paper in context within the field, alerting the editor to related papers in the field, or explaining choices of preferred or non-preferred reviewers.
+
 \vspace{2\parskip}

### 2021-10-14 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-    Bottom-up effects of disturbance to basal resources can strongly affect the persistence of consumers.
-    However, predicting which species will be most affected by increasing disturbance is not straightforward.
-    In-degree and trophic level are known to affect a consumer's vulnerability to bottom-up disturbance, but these are relatively coarse measures of a species' place in a food web.
-    Here we explore how roles based on participation in three-species motifs are related to persistence, and how these motif-based roles relate to simpler measures.
-    We show that a consumer species' frequency of participation in different motifs is related to its probability of persistence.
-Similarly, the overall motif profile of a network was related to the average persistence of its consumer species.
-    Notably, we found that networks containing a higher proportion of the `omnivory' motif tend to have lower mean persistence following disturbance to basal species, while more frequent participation in the omnivory motif was associated with higher probability of persistence.
+    Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.
+    Coarse measures of a species' structural role, like in-degree and trophic level, are known to affect a consumer's vulnerability to bottom-up disturbance.
+    We are interested in whether a richer set of species roles based on participation in three-species motifs is also related to persistence, and how these motif-based roles relate to simpler measures.
+    We show that a consumer species' participation in different motifs is related to its probability of persistence. 
+    Similarly, the overall motif profile of a network was related to the average persistence of consumer species.
+    Notably, we found that networks containing a higher proportion of the `omnivory' motif tend to have lower mean persistence following disturbance, while more frequent participation in the omnivory motif was associated with higher species-level probabilities of persistence.
     Additionally, we find that relationships between motif participation and persistence can synthesize the information provided by in-degree and trophic level.
-    Motif participation therefore provides a relatively rich description of a species' structural role in its community and captures important trends in vulnerability to extinction.
+    Motif participation therefore captures important trends in vulnerability to extinction and provides a rich description of species' structural roles in their communities.
 
@@
         To correctly interpret any relationships between motifs and persistence, we must consider the context of these other relationships (Appendix \emph{S3.5}).
-        To identify the relationship between global network structure and mean probability of persistence within a network, we fit a linear regression:
-            \begin{equation}
-                \overline{\Psi_{nk}} \approx \Sigma_{n} + \zeta_{n} + \pi_k + \Sigma_{n}:\zeta_{n} + \Sigma_{n}\pi_k + \zeta_{n}\pi_k + \Sigma_{n}\zeta_{n}\pi_k,
-                \label{SCeq}
-            \end{equation}
-        \noindent where all symbols are as in equation~\ref{propreq} and~\ref{profile_global}.
+        To identify the relationship between global network structure and mean probability of persistence within a network, we fit a linear regression relating mean persistence to network size, connectance, disturbance, and their interactions (Eg.
+S9).
         As the three-way interaction was significant, we did not simplify the model.
@@
         Persistence is also known to vary with local-scale measures of network structure. 
-        To identify the relationships between local-scale properties and species' probabilities of persistence, we fit two final LMMs:
-        \begin{equation}
-                \Psi_{ink} \approx \delta_{i} + \pi_k + \delta_{i}\pi_k +
-                S_{n}C_{n} + N_n,
-                \label{degeq}
-            \end{equation}
-        \begin{equation}
-                \Psi_{ink} \approx \tau_{i} + \pi_k + \tau_{i}\pi_k +
-                S_{n}C_{n} + N_n,
-                \label{TLeq}
-            \end{equation}
-        \noindent where $\delta_{i}$ and $\tau_i$ are the in-degree and trophic level of species $i$, respectively, and all other terms are as in equation~\ref{propreq}.
+        To identify the relationships between local-scale properties and species' probabilities of persistence, we fit two final LMMs relating a consumer's persistence to disturbance, in-degree or STL, and the interaction between disturbance and the local-scale measure (Eq.
+S10-S11).
+        These LMMs also included random effects to account for the effects of global network structure and network ID on persistence.
 
@@
     These results contrast with earlier work showing that higher numbers of omnivory motifs and lower numbers of apparent competition motifs are associated with greater food-web persistence~\citep{Stouffer2010b}.
-    By using counts of motifs rather than proportions,~\citet{Stouffer2010b} may be combining effects of total numbers of motifs with effects of motif proportions.
+    By using counts of motifs rather than proportions,~\citet{Stouffer2010b} may be mixing effects of total numbers of motifs with effects of motif proportions.
     Moreover, the effect of omnivory on stability is known to be highly context-dependent \citep{bascompte2005simple, Monteiro2016,McLeod2021}. 

### 2021-10-14 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
         \centering
-        \includegraphics[height=.5\textheight]{figures/participation_vs_SC.eps}
+        \includegraphics[height=.4\textheight]{figures/participation_vs_SC.eps}
         \caption{The proportion of each motif in a consumer's participation vector varied with network size, connectance, and their interaction.
@@
      \centering
-     \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}
+     \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}
      \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
@@
      \centering
-     \includegraphics[width=0.8\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
+     \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
      \caption{Basal resources experience a disturbance increasing their probability to go extinct to $\pi = 0.5$.
@@
         \centering
-            \includegraphics[width=0.75\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
+            \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
             \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all connectances - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
@@
         \centering
-        \includegraphics[width=0.95\textwidth]{figures/persistence_vs_SC_lm.eps}
+        \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
         \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while network size, connectance, and their interaction had smaller effects.
@@
     \begin{figure}[h!]
-      \includegraphics[width=0.95\textwidth]{figures/heatmap_CS_BPcompare.eps}
+      \includegraphics[width=\textwidth]{figures/heatmap_CS_BPcompare.eps}
        \caption{Heat maps of average consumer persistence in networks with varying connectance (y-axis) and network size (x-axis) at two levels of disturbance to basal resources.

### 2021-10-21 code/plot_roles_vs_TL.py
rescaled eps figs

@@ -190,10 +190,10 @@ def populate_persgraph(graph,persdict,simple):
       data.legend=str(basal_p)
 
   if simple=='TL':
-    graph.add_drawing_object(DrawText,text='Basal species',x=5.25,y=0.58,loctype='world',char_size=.75) 
-    graph.add_drawing_object(DrawText,text='extinction',x=5.25,y=0.52,loctype='world',char_size=.75) 
-    graph.add_drawing_object(DrawText,text='probability',x=5.25,y=0.47,loctype='world',char_size=.75) 
-    graph.legend.configure(char_size=.75,loc=(5.25,0.45),loctype='world',box_linestyle=0,fill_pattern=0)
+    graph.add_drawing_object(DrawText,text='Basal species',x=5.25,y=0.71,loctype='world',char_size=.75) 
+    graph.add_drawing_object(DrawText,text='extinction',x=5.25,y=0.65,loctype='world',char_size=.75) 
+    graph.add_drawing_object(DrawText,text='probability',x=5.25,y=0.58,loctype='world',char_size=.75) 
+    graph.legend.configure(char_size=.75,loc=(5.25,0.55),loctype='world',box_linestyle=0,fill_pattern=0)
 
 
   return graph
@@ -228,10 +228,10 @@ for graphtype in ['Proportion','Persistence']:
     
 grace.multi(rows=2,cols=2,vgap=.05,hgap=.05)
 grace.hide_redundant_labels()
-# grace.set_row_xaxislabel(label='in-degree (number of prey)',row=0,colspan=(None,None),char_size=1,perpendicular_offset=.05)
-# grace.set_row_xaxislabel(label='Trophic level (STL)',row=1,colspan=(None,None),char_size=1,perpendicular_offset=.05)
-# grace.set_col_yaxislabel(label='Count of motif',col=0,rowspan=(None,None),char_size=1,perpendicular_offset=.07)
-# grace.set_col_yaxislabel(label='Proportion of motif role',col=1,rowspan=(None,None),char_size=1,perpendicular_offset=.07)
+grace.graphs[0].set_view(0.10, 0.55, 0.5, 0.85)
+grace.graphs[1].set_view(0.55, 0.55, 0.95, 0.85)
+grace.graphs[2].set_view(0.10, 0.2, 0.5, 0.5)
+grace.graphs[3].set_view(0.55, 0.2, 0.95, 0.5)
 
 grace.write_file('../manuscript/figures/roles_vs_TL.eps')
 

### 2021-10-21 code/stat_analysis/persistence_vs_profiles.py

@@ -173,6 +173,8 @@ graph2.add_drawing_object(DrawText,x=0.5,y=0.34,loctype='world',text='Highest di
 
 graph2.legend.configure(char_size=1,loc=(0.05,0.685),loctype='world',box_linestyle=0,fill_pattern=0)
 
+grace.graphs[0].set_view(0.07,0.07,0.95,0.7)
+
 grace.write_file('../../manuscript/figures/persistence_motif_profiles.eps')
 
 

### 2021-10-21 manuscript/cover_letter.tex

--- 2021-10-08 manuscript/cover_letter.tex
@@
-\documentclass[11pt]{letter}
+\documentclass[10.5pt]{letter}
 
@@
 
-Please find enclosed the manuscript ``Does motif participation provide unique information about species risk of extinction?'' by Anna {\AA}kesson et al., which we would like to be considered for publication in \emph{Proceedings of the Royal Society B} as an Article.
+We invite you to consider the enclosed ``Does motif participation provide unique information about species risk of extinction?'' by Anna {\AA}kesson et al., for publication in \emph{Proceedings of the Royal Society B} as an Article.
+Dr.
+Cirtwill was approached by the Proceedings B Preprint Editorial Team to consider publishing a thematically similar paper in this journal.
+That manuscript is currently in revisions elsewhere, but we believe that this manuscript also would be a good topical fit.
+%Both manuscripts deal with the relationship between species' motif roles and risk of extinction, although in different modelling frameworks and using different types of disturbance.
 
-Alyssa Cirtwill has earlier been approached by the Proceedings B Preprint Editorial Team, considering another paper regarding motifs.
+Ecological effects of disturbances in food webs are known to depend on global (e.g.
+food web connectance) and local (e.g.
+trophic level) network properties.
+Disturbances to the basal level of a food web have a particularly strong impact on the persistence of all species.
+However, the locations for secondary extinctions are difficult to identify as both direct and indirect interactions among species are important.
+As global network properties are too coarse-grained to include interactions, and local network properties disregard any indirect effects, we must consider other structural properties in order to fully understand secondary extinctions.
+Meso-scale network properties, such as smaller sub-graphs within a large network (motifs), offer a potentially valuable middle-ground approach.
+Motifs have been shown to be important for community-level stability, but how motifs relate to risk of secondary extinctions is so far unexplored.
 
-While there is pressing need to understand ...
-Here we study the effects of these processes ... 
-Several key results emerge ...
+Here, we use a Bayesian network approach to simulate different levels of disturbance in complex food webs.
+In this framework, we analyse the relationships between consumers' participation in different three-species motifs and extinction risks, under the influence of different levels of disturbance to the basal species.
+Moreover, we analyze whether the motif profile of the full food web affects the extinction risk of consumer species.
+This extra layer of analysis places motifs as a bridge between species-level and network-level properties affecting extinction risk.
+Several novel key results emerge.
 
-First ...
+First, we show that a species' participation in different motifs is clearly related to its probability of extinction.
+Interestingly, this relationship is strongly influenced by the level of the disturbance, which indicates important 'tipping points' where increased level of disturbance can turn a previously beneficial characteristic into a risk factor.
+Second, we demonstrate that relationships between motif participation and extinction risk synthesize the information provided by simpler local properties, such as trophic level and in-degree. 
+Third, we show that the overall motif profile of a network is related to the average persistence of consumer species, capturing the same information as network size and connectance.
 
-Second ...
-
-Third ...
-
-We reach our conclusions by ...
-
-Due to the relevance of our results to understanding ..., for which \emph{ Proceedings of the Royal Society B} would be an ideal platform.
+As we take an important step towards understanding of how secondary effects propagate in complex food webs, we believe our method and results will interest a wide audience of ecologists including both theoreticians and empiricists.
+We therefore think \emph{ Proceedings of the Royal Society B} would be an ideal platform for our research.
 
@@
 
+\vspace{4\parskip}
+\centering
+\includegraphics[width=60mm]{manuscript/figures/LiU_primary_black.pdf}
+
 \end{letter} 

### 2021-10-21 manuscript/reorder_main_text.tex

@@
 
-\title{Does motif participation provide unique information about species risk of extinction?}
+\title{Species motif participation provides unique information about species risk of extinction}
 
@@
 \maketitle 
+\linenumbers
 \raggedright
@@
         In our baseline scenario, basal resources had an extinction probability of $\pi = 0.1$. 
-		To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase the affected species' probability of extinction ($\pi$). 
-		We increased $\pi$ for basal resources from $0.10$ to $0.5$, in steps of $0.08$. 
+		To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase the affected species' probability of extinction from $0.10$ to $0.5$, in steps of $0.08$.
 		The highest disturbance level, $\pi = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
@@
         To correctly interpret any relationships between motifs and persistence, we must consider the context of these other relationships (Appendix \emph{S3.5}).
-        To identify the relationship between global network structure and mean probability of persistence within a network, we fit a linear regression relating mean persistence to network size, connectance, disturbance, and their interactions (Eg.
+        To identify the relationship between global network structure and mean probability of persistence within a network, we fit a linear regression relating mean persistence to network size, connectance, disturbance, and their interactions (Eq.
 S9).
@@
 
-    The authors thank György Barabás for excellent discussion of the ideas in this manuscript. ARC is supported by a Finnish Academy postdoctoral research grant (\#332999).
+    The authors thank György Barabás for excellent discussion of the ideas in this manuscript.
+AÅ is supported by the Swedish Research Council for Sustainable Development Formas (\#2015-01262) granted to AE.
+ARC is supported by a Finnish Academy postdoctoral research grant (\#332999).
 
 \clearpage
-
-\section*{Figure captions}
 

### 2022-02-14 manuscript/response_PRSB_round1.tex
Updates from Overleaf

\documentclass[12pt]{article} 
\usepackage{amsmath} 
\usepackage[dvips]{graphicx}
\usepackage{multirow} 
\usepackage{geometry} 
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage[labelfont=bf]{caption} 
\usepackage{setspace}
\usepackage[running]{lineno} 
% \usepackage[numbers,sort]{natbib}
\usepackage[round]{natbib} 
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{xr}

\newcommand{\us}{\rm \setlength{\leftskip}{0.3cm} \setlength{\rightskip}{0.3cm}}
\newcommand{\them}{\it \setlength{\leftskip}{0cm} \setlength{\rightskip}{0cm}}

\topmargin -1.5cm % 0.0cm 
\oddsidemargin 0.0cm % 0.2cm 
\textwidth 6.5in
\textheight 9.0in % 21cm
\footskip 1.0cm % 1.0cm

\usepackage{authblk}

\setlength{\parskip}{0.8em}
\setlength{\parindent}{0em}
\bibpunct{(}{)}{,}{a}{}{,}
\clubpenalty = 10000
\widowpenalty = 10000

\title{RESPONSE LETTER \\Species motif participation provides unique information about species risk of extinction}

\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{2}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{1}$} 
\date{\small$^1$Department of Theoretical Biology, Chemistry, and Physics\\ 
Link\"{o}ping University\\
Link\"{o}ping, Sweden\\
\medskip
\small$^2$Department of Agricultural Sciences\\
University of Helsinki\\
Helsinki, Finland\\
\medskip
\small$^3$ BioFrontiers Institute\\
University of Colorado, Boulder\\
U.S.A.
\medskip
$^\dagger$ Corresponding author:\\
}

\begin{document} 
\maketitle 
\linenumbers
\raggedright

\setlength{\parindent}{15pt}

\subsection*{Associate Editor}
\them
I enjoyed reading this paper, as did the two reviewers, and I would argue that there is the potential to make a large contribution to the literature on this new way of looking at trophic webs. I found the argument for using motifs much more compelling than previous generations of methods in this area.  The major point for improvement however is to make this manuscript more approachable by and useful to empiricists.

\us
bla

\them 
The following highlights key areas for improvement: first, as reviewer 1 argues, the text can be clarified and made more direct towards the specific goals of the paper, as well as making some of the terminology more transparent.  Second, as reviewer 2 argues there is scope for including a figure about the motifs in the context of their food web to facilitate the understanding about their structure and ecological meaning.  The top of Figure 1 helps, but there is room to expand.  Last, the authors generate what they argue are "a range of plausible network architectures".  But there is never any comparison of these architecture to empirical networks.  Nor is there any calculation of the key motif parameters on empirical data (for example the x-axis of figure 1). It will expand the audience for this manuscript greatly to include some empirical data to help empiricists understand for example in well-studied systems, what is the proportion of these different motifs.  This comparison does not need to be exhaustive, but it will serve two main roles in the manuscript: (1) provide support for the assertion that the authors have generated "a range of plausible network architectures" and (2) allow empiricists to understand how these simulations relate to real ecosystems.

\subsection*{Referee: 1}

Comments to the Author(s)
Thank you for the opportunity to review this paper. I found it very interesting and appreciated the comparison between motifs and more “classic” network metrics. I have a couple major concerns which I do not anticipate being that difficult to address, and some smaller comments.
First, I think the introduction could be clearer. I think it spent too much time discussing bottom up effects, which I appreciate because that is one of the major limitations to this approach, however, I think some of the text could be better spent setting up some of the additional questions of the paper and clarifying various terminology (e.g. local, meso, and global scale gets thrown about, but is a bit of a surprise).

Second, I found the analysis, particularly the divvying up into 5 parts rather circular. We know that connectance and species richness has a large impact on both motif expression and on persistence. Instead of multiple separate analyses, I think these can all be included in one large model. Moreover, I would be interested in an analysis which pits trophic level and in-degree against the motifs. In particular, collecting data on trophic level is much easier than collecting entire food web data for motif roles.

Third, I realize that this technique is already published on, however, some of the limitations of the technique may impact the utility of the results. I think the study would be greatly enhanced by a smaller comparison of a more traditional dynamic food web model approach and this approach – even just placed in the appendix. I note that the authors rely quite heavily on a paper on BioRxivs which has done something similar, so anticipate this wouldn’t be that difficult, but would be very convincing.

Finally, I appreciate that the authors have used similar methods in a paper on BioRxiv – however, the current study relies quite heavily on it in the methodology. To my knowledge this paper is not peer-reviewed (or at least not yet) and thus I am less inclined to accept these citations as appropriate rationale for various decisions.

Smaller comments

Line 65 – 70: I agree that in-degree and trophic level may be coarse description of a species’ role within its food-web. However, the examples provided do not convince me that species’ roles based on motifs are a more nuanced way of describing a species’ role within its food web since the two examples provided refer to the influence disturbance may have on a basal resource – an influence that I’m not convinced a species’ role based on motifs will impact.  
I do think that species’ roles based on motifs are a more nuanced way of describing species’ roles, and I am interested in the central question of this paper – however, this rationale would not convince me if I was unfamiliar with motifs/species roles/finer scale network descriptors.  As this is a central tenet of the paper I think the argument needs to be constructed better. Furthermore, just because a metric is coarse does not mean it is not an effective predictor – and when applying theoretical/computational results to empirical systems sometimes coarse is better. It is easier to measure the trophic position of a species using stable isotopes and approximate extinction probabilities than it is to construct an entire food web.

Line 82-84: This sentence doesn’t quite make sense to me.

Line 86: is this “focal” species the consumer mentioned in line 85?

Line 86: the impact of the motif profile of the full food web on the consumer’s extinction risk has come as a bit of a surprise to me. I think this could have been introduced better in the introduction – not extensively because I think it is probably a secondary question, but at this point I have no hypotheses for why I would even expect the motif profile of the full food web to out perform the motifs in which a focal species participates.

Line 87 – 89: similar to comment above – I think the idea of global and local food web properties has not been introduced at all, nor the classic food web properties (which ones are being measured? Only the four specific motifs thought to contribute to stability are mentioned).

Line 92 – 95: this seems to be a flaw to the Bayesian network simulations – we know that mutual predation happens in food webs – especially if the food web isn’t explicitly size structured, and when nested within the broader food web context these feedback loops may have stabilizing, or destabilizing effects.

Line 149 – 152: these seem like large limitations to me.

Line 164 – 166: Does this normalization control for differences in connectance? It seems like the more connected the networks are the higher degree species will have, and then connectance (definitively known to impact species persistence) will be a major driver

Line 182: what is network mean persistence? Species persistence is the probability of species persistence referred to in the methods? Be explicit. Also, what are the “other network properties”?

Line 192: the idea of disturbance level came as a surprise – a careful reading of the methods pinpointed where you discuss this, however it is not discussed in the statistical analysis part, or in the main objectives of the paper at the end of the introduction.

Line 193: Does it make sense to include Sn and Cn as random intercepts if it is known that both of these values impact 1) the number of motifs (which I don’t think has been rigorously controlled for – see above) and 2) network persistence? (same with eq’n S3)

Line 197 – 199: This sounds more statistically rigorous here than it does in the appendix where these are just visually compared.

Line 201: I have no idea what local and meso-scale refer to. Likewise global.

Line 200-206: The statistical analyses of the various components of this study seem very circular to me. Various local metrics depend on global metrics but treated differently in each step. There must be a more appropriate statistical technique for considering all of these moving parts in a comprehensive manner – even using model selection to compare.

Line 214: Why mean, and why not just that network persistence?

Line 247-254: This would all be greatly enhanced if there were some error bars on the figures.

Line 259-262: This result seems odd to me. Is this just because all of the potential apparent competition interactions are filled first, and really is more of an artefact of the constraints about the types of motifs that are possible within these networks? And perhaps the connectances being explored are unrealistic given these constraints?

Line 289: Odd use of direct competition when most other papers use exploitative competition

Line 334: in-degree and trophic level are not sufficient to explain the trends, yet there is a strong negative correlation between apparent competition and in-degree. This seems contradictory.
SI references incomplete (search (?))

\subsection*{Referee: 2}

Comments to the Author(s)
The authors present an interesting study about how motif based measures can improve our understanding of species persistence in food webs after bottom up disturbances. The idea of this study is original and the authors made a great effort to explain a novel methodology, such as motifs, in a clear way. In addition, I consider the exploration of relationships among level of disturbances, network structural properties, species level indices and species persistence very important and necessary to understand food webs stability and to help future authors to select the most appropriate variables according to their goals.
I found the introduction very easy to follow with an appropriate background supporting the goals of the study allowing to link concepts such as secondary extinction, network structure and motifs. Perhaps authors might reduce the number of paragraphs of the introduction by simply joining some of them. For example, I would suggest to join the second and third paragraph since both refer to the approaches for studying secondary extinctions. Similarly, I would suggest to join the fourth and fifth paragraph because both explain how the structure of different scales (whole-network and species level) affect the response to disturbances.
Furthermore, the methods and statistical analysis were consistent with the proposed goals and very well written. I would like to highlight the correct use of motifs that authors made in the ms, which allows them to deduce new ecological information in addition to common network indices. Perhaps authors might include a figure with the used motifs in the appendix to facilitate the understanding about their structure and ecological meaning. Regarding the results and discussion section, I found them very well written and fluent. However, because of the great number of results, I would suggest keeping consistency between the subtitles in the section methods and results referring to the same goal.

Minor revisions

Line 16: I suggest changing the word “vulnerability” to “species persistence” to keep consistency along the manuscript.

Line 23-26: this sentence could be clearer. Perhaps authors could rephrase to: “The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole (Proulx et al., 2010; Scherber et al., 2010, Rosenblatt & Schmitz, 2016). For example, the loss of diversity…”

Line 57-59: This sentence seems counterintuitive to me. How could more connected networks be more resistant to secondary extinctions than less connected ones? A higher connectance indicates more pathways available where the energy flows, promoting the propagation of a disturbance (Vieira & Almeida-Neto, 2015). Please justify your statement.
Vieira, M. C., & Almeida‐Neto, M. (2015). A simple stochastic model for complex coextinctions in mutualistic networks: robustness decreases with connectance. Ecology Letters, 18(2), 144-152.

Line 61: Please add new cites to sustain the statement. For example:
Gilarranz, L. J., Rayfield, B., Liñán-Cembrano, G., Bascompte, J., & Gonzalez, A. (2017). Effects of network modularity on the spread of perturbation impact in experimental metapopulations. Science, 357(6347), 199-201.

Line 112: Please add “, respectively” after “network”.

Line 161: Perhaps authors might add a figure of the four three-species motifs in the appendix explaining the ecological meaning of each structure. I think this could help readers to understand motifs.

Line 187: Please add “Species” before “persistence” in the title to improve clarity.

Line 200: Please add “Motif” before “participation” in the sentence.

Line 207: Add “Network” before “persistence” in the title.

Line 240: Please keep consistency between the subtitles in the section methods and results referring to the same goal.

Line 270-272: Interesting. This could be related to the connectance of the species. Consumers with higher number of prey might have lower persistence at higher levels of disturbances because the disturbance has more pathways to propagate and reach the consumer.

Line 330-331: Unclear the expression “, and interactions among,”. Please rephrase.

Line 357-358: Unclear the expression “than directly”. Please explain.

Line 376: Please remove the italic formatting of the word “and”.

Figure 5: Please rephrase the first sentence to: “The proportions of the four motifs in a network's motif profile were related to network size, connectance, and their interaction.”

Appendix:

Page 3: Please remove the expression “(?)” from the sentence in the third paragraph.

Page 7: Please remove the expression “(?)” from the last sentence.

Page 10: Please remove the expression “(?)” from the first paragraph of the section S3.3.

\end{document}

### 2022-02-14 manuscript/cover_letter.tex
Updates from Overleaf

@@
 
-\date{October XX, 2021}
+\date{October 21, 2021}
 
@@
 
-Dr.~XX\\
+Prof.~Barrett\\
 Editor-in-Chief\\
@@
 
-\opening{Dear Dr.~XX,}
+\opening{Dear Prof.~Barrett,}
 
-We invite you to consider the enclosed ``Does motif participation provide unique information about species risk of extinction?'' by Anna {\AA}kesson et al., for publication in \emph{Proceedings of the Royal Society B} as an Article.
+We invite you to consider the enclosed ``Motif participation provides unique information about species risk of extinction'' by Anna {\AA}kesson et al., for publication in \emph{Proceedings of the Royal Society B} as an Article.
 Dr.
@@
 That manuscript is currently in revisions elsewhere, but we believe that this manuscript also would be a good topical fit.
-%Both manuscripts deal with the relationship between species' motif roles and risk of extinction, although in different modelling frameworks and using different types of disturbance.
 

### 2022-02-14 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 
-\title{Do motif roles provide unique information about species risk of extinction?\\ \medskip Supplementary Information}
+\title{Motif roles provides unique information about species risk of extinction\\ \medskip Supplementary Information}
 
@@
 
-    % Updated with new random effect
+    % Updated with new random effect, converted to data-scale
     \begin{table}[h!]
@@
         \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as random effects to account for differences in persistence across levels of network size, connectance, and network ID  (Equation~\ref{propreq}).
-Predictors were centered and scaled before model fitting.
+Predictors were centered and scaled before model fitting; coefficients refer to a unit increase in each parameter.
 We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
@@
         \hline
-        3-sp Chain & 0.574 & \textless0.001 & -0.00492 & \textless0.001 & -0.139 & \textless0.001 & -0.00337 & \textless0.001 & 0.900 & 0.912 \\
-        App. Com. & 0.574 & \textless0.001 & -0.00390 & \textless0.001 & -0.139 & \textless0.001 & 0.00410 & \textless0.001 & 0.900 & 0.911 \\
-        Dir. Com. & 0.574 & \textless0.001 & 0.00607 & \textless0.001 & -0.139 & \textless0.001 & 0.000256 & \textless0.001 & 0.901 & 0.911 \\
-        Omnivory & 0.575 & \textless0.001 & 0.00806 & \textless0.001 & -0.139 & \textless0.001 & -0.00198 & \textless0.001 & 0.894 & 0.912 \\
+        3-sp Chain & 0.578 & \textless0.001 & -0.00360 & \textless0.001 & -0.0154 & \textless0.001 & 8.39$\times10^{-5}$& \textless0.001 & 0.866 & 0.914 \\
+        App. Com. & 0.579 & \textless0.001 & -0.00239 & \textless0.001 & -0.0154 & \textless0.001 & -2.70$\times10^{-5}$ & \textless0.001 & 0.822 & 0.882 \\
+        Dir. Com. & 0.577 & \textless0.001 & 7.74$\times10^{-4}$ & \textless0.001 & -0.0154 & \textless0.001 & -3.76$\times10^{-5}$ & \textless0.001 & 0.833 & 0.866 \\
+        Omnivory & 0.577 & \textless0.001 & -0.00146 & \textless0.001 & -0.0154 & \textless0.001 & 4.22$\times10^{-5}$ & \textless0.001 & 0.842 & 0.877 \\
         \end{tabular}

### 2022-02-23 manuscript/response_PRSB_round2.tex

--- 2022-02-14 manuscript/response_PRSB_round1.tex
@@
 
-\title{RESPONSE LETTER \\Species motif participation provides unique information about species risk of extinction}
+\begin{document}
 
-\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{2}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{1}$} 
-\date{\small$^1$Department of Theoretical Biology, Chemistry, and Physics\\ 
-Link\"{o}ping University\\
-Link\"{o}ping, Sweden\\
-\medskip
-\small$^2$Department of Agricultural Sciences\\
-University of Helsinki\\
-Helsinki, Finland\\
-\medskip
-\small$^3$ BioFrontiers Institute\\
-University of Colorado, Boulder\\
-U.S.A.
-\medskip
-$^\dagger$ Corresponding author:\\
-}
+\section*{Response to reviewer comments}
 
-\begin{document} 
-\maketitle 
-\linenumbers
-\raggedright
-
-\setlength{\parindent}{15pt}
+\begin{flushleft}
+\textbf{Manuscript number: } \texttt{RSPB-2021-2331} \\
+``Species motif participation provides unique information about species risk of extinction''\\
+Anna {\AA}kesson, Alyssa R. Cirtwill, Kate L. Wootton \& Anna Ekl\"of
+\end{flushleft}
 
@@
 
-\subsection*{Referee: 1}
+\subsection*{Referee 1}
 
@@
 I have a couple major concerns which I do not anticipate being that difficult to address, and some smaller comments.
+
+\us
+We thank the reviewer for the thorough engagement with our work, the encouraging words, and the constructive criticisms.
+
+\them
 First, I think the introduction could be clearer.
@@
 
+\us
+
+\them
 Second, I found the analysis, particularly the divvying up into 5 parts rather circular.
@@
 
+\us
+
+\them
 Third, I realize that this technique is already published on, however, some of the limitations of the technique may impact the utility of the results.
@@
 
+\us
+
+\them
 Finally, I appreciate that the authors have used similar methods in a paper on BioRxiv – however, the current study relies quite heavily on it in the methodology.
@@
 
+\us
+
+\them
 Smaller comments
@@
 
+\us
+
+\them
 Line 82-84: This sentence doesn’t quite make sense to me.
 
+\us
+
+\them
 Line 86: is this “focal” species the consumer mentioned in line 85?
 
+\us
+
+\them
 Line 86: the impact of the motif profile of the full food web on the consumer’s extinction risk has come as a bit of a surprise to me.
@@
 
+\us
+
+\them
 Line 87 – 89: similar to comment above – I think the idea of global and local food web properties has not been introduced at all, nor the classic food web properties (which ones are being measured?
@@
 
+\us
+
+\them
 Line 92 – 95: this seems to be a flaw to the Bayesian network simulations – we know that mutual predation happens in food webs – especially if the food web isn’t explicitly size structured, and when nested within the broader food web context these feedback loops may have stabilizing, or destabilizing effects.
 
+\us
+
+\them
 Line 149 – 152: these seem like large limitations to me.
 
+\us
+
+\them
 Line 164 – 166: Does this normalization control for differences in connectance?
@@
 
+\us
+
+\them
 Line 182: what is network mean persistence? Species persistence is the probability of species persistence referred to in the methods? Be explicit. Also, what are the “other network properties”?
 
+\us
+
+\them
 Line 192: the idea of disturbance level came as a surprise – a careful reading of the methods pinpointed where you discuss this, however it is not discussed in the statistical analysis part, or in the main objectives of the paper at the end of the introduction.
 
+\us
+
+\them
 Line 193: Does it make sense to include Sn and Cn as random intercepts if it is known that both of these values impact 1) the number of motifs (which I don’t think has been rigorously controlled for – see above) and 2) network persistence?
@@
 
+\us
+
+\them
 Line 197 – 199: This sounds more statistically rigorous here than it does in the appendix where these are just visually compared.
 
+\us
+
+\them
 Line 201: I have no idea what local and meso-scale refer to. Likewise global.
 
+\us
+
+\them
 Line 200-206: The statistical analyses of the various components of this study seem very circular to me.
@@
 
+\us
+
+\them
 Line 214: Why mean, and why not just that network persistence?
 
+\us
+
+\them
 Line 247-254: This would all be greatly enhanced if there were some error bars on the figures.
 
+\us
+
+\them
 Line 259-262: This result seems odd to me.
@@
 
+\us
+
+\them
 Line 289: Odd use of direct competition when most other papers use exploitative competition
 
+\us
+
+\them
 Line 334: in-degree and trophic level are not sufficient to explain the trends, yet there is a strong negative correlation between apparent competition and in-degree. This seems contradictory.
@@
 
-\subsection*{Referee: 2}
+\subsection*{Referee 2}
 
@@
 In addition, I consider the exploration of relationships among level of disturbances, network structural properties, species level indices and species persistence very important and necessary to understand food webs stability and to help future authors to select the most appropriate variables according to their goals.
+
+\us
+We thank the Reviewer for the thorough reading of our manuscript and the helpful comments.
+
+\them
 I found the introduction very easy to follow with an appropriate background supporting the goals of the study allowing to link concepts such as secondary extinction, network structure and motifs.
@@
 Similarly, I would suggest to join the fourth and fifth paragraph because both explain how the structure of different scales (whole-network and species level) affect the response to disturbances.
+
+\us
+
+\them
 Furthermore, the methods and statistical analysis were consistent with the proposed goals and very well written.
@@
 
+\us
+
+\them
 Minor revisions
@@
 
+\us
+
+\them
 Line 23-26: this sentence could be clearer.
@@
 
+\us
+
+\them
 Line 57-59: This sentence seems counterintuitive to me.
 How could more connected networks be more resistant to secondary extinctions than less connected ones?
-A higher connectance indicates more pathways available where the energy flows, promoting the propagation of a disturbance (Vieira & Almeida-Neto, 2015).
+A higher connectance indicates more pathways available where the energy flows, promoting the propagation of a disturbance (Vieira \& Almeida-Neto, 2015).
 Please justify your statement.
-Vieira, M. C., & Almeida‐Neto, M. (2015). A simple stochastic model for complex coextinctions in mutualistic networks: robustness decreases with connectance. Ecology Letters, 18(2), 144-152.
+Vieira, M. C., \& Almeida‐Neto, M. (2015). A simple stochastic model for complex coextinctions in mutualistic networks: robustness decreases with connectance. Ecology Letters, 18(2), 144-152.
 
+\us
+
+\them
 Line 61: Please add new cites to sustain the statement. For example:
 Gilarranz, L.
-J., Rayfield, B., Liñán-Cembrano, G., Bascompte, J., & Gonzalez, A.
+J., Rayfield, B., Liñán-Cembrano, G., Bascompte, J., \& Gonzalez, A.
 (2017).
@@
 
+\us
+
+\them
 Line 112: Please add “, respectively” after “network”.
 
+\us
+
+\them
 Line 161: Perhaps authors might add a figure of the four three-species motifs in the appendix explaining the ecological meaning of each structure. I think this could help readers to understand motifs.
 
+\us
+
+\them
 Line 187: Please add “Species” before “persistence” in the title to improve clarity.
 
+\us
+
+\them
 Line 200: Please add “Motif” before “participation” in the sentence.
 
+\us
+
+\them
 Line 207: Add “Network” before “persistence” in the title.
 
+\us
+
+\them
 Line 240: Please keep consistency between the subtitles in the section methods and results referring to the same goal.
 
+\us
+
+\them
 Line 270-272: Interesting.
@@
 
+\us
+
+\them
 Line 330-331: Unclear the expression “, and interactions among,”. Please rephrase.
 
+\us
+
+\them
 Line 357-358: Unclear the expression “than directly”. Please explain.
 
+\us
+
+\them
 Line 376: Please remove the italic formatting of the word “and”.
 
+\us
+
+\them
 Figure 5: Please rephrase the first sentence to: “The proportions of the four motifs in a network's motif profile were related to network size, connectance, and their interaction.”
 
+\us
+
+\them
 Appendix:
@@
 Page 10: Please remove the expression “(?)” from the first paragraph of the section S3.3.
+\us
 

### 2022-02-24 manuscript/response_PRSB_round2.tex

@@
 \us
-bla
+thanks etc etc
 
@@
 The following highlights key areas for improvement: first, as reviewer 1 argues, the text can be clarified and made more direct towards the specific goals of the paper, as well as making some of the terminology more transparent.
+
+\us
+
+\them
 Second, as reviewer 2 argues there is scope for including a figure about the motifs in the context of their food web to facilitate the understanding about their structure and ecological meaning.
 The top of Figure 1 helps, but there is room to expand.
+
+\us
+Anna E
+
+\them
 Last, the authors generate what they argue are "a range of plausible network architectures".
@@
 This comparison does not need to be exhaustive, but it will serve two main roles in the manuscript: (1) provide support for the assertion that the authors have generated "a range of plausible network architectures" and (2) allow empiricists to understand how these simulations relate to real ecosystems.
+
+\us
+Alyssa
+Try with Serengeti as well?
+Webs referenced in the niche model vs empirical motifs paper
+
+The niche model has been shown in previous work to simulate empirical networks reasonably well~\citep{Stouffer2005a,Stouffer2006}.
+The specific method we use to simulate networks, developed in~\citet{Delmas2017}, has been used in several published studies already (e.g.,~\citep{})
 
@@
 \us
+Anna E
+- add point that once we have the network, we have the motif profiles and roles so might as well use them and not just degree/TL.
 
@@
 Instead of multiple separate analyses, I think these can all be included in one large model.
+
+\us
+a written response here (and something brief in the paper)
+
+\them
 Moreover, I would be interested in an analysis which pits trophic level and in-degree against the motifs.
@@
 \us
+Additional analysis - maybe use motifs to explain differences between species with the same degree, trophic level?
+if you have TL and degree you also have the motif profile of a species (disagree that trophic level is "easier" - once the network is characterised all three role types are available) put in the intro as well
 
@@
 
-\us
+\us Alyssa - written response that including a whole second paper is beyond the scope and that paper is under review
 
@@
 
-\us
+\us Alyssa - change citations to original papers, not BiorXiv
 
@@
 
-\us
+\us Degree is better than TL and then you still need the whole web - mention limits of stable isotopes (stable isotope guy maybe?) ... Address this in the new figure.
 
@@
 
-\us
+\us revise, add references to Allesina lab papers, don't end intro on limitation. Anna E
 
@@
 
-\us
+\us tone down discussion of limitations a bit, deal more in Discussion. Anna E
 
@@
 
-\us
+\us not explicitly, control for the effect of C as a random effect. Most interested in comparing species within the same web. Alyssa
 
@@
 \us
-
+Add explanations.
 \them
@@
 
-\us
+\us Mention disturbance earlier - intro if possible
 
@@
 
-\us
+\us Frequency normalisation explicitly controls for total number of motifs, removing that potential (not proven to our knowledge) effect of S and C.
+To our knowledge, we are the first to explore whether S and C affect motif profiles ...
+We have them as random intercepts to account for 2 since this is indeed a known effect and we aren't interested in re-inventing the wheel
 
@@
 
-\us
+\us Probably move this line to appendix
 
@@
 
-\us
+\us Add a glossary in the Appendix for those who forget the introduction
 
@@
 
-\us
+\us See response to XX ... probably need to repeat that our focus is not to throw away degree
 
@@
 
-\us
+\us See response to XX. Add to glossary.
 
@@
 
-\us
+\us Annoying but maybe possible
 
@@
 
-\us
+\us Empirical comparison should take care of this. ... citations to connectance range papers.
 
@@
 
-\us
+\us All other motif papers use direct competition ... we do not say whether it's exploitative or interference...
 
@@
 Line 334: in-degree and trophic level are not sufficient to explain the trends, yet there is a strong negative correlation between apparent competition and in-degree. This seems contradictory.
+
+\us Clarify our writing ... is there a way to connect the figures better?
+
+\them
 SI references incomplete (search (?))
+
+\us
 
@@
 
-\us
+\us TBD after answer to reviewer 1
 
@@
 Perhaps authors might include a figure with the used motifs in the appendix to facilitate the understanding about their structure and ecological meaning.
+
+\us Same figure requested by Editor, good idea to add
+
+\them
 Regarding the results and discussion section, I found them very well written and fluent.
@@
 
-\us
+\us Double-check subheadings consistent
 
@@
 
-\us
+\us yep
 
@@
 
-\us
+\us Expand explanation of why high connectance can be good sometimes (more prey available to switch to, albeit more disturbance pathways).
+Probably a good way to introduce size of disturbance - many pathways of a small disturbance may not matter, many pathways of a large more severe
 
@@
 
-\us
+\us Same figure as above, if possible?
 

### 2022-02-28 code/collect_empirical_roles.py
changing to nonlinear simulations

@@ -162,10 +162,15 @@ def main():
   # print 'TLs and degrees read'
 
 
-  linear_TLdict,linear_degdict,linear_perdict=read_persistence_file('../data/all_empirical.tsv',partic_dict)
+  # linear_TLdict,linear_degdict,linear_perdict=read_persistence_file('../data/all_empirical_linear.tsv',partic_dict)
+  # print 'persistence read'
+  # collect_and_print_3sp('../data/empirical_3sp_roles_participation_linear.tsv',role_dict,good_positions,partic_dict,good_motifs,linear_degdict,linear_TLdict,linear_perdict)
+  # print '3sp role files complete'
+
+  nonlinear_TLdict,nonlinear_degdict,nonlinear_perdict=read_persistence_file('../data/all_empirical_nonlinear.tsv',partic_dict)
   print 'persistence read'
-  collect_and_print_3sp('../data/empirical_3sp_roles_participation.tsv',role_dict,good_positions,partic_dict,good_motifs,linear_degdict,linear_TLdict,linear_perdict)
-  print '3sp role files complete'
+  collect_and_print_3sp('../data/empirical_3sp_roles_participation_nonlinear.tsv',role_dict,good_positions,partic_dict,good_motifs,nonlinear_degdict,nonlinear_TLdict,nonlinear_perdict)
+  print 'nonlinear 3sp role files complete'
       
 
 if __name__ == '__main__':

### 2022-02-28 code/stat_analysis/empirical_statistics.R

@@ -11,7 +11,7 @@ library(vegan)
 library(MuMIn)
 
 
-data<-read.table(file = '../../data/empirical_3sp_roles_participation.tsv', sep = '\t', header = TRUE)
+data<-read.table(file = '../../data/empirical_3sp_roles_participation_nonlinear.tsv', sep = '\t', header = TRUE)
 data$Disturbance<-0.1+0.4*data$Basal_p
 # ONLY FIT MODELS ON CONSUMERS!!!
 consumers<-data[which(data$in_Degree>0),]
@@ -83,7 +83,7 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
         interdata2$S[r]=strsplit(levelcode,':')[[1]][1]
         interdata2$C[r]=strsplit(levelcode,':')[[1]][2]
     }
-    write.table(interdata2,file='empirical_proportion_variability_SC.tsv',sep='\t')
+    write.table(interdata2,file='empirical_proportion_variability_SC_nonlinear.tsv',sep='\t')
 
     # LMs - no need for network-level random effect since it's within-network motif profiles
     pchainlm=with(motdata,lm(prop_chain~Size*Connectance))
@@ -136,7 +136,7 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     # SCper=with(consumers,lm(Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)))
     SCper_mean=with(simdata,lm(per~scale(Size)*scale(Connectance)*scale(dist)))
 
-# Section 4: How does persistence vary with proportions?
+# Section 4: How does persistence vary with proportions? Note that global==netID for the empirical webs
     # # lmer with random intercepts for each level of S:C
     consumers$Global=paste0(consumers$Size,':',consumers$Connectance)
     propchain_lmer1<-with(consumers,lmer(Persistence~scale(prop_chain)*scale(Disturbance)+(1|Global)))
@@ -148,16 +148,6 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     R2app=r.squaredGLMM(propapparent_lmer1,null=lmer(consumers$Persistence~(1|Global)))
[…]

### 2022-02-28 manuscript/reorder_main_text.tex

@@
 
-% Maybe Proc B - approached Alyssa re: other motifs paper, flexible length limit (but pay extra over X pages).
-
 \section*{Abstract}
@@
     Additionally, we find that relationships between motif participation and persistence can synthesize the information provided by in-degree and trophic level.
-    Motif participation therefore captures important trends in vulnerability to extinction and provides a rich description of species' structural roles in their communities.
+    Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities.
 
@@
     % basic idea behind BN modeling
-        Bayesian networks can be used to represent species and their feeding interactions as the nodes and links of a network (see Appendix \emph{S2} for full details).
+        Bayesian networks can be used to represent species and their feeding interactions as the nodes and links of a network, respectively (see Appendix \emph{S2} for full details).
         In our setting, each species (node) is a Bernoulli random variable, being either extant or extinct. 
@@
 
-    \subsubsection*{1) Persistence and motif participation}
+    \subsubsection*{1) Species persistence and motif participation}
 
@@
 
-    \subsubsection*{2) Participation and other network properties}    
+    \subsubsection*{2) Motif participation and other network properties}    
         Different scales of network structure (e.g., local and meso-scale) are not independent. 
@@
 
-    \subsubsection*{3) Persistence and motif profiles}
+    \subsubsection*{3) Network persistence and motif profiles}
 
@@
     % High C webs and omnivory
-    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks \emph{and} the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}) as in~\citep{McLeod2021}.
+    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks and the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}) as in~\citep{McLeod2021}.
     These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
@@
         \includegraphics[width=\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
-        \caption{The proportions of the four motifs in a network's motif profile were network size, connectance, and their interaction.
+        \caption{The proportions of the four motifs in a network's motif profile were related to network size, connectance, and their interaction.
 The effect of connectance was much stronger than the main effect or interaction with network size.

### 2022-03-01 code/plot_motif_profile_comparison.py
comparing motif profiles in empirical webs

@@ -1,0 +1,156 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Spectral', reverse=True,n=180)  # The blue is very beautiful but maybe harder to see.
+colors.add_color(173,132,191,'Purple')
+colors.add_color(34,90,34,'Green')
+# colors.add_color(120,120,120,'grey')
+# colors.add_color(255,125,125,'lightish_red')
+# colors.add_color(200,200,200,'lightgrey')
+
+# Want to use AC, DC as x's, chain, omni as y's
+
+def read_motif_profiles(motfile):
+  profiles={'AC':{'chain':{},'omnivory':{}},'DC':{'chain':{},'omnivory':{}}}
+  netinfo={}
+  f=open(motfile,'r')
+  for line in f:
+    if line.split()[0]!='Size':
+      S=int(line.split()[0])
+      C=float(line.split()[1])
+      ID=line.split()[2]
+      AC=int(line.split()[3])
+      chain=int(line.split()[4])
+      DC=int(line.split()[5])
+      omnivory=int(line.split()[6])
+      total=int(line.split()[7])
+      profiles['AC']['chain'][ID]=[((float(AC)/float(total),float(chain)/float(total)))]
+      profiles['AC']['omnivory'][ID]=[((float(AC)/float(total),float(omnivory)/float(total)))]
+      profiles['DC']['chain'][ID]=[((float(DC)/float(total),float(chain)/float(total)))]
+      profiles['DC']['omnivory'][ID]=[((float(DC)/float(total),float(omnivory)/float(total)))]
+      netinfo[ID]=((S,C))
+
+  f.close()
+
+  return profiles,netinfo
+
[…]

### 2022-03-01 code/plot_motif_role_comparison.py

@@ -1,0 +1,139 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Spectral', reverse=True,n=180)  # The blue is very beautiful but maybe harder to see.
+colors.add_color(173,132,191,'Purple')
+colors.add_color(34,90,34,'Green')
+# colors.add_color(120,120,120,'grey')
+# colors.add_color(255,125,125,'lightish_red')
+# colors.add_color(200,200,200,'lightgrey')
+
+# Want to use AC, DC as x's, chain, omni as y's
+
+def read_motif_profiles(motfile):
+  profiles={'AC':{'chain':{},'omnivory':{}},'DC':{'chain':{},'omnivory':{}}}
+  netinfo={}
+  f=open(motfile,'r')
+  for line in f:
+    if line.split()[0]!='Size':
+      S=int(line.split()[0])
+      C=float(line.split()[1])
+      ID=line.split()[2]
+      AC=int(line.split()[3])
+      chain=int(line.split()[4])
+      DC=int(line.split()[5])
+      omnivory=int(line.split()[6])
+      total=int(line.split()[7])
+      profiles['AC']['chain'][ID]=[((float(AC)/float(total),float(chain)/float(total)))]
+      profiles['AC']['omnivory'][ID]=[((float(AC)/float(total),float(omnivory)/float(total)))]
+      profiles['DC']['chain'][ID]=[((float(DC)/float(total),float(chain)/float(total)))]
+      profiles['DC']['omnivory'][ID]=[((float(DC)/float(total),float(omnivory)/float(total)))]
+      netinfo[ID]=((S,C))
+
+  f.close()
+
+  return profiles,netinfo
+
[…]

### 2022-03-01 manuscript/supplement_motifs.tex

@@
     While niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
-    We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009}.
+    We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009functional}.
     This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009functional}.
@@
 
-\bibliography{anna_bib} % Abbreviate journal titles.
+\bibliography{anna_bib_new} % Abbreviate journal titles.
 \bibliographystyle{ecollett}

### 2022-03-02 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
     The global (whole-network) structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002, Eklof2006}.
-    For example, more highly-connected networks are more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{Krause2003, Teng2004}.
+    For example, more highly-connected networks are more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{Krause2003, Teng2004, Gilarranz2017}.
 
@@
     Highly-connected networks tended to contain higher frequencies of the omnivory motif, which includes more links than the other three-species motifs considered here.
-    Global-scale properties may thus  have stronger effects on network persistence by changing meso-scale properties than directly.
+    Global-scale properties may therefore indirectly affect network persistence by changing meso-scale properties, such as the distribution of motifs.
 

### 2022-03-02 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
             \end{itemize}
+
+        \item Glossary    
     \end{enumerate}    
@@
             \rho_{in} \approx \tau_{i} + S_{n}C_{n} + N_n,
-            \label{partic_deg}
+            \label{partic_tl}
         \end{equation}
@@
         To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
-        As with the PERMANOVA relating motif participation to network properties, we fit the PERMANOVA using the R~\citep{R} function `vegan' from the package \emph{adonis}~\citep{adonis} and calculated significance based on 9999 permutations.
+        As with the PERMANOVA relating motif participation to network properties, we fit the PERMANOVA using the R~\citep{R} function `vegan' from the package \emph{adonis} and calculated significance based on 9999 permutations.
 
@@
         \end{tabular}
+    \end{table}
+
+\clearpage
+
+\section{Glossary}
+
+ \begin{table}[h!]
+ \label{glossary}
+ \caption{Glossary of terms relating to motifs and Bayesian networks}
+     \footnotesize{
+ \begin{tabular}{l|l}
+     Term & Definition \\
+     \hline
+     Motif & Set of $n$ interacting species. In this case, $n=3$ \\
+     Network motif profile & Vector describing the frequency of each motif in the network. \\
+     & Normalised by dividing counts of each motif by the total across all motifs.\\
+     Motif participation role & Vector describing the frequency with which a focal species appears in each motif.\\
+     & Normalised by dividing counts for each motif by the total across all motifs. \\
+     Bayesian network & A directed acyclic graph, used to predict species' likelihood of persistence. \\
+     Network persistence & The mean likelihood of consumers in a network not going extinct.\\
+     Species persistence & The individual likelihood of a species not going extinct.\\
+     In-degree & Number of prey species to a consumer species.\\
+     Trophic level (STL) & The shortest food chain between the focal species and any basal species.\\
+     Disturbance ($\pi_{disturbed}$) & Probability of extinction of a basal resource when extra disturbance is added. \\
+     Baseline extinction &  When no disturbance is applied, $\pi_{base}$=0.1.\\
+ \end{tabular}}
  \end{table}

### 2022-03-02 manuscript/supplement_motifs.tex
added details of network generation

--- 2022-03-01 manuscript/supplement_motifs.tex
@@
 
+    \subsection*{Generation of simulated networks}
+
+        We simulated a suite of food webs based on the probabilistic niche model, which assigns predator-prey links based on the body-mass ratios between individuals of different species~\citep{Williams2000,Delmas2017}.
+The meso-scale structure of niche-model networks closely mimics that of empirical food webs~\citep{Stouffer2007}.
+To ensure that we captured a variety of realistic community sizes and structures, we generated networks ranging between 50 and 100 species (in steps of 10) with connectance values between 0.02 and 0.2 (in steps of 0.02).
+The range of network sizes was chosen to reflect moderately well-sampled empirical webs while working within our computational limits, while the range of connectance values was chosen to cover that observed in most empirical food webs~\citep{Dunne2002e}.
+We generated a total of 100 networks with each combination of parameters, for a total of 6000 networks.
+All networks were generated using the function "nichemodel" within the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
+If a simulated network contained any disconnected species (species without predators or prey) or disconnected components (a group of species connected among themselves but not to the rest of the network), the network was rejected and a new network simulated.
+Finally, networks where the path lengths between each species and a basal resource could not be resolved (i.e., trophic levels were undefined) were rejected and new networks simulated.
+
+        After generating the network structure, we simulated community dynamics using the function ``simulate'' from the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
+This function uses Lotka-Volterra predator-prey models including density dependence and type 2 functional responses for all species (please see~\citet{Delmas2017} for full details).
+        All non-basal species were designated as vertebrates to ensure a good match between metabolic and predator-prey body-mass ratio values.
+Metabolic rates in the Lotka-Volterra model are based on each species' body mass (i.e.
+mass of a single individual).
+We assigned relative body masses based on each species' trophic level, which was, in turn, calculated based on the food-web structure provided by the niche model.
+After basal species were assigned a body mass of 1, we used a predator-prey body-mass ratio of 3.065 to calculate the relative body masses of higher trophic levels.
+We selected this ratio based on the estimate for vertebrates (averaged across ecosystem and metabolic types) in~\citet{Brose2006}.
+We excluded reported body-mass ratios for invertebrates as these could include parasites and parasitoids, which are generally smaller than their prey, and because interactions among vertebrates are better represented in the food-web literature than interactions involving invertebrates.
+
+        The persistence of each species in our simulated networks also depends on its population biomass. 
+        We randomly assigned initial population biomasses (i.e.
+cumulative biomass across all individuals of a species) for each species from a uniform distribution [0,1].
+Note that population biomasses and individual body masses are not calculated on the same scale.
+We then simulated community dynamics for 1000 time steps to obtain an `equilibrium' community.
+To ensure that species did not `recover' from unrealistically low biomasses during the simulation, we considered a species extinct if it dropped below an arbitrary threshold biomass of 1$\times10^{-5}$.
+When simulating initial (i.e.
+pre-disturbance or equilibrium) dynamics, we rejected any network where one or more species dropped below this biomass threshold.
+Consumers were assumed to have no preferences such that the consumption rate $w_{ij}$ of predator $i$ eating prey $j$ is equal to $\frac{1}{n}$, where $n$ is the number of prey for predator $i$.
+If the network did not retain all species for 1000 time steps, a new set of initial population biomasses was applied and the simulation repeated.
+        If a set of biomasses where all species persist still had not been reached after 100 sets of randomly-assigned initial biomasses, we discarded the network and simulated another to replace it.
+
+    \subsection*{Filtering of biologically unrealistic networks}
+
         After simulating our initial niche-model networks, we filtered networks for biological plausibility.
@@
         Any removed network was replaced with a fresh simulation and re-checked until 100 realistic networks for each size-connectance combination had been obtained.
+
+    \subsection*{Conversion of networks to acyclic form}
 

### 2022-03-02 manuscript/supplement_motifs.tex

@@
 
+\section{Glossary}
+
+ \begin{table}[h!]
+ \label{glossary}
+ \caption{Glossary of terms relating to motifs and Bayesian networks}
+     \footnotesize{
+ \begin{tabular}{l|l}
+     Term & Definition \\
+     \hline
+     Motif & Set of $n$ interacting species. In this case, $n=3$ \\
+     Network motif profile & Vector describing the frequency of each motif in the network. \\
+     & Normalised by dividing counts of each motif by the total across all motifs.\\
+     Motif participation role & Vector describing the frequency with which a focal species appears in each motif.\\
+     & Normalised by dividing counts for each motif by the total across all motifs. \\
+     Bayesian network & A directed acyclic graph, used to predict species' likelihood of persistence. \\
+     Network persistence & The mean likelihood of consumers in a network not going extinct.\\
+     Species persistence & The individual likelihood of a species not going extinct.\\
+     In-degree & Number of prey species to a consumer species.\\
+     Trophic level (STL) & The shortest food chain between the focal species and any basal species.\\
+     Disturbance ($\pi_{disturbed}$) & Probability of extinction of a basal resource when extra disturbance is added. \\
+     Baseline extinction &  When no disturbance is applied, $\pi_{base}$=0.1.\\
+ \end{tabular}}
+ \end{table}
+
+\clearpage
+
 % \section{SH: Persistence and motif roles}

### 2022-03-02 manuscript/reorder_main_text.tex

@@
 		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
-		For full details, see~\citet{Cirtwill2021_inprep}.
+		For full details, se \emph{Appendix S1}.
 		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
@@
         As we are interested in relationships between the type of motif participated in rather than the total number of motifs, we normalized participation vectors by dividing each count by the total number of motifs the species appears in.
-        Note, however, that this normalization does \emph{not} control for differences due to degree \emph{per se} as high-degree species also tend to participate in different types of motifs than low-degree species~\citep{Cirtwill2021_inprep}.
+        Note, however, that this normalization does not necessarily control for differences due to degree \emph{per se} as high- and low-degree species may also tend to participate in different types of motifs.
 

### 2022-03-02 manuscript/response_PRSB_round2.tex
Updates from Overleaf

@@
 
-\us Alyssa - written response that including a whole second paper is beyond the scope and that paper is under review
+\us We agree that a comparison between Bayesian and dynamical simulations could be interesting, however performing such a comparison is beyond the scope of this manuscript.
+As the Reviewer notes, a subset of the authors of this manuscript are working on an analysis of the relationships between motif participation and persistence in a dynamical framework (currently under review).
+We do not believe that including the material from this second paper in an appendix (where few people will see it) is justified when there are already several comparisons of extinctions simulated by Bayesian and dynamical approaches (e.g.,~\citealp[]{Eklof2013a,others}).
+
+In addition, we feel that the Reviewer has over-estimated our reliance on the BioRxiv pre-print currently under review.
+Our previous draft had only three references to this pre-print: when describing the generation of simulated networks, when noting that degree and motif participation are not independent, and when briefly comparing the dynamical and Bayesian results in the discussion.
+To address this mis-perception, we have now removed two of these references.
+We present the full network-generation methods in \emph{Appendix 1} and, to avoid confusion by referring to results presented in the preprint, have softened our language to say that degree and motif participation \emph{may} be correlated.
+The latter is not a major point in our manuscript; we mention it only to emphasize that controlling for the total number of motifs in which a species appears does not necessarily control for degree. 
+Thus, we see some overlap in the associations of persistence with degree and with normalised motifs.
+
+We have retained the reference to the pre-print in the discussion because this brief comparison of results using dynamical and Bayesian simulations appears to be exactly what the Reviewer has asked for.
+As the Reviewer points out, different approaches to simulating disturbance can have different results, and we explicit mention the main differences in the Discussion. 
+As the pre-print is freely available to any who wish to critically examine it, we believe that this is a good compromise between maintaining conciseness of the present manuscript and comparing different simulation approaches.
 
@@
 
-\us Alyssa - change citations to original papers, not BiorXiv
+\us See above.
+The only methodological tie between the two manuscripts was when simulating the networks.
+The underlying software used is peer-reviewed~\citep{Delmas2017} and has been used in several other published studies, as well as the pre-print originally referenced (currently under review).
+To avoid the mis-perception that we rely on unproven methods, we now present the full network-generation methods in an Appendix.
 
@@
 
-\us not explicitly, control for the effect of C as a random effect. Most interested in comparing species within the same web. Alyssa
+\us To the extent that species in higher-connectance networks tend to participate in more motifs, this normalization does address connectance, however that is not its main purpose and it has been tested whether normalizing motif roles in this way fully controls for connectance.
+Since this normalization does not remove correlations between degree and motif participation (as discussed in~\citealp[]{Cirtwill2021_inprep}), we are not confident in making such a claim.
+Instead, to control for connectance explicitly, we include it (and species richness) in a random effect in all regressions where it is not a main effect. 
+We take this approach because re-evaluating the known effects of connectance and species richness is not a major theme in the manuscript.
+We are more interested in comparing extinction risk for species within the same food web (where connectance is constant) than between food webs (where connectance, species richness, and --in empirical systems-- local conditions will have large effects).
+We have lightly revised the methods section to make this focus clearer and to state that, while our motif role normalization does not control for degree or connectance, we do explicitly consider how degree and connectance (as well as species richness and trophic level) may be related.
+
+    ``Note, however, that this normalization does not necessarily control for differences due to degree or connectance as high- and low-degree species (or species in more- and less-connected networks) may also tend to participate in different types of motifs.
+    We address the potential associations between motif participation, degree, and connectance explicitly below.''
 
@@
 Add explanations.
+
 \them
@@
 \us
+Done.
 
@@
 
-\us yep
+\us
+Changed.
 
@@
 
-\us yep
+\us Changed as suggested.
 
@@
 
-\us Expand explanation of why high connectance can be good sometimes (more prey available to switch to, albeit more disturbance pathways).
+\us Alyssa: Expand explanation of why high connectance can be good sometimes (more prey available to switch to, albeit more disturbance pathways).
 Probably a good way to introduce size of disturbance - many pathways of a small disturbance may not matter, many pathways of a large more severe
@@
 \us
+Added, thanks for mentioning!
 
@@
 
-\us
+\us We have revised the subtitles to be consistent. Thank you for the suggestion.
 
@@
 
-\us
+\us Need to check the line, but if this is talking about degree then I'm not sure about the difference between degree and connectance at the species level
 
@@
 
-\us
+\us We have revised this line to ``Motifs can thus act as a tool for synthesizing the information provided by other measures of a species' place in its community and for understanding the interactions among these measures.'' and hope that this is now clear.
 
@@
 \us
+Rephrased.
+Here, we meant that although we did not see a strong effect on network mean persistence of global-scale properties (such as connectance and size), we found that, in particular, connectance altered the distribution of motifs which in turn affected the network mean persistence.
 
@@
 \us
+Changed.
 
@@
 \us
+Changed, thanks for catching!
 
@@
 Page 10: Please remove the expression “(?)” from the first paragraph of the section S3.3.
+
 \us
+Done.
 

### 2022-03-02 manuscript/reorder_main_text.tex

@@
 Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-The stability of the plant community is crucial for sustaining a healthy energy flow in the ecosystem as a whole \citep{Rosenblatt2016} and the structure of the plant community affects stability across all levels of ecosystem organization \citep{proulx2010diversity,scherber2010bottom}.
-Loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
+     The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity,scherber2010bottom,Rosenblatt2016}.
+     For example, the loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
 \citep{scherber2010bottom}.
@@
         As we are interested in relationships between the type of motif participated in rather than the total number of motifs, we normalized participation vectors by dividing each count by the total number of motifs the species appears in.
-        Note, however, that this normalization does not necessarily control for differences due to degree \emph{per se} as high- and low-degree species may also tend to participate in different types of motifs.
+        Note, however, that this normalization does not necessarily control for differences due to degree or connectance as high- and low-degree species (or species in more- and less-connected networks) may also tend to participate in different types of motifs.
+        We address the potential associations between motif participation, degree, and connectance explicitly below.
 
@@
 
-    \subsubsection*{3) Network persistence and motif profiles}
+    \subsubsection*{3) Network mean persistence and motif profiles}
 
@@
 
-    \subsection*{Species motif participation and species persistence} 
+    \subsection*{Species persistence and motif participation} 
     % Coefficients changed slightly with network-level random effect, but conclusions similar
@@
 
-    \subsection*{Consistency in effects of motif participation on species persistence}
+        \subsubsection*{Consistency in effects of motif participation on species persistence}
 
@@
 
-    \subsection*{Network motif profiles and network mean persistence}
+    \subsection*{Network mean persistence and motif profiles}
         % Updated to use means instead of all species
@@
 
-    \subsection*{Network motif profiles and global network properties}
+    \subsection*{Motif profiles and network properties}
 
@@
     This is consistent with a strong negative correlation between omnivory and trophic level (leading to the overall positive effect of omnivory) and a strong positive correlation with in-degree (leading to the reduced benefit of omnivory at high levels of disturbance).
-    Motifs can thus act as a tool for synthesizing the information provided by, and interactions among, other measures of a species' place in its community.
+    Motifs can thus act as a tool for synthesizing the information provided by other measures of a species' place in its community and for understanding the interactions among these measures.
 

### 2022-03-02 manuscript/response_PRSB_round2.tex
Updates from Overleaf

@@
 
-\us Add a glossary in the Appendix for those who forget the introduction
+\us These terms are explained at their first use in the introduction or methods. To avoid confusion, we also add a glossary in \emph{Appendix S11} so that readers can easily look up these terms.
 
@@
 
-\us See response to XX. Add to glossary.
+\us We use the term ``network mean persistence'' because this is the mean persistence for all species within the network.
+``Network persistence'' could easily be misunderstood as referring to R50 or some other network-level measure of stability.
+To avoid confusion, we have added the term to the glossary in \emph{Appendix S11}
 
@@
 
-\us All other motif papers use direct competition ... we do not say whether it's exploitative or interference...
+\us We are not aware of any other motif-focused papers calling this motif ``exploitative competition'' and follow the nomenclature from the introduction of motifs to food webs~\citep{Stouffer2007}.
+Note that ``direct competition'' refers simply to two species which consume the same prey and is agnostic as to whether this is exploitative or interference competition.
 
@@
 
-\us Double-check subheadings consistent
+\us We have revised the subheadings to ensure consistency and thank the Reviewer for the reminder.
 
@@
 
-\us Need to check the line, but if this is talking about degree then I'm not sure about the difference between degree and connectance at the species level
+\us We are not sure what the distinction between a species' degree and its connectance might be - we are used to connectance referring to network-level link density. 
+However, we agree that this is an interesting result and had similar thoughts to the Reviewer that the interaction between degree and disturbance might indicate some kind of trade-off between the benefits of having many prey and the drawbacks of being exposed to disturbance through many pathways.
+We had minimized this theme in the Discussion in order to meet word count limits, but have added a brief example in the final paragraph.
+
+% Original line: In-degree shows an even stronger interaction with disturbance; having more prey species is associated with higher persistence without  disturbance, but lower persistence at high levels of disturbance.
 

### 2022-03-02 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
         To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
-        As with the PERMANOVA relating motif participation to network properties, we fit the PERMANOVA using the R~\citep{R} function `vegan' from the package \emph{adonis} and calculated significance based on 9999 permutations.
+        As with the PERMANOVA relating motif participation to network properties, we fit the PERMANOVA using the R~\citep{R} function `adonis' from the package \emph{vegan}~\citep{vegan} and calculated significance based on 9999 permutations.
 
@@
     % Updated with new random effect, converted to data-scale
-    \begin{table}[h!]
+    \begin{table}[ht!]
         \centering
@@
     % Updated with network-level random effect
-    \begin{figure}[h!]
+    \begin{figure}[ht!]
         \centering
@@
     % Updated with network-level random effect
-    \begin{table}[h!]
+    \begin{table}[hb!]
         \caption{Coeffcients ($\beta$) for linear mixed-effect models of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction, as well as a random effect of network ID.
@@
 
-    \begin{figure}[h!]
+    \begin{figure}[hb!]
      \centering
@@
     % Updated to use mean instead of all species
-    \begin{table}[h!]
+    \begin{table}[hb!]
         \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
@@
 
-   \begin{figure}[h!]
+   \begin{figure}[hb!]
        \centering
@@
     % updated to use network means rather than all species
-    \begin{figure}[h!]
+    \begin{figure}[hb!]
         \centering
@@
 
-    \begin{figure}[h!]
+    \begin{figure}[hb!]
       \includegraphics[width=\textwidth]{figures/heatmap_CS_BPcompare.eps}
@@
 
- \begin{table}[h!]
+ \begin{table}[hb!]
  \label{glossary}
@@
      \footnotesize{
- \begin{tabular}{l|l}
+ \begin{tabular}{l|m{11cm}}
      Term & Definition \\
      \hline
+     Global property & A property of the whole network (e.g., species richness) \\
+     Local property & A property of a single species, generally referring only to direct interactions (e.g., degree) \\
+     Meso-scale property &  A property of a single species which includes direct and indirect interactions; the species' immediate `neighbourhood' in the network (e.g., motif participation) \\
      Motif & Set of $n$ interacting species. In this case, $n=3$ \\
@@
      Bayesian network & A directed acyclic graph, used to predict species' likelihood of persistence. \\
-     Network persistence & The mean likelihood of consumers in a network not going extinct.\\
+     Network mean persistence & The mean likelihood of persistence for all consumers in a network.\\
      Species persistence & The individual likelihood of a species not going extinct.\\

### 2022-05-24 manuscript/response_PRSB_round2.tex
Updates from Overleaf

@@
 \us
-Alyssa
-Try with Serengeti as well?
+Alyssa - this week (or bother me why not)
 Webs referenced in the niche model vs empirical motifs paper
@@
 \us
-Anna E
-- add point that once we have the network, we have the motif profiles and roles so might as well use them and not just degree/TL.
+The introduction has been re-written, and is hopefully clearer now regarding terminology and the questions asked. Additionally, we have now included a glossary in \emph{Appendix S11}
 
@@
 
-\us
+\us - Alyssa: S and C important but not main focus, so want to address them and then move on to more detailed things.
 a written response here (and something brief in the paper)
@@
 \us
-Additional analysis - maybe use motifs to explain differences between species with the same degree, trophic level?
+Additional analysis - maybe use motifs to explain differences between species with the same degree, trophic level? (try explanation first)
 if you have TL and degree you also have the motif profile of a species (disagree that trophic level is "easier" - once the network is characterised all three role types are available) put in the intro as well
@@
 
-\us Degree is better than TL and then you still need the whole web - mention limits of stable isotopes (stable isotope guy maybe?) ... Address this in the new figure.
+\us Degree is better than TL and then you still need the whole web - mention limits of stable isotopes (stable isotope guy maybe?) ...
+Address this in the new (conceptual?) figure.
+Show motif profiles that should experience strong and weak effects of bottom-up disturbance
 
@@
 \us
+This sentence was removed when re-writing the introduction.
 
@@
 \us
+Rephrased, yes it is the same species, i.e. do the extinction risk of a consumer depend upon the motifs in which it participates in?
 
@@
 \us
+We have now revised the introduction to introduce the motif profile of the whole network and its potential relevance to species-level extinction risk [[check]] and hope that the purpose of these analyses is now clearer.
+[[edit to be nicer]]We did not expect that the motif profile of a full food web to outperform the impact of a species motif participation, instead, this was done to investigate whether the full profile of a web would give some additional information, in "average" terms.
 
@@
 
-\us
+\us We have revised the introduction to more clearly introduce these concepts.
 
@@
 
-\us revise, add references to Allesina lab papers, don't end intro on limitation. Anna E
+\us This is now re-written, and we describe this limitation of the Bayesian network framework in the Methods section [[might need to make this more explicit in the methods section, it is now mentioned a bit here and there, also add references to Allesina lab papers if we have some additional?]]
 
@@
 \us
-Add explanations.
+Added clarifications to to point 3 and 5 where we had "network mean persistence" and "other network properties", please check if ok.
 
@@
 
-\us Mention disturbance earlier - intro if possible
+\us This concept is now introduced in the introduction [[but should we discuss it a bit more also in the statistical analysis?
+we mention the different disturbance levels in the end of the methods section Calculating persistence]]
 
@@
 
-\us Probably move this line to appendix
+\us We have added the word ``visually'' to the main text.
 
@@
 
-\us These terms are explained at their first use in the introduction or methods. To avoid confusion, we also add a glossary in \emph{Appendix S11} so that readers can easily look up these terms.
+\us These terms are explained at their first use in the introduction or methods.
+To avoid confusion, we also add a glossary in \emph{Appendix S11} so that readers can easily look up these terms.
+We have also revised the paragraph starting on line 201 to include parenthetical definitions and clarify the scale which we are considering.
 

### 2022-05-24 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 \clearpage
-\section*{Introduction}
+\section*{Introduction} %New
 
@@
 Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-     The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity,scherber2010bottom,Rosenblatt2016}.
+     The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
+     For example, the loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
+\citep{scherber2010bottom}.
+
+    %\subsection*{2 Global and local network properties are known to affect stability}
+    How population decline or extinction of one species leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances. 
+%network functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002, Eklof2006}. 
+    Two well-known, influential global properties are network size, the number of species present in the network, and network connectance, the number of realized feeding interactions.
+    Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
+    The presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
+    Highly-connected networks may be more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} because consumers, on average, have access to a larger number of prey species.
+However, in a well-connected network a disturbance may spread more easily and cause more secondary extinctions due to the presence of many possible pathways.
+    The net effect of connectance on extinction therefore likely depends on the severity of disturbance and the ability of consumers to subsist on a reduced set of prey.
+
+    Local properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
+    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}. 
+    However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}.
+For example, a consumer with low trophic level may be at high risk of extinction if it consumes prey which are themselves very likely to be affected by disturbance,and a specialist may not necessarily have a high risk of extinction if it consumes a basal resource which is unlikely to be affected by disturbance.
+
+    %\subsection*{3 Motif participation may similarly affect species' particular extinction risk}
+    One way to incorporate more nuance into descriptions of how species fit into their communities, and provide information linking global-scale and local-scale properties, is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
+    Such motif profiles of species are easily calculated once we know the network structure.
+    Importantly, these `motif participation' roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a}.
+    This means that two species with the same in-degrees and trophic levels can still participate in different sets of motifs if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
+
+    This extra detail may be useful when predicting extinction risks of single species, but to our knowledge this has not yet been investigated. 
+    However, at the network level, it has been shown that the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
+    While these studies uncovered the potential influence of network-level motif frequencies on network stability, a disturbance may likely alter the network and the interactions between species in different ways.
+    This can affect the influence of motif frequencies on network stability, as it has been demonstrated that an unstable motif may appear stabilizing dependent on interactions with other species \citep{Kratina2012}.
+
+%\subsection*{Here we test...}
+    Therefore, we are interested in whether a consumer's extinction risk depends on 1) the motifs in which the species participates and 2) the motif profile of the full food web, in scenarios with various levels of disturbance targeting the basal level.
+    In addition, we want to understand how the information provided by motifs is related to other global and local food web properties in order to better synthesize information provided by different scales of network structure.
+    We model extinction probabilities using Bayesian network modeling \citep{Eklof2013, Haussler2020}
+    Of the 13 three-species motifs that can occur in empirical food webs, four motifs are over-represented in empirical networks and are thought to contribute to stability; the three-species chain, omnivory, apparent competition, and direct competition \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
+    Therefore, we focus on these four particularly important motifs.
+
+     %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
+
+% OLD INTRODUCTION BELOW
+\section*{Introduction} %Old, for comparison
+
+    %\subsection*{1 Why are bottom-up effects important for understanding stability?}
+     The plant community is the foundation upon which the myriad species in terrestrial food webs depend for their survival.
+Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
+     The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
      For example, the loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
@@
     One way to incorporate more nuance into descriptions of how species fit into their communities is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
+This can be referred to as meso-scale properties and describe how a species is interacting with its closest neighbours.
+As such, meso-scale properties is thus one a scale between local properties and global properties.
 Importantly, these `motif participation' roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a}.
@@
 \section*{Methods}
+
+    \subsection*{Bayesian network framework}	Traditionally, there are two main approaches for studying secondary extinctions.
+First, there are topological models, based only on food web structure \citep{dunne2009cascading}.
+Here, extinctions only affect other species in the network once a consumer has lost all of its prey and therefore must go extinct.
+    The second approach uses dynamical models which explicitly simulate population dynamics using a system of differential equations \citep{binzer2011susceptibility}.
+Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network.
+    This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are more parameter intensive, and simulations are more time-consuming, than topological models.
+
+    A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks \citep{Eklof2013}. 
+    In this framework, a consumer's probability of extinction depends on the fraction of resources lost and on a baseline probability of extinction which captures causes unrelated to the network structure (e.g., disease, stochastic extinction of small populations).
+    Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
+    As consumers do not affect the extinction probabilities of their resources, Bayesian networks can only capture bottom-up effects of disturbances within networks. 
+    However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
+    This emphasizes the importance of bottom-up effects in food webs and makes Bayesian networks the ideal framework to use when studying them.
 
@@
 
-	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties.
+	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, level of disturbance, and their interactions).
 	See Appendix \emph{3.1 - 3.5} for details. 
@@
         All predictors were centered and scaled before fitting the models.         
-        To test whether these general trends were consistent across networks, we examined the distribution of slopes of simplified regressions fit to a single network and level of disturbance (Appendix \emph{3.1}).
+        To test whether these general trends were consistent across networks, we visually examined the distribution of slopes of simplified regressions fit to a single network and level of disturbance (Appendix \emph{3.1}).
 
     \subsubsection*{2) Motif participation and other network properties}
-        Different scales of network structure (e.g., local and meso-scale) are not independent. 
-        To identify correlations between global network structure and participation in particular motifs, we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
-        We fit a similar set of  LMMs relating motif participation to in-degree or trophic level (Appendix \emph{S3.2}).
+
+        Properties which refer to different scales of network structure, including different amounts of direct and indirect interactions with the focal species, are not independent. 
+        To identify correlations between global network structure (i.e., the whole network) and participation in particular motifs (a meso-scale description of species roles), we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
+        We fit a similar set of  LMMs relating motif participation to in-degree or trophic level (i.e., local measures of a species' role; Appendix \emph{S3.2}).
         Taken together, these supplemental models demonstrate how motif participation relates to other measures of network structure.
@@
     These simulated results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
+    For example, at low levels of disturbance a high degree may be protective since species can compensate if one or a few prey are lost. 
+    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and have lower persistence.
     These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.

### 2022-05-24 manuscript/response_PRSB_round2.tex
a few edits to the response

@@
 \us
+
+This definitely sounds like the AE just wants a comparison of motif frequencies with empirical networks - much easier than showing same results!
+
 Alyssa - this week (or bother me why not)
@@
 
-\us - Alyssa: S and C important but not main focus, so want to address them and then move on to more detailed things.
+\us [[May be related to some of the confusion about the purpose of different tests. Make sure the question asked by each analysis is clearer now.]] 
+- Alyssa: S and C important but not main focus, so want to address them and then move on to more detailed things.
 a written response here (and something brief in the paper)
@@
 However, the examples provided do not convince me that species’ roles based on motifs are a more nuanced way of describing a species’ role within its food web since the two examples provided refer to the influence disturbance may have on a basal resource – an influence that I’m not convinced a species’ role based on motifs will impact.
+
+\us We have revised the introduction throughout, and have taken care to clearly link the extra information given by motifs to our intuition about extinction risk.
+In particular, motifs give information about a consumer's prey that may reflect the prey's risk of extinction and thereby the predators' risk (e.g., a predator consuming generalist prey will participate in many three-species chain motifs).
+
 I do think that species’ roles based on motifs are a more nuanced way of describing species’ roles, and I am interested in the central question of this paper – however, this rationale would not convince me if I was unfamiliar with motifs/species roles/finer scale network descriptors.
@@
 
-\us Degree is better than TL and then you still need the whole web - mention limits of stable isotopes (stable isotope guy maybe?) ...
-Address this in the new (conceptual?) figure.
-Show motif profiles that should experience strong and weak effects of bottom-up disturbance
+\us In our analysis, in-degree was a much stronger predictor of extinction risk than trophic level.
+We thus do not escape the need for full food-web information to predict extinction risk using coarse measures, even if we accept that stable isotope analyses are much less work than constructing a full food web (and whether stable isotopes are easier than constructing a full food web seems to largely depend on how many species are of interest ...
+we are not convinced that obtaining SI trophic levels for 100 species would be easier that estimating the full food web).
+Regardless, we are not suggesting that motifs replace degree or trophic level as all four metrics can be calculated using the same data.
+We are interested in showing that motifs can provide information about extinction risk that is not given by degree and trophic level.
+
+[[Make sure conceptual figure shows how species with same degree, TL can participate in different motifs and have different risks - consumer with degree=2 consuming two specialists vs.
+consuming two generalists should make the point]]
 
@@
 \us
-We have now revised the introduction to introduce the motif profile of the whole network and its potential relevance to species-level extinction risk [[check]] and hope that the purpose of these analyses is now clearer.
-[[edit to be nicer]]We did not expect that the motif profile of a full food web to outperform the impact of a species motif participation, instead, this was done to investigate whether the full profile of a web would give some additional information, in "average" terms.
+We have now revised the introduction to introduce the motif profile of the whole network and its potential relevance to species-level extinction risk [[check - yes, I think so]] and hope that the purpose of these analyses is now clearer.
+Note that we do not expect the frequencies of motifs at the network to out-perform species-level motif participation.
+In fact, we do not compare the predictive abilities of motif profiles and motif participation as these models relate to different things.
+We connect network-level motif profiles to the network-level average extinction risk of all consumers and species-level motif participation to species-level extinction risk of one consumer at a time.
+We regret the confusion about the goals of our different analyses in the previous draft and hope that it is now clear how our analyses fit together.
 
@@
 
-\us We have revised the introduction to more clearly introduce these concepts.
+\us We have revised the introduction to more clearly introduce these concepts. [[yes]]
 
@@
 
-\us This is now re-written, and we describe this limitation of the Bayesian network framework in the Methods section [[might need to make this more explicit in the methods section, it is now mentioned a bit here and there, also add references to Allesina lab papers if we have some additional?]]
+\us In revising the introduction, we now introduce the Bayesian network approach and its limitations in the methods section.
+[[might need to make this more explicit in the methods section, it is now mentioned a bit here and there, also add references to Allesina lab papers if we have some additional?]] [[I think it would be good to reiterate around this line that most secondary extinctions are captured well by the Bayesian framework.
+It does seem a little hard to believe that acyclic webs work so weel if you're new.]]
 
@@
 
-\us This concept is now introduced in the introduction [[but should we discuss it a bit more also in the statistical analysis?
+\us This concept is now introduced in the introduction [[Disturbance is introduced, but different effects of different severities not clear yet.]] [[but should we discuss it a bit more also in the statistical analysis?
 we mention the different disturbance levels in the end of the methods section Calculating persistence]]
@@
 
-\us Frequency normalisation explicitly controls for total number of motifs, removing that potential (not proven to our knowledge) effect of S and C.
-To our knowledge, we are the first to explore whether S and C affect motif profiles ...
-We have them as random intercepts to account for 2 since this is indeed a known effect and we aren't interested in re-inventing the wheel
+\us We explicitly control for differences in the number of motifs through the frequency normalisation discussed above.
+We include the random intercepts to account for potential effects of Sn and Cn on network persistence that are \emph{not} due to effects on the total number of motifs.
+We agree with the Reviewer that the effects of Sn and Cn are well-known and important to account for, but they are not the main focus of our manuscript.
+We are therefore comfortable with keeping them as random effects in these equations.
 
@@
 
-\us See response to XX ... probably need to repeat that our focus is not to throw away degree
+\us See response to XX ... probably need to repeat that our focus is not to throw away degree [[Definitely need to clarify our modelling framework]]
 
@@
 \us Alyssa: Expand explanation of why high connectance can be good sometimes (more prey available to switch to, albeit more disturbance pathways).
-Probably a good way to introduce size of disturbance - many pathways of a small disturbance may not matter, many pathways of a large more severe
+Probably a good way to introduce size of disturbance - many pathways of a small disturbance may not matter, many pathways of a large more severe [[I think the current version does a succinct job of showing the good and bad of connectance]]
 

### 2022-05-24 manuscript/reorder_main_text.tex

@@
     Local properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
-    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}. 
+    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, including secondary extinction due to reduced plant populations.
     However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
-For example, a consumer with low trophic level may be at high risk of extinction if it consumes prey which are themselves very likely to be affected by disturbance,and a specialist may not necessarily have a high risk of extinction if it consumes a basal resource which is unlikely to be affected by disturbance.
+    As well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
+    For example, a generalist consuming many specialist (and therefore high-risk) prey may, for example, be at greater risk of extinction than a specialist which consumes a highly generalist resource that is very unlikely to go extinct.
+    %For example, a consumer with low trophic level may be at high risk of extinction if it consumes prey which are themselves very likely to be affected by disturbance,and a specialist may not necessarily have a high risk of extinction if it consumes a basal resource which is unlikely to be affected by disturbance.
 
     %\subsection*{3 Motif participation may similarly affect species' particular extinction risk}
-    One way to incorporate more nuance into descriptions of how species fit into their communities, and provide information linking global-scale and local-scale properties, is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
-    Such motif profiles of species are easily calculated once we know the network structure.
-    Importantly, these `motif participation' roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a}.
-    This means that two species with the same in-degrees and trophic levels can still participate in different sets of motifs if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
+    It is possible to capture information about how a consumer and its prey fit together within a community by defining participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
+    Such motif profiles of species are easily calculated once we know the network structure and require no more information than is used to calculate degree.
+    These `motif participation' roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
+    By allowing researchers to distinguish between, for example, consumers of specialists and generalists, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
 
-    This extra detail may be useful when predicting extinction risks of single species, but to our knowledge this has not yet been investigated. 
-    However, at the network level, it has been shown that the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
+    While the potential relationships between motif participation and species-level extinction risk have not yet been tested to our knowledge, it has been shown that the frequencies of different motifs within a whole network are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
     While these studies uncovered the potential influence of network-level motif frequencies on network stability, a disturbance may likely alter the network and the interactions between species in different ways.

### 2022-05-25 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-     The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
+     As such, the structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
      For example, the loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
 \citep{scherber2010bottom}.
+Secondary extinctions and plants...
 
     %\subsection*{2 Global and local network properties are known to affect stability}
-    How population decline or extinction of one species leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    How disturbances, such as population decline or extinction, lead to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
     Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances.
-%network functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002, Eklof2006}. 
 Two well-known, influential global properties are network size, the number of species present in the network, and network connectance, the number of realized feeding interactions.
@@
     Highly-connected networks may be more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} because consumers, on average, have access to a larger number of prey species.
-However, in a well-connected network a disturbance may spread more easily and cause more secondary extinctions due to the presence of many possible pathways.
+However, in a well-connected network a disturbance may spread more easily and cause more secondary extinctions due to the presence of many possible pathways \citep{}.
     The net effect of connectance on extinction therefore likely depends on the severity of disturbance and the ability of consumers to subsist on a reduced set of prey.
@@
     Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, including secondary extinction due to reduced plant populations.
+
     However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
@@
     For example, a generalist consuming many specialist (and therefore high-risk) prey may, for example, be at greater risk of extinction than a specialist which consumes a highly generalist resource that is very unlikely to go extinct.
-    %For example, a consumer with low trophic level may be at high risk of extinction if it consumes prey which are themselves very likely to be affected by disturbance,and a specialist may not necessarily have a high risk of extinction if it consumes a basal resource which is unlikely to be affected by disturbance.
 
@@
     While the potential relationships between motif participation and species-level extinction risk have not yet been tested to our knowledge, it has been shown that the frequencies of different motifs within a whole network are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
+This indicate that motifs may provide additional information on network resilience to disturbances.
+
+    The severity of disturbances impacting structure forming species, i.e., plant species, is important for how the food web respond \citep{}.
+Naturally, the magnitude of the effect will increase with severity of the disturbance \citep{}, but additionally disturbances impacting structure forming species, i.e., plant species, may fundamentally alter food web structure \citep{byrnes2011climate}.
+This has been shown in several studies where the magnitude of the disturbances alter the competition between species in the benthic community \citep{detmer2021variation} and change the percent coverage of different taxonomic groups \citep{carnell2020more} in kelp forests.
+
     While these studies uncovered the potential influence of network-level motif frequencies on network stability, a disturbance may likely alter the network and the interactions between species in different ways.
-    This can affect the influence of motif frequencies on network stability, as it has been demonstrated that an unstable motif may appear stabilizing dependent on interactions with other species \citep{Kratina2012}.
+    This can affect the influence of motif frequencies on network stability, as it has been demonstrated that a previously unstable motif may appear stabilizing dependent on interactions with other species \citep{Kratina2012}.
 
 %\subsection*{Here we test...}
-    Therefore, we are interested in whether a consumer's extinction risk depends on 1) the motifs in which the species participates and 2) the motif profile of the full food web, in scenarios with various levels of disturbance targeting the basal level.
+   Here, we are interested in whether a consumer's extinction risk of going secondary extinct depends on 1) the motifs in which the species participates, 2) the motif profile of the full food web and 3) the severity of the disturbance.
+%We address this in scenarios with various levels of disturbance targeting the basal level.
     In addition, we want to understand how the information provided by motifs is related to other global and local food web properties in order to better synthesize information provided by different scales of network structure.
-    We model extinction probabilities using Bayesian network modeling \citep{Eklof2013, Haussler2020}
-    Of the 13 three-species motifs that can occur in empirical food webs, four motifs are over-represented in empirical networks and are thought to contribute to stability; the three-species chain, omnivory, apparent competition, and direct competition \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
-    Therefore, we focus on these four particularly important motifs.
+    We model secondary extinction probabilities using Bayesian networks \citep{Eklof2013, Haussler2020}. 
+    We focus our analyse on the four network motifs that are over-represented in empirical networks and additionally have been shown to contribute to network stability; the three-species chain, omnivory, apparent competition, and direct competition \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
 

### 2022-05-27 code/calculate_empirical_motifs_and_roles.py
comparing empirical motifs and roles

@@ -4,10 +4,10 @@ import pymfinder as py
 
 def main():
 
-    webdir='../data/empirical/edgelists/'
-    roledir='../data/roles/empirical/'
-    motifdir='../data/motif_frequencies/empirical/'
-    pardir='../data/motif_participation/empirical/'
+    webdir='../data/empirical/global_verts/edgelists/'
+    roledir='../data/roles/empirical/global_verts/'
+    motifdir='../data/motif_frequencies/empirical/global_verts/'
+    pardir='../data/motif_participation/empirical/global_verts/'
 
     for web in os.listdir(webdir):
       infile=webdir+web

### 2022-05-27 code/calculate_motifs_and_roles.py

@@ -4,10 +4,10 @@ import pymfinder as py
 
 def main():
 
-    webdir='../data/edgelists/acyclic/'
-    motifdir='../data/motif_frequencies/acyclic/'
-    roledir='../data/roles/acyclic/'
-    pardir='../data/motif_participation/acyclic/'
+    webdir='../data/edgelists/original/'
+    motifdir='../data/motif_frequencies/original/'
+    roledir='../data/roles/original/'
+    pardir='../data/motif_participation/original/'
 
     for s in sorted(os.listdir(webdir)):
       try:

### 2022-05-27 code/collect_empirical_motif_profiles.py

@@ -1,6 +1,23 @@
 import os
 import sys
 
+def CSreader(roledir): # Read the edgelists, calculate S and C, make a dict.
+  Cdict={}
+  for web in os.listdir(roledir):
+    S=set()
+    L=set()
+    f=open(roledir+web,'r')
+    for line in f:
+      (pred,prey)=line.split()
+      S.add(pred)
+      S.add(prey)
+      L.add((pred,prey))
+    f.close()
+    C=float(len(L))/float(len(S)**2)
+    Cdict[web.split('.tsv')[0]]=((len(S),C))
+  return Cdict
+
+
 # Creates a dictionary with motif counts listed by species, connectance, network, motif.
 def read_role_file(roledir):
   role_dict={}
@@ -23,7 +40,7 @@ def read_role_file(roledir):
   return role_dict
 
 
-def write_motif_file(role_dict,outfile):
+def write_motif_file(role_dict,Cdict,outfile):
   f=open(outfile,'w')
   f.write('Size\tConnectance\tNetwork\tm6\tm12\tm36\tm38\tTotal\n')
   for web in role_dict:
@@ -37,16 +54,17 @@ def write_motif_file(role_dict,outfile):
 
   return
 
-# Calculated S and C=L/(S*(S-1)) for empirical webs
-Cdict={'kongsfjorden':(260,0.024),'loughhyne':(341,0.043),'reef_noSynodus':(242,0.056),
-    'stmarks':(141,0.089),'weddell':(490,0.060),'ythanjacob':(88,0.051)}
+# # Calculated S and C=L/(S*(S-1)) for empirical webs
+# Cdict={'kongsfjorden':(260,0.024),'loughhyne':(341,0.043),'reef_noSynodus':(242,0.056),
+#     'stmarks':(141,0.089),'weddell':(490,0.060),'ythanjacob':(88,0.051)}
 
 def main():
+  Cdict=CSreader('../data/empirical/global_verts/edgelists/')
   # role_dict takes s, c, network, motif: count
   # only motifs 6, 12, 36, 38 exist in the Bayesian networks
-  motif_dict=read_role_file('../data/motif_frequencies/empirical/')
+  motif_dict=read_role_file('../data/motif_frequencies/empirical/global_verts/')
 
-  write_motif_file(motif_dict,'../data/3sp_empirical_motif_profiles.tsv')
+  write_motif_file(motif_dict,Cdict,'../data/3sp_global_verts_motif_profiles.tsv')
   print 'Motif roles read'
 
 

### 2022-05-27 code/collect_empirical_roles.py

@@ -1,6 +1,22 @@
 import os
 import sys
 
+def CSreader(roledir): # Read the edgelists, calculate S and C, make a dict.
+  Cdict={}
+  for web in os.listdir(roledir):
+    S=set()
+    L=set()
+    f=open(roledir+web,'r')
+    for line in f:
+      (pred,prey)=line.split()
+      S.add(pred)
+      S.add(prey)
+      L.add((pred,prey))
+    f.close()
+    C=float(len(L))/float(len(S)**2)
+    Cdict[web.split('.tsv')[0]]=((len(S),C))
+  return Cdict
+
 # Then bin the species according to ... things, like in Benno's paper.
 # Then do some stats and send to Anna.
 def read_persistence_file(persfile,role_dict):
@@ -26,7 +42,10 @@ def read_persistence_file(persfile,role_dict):
       sp='"sp'+''.join(line.split()[2].split('"'))+'"'
       indeg=int(line.split()[3])
       outdeg=int(line.split()[4])
-      STL=int(line.split()[5])
+      try:
+        STL=int(line.split()[5])
+      except ValueError:
+        STL=10000
       basal_p=float(line.split()[6])
       web=line.split()[7].split('.')[0].split('"')[1]
       persdict[web][basal_p][sp]=(str(per))
@@ -120,56 +139,70 @@ def sort_nonzero_roles(role_dict,full_positions):
   return good_motifs
 
 
-def collect_and_print_3sp(outfile,role_dict,good_positions,partic_dict,good_motifs,degdict,TLdict,persdict):
+def collect_and_print_3sp(outfile,role_dict,Cdict,good_positions,partic_dict,good_motifs,degdict,TLdict,persdict):
+
+  nomotif_species=set()
   f=open(outfile,'w')
   f.write('Size\tConnectance\tBasal_p\tNetwork\tSpecies\tPersistence\tin_Degree\tout_Degree\tSTL\tm')
   f.write('\tm'.join(sorted(good_motifs[3]))+'\tp')
   f.write('\tp'.join(sorted(good_positions))+'\n')
   for web in persdict:
+    print web
     for bp in [0.0,0.2,0.4,0.6,0.8,1.0]:
       for sp in persdict[web][bp]: # Seems to be only non-basal species?
+        # Particdict doesn't include all species - some may not be in 3sp motifs
+        # Especially in webs
         f.write(str(Cdict[web][0])+'\t'+str(Cdict[web][1])+'\t')
         f.write('\t'.join([str(bp),web,sp]))
[…]

### 2022-05-27 code/convert_empirical_webs.R

@@ -19,10 +19,10 @@ topo_sort_acyclic <- function(A) {
     return(A[o,o])
 }
 
-goodfiles=as.character(list.files(path="../data/empirical/matrix/", full.names=FALSE, pattern="*.tsv"))
+goodfiles=as.character(list.files(path="../data/empirical/global_verts/matrix/", full.names=FALSE, pattern="*.tsv"))
 for (g in goodfiles){
     print(g)
-    A <- read.table(paste("../data/empirical/matrix/",g,sep=''), sep="\t", row.names=1, header=TRUE)
+    A <- read.table(paste("../data/empirical/global_verts/matrix/",g,sep=''), sep="\t", row.names=1, header=TRUE)
     rownames(A) <- colnames(A) <- 1:nrow(A) 
     # image(t(apply(A,2,rev)))
     
@@ -38,9 +38,9 @@ for (g in goodfiles){
     # Let's only keep the ones with max PATL<6
     # Also indexing by "good webs" so I know how many to re-run
 
-    write.table(A, file=paste("../data/empirical/acyclic_matrix/",g,sep=''), sep=",")
+    write.table(A, file=paste("../data/empirical/global_verts/acyclic_matrix/",g,sep=''), sep=",")
     edges=get.edgelist(graph.adjacency(as.matrix(A)))
-    write.table(edges,file=paste0('../data/empirical/edgelists/',g,sep=''),sep='\t',row.names=FALSE,col.names=FALSE)
+    write.table(edges,file=paste0('../data/empirical/global_verts/edgelists/',g,sep=''),sep='\t',row.names=FALSE,col.names=FALSE)
 }
 
 

### 2022-05-27 code/plot_profiles_vs_empirical.py

@@ -1,0 +1,362 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Spectral', reverse=True,n=18)  # The blue is very beautiful but maybe harder to see.
+colors.add_color(173,132,191,'Purple')
+colors.add_color(34,90,34,'Green')
+# colors.add_color(120,120,120,'grey')
+# colors.add_color(255,125,125,'lightish_red')
+# colors.add_color(200,200,200,'lightgrey')
+
+# Want to use AC, DC as x's, chain, omni as y's
+
+def read_motif_roles(motfile):
+  profiles={}
+  f=open(motfile,'r')
+  for line in f:
+    if line.split()[0]!='Size':
+      S=int(line.split()[0])
+      if S not in profiles:
+        profiles[S]={}
+      C=float(line.split()[1])
+      if C not in profiles[S]:
+        profiles[S][C]={'AC':[],'DC':[],'chain':[],'omnivory':[]}
+      basal_p=float(line.split()[2])
+      ID=line.split()[3]
+      sp=line.split()[4]
+      persis=line.split()[5]
+      deg=float(line.split()[6])
+      out_deg=line.split()[7]
+      STL=line.split()[8]
+      PATL=line.split()[9]
+      chain=int(line.split()[10])
+      DC=int(line.split()[11])
+      omnivory=int(line.split()[12])
+      AC=int(line.split()[13])
+      total=chain+DC+omnivory+AC
+      if basal_p==1.0 and deg>0 and S>50: # Only want one per network
[…]

### 2022-05-27 code/stat_analysis/empirical_statistics.R

@@ -11,7 +11,8 @@ library(vegan)
 library(MuMIn)
 
 
-data<-read.table(file = '../../data/empirical_3sp_roles_participation_nonlinear.tsv', sep = '\t', header = TRUE)
+# data<-read.table(file = '../../data/global_verts_3sp_roles_participation_nonlinear.tsv', sep = '\t', header = TRUE)
+data<-read.table(file = '../../data/global_verts_3sp_roles_participation_linear.tsv', sep = '\t', header = TRUE)
 data$Disturbance<-0.1+0.4*data$Basal_p
 # ONLY FIT MODELS ON CONSUMERS!!!
 consumers<-data[which(data$in_Degree>0),]
@@ -36,7 +37,7 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
 
 
 # Section 2: How do motif profiles vary with S and C?
-    motdata=read.table('../../data/3sp_empirical_motif_profiles.tsv',header=TRUE)
+    motdata=read.table('../../data/3sp_global_verts_motif_profiles.tsv',header=TRUE)
     motdata$mot_total=rowSums(motdata[,4:7])
     motdata$prop_chain=motdata$m12/motdata$mot_total
     motdata$prop_apparent=motdata$m6/motdata$mot_total
@@ -83,7 +84,8 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
         interdata2$S[r]=strsplit(levelcode,':')[[1]][1]
         interdata2$C[r]=strsplit(levelcode,':')[[1]][2]
     }
-    write.table(interdata2,file='empirical_proportion_variability_SC_nonlinear.tsv',sep='\t')
+    write.table(interdata2,file='global_verts_proportion_variability_SC_linear.tsv',sep='\t')
+    # write.table(interdata2,file='global_verts_proportion_variability_SC_nonlinear.tsv',sep='\t')
 
     # LMs - no need for network-level random effect since it's within-network motif profiles
     pchainlm=with(motdata,lm(prop_chain~Size*Connectance))
@@ -158,8 +160,10 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     TLper=with(consumers,lmer(Persistence~scale(STL)*scale(Disturbance)+(1|Global)))
     Degper=with(consumers,lmer(Persistence~scale(in_Degree)*scale(Disturbance)+(1|Global)))
 
[…]

### 2022-05-27 manuscript/supplement_motifs.tex
refining SI

@@
             \item Conversion of networks to acyclic form
+            \item Comparison of empirical and simulated networks
             \end{itemize}
@@
 
+        \item Motif profiles and global properties
+
+            \begin{itemize}
+                \item Contains Figure S5
+                \item Contains Table S5
+            \end{itemize}
+
+        \item Persistence and global network properties
+            \begin{itemize}
+                \item Contains Figures S6, S7
+                \item Contains Table S6
+            \end{itemize}
+
         \item Glossary
+
+            \begin{itemize}
+                \item Contains Table S7
+            \end{itemize}
+
     \end{enumerate}    
@@
 \clearpage
+
+    \subsection*{Comparison of empirical and simulated networks}
+
+        Although previous work has shown that the niche model replicates the motif structure of empirical networks~\citep{Stouffer2005a,Stouffer2006}, we provide a comparison of the motif profiles and consumer motif participation for our simulated networks and the highest-quality empirical networks we could find within the same size range.
+        For this comparison, we obtained all networks containing 50-100 species from \citet{globalwebdb}.
+        We then removed networks where sets of predators and prey were completely different (i.e., bipartite networks) or links were not clear (i.e., +/- effects rather than predation specifically).
+        This yielded five vertebrate-dominated food webs describing the California coast in different years, four invertebrate-dominated food webs describing mainland US streams, 23 invertebrate-dominated food webs describing New Zealand streams, and five other invertebrate-dominated food webs.
+        We note that most of these webs were compiled by just three research groups.
+        Moreover, many of the webs are repeated samples of the same site  or closely-grouped sites.
+        This means that these empirical networks cannot be considered fully independent samples.
+
+        \begin{figure}
+            \includegraphics[width=\textwidth]{figures/motif_profiles_participation_vs_empirical.eps}
+            \caption{\textbf{A)} Proportions of acyclic motifs in the motif profiles of empirical and simulated networks.
+Mean proportions ($\pm$SD) in the simulated networks are indicated in black, with the range of proportions in the simulated networks indicated by black boxes.
+\textbf{B)} Proportions of acyclic motifs in the motif participation vectors of consumers in simulated and empirical networks.}
+            \label{empirical_compare}
+        \end{figure}
+
+    \subsubsection*{Comparing motif profiles}
+
+      The proportions of omnivory and apparent competition in the motif profiles of simulated networks matched those of the empirical webs fairly well, but many of the empirical webs had higher proportions of direct competition and lower proportions of the three-species chain than any simulated web.
+      The empirical webs with high proportions of direct competition tend to be invertebrate-dominated webs representing small streams.
+      High proportions of direct competition in these system suggest large numbers of fairly generalist herbivore/detritivores, with relatively few secondary consumers present to create chains.
+
+      A few webs had lower proportions of apparent competition than any simulated web but similar proportions of the other motifs.
+      Most of these webs represented the California coast in different years (2003-2007).
+      Basal resources in these webs were highly aggregated into only two size classes of phytoplankton and three types of detritus.
+      This aggregation at the basal level leads to artificial ``specialisation'' in consumers with low trophic levels who may consume, for example, many species of phytoplankton that are grouped into a single size class.
+      As the basal resources in our simulated networks are not aggregated, our webs feature higher proportions of apparent competition as seen in the other empirical webs.
+
+    \subsubsection*{Comparing motif participation}
+
+      Despite the differences in network motif profiles between some of the empirical food webs and the simulated webs, we note that the motif participation of consumers in the simulated webs entirely spans the range of motif participation for consumers in the empirical webs.
+      Consumers in the NZ stream food webs had higher proportions of direct competition and lower proportions of apparent competition than the mean of the simulated webs, consistent with the network motif profiles.
+      Nonetheless, the greater differences in network profiles than consumer motif participation suggests that the largest differences between the simulated and empirical networks are in the roles of basal resources.
+      In empirical webs, basal resources are often highly aggregated. 
+      Improved resolution at the basal level should improve the correspondence between simulated and empirical webs and allow better modelling of empirical communities.
 

### 2022-05-27 manuscript/response_PRSB_round2.tex
small edits to reply

@@
 
-This definitely sounds like the AE just wants a comparison of motif frequencies with empirical networks - much easier than showing same results!
-
-Alyssa - this week (or bother me why not)
-Webs referenced in the niche model vs empirical motifs paper
+Alyssa to plot box and whiskers for simulated motif profiles and participation plus dots for empirical webs (all globalweb with 50-100 species).
 
@@
 
-\us Clarify our writing ... is there a way to connect the figures better?
+\us We agree that there is a strong correlation between apparent competition and in-degree, and state that degree appears able to explain the trend in apparent competition.
+However, not all motifs show such strong relationships, meaning that degree is not sufficient to
+
+Clarify our writing ... is there a way to connect the figures better?
 

### 2022-05-27 code/plot_profiles_vs_empirical.py

@@ -227,6 +227,8 @@ def populate_graph(graph,sim_profiles,emp_profiles):
         if net=='WEB242' and mot=='AC':
           dat.legend='NZ Stream'
       else:
+        if mot=='AC':
+          print net
         dat.symbol.configure(shape=2,color=14,fill_pattern=0,linewidth=1)
         if net=='WEB181' and mot=='AC':
           dat.legend='Other invert-dominated'

### 2022-05-27 manuscript/reorder_main_text.tex

@@
 
-\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{2}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{1}$} 
+\author{Anna \r{A}kesson$^{1*}$, Alyssa R. Cirtwill$^{2*\dagger}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{1}$} 
 \date{\small$^1$Department of Theoretical Biology, Chemistry, and Physics\\ 
@@
 \medskip
+* Joint first authors
+\medskip
 $^\dagger$ Corresponding author:\\
+
 }
@@
     While the potential relationships between motif participation and species-level extinction risk have not yet been tested to our knowledge, it has been shown that the frequencies of different motifs within a whole network are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
-This indicate that motifs may provide additional information on network resilience to disturbances.
+    Given these relationships between network-level motif frequencies and overall community stability, it is worth investigating whether a species' participation in different motifs also affects its extinction risk.
+    We are also interested in whether the relationships between network structure and extinction risk depend on the severity of disturbance experienced by the community.
 
-    The severity of disturbances impacting structure forming species, i.e., plant species, is important for how the food web respond \citep{}.
-Naturally, the magnitude of the effect will increase with severity of the disturbance \citep{}, but additionally disturbances impacting structure forming species, i.e., plant species, may fundamentally alter food web structure \citep{byrnes2011climate}.
-This has been shown in several studies where the magnitude of the disturbances alter the competition between species in the benthic community \citep{detmer2021variation} and change the percent coverage of different taxonomic groups \citep{carnell2020more} in kelp forests.
+    Naturally, a larger disturbance tends to lead to more secondary extinctions \citep{}.
+    Disturbances to plant species can also affect some consumers more strongly than others  \citep{byrnes2011climate}. 
+    In empirical studies of kelp forests, the magnitude of the disturbances to basal resources can alter the competition between species in the benthic community \citep{detmer2021variation} and change the percent coverage of different taxonomic groups \citep{carnell2020more}.
 
-    While these studies uncovered the potential influence of network-level motif frequencies on network stability, a disturbance may likely alter the network and the interactions between species in different ways.
-    This can affect the influence of motif frequencies on network stability, as it has been demonstrated that a previously unstable motif may appear stabilizing dependent on interactions with other species \citep{Kratina2012}.
+    %While these studies uncovered the potential influence of network-level motif frequencies on network stability, a disturbance may likely alter the network and the interactions between species in different ways.
+    %This can affect the influence of motif frequencies on network stability, as it has been demonstrated that omnivory may appear stabilizing dependent on interactions with other species \citep{Kratina2012}.
 
 %\subsection*{Here we test...}
-   Here, we are interested in whether a consumer's extinction risk of going secondary extinct depends on 1) the motifs in which the species participates, 2) the motif profile of the full food web and 3) the severity of the disturbance.
+   Here, we are interested in whether a consumer's extinction risk of going secondarily extinct depends on 1) the motifs in which the species participates, 2) the motif profile of the full food web and 3) the severity of the disturbance to the basal level.
 %We address this in scenarios with various levels of disturbance targeting the basal level.
@@
 
-    \subsection*{Bayesian network framework}	Traditionally, there are two main approaches for studying secondary extinctions.
+	\subsection*{Initial network construction}
+
+		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}.
+		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
+		For full details, se \emph{Appendix S1}.
+		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
+
+    \subsection*{Modelling secondary extinction using Bayesian networks}
+
+    Traditionally, there are two main approaches for studying secondary extinctions.
 First, there are topological models, based only on food web structure \citep{dunne2009cascading}.
@@
     Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
+    This increased speed allows us to simulate a range of disturbance severity rather than single species removals as in many previous studies~\citep{}.
+    Moreover, Bayesian network simulations are less parameter-intensive than dynamical models, meaning that they are less sensitive to assumptions made in the modelling process~\citep{}.
+
     As consumers do not affect the extinction probabilities of their resources, Bayesian networks can only capture bottom-up effects of disturbances within networks. 
     However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
-    This emphasizes the importance of bottom-up effects in food webs and makes Bayesian networks the ideal framework to use when studying them.
-
-	\subsection*{Initial network construction}
-
-		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}.
-		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
-		For full details, se \emph{Appendix S1}.
-		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
-		All networks were rendered acyclic following~\citet{Allesina2009functional} and topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).
-		These processing steps allow the strictly bottom-up calculation of probability of persistence used in the Bayesian network framework.
+    [[Something about strong bottom-up effects in empirical studies]]
+	To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).
+	These processing steps allow the strictly bottom-up calculation of probability of persistence.
 

### 2022-06-01 code/Fig1_errorbars.py
figure 1 with error bars

@@ -1,0 +1,185 @@
+import sys
+import os
+import math
+import random
+from decimal import *
+import numpy as np
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Greys',n=10)  # The blue is very beautiful but maybe harder to see.
+# Match Anna's ggplot colours
+colors.add_color(94,79,162,'Rpurple')
+colors.add_color(50,136,189,'Rblue')
+colors.add_color(102,194,165,'Rgreen1')
+colors.add_color(171,221,164,'Rgreen2')
+colors.add_color(230,245,152,'Rgreen3')
+colors.add_color(255,255,191,'Ryellow1')
+colors.add_color(254,224,139,'Ryellow2')
+colors.add_color(253,174,97,'Rorange1')
+colors.add_color(244,109,67,'Rorange2')
+colors.add_color(213,62,79,'Rred1')
+colors.add_color(158,1,66,'Rred2')
+
+# colors.add_color(120,120,120,'grey')
+# colors.add_color(255,125,125,'lightish_red')
+# colors.add_color(200,200,200,'lightgrey')
+
+motnames={'apparent':'Apparent comp.','direct':'Diect comp.','chain':'Three-sp. chain','omnivory':'Omnivory'}
+
+scaledict={'dist':(0.3,0.1366261),'apparent':(0.4051135,0.1461447),'direct':(0.1923225,0.1143021),
+  'chain':(0.2489432,0.1272357),'omnivory':(0.1536209,0.1007478)}
+
+
+def read_lmfile(lmfile):
+  lmdict={'(Intercept)':[],'motif':[],'scale(Disturbance)':[],'inter':[]} # estimate, SD
+
+  f=open(lmfile,'r')
+  for line in f:
+    if line.split()[0]!='"Estimate"':
[…]

### 2022-06-01 code/Fig1_errorbars.py
better grey maybe

@@ -127,7 +127,7 @@ def populate_graph(graph,motif,lmdict):
     uperr=graph.add_dataset(upper)
     uperr.symbol.shape=0
     uperr.line.linestyle=0
-    uperr.fill.configure(type=2,color=5,pattern=1)
+    uperr.fill.configure(type=2,color=4,pattern=1)
 
     derr=graph.add_dataset(lower)
     derr.symbol.shape=0

### 2022-06-02 manuscript/reorder_main_text.tex
main text revisions per R1

@@
     The presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
-    Highly-connected networks may be more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} because consumers, on average, have access to a larger number of prey species.
-However, in a well-connected network a disturbance may spread more easily and cause more secondary extinctions due to the presence of many possible pathways \citep{}.
-    The net effect of connectance on extinction therefore likely depends on the severity of disturbance and the ability of consumers to subsist on a reduced set of prey.
+    Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2020}.
+However, a well-connected network also includes more pathways through which a disturbance can spread~\citep{Vieira2015}.
+    Depending on the strength and arrangements of interactions, this means that species may be impacted through several paths, amplifying the original disturbance and causing more secondary extinctions.
+    The net effect of connectance on extinction therefore likely depends on finer-scale network structure, as well as the type and severity of disturbance.
 
@@
     Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, including secondary extinction due to reduced plant populations.
-
     However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
@@
     While the potential relationships between motif participation and species-level extinction risk have not yet been tested to our knowledge, it has been shown that the frequencies of different motifs within a whole network are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
-    Given these relationships between network-level motif frequencies and overall community stability, it is worth investigating whether a species' participation in different motifs also affects its extinction risk.
-    We are also interested in whether the relationships between network structure and extinction risk depend on the severity of disturbance experienced by the community.
+    These relationships between network-level motif frequencies (`motif profiles') and overall community stability mean that a network's motif profile could affect the average extinction risk for consumers within the network.
+    Moreover, if network motif profiles are related to average extinction risk, it is worth investigating whether a species' participation in different motifs also affects its extinction risk.
 
-    Naturally, a larger disturbance tends to lead to more secondary extinctions \citep{}.
-    Disturbances to plant species can also affect some consumers more strongly than others  \citep{byrnes2011climate}. 
-    In empirical studies of kelp forests, the magnitude of the disturbances to basal resources can alter the competition between species in the benthic community \citep{detmer2021variation} and change the percent coverage of different taxonomic groups \citep{carnell2020more}.
+    As mentioned above, the severity of a disturbance may interact with network structure to affect secondary extinctions.
+    A mild disturbance (e.g., a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
+    Empirical studies have shown that disturbances to plant species can also affect some consumers more strongly than others  \citep{byrnes2011climate}, and that the strength of disturbance also affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}.
+    By taking advantage of recent advances in rapid simulation of ecological networks, we are able to simulate a range of disturbance intensities and test whether meso-scale network structure also interacts with disturbance.
 
@@
 
-	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, level of disturbance, and their interactions).
+    	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, in-degree, and trophic level).
+        Where applicable, we also consider how interactions between network structure and disturbance level might affect extinction risk.
     	See Appendix \emph{3.1 - 3.5} for details. 
@@
         \centering
+        \includegraphics[width=0.85\textwidth]{figures/persistence_vs_motifs.eps}
+        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).
+The effect of participating in each motif is based on the fixed effects in Equation~\ref{propreq}.
+The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance).
+95\% confidence intervals for each line are shown in grey.
+Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
+    \label{fig:prop_lmer_all}
+    \end{figure}
+
+    \begin{figure}[hb!]
+        \centering
         \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}

### 2022-06-02 manuscript/reorder_main_text.tex
Updates from Overleaf

--- 2022-05-27 manuscript/reorder_main_text.tex
@@
 %We address this in scenarios with various levels of disturbance targeting the basal level.
-    In addition, we want to understand how the information provided by motifs is related to other global and local food web properties in order to better synthesize information provided by different scales of network structure.
+    In addition, we want to understand how the information provided by motifs is related to global and local food web properties with known relationships to extinction risk in order to better synthesize information provided by different scales of network structure.
     We model secondary extinction probabilities using Bayesian networks \citep{Eklof2013, Haussler2020}. 
@@
 
-% OLD INTRODUCTION BELOW
-\section*{Introduction} %Old, for comparison
+% % OLD INTRODUCTION BELOW
+% \section*{Introduction} %Old, for comparison
 
-    %\subsection*{1 Why are bottom-up effects important for understanding stability?}
-     The plant community is the foundation upon which the myriad species in terrestrial food webs depend for their survival.
+%     %\subsection*{1 Why are bottom-up effects important for understanding stability?}
+%      The plant community is the foundation upon which the myriad species in terrestrial food webs depend for their survival.
 Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-     The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
-     For example, the loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
+%      The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
+%      For example, the loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
 \citep{scherber2010bottom}.
 
-     %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
-    How population decline or extinction of one species leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+%      %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
+%     How population decline or extinction of one species leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
 Traditionally, there are two main approaches for studying secondary extinctions.
@@
 Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network.
-    This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are more parameter intensive, and simulations are more time-consuming, than topological models.
+%     This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are more parameter intensive, and simulations are more time-consuming, than topological models.
 
-    A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks \citep{Eklof2013}. 
-    In this framework, a consumer's probability of extinction depends on the fraction of resources lost and on a baseline probability of extinction which captures causes unrelated to the network structure (e.g., disease, stochastic extinction of small populations).
-    Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
-    As consumers do not affect the extinction probabilities of their resources, Bayesian networks can only capture bottom-up effects of disturbances within networks. 
-    However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
-    This emphasizes the importance of bottom-up effects in food webs and makes Bayesian networks the ideal framework to use when studying them.
+%     A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks \citep{Eklof2013}. 
+%     In this framework, a consumer's probability of extinction depends on the fraction of resources lost and on a baseline probability of extinction which captures causes unrelated to the network structure (e.g., disease, stochastic extinction of small populations).
+%     Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
+%     As consumers do not affect the extinction probabilities of their resources, Bayesian networks can only capture bottom-up effects of disturbances within networks. 
+%     However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
+%     This emphasizes the importance of bottom-up effects in food webs and makes Bayesian networks the ideal framework to use when studying them.
 
-    %\subsection*{3 Global scale network properties are known to affect stability}
-    Using Bayesian networks, researchers can explore how the structures of realistically-large networks shape their responses to disturbance.
-    The global (whole-network) structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002, Eklof2006}.
-    For example, more highly-connected networks are more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{Krause2003, Teng2004, Gilarranz2017}.
+%     %\subsection*{3 Global scale network properties are known to affect stability}
+%     Using Bayesian networks, researchers can explore how the structures of realistically-large networks shape their responses to disturbance.
+%     The global (whole-network) structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002, Eklof2006}.
+%     For example, more highly-connected networks are more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{Krause2003, Teng2004, Gilarranz2017}.
 
-    %\subsection*{4 meso-scale nw propoerties also affect stability}
-    Structural properties on the local (single-species) scale are also known to affect response to disturbance. 
-    Species with high trophic levels (long paths to basal resources), low in-degree (few prey), and low abundances are generally at higher risks of secondary extinction \citep{binzer2011susceptibility}.
-    However, in-degree and trophic level are both relatively coarse descriptions of a species' role within its food-web~\citep{Cirtwill2018FoodWebs}. 
-    For example, a specialist may not necessarily have a high risk of extinction if it consumes a basal resource which is unlikely to be affected by disturbance. 
-    On the other hand, a consumer with low trophic level may be at high risk of extinction if it consumes prey which are themselves very likely to be affected by disturbance.
+%     %\subsection*{4 meso-scale nw propoerties also affect stability}
+%     Structural properties on the local (single-species) scale are also known to affect response to disturbance. 
+%     Species with high trophic levels (long paths to basal resources), low in-degree (few prey), and low abundances are generally at higher risks of secondary extinction \citep{binzer2011susceptibility}.
+%     However, in-degree and trophic level are both relatively coarse descriptions of a species' role within its food-web~\citep{Cirtwill2018FoodWebs}. 
+%     For example, a specialist may not necessarily have a high risk of extinction if it consumes a basal resource which is unlikely to be affected by disturbance. 
+%     On the other hand, a consumer with low trophic level may be at high risk of extinction if it consumes prey which are themselves very likely to be affected by disturbance.
 
-    %\subsection*{5 Motif participation may similarly affect species' particular extinction risk}
-    One way to incorporate more nuance into descriptions of how species fit into their communities is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
+%     %\subsection*{5 Motif participation may similarly affect species' particular extinction risk}
+%     One way to incorporate more nuance into descriptions of how species fit into their communities is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
 This can be referred to as meso-scale properties and describe how a species is interacting with its closest neighbours.
@@
 Importantly, these `motif participation' roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a}.
-    This means that two species with the same in-degrees and trophic levels can still participate in different sets of motifs if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
-    This extra detail may be useful when predicting  extinction risk since, at the network level, the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
-    We therefore expect that both the motif structure of a network and species' participation in that structure should be related to risks due to disturbance to basal resources.
+%     This means that two species with the same in-degrees and trophic levels can still participate in different sets of motifs if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
+%     This extra detail may be useful when predicting  extinction risk since, at the network level, the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
+%     We therefore expect that both the motif structure of a network and species' participation in that structure should be related to risks due to disturbance to basal resources.
 
-           %\subsection*{Here we test...}
-    Specifically, we are interested in whether a consumer's extinction risk depends on 1) the motifs in which a focal species participates and 2) the motif profile of the full food web. 
-    In addition, we want to understand how the information provided by motifs is related to classical global and local food web properties in order to better synthesize information provided by different scales of network structure.
-    Of the 13 three-species motifs that can occur in empirical food webs, there are four motifs that are over-represented in empirical networks and are thought to contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
+%           %\subsection*{Here we test...}
+%     Specifically, we are interested in whether a consumer's extinction risk depends on 1) the motifs in which a focal species participates and 2) the motif profile of the full food web. 
+%     In addition, we want to understand how the information provided by motifs is related to classical global and local food web properties in order to better synthesize information provided by different scales of network structure.
+%     Of the 13 three-species motifs that can occur in empirical food webs, there are four motifs that are over-represented in empirical networks and are thought to contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
 These four motifs, the three-species chain, omnivory, apparent competition, and direct competition, are also the only ones which can occur in the acyclic networks required for Bayesian network simulations~\citep{Eklof2013}.
-    Therefore, we focus on these four particularly important motifs.
+%     Therefore, we focus on these four particularly important motifs.
 
@@
         However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
+        This implies that bottom-up effects drive most secondary extinctions in simulations, echoing the strong bottom-up effects observed in empirical studies~\citep{}.
         [[Something about strong bottom-up effects in empirical studies]]
@@
 
-	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, level of disturbance, and their interactions).
+        We are primarily interested in whether a consumer's extinction risk is related to its motif participation or the motif profile of the network and, secondarily, also wish to connect the information provided by motifs to global and local network properties that have known effects on extinction risk.
+    	To address these questions, we tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (in-degree, STL, S, C, and level of disturbance).
     	See Appendix \emph{3.1 - 3.5} for details. 
@@
 
+            Our main question of interest is whether appearing in different motifs is associated with differences in extinciton risk. 
             To test if participating more often in a certain motif affected a species' probability of persistence, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks):
@@
 
-        Properties which refer to different scales of network structure, including different amounts of direct and indirect interactions with the focal species, are not independent. 
-        To identify correlations between global network structure (i.e., the whole network) and participation in particular motifs (a meso-scale description of species roles), we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
+            Properties which refer to different scales of network structure (e.g., local, global, and meso-scale) are not independent. 
+            This means that, to understand relationships between motif participation (a meso-scale network property) and extinction risk, we must also consider how motif participation varies with other properties known to affect extinction risk.
+            To identify correlations between global network structure (i.e., the whole network) and participation in particular motifs, we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
             We fit a similar set of  LMMs relating motif participation to in-degree or trophic level (i.e., local measures of a species' role; Appendix \emph{S3.2}).
@@
 
-        As well as species-specific probabilities of persistence, the overall frequency of motifs in a network may provide information about the overall probability of any species going extinct (since they incorporate information on indirect effects).
+            As well as relating species-specific probabilities of extinction to species-specific motif participation, we are interested in whether the overall frequency of motifs in a network provides information about the overall probability of any species going extinct.
+            Just as other whole-network properties (S, C, etc.) are known to affect the overall stability of a network, it is possible that the arrangement of interactions into different motifs could reflect different average probabilities of persistence across all consumers.
             We tested whether networks with similar motif profiles tend to have similar mean persistences among consumers using a PERMANOVA (Appendix \emph{S3.3}).
@@
 
-        To test whether networks with the the same size and connectance have more similar motif profiles, we used a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
-        In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMs:
+            Just like motif participation, a network's motif profile is not likely to be independent of other levels of structure. 
+            To formally test whether networks with the the same size and connectance have more similar motif profiles, we used a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
+            In addition, to understand how motif profiles may vary with global network structure in more detail, we identify correlations between size, connectance, and particular motifs using a set of four LMs:
             \begin{equation}

### 2022-06-02 manuscript/response_PRSB_round2.tex
Updates from Overleaf

@@
 
-\subsection*{Associate Editor}
+\subsection*{Associate Editor} [[Add conceptual figure, review Alyssa's new text]]
     \them
@@
     \us
-thanks etc etc
+    We thank the Associate Editor for their kind words and are especially pleased to have made a stronger case for motifs than some previous efforts have managed.
+We have taken the Associate Editor and Reviewers' comments to heart and hope that the revised manuscript is more approachable.
 
@@
     \us
+    We have thoroughly revised the introduction for clarity and directness, added a glossary in case of any remaining confusion about our terminology, and added explanations of the relevance of each statistical test we conduct.
+We hope that this makes our goals clearer and results easier to interpret.
+We have not adjusted our analyses as we note that both the Associate Editor and Reviewer 2 found the current methods suitable.
 
@@
     \us
+    We have added a conceptual figure showing the four motifs within a food web, with particular attention to how species with the same degree and trophic level can have different motif participation.
+We hope that this makes the text as a whole more approachable.
     Anna E
@@
     \us
+    While the niche model has been shown in previous work to simulate empirical networks reasonably well~\citep{Stouffer2005a,Stouffer2006}, we are happy to provide further confirmation in the supplemental information.
+    Specifically, we compare the motif profiles and consumers' motif participation of our simulated webs to the  unipartite webs in the same size range (excluding those where predators and prey were differently resolved) available from the globalweb database.
+    Not all of the empirical webs have motif profiles within the simulated range; this could be because of flaws in our simulation or because of flaws in the empirical webs (e.g., poorly-resolved basal resources).
+    While we cannot say conclusively which is the case, we are encouraged by the fact that the consumer motif participations in our simulated webs span the motif participations in the empirical web.
+    We hope that this comparison will improve empiricists' confidence in our simulations.
 
-Alyssa to plot box and whiskers for simulated motif profiles and participation plus dots for empirical webs (all globalweb with 50-100 species).
-
-The niche model has been shown in previous work to simulate empirical networks reasonably well~\citep{Stouffer2005a,Stouffer2006}.
-The specific method we use to simulate networks, developed in~\citet{Delmas2017}, has been used in several published studies already (e.g.,~\citep{})
+\clearpage
 
@@
 
-Comments to the Author(s)
+    \them
     Thank you for the opportunity to review this paper.
@@
 
+    \subsubsection*{Major comments}
+
         \them
@@
         \us
-The introduction has been re-written, and is hopefully clearer now regarding terminology and the questions asked. Additionally, we have now included a glossary in \emph{Appendix S11}
+        The introduction has been re-written, and is hopefully clearer now in terms of both terminology and our questions of interest.
+Additionally, we have now included a glossary in \emph{Appendix S11}
 
@@
 
-\us [[May be related to some of the confusion about the purpose of different tests. Make sure the question asked by each analysis is clearer now.]] 
-- Alyssa: S and C important but not main focus, so want to address them and then move on to more detailed things.
-a written response here (and something brief in the paper)
+        \us One large model will not serve the purpose of our manuscript.
+We agree that connectance and species richness have a large impact on persistence, but re-iterating these well-known effects is not our main focus.
+Instead, our main goal is to show whether, as is frequently proposed, information about motif participation or profiles can be used to comment on extinction risk.
+Secondarily, we hope to place these relationships in the context of known relationships between extinction risk and other network properties.
+We believe that separate, simpler analyses are more effective for this purpose than constructing one large model that hides the relationships between network properties.
+
+        We appreciate the Reviewer's general point that the purpose of these separate analyses was not clear.
+To remedy this, we have restated our questions of interest at the beginning of the `statistical analyses' section and now preface our description of each test with a statement of how it fits into the framework of our manuscript.
+We hope that the question addressed by each test is now clearer.
 
@@
         \us
-Additional analysis - maybe use motifs to explain differences between species with the same degree, trophic level? (try explanation first)
-if you have TL and degree you also have the motif profile of a species (disagree that trophic level is "easier" - once the network is characterised all three role types are available) put in the intro as well
+        We are not convinced that collecting data on trophic level is easier than collecting entire food web data.
+If the Reviewer is referring to stable-isotope trophic levels, then we must point out that calculating trophic levels requires a good background understanding of the food sources for the focal species and rates of incorporation of different isotopes across trophic levels (and how these may differ between taxa).
+For a large community (e.g., 100 species), it could well be simpler to at least approximate the entire food web (giving trophic level, degree, and motifs simultaneously) than to calculate stable isotope-based trophic levels for each species.
+Indeed, the larger set of empirical food webs than comprehensive stable-isotope data suggests that food webs are easier to estimate at scale than trophic levels alone.
+
+        To address this concern, we have added a brief statement in the introduction to make it clear to readers that researchers who have enough information to calculate in-degree and trophic level can also calculate motifs.
+Motif roles, therefore, provide additional information than degree and trophic level alone without requiring extra sampling.
 
@@
 
+    \subsubsection*{Minor comments}
+
         \them
-Smaller comments
-
         Line 65 – 70: I agree that in-degree and trophic level may be coarse description of a species’ role within its food-web.
@@
 
-\us We have revised the introduction throughout, and have taken care to clearly link the extra information given by motifs to our intuition about extinction risk.
+        \us
+        We have revised the introduction throughout, and have taken care to clearly link the extra information given by motifs to our intuition about extinction risk.
 In particular, motifs give information about a consumer's prey that may reflect the prey's risk of extinction and thereby the predators' risk (e.g., a predator consuming generalist prey will participate in many three-species chain motifs).
 
+        \them
         I do think that species’ roles based on motifs are a more nuanced way of describing species’ roles, and I am interested in the central question of this paper – however, this rationale would not convince me if I was unfamiliar with motifs/species roles/finer scale network descriptors.
@@
         \us In our analysis, in-degree was a much stronger predictor of extinction risk than trophic level.
-We thus do not escape the need for full food-web information to predict extinction risk using coarse measures, even if we accept that stable isotope analyses are much less work than constructing a full food web (and whether stable isotopes are easier than constructing a full food web seems to largely depend on how many species are of interest ...
-we are not convinced that obtaining SI trophic levels for 100 species would be easier that estimating the full food web).
+We thus do not escape the need for full food-web information to predict extinction risk using coarse measures, even if we accept that stable isotope analyses are much less work than constructing a full food web (and we are not fully convinced of this, see response above).
         Regardless, we are not suggesting that motifs replace degree or trophic level as all four metrics can be calculated using the same data.
@@
 
-[[Make sure conceptual figure shows how species with same degree, TL can participate in different motifs and have different risks - consumer with degree=2 consuming two specialists vs.
+        To more clearly illustrate how motifs can provide more information than degree and trophic level, we have added a conceptual figure.
+This figure illustrates the four motifs and shows how two species with the same degree and trophic level can have different motif participation.
+We hope that this figure and the revised introduction are more persuasive.
+
+        % [[Make sure conceptual figure shows how species with same degree, TL can participate in different motifs and have different risks - consumer with degree=2 consuming two specialists vs.
 consuming two generalists should make the point]]
@@
         \us
-We have now revised the introduction to introduce the motif profile of the whole network and its potential relevance to species-level extinction risk [[check - yes, I think so]] and hope that the purpose of these analyses is now clearer.
+        We have now revised the introduction to introduce the motif profile of the whole network earlier and make its potential relevance to species-level extinction risk clearer.
         Note that we do not expect the frequencies of motifs at the network to out-perform species-level motif participation.
@@
 
-\us We have revised the introduction to more clearly introduce these concepts. [[yes]]
+        \us We have revised the introduction to more clearly introduce these concepts.
 
@@
         Line 92 – 95: this seems to be a flaw to the Bayesian network simulations – we know that mutual predation happens in food webs – especially if the food web isn’t explicitly size structured, and when nested within the broader food web context these feedback loops may have stabilizing, or destabilizing effects.
+        \smallskip
+        Line 149 – 152: these seem like large limitations to me.
 
         \us In revising the introduction, we now introduce the Bayesian network approach and its limitations in the methods section.
-[[might need to make this more explicit in the methods section, it is now mentioned a bit here and there, also add references to Allesina lab papers if we have some additional?]] [[I think it would be good to reiterate around this line that most secondary extinctions are captured well by the Bayesian framework.
-It does seem a little hard to believe that acyclic webs work so weel if you're new.]]
-
-\them
-Line 149 – 152: these seem like large limitations to me.
-
-\us tone down discussion of limitations a bit, deal more in Discussion. Anna E
+There, we clearly state that the lack of top-down effects in the Bayesian networks is a limitation.
+However, we also point out that in prior studies Bayesian simulations have captured most secondary extinctions identified using dynamical simulations.
+This suggests that, in many cases, bottom-up effects are the main drivers of secondary extinction.
+We also emphasize the major gain in computational speed when using the Bayesian approach, which allows us to investigate different levels of disturbance rather than simple species removals.
 
@@
         \them
-Line 182: what is network mean persistence? Species persistence is the probability of species persistence referred to in the methods? Be explicit. Also, what are the “other network properties”?
+        Line 182: what is network mean persistence?
+Species persistence is the probability of species persistence referred to in the methods?
+Be explicit.
+Also, what are the “other network properties”?
 
@@
 
-\us This concept is now introduced in the introduction [[Disturbance is introduced, but different effects of different severities not clear yet.]] [[but should we discuss it a bit more also in the statistical analysis?
-we mention the different disturbance levels in the end of the methods section Calculating persistence]]
+        \us This concept is now introduced in the introduction.
+We give examples from other theoretical studies and empirical studies briefly indicating how intensity of disturbance can affect extinction risk.
+We also explicitly mention disturbance level in our objectives paragraph and at the start of our statistical analysis section.
+However, we do note that disturbance level was included as a predictor in the equations in our previous version of the manuscript.
+We hope that the current revision makes this parameter a more obvious part of our approach.
 
@@
 
-\us We explicitly control for differences in the number of motifs through the frequency normalisation discussed above.
+        \us
+        We explicitly control for differences in the number of motifs through the frequency normalisation discussed above.
 We include the random intercepts to account for potential effects of Sn and Cn on network persistence that are \emph{not} due to effects on the total number of motifs.
@@
 
-\us See response to XX ... probably need to repeat that our focus is not to throw away degree [[Definitely need to clarify our modelling framework]]
+        \us See response to the Reviewer's second major point, above.
+We reiterate that our goal is not to prove that motifs should supplant other network metrics for predicting extinction risk.
+Rather, we want to test the proposal in earlier motif-based papers that motifs are related to extinction risk \emph{as well as} other network properties.
+Model selection is therefore not of particular interest for our purposes.
 
@@
 
-\us Annoying but maybe possible
+        \us We have added error bars representing 95\% confidence intervals to this figure.
 
@@
 
-\us Empirical comparison should take care of this. ... citations to connectance range papers.
+        \us While we do not use a generative model that fills some motifs/interactions before others, we do not see any particular reason why apparent competition motifs should be filled first.
+We can confirm, however, that the connectances we explore are realistic.
+We compare the structure of our simulated networks to the largest set of high-quality, similarly-sized empirical networks we could find.
+These networks had 50-77 species and connectances of 0.2-0.214.
+This is at the upper range of connectances we consider (probably due to aggregation of basal resources into taxonomically coarse nodes in the empirical webs).
+We also, per the Associate Editor's request, compare the motif profiles and motif participation of these empirical webs to our simulated webs.
+We hope that this comparison will demonstrate the realism of our simulated networks.
 
@@
 
-\us We are not aware of any other motif-focused papers calling this motif ``exploitative competition'' and follow the nomenclature from the introduction of motifs to food webs~\citep{Stouffer2007}.
+        \us We are not aware of any other motif-focused papers calling this motif ``exploitative competition''.
+Our terminology follows the nomenclature from the introduction of motifs to food webs~\citep{Stouffer2007}.
         Note that ``direct competition'' refers simply to two species which consume the same prey and is agnostic as to whether this is exploitative or interference competition. 
+        As exploitative and interference competition are not the same ecologically, we have maintained the conventional motif terminology so as not to imply an overly specific interpretation.
 
@@
 
-Comments to the Author(s)
+    \them
     The authors present an interesting study about how motif based measures can improve our understanding of species persistence in food webs after bottom up disturbances.
@@
 
-\us TBD after answer to reviewer 1
+    \us We have thoroughly revised the introduction in response to Reviewer 1's comments and, in the process, have condensed the sections referring to secondary extinctions and reworked the sections referring to different scales.
+We have also somewhat expanded our descriptions of different scales and how they can affect response to description.
+We hope that local-scale and global-scale properties now more obviously justify their own paragraphs.
 
@@
 
-\us Same figure requested by Editor, good idea to add
+    \us  We thank the Reviewer for following our methods and analysis and are glad that they match our goals.
+We have added a conceptual figure showing the four motifs within a food web, with particular attention to how species with the same degree and trophic level can have different motif participation.
+We hope that this makes the text as a whole more approachable.
 
@@
 
+    \subsubsection*{Minor revisions}
+
         \them
-Minor revisions
-
         Line 16: I suggest changing the word “vulnerability” to “species persistence” to keep consistency along the manuscript.
@@
         Line 23-26: this sentence could be clearer.
-Perhaps authors could rephrase to: “The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole (Proulx et al., 2010; Scherber et al., 2010, Rosenblatt & Schmitz, 2016).
+Perhaps authors could rephrase to: “The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole (Proulx et al., 2010; Scherber et al., 2010, Rosenblatt \& Schmitz, 2016).
 For example, the loss of diversity…”
@@
 
-\us Alyssa: Expand explanation of why high connectance can be good sometimes (more prey available to switch to, albeit more disturbance pathways).
-Probably a good way to introduce size of disturbance - many pathways of a small disturbance may not matter, many pathways of a large more severe [[I think the current version does a succinct job of showing the good and bad of connectance]]
+        \us Connectance can be a double-edged sword depending on the type of disturbance and structure of the rest of the web.
+Although higher connectance does allow more pathways for a disturbance to propagate, on average a higher connectance also means that each species has more prey.
+If species can switch between prey, this can allow consumers to persist until the effects of the disturbance has abated.
+This is easiest to understand when the focal disturbance is species removal.
+In a low-connectance network, the consumer(s) of a removed species are less likely to have other resources.
+If a consumer's only prey is removed, it must necessarily go secondarily extinct.
+In a high-connectance network where most species have many prey, the consumer's fate is much less certain following a single-species removal.
+We appreciate the prompt to clarify these opposing potential effects of connectance, and have added a reference to the Vieira \& Almeida-Neto paper suggested.
+These lines now read:
+
+            \begin{quotation}
+                 Highly-connected networks may be more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} because consumers, on average, have access to a larger number of prey species and can switch from a declining prey to another resource.
+However, a well-connected network also includes more pathways through which a disturbance can spread, affecting more species and potentially causing more secondary extinctions~\citep{Vieira2015}.
+                    The net effect of connectance on extinction therefore likely depends on the severity of disturbance and the ability of consumers to subsist on a reduced set of prey.    
+            \end{quotation}
 
@@
         \them
-Line 161: Perhaps authors might add a figure of the four three-species motifs in the appendix explaining the ecological meaning of each structure. I think this could help readers to understand motifs.
+        Line 161: Perhaps authors might add a figure of the four three-species motifs in the appendix explaining the ecological meaning of each structure.
+I think this could help readers to understand motifs.
 
-\us Same figure as above, if possible?
+        \us We have added these motifs to the conceptual figure described above and hope that this helps readers understand.
 

### 2022-06-02 manuscript/supplement_motifs.tex

@@
         For this comparison, we obtained all networks containing 50-100 species from \citet{globalwebdb}.
+        These webs ranged in connectance from 0.024-0.089
         We then removed networks where sets of predators and prey were completely different (i.e., bipartite networks) or links were not clear (i.e., +/- effects rather than predation specifically).
         This yielded five vertebrate-dominated food webs describing the California coast in different years, four invertebrate-dominated food webs describing mainland US streams, 23 invertebrate-dominated food webs describing New Zealand streams, and five other invertebrate-dominated food webs.
+        This final set of webs had 50-77 species and connectances of 0.2-0.21 - the upper range of connectance in our simulations.
+        This may be partly due to relatively poor resolution of basal resources in these empirical webs: aggregation of many taxa into nodes such as `phytoplankton' will tend to decrease the number of species but leave the number of links unchanged, raising connectance.
         We note that most of these webs were compiled by just three research groups.

### 2022-06-07 manuscript/response_PRSB_round2.tex
Updates from Overleaf

@@
 
-        \us We agree that there is a strong correlation between apparent competition and in-degree, and state that degree appears able to explain the trend in apparent competition.
-However, not all motifs show such strong relationships, meaning that degree is not sufficient to
+        \us The Reviewer is correct that there is a strong  correlation between apparent competition and in-degree, as we state in the text.
+It is also true that apparent competition and in-degree show opposite relationships with persistence, such that the apparent competition-persistence relationship could be described by the in-degree-persistence relationship.
+However, as detailed later in the paragraph, this ignores the strong correlation between apparent competition and trophic level which would suggest that persistence should show different relationships to apparent competition.
+This means that either trophic level is unrelated to persistence (which we know is not true) or some other factor is over-riding the ``effect" (there is no clear causality from trophic level to motifs or vice versa) of trophic level on apparent competition.
 
-        Clarify our writing ... is there a way to connect the figures better?
+        We have revised this paragraph to stop readers from jumping to conclusions before reading the entire section.
+
+        \begin{quotation}
+            For apparent competition, there were strong and opposing correlations with in-degree and trophic level but the relationship between apparent competition and persistence only matches the trends for in-degree.
+            This suggests that some other factor is over-riding the relationship between apparent competition and trophic level.        
+        \end{quotation}
 

### 2022-06-07 manuscript/reorder_main_text.tex
trying to fix R1s poor reading comprehension
Updates from Overleaf

@@
     The presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
-    Highly-connected networks may be more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} because consumers, on average, have access to a larger number of prey species.
-However, in a well-connected network a disturbance may spread more easily and cause more secondary extinctions due to the presence of many possible pathways \citep{}.
-    The net effect of connectance on extinction therefore likely depends on the severity of disturbance and the ability of consumers to subsist on a reduced set of prey.
+    Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2020}.
+However, a well-connected network also includes more pathways through which a disturbance can spread~\citep{Dunne2004,Vieira2015}.
+    Depending on the strength and arrangements of interactions, this means that species may be impacted through several paths, amplifying the original disturbance and causing more secondary extinctions.
+    The net effect of connectance on extinction therefore likely depends on finer-scale network structure, as well as the type and severity of disturbance.
 
@@
     While the potential relationships between motif participation and species-level extinction risk have not yet been tested to our knowledge, it has been shown that the frequencies of different motifs within a whole network are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
-    Given these relationships between network-level motif frequencies and overall community stability, it is worth investigating whether a species' participation in different motifs also affects its extinction risk.
-    We are also interested in whether the relationships between network structure and extinction risk depend on the severity of disturbance experienced by the community.
+    These relationships between network-level motif frequencies (`motif profiles') and overall community stability mean that a network's motif profile could affect the average extinction risk for consumers within the network.
+    Moreover, if network motif profiles are related to average extinction risk, it is worth investigating whether a species' participation in different motifs also affects its extinction risk.
 
-    Naturally, a larger disturbance tends to lead to more secondary extinctions \citep{}.
-    Disturbances to plant species can also affect some consumers more strongly than others  \citep{byrnes2011climate}. 
-    In empirical studies of kelp forests, the magnitude of the disturbances to basal resources can alter the competition between species in the benthic community \citep{detmer2021variation} and change the percent coverage of different taxonomic groups \citep{carnell2020more}.
-
-    %While these studies uncovered the potential influence of network-level motif frequencies on network stability, a disturbance may likely alter the network and the interactions between species in different ways.
-    %This can affect the influence of motif frequencies on network stability, as it has been demonstrated that omnivory may appear stabilizing dependent on interactions with other species \citep{Kratina2012}.
+    As mentioned above, the severity of a disturbance may interact with network structure to affect secondary extinctions.
+    A mild disturbance (e.g., a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
+    Empirical studies have shown that disturbances to plant species can also affect some consumers more strongly than others  \citep{byrnes2011climate}, and that the strength of disturbance also affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}.
+    By taking advantage of recent advances in rapid simulation of ecological networks, we are able to simulate a range of disturbance intensities and test whether meso-scale network structure also interacts with disturbance.
 
@@
 %We address this in scenarios with various levels of disturbance targeting the basal level.
-    In addition, we want to understand how the information provided by motifs is related to global and local food web properties with known relationships to extinction risk in order to better synthesize information provided by different scales of network structure.
+    In addition, we want to understand how the information provided by motifs is related to other global and local food web properties in order to better synthesize information provided by different scales of network structure.
     We model secondary extinction probabilities using Bayesian networks \citep{Eklof2013, Haussler2020}. 
     We focus our analyse on the four network motifs that are over-represented in empirical networks and additionally have been shown to contribute to network stability; the three-species chain, omnivory, apparent competition, and direct competition \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
-
-     %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
 
@@
 		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
-		For full details, se \emph{Appendix S1}.
+		For full details, see \emph{Appendix S1}.
 		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
@@
         However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
-        This implies that bottom-up effects drive most secondary extinctions in simulations, echoing the strong bottom-up effects observed in empirical studies~\citep{}.
         [[Something about strong bottom-up effects in empirical studies]]
@@
 
-        We are primarily interested in whether a consumer's extinction risk is related to its motif participation or the motif profile of the network and, secondarily, also wish to connect the information provided by motifs to global and local network properties that have known effects on extinction risk.
-    	To address these questions, we tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (in-degree, STL, S, C, and level of disturbance).
+    	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, in-degree, and trophic level).
+        Where applicable, we also consider how interactions between network structure and disturbance level might affect extinction risk.
     	See Appendix \emph{3.1 - 3.5} for details. 
@@
 
-            Our main question of interest is whether appearing in different motifs is associated with differences in extinciton risk. 
             To test if participating more often in a certain motif affected a species' probability of persistence, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks):
@@
 
-            Properties which refer to different scales of network structure (e.g., local, global, and meso-scale) are not independent. 
-            This means that, to understand relationships between motif participation (a meso-scale network property) and extinction risk, we must also consider how motif participation varies with other properties known to affect extinction risk.
-            To identify correlations between global network structure (i.e., the whole network) and participation in particular motifs, we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
+            Properties which refer to different scales of network structure, including different amounts of direct and indirect interactions with the focal species, are not independent. 
+            To identify correlations between global network structure (i.e., the whole network) and participation in particular motifs (a meso-scale description of species roles), we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
             We fit a similar set of  LMMs relating motif participation to in-degree or trophic level (i.e., local measures of a species' role; Appendix \emph{S3.2}).
@@
 
-            As well as relating species-specific probabilities of extinction to species-specific motif participation, we are interested in whether the overall frequency of motifs in a network provides information about the overall probability of any species going extinct.
-            Just as other whole-network properties (S, C, etc.) are known to affect the overall stability of a network, it is possible that the arrangement of interactions into different motifs could reflect different average probabilities of persistence across all consumers.
+            As well as species-specific probabilities of persistence, the overall frequency of motifs in a network may provide information about the overall probability of any species going extinct (since they incorporate information on indirect effects).
             We tested whether networks with similar motif profiles tend to have similar mean persistences among consumers using a PERMANOVA (Appendix \emph{S3.3}).
@@
 
-            Just like motif participation, a network's motif profile is not likely to be independent of other levels of structure. 
-            To formally test whether networks with the the same size and connectance have more similar motif profiles, we used a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
-            In addition, to understand how motif profiles may vary with global network structure in more detail, we identify correlations between size, connectance, and particular motifs using a set of four LMs:
+            To test whether networks with the the same size and connectance have more similar motif profiles, we used a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
+            In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMs:
             \begin{equation}

### 2022-06-07 manuscript/reorder_main_text.tex
minor

--- 2022-06-02 manuscript/reorder_main_text.tex
@@
 \medskip
-$^\dagger$ Corresponding author:\\
-
+$^\dagger$ Corresponding author\\
 }
@@
     In these cases, in-degree and trophic level are not sufficient to explain the trends we observe in the motifs.
-    Participation in the apparent competition motif was strongly negatively correlated with in-degree and shows a similar relationship to persistence.
-    However, participation in the apparent competition motif was also strongly and positively correlated with trophic level, and the negative effect of high trophic levels on persistence is not reflected in the relationship between persistence and participation in the apparent competition motif.
+    For apparent competition, there were strong and opposing correlations with in-degree and trophic level but the relationship between apparent competition and persistence only matches the trends for in-degree.
+    This suggests that some other factor is over-riding the relationship between apparent competition and trophic level.
 

### 2022-06-07 manuscript/reorder_main_text.tex

@@
     The presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
-    Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2020}.
+    Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2015}.
 However, a well-connected network also includes more pathways through which a disturbance can spread~\citep{Dunne2004,Vieira2015}.
@@
         \centering
-        \includegraphics[width=0.85\textwidth]{figures/prop_lmer_allCS.pdf}
-        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).
-The effect of participating in each motif is based on linear mixed-effect models following Equation~\ref{propreq}, except that models for each  disturbance level are fitted separately.
-The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance).
-Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
-    \label{fig:prop_lmer_all}
+        \includegraphics[width=0.85\textwidth]{manuscript/figures/concept_motif.png}
+        \caption{Conceptual figure of a toy food web showing how species with similar degree and trophic level can have different motif participation. }
+    \label{fig:concept}
     \end{figure}
@@
         \centering
-        \includegraphics[width=0.85\textwidth]{figures/persistence_vs_motifs.eps}
+        \includegraphics[width=0.85\textwidth]{figures/persistence_vs_motifs_labs.eps}
         \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).

### 2022-06-07 manuscript/reorder_main_text.tex

@@
         Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
-        This increased speed allows us to simulate a range of disturbance severity rather than single species removals as in many previous studies~\citep{}.
-        Moreover, Bayesian network simulations are less parameter-intensive than dynamical models, meaning that they are less sensitive to assumptions made in the modelling process~\citep{}.
+        This increased speed allows us to simulate a range of disturbance severity rather than single species removals as in many previous studies~\citealp[e.g.,]{Memmott2004,Staniczenko2010,Dunne2004,Cirtwill2021_inprep}.
+        Moreover, Bayesian network simulations are less parameter-intensive than dynamical models, meaning that they are less sensitive to assumptions made in the modelling process~\citep{Eklof2013}.
 
@@
         However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
-        [[Something about strong bottom-up effects in empirical studies]]
+        Moreover, bottom-up effects are often strong in empirical systems.
+        For example, bottom-up effects regulate zoo-plankton communities in freshwater lakes, regulating zoo-plankton communities \citep{li2020bottom} and control abundances and biomass of vertebrate species in several terrestrial systems \citep{Dobson2009food, Mduma1999food, Georgiadis2007}.
+
     	To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).
@@
         \centering
-        \includegraphics[width=0.85\textwidth]{manuscript/figures/concept_motif.png}
-        \caption{Conceptual figure of a toy food web showing how species with similar degree and trophic level can have different motif participation. }
+        \includegraphics[width=1.0\textwidth]{manuscript/figures/concept_motifs_ab.png}
+        \caption{Conceptual figure of a toy food web.
+\textbf{a)} This web contains examples of the four motifs present in an acyclic Bayesian network; apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow).
+Note that the nodes participate in more motifs than the ones highlighted.
+\textbf{b)} In the same food web, species with similar degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than tree times more frequently.
+Inserted table shows the number of times A and B, respectively, participate in the four different motifs.}
     \label{fig:concept}

### 2022-06-08 code/Fig1_errorbars.py
changed errorbar colours

@@ -30,6 +30,13 @@ colors.add_color(253,174,97,'Rorange1')
 colors.add_color(244,109,67,'Rorange2')
 colors.add_color(213,62,79,'Rred1')
 colors.add_color(158,1,66,'Rred2')
+colors.add_color(204,199,242,'Gpurple')
+colors.add_color(210,246,249,'Gblue')
+colors.add_color(192,244,235,'Ggreen')
+colors.add_color(211,250,204,'Ggreen2')
+colors.add_color(240,255,222,'Ggreen3')
+colors.add_color(245,245,221,'Gyellow')
+
 
 # colors.add_color(120,120,120,'grey')
 # colors.add_color(255,125,125,'lightish_red')
@@ -127,7 +134,7 @@ def populate_graph(graph,motif,lmdict):
     uperr=graph.add_dataset(upper)
     uperr.symbol.shape=0
     uperr.line.linestyle=0
-    uperr.fill.configure(type=2,color=4,pattern=1)
+    uperr.fill.configure(type=2,color=j+11,pattern=1)
 
     derr=graph.add_dataset(lower)
     derr.symbol.shape=0

### 2022-06-08 manuscript/reorder_main_text.tex

@@
     %\subsection*{1 Why are bottom-up effects important for understanding stability?}
-     The plant community is the foundation upon which the myriad species in terrestrial food webs depend for their survival.
+     The plant community is the foundation upon which the myriad species in food webs depend for their survival.
 Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
@@
 \citep{scherber2010bottom}.
-Secondary extinctions and plants...
 
     %\subsection*{2 Global and local network properties are known to affect stability}
-    How disturbances, such as population decline or extinction, lead to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    How disturbances, such as population decline or extinction, lead to loss of other species via direct and indirect effects (i.e., secondary extinctions) has for long been vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
     Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances.
@@
     Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
-    The presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
+    The undisputable presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
     Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2015}.
-However, a well-connected network also includes more pathways through which a disturbance can spread~\citep{Dunne2004,Vieira2015}.
+However, a well-connected network also includes more pathways through which a disturbance can spread~\citep{Dunne2004, Vieira2015}.
 Depending on the strength and arrangements of interactions, this means that species may be impacted through several paths, amplifying the original disturbance and causing more secondary extinctions.
@@
 
-    Local properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
-    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, including secondary extinction due to reduced plant populations.
+    Species properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
+This can be referred to local properties in the network.
+    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and secondary extinctions due to reduced plant populations are particularly detrimental \citep{Curtsdotter2011}.
+%including secondary extinction due to reduced plant populations.
     However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
-    As well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
-    For example, a generalist consuming many specialist (and therefore high-risk) prey may, for example, be at greater risk of extinction than a specialist which consumes a highly generalist resource that is very unlikely to go extinct.
+    This since, as well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
+    For example, a generalist consuming many specialist (and therefore high-risk) prey may, for example, be at greater risk of extinction than a specialist which consumes a highly generalist resource that is more unlikely to go extinct.
 
@@
     Such motif profiles of species are easily calculated once we know the network structure and require no more information than is used to calculate degree.
-    These `motif participation' roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
+    These `motif participation' roles capture information about species' direct as well as indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
     By allowing researchers to distinguish between, for example, consumers of specialists and generalists, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
@@
 
+% Severity of disturbance
     As mentioned above, the severity of a disturbance may interact with network structure to affect secondary extinctions.
-    A mild disturbance (e.g., a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
+    A mild disturbance (e.g., leading to a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
     Empirical studies have shown that disturbances to plant species can also affect some consumers more strongly than others  \citep{byrnes2011climate}, and that the strength of disturbance also affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}.
-    By taking advantage of recent advances in rapid simulation of ecological networks, we are able to simulate a range of disturbance intensities and test whether meso-scale network structure also interacts with disturbance.
 
@@
     In addition, we want to understand how the information provided by motifs is related to other global and local food web properties in order to better synthesize information provided by different scales of network structure.
-    We model secondary extinction probabilities using Bayesian networks \citep{Eklof2013, Haussler2020}. 
+    We use synthetic networks and model secondary extinctions using Bayesian networks \citep{Eklof2013, Haussler2020}.
+As Bayesian networks are computationally efficient (see Methods), we are able to simulate a wide range of disturbance intensities and test whether meso-scale network structure also interacts with severity of the disturbance.
     We focus our analyse on the four network motifs that are over-represented in empirical networks and additionally have been shown to contribute to network stability; the three-species chain, omnivory, apparent competition, and direct competition \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
@@
 
-	\subsection*{Initial network construction}
-
-		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}.
-		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
-		For full details, see \emph{Appendix S1}.
-		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
-
     \subsection*{Modelling secondary extinction using Bayesian networks}
@@
 Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network.
-        This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are more parameter intensive, and simulations are more time-consuming, than topological models.
+        This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are extensively more parameter intensive, and simulations are much more time-consuming, than topological models.
 
-        A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks \citep{Eklof2013}. 
-        In this framework, a consumer's probability of extinction depends on the fraction of resources lost and on a baseline probability of extinction which captures causes unrelated to the network structure (e.g., disease, stochastic extinction of small populations).
+        A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks \citep{Eklof2013, Haussler2020}. 
+        In this framework, a consumer's probability of extinction depends on the fraction of resources lost and on a baseline probability of extinction which captures causes affecting the respective species extinction probability that are unrelated to the network structure (e.g., disease, stochastic extinction of small populations).
         Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
-        This increased speed allows us to simulate a range of disturbance severity rather than single species removals as in many previous studies~\citealp[e.g.,]{Memmott2004,Staniczenko2010,Dunne2004,Cirtwill2021_inprep}.
+        This increased speed allows us to simulate a range of disturbance severity rather than single species removals as in many previous studies~\citealp[e.g.,]{Memmott2004,Staniczenko2010,Dunne2004}.
         Moreover, Bayesian network simulations are less parameter-intensive than dynamical models, meaning that they are less sensitive to assumptions made in the modelling process~\citep{Eklof2013}.
 
-        As consumers do not affect the extinction probabilities of their resources, Bayesian networks can only capture bottom-up effects of disturbances within networks. 
+       % In Bayesian networks consumers do not affect the extinction probabilities of their resources, and as such we can here only capture bottom-up effects of disturbances.
 However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
@@
 
-    	To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).
-    	These processing steps allow the strictly bottom-up calculation of probability of persistence.
-
-	\subsection*{Calculating persistence}	
+	%\subsection*{Calculating persistence}	
     % basic idea behind BN modeling
@@
 
-		Modelling species extinctions in food webs using Bayesian networks has two main limitations. First, due to the strict bottom-up approach the food webs must be acyclic. 
-		Second, as consumers are influenced by the state of their resources, but not vice versa, the method does not capture top-down effects.
+		% BN limitations
+		Modelling species extinctions in food webs using Bayesian networks has two main limitations.
+First, in Bayesian networks consumers do not affect the extinction probabilities of their resources, and as such we can here only capture bottom-up effects of disturbances.
+However, Bayesian networks has been shown to capture the majority of secondary extinctions produced fully-fledged dynamical models \citep{Eklof2013}.
+Moreover, bottom-up effects are often the main force of secondary extinctions in empirical systems.
+For example, bottom-up effects regulate zoo-plankton communities in freshwater lakes, regulating zoo-plankton communities \citep{li2020bottom} and control abundances and biomass of vertebrate species in several terrestrial systems \citep{Dobson2009food, Mduma1999food, Georgiadis2007}.
+Second, due to the strict bottom-up approach of Bayesian networks the networks must be acyclic.
+Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited.
+Thus, making the networks acyclic in order to use the Bayesian network approach will not strongly affect extinction scenarios.
 
-    % various disturbance scenarios
-        In our baseline scenario, basal resources had an extinction probability of $\pi = 0.1$. 
-		To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase the affected species' probability of extinction from $0.10$ to $0.5$, in steps of $0.08$.
+    \subsection*{Disturbance scenarios}
+        In our baseline scenario, all species had an extinction probability of $\pi = 0.1$. 
+		To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase the affected basal species' probability of extinction from $0.10$ to $0.5$, in steps of $0.08$.
 		The highest disturbance level, $\pi = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
 		Consumer species retained $\pi=0.1$ in all cases.
+
+	\subsection*{Initial network construction}
+
+		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}.
+		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
+		For full details, see \emph{Appendix S1}.
+		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
+
+    	To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and species topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).
+These processing steps allow the strictly bottom-up calculation of probability of persistence.
 
@@
 
+ %   \begin{figure}[hb!]
+ %      \centering
+ %       \includegraphics[width=1.0\textwidth]{manuscript/figures/concept_motifs_ab.png}
+ %       \caption{Conceptual figure of a toy food web.
+\textbf{a)} This web contains examples of the four motifs present in an acyclic Bayesian network: apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow).
+These motifs capture different ways in which disturbances to basal resources can propagate through the network.
+For example, a disturbance to the bottom species in a three-species chain will have a direct effect on the middle species but only indirectly affect the top species while a disturbance to the bottom species in an omnivory motif will have a direct effect on the middle species and direct and indirect effects on the top species.
+Note that the nodes participate in more motifs than the ones highlighted.
+\textbf{b)} In the same food web, species with similar degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than tree times more frequently.
+Inserted table shows the number of times A and B, respectively, participate in the four different motifs.}
+ %   \label{fig:concept}
+ %   \end{figure}
+
         \begin{figure}[hb!]
         \centering
-        \includegraphics[width=1.0\textwidth]{manuscript/figures/concept_motifs_ab.png}
+        \includegraphics[width=1.0\textwidth]{manuscript/figures/concept_fig_ver2.pdf}
         \caption{Conceptual figure of a toy food web.
-\textbf{a)} This web contains examples of the four motifs present in an acyclic Bayesian network; apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow).
+\textbf{a)} This web contains examples of the four motifs present in an acyclic Bayesian network: apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow).
+These motifs capture different ways in which disturbances to basal resources can propagate through the network.
+For example, a disturbance to the bottom species in a three-species chain will have a direct effect on the middle species but only indirectly affect the top species while a disturbance to the bottom species in an omnivory motif will have a direct effect on the middle species and direct and indirect effects on the top species.
         Note that the nodes participate in more motifs than the ones highlighted.

### 2022-06-08 manuscript/response_PRSB_round2.tex
Updates from Overleaf

@@
 
-\subsection*{Associate Editor} [[Add conceptual figure, review Alyssa's new text]]
+\subsection*{Associate Editor} [[review Alyssa's new text]]
     \them
@@
 We hope that this makes the text as a whole more approachable.
-    Anna E
 
@@
 As the Reviewer notes, a subset of the authors of this manuscript are working on an analysis of the relationships between motif participation and persistence in a dynamical framework (currently under review).
-We do not believe that including the material from this second paper in an appendix (where few people will see it) is justified when there are already several comparisons of extinctions simulated by Bayesian and dynamical approaches (e.g.,~\citealp[]{Eklof2013a,others}).
+We do not believe that including the material from this second paper in an appendix is justified when there are already a published comparisons of extinctions simulated by Bayesian and dynamical approaches (\citealp[]{Eklof2013a}).
+That paper focused on comparing the number and identity of secondary extinctions between the two approaches, and additionally evaluated which functional response of consumers to loss of prey that most accurately predicted the secondary extinctions from a fully-fledged dynamical model.
+We use this information in the current manuscript.
 
@@
 In particular, motifs give information about a consumer's prey that may reflect the prey's risk of extinction and thereby the predators' risk (e.g., a predator consuming generalist prey will participate in many three-species chain motifs).
+We have also added a conceptual figure illustrating the four motifs and explaining how different motifs imply different pathways for bottom-up disturbances.
 
@@
 We hope that this figure and the revised introduction are more persuasive.
-
-        % [[Make sure conceptual figure shows how species with same degree, TL can participate in different motifs and have different risks - consumer with degree=2 consuming two specialists vs.
-consuming two generalists should make the point]]
 
@@
         \us In revising the introduction, we now introduce the Bayesian network approach and its limitations in the methods section.
-There, we clearly state that the lack of top-down effects in the Bayesian networks is a limitation.
+We want to stress that the Reviewer is clearly correct that these are important limitations and we have further clarified that in the text.
+At the same time we have clarified the clear advantages that the Bayesian network approach provides.
+        First, we clearly state that the lack of top-down effects in the Bayesian networks is a limitation.
 However, we also point out that in prior studies Bayesian simulations have captured most secondary extinctions identified using dynamical simulations.
 This suggests that, in many cases, bottom-up effects are the main drivers of secondary extinction.
-We also emphasize the major gain in computational speed when using the Bayesian approach, which allows us to investigate different levels of disturbance rather than simple species removals.
+We also emphasize the major gain in computational speed when using the Bayesian approach, which allows us to investigate different levels of disturbance rather than simple species removals that needs to be repeated numerous times.
+Finally, we have also added text to clarify how the process of making the food webs acyclic can affect the results, but that food web robustness rarely is affected.
 

### 2022-06-10 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
     Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
-    The undisputable presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
+    The indisputable presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
     Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2015}.
@@
     Species properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
-This can be referred to local properties in the network.
+This can be referred to local properties of the network.
     Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and secondary extinctions due to reduced plant populations are particularly detrimental \citep{Curtsdotter2011}.
@@
     It is possible to capture information about how a consumer and its prey fit together within a community by defining participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
-    Such motif profiles of species are easily calculated once we know the network structure and require no more information than is used to calculate degree.
+    Such motif profiles of species are easily calculated once we know the network structure and require no more information than is used to calculate degree or trophic level.
     These `motif participation' roles capture information about species' direct as well as indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
-    By allowing researchers to distinguish between, for example, consumers of specialists and generalists, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
+We refer to these motifs as meso-scale structures, to emphasise their role as structures between global (network) and local (species) properties.
+A species' motif participation profile shows the second step indirect interactions the species are affected by, and thereby capture important structures that neither global or local properties do.
+    Therefore, by allowing researchers to distinguish between, for example, consumers of specialists and generalists, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
 
@@
     A mild disturbance (e.g., leading to a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
-    Empirical studies have shown that disturbances to plant species can also affect some consumers more strongly than others  \citep{byrnes2011climate}, and that the strength of disturbance also affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}.
+    Empirical studies have shown that disturbances to plant species can affect some consumers more strongly than others  \citep{byrnes2011climate}, that the strength of disturbance affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}, and that the resilience and return time of the ecological community can be strongly influenced by disturbance severity \citep{rydgren2004disturbance}.
+The severity disturbance therefore clearly has an effect on how the ecological community will respond.
 
@@
 However, Bayesian networks has been shown to capture the majority of secondary extinctions produced fully-fledged dynamical models \citep{Eklof2013}.
+\cite{Eklof2013} analyzed which functional form of consumers response to loss of resources that most accurately predicted the secondary extinctions realized in the dynamical model, and we here use that functional form in our analyzes.
 Moreover, bottom-up effects are often the main force of secondary extinctions in empirical systems.
@@
 
-		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}.
+		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}.
 		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
-		For full details, see \emph{Appendix S1}.
+		For full details and a comparison to empirical networks, see \emph{Appendix S1}.
 		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
@@
 
-    	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, in-degree, and trophic level).
+    	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and trophic level, STL), 3) how network mean persistence (the average persistence of all consumers in the network) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, in-degree, and trophic level).
         Where applicable, we also consider how interactions between network structure and disturbance level might affect extinction risk.
@@
 
-        At high levels of disturbance, a species' persistence decreased as the proportion of the three-species chain increased, but persistence increased as the proportion of any other motif increased (Figure~\ref{fig:prop_lmer_all}; lower lines in each panel).
-        This means that increased disturbance strengthened the effects of the three-species chain motif and changed the direction of the effect of the apparent competition motif  (Table S1, \emph{Appendix S4}).
+        At high levels of disturbance, a species' persistence decreased as the proportion of the three-species chain in its motif profile increased.
+Persistence increased as the proportion of any other motif increased (Figure~\ref{fig:prop_lmer_all}; lower lines in each panel).
+        This means that increased disturbance strengthened the effects of the three-species chain motif and changed the direction of the effect of the apparent competition motif  (Figure~\ref{fig:prop_lmer_all}, Table S1, \emph{Appendix S4}).
         The relationship between omnivory and persistence was still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.
@@
         The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=645, $p$\textless0.001), although this result must be considered tentative due to non-homogeneous distribution of roles (\emph{Appendix S8}).
-        Average persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S4, \emph{Appendix S8}). 
+        Average persistence was higher in networks with higher proportions of the three-species chain, direct competition, and apparent competition motifs (Table S4, \emph{Appendix S8}). 
         Conversely, average persistence was lower in networks with higher proportions of the omnivory motif. 
@@
     As well as clarifying the relationships between network structure and the effect of bottom-up disturbance on particular species, motifs were also related to the mean probability of persistence across all species in a network.
-    In fact, a network's motif profile had a stronger relationship to mean persistence compared to network size or connectance. 
+    In fact, a network's motif profile had a stronger relationship to mean persistence compared to both network size and connectance. 
     This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.
@@
     % High C webs and omnivory
-    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks and the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}) as in~\citep{McLeod2021}.
+    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks and the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}) as in~\cite{McLeod2021}.
     These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
@@
     In summary, our results show that participating in different three-species motifs is related to a species' risk of extinction following a bottom-up disturbance, and that these relationships synthesize information provided by other, simpler measures of species roles.
-    Our results also underscore the importance of considering the level of disturbance when identifying which species are most at risk of extinction.
+    Most interestingly, our results also underscore the importance of considering the level of disturbance when identifying which species are most at risk of extinction.
     There were strong interactions between the level of disturbance and the relationships between persistence and multiple measures of species roles (motifs and in-degree).
@@
         \centering
-        \includegraphics[width=1.0\textwidth]{manuscript/figures/concept_fig_ver2.pdf}
+        \includegraphics[width=1.0\textwidth]{manuscript/figures/concept_fig_ver2.png}
         \caption{Conceptual figure of a toy food web.
-\textbf{a)} This web contains examples of the four motifs present in an acyclic Bayesian network: apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow).
+\textbf{a)} This food web contains examples of the four motifs (present in an acyclic directed networks): apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow).
 These motifs capture different ways in which disturbances to basal resources can propagate through the network.
@@
 \textbf{b)} In the same food web, species with similar degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than tree times more frequently.
-Inserted table shows the number of times A and B, respectively, participate in the four different motifs.}
+Inserted table shows the number of times A and B, respectively, participate in the four different motifs.
+This example highlight that species A may then have a higher risk of going secondary extinct due to its dependence on specialist prey species, compared to species B who depends om generalist prey species.
+This relationships are captured by the species motif profiles, but not their trophic level or degree.}
     \label{fig:concept}
@@
     \label{fig:prop_lmer_all}
+    \end{figure}
+
+        \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=0.85\textwidth]{figures/persistence_vs_motifs.eps}
+        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).
+The effect of participating in each motif is based on the fixed effects in Equation~\ref{propreq}.
+The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance).
+95\% confidence intervals for each line are shown in grey.
+Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
+    \label{fig:prop_lmer_2}
     \end{figure}

### 2022-06-10 manuscript/response_PRSB_round2.tex

@@
 That paper focused on comparing the number and identity of secondary extinctions between the two approaches, and additionally evaluated which functional response of consumers to loss of prey that most accurately predicted the secondary extinctions from a fully-fledged dynamical model.
-We use this information in the current manuscript.
+We use this information in the current manuscript and make our simulations with the functional form that was shown to most accurately predict the same secondary extinctions as the dynamical model.
+We have now calrified this in the manuscript (line XX).
+However, we would also like to point out that there are no guarantees that advanced (as in parameter intensive) dynamical models describe the ecological reality better, as comparisons with extensive empirical extinction scenarios are still lacking.
 
-        In addition, we feel that the Reviewer has over-estimated our reliance on the BioRxiv pre-print currently under review.
+        In addition, we believe that the Reviewer has over-estimated our reliance on the BioRxiv pre-print currently under review.
 Our previous draft had only three references to this pre-print: when describing the generation of simulated networks, when noting that degree and motif participation are not independent, and when briefly comparing the dynamical and Bayesian results in the discussion.
-        To address this mis-perception, we have now removed two of these references.
+        We have now removed two of these references.
         We present the full network-generation methods in \emph{Appendix 1} and, to avoid confusion by referring to results presented in the preprint, have softened our language to say that degree and motif participation \emph{may} be correlated.
@@
 We want to stress that the Reviewer is clearly correct that these are important limitations and we have further clarified that in the text.
-At the same time we have clarified the clear advantages that the Bayesian network approach provides.
+At the same time we have clarified the advantages that the Bayesian network approach provides.
         First, we clearly state that the lack of top-down effects in the Bayesian networks is a limitation.
@@
 
-        We have revised this paragraph to stop readers from jumping to conclusions before reading the entire section.
+        We have revised this paragraph with particular focus on these relationships and hope that clarifies our reasoning .
+%to stop readers from jumping to conclusions before reading the entire section.
 

### 2022-06-13 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-\author{Anna \r{A}kesson$^{1*}$, Alyssa R. Cirtwill$^{2*\dagger}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{1}$} 
-\date{\small$^1$Department of Theoretical Biology, Chemistry, and Physics\\ 
+\author{Alyssa R. Cirtwill$^{1*\dagger}$, Anna \r{A}kesson$^{2*}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{2}$} 
+\date{
+\small$^1$Spatial Foodweb Ecology Group\\
+Research Centre for Ecological Change\\
+Organismal and Evolutionary Biology Research Programm\\
+Faculty of Biological and Environmental Sciences\\
+University of Helsinki
+\medskip
+\small$^2$Department of Theoretical Biology, Chemistry, and Physics\\ 
 Link\"{o}ping University\\
 Link\"{o}ping, Sweden\\
-\medskip
-\small$^2$Department of Agricultural Sciences\\
-University of Helsinki\\
-Helsinki, Finland\\
 \medskip
@@
 Depending on the strength and arrangements of interactions, this means that species may be impacted through several paths, amplifying the original disturbance and causing more secondary extinctions.
-    The net effect of connectance on extinction therefore likely depends on finer-scale network structure, as well as the type and severity of disturbance.
+    The net effect of connectance on extinction therefore likely depends on finer-scale (species-level) network structure, as well as the type and severity of disturbance.
 
-    Species properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
-This can be referred to local properties of the network.
-    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and secondary extinctions due to reduced plant populations are particularly detrimental \citep{Curtsdotter2011}.
+    Species-level properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
+    These can also be referred to as local properties of the network.
+    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and disturbances to plant populations cause particularly many econdary extinctions \citep{Curtsdotter2011}.
 %including secondary extinction due to reduced plant populations.
     However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
-    This since, as well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
-    For example, a generalist consuming many specialist (and therefore high-risk) prey may, for example, be at greater risk of extinction than a specialist which consumes a highly generalist resource that is more unlikely to go extinct.
+    A more detailed perspective is desirable since, as well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
+    For example, a generalist consuming many specialist (and therefore high-risk) prey may be at greater risk of extinction than a specialist which consumes a highly generalist resource that is more unlikely to go extinct (Fig.~\ref{fig:concept}).
 
@@
     These `motif participation' roles capture information about species' direct as well as indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
-We refer to these motifs as meso-scale structures, to emphasise their role as structures between global (network) and local (species) properties.
-A species' motif participation profile shows the second step indirect interactions the species are affected by, and thereby capture important structures that neither global or local properties do.
+    We refer to these motifs as meso-scale structures to emphasise their role as structures between global (network-level) and local (species-level) properties.  
+    A species' motif participation profile shows the second-step indirect interactions the species are affected by, thereby capturing important structures that are missed by global and local properties.
     Therefore, by allowing researchers to distinguish between, for example, consumers of specialists and generalists, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
@@
     Empirical studies have shown that disturbances to plant species can affect some consumers more strongly than others \citep{byrnes2011climate}, that the strength of disturbance affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}, and that the resilience and return time of the ecological community can be strongly influenced by disturbance severity \citep{rydgren2004disturbance}.
-The severity disturbance therefore clearly has an effect on how the ecological community will respond.
+    The severity of disturbance therefore clearly has an effect on how the ecological community will respond.
 
@@
     As Bayesian networks are computationally efficient (see Methods), we are able to simulate a wide range of disturbance intensities and test whether meso-scale network structure also interacts with severity of the disturbance.
-    We focus our analyse on the four network motifs that are over-represented in empirical networks and additionally have been shown to contribute to network stability; the three-species chain, omnivory, apparent competition, and direct competition \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
-
-% % OLD INTRODUCTION BELOW
-% \section*{Introduction} %Old, for comparison
-
-%     %\subsection*{1 Why are bottom-up effects important for understanding stability?}
-%      The plant community is the foundation upon which the myriad species in terrestrial food webs depend for their survival.
-Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-%      The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
-%      For example, the loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
-\citep{scherber2010bottom}.
-
-%      %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
-%     How population decline or extinction of one species leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
-Traditionally, there are two main approaches for studying secondary extinctions.
-First, there are topological models, based only on food web structure \citep{dunne2009cascading}.
-Here, extinctions only affect other species in the network once a consumer has lost all of its prey and therefore must go extinct.
-The second approach uses dynamical models which explicitly simulate population dynamics using a system of differential equations \citep{binzer2011susceptibility}.
-Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network.
-%     This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are more parameter intensive, and simulations are more time-consuming, than topological models.
-
-%     A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks \citep{Eklof2013}. 
-%     In this framework, a consumer's probability of extinction depends on the fraction of resources lost and on a baseline probability of extinction which captures causes unrelated to the network structure (e.g., disease, stochastic extinction of small populations).
-%     Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
-%     As consumers do not affect the extinction probabilities of their resources, Bayesian networks can only capture bottom-up effects of disturbances within networks. 
-%     However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
-%     This emphasizes the importance of bottom-up effects in food webs and makes Bayesian networks the ideal framework to use when studying them.
-
-%     %\subsection*{3 Global scale network properties are known to affect stability}
-%     Using Bayesian networks, researchers can explore how the structures of realistically-large networks shape their responses to disturbance.
-%     The global (whole-network) structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002, Eklof2006}.
-%     For example, more highly-connected networks are more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{Krause2003, Teng2004, Gilarranz2017}.
-
-%     %\subsection*{4 meso-scale nw propoerties also affect stability}
-%     Structural properties on the local (single-species) scale are also known to affect response to disturbance. 
-%     Species with high trophic levels (long paths to basal resources), low in-degree (few prey), and low abundances are generally at higher risks of secondary extinction \citep{binzer2011susceptibility}.
-%     However, in-degree and trophic level are both relatively coarse descriptions of a species' role within its food-web~\citep{Cirtwill2018FoodWebs}. 
-%     For example, a specialist may not necessarily have a high risk of extinction if it consumes a basal resource which is unlikely to be affected by disturbance. 
-%     On the other hand, a consumer with low trophic level may be at high risk of extinction if it consumes prey which are themselves very likely to be affected by disturbance.
-
-%     %\subsection*{5 Motif participation may similarly affect species' particular extinction risk}
-%     One way to incorporate more nuance into descriptions of how species fit into their communities is to define species' roles based on their participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
-This can be referred to as meso-scale properties and describe how a species is interacting with its closest neighbours.
-As such, meso-scale properties is thus one a scale between local properties and global properties.
-Importantly, these `motif participation' roles capture information about species' direct and indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a}.
-%     This means that two species with the same in-degrees and trophic levels can still participate in different sets of motifs if, for example, one interacts with specialists and the other interacts with generalists~\citep{Cirtwill2018FoodWebs}.
-%     This extra detail may be useful when predicting  extinction risk since, at the network level, the frequencies of different motifs are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
-%     We therefore expect that both the motif structure of a network and species' participation in that structure should be related to risks due to disturbance to basal resources.
-
-%           %\subsection*{Here we test...}
-%     Specifically, we are interested in whether a consumer's extinction risk depends on 1) the motifs in which a focal species participates and 2) the motif profile of the full food web. 
-%     In addition, we want to understand how the information provided by motifs is related to classical global and local food web properties in order to better synthesize information provided by different scales of network structure.
-%     Of the 13 three-species motifs that can occur in empirical food webs, there are four motifs that are over-represented in empirical networks and are thought to contribute to stability \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
-These four motifs, the three-species chain, omnivory, apparent competition, and direct competition, are also the only ones which can occur in the acyclic networks required for Bayesian network simulations~\citep{Eklof2013}.
-%     Therefore, we focus on these four particularly important motifs.
+    We focus our analyse on the four network motifs that are over-represented in empirical networks and additionally have been shown to contribute to network stability: the three-species chain, omnivory, apparent competition, and direct competition \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
 
@@
 Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network.
-        This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are extensively more parameter intensive, and simulations are much more time-consuming, than topological models.
+        This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are highly parameter intensive and simulations are much more time-consuming than topological models.
 
-        A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks \citep{Eklof2013, Haussler2020}. 
-        In this framework, a consumer's probability of extinction depends on the fraction of resources lost and on a baseline probability of extinction which captures causes affecting the respective species extinction probability that are unrelated to the network structure (e.g., disease, stochastic extinction of small populations).
+        A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks \citep{Eklof2013,Haussler2020}. 
+        In this framework, a consumer's probability of extinction depends on the fraction of resources lost ($f = k/n$ for a species with $n$ resources, of which $k$ have gone extinct) and on a baseline probability of extinction ($\pi$) which captures factors affecting the focal species' extinction probability that are not related to the network structure (e.g., disease, stochastic extinction of small populations).
         Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
-        This increased speed allows us to simulate a range of disturbance severity rather than single species removals as in many previous studies~\citealp[e.g.,]{Memmott2004,Staniczenko2010,Dunne2004}.
+        This increased speed allows us to simulate a range of disturbance severity rather than single species removals as in many previous studies (e.g.,~\citealp[]{Memmott2004,Staniczenko2010,Dunne2004}).
         Moreover, Bayesian network simulations are less parameter-intensive than dynamical models, meaning that they are less sensitive to assumptions made in the modelling process~\citep{Eklof2013}.
-
-       % In Bayesian networks consumers do not affect the extinction probabilities of their resources, and as such we can here only capture bottom-up effects of disturbances.
-However, Bayesian networks do capture the majority of secondary extinctions captured by dynamical models \citep{Eklof2013}.
-Moreover, bottom-up effects are often strong in empirical systems.
-For example, bottom-up effects regulate zoo-plankton communities in freshwater lakes, regulating zoo-plankton communities \citep{li2020bottom} and control abundances and biomass of vertebrate species in several terrestrial systems \citep{Dobson2009food, Mduma1999food, Georgiadis2007}.
-
-	%\subsection*{Calculating persistence}	
-    % basic idea behind BN modeling
-        Bayesian networks can be used to represent species and their feeding interactions as the nodes and links of a network, respectively (see Appendix \emph{S2} for full details).
-        In our setting, each species (node) is a Bernoulli random variable, being either extant or extinct. 
-        The probability that a particular species is extinct is a function of the states of its resources.
-
-		% probability of extinction equation
-		%Consider a species with $n$ resource species. 
-		Before taking food-web structure into account, all species already have a baseline probability of extinction, $\pi$. 
-		$\pi$ represents the risk of going extinct due to diseases, stochastic extinction of small populations, or other factors not related to the food web.
-		A species probability of extinction also depends on the state of its resources. 
-        If the focal species is a consumer with $n$ resources, of which $k$ have gone extinct, its probability of extinction depends on $\pi$ and on the fraction $f = k/n$.
 
@@
 This will in turn affect the extinction probability $P(\lnot C|f)$.
-		For basal resources, we assume that abiotic resources are always available and $f=0$. Following Equation \ref{betafunc}, the extinction will only depend on $\pi$.
+		For basal resources, we assume that abiotic resources are always available and $f=0$. 
+		Following Equation \ref{betafunc}, the extinction probability of a basal resource will only depend on $\pi$.
 		Note that we here assume that only the fraction, and not the identity, of the resources lost is important. 
@@
         Species are considered extinct if a random number drawn from a uniform distribution $[0-1]$ is less than $P(\lnot i|f)$.
-
         These simulated extinctions are used when calculating the fraction of resources lost $f$ for each consumer.
@@
 		Modelling species extinctions in food webs using Bayesian networks has two main limitations. 
-First, in Bayesian networks consumers do not affect the extinction probabilities of their resources, and as such we can here only capture bottom-up effects of disturbances.
+		First, in Bayesian networks consumers do not affect the extinction probabilities of their resources and as such we can here only capture bottom-up effects of disturbances. 
 		However, Bayesian networks has been shown to capture the majority of secondary extinctions produced fully-fledged dynamical models \citep{Eklof2013}.
@@
 		Thus, making the networks acyclic in order to use the Bayesian network approach will not strongly affect extinction scenarios.
+		Moreover, the cyclic motifs which are excluded from an acyclic network are rare in empirical networks~\citep{Stouffer2007} and therefore unlikely to strongly affect extinction risk.
 
     \subsection*{Disturbance scenarios}
+
         In our baseline scenario, all species had an extinction probability of $\pi = 0.1$. 
-		To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase the affected basal species' probability of extinction from $0.10$ to $0.5$, in steps of $0.08$.
+		To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase basal species' probability of extinction from $0.10$ to $0.5$, in steps of $0.08$. 
 		The highest disturbance level, $\pi = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
@@
 		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
-
 		To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and species topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).

### 2022-06-14 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
     % Degree and TL are both important for omnivory
-    The relationship between participation in the omnivory motif and persistence connects motif participation with both in-degree and trophic level.
+    The relationship between participation in the omnivory motif and persistence connects motif participation with both in-degree and trophic level (STL).
     Species participation in the omnivory motif was associated with greater persistence, especially at low levels of disturbance.
@@
     % High C webs and omnivory
-    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks and the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}) as in~\cite{McLeod2021}.
+    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks and the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}) as in~\citet{McLeod2021}.
     These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
@@
     In summary, our results show that participating in different three-species motifs is related to a species' risk of extinction following a bottom-up disturbance, and that these relationships synthesize information provided by other, simpler measures of species roles.
-    Most interestingly, our results also underscore the importance of considering the level of disturbance when identifying which species are most at risk of extinction.
+    Critically, our results also underscore the importance of considering the level of disturbance when identifying which species are most at risk of extinction.
     There were strong interactions between the level of disturbance and the relationships between persistence and multiple measures of species roles (motifs and in-degree).
@@
         \caption{Conceptual figure of a toy food web.
-\textbf{a)} This food web contains examples of the four motifs (present in an acyclic directed networks): apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow).
+\textbf{a)} This food web contains examples of the four motifs: apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow).
 These motifs capture different ways in which disturbances to basal resources can propagate through the network.
@@
         Note that the nodes participate in more motifs than the ones highlighted.
-\textbf{b)} In the same food web, species with similar degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than tree times more frequently.
-Inserted table shows the number of times A and B, respectively, participate in the four different motifs.
-This example highlight that species A may then have a higher risk of going secondary extinct due to its dependence on specialist prey species, compared to species B who depends om generalist prey species.
-This relationships are captured by the species motif profiles, but not their trophic level or degree.}
+\textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than tree times more frequently (see inserted table for counts of participation in all four motifs).
+        Despite having equal degree and trophic level, species A may be at greater risk of going extinct than species B because of its dependence on specialist prey species.
+Two-thirds of the prey of species B are generalists and therefore more likely to survive a disturbance to basal resources.
+By including information about indirect interactions, motif profiles reflect this difference between species A and B while trophic level and degree do not.}
     \label{fig:concept}

### 2022-06-14 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 % \usepackage[numbers,sort]{natbib}
-\usepackage[round]{natbib} 
+\usepackage[numbers]{natbib} 
 \usepackage{array}
@@
 
-		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}.
+		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.
 		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
@@
         \centering
-        \includegraphics[width=0.85\textwidth]{figures/persistence_vs_motifs_labs.eps}
+        \includegraphics[width=0.85\textwidth]{figures/persistence_vs_motifs.eps}
         \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).
@@
     \label{fig:prop_lmer_all}
-    \end{figure}
-
-        \begin{figure}[hb!]
-        \centering
-        \includegraphics[width=0.85\textwidth]{figures/persistence_vs_motifs.eps}
-        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).
-The effect of participating in each motif is based on the fixed effects in Equation~\ref{propreq}.
-The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance).
-95\% confidence intervals for each line are shown in grey.
-Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
-    \label{fig:prop_lmer_2}
     \end{figure}
@@
 
-\bibliographystyle{ecollett} 
+\bibliographystyle{unsrtnat} 
 \bibliography{anna_bib_new} % Abbreviate journal titles.

### 2022-06-14 manuscript/response_PRSB_round2.tex

@@
 
-\subsection*{Associate Editor} [[review Alyssa's new text]]
+\subsection*{Associate Editor}
     \them
@@
 We hope that this makes our goals clearer and results easier to interpret.
-We have not adjusted our analyses as we note that both the Associate Editor and Reviewer 2 found the current methods suitable.
 
@@
     \them
-    Last, the authors generate what they argue are "a range of plausible network architectures".
+    Last, the authors generate what they argue are ``a range of plausible network architectures''.
 But there is never any comparison of these architecture to empirical networks.
@@
 It will expand the audience for this manuscript greatly to include some empirical data to help empiricists understand for example in well-studied systems, what is the proportion of these different motifs.
-This comparison does not need to be exhaustive, but it will serve two main roles in the manuscript: (1) provide support for the assertion that the authors have generated "a range of plausible network architectures" and (2) allow empiricists to understand how these simulations relate to real ecosystems.
+This comparison does not need to be exhaustive, but it will serve two main roles in the manuscript: (1) provide support for the assertion that the authors have generated ``a range of plausible network architectures'' and (2) allow empiricists to understand how these simulations relate to real ecosystems.
 
@@
     Specifically, we compare the motif profiles and consumers' motif participation of our simulated webs to the  unipartite webs in the same size range (excluding those where predators and prey were differently resolved) available from the globalweb database.
-    Not all of the empirical webs have motif profiles within the simulated range; this could be because of flaws in our simulation or because of flaws in the empirical webs (e.g., poorly-resolved basal resources).
+    Not all of the empirical webs have motif profiles within the simulated range; this could be because of flaws in our simulation or because of flaws in the empirical webs (e.g., poorly-resolved basal resources, a common problem in even high-quality empirical webs).
     While we cannot say conclusively which is the case, we are encouraged by the fact that the consumer motif participations in our simulated webs span the motif participations in the empirical web.
@@
     Thank you for the opportunity to review this paper.
-I found it very interesting and appreciated the comparison between motifs and more “classic” network metrics.
+I found it very interesting and appreciated the comparison between motifs and more ``classic'' network metrics.
 I have a couple major concerns which I do not anticipate being that difficult to address, and some smaller comments.
@@
         \us
-        The introduction has been re-written, and is hopefully clearer now in terms of both terminology and our questions of interest.
+        The introduction has been thoroughly re-written, and is hopefully clearer now in terms of both terminology and our questions of interest.
 Additionally, we have now included a glossary in \emph{Appendix S11}
@@
         \us One large model will not serve the purpose of our manuscript.
-We agree that connectance and species richness have a large impact on persistence, but re-iterating these well-known effects is not our main focus.
+We agree that connectance and species richness have a large impact on persistence, but re-proving these well-known effects is not our main focus.
 Instead, our main goal is to show whether, as is frequently proposed, information about motif participation or profiles can be used to comment on extinction risk.
@@
 
+        In the same vein, since we cannot imagine a scenario in which researchers have either degree and trophic level or motif participation information but not both (as calculating degrees of all species requires the same diet information used to calculate motifs, even if we assume trophic levels are derived from stable isotopes), we do not believe that it is productive to pit motifs against trophic level.
+Our aim is to add motif participation to the toolbox of researchers wishing to predict extinction risk, not attempt to supplant trophic level, which we agree provides a large amount of useful information and should be taken into account.
+
         \them
@@
 I think the study would be greatly enhanced by a smaller comparison of a more traditional dynamic food web model approach and this approach – even just placed in the appendix.
-I note that the authors rely quite heavily on a paper on BioRxivs which has done something similar, so anticipate this wouldn’t be that difficult, but would be very convincing.
+I note that the authors rely quite heavily on a paper on BioRxivs which has done something similar, so anticipate this wouldn't be that difficult, but would be very convincing.
 
@@
 We use this information in the current manuscript and make our simulations with the functional form that was shown to most accurately predict the same secondary extinctions as the dynamical model.
-We have now calrified this in the manuscript (line XX).
-However, we would also like to point out that there are no guarantees that advanced (as in parameter intensive) dynamical models describe the ecological reality better, as comparisons with extensive empirical extinction scenarios are still lacking.
+We have now clarified this in the manuscript (line 126-127).
+However, we would also like to point out that there are no guarantees that advanced (as in parameter intensive) dynamical models describe the ecological reality better, as comparisons with extensive, highly-resolved empirical extinction scenarios are still lacking.
 
         In addition, we believe that the Reviewer has over-estimated our reliance on the BioRxiv pre-print currently under review.
-Our previous draft had only three references to this pre-print: when describing the generation of simulated networks, when noting that degree and motif participation are not independent, and when briefly comparing the dynamical and Bayesian results in the discussion.
+Our previous draft had only three references to this pre-print: when describing the generation of simulated networks, when noting that degree and motif participation are not independent (which we also demonstrate \emph{de novo} in the present manuscript), and when briefly comparing the dynamical and Bayesian results in the discussion.
         We have now removed two of these references.
@@
         As the pre-print is freely available to any who wish to critically examine it, we believe that this is a good compromise between maintaining conciseness of the present manuscript and comparing different simulation approaches.
+        If, all going well, the preprint is accepted for publication by the time the present manuscript is accepted, we will of course update the reference to provide the peer-reviewed version of the text.
 
@@
         \them
-        Line 65 – 70: I agree that in-degree and trophic level may be coarse description of a species’ role within its food-web.
-However, the examples provided do not convince me that species’ roles based on motifs are a more nuanced way of describing a species’ role within its food web since the two examples provided refer to the influence disturbance may have on a basal resource – an influence that I’m not convinced a species’ role based on motifs will impact.
+        Line 65 – 70: I agree that in-degree and trophic level may be coarse description of a species' role within its food-web.
+However, the examples provided do not convince me that species' roles based on motifs are a more nuanced way of describing a species' role within its food web since the two examples provided refer to the influence disturbance may have on a basal resource – an influence that I'm not convinced a species' role based on motifs will impact.
 
@@
         \them
-        I do think that species’ roles based on motifs are a more nuanced way of describing species’ roles, and I am interested in the central question of this paper – however, this rationale would not convince me if I was unfamiliar with motifs/species roles/finer scale network descriptors.
+        I do think that species' roles based on motifs are a more nuanced way of describing species' roles, and I am interested in the central question of this paper – however, this rationale would not convince me if I was unfamiliar with motifs/species roles/finer scale network descriptors.
 As this is a central tenet of the paper I think the argument needs to be constructed better.
@@
         \them
-        Line 82-84: This sentence doesn’t quite make sense to me.
+        Line 82-84: This sentence doesn't quite make sense to me.
 
@@
         \them
-        Line 86: is this “focal” species the consumer mentioned in line 85?
+        Line 86: is this ``focal'' species the consumer mentioned in line 85?
 
@@
         \them
-        Line 86: the impact of the motif profile of the full food web on the consumer’s extinction risk has come as a bit of a surprise to me.
+        Line 86: the impact of the motif profile of the full food web on the consumer's extinction risk has come as a bit of a surprise to me.
 I think this could have been introduced better in the introduction – not extensively because I think it is probably a secondary question, but at this point I have no hypotheses for why I would even expect the motif profile of the full food web to out perform the motifs in which a focal species participates.
@@
         \them
-        Line 92 – 95: this seems to be a flaw to the Bayesian network simulations – we know that mutual predation happens in food webs – especially if the food web isn’t explicitly size structured, and when nested within the broader food web context these feedback loops may have stabilizing, or destabilizing effects.
+        Line 92 – 95: this seems to be a flaw to the Bayesian network simulations – we know that mutual predation happens in food webs – especially if the food web isn't explicitly size structured, and when nested within the broader food web context these feedback loops may have stabilizing, or destabilizing effects.
         \smallskip
@@
         \us In revising the introduction, we now introduce the Bayesian network approach and its limitations in the methods section.
-We want to stress that the Reviewer is clearly correct that these are important limitations and we have further clarified that in the text.
-At the same time we have clarified the advantages that the Bayesian network approach provides.
+We want to stress that the Reviewer is absolutely correct that these are important limitations and we have further clarified that in the text.
+At the same time we have clarified the advantages that the Bayesian network approach provides in order to better make the case that the Bayesian approach is useful in this case.
         First, we clearly state that the lack of top-down effects in the Bayesian networks is a limitation.
@@
         We are more interested in comparing extinction risk for species within the same food web (where connectance is constant) than between food webs (where connectance, species richness, and --in empirical systems-- local conditions will have large effects).
-        We have lightly revised the methods section to make this focus clearer and to state that, while our motif role normalization does not control for degree or connectance, we do explicitly consider how degree and connectance (as well as species richness and trophic level) may be related.
+        We have lightly revised the methods section (quoted below) to make this focus clearer and to state that, while our motif role normalization does not control for degree or connectance, we do explicitly consider how degree and connectance (as well as species richness and trophic level) may be related.
 
@@
 Be explicit.
-Also, what are the “other network properties”?
+Also, what are the ``other network properties''?
 
         \us
-        Added clarifications to to point 3 and 5 where we had "network mean persistence" and "other network properties", please check if ok.
+        We have revised the methods text to clarify that network mean persistence is the mean probability of persistence across all consumers in the network.
+We also now specify that ``other network properties'' refer to species richness, connectance, in-degree, and trophic level.
 
@@
         \them
-        Line 193: Does it make sense to include Sn and Cn as random intercepts if it is known that both of these values impact 1) the number of motifs (which I don’t think has been rigorously controlled for – see above) and 2) network persistence?
-(same with eq’n S3)
+        Line 193: Does it make sense to include Sn and Cn as random intercepts if it is known that both of these values impact 1) the number of motifs (which I don't think has been rigorously controlled for – see above) and 2) network persistence?
+(same with eq'n S3)
 
@@
         We explicitly control for differences in the number of motifs through the frequency normalisation discussed above.
-We include the random intercepts to account for potential effects of Sn and Cn on network persistence that are \emph{not} due to effects on the total number of motifs.
-We agree with the Reviewer that the effects of Sn and Cn are well-known and important to account for, but they are not the main focus of our manuscript.
-We are therefore comfortable with keeping them as random effects in these equations.
+This is a more direct approach than controlling for effects of S and C on the total number of motifs, which are likely but have not been thoroughly explored to date.
+We include the random intercepts to account for potential effects of S and C on network persistence that are \emph{not} due to effects on the total number of motifs.
+We agree with the Reviewer that the effects of S and C are well-known and important to account for, but they are not the main focus of our manuscript.
+We therefore choose to keep them as random effects in these equations.
 
@@
         \us While we do not use a generative model that fills some motifs/interactions before others, we do not see any particular reason why apparent competition motifs should be filled first.
+More likely, the high frequencies of apparent competition reflects the tendency for resources to outnumber consumers, creating the classic ``pyramid shape'' food web.
 We can confirm, however, that the connectances we explore are realistic.
@@
 Our terminology follows the nomenclature from the introduction of motifs to food webs~\citep{Stouffer2007}.
-        Note that ``direct competition'' refers simply to two species which consume the same prey and is agnostic as to whether this is exploitative or interference competition. 
+        Note that ``direct competition'' refers simply to two species which consume the same prey and is agnostic as to whether or not competition involves interference or is purely exploitative. 
         As exploitative and interference competition are not the same ecologically, we have maintained the conventional motif terminology so as not to imply an overly specific interpretation.
@@
 However, as detailed later in the paragraph, this ignores the strong correlation between apparent competition and trophic level which would suggest that persistence should show different relationships to apparent competition.
-This means that either trophic level is unrelated to persistence (which we know is not true) or some other factor is over-riding the ``effect" (there is no clear causality from trophic level to motifs or vice versa) of trophic level on apparent competition.
-
+This means that either trophic level is unrelated to persistence (which we know is not true) or some other factor is over-riding the ``effect'' (there is no clear causality from trophic level to motifs or vice versa) of trophic level on apparent competition.
         We have revised this paragraph with particular focus on these relationships and hope that clarifies our reasoning .
-%to stop readers from jumping to conclusions before reading the entire section.
 
@@
         \them
-        Line 16: I suggest changing the word “vulnerability” to “species persistence” to keep consistency along the manuscript.
+        Line 16: I suggest changing the word ``vulnerability'' to ``species persistence'' to keep consistency along the manuscript.
 
@@
         Line 23-26: this sentence could be clearer.
-Perhaps authors could rephrase to: “The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole (Proulx et al., 2010; Scherber et al., 2010, Rosenblatt \& Schmitz, 2016).
-For example, the loss of diversity…”
+Perhaps authors could rephrase to: ``The structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole (Proulx et al., 2010; Scherber et al., 2010, Rosenblatt \& Schmitz, 2016).
+For example, the loss of diversity…''
 
@@
         \them
-        Line 112: Please add “, respectively” after “network”.
+        Line 112: Please add ``, respectively'' after ``network''.
 
@@
         \them
-        Line 187: Please add “Species” before “persistence” in the title to improve clarity.
+        Line 187: Please add ``Species'' before ``persistence'' in the title to improve clarity.
 
@@
         \them
-        Line 200: Please add “Motif” before “participation” in the sentence.
+        Line 200: Please add ``Motif'' before ``participation'' in the sentence.
 
@@
         \them
-        Line 207: Add “Network” before “persistence” in the title.
+        Line 207: Add ``Network'' before ``persistence'' in the title.
 
@@
         \them
-        Line 330-331: Unclear the expression “, and interactions among,”. Please rephrase.
+        Line 330-331: Unclear the expression ``, and interactions among,''. Please rephrase.
 
@@
         \them
-        Line 357-358: Unclear the expression “than directly”. Please explain.
+        Line 357-358: Unclear the expression ``than directly''. Please explain.
 
@@
         \them
-        Line 376: Please remove the italic formatting of the word “and”.
+        Line 376: Please remove the italic formatting of the word ``and''.
 
@@
         \them
-        Figure 5: Please rephrase the first sentence to: “The proportions of the four motifs in a network's motif profile were related to network size, connectance, and their interaction.”
+        Figure 5: Please rephrase the first sentence to: ``The proportions of the four motifs in a network's motif profile were related to network size, connectance, and their interaction.''
 
@@
 
-        Page 3: Please remove the expression “(?)” from the sentence in the third paragraph.
+        Page 3: Please remove the expression ``(?)'' from the sentence in the third paragraph.
 
-        Page 7: Please remove the expression “(?)” from the last sentence.
+        Page 7: Please remove the expression ``(?)'' from the last sentence.
 
-        Page 10: Please remove the expression “(?)” from the first paragraph of the section S3.3.
+        Page 10: Please remove the expression ``(?)'' from the first paragraph of the section S3.3.
 
         \us
-        Done.
+        Done. These were incorrect references and have now been corrected.
 
+\clearpage
+\bibliographystyle{vancouver}
+\bibliography{anna_bib_new}
 \end{document}

### 2022-06-14 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-        The response of the consumer to changes in $f$ can take any form (Appendix \emph{S2}). Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$).
+        The response of the consumer to changes in $f$ can take any form (\emph{Appendix S1}). Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$).
 		This type of response most accurately captures the secondary extinctions produced by dynamical models (\citealp[]{Eklof2013}).
@@
 		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
-		For full details and a comparison to empirical networks, see \emph{Appendix S1}.
-		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
-		To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and species topologically sorted from lowest to highest trophic level (Appendix \emph{S1}).
+		For full details and a comparison to empirical networks, see \emph{Appendix S2}.
+		We removed any biologically unlikely networks ( \emph{Appendix S2}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
+		To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and species topologically sorted from lowest to highest trophic level (\emph{Appendix S2}).
 		 These processing steps allow the strictly bottom-up calculation of probability of persistence.
@@
 
-\bibliographystyle{unsrtnat} 
+\bibliographystyle{vancouver} 
 \bibliography{anna_bib_new} % Abbreviate journal titles.

### 2022-06-14 manuscript/supplement_motifs.tex
reordered SI
converted all figs to eps

@@
 
+        \item Calculating persistence
+
+            \begin{itemize}
+                \item Details on functional response of consumers
+                \item Contains Eqn. S1
+            \end{itemize}
+
         \item Network creation \\
@@
             \item Comparison of empirical and simulated networks
-            \end{itemize}
-
-        \item Calculating persistence
-
-            \begin{itemize}
-                \item Details on functional response of consumers
-                \item Contains Eqn. S1
             \end{itemize}
@@
 
-\section{Network creation}
-
-    \subsection*{Generation of simulated networks}
-
-        We simulated a suite of food webs based on the probabilistic niche model, which assigns predator-prey links based on the body-mass ratios between individuals of different species~\citep{Williams2000,Delmas2017}.
-The meso-scale structure of niche-model networks closely mimics that of empirical food webs~\citep{Stouffer2007}.
-To ensure that we captured a variety of realistic community sizes and structures, we generated networks ranging between 50 and 100 species (in steps of 10) with connectance values between 0.02 and 0.2 (in steps of 0.02).
-The range of network sizes was chosen to reflect moderately well-sampled empirical webs while working within our computational limits, while the range of connectance values was chosen to cover that observed in most empirical food webs~\citep{Dunne2002e}.
-We generated a total of 100 networks with each combination of parameters, for a total of 6000 networks.
-All networks were generated using the function "nichemodel" within the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
-If a simulated network contained any disconnected species (species without predators or prey) or disconnected components (a group of species connected among themselves but not to the rest of the network), the network was rejected and a new network simulated.
-Finally, networks where the path lengths between each species and a basal resource could not be resolved (i.e., trophic levels were undefined) were rejected and new networks simulated.
-
-        After generating the network structure, we simulated community dynamics using the function ``simulate'' from the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
-This function uses Lotka-Volterra predator-prey models including density dependence and type 2 functional responses for all species (please see~\citet{Delmas2017} for full details).
-        All non-basal species were designated as vertebrates to ensure a good match between metabolic and predator-prey body-mass ratio values.
-Metabolic rates in the Lotka-Volterra model are based on each species' body mass (i.e.
-mass of a single individual).
-We assigned relative body masses based on each species' trophic level, which was, in turn, calculated based on the food-web structure provided by the niche model.
-After basal species were assigned a body mass of 1, we used a predator-prey body-mass ratio of 3.065 to calculate the relative body masses of higher trophic levels.
-We selected this ratio based on the estimate for vertebrates (averaged across ecosystem and metabolic types) in~\citet{Brose2006}.
-We excluded reported body-mass ratios for invertebrates as these could include parasites and parasitoids, which are generally smaller than their prey, and because interactions among vertebrates are better represented in the food-web literature than interactions involving invertebrates.
-
-        The persistence of each species in our simulated networks also depends on its population biomass. 
-        We randomly assigned initial population biomasses (i.e.
-cumulative biomass across all individuals of a species) for each species from a uniform distribution [0,1].
-Note that population biomasses and individual body masses are not calculated on the same scale.
-We then simulated community dynamics for 1000 time steps to obtain an `equilibrium' community.
-To ensure that species did not `recover' from unrealistically low biomasses during the simulation, we considered a species extinct if it dropped below an arbitrary threshold biomass of 1$\times10^{-5}$.
-When simulating initial (i.e.
-pre-disturbance or equilibrium) dynamics, we rejected any network where one or more species dropped below this biomass threshold.
-Consumers were assumed to have no preferences such that the consumption rate $w_{ij}$ of predator $i$ eating prey $j$ is equal to $\frac{1}{n}$, where $n$ is the number of prey for predator $i$.
-If the network did not retain all species for 1000 time steps, a new set of initial population biomasses was applied and the simulation repeated.
-        If a set of biomasses where all species persist still had not been reached after 100 sets of randomly-assigned initial biomasses, we discarded the network and simulated another to replace it.
-
-    \subsection*{Filtering of biologically unrealistic networks}
-
-        After simulating our initial niche-model networks, we filtered networks for biological plausibility.
-        Specifically, we removed any network containing disconnected components
-        (species or groups of species not connected to the rest of the network) 
-        or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
-        Any removed network was replaced with a fresh simulation and re-checked until 100 realistic networks for each size-connectance combination had been obtained.
-
-    \subsection*{Conversion of networks to acyclic form}
-
-        A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}.
-        All persistence calculations therefore begin by determining the status of the basal resources (who do not depend on other species).
-        Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on basal resources), and so on up the network.
-
-        While niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
-        We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009functional}.
-        This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009functional}.
-        These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
-        Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citet{Tarjan1972} and \citet{Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
-        After these steps, calculations of persistence probabilities can begin.
-
-\clearpage
-
-    \subsection*{Comparison of empirical and simulated networks}
-
-        Although previous work has shown that the niche model replicates the motif structure of empirical networks~\citep{Stouffer2005a,Stouffer2006}, we provide a comparison of the motif profiles and consumer motif participation for our simulated networks and the highest-quality empirical networks we could find within the same size range.
-        For this comparison, we obtained all networks containing 50-100 species from \citet{globalwebdb}.
-        These webs ranged in connectance from 0.024-0.089
-        We then removed networks where sets of predators and prey were completely different (i.e., bipartite networks) or links were not clear (i.e., +/- effects rather than predation specifically).
-        This yielded five vertebrate-dominated food webs describing the California coast in different years, four invertebrate-dominated food webs describing mainland US streams, 23 invertebrate-dominated food webs describing New Zealand streams, and five other invertebrate-dominated food webs.
-        This final set of webs had 50-77 species and connectances of 0.2-0.21 - the upper range of connectance in our simulations.
-        This may be partly due to relatively poor resolution of basal resources in these empirical webs: aggregation of many taxa into nodes such as `phytoplankton' will tend to decrease the number of species but leave the number of links unchanged, raising connectance.
-        We note that most of these webs were compiled by just three research groups.
-        Moreover, many of the webs are repeated samples of the same site  or closely-grouped sites.
-        This means that these empirical networks cannot be considered fully independent samples.
-
-        \begin{figure}
-            \includegraphics[width=\textwidth]{figures/motif_profiles_participation_vs_empirical.eps}
-            \caption{\textbf{A)} Proportions of acyclic motifs in the motif profiles of empirical and simulated networks.
-Mean proportions ($\pm$SD) in the simulated networks are indicated in black, with the range of proportions in the simulated networks indicated by black boxes.
-\textbf{B)} Proportions of acyclic motifs in the motif participation vectors of consumers in simulated and empirical networks.}
-            \label{empirical_compare}
-        \end{figure}
-
-    \subsubsection*{Comparing motif profiles}
-
-      The proportions of omnivory and apparent competition in the motif profiles of simulated networks matched those of the empirical webs fairly well, but many of the empirical webs had higher proportions of direct competition and lower proportions of the three-species chain than any simulated web.
-      The empirical webs with high proportions of direct competition tend to be invertebrate-dominated webs representing small streams.
-      High proportions of direct competition in these system suggest large numbers of fairly generalist herbivore/detritivores, with relatively few secondary consumers present to create chains.
-
-      A few webs had lower proportions of apparent competition than any simulated web but similar proportions of the other motifs.
-      Most of these webs represented the California coast in different years (2003-2007).
-      Basal resources in these webs were highly aggregated into only two size classes of phytoplankton and three types of detritus.
-      This aggregation at the basal level leads to artificial ``specialisation'' in consumers with low trophic levels who may consume, for example, many species of phytoplankton that are grouped into a single size class.
-      As the basal resources in our simulated networks are not aggregated, our webs feature higher proportions of apparent competition as seen in the other empirical webs.
-
-    \subsubsection*{Comparing motif participation}
-
-      Despite the differences in network motif profiles between some of the empirical food webs and the simulated webs, we note that the motif participation of consumers in the simulated webs entirely spans the range of motif participation for consumers in the empirical webs.
-      Consumers in the NZ stream food webs had higher proportions of direct competition and lower proportions of apparent competition than the mean of the simulated webs, consistent with the network motif profiles.
-      Nonetheless, the greater differences in network profiles than consumer motif participation suggests that the largest differences between the simulated and empirical networks are in the roles of basal resources.
-      In empirical webs, basal resources are often highly aggregated. 
-      Improved resolution at the basal level should improve the correspondence between simulated and empirical webs and allow better modelling of empirical communities.
-
+\linenumbers
 \section{Calculating persistence}
@@
         To approximate persistence probabilities through numerical simulation are shown to be highly efficient and produce the same result as exact methods of solving Bayesian networks \citep{Haussler2020}.
+
+\clearpage
+
+\section{Network creation}
+
+    \subsection*{Generation of simulated networks}
+
+        We simulated a suite of food webs based on the probabilistic niche model, which assigns predator-prey links based on the body-mass ratios between individuals of different species~\citep{Williams2000,Delmas2017}.
+The meso-scale structure of niche-model networks closely mimics that of empirical food webs~\citep{Stouffer2007}.
+To ensure that we captured a variety of realistic community sizes and structures, we generated networks ranging between 50 and 100 species (in steps of 10) with connectance values between 0.02 and 0.2 (in steps of 0.02).
+The range of network sizes was chosen to reflect moderately well-sampled empirical webs while working within our computational limits, while the range of connectance values was chosen to cover that observed in most empirical food webs~\citep{Dunne2002e}.
+We generated a total of 100 networks with each combination of parameters, for a total of 6000 networks.
+All networks were generated using the function "nichemodel" within the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
+If a simulated network contained any disconnected species (species without predators or prey) or disconnected components (a group of species connected among themselves but not to the rest of the network), the network was rejected and a new network simulated.
+Finally, networks where the path lengths between each species and a basal resource could not be resolved (i.e., trophic levels were undefined) were rejected and new networks simulated.
+
+        After generating the network structure, we simulated community dynamics using the function ``simulate'' from the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
+This function uses Lotka-Volterra predator-prey models including density dependence and type 2 functional responses for all species (please see~\citet{Delmas2017} for full details).
+        All non-basal species were designated as vertebrates to ensure a good match between metabolic and predator-prey body-mass ratio values.
+Metabolic rates in the Lotka-Volterra model are based on each species' body mass (i.e.
+mass of a single individual).
+We assigned relative body masses based on each species' trophic level, which was, in turn, calculated based on the food-web structure provided by the niche model.
+After basal species were assigned a body mass of 1, we used a predator-prey body-mass ratio of 3.065 to calculate the relative body masses of higher trophic levels.
+We selected this ratio based on the estimate for vertebrates (averaged across ecosystem and metabolic types) in~\citet{Brose2006}.
+We excluded reported body-mass ratios for invertebrates as these could include parasites and parasitoids, which are generally smaller than their prey, and because interactions among vertebrates are better represented in the food-web literature than interactions involving invertebrates.
+
+        The persistence of each species in our simulated networks also depends on its population biomass. 
+        We randomly assigned initial population biomasses (i.e.
+cumulative biomass across all individuals of a species) for each species from a uniform distribution [0,1].
+Note that population biomasses and individual body masses are not calculated on the same scale.
+We then simulated community dynamics for 1000 time steps to obtain an `equilibrium' community.
+To ensure that species did not `recover' from unrealistically low biomasses during the simulation, we considered a species extinct if it dropped below an arbitrary threshold biomass of 1$\times10^{-5}$.
+When simulating initial (i.e.
+pre-disturbance or equilibrium) dynamics, we rejected any network where one or more species dropped below this biomass threshold.
+Consumers were assumed to have no preferences such that the consumption rate $w_{ij}$ of predator $i$ eating prey $j$ is equal to $\frac{1}{n}$, where $n$ is the number of prey for predator $i$.
+If the network did not retain all species for 1000 time steps, a new set of initial population biomasses was applied and the simulation repeated.
+        If a set of biomasses where all species persist still had not been reached after 100 sets of randomly-assigned initial biomasses, we discarded the network and simulated another to replace it.
+
+    \subsection*{Filtering of biologically unrealistic networks}
+
+        After simulating our initial niche-model networks, we filtered networks for biological plausibility.
+        Specifically, we removed any network containing disconnected components
+        (species or groups of species not connected to the rest of the network) 
+        or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
+        Any removed network was replaced with a fresh simulation and re-checked until 100 realistic networks for each size-connectance combination had been obtained.
+
+    \subsection*{Conversion of networks to acyclic form}
+
+        A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}.
+        All persistence calculations therefore begin by determining the status of the basal resources (who do not depend on other species).
+        Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on basal resources), and so on up the network.
+
+        While niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
+        We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009functional}.
+        This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009functional}.
+        These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
+        Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citet{Tarjan1972} and \citet{Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
+        After these steps, calculations of persistence probabilities can begin.
+
+    \subsection*{Comparison of empirical and simulated networks}
+
+        Although previous work has shown that the niche model replicates the motif structure of empirical networks~\citep{Stouffer2005a,Stouffer2006}, we provide a comparison of the motif profiles and consumer motif participation for our simulated networks and the highest-quality empirical networks we could find within the same size range.
+        For this comparison, we obtained all networks containing 50-100 species from \citet{globalwebdb}.
+        These webs ranged in connectance from 0.024-0.089
+        We then removed networks where sets of predators and prey were completely different (i.e., bipartite networks) or links were not clear (i.e., +/- effects rather than predation specifically).
+        This yielded five vertebrate-dominated food webs describing the California coast in different years, four invertebrate-dominated food webs describing mainland US streams, 23 invertebrate-dominated food webs describing New Zealand streams, and five other invertebrate-dominated food webs.
+        This final set of webs had 50-77 species and connectances of 0.2-0.21 - the upper range of connectance in our simulations.
+        This may be partly due to relatively poor resolution of basal resources in these empirical webs: aggregation of many taxa into nodes such as `phytoplankton' will tend to decrease the number of species but leave the number of links unchanged, raising connectance.
+        We note that most of these webs were compiled by just three research groups.
+        Moreover, many of the webs are repeated samples of the same site  or closely-grouped sites.
+        This means that these empirical networks cannot be considered fully independent samples.
+
+        \begin{figure}
+            \includegraphics[width=\textwidth]{figures/motif_profiles_participation_vs_empirical.eps}
+            \caption{\textbf{A)} Proportions of acyclic motifs in the motif profiles of empirical and simulated networks.
+Mean proportions ($\pm$SD) in the simulated networks are indicated in black, with the range of proportions in the simulated networks indicated by black boxes.
+\textbf{B)} Proportions of acyclic motifs in the motif participation vectors of consumers in simulated and empirical networks.}
+            \label{empirical_compare}
+        \end{figure}
+
+    \subsubsection*{Comparing motif profiles}
+
+      The proportions of omnivory and apparent competition in the motif profiles of simulated networks matched those of the empirical webs fairly well, but many of the empirical webs had higher proportions of direct competition and lower proportions of the three-species chain than any simulated web.
+      The empirical webs with high proportions of direct competition tend to be invertebrate-dominated webs representing small streams.
+      High proportions of direct competition in these system suggest large numbers of fairly generalist herbivore/detritivores, with relatively few secondary consumers present to create chains.
+
+      A few webs had lower proportions of apparent competition than any simulated web but similar proportions of the other motifs.
+      Most of these webs represented the California coast in different years (2003-2007).
+      Basal resources in these webs were highly aggregated into only two size classes of phytoplankton and three types of detritus.
+      This aggregation at the basal level leads to artificial ``specialisation'' in consumers with low trophic levels who may consume, for example, many species of phytoplankton that are grouped into a single size class.
+      As the basal resources in our simulated networks are not aggregated, our webs feature higher proportions of apparent competition as seen in the other empirical webs.
+
+    \subsubsection*{Comparing motif participation}
+
+      Despite the differences in network motif profiles between some of the empirical food webs and the simulated webs, we note that the motif participation of consumers in the simulated webs entirely spans the range of motif participation for consumers in the empirical webs.
+      Consumers in the NZ stream food webs had higher proportions of direct competition and lower proportions of apparent competition than the mean of the simulated webs, consistent with the network motif profiles.
+      Nonetheless, the greater differences in network profiles than consumer motif participation suggests that the largest differences between the simulated and empirical networks are in the roles of basal resources.
+      In empirical webs, basal resources are often highly aggregated. 
+      Improved resolution at the basal level should improve the correspondence between simulated and empirical webs and allow better modelling of empirical communities.
 
@@
      \centering
-     \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP0.pdf}
+     \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP0.eps}
      \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
@@
      \centering
-     \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP1.pdf}
+     \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP1.eps}
      \caption{Basal resources experience a disturbance increasing their probability to go extinct to $\pi = 0.5$.
@@
         \centering
-            \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.pdf}
+            \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.eps}
             \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all connectances - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.

### 2022-06-14 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
         \centering
-        \includegraphics[width=1.0\textwidth]{manuscript/figures/concept_fig_ver2.png}
+        \includegraphics[width=1.0\textwidth]{figures/concept_fig_ver2.eps}
         \caption{Conceptual figure of a toy food web.
@@
     \centering
-        \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.pdf}
+        \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.eps}
         \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
@@
         \centering
-        \includegraphics[width=\textwidth]{manuscript/figures/motif_proportion_lms.pdf}
+        \includegraphics[width=\textwidth]{figures/motif_proportion_lms.eps}
         \caption{The proportions of the four motifs in a network's motif profile were related to network size, connectance, and their interaction.

### 2022-06-15 manuscript/response_PRSB_round2.tex
Updates from Overleaf

@@
 ``Species motif participation provides unique information about species risk of extinction''\\
-Anna {\AA}kesson, Alyssa R. Cirtwill, Kate L. Wootton \& Anna Ekl\"of
+Alyssa R. Cirtwill, Anna {\AA}kesson, Kate L. Wootton \& Anna Ekl\"of
 \end{flushleft}

### 2022-06-15 manuscript/cover_letter_2.tex
Updates from Overleaf

\documentclass[12pt]{letter}
\usepackage[britdate]{UH-letter}
\usepackage{times}
\usepackage{letterbib}
\usepackage{geometry}
\usepackage[round]{natbib}
\usepackage{graphicx}
\geometry{a4paper}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage[running]{lineno}
\usepackage{amsmath,amsfonts,amssymb}
% \usepackage[margin=10pt,font=small,labelfont=bf]{caption}

\newenvironment{refquote}{\bigskip \begin{it}}{\end{it}\smallskip}

\newenvironment{figure}{}

\position{Postdoctoral fellow}
\department{Department of Agricultural Sciences}
\location{University of Helsinki}
\cityzip{Helsinki, Finland}
\telephone{}
\fax{}
\email{alyssa.cirtwill@helsinki.fi}
\url{http://cirtwill.github.io}
\name{Dr. Alyssa R. Cirtwill}

\begin{document}

\begin{letter}{\bf
Prof.~Barrett\\
Editor-in-Chief\\
Proceedings of the Royal Society B} % Name/title of the addressee

\opening{Dear Prof. Barrett,}

We appreciate the opportunity to provide a revision of our manuscript ``Species motif participation provides unique information about species risk of extinction'' for publication in \emph{Proceedings of the Royal Society B}. 
The feedback by the Associate Editor and two external Reviewers was insightful and, we feel, has strengthened our manuscript.
We respond in detail to all comments in a separate response letter (attached).

Please note that the initial corresponding author, Anna \r{A}kesson, is leaving academia and will have limited ability to respond to any reader queries.
We have therefore, with the consent of all concerned, changed corresponding authorship to the other joint first author, Alyssa Cirtwill.
At the same time, we have reordered these joint first authors so that the corresponding author continues to appear first.
We hope that this change does not create any administrative issues.

We thank you once again for the helpful comments on our previous revision and for the opportunity to submit a revision.

On behalf of Anna {\AA}kesson, Kate Wootton, and Anna Ekl\"of,

\closing{Sincerely,}
\vspace{4\parskip}
\centering

\end{letter} 
\end{document}

### 2022-06-16 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 Faculty of Biological and Environmental Sciences\\
-University of Helsinki
+University of Helsinki, Helsinki, Finland\\
 \medskip
 \small$^2$Department of Theoretical Biology, Chemistry, and Physics\\ 
-Link\"{o}ping University\\
-Link\"{o}ping, Sweden\\
+Link\"{o}ping University, Link\"{o}ping, Sweden\\
 \medskip
 \small$^3$ BioFrontiers Institute\\
-University of Colorado, Boulder\\
-U.S.A.
+University of Colorado, Boulder, Boulder, U.S.A.\\
 \medskip
-* Joint first authors
+* Joint first authors\\
 \medskip

### 2022-06-16 manuscript/supplement_motifs.tex
added globalweb ref
Updates from Overleaf

@@
 
-\title{Motif roles provides unique information about species risk of extinction\\ \medskip Supplementary Information}
+\title{Species motif participation provides unique information about species risk of extinction}
 
-\author{Anna \r{A}kesson$^{1\dagger}$, Alyssa R. Cirtwill$^{2}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{1}$} 
-\date{\small$^1$Department of Theoretical Biology, Chemistry, and Physics\\ 
-Link\"{o}ping University\\
-Link\"{o}ping, Sweden\\
+\author{Alyssa R. Cirtwill$^{1*\dagger}$, Anna \r{A}kesson$^{2*}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{2}$} 
+\date{
+\small$^1$Spatial Foodweb Ecology Group\\
+Research Centre for Ecological Change\\
+Organismal and Evolutionary Biology Research Programm\\
+Faculty of Biological and Environmental Sciences\\
+University of Helsinki, Helsinki, Finland\\
 \medskip
-\small$^2$Department of Agricultural Sciences\\
-University of Helsinki\\
-Helsinki, Finland\\
+\small$^2$Department of Theoretical Biology, Chemistry, and Physics\\ 
+Link\"{o}ping University, Link\"{o}ping, Sweden\\
 \medskip
 \small$^3$ BioFrontiers Institute\\
-University of Colorado, Boulder\\
-U.S.A.
+University of Colorado, Boulder, Boulder, U.S.A.\\
 \medskip
-$^\dagger$ Corresponding author:\\
+* Joint first authors\\
+\medskip
+$^\dagger$ Corresponding author\\
 }
@@
 \section*{Table of contents}
+    {\footnotesize
     \begin{spacing}{1.0}
@@
 \end{spacing}
+}
 \clearpage
@@
         In our ecological setting, the nodes of the network represent species and links represent feeding interactions between species.
-        In network-theory terms, the species are Bernoulli random variables, and the links represent conditional dependencies among these variables. 
-        Each node has a set of local probability dependencies, conditional on the state of the parent nodes. 
+        In network-theory terms, the species are Bernoulli random variables and the links represent conditional dependencies among these variables. 
+        Each node has a set of local probability dependencies that are conditional on the state of the parent (resource) nodes. 
         Since the nodes are species, their state can be either extant or extinct.
-        The extinction probability of each species is a function of the state of its resources and also depends on a baseline probability of extinction ($\pi$) representing risk of extinction due to factors unrelated to network structure.
+        The extinction probability of each species is therefore a function of the state of its resources and also depends on a baseline probability of extinction ($\pi$) representing risk of extinction due to factors unrelated to network structure.
         In our baseline scenario, basal resources had a baseline extinction probability of $\pi = 0.1$. 
@@
         This will in turn affect the extinction probability $P(\lnot C|f)$. 
-        For a basal species, we assume that abiotic resources are always available. Therefore, a primary producer is independent of resource species and will always have $f = 0$.
+        For a basal species, we assume that abiotic resources are always available. 
+        Therefore, a primary producer is independent of resource species and will always have $f = 0$.
 
@@
         We draw a random number $r$ from a uniform distribution between $0-1$.
-The basal species would be considered extinct if $r < 0.1$, and extant otherwise.
+        The basal species would be considered extinct if $r < 0.1$ and extant otherwise.
 
@@
         We generated a total of 100 networks with each combination of parameters, for a total of 6000 networks. 
-All networks were generated using the function "nichemodel" within the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
+        All networks were generated using the function ``nichemodel'' within the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}. 
         If a simulated network contained any disconnected species (species without predators or prey) or disconnected components (a group of species connected among themselves but not to the rest of the network), the network was rejected and a new network simulated.
@@
 
-        After generating the network structure, we simulated community dynamics using the function ``simulate'' from the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
+        As a further check that our simulated networks are realistic (e.g., could plausibly retain all species), we simulated community dynamics for 500 timesteps and discarded any networks which did not retain all species.
+        We simulated dynamics using the function ``simulate'' from the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
         This function uses Lotka-Volterra predator-prey models including density dependence and type 2 functional responses for all species (please see~\citet{Delmas2017} for full details).
@@
         The persistence of each species in our simulated networks also depends on its population biomass. 
-        We randomly assigned initial population biomasses (i.e.
-cumulative biomass across all individuals of a species) for each species from a uniform distribution [0,1].
+        We randomly assigned initial population biomasses (i.e. cumulative biomass across all individuals of a species) for each species from a uniform distribution [0,1]. 
         Note that population biomasses and individual body masses are not calculated on the same scale. 
@@
         To ensure that species did not `recover' from unrealistically low biomasses during the simulation, we considered a species extinct if it dropped below an arbitrary threshold biomass of 1$\times10^{-5}$.
-When simulating initial (i.e.
-pre-disturbance or equilibrium) dynamics, we rejected any network where one or more species dropped below this biomass threshold.
+        We rejected any network where one or more species dropped below this biomass threshold. 
         Consumers were assumed to have no preferences such that the consumption rate $w_{ij}$ of predator $i$ eating prey $j$ is equal to $\frac{1}{n}$, where $n$ is the number of prey for predator $i$.
@@
         After simulating our initial niche-model networks, we filtered networks for biological plausibility.
-        Specifically, we removed any network containing disconnected components
-        (species or groups of species not connected to the rest of the network) 
-        or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
+        Specifically, we removed any network containing disconnected components (species or groups of species not connected to the rest of the network) or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
         Any removed network was replaced with a fresh simulation and re-checked until 100 realistic networks for each size-connectance combination had been obtained.
@@
         All persistence calculations therefore begin by determining the status of the basal resources (who do not depend on other species).
-        Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on basal resources), and so on up the network.
+        Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on basal resources), then herbivores, and so on up the network.
 
@@
         \begin{figure}
+            \centering
             \includegraphics[width=\textwidth]{figures/motif_profiles_participation_vs_empirical.eps}
@@
     \begin{figure}[hb!]
+        \centering
       \includegraphics[width=\textwidth]{figures/heatmap_CS_BPcompare.eps}
@@
 
-% \section{SH: Persistence and motif roles}
-
-%     \subsection*{Methods}
-
-%         As one more level of detail, we can consider the frequency with which a species appears in each \emph{position} in each motif. 
-%         Different positions are associated with different numbers of predators and prey, making the similarities and differences between motifs and degree more explicit.
-%         The full set of these positions, across all motifs, is the species' \emph{motif role}~\citep{Stouffer2010,Cirtwill2017}.
-%         As with motif participation, we normalized motif roles by dividing the number of times a species appeared in each position by the total number of times it appeared in all positions~\citep{Baker2015,Cirtwill2017}.
-
-%         [[Could add a PERMANOVA to show that species' roles overal affect their persistence. Haven't done so yet.]]
-
-%         To test whether participating in different motif positions was related to a species' probability of persistence, we fit similar models to those for motif participation.
-%         For each position, we fit a linear mixed-effect model relating persistence to the frequency of the position in a species' motif role, disturbance, and their interaction.
-%         To account for differences in persistence across networks of different size and connectance, we included a random effect of the interaction between size and connectance.
-%         All regressions were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-
-%     \subsection*{Results}
-
-%     \begin{table}[h!]
-%         \caption{Coefficients in a series of regressions relating probability of persistence to the frequency of a position in a species' motif role, disturbance, and their interaction.
-For ease of comparison between motifs, we did not center or scale predictors.
-Note, however, that some motifs and positions have higher frequencies than others.}
-%         \label{tab:persistence_vs_positions}
-%         \centering
-%         \begin{tabular}{l | c c c c}
-%         Position    & Intercept & Disturbance   & Position  & Interaction \\
-%         \hline
-%         Chain: top  & 0.880 &   -0.985 & -0.00544 & -0.332 \\
-%         Chain: bottom   & 0.883 &   -1.07 & -0.0510 &   0.647 \\
-%         Chain: middle   & 0.862 & -1.00 &   0.246 & -0.244 \\
-%         \hline
-%         Dir. Comp.: top & 0.872 &   -0.995  & 0.0549 &  -0.178 \\
-%         Dir. Comp.: bottom & 0.886 &    -1.07 & -0.113 &    0.939 \\
-%         \hline
-%         Omnivory: top   & 0.872 &   -0.984 &    0.180 & -0.818 \\
-%         Omnivory: middle & 0.873 & -1.00 &  0.119 & -0.324 \\
-%         Omnivory: bottom & 0.882 & -1.06 &  -0.0449 &   0.845 \\
-%         \hline
-%         App. Comp.: top & 0.865 &   -0.966 &    0.140 & -0.520 \\
-%         App. Comp.: bottom & 0.910 &    -1.12 & -0.100 &    0.323 \\
-%         \hline
-%         \end{tabular}
-%     \end{table}
-
-%     \begin{figure}
-%         \centering
-%         \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Apparent.eps}
-%         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
-Here we show the relationships between persistence and the proportions of the three positions in the apparent competition motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
-Line style indicates the motif  position.
-}
-%         \label{fig:apparent_positions}
-%     \end{figure}
-
-%     \begin{figure}
-%         \centering
-%         \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Direct.eps}
-%         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
-Here we show the relationships between persistence and the proportions of the three positions in the direct competition motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
-Line style indicates the motif  position.
-}
-%         \label{fig:direct_positions}
-%     \end{figure}
-
-%     \begin{figure}
-%         \centering
-%         \includegraphics[width=0.75\textwidth]{figures/persistence_positions_Three-species.eps}
-%         \caption{A species' probability of persistence varied with the frequency of different motif positions in its role.
-Here we show the relationships between persistence and the proportions of the three positions in the three-species chain motif at three levels of probability of extinction for basal resources (p=0.1, 0.26, and 0.42), indicated by line colour.
-Line style indicates the motif  position.
-}
-%         \label{fig:chain_positions}
-%     \end{figure}
-
-% \clearpage
-
-% \section{SI: Motif roles and network properties}
-
-%     \subsection*{Methods}
-
-%         As with network motif profiles and motif participation, species' motif roles may vary with network size and connectance and with the species' in degree and trophic level.
-%         To identify these relationships, we fit two linear regressions for each motif position.
-%         The first related the frequency of the position in species' roles to network size, connectance, and their interaction.
-%         The second related the frequency of the position to in-degree, STL, and their interaction.
-%         All regressions were fit using the R~\citep{R} base function `lm'.
-
-%     \subsection*{Results}
-%         \subsection*{Degree and trophic level}
-
-%             Positions in all three motifs were strongly related to in-degree and/or trophic level (Table~\ref{DT_tab}).
-%             The top position in the omnivory motif increased with increasing degree across all trophic levels.
-%             The frequency of the middle position increased with the interaction of degree and trophic level.
-%             The bottom position was generaly rare, especially for species with high degree or high trophic level (this suggests that the bottom position is frequently a basal resource).
-
-%             The frequency of the bottom position in the apparent competition motif decreased strongly with increasing degree, especially for species at a high trophic level.
-%             The frequency of the top position increased strongly with increasing degree and was not strongly related to trophic level.
-
-%             The frequency of the bottom position in the direct competition motif increased with increasing degree, especially for species at a high trophic level.
-%             The bottom position was very rare, especially for species with high degree (also suggesting that this position is usually a basal resource).
-
-%             The bottom position in the three-species chain was generally rare, especially for species with high degree or trophic level (again, this is likely to commonly be basal resources).
-%             The middle position in the three-species chain did not vary strongly with degree and declined slightly with increasing trophic level.
-%             The top position in the three-species chain increased with trophic level, degree, and their interaction.
-
-%             \begin{table}[h!]
-%                 \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against in-degree, STL, and their interaction.}
-%                 \label{DT_tab}
-%                 \footnotesize
-%                 \begin{tabular}{c c | c c | cc | cc |}
-%                 Motif & Position & \multicolumn{2}{c}{Degree} & \multicolumn{2}{c}{STL} & \multicolumn{2}{c}{Interaction} \\
-%                 & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
-%                 \hline
-%                 Chain & bottom & -1.47 $\times10^{-5}$ & 0.769 & -0.0218 & \textless0.001 & -0.00160 & \textless0.001 \\
-%                 Chain & middle & 3.13$\times10^{-4}$ & \textless0.001 & -0.0353 & \textless0.001 & 1.85$\times10^{-5}$ & 0.275 \\
-%                 Chain & top & -0.0172 & \textless0.001 & 0.0259 & \textless0.001 & 0.00919 & \textless0.001 \\
-%                 \hline
-%                 Omnivory & bottom & 6.14$\times10^{-4}$ & \textless0.001 & -0.00270 & \textless0.001 & -0.00112 & \textless0.001 \\
-%                 Omnivory & middle & 3.13$\times10^{-4}$ & \textless0.001 & -0.0353 & \textless0.001 & 1.85$\times10^{-5}$ & \textless0.001 \\
-%                 Omnivory & top & 0.00413 & \textless0.001 & 0.00381 & \textless0.001 & 9.96$\times10^{-5}$ & \textless0.001 \\
-%                 \hline
-%                 App. comp. & bottom & -4.74$\times10^{-4}$ & \textless0.001 & -0.663 & \textless0.001 & 0.00224 & \textless0.001 \\
-%                 App. comp. & top & 2.12$\times10^{-5}$ & 0.183 & -0.112 & \textless0.001 & -4.47$\times10^{-4}$ & \textless0.001 \\
-%                 \hline
-%                 Dir. comp. & bottom & 8.12$\times10^{-4}$ & \textless0.001 & -5.57$\times10^{-3}$ & \textless0.001 & -1.37$\times10^{-3}$ & \textless0.001 \\
-%                 Dir. comp. & top & -0.00405 & \textless0.001 & -0.0435 & \textless0.001 & 0.00260 & \textless0.001 \\
-%                 \hline
-%                 \end{tabular}
-%                 \end{table}
-
-%             \begin{figure}[h!]
-%                 \centering
-%                 \includegraphics[width=.75\textwidth]{figures/positions_byTL_Omnivory.eps}
-%                 \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
-%                 \label{omnivory_DT}
-%             \end{figure}
-
-%             \begin{figure}[h!]
-%                 \centering
-%                 \includegraphics[width=.75\textwidth]{figures/positions_byTL_Apparentcompetition.eps}
-%                 \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
-%                 \label{appcomp_DT}
-%             \end{figure}
-
-%             \begin{figure}[h!]
-%                 \centering
-%                 \includegraphics[width=.75\textwidth]{figures/positions_byTL_Directcompetition.eps}
-%                 \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
-%                 \label{dircomp_DT}
-%             \end{figure}
-
-%             \begin{figure}[h!]
-%                 \centering
-%                 \includegraphics[width=.75\textwidth]{figures/positions_byTL_Three-specieschain.eps}
-%                 \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
-%                 \label{chain_DT}
-%             \end{figure}
-
-%         \subsection*{Network size and connectance}
-
-%             % Many of these are significant, but the slopes tend to be very shallow. TL and degree are more interesting - this is DEFINITELY SI.
-%             The frequency of the top position in the apparent competition motif did not vary strongly with network size or connectance (Table~\ref{SC_tab}).
-%             The frequency of the bottom position in the apparent competition motif decreased with increasing connectance but did not vary strongly vary with network size.
-
-%             The frequency of the top position in the direct competition motif increased slightly with the interaction of network size and connectance (i.e., in large and highly-connected networks).
-%             The frequency of the bottom position in the direct competition motif was low in all networks.
-
-%             The frequency of all three positions in the omnivory motif showed similar relationships with network size and connectance.
-%             Frequencies of these motifs increased with increasing connectance and were not strongly affected by network size.
-%             None of the positions in the three-species chain varied strongly with network size or connectance.
-
-%             \begin{table}[h!]
-%                 \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against network size, connectance, and their interaction.}
-%                 \label{SC_tab}
-%                 \footnotesize
-%                 \begin{tabular}{c c | c c | cc | cc |}
-%                 Motif & Position & \multicolumn{2}{c}{Size} & \multicolumn{2}{c}{Connectance} & \multicolumn{2}{c}{Interaction} \\
-%                 & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
-%                 \hline
-%                 Chain & bottom & -4.63$\times10^{-5}$ & \textless0.001 & -0.108 & \textless0.001 & -1.94$\times10^{-4}$ & 0.035 \\
-%                 Chain & middle & 1.68$\times10^{-4}$ & \textless0.001 & -0.0530 & \textless0.001 & -0.00212 & \textless0.001 \\
-%                 Chain & top & 1.00$\times10^{-4}$ & \textless0.001 & 0.0304 & 0.015 & -8.79$\times10^{-4}$ & \textless0.001 \\
-%                 \hline
-%                 Omnivory & bottom & 2.00$\times10^{-4}$ & \textless0.001 & 0.575 & \textless0.001 & -0.00240 & \textless0.001 \\
-%                 Omnivory & middle & 1.32$\times10^{-4}$ & \textless0.001 & 0.551 & \textless0.001 & -0.00215 & \textless0.001 \\
-%                 Omnivory & top & 6.42$\times10^{-5}$ & \textless0.001 & 0.448 & \textless0.001 & -0.00115 & \textless0.001 \\
-%                 \hline
-%                 App. comp. & bottom & -4.74$\times10^{-4}$ & \textless0.001 & -0.663 & \textless0.001 & 0.00224 & \textless0.001 \\
-%                 App. comp. & top & 2.12$\times10^{-5}$ & 0.183 & -0.112 & \textless0.001 & -4.47$\times10^{-4}$ & \textless0.001 \\
-%                 \hline
-%                 Dir. comp. & bottom & 7.94$\times10^{-5}$ & \textless0.001 & -0.0663 & \textless0.001 & 0.00101 & \textless0.001 \\
-%                 Dir. comp. & top & -2.45$\times10^{-4}$ & \textless0.001 & -0.602 & \textless0.001 & 0.00610 & \textless0.001 \\
-%                 \hline
-%                 \end{tabular}
-%                 \end{table}
-
-%             \begin{figure}[h!]
-%                 \centering
-%                 \includegraphics[width=.75\textwidth]{figures/positions_bySC_Omnivory.eps}
-%                 \caption{The frequencies of positions within the omnivory motif varied with degree, trophic level, and their interaction.}
-%                 \label{omnivory_SC}
-%             \end{figure}
-
-%             \begin{figure}[h!]
-%                 \centering
-%                 \includegraphics[width=.75\textwidth]{figures/positions_bySC_Apparentcompetition.eps}
-%                 \caption{The frequencies of positions within the apparent competition motif varied with degree, trophic level, and their interaction.}
-%                 \label{appcomp_SC}
-%             \end{figure}
-
-%             \begin{figure}[h!]
-%                 \centering
-%                 \includegraphics[width=.75\textwidth]{figures/positions_bySC_Directcompetition.eps}
-%                 \caption{The frequencies of positions within the direct competition motif varied with degree, trophic level, and their interaction.}
-%                 \label{dircomp_SC}
-%             \end{figure}
-
-%             \begin{figure}[h!]
-%                 \centering
-%                 \includegraphics[width=.75\textwidth]{figures/positions_bySC_Three-specieschain.eps}
-%                 \caption{The frequencies of positions within the three-species chain motif varied with degree, trophic level, and their interaction.}
-%                 \label{chain_SC}
-%             \end{figure}
-
-%     \clearpage
-
-% \clearpage
-
-% \section*{SJ: Proportion of basal resources in the network}
-
-%     \subsection*{Basal resources and mean persistence}
-
-%         We fit a linear regression relating consumer persistence to network size, connectance, disturbance, the proportion of the network made up of basal resources (pBasal), and all interactions between them.
-All terms were significant, indicating that the pBasal affects a consumer's probability of persistence but that this effect depends upon other aspects of network structure.
-%         As with the effects of network size and connectance, however, the effects of pBasal were small in the range of networks we considered here (Fig.~\ref{fig:lm_BCS}).
-%         % Table is from basal_models.Rdata: BCper, if we want it.
-
-%         \begin{figure}[h!]
-%             \centering
-%             \includegraphics[height=0.75\textheight]{figures/persistence_vs_BSC_lm.eps}
-%             \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while the proportion of the network made up of basal resources, network size, connectance, and their interactions had smaller effects.
-We show relationships with persistence for low disturbance ($\pi$=0.1) and high disturbance($\pi$=0.5).
-For each predictor other than disturbance, we show the relationship between the focal predictor and persistence for combinations of the lowest and highest observed values of the other predictors.
-For proportion of basal resources, these values were 0.02 and 0.5; for network size, 50 and 100; and for connectance, 0.02 and 0.18.}
-%             \label{fig:lm_BCS}
-%         \end{figure}
-
-%     \subsection*{Motif profiles and proportion of basal resources}
-
-%         We repeated our analyses testing whether motif profiles were related to network size and connectance, replacing network size with the proportion of the network made up by basal resources (pBasal).
-%         A network's overall motif profile was significantly related to pBasal ($F_{1,2996}$=196, $p$\textless0.001, connecance ($F_{1,2996}$=10.0, $p$\textless0.001), and their interaction ($F_{1,2996}$=27.3, $p$\textless0.001).
-%         However, this result could be a false positive as the dispersion of motif profiles was non-homogeneous across connectance ($F_{4,2995}$=5.17, $p$\textless0.001), pBasal ($F_{128,2871}$=19.0, $p$\textless0.001), and their interaction ($F_{291,2708}$=13.4, $p$\textless0.001).
-%         Distance from group centroids was low for networks with extremely low or extremely high pBasal, perhaps because there were few such networks.
-%         The ranges and median distances from group centroids was similar across levels of connectance.
-
-%         Considered separately, the proportions of all four motifs in a network's motif profile were related to the proportion of basal resources in a network (Fig.~\ref{fig:Blms}).
-%         The proportion of the three-species chain decreased in networks with more basal resources ($\beta_{pBasal}$=-0.080, $p$\textless0.001) and networks with higher connectance ($\beta_{C}$=0.498, $p$\textless0.001).
-%         The interaction between pBasal and connectance was not significant ($\beta_{pBasal:C}$=0.0554, $p$=0.766).
-
-%         The proportion of the omnivory motif decreased in networks with more basal resources ($\beta_{pBasal}$=-0.198, $p$\textless0.001), increased in networks with higher connectance ($\beta_{C}$=0.895, $p$\textless0.001), and increased most strongly in networks with high connectance and many basal resources ($\beta_{pBasal:C}$=2.53, $p$\textless0.001).
-
-%         The proportion of the apparent competition motif increased with in networks with more basal resources ($\beta_{pBasal}$=0.202, $p$\textless0.001) and decreased in more connected networks ($\beta_C$=-0.596, $p$\textless0.001).
-%         The interaction between the proportion of basal resources and connectance was not significant ($\beta_{pBasal:C}$=0.258, $p$\textless0.001).
-
-%         The proportion of the direct competition motif increased in networks with more basal resources ($\beta_{pBasal}$=0.0759, $p$\textless0.001) and more-connected networks ($\beta_C$=0.199, $p$\textless0.001), but decreased with the interaction between them ($\beta_{pBasal:C}$=-2.85, $p$\textless0.001).
-
-%         \begin{figure}
-%             \centering
-%             \includegraphics[width=.75\textwidth]{figures/motif_proportion_basal_lms.eps}
-%             \caption{Mean proportion of each motif in a network's motif profile varied with connectance and the proportion of basal resources in the network.
-For direct competition and omnivory, the frequency of the motif also varied with the interaction between connectance and basal resources.}
-%             \label{fig:Blms}
-%         \end{figure}
-
-%     \subsection*{Motif participation and proportion of basal resources}
-
-%         We fit regressions relating the proportion of each motif in a consumer's motif participation vector to the proportion of basal resources in the network (pBasal), connectance, and the interaction between them.
-%         Participation in all four motifs was significantly related to pBasal, connectance, and the interaction between them (Fig.~\ref{fig:basal_participation}).
-%         These interactions were strongest for apparent and direct competition.
-
-%         \begin{figure}
-%             \centering
-%             \includegraphics[height=.65\textheight]{figures/participation_vs_BC.eps}
-%             \caption{The proportion of each motif in a consumer's participation vector varied with the proportion of the network made up by basal resources, connectance, and their interaction.}
-%             \label{fig:basal_participation}
-%         \end{figure}
-
-%         Participation in the three-species chain increased in networks with more basal resources ($\beta_{pBasal}$=0.0447, $p$\textless0.001), decreased in more-connected networks ($\beta_C$=-0.296, $p$\textless0.001), and decreased with the interaction between them ($\beta_{pBasal:C}$=-0.525, $p$\textless0.001).
-%         Participation in the omnivory motif decreased in networks with more basal resources ($\beta_{pBasal}$=-0.198, $p$\textless0.001), increased in more-connected networks ($\beta_C$=0.701, $p$\textless0.001), and increased strongly with the interaction between them ($\beta_{pBasal:C}$=2.81, $p$\textless0.001).
-%         Participation in the apparent competition motif decreased in networks with more basal resources and more-connected networks ($\beta_{pBasal}$=-0.0825, $p$\textless0.001 and $\beta_C$=-0.846, $p$\textless0.001) but increased strongly with the interaction between them ($\beta_{pBasal:C}$=1.66, $p$\textless0.001).
-%         Participation in the direct competition motif increased in networks with more basal resources and more-connected networks ($\beta_{pBasal}$=0.235, $p$\textless0.001 and $\beta_C$=0.442, $p$\textless0.002), but decreased strongly with the interaction between them ($\beta_{pBasal:C}$=-3.94, $p$\textless0.001).
-
-%     \subsection*{Frequencies of positions and proportion of basal resources}
-
-%         As with the frequencies of each motif in a species' motif participation vector, the frequencies of positions in a species' motif role varied with the proportion of basal resources in the network, connectance, and their interaction (Table~\ref{BC_tab}).
-
-%         \begin{table}[h!]
-%             \caption{Coefficients ($\beta$) and $p$-values for linear regressions of the frequency of each motif position against the proportion of the network made up of basal resources (pBasal), connectance, and their interaction.}
-%             \label{BC_tab}
-%             \footnotesize
-%             \begin{tabular}{c c | c c | cc | cc |}
-%             Motif & Position & \multicolumn{2}{c}{pBasal} & \multicolumn{2}{c}{Connectance} & \multicolumn{2}{c}{Interaction} \\
-%             & & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
-%             \hline
-%             Chain & bottom & -0.116 & \textless0.001 & -0.266 & \textless0.001 & 0.217 & \textless0.001 \\
-%             Chain & middle & 0.0779 & \textless0.001 & -0.107 & \textless0.001 & -0.386 & \textless0.001 \\
-%             Chain & top & 0.0829 & \textless0.001 & 0.0776 & \textless0.001 & -0.356 & \textless0.001 \\
-%             \hline
-%             Omnivory & bottom & -0.104 & \textless0.001 & 0.203 & \textless0.001 & 0.937 & \textless0.001 \\
-%             Omnivory & middle & -0.0668 & \textless0.001 & 0.217 & \textless0.001 & 1.24 & \textless0.001 \\
-%             Omnivory & top & -0.0269 & \textless0.001 & 0.281 & \textless0.001 & 0.638 & \textless0.001 \\
-%             \hline
-%             App. comp. & bottom & -0.237 & \textless0.001 & -0.874 & \textless0.001 & 1.76 & \textless0.001 \\
-%             App. comp. & top & 0.154 & \textless0.001 & 0.0279 & \textless0.001 & -0.0947 & 0.013 \\
-%             \hline
-%             Dir. comp. & bottom & -0.0631 & \textless0.001 & -0.0247 & \textless0.001 & -0.418 & \textless0.001 \\
-%             Dir. comp. & top & 0.298 & \textless0.001 & 0.466 & \textless0.001 & =3.53 & \textless0.001 \\
-%             \hline
-%             \end{tabular}
-%             \end{table}
-% \clearpage
-
-% \section*{SK: Motif profiles in different network sets}
-
-%     Not all motifs were equally likely to occur in the networks we considered. 
-%     Across the initially stable, acyclic networks (i.e., those used in the main text), the omnivory motif tended to make up a smaller proportion of the motif profile than other motifs.
-%     The omnivory motif was also rarer than the three-species chain and apparent competition motifs before networks were rendered acyclic, but made up a similar proportion to the direct competition motif (Table~\ref{tab:acyclic_proportions}).
-%     The omnivory motif made up a larger proportion of the full set of initially stable networks, including those which were excluded from our final analysis due to overly high trophic levels.
-%     Nevertheless, in each group of networks the apparent competition and three-species chain motifs made up the largest proportion of the motif profile.
-%     This suggests that, while rendering networks acyclic does change proportions slightly, stable networks created by the niche model appear to be generally biased towards the chain and apparent competition motifs.
-
-%     \begin{table}[h!]
-%         \centering
-%         \caption{Proportions of the four motifs which appear in Bayesian networks (three-species chain: `Chain', apparent competition: `AC', direct competition `DC', and omnivory: `Omni') and other motifs (`Other') in the motif profiles of acyclic, filtered, and original networks.
-Acyclic networks are those used in the main text analyses.
-Filtered networks are those used in the main text analyses \emph{before} being rendered acyclic.
-Original networks are all simulated networks, including those removed because of overly high maximum trophic levels.}
-%         \label{tab:acyclic_proportions}            \footnotesize
-%         \begin{tabular}{c|l l l | l l l | l l l |}
-%         & \multicolumn{3}{c|}{Acyclic networks} & \multicolumn{3}{c|}{Filtered networks} & \multicolumn{3}{c|}{Original networks} \\
-%         Motif & Mean (SD) & Min & Max & Mean (SD) & Min & Max & Mean (SD) & Min & Max \\
-%         \hline
-%         Chain & 0.235 (0.042) & 0.109 & 0.412 & 0.225 (0.045) & 0.100 & 0.412 & 0.218 (0.043) & 0.096 & 0.413 \\
-%         AC & 0.422 (0.064) & 0.250 & 0.765 &
-%         0.417 (0.067) & 0.233 & 0.765 & 0.397 (0.064) & 0.231 & 0.765 \\
-%         DC & 0.179 (0.041) & 0.050 & 0.359 & 0.155 (0.046) & 0.040 & 0.359 & 0.140 (0.046) & 0.031 & 0.359 \\
-%         Omni & 0.164 (0.079) & 0.006 & 0.346 & 0.156 (0.075) & 0.006 & 0.319 & 0.184 (0.073) & 0.006 & 0.329 \\
-%         \hline
-%         Other & \multicolumn{3}{c|}{NA} & 0.047 (0.042) & 0.00 & 0.200 &
-%         0.062 (0.046) & 0.00 & 0.258 \\
-%         \end{tabular}
-%     \end{table}
-% \clearpage
-
+\bibliographystyle{vancouver} 
 \bibliography{anna_bib_new} % Abbreviate journal titles.
-\bibliographystyle{ecollett}
 

### 2022-06-16 manuscript/response_PRSB_round2.tex
corrected empirical connectance range

@@
 We compare the structure of our simulated networks to the largest set of high-quality, similarly-sized empirical networks we could find.
-These networks had 50-77 species and connectances of 0.2-0.214.
-This is at the upper range of connectances we consider (probably due to aggregation of basal resources into taxonomically coarse nodes in the empirical webs).
+These networks had 50-77 species and connectances of 0.024-0.089, squarely within the range we simulate.
 We also, per the Associate Editor's request, compare the motif profiles and motif participation of these empirical webs to our simulated webs.

### 2022-06-16 manuscript/supplement_motifs.tex

@@
 
-        Although previous work has shown that the niche model replicates the motif structure of empirical networks~\citep{Stouffer2005a,Stouffer2006}, we provide a comparison of the motif profiles and consumer motif participation for our simulated networks and the highest-quality empirical networks we could find within the same size range.
+        Although previous work has shown that the niche model replicates the motif structure of empirical networks~\citep{Stouffer2005a,Stouffer2006}, we additionally provide a comparison of the motif profiles and consumer motif participation for our simulated networks and the highest-quality empirical networks we could find within the same size range.
         For this comparison, we obtained all networks containing 50-100 species from \citet{globalwebdb}.
-        These webs ranged in connectance from 0.024-0.089
+        These webs ranged in connectance from 0.024-0.089.
         We then removed networks where sets of predators and prey were completely different (i.e., bipartite networks) or links were not clear (i.e., +/- effects rather than predation specifically).

### 2022-06-16 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
         For this comparison, we obtained all networks containing 50-100 species from \citet{globalwebdb}.
-        These webs ranged in connectance from 0.024-0.089.
         We then removed networks where sets of predators and prey were completely different (i.e., bipartite networks) or links were not clear (i.e., +/- effects rather than predation specifically).
         This yielded five vertebrate-dominated food webs describing the California coast in different years, four invertebrate-dominated food webs describing mainland US streams, 23 invertebrate-dominated food webs describing New Zealand streams, and five other invertebrate-dominated food webs.
-        This final set of webs had 50-77 species and connectances of 0.2-0.21 - the upper range of connectance in our simulations.
+        This final set of webs had 50-77 species and connectances (after rendering the networks acyclic) of 0.095-0.227 - covering the upper half of the connectance range we simulate, as well as slightly higher connectances.
         This may be partly due to relatively poor resolution of basal resources in these empirical webs: aggregation of many taxa into nodes such as `phytoplankton' will tend to decrease the number of species but leave the number of links unchanged, raising connectance.

### 2022-06-17 code/approx.R
reran models for nonlinear

@@ -82,7 +82,7 @@ for (i in 1:length(species)){
                 # persistence calculations
                 persistence <- tibble(
                     sp_no=rownames(A), ## species numbers
-                    per = bayes_ntw_MC(A, Pb, 1, 1, 100000) %>% # input: number of species, 
+                    per = bayes_ntw_MC(A, Pb, 3, 3, 100000) %>% # input: number of species, 
                                                                 # probability of extinction vector, 
                                                                 # alpha and beta for the functional response of consumers, 
                                                                 # 100k evals of BN
@@ -100,7 +100,7 @@ for (i in 1:length(species)){
                 } else{
                     collect <- bind_rows(collect, df2)
                 }
-                save(collect, file=paste0("Results_restricted/Linear/", sp_no, "/", connectance, "/TESTweb_",web_no, ".RData"))
+                save(collect, file=paste0("Results_restricted/Nonlinear/", sp_no, "/", connectance, "/TESTweb_",web_no, ".RData"))
             }
         }
     }

### 2022-06-17 manuscript/response_PRSB_round2.tex
Updates from Overleaf

@@
         We have retained the reference to the pre-print in the discussion because this brief comparison of results using dynamical and Bayesian simulations appears to be exactly what the Reviewer has asked for.
-        As the Reviewer points out, different approaches to simulating disturbance can have different results, and we explicit mention the main differences in the Discussion. 
+        As the Reviewer points out, different approaches to simulating disturbance can have different results, and we explicitly mention the main differences in the Discussion. 
         As the pre-print is freely available to any who wish to critically examine it, we believe that this is a good compromise between maintaining conciseness of the present manuscript and comparing different simulation approaches.
@@
 We compare the structure of our simulated networks to the largest set of high-quality, similarly-sized empirical networks we could find.
-These networks had 50-77 species and connectances of 0.024-0.089, squarely within the range we simulate.
+These networks had 50-77 species and connectances of 0.095-0.227, covering the upper half of connectances we simulate.
 We also, per the Associate Editor's request, compare the motif profiles and motif participation of these empirical webs to our simulated webs.

### 2022-06-17 manuscript/reorder_main_text.tex

@@
     We show that a consumer species' participation in different motifs is related to its probability of persistence. 
-    Similarly, the overall motif profile of a network was related to the average persistence of consumer species.
+    Similarly, the overall motif profile of a network is related to the average persistence of consumer species.
     Notably, we found that networks containing a higher proportion of the `omnivory' motif tend to have lower mean persistence following disturbance, while more frequent participation in the omnivory motif was associated with higher species-level probabilities of persistence.
@@
      The plant community is the foundation upon which the myriad species in food webs depend for their survival.
-     Disturbances in the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
+     Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
      As such, the structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
@@
     %\subsection*{2 Global and local network properties are known to affect stability}
-    How disturbances, such as population decline or extinction, lead to loss of other species via direct and indirect effects (i.e., secondary extinctions) has for long been vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    How disturbances, such as population decline or extinction, lead to loss of other species via direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
     Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances.
@@
 
-    Species-level properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
+    Species-level properties known to affect species' responses to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
     These can also be referred to as local properties of the network.
-    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and disturbances to plant populations cause particularly many econdary extinctions \citep{Curtsdotter2011}.
+    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and disturbances to plant populations cause particularly many secondary extinctions \citep{Curtsdotter2011}.
 %including secondary extinction due to reduced plant populations.
@@
     A more detailed perspective is desirable since, as well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
-    For example, a generalist consuming many specialist (and therefore high-risk) prey may be at greater risk of extinction than a specialist which consumes a highly generalist resource that is more unlikely to go extinct (Fig.~\ref{fig:concept}).
+    For example, a generalist consuming many specialist (and therefore high-risk) prey may be at greater risk of extinction than a specialist which consumes a highly generalist resource that is less likely to go extinct (Fig.~\ref{fig:concept}).
 
@@
         Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
-        This increased speed allows us to simulate a range of disturbance severity rather than single species removals as in many previous studies (e.g.,~\citealp[]{Memmott2004,Staniczenko2010,Dunne2004}).
+        This increased speed allows us to simulate a range of disturbance severities rather than single species removals as in many previous studies (e.g.,~\citealp[]{Memmott2004,Staniczenko2010,Dunne2004}).
         Moreover, Bayesian network simulations are less parameter-intensive than dynamical models, meaning that they are less sensitive to assumptions made in the modelling process~\citep{Eklof2013}.
@@
 		% how we calculate persistences 
-        Beginning with basal resources and working systematically up through the food web, we use Equation \ref{betafunc} to calculate each species' $i$ probability of extinction $P(\lnot i|f)$.
+        Beginning with basal resources and working systematically up through the food web, we use Equation \ref{betafunc} to calculate each species $i$'s probability of extinction $P(\lnot i|f)$.
         Using this probability, we then perform Bernoulli trials to determine whether the species goes extinct or not. 
@@
 		First, in Bayesian networks consumers do not affect the extinction probabilities of their resources and as such we can here only capture bottom-up effects of disturbances. 
-		However, Bayesian networks has been shown to capture the majority of secondary extinctions produced fully-fledged dynamical models \citep{Eklof2013}.
-		\cite{Eklof2013} analyzed which functional form of consumers response to loss of resources that most accurately predicted the secondary extinctions realized in the dynamical model, and we here use that functional form in our analyzes.
+		However, Bayesian networks has been shown to capture the majority of secondary extinctions produced by fully-fledged dynamical models \citep{Eklof2013}.
+		\cite{Eklof2013} analyzed which functional form of consumers' response to loss of resources most accurately predicted the secondary extinctions realized in the dynamical model. 
+		We here use that functional form in our analyzes.  
 		Moreover, bottom-up effects are often the main force of secondary extinctions in empirical systems. 
-		For example, bottom-up effects regulate zoo-plankton communities in freshwater lakes, regulating zoo-plankton communities \citep{li2020bottom} and control abundances and biomass of vertebrate species in several terrestrial systems \citep{Dobson2009food, Mduma1999food, Georgiadis2007}.
-		Second, due to the strict bottom-up approach of Bayesian networks the networks must be acyclic.
+		For example, bottom-up effects regulate zooplankton communities in freshwater lakes, regulating zooplankton communities \citep{li2020bottom} and control abundances and biomass of vertebrate species in several terrestrial systems \citep{Dobson2009food, Mduma1999food, Georgiadis2007}.
+		Second, due to the strict bottom-up approach of Bayesian networks, the networks must be acyclic.
 		Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited.
@@
         In our baseline scenario, all species had an extinction probability of $\pi = 0.1$. 
-		To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase basal species' probability of extinction from $0.10$ to $0.5$, in steps of $0.08$. 
+		To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase all basal species' probability of extinction from $0.10$ to $0.5$, in steps of $0.08$. 
 		The highest disturbance level, $\pi = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
@@
 		For full details and a comparison to empirical networks, see \emph{Appendix S2}.
-		We removed any biologically unlikely networks ( \emph{Appendix S2}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
+		We removed any biologically unlikely networks (\emph{Appendix S2}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
 		To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and species topologically sorted from lowest to highest trophic level (\emph{Appendix S2}).
@@
 
-    The authors thank György Barabás for excellent discussion of the ideas in this manuscript.
-AÅ is supported by the Swedish Research Council for Sustainable Development Formas (\#2015-01262) granted to AE.
+    The authors thank Gy\"{o}rgy Barab\'{a}s for excellent discussion of the ideas in this manuscript.
+A{\AA} is supported by the Swedish Research Council for Sustainable Development Formas (\#2015-01262) granted to AE.
 ARC is supported by a Finnish Academy postdoctoral research grant (\#332999).
@@
 \section*{Figures}
-
- %   \begin{figure}[hb!]
- %      \centering
- %       \includegraphics[width=1.0\textwidth]{manuscript/figures/concept_motifs_ab.png}
- %       \caption{Conceptual figure of a toy food web.
-\textbf{a)} This web contains examples of the four motifs present in an acyclic Bayesian network: apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow).
-These motifs capture different ways in which disturbances to basal resources can propagate through the network.
-For example, a disturbance to the bottom species in a three-species chain will have a direct effect on the middle species but only indirectly affect the top species while a disturbance to the bottom species in an omnivory motif will have a direct effect on the middle species and direct and indirect effects on the top species.
-Note that the nodes participate in more motifs than the ones highlighted.
-\textbf{b)} In the same food web, species with similar degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than tree times more frequently.
-Inserted table shows the number of times A and B, respectively, participate in the four different motifs.}
- %   \label{fig:concept}
- %   \end{figure}
 

### 2022-06-17 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 
-        Species' motif participation may covary with other measures of network structure and species' roles within the network.
+        Species' motif participation may co-vary with other measures of network structure and species' roles within the network.
         To determine the relationships between motif participation and global network structure, we fit a set of four linear mixed-effect models (one per motif):
@@
         \label{tab:proportion}                \footnotesize
-        \begin{tabular}{c|c c | c c | c c | c c | c c |}
-        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c|}{$R^2$} \\
+        \begin{tabular}{c|c c | c c | c c | c c | c c }
+        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c}{$R^2$} \\
         Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
         \hline
-        3-sp Chain & 0.578 & \textless0.001 & -0.00360 & \textless0.001 & -0.0154 & \textless0.001 & 8.39$\times10^{-5}$& \textless0.001 & 0.866 & 0.914 \\
-        App. Com. & 0.579 & \textless0.001 & -0.00239 & \textless0.001 & -0.0154 & \textless0.001 & -2.70$\times10^{-5}$ & \textless0.001 & 0.822 & 0.882 \\
-        Dir. Com. & 0.577 & \textless0.001 & 7.74$\times10^{-4}$ & \textless0.001 & -0.0154 & \textless0.001 & -3.76$\times10^{-5}$ & \textless0.001 & 0.833 & 0.866 \\
-        Omnivory & 0.577 & \textless0.001 & -0.00146 & \textless0.001 & -0.0154 & \textless0.001 & 4.22$\times10^{-5}$ & \textless0.001 & 0.842 & 0.877 \\
+        3-sp Chain & 0.574 & \textless0.001 & -6.25$\times10^{-4}$ & \textless0.001 & -0.0190 & \textless0.001 & -5.86$\times10^{-5}$& \textless0.001 & 0.900 & 0.912 \\
+        App. Com. & 0.574 & \textless0.001 & -5.71$\times10^{-4}$ & \textless0.001 & -0.0190 & \textless0.001 & 8.19$\times10^{-5}$ & \textless0.001 & 0.900 & 0.911 \\
+        Dir. Com. & 0.574 & \textless0.001 & 6.95$\times10^{-4}$ & \textless0.001 & -0.0190 & \textless0.001 & 4.01$\times10^{-6}$ & \textless0.001 & 0.901 & 0.911 \\
+        Omnivory & 0.574 & \textless0.001 & 8.12$\times10^{-4}$ & \textless0.001 & -0.0190 & \textless0.001 & -2.73$\times10^{-5}$ & \textless0.001 & 0.894 & 0.912 \\
         \end{tabular}
@@
         \begin{tabular}{c|c c c | c c | c c | c c}
-             &  &  &  & \multicolumn{2}{c}{Size} & \multicolumn{2}{|c}{Connectance} & \multicolumn{2}{|c}{Interaction} \\
+             &  &  &  & \multicolumn{2}{c|}{Size} & \multicolumn{2}{c|}{Connectance} & \multicolumn{2}{c}{Interaction} \\
             Motif & Min & Max & Mean (SD) & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
@@
 All $p$-values were \textless0.001.
-The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.}
+The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.
+We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
         \label{tab:per_vs_TLdeg}
         \centering
-        \begin{tabular}{l|c  c  c |}
-            Simple role measure & Role & Disturbance & Interaction \\
+        \begin{tabular}{l|c  c  c | c c  }
+            Simple role measure & Role & Disturbance & Interaction & R$^2_M$ & R$^2_C$ \\
             \hline
-            STL & -0.0210 & -0.139 & 0.00508 \\
-            in-degree & -0.00188 & -0.139 & -0.00908 \\
+            STL & -0.0101 & -0.0190 & 3.34$\times10^{4}$ & 0.922 & 0.95 \\
+            in-degree & -0.0236 & -0.0190 & -0.0155 & 0.903 & 0.936 \\
         \end{tabular}
@@
      \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
-The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
+The panels show final persistence for consumers with different shortest trophic level, for increasing connectance (y-axis) and network size (x-axis).
 Persistence decreases with darker color in the heat map.}
@@
      \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP1.eps}
-     \caption{Basal resources experience a disturbance increasing their probability to go extinct to $\pi = 0.5$.
+     \caption{Basal resources experience a disturbance increasing their probability of extinction to $\pi = 0.5$.
 All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
-The sub figures show final persistence for consumers with different shortest trophic length, for increasing connectance (y-axis) and network size (x-axis).
+The panels show final persistence for consumers with different shortest trophic level, for increasing connectance (y-axis) and network size (x-axis).
 Persistence decreases with darker color in the heat map.}
@@
 3 and Table 1, \emph{Main Text}).
-    The same pattern holds overall for LRs fit within a network  (Figure~\ref{fig:density_prop_S})
-    : the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance. 
+    The same pattern holds overall for LRs fit within a network  (Figure~\ref{fig:density_prop_S}): the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance.
     This trend is most pronounced in large networks. (bottom panel, Figure~\ref{fig:density_prop_S}). 
@@
             \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.eps}
-            \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all connectances - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
+            \caption{Here we show the density (y-axis) of slopes (x-axis) for regressions of probability of persistence against proportion of different motifs for all simulated webs of all connectances - a visualization of how an increased proportion of each motif (different colored lines) affects the probability of persistence for consumer species.
 Columns show the result for various disturbances on the basal level, from $\pi = 0.1$ (left) to $\pi = 0.5$ (right).
@@
 A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
-The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend).
+The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot.
+Line color corresponds to the motif.
 }
@@
     \begin{table}[hb!]
-        \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
-Predictors were centered and scaled before model fitting.
-All effects were significant ($p$\textless0.05).
-We also provide the $R^2$ (multiple $R^2$).
+        \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile made up by a focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
+Predictors were centered and scaled before model fitting; coefficients refer to a unit increase in predictors.
+We also provide the marginal $R^2$ value for each regression.
 }
@@
         \centering
-        \begin{tabular}{c|c c c c c | c}
-            Motif & Intercept & Proportion & Disturbance & Interaction & $R^2$\\
+        \footnotesize
+        \begin{tabular}{c|c c | c c | c c | c c | c }
+        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \\
+        Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ \\
             \hline
-            Omnivory & 0.574 & -0.0140 & -0.139 & -5.59$\times10^{-4}$ & 0.968 \\
-            Apparent competition & 0.574 & 0.0112 & -0.139 & 5.90$\times10^{-4}$ & 0.964 \\
-            Direct competition & 0.574 & 3.38$\times10^{-3}$ & -0.139 & 1.81$\times10^{-4}$ & 0.958 \\
-            Three-species chain & 0.574 & 5.80$\times10^{-3}$ & -0.138 & -3.07$\times10^{-5}$ & 0.960 \\ 
+            Omnivory & 0.574 & \textless0.001 & 5.81$\times10^{-3}$ & \textless0.001 & -0.0189 & \textless0.001 & -1.775$\times10^{-7}$ & 0.885 & 0.960 \\
+            Apparent competition & 0.574 & \textless0.001 & 7.22$\times10^{-4}$ & \textless0.001 & -0.0189 & \textless0.001 & 5.20$\times10^{-6}$ & 0.003 & 0.964 \\
+            Direct competition & 0.574 & \textless0.001 &  1.38$\times10^{-4}$ & \textless0.001 & -0.0189 & \textless0.001 & 1.01$\times10^{-6}$ & 0.400 & 0.964 \\
+            Three-species chain & 0.574 & \textless0.001 & 7.22$\times10^{-4}$ & \textless0.001 & -0.0189 & \textless0.001 & 5.20$\times10^{-6}$ & 0.003 & 0.964 \\ 
         \end{tabular}
@@
     A network's normalized motif profile (i.e., proportions of the total made up by each motif) was related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
-    However, normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).These differences in variability can cause false positives in the PERMANOVA test.
-Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).
+    However, normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).
+    These differences in variability can cause false positives in the PERMANOVA test. Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).
 
@@
 Here we show the effect sizes ($\beta$) and $p$-values for the regression described above.
-All predictors were centered and scaled before fitting the model.
+All predictors were centered and scaled before fitting the model; coefficients refer to a unit increase in predictors.
 The level of disturbance corresponds to basal resources having a probability of extinction ranging between $0.1 - 0.5$ (see \emph{Methods}).
+The regression had $R^2$=
         }
@@
             Intercept & 0.574 & \textless0.001 \\
-            Size & -4.82$\times10^{-4}$  & 0.011 \\
-            Connectance & -0.0137 & \textless0.001 \\
-            Disturbance & -0.139 & \textless0.001 \\
-            S:C & -9.77$\times10^{-4}$ & \textless0.001 \\
-            S:D & -3.46$\times10^{-3}$ & \textless0.001 \\
-            C:D & -2.25$\times10^{-4}$ & 0.233 \\
-            S:C:D: & -3.97$\times10^{-4}$ & 0.036 \\
+            Size & -8.22$\times10^{-3}$  & 0.011 \\
+            Connectance & -7.74 & \textless0.001 \\
+            Disturbance & -0.0189 & \textless0.001 \\
+            S:C & -9.43$\times10^{-4}$ & \textless0.001 \\
+            S:D & -8.07$\times10^{-3}$ & \textless0.001 \\
+            C:D & -1.74$\times10^{-6}$ & 0.233 \\
+            S:C:D: & -5.24$\times10^{-5}$ & 0.036 \\
         \end{tabular}
@@
      Motif & Set of $n$ interacting species. In this case, $n=3$ \\
-     Network motif profile & Vector describing the frequency of each motif in the network. \\
-     & Normalised by dividing counts of each motif by the total across all motifs.\\
-     Motif participation role & Vector describing the frequency with which a focal species appears in each motif.\\
-     & Normalised by dividing counts for each motif by the total across all motifs. \\
-     Bayesian network & A directed acyclic graph, used to predict species' likelihood of persistence. \\
+     Network motif profile & Vector describing the frequency of each motif in the network (normalised by dividing counts of each motif by the total participation in  all motifs).\\
+     Motif participation role & Vector describing the frequency with which a focal species appears in each motif  (normalised by dividing counts for each motif by the total of all motifs). \\
+     Bayesian network & A directed acyclic graph, used to efficiently predict species' likelihood of persistence based on the state of their resources. \\
      Network mean persistence & The mean likelihood of persistence for all consumers in a network.\\
-     Species persistence & The individual likelihood of a species not going extinct.\\
+     Species persistence & The likelihood of a particular focal species not going extinct.\\
      In-degree & Number of prey species to a consumer species.\\
-     Trophic level (STL) & The shortest food chain between the focal species and any basal species.\\
-     Disturbance ($\pi_{disturbed}$) & Probability of extinction of a basal resource when extra disturbance is added. \\
-     Baseline extinction &  When no disturbance is applied, $\pi_{base}$=0.1.\\
+     Trophic level (STL) & Length of shortest food chain between the focal species and any basal resource.\\
+     Disturbance ($\pi_{disturbed}$) & Probability of extinction of a basal resource after a disturbance is added. \\
+     Baseline extinction &  Probability of extinction when no disturbance is applied, $\pi_{base}$=0.1.\\
  \end{tabular}}

### 2022-06-19 manuscript/response_PRSB_round2.tex
removing dead comments

@@
         As the Reviewer points out, different approaches to simulating disturbance can have different results, and we explicitly mention the main differences in the Discussion. 
-        As the pre-print is freely available to any who wish to critically examine it, we believe that this is a good compromise between maintaining conciseness of the present manuscript and comparing different simulation approaches.
-        If, all going well, the preprint is accepted for publication by the time the present manuscript is accepted, we will of course update the reference to provide the peer-reviewed version of the text.
+        As the pre-print is freely available to any who wish to critically examine it (even if the peer-reviewed version is delayed), we believe that this is a good compromise between maintaining conciseness of the present manuscript and comparing different simulation approaches.
+        If, all going well, the preprint is accepted for publication by the time the present manuscript is accepted, we will of course update the reference to the peer-reviewed version of the text.
 

### 2022-06-19 manuscript/reorder_main_text.tex
removing dead comments

@@
     Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and disturbances to plant populations cause particularly many secondary extinctions \citep{Curtsdotter2011}.
-%including secondary extinction due to reduced plant populations.
     However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
@@
     Here, we are interested in whether a consumer's extinction risk of going secondarily extinct depends on 1) the motifs in which the species participates, 2) the motif profile of the full food web and 3) the severity of the disturbance to the basal level.
-%We address this in scenarios with various levels of disturbance targeting the basal level.
     In addition, we want to understand how the information provided by motifs is related to other global and local food web properties in order to better synthesize information provided by different scales of network structure.
@@
     \subsection*{Species persistence and motif participation}
-    % Coefficients changed slightly with network-level random effect, but conclusions similar
+
        The relationship between frequency of participation in a motif and a species' probability of persistence depended on the level of disturbance. 
@@
     \subsection*{Motif participation and other network properties}
-        % Updated with network-level random effect. No visible change.
+
         Species' motif participation varied depending on the global properties of the network: size, connectance, and their interaction (Fig. S1, Table S2; \emph{Appendix S5}).
@@
 
-        % Updated with network-level random effect. No visible change.    
         Species' motif participation also varied with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
@@
         Species with higher trophic levels tended to have higher proportions of apparent competition and three-species chains. 
-        % Updated with network-level random effect. No visible change.   
         A species' persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix S6}). 
@@
 
-            %Update not needed
             When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally showed more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
@@
     \subsection*{Network mean persistence and motif profiles}
-        % Updated to use means instead of all species
+
         The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=645, $p$\textless0.001), although this result must be considered tentative due to non-homogeneous distribution of roles (\emph{Appendix S8}).
@@
         Conversely, average persistence was lower in networks with higher proportions of the omnivory motif. 
-        These trends were consistent across levels of disturbance (Fig.~\ref{fig:motif_profile_persistence}, Table S4), \emph{Appendix S8}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001).
-For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).
+        These trends were consistent across levels of disturbance (Fig.~\ref{fig:motif_profile_persistence}, Table S4), \emph{Appendix S8}).
 
@@
         Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S5, \emph{Appendix S9}).
-
-% Appendices SA (persistence vs.
-S,C), SH-I (analyses with motif positions), SJ-K (repeated analyses with number of basal resources instead of disturbance) not referred to in text.
-Keeping SH-K in commented-out form in case reviewers ask.
-We should refer to SA somewhere in Discussion (will be Appendix 10).
 

### 2022-06-19 manuscript/reorder_main_text.tex
Updates from Overleaf

--- 2022-06-17 manuscript/reorder_main_text.tex
@@
         Note that the nodes participate in more motifs than the ones highlighted.
-\textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than tree times more frequently (see inserted table for counts of participation in all four motifs).
+\textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than three times more frequently (see inserted table for counts of participation in all four motifs).
         Despite having equal degree and trophic level, species A may be at greater risk of going extinct than species B because of its dependence on specialist prey species.

### 2022-06-20 code/motif_SC_lms.py
fixed legend centering, submitted to PRSB

@@ -159,16 +159,16 @@ for lm in ['apparent','direct','omnivory','chain']:
   graph=populate_graph(graph,lm)
   graph.panel_label.configure(placement='ouc',char_size=1,dx=.03,dy=.01)
 
-grace.multi(rows=2,cols=2,vgap=.03,hgap=.04)
+grace.multi(rows=2,cols=2,vgap=.05,hgap=.04)
 grace.hide_redundant_labels()
-grace.set_row_xaxislabel(label='Connectance',row=1,colspan=(None,None),char_size=1,perpendicular_offset=.05,just=2)
-grace.set_col_yaxislabel(label='Proportion of total motifs',col=0,rowspan=(None,None),char_size=1,perpendicular_offset=.05,just=2)
+grace.set_col_yaxislabel(label='Proportion of total motifs',col=0,rowspan=(None,None),char_size=1,perpendicular_offset=.06,just=2)
 # grace.set_row_xaxislabel(label='Day of year',row=2,colspan=(None,None),char_size=1,perpendicular_offset=.05)
 for graph in grace.graphs:
   print graph.get_view()
-grace.graphs[0].set_view(0.15,0.45,0.53,0.70)
-grace.graphs[1].set_view(0.57,0.45,0.95,0.70)
-grace.graphs[2].set_view(0.15,0.15,0.53,0.40)
-grace.graphs[3].set_view(0.57,0.15,0.95,0.40)
+grace.graphs[0].set_view(0.15, 0.6010588235294116,0.53,0.95)
+grace.graphs[1].set_view(0.57, 0.6010588235294116,0.95,0.95)
+grace.graphs[2].set_view(0.15,0.18211764705882338,0.53,0.5310588235294116)
+grace.graphs[3].set_view(0.57,0.18211764705882338,0.95,0.5310588235294116)
+# grace.add_drawing_object(DrawText,text='Connectance',loctype='view',x=.55,y=0.08,char_size=1,just=2)
 grace.write_file('../manuscript/figures/motif_proportion_lms.eps')
 

### 2022-06-20 code/stat_analysis/persistence_vs_profiles.py

@@ -173,7 +173,7 @@ graph2.add_drawing_object(DrawText,x=0.5,y=0.34,loctype='world',text='Highest di
 
 graph2.legend.configure(char_size=1,loc=(0.05,0.685),loctype='world',box_linestyle=0,fill_pattern=0)
 
-grace.graphs[0].set_view(0.07,0.07,0.95,0.7)
+grace.graphs[0].set_view(0.15,0.15,0.95,0.8)
 
 grace.write_file('../../manuscript/figures/persistence_motif_profiles.eps')
 

### 2022-12-02 manuscript/reorder_main_text.tex

@@
     Coarse measures of a species' structural role, like in-degree and trophic level, are known to affect a consumer's vulnerability to bottom-up disturbance.
-    We are interested in whether a richer set of species roles based on participation in three-species motifs is also related to persistence, and how these motif-based roles relate to simpler measures.
+    We are interested in whether a richer set of species roles --based on participation in three-species motifs-- is also related to persistence and how these motif-based roles relate to simpler measures.
     We show that a consumer species' participation in different motifs is related to its probability of persistence. 
@@
     %\subsection*{2 Global and local network properties are known to affect stability}
-    How disturbances, such as population decline or extinction, lead to loss of other species via direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    How disturbances such as population decline or extinction lead to the loss of species via direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
     Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances.
@@
     Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
-    The indisputable presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
+    The indisputable presence of large, persistent empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
     Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2015}.
@@
 		Similarly, if all resources are extinct ($f = 1$), the extinction probability $P(\lnot C|f)$ will be equal to 1 and the consumer will go extinct.
-		If the fraction $f$ is not 1 nor 0, the non-linear sigmoid curve will determine the response of the consumer, which will be more severe if the fraction resources lost is high and less severe if the fraction is small.
+		If the fraction $f$ is neither 1 nor 0, the non-linear sigmoid curve will determine the response of the consumer, which will be more severe if the fraction resources lost is high and less severe if the fraction is small.
 This will in turn affect the extinction probability $P(\lnot C|f)$.
@@
     This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.
-    Connectance did, however, affect which motifs were most common within a network. 
+    Connectance did, however, affect which motifs were most common within a network and therefore cannot be entirely ignored. 
     Highly-connected networks tended to contain higher frequencies of the omnivory motif, which includes more links than the other three-species motifs considered here.
@@
     If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
-    Ongoing work using a dynamical model shows that time to extinction for species following disturbance is tightly coupled to the motifs a species participates in, and that adding top-down effects can lead omnivory to be negatively correlated with persistence \citep{Cirtwill2021_inprep}.
+    For example, one recent study using a dynamical model shows that time to extinction for species following disturbance is tightly coupled to the motifs a species participates in, and that adding top-down effects can lead omnivory to be negatively correlated with persistence \citep{Cirtwill2021_inprep}.
     If top-down effects are not crucial, the Bayesian network approach could be extended by explicitly incorporating information on species traits.

### 2022-12-06 manuscript/reorder_main_text.tex
formatted abstract for JAE

@@
 % \usepackage[numbers,sort]{natbib}
-\usepackage[numbers]{natbib} 
+\usepackage[sort]{natbib} 
 \usepackage{array}
@@
 \usepackage{xr}
+\usepackage{indentfirst}
 
@@
 \medskip
-$^\dagger$ Corresponding author\\
+$^\dagger$ Corresponding author email: alyssa.cirtwill@gmail.com\\
 }
@@
 \maketitle 
-\linenumbers
 \raggedright
@@
 
-\section*{Abstract}
+\section*{Running headline}
+Motif participation describes extinction risk
 
-    Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.
-    Coarse measures of a species' structural role, like in-degree and trophic level, are known to affect a consumer's vulnerability to bottom-up disturbance.
+\clearpage
+
+\section*{Summary}
+
+    \indent{1.
+Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.} \\
+    2. Coarse measures of a species' structural role, like in-degree and trophic level, are known to affect a consumer's vulnerability to bottom-up disturbance. \\
+    3.
 We are interested in whether a richer set of species roles --based on participation in three-species motifs-- is also related to persistence and how these motif-based roles relate to simpler measures.
-    We show that a consumer species' participation in different motifs is related to its probability of persistence. 
-    Similarly, the overall motif profile of a network is related to the average persistence of consumer species.
+\\
+    4. We show that a consumer species' participation in different motifs is related to its probability of persistence. \\ 
+    5. Similarly, the overall motif profile of a network is related to the average persistence of consumer species. \\
+    6.
 Notably, we found that networks containing a higher proportion of the `omnivory' motif tend to have lower mean persistence following disturbance, while more frequent participation in the omnivory motif was associated with higher species-level probabilities of persistence.
-    Additionally, we find that relationships between motif participation and persistence can synthesize the information provided by in-degree and trophic level.
-    Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities.
+\\
+    7. Additionally, we find that relationships between motif participation and persistence can synthesize the information provided by in-degree and trophic level. \\
+    8. Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities.
+
+\section*{Key-words}
+    food web, degree, trophic level, omnivory, competition
 
 \begin{spacing}{2.0}
+\linenumbers
 
-\clearpage
 \section*{Introduction} %New
@@
 
-\bibliographystyle{vancouver} 
+\bibliographystyle{jae} 
 \bibliography{anna_bib_new} % Abbreviate journal titles.

### 2023-03-31 manuscript/cover_letter.tex

@@
 
-\address{Anna {\AA}kesson \\ Link\"oping University \\ Division of Theoretical Biology, Dept.~IFM \\ SE-58183 Link\"oping, Sweden}
+\address{Alyssa Cirtwill \\ University of Helsinki \\ Spatial Foodweb Ecology Group \\ Helsinki, Finland}
 
-\date{October 21, 2021}
+\date{October XX, 2022}
 
@@
 
-Prof.~Barrett\\
-Editor-in-Chief\\
-Proceedings of the Royal Society B} % Name/title of the addressee
+Editorial Board\\
+Journal of Animal Ecology} % Name/title of the addressee
 
-\opening{Dear Prof.~Barrett,}
+\opening{Dear Editor,}
 
-We invite you to consider the enclosed ``Motif participation provides unique information about species risk of extinction'' by Anna {\AA}kesson et al., for publication in \emph{Proceedings of the Royal Society B} as an Article.
-Dr.
-Cirtwill was approached by the Proceedings B Preprint Editorial Team to consider publishing a thematically similar paper in this journal.
-That manuscript is currently in revisions elsewhere, but we believe that this manuscript also would be a good topical fit.
+We invite you to consider the enclosed manuscript ``Species motif participation provides unique information about species risk of extinction'' for publication in the \emph{Journal of Animal Ecology} as an Article.
 
@@
 However, the locations for secondary extinctions are difficult to identify as both direct and indirect interactions among species are important.
-As global network properties are too coarse-grained to include interactions, and local network properties disregard any indirect effects, we must consider other structural properties in order to fully understand secondary extinctions.
+As global network properties are too coarse-grained to include interactions and local network properties disregard any indirect effects, we must consider other structural properties in order to fully understand secondary extinctions.
 Meso-scale network properties, such as smaller sub-graphs within a large network (motifs), offer a potentially valuable middle-ground approach.
@@
 As we take an important step towards understanding of how secondary effects propagate in complex food webs, we believe our method and results will interest a wide audience of ecologists including both theoreticians and empiricists.
-We therefore think \emph{ Proceedings of the Royal Society B} would be an ideal platform for our research.
+We therefore think the \emph{Journal of Animal Ecology} would be an ideal platform for our research.
 
@@
 
-\hspace{4\parskip} Anna {\AA}kesson, Alyssa R. Cirtwill, Kate Wootton, and Anna Ekl\"of
+\hspace{4\parskip} Alyssa R. Cirtwill, Anna {\AA}kesson,  Kate Wootton, and Anna Ekl\"of
 
@@
 \centering
-\includegraphics[width=60mm]{manuscript/figures/LiU_primary_black.pdf}
+\includegraphics[width=60mm]{HY_logo_Eng.eps}
 

### 2023-03-31 manuscript/JAE_reply.tex
setting up reply, preserving a copy of initial submission to JAE

\documentclass[12pt]{article} 
\usepackage{amsmath} 
\usepackage[dvips]{graphicx}
\usepackage{multirow} 
\usepackage{geometry} 
\usepackage{pdflscape}
\usepackage[labelfont=bf]{caption} 
\usepackage{setspace}
\usepackage[running]{lineno} 
% \usepackage[numbers,sort]{natbib}
\usepackage[round]{natbib} 
\usepackage{array}
\usepackage[table]{xcolor}

\topmargin -1.5cm % 0.0cm 
\oddsidemargin 0.0cm % 0.2cm 
\textwidth 6.5in
\textheight 9.0in % 21cmhttps://www.overleaf.com/project/6048001a3ac22e3aff2ced4a
\footskip 1.0cm % 1.0cm

\newenvironment{refquote}{\bigskip \begin{it}}{\end{it}\medskip}

\usepackage{authblk}

\begin{document}

\section*{Reply to Editor}

    \begin{refquote}
        Decision by the Editor (Prof. Darren Evans):

        Thank you for submitting your manuscript to Journal of Animal Ecology.

        I have now received reviewers' reports and the Associate Editor's comments on your manuscript and looked at it myself. As you can see from their comment below, they liked aspects of your work but raised a number of issues around context and clarity. As a result, I cannot accept this current version of the manuscript but I am giving you the option to submit a revised version of the manuscript for consideration.

        Although we are giving you the option to complete a major revision on this manuscript, please note that acceptance and publication is not guaranteed. Your revised manuscript is due on 18-Jun-2023. If you foresee any problems with meeting this deadline, please contact the editorial office at: admin@journalofanimalecology.org

    \end{refquote}

    \textbf{R:}

\clearpage

\section*{Reply to Associate Editor}

    \begin{refquote}

        Associate Editor Comments for Authors:
        Dear Alyssa Cirtwill,

        Thank you for submitting the above manuscript to Journal of Animal Ecology. The manuscript has now been evaluated by two independent referees. Both of them agree that this study is interesting. However, they have raised some issues that should be addressed in a revision. Among these, lack of clarity in the multitude of concepts used throughout the manuscript has been highlighted by both referees. Also, certain results need more contextualisation. The writing structure, mainly in the introduction and discussion sections, are not easy to follow, and some figures are difficult to interpret (e.g. Figure 4). In sum, the manuscript would benefit from increasing clarity, context and making the study goal more focused. One of the referees also points to certain aspects of the statistical analysis that need to be clarified. These issues should be addressed in a revised manuscript.

        I hope that these comments and those of the reviewers assist the authors in preparing a revision.

        Best wishes,

        Daniel Montoya
        Associate Editor of Journal of Animal Ecology

    \end{refquote}

    \textbf{R:}

\clearpage

\section*{Reply to Reviewer \#1}

    \begin{refquote}

        In this paper the authors examine how a species' participation in different three-species motifs within a food web impacts their probability of persistence. They apply a Bayesian network approach to determine whether a species goes extinct following a disturbance, focusing on disturbance to the basal trophic levels.

        Overall I felt this was a very good paper. The introduction laid out the context in a clear and concise manner, and I thought the methods were well described, both in terms of what was done as well as potential limitations of the approach. I also liked the different links that were tested between motif profiles and both global and local network properties.

    \end{refquote}

    \subsection*{Comments}

        \begin{enumerate}

            \item One concern I had is in regards to the statistical analysis. Given the use of lmer, I assume that the error distribution was assumed to be Gaussian, though the response variable is constrained to be between 0 and 1. It may be more appropriate to use a binomial or quasi-binomial error distribution (using glmer).

            \textbf{R:}

            \item The only other comment I had was that Fig. 4 was a bit difficult to interpret. I was wondering if it might not be better to facet across motif and connectance, with each panel showing that motif's slope distribution at different disturbance levels. I think this would make it easier to compare the distributions for each motif, which is what the text of the results seemed to focus on.

            \textbf{R:}

        \end{enumerate}

\clearpage

\section*{Reply to Reviewer \#2}

    \begin{refquote}

        The theme of this study certainly captured my attention and interest, and the overall concepts explored in the article speaks to a broad audience of researchers. Although the topics introduced in the article are related, I found that the authors incorporated a few too many high-level ideas
        and did not unpack all of them, causing the goal of the study to be somewhat difficult to determine. For example, in the introduction, the authors begin their first paragraph by talking about disturbance in plant communities. Although disturbance is clearly relevant to consumer persistence, it would be clearer if the authors began by providing insight and definitions of what a motif is, why it is important, and how it relates to consumer persistence. With this jumping off point, the authors could mention that disturbance in plant communities influences consumers differentially depending on their structural configuration (i.e., motif). As it is in its current form, it feels like the authors are jumping from one idea to the next without clarifying the details of any one of them.

        I feel that this study includes a lot of work and analysis that is meaningful. For example, at line 195-196, the authors state “This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.”. To me, this sentence highlights how their results fit into an up-and-coming area of ecological research: monitoring the likelihood of consumer persistence in efficient ways. Nonetheless, I found the structure of the writing in the introduction and discussion difficult to follow and some of the key points from the results were not easy to interpret because the authors have so many different conceptual ideas incorporated throughout the article.

    \end{refquote}

    \textbf{R:}

    \subsection*{Specific comments}

        \begin{enumerate}

            \item Summary: I found that the ideas explored in the paper were somewhat unclear from the summary. Is the overall goal of the study to understand participation in a motif (what I understood from the abstract and title) or is it to understand how the presence of particular sub-webs (i.e., motifs) influences persistence? It seems like it is both from the results presented, but the authors highlight the former extensively and leave the latter as more of an after-thought. This may be addressed by simply taking out words like “additionally” from the start of the last few summary points.

            \textbf{R:}

            \item Introduction Line 13-16: The wording here makes it seem like species loss is both the stressor/disturbance and the response variable. This would be clearer if the authors clarify whether they are studying the effects of disturbance at the primary producer level on consumer persistence right away. It wouldalso be helpful to define exactly what definition of disturbance they are referring to in the introduction.

            \textbf{R:}

            \item Line 42-44: I am uncertain of what this sentence is trying to convey. In the sentence prior to this one, the authors suggest that trophic level and in-degree are too coarse of descriptors. Then, instead of providing insight into why or how more detailed metrics would be desirable, they mention that the consumer’s risk of extinction is dependent on its prey. It is unclear how prey-dependent extinction of consumers is related to the coarseness of the metrics used to analyze food webs. This sentence might be highlighting a key point in the article, but as it is currently written it is very difficult to understand what this sentence intends to describe.

            \textbf{R:}

            \item Line 48: This paragraph contains important information to understanding the concept of motifs; however, I found that the description of motif participation is vague and difficult to interpret in the context of this study. For example, the “second-step indirect interactions the species are affected by, thereby capturing important structures that are missed by global and local properties” does not provide any details that would help the reader understand why motif participation is a relevant metric. Specifically, what important structures are the authors referring to here, and how are they captured by the second-step indirect interactions? Further, is there a better way to describe what a second-step indirect interaction is? I am unsure what the authors mean by this.

            \textbf{R:}

            \item Line 87-88: Local and global food web properties has been mentioned more than once but it is unclear what these properties are specifically. Perhaps defining what this is referring to earlier on in the introduction would make it more clear to the reader.

            \textbf{R:}

            \item Line 92-95: The authors bring up many conceptual topics in their introduction, but perhaps the most relevant topic: network motif structures, has been almost entirely overlooked in terms of providing background information to the reader about which motifs are most common, how they influence energy transfer, and what encourages the establishment of the different motif structures across food webs. The authors mention that they will be investigating food chain, omnivory, apparent and direct competition motifs only briefly at the end of the introduction. From my perspective, the authors are missing key details about what is already known about how these different structural configurations of sub-webs can influence stability and consumer persistence. For example, networks with a high degree of omnivory may be more stable than simple food chains, but this depends on the strength of the omnivorous interactions among trophic guilds. It would be helpful to provide more insight into the gaps in knowledge surrounding the prevalence of these motifs in food webs more generally, and why this would help food web researchers understand consumer persistence and how it is related to resilience/stability since this is where I see the greatest amount of novelty from this article.

            \textbf{R:}

            \item Methods: Line 98: The authors have not specified what species/trophic guilds are being considered as secondary extinctions in this study. Is a primary extinction at the primary producer level? Is a secondary extinction referring to an extinction of a species that relies on a species that has faced disturbance-induced extinction? This should be clarified.

            \textbf{R:}

            \item Line 104-107: I am not convinced that this is always the case. Many dynamical models are simplified and allow researchers to use key parameters to model food webs and understand how major flows of energy transfer influence consumer persistence.

            \textbf{R:}

            \item Line 108: I am not well-versed in Bayesian statistics and will therefore refrain from evaluating the methodology used here. I am hopeful that the other reviewers will be able to provide feedback on this section of the article.

            \textbf{R:}

            \item Line 185: The authors should provide information about how they have determined what networks would be more biologically likely than others.

            \textbf{R:}

            \item Line 192: It is not obvious how the authors have decided to define a focal species in their study. It should be made clear what trophic guild or species they are specifically looking at in each of these networks as the wording now is quite vague.

            \textbf{R:}

            \item Line 194-195: It is unclear whether this is a general phenomenon or specific to this study. I am also uncertain of whether the vectors that the authors are referring to related to the number of species in the motifs or the number of motifs in a network.

            \textbf{R:}

            \item Results: Overall, I found the results difficult to interpret, mainly because of the number of results presented. I had issue understanding what figure 2 was showing, since the x-axis “proportion of role” was not alluded to described as an explanatory variable anywhere in the text. I wasn’t able to determine how the authors expected the probability of persistence to change across this gradient for each of the motifs, and how the results may have deviated from that expectation. This problem was mirrored in most of the figures presented.

            \textbf{R:}

            \item Network mean persistence and motif profiles: The ways in which persistence is related to motif profiles is interesting and valuable to explore; however, I am concerned with some of the interpretation of these results. Specifically, higher proportions of omnivory may lead to reduced average persistence if the average interaction strengths among organisms are strong; however, many food webs are composed of weakly interacting species where omnivory increases the likelihood of community-level persistence. Similarly, apparent competition is known to be destabilizing to food webs when energy is drastically diverted away from one species. The results presented here counter those examples, so it is important that the authors provide well-founded context for these findings.

            \textbf{R:}

            \item Discussion: I found the discussion section relatively well-written, and the authors pointed out how their analyses may be relevant in understanding how motif participation may be an important metric to consider when aiming to conserve species and prevent extinction.

            \textbf{R:}

            \item However, I am concerned that the authors do not provide any analysis for top-down effects on persistence. Although they are forthcoming about this (line 428-430), the interpretation of their analysis may be less relevant in the context of managing real food webs since predation does of course influence persistence.

            \textbf{R:}

            \item I feel that the discussion would be greatly improved if the authors focus in on one idea rather than three at once (i.e., disturbance strength; mean network persistence; motif participation). I am left trying to discern how motif participation affects persistence and have not been able to fully understand this, and I initially was under the impression that this was the ultimate goal of the study.

            \textbf{R:}

        \end{enumerate}

\clearpage

    \bibliographystyle{ecollett} 
    \bibliography{manual} % Do not abbreviate journal titles, papers with >

\end{document}

### 2023-03-31 manuscript/JAE_reply.tex
starting on the reply

@@
 \begin{document}
+
+% Due June 18.
 
@@
 
+Proposed intro structure: 
+1. What a motif is and why we care. 2. How do motifs relate to persistence? 3. Disturbance to plant communities influences consumers differently dependeing on network structure
+
+- but that introduces persistence before we talk about disturbance ... bit circular. 
+- main thrust is that we need to introduce fewer concepts, in more depth.
+
     \textbf{R:}
@@
 In the sentence prior to this one, the authors suggest that trophic level and in-degree are too coarse of descriptors.
-Then, instead of providing insight into why or how more detailed metrics would be desirable, they mention that the consumer’s risk of extinction is dependent on its prey.
+Then, instead of providing insight into why or how more detailed metrics would be desirable, they mention that the consumer's risk of extinction is dependent on its prey.
 It is unclear how prey-dependent extinction of consumers is related to the coarseness of the metrics used to analyze food webs.
@@
 
-            \textbf{R:}
+                \textbf{R:} A primary extinction is the extinction of a species that is directly disturbed (or removed, in the case of simulated species removal) while a secondary extinction is the extinction of any species not directly disturbed (as a result of the loss of its resources to primary or earlier secondary extinctions, as the Reviewer says).
+As we disturb primary producers and do not simulate top-down effects, only consumers can go secondarily extinction in our framework.
+We now define secondary extinction parenthetically in the introduction (line XX) and methods (line XX) and hope that this is clearer.
 
@@
 
-            \textbf{R:}
+                \textbf{R:} If the Reviewer is thinking of EcoPath with EcoSim, that type of model only works with quite a small food web with very well-understood interaction strengths and initial population sizes.
+We are not aware of any dynamical model that is fast over a large food web - certainly this was not the case when some of the authors conducted dynamical simulations of webs with 50-100 species.
+[[Anyone have a better answer?
+From the focus on major energy flows I think it's EcoSim they're thinking of...]]
 
@@
 
-            \textbf{R:}
+                \textbf{R:} [[Alyssa doesn't know how to respond to this one]]
 
@@
 
-            \textbf{R:}
+                \textbf{R:} [[Refers to: To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
+]] [[Based on many collective years of experience working with empirical food webs...
+is that acceptable to say?]]
 
@@
 
-            \textbf{R:}
+                \textbf{R:} A focal species is simply that for which the motif role is currently being calculated ...
+all species are focal species in turn.
+Note that as these networks are simulated the species (nodes) in them do not correspond to any particular focal species or trophic guild.
+The vagueness is intentional as there are no characteristics of, say, invertebrates vs.
+vertebrates or mammals vs.
+reptiles incorporated in the networks.
+[[Another really basic misunderstanding ]][[Seems to be this line: A species’ ``motif participation role'' is the frequency with which a focal species appears in each of the motifs present in a network (Stouffer et al., 2012).
+Is it clearer if we say "A species’ ``motif participation role'' is the frequency with which it appears in each of the motifs present in a network"?]]
 
@@
 
-            \textbf{R:}
+                \textbf{R:} We only considered three-species motifs in this study, as stated in lines XX, XX, and XX.
+The fact that only four \emph{unique} three-species motifs are possible in acyclic networks is a general feature of all acyclic networks; the other 9 possible three-species motifs contain cycles of two or three species (see~\citep{Stouffer2007} for a full list of these motifs).
+Note that each motif can appear within the network many times with a different set of three interacting species (e.g., a network might contain the three-species-chains `hawk-mouse-oat',  `hawk-mouse-grasshopper', and `mouse-grasshopper-oat') so that the \emph{total occurrences} of all three-species motifs in a network will be much greater than four.
+                [[I think it's this line: In our case, these are four-dimensional vectors since only four three-species motifs can appear in acyclic networks]]
+                [[How can we clarify this line?]]
 
             \item Results: Overall, I found the results difficult to interpret, mainly because of the number of results presented.
-I had issue understanding what figure 2 was showing, since the x-axis “proportion of role” was not alluded to described as an explanatory variable anywhere in the text.
-I wasn’t able to determine how the authors expected the probability of persistence to change across this gradient for each of the motifs, and how the results may have deviated from that expectation.
+I had issue understanding what figure 2 was showing, since the x-axis ``proportion of role'' was not alluded to described as an explanatory variable anywhere in the text.
+I wasn't able to determine how the authors expected the probability of persistence to change across this gradient for each of the motifs, and how the results may have deviated from that expectation.
 This problem was mirrored in most of the figures presented.
 
-            \textbf{R:}
+                \textbf{R:} Proportion of role is the proportion of the species' motif role made up by the particular motif in each panel. [[Ummm this is kind of dumb. Will a caption rewrite fix it?]]
 
@@
 
-            \textbf{R:}
+                \textbf{R:} We thank the Reviewer for their compliments.
 
@@
 
-            \textbf{R:}
+                \textbf{R:} Using the Bayesian network framework, we simply cannot include top-down effects on persistence.
+However, in a parallel study (started around the same time but by chance published earlier) some of the authors of the present study did simulate both top-down and bottom-up effects in a dynamical model.
+The dynamical model has its own caveats (in general, dynamical models are sensative to parameter choice and, due to the time and computational resources required, a full sensitivity analysis could not be performed).
+However, [[something comparing the results between papers - somewhat similar?]]
+
+                It may also reassure the Reviewer to know that in the initial studies introducing the Bayesian network framework~\citep{}, Bayesian networks captured XX\% of the extinctions identified by dynamical models run on the same set of networks.
+As we use the same modelling framework as in those studies, we can reasonably expect similar results.
+
+                Finally, consider the situation from an ecological perspective.
+We simulate disturbance to primary producers only (and appreciate that this may not have been clear in the previous revision).
+While predation certainly does influence persistence, the disturbance we introduce will first act to remove food sources from some herbivores.
+Some fraction of these herbivores will then go extinct, reducing resources for their predators.
+Some fraction of these predators will also go extinct due to the reduction in their food.
+It is only those predators which are affected by declining resources, do not go extinct, and are able to increase predation on their remaining prey which can cause top-down extinctions.
+While this obviously does occur in nature, logically this is a smaller number of extinctions than the set of herbivore and predator extinctions caused by bottom-up effects.
+This is also reflected in the similar results from Bayesian networks and dynammical models in the studies cited above - if top-down effects were responsible for the majority of secondary extinctions we would see quite different performance.
+
+                None of this is to discount the reality of top-down extinctions or their interest from a scientific standpoint.
+However, given that our interest is in bottom-up extinctions in particular and that we are up front about the absence of top-down extinctions in our model [[]]
+
+                [[Add a comment about how strong top-down effects could change our interpretation to discussion?]]
 
@@
                 \textbf{R:}
+                [[refocus on motif participation is definitely a good idea. Though there were interaction effects with disturbance strength, so we can't leave that out.]]
 

### 2023-04-03 code/plot_roles_vs_TL.py
fixed axis fig. 3

@@ -1,9 +1,9 @@
 import sys
 import os
-import math
-import random
-from decimal import *
-import numpy as np
+# import math
+# import random
+# from decimal import *
+# import numpy as np
 
 #Pygrace libraries
 from PyGrace.grace import Grace
@@ -91,7 +91,7 @@ def format_graph(graph,simple,roletype):
     graph.world.ymax=.600000001
     graph.yaxis.tick.major=.2
     graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
-    graph.yaxis.label.text='Proportion of motif role'
+    graph.yaxis.label.text='Proportion of motif in role'
   elif roletype=='Count':
     graph.world.ymax=600
     graph.yaxis.tick.major=200

### 2023-04-03 code/stat_analysis/persistence_vs_profiles.py
fixed legend in fig. 5

@@ -1,9 +1,9 @@
 import sys
 import os
-import math
-import random
-from decimal import *
-import numpy as np
+# import math
+# import random
+# from decimal import *
+# import numpy as np
 
 #Pygrace libraries
 from PyGrace.grace import Grace
@@ -97,7 +97,7 @@ def format_graph(graph,simple):
     graph.yaxis.tick.major=.2
     graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
     graph.yaxis.label.text='Mean persistence'
-    graph.xaxis.label.text='Proportion of network profile'
+    graph.xaxis.label.text='Proportion of motif in network profile'
 
 
   graph.xaxis.ticklabel.configure(format='decimal',prec=1,char_size=1)

### 2023-04-03 manuscript/reorder_main_text.tex

@@
 %\subsection*{Here we test...}
-    Here, we are interested in whether a consumer's extinction risk of going secondarily extinct depends on 1) the motifs in which the species participates, 2) the motif profile of the full food web and 3) the severity of the disturbance to the basal level.
+    Here, we are interested in whether a consumer's extinction risk after disturbance to the plant community (i.e., its risk of secondary extinction) depends on 1) the motifs in which the species participates, 2) the motif profile of the full food web and 3) the severity of the disturbance to the basal level.
     In addition, we want to understand how the information provided by motifs is related to other global and local food web properties in order to better synthesize information provided by different scales of network structure.
@@
     As Bayesian networks are computationally efficient (see Methods), we are able to simulate a wide range of disturbance intensities and test whether meso-scale network structure also interacts with severity of the disturbance.
-    We focus our analyse on the four network motifs that are over-represented in empirical networks and additionally have been shown to contribute to network stability: the three-species chain, omnivory, apparent competition, and direct competition \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
+    We focus our analysis on the four network motifs that are over-represented in empirical networks and additionally have been shown to contribute to network stability: the three-species chain, omnivory, apparent competition, and direct competition \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
 
@@
 
-        Traditionally, there are two main approaches for studying secondary extinctions.
+        Traditionally, there are two main approaches for studying secondary extinctions (i.e., extinctions of species that were not directly disturbed).
 First, there are topological models, based only on food web structure \citep{dunne2009cascading}.

### 2023-04-03 manuscript/JAE_reply.tex

@@
 
-                \textbf{R:} Proportion of role is the proportion of the species' motif role made up by the particular motif in each panel. [[Ummm this is kind of dumb. Will a caption rewrite fix it?]]
+                \textbf{R:} Proportion of role is the proportion of the species' motif role made up by the particular motif in each panel.
+This misunderstanding seems linked to a general difficulty in understanding precisely how we define motif roles.
+To help remedy this, we have now added motif role vectors to Figure 1 to make it explicit how counts of motifs are translated into vectors of normalised motif frequencies.
+These frequencies are equivalent to the proportion of a species' total motif participation made up by each motif.
+We also now use the perminology ``proportion of motif in role'' and ``proportion of motif in network profile'' in the axis labels of Figs.
+2, 3, and 5 for consistency.
 

### 2023-04-04 code/collect_empirical_data.R
git roundup and editing figure 6

@@ -1,0 +1,50 @@
+require(tidyverse) ## for efficient data manipulation and plotting
+
+web_nos=as.character(list.files(path="../data/empirical/global_verts/matrix/", full.names=FALSE, pattern="*.tsv"))
+
+# for (k in 1:length(web_nos)){
+#     web_no <- web_nos[k]
+#     load(file=paste0("Results_restricted/Linear/",web_no, ".RData"))
+#     final <- mutate(collect, web = web_no)
+#    # combines the result from all webs (of a particular S and C) in a single df (tab)
+#      if (k==1){
+#         tab <- final
+#     } else {
+#         tab <- bind_rows(tab, final)
+#     }                   
+# }
+# tab$per=unlist(tab$per)
+# write.table(tab,file='../data/all_empirical_linear.tsv')
+# save(tab, file="Results_restricted/Linear/empirical.RData")
+
+
+for (k in 1:length(web_nos)){
+    web_no <- strsplit(web_nos[k],'.tsv')[[1]][1]
+    load(file=paste0("Results_restricted/Nonlinear/global_verts/",web_no, ".RData"))
+    final <- mutate(collect, web = web_no)
+   # combines the result from all webs (of a particular S and C) in a single df (tab)
+     if (k==1){
+        tab <- final
+    } else {
+        tab <- bind_rows(tab, final)
+    }                   
+}
+tab$per=unlist(tab$per)
+write.table(tab,file='../data/global_verts_nonlinear.tsv')
+save(tab, file="Results_restricted/Nonlinear/global_verts.RData")
+
+for (k in 1:length(web_nos)){
+    web_no <- strsplit(web_nos[k],'.tsv')[[1]][1]
+    load(file=paste0("Results_restricted/Linear/global_verts/",web_no, ".RData"))
+    final <- mutate(collect, web = web_no)
+   # combines the result from all webs (of a particular S and C) in a single df (tab)
+     if (k==1){
+        tab <- final
+    } else {
+        tab <- bind_rows(tab, final)
+    }                   
+}
+tab$per=unlist(tab$per)
+write.table(tab,file='../data/global_verts_linear.tsv')
+save(tab, file="Results_restricted/Linear/global_verts.RData")
+

### 2023-04-04 code/convert_ecowebs.py

@@ -1,0 +1,75 @@
+import os
+import sys
+
+# igraph expects predators on rows/
+
+def read_matrix(infile):
+  species=set()
+  linkdict={}
+  f=open(infile,'r')
+  for line in f:
+    comsplit=line.split(',')
+    if 'WEB239' in infile:
+      print comsplit
+    if comsplit[-5:]==['','','','','\r\n']: # Ignore headers
+      pass
+    else:
+      prey=line.split('\t')[0]
+      links=line.split('\t')[1:]
+      if '0' in links or 0 in links:
+        species.add(prey)
+        for i in range(0,len(links)):
+          pred=preds[i].split('\n')[0]
+          if 'WEB239' in infile:
+            print pred
+          species.add(pred)
+          if float(links[i].split('\n')[0])>0:
+            try:
+              linkdict[pred].append(prey)
+            except KeyError:
+              linkdict[pred]=[prey]
+      else:
+        preds=links
+  f.close()
+
+  if 'WEB239' in infile:
+    print species
+
+  return species, linkdict
+
+def write_matrix(species,linkdict,net):
+
+  f=open('../data/empirical/global_verts/matrix/'+net.split('_mod')[0]+'.tsv','w')
+  f.write('\t'+'\t'.join(sorted(species))+'\n')
+  for pred in sorted(species):
+    f.write(pred)
+    if pred in linkdict:
+      for prey in sorted(species):
+        if prey in linkdict[pred]:
+          f.write('\t1')
+        else:
+          f.write('\t0')
+      f.write('\n')
+    else:
+      for spec in species:
+        f.write('\t0')
+      f.write('\n')      
+
+  f.close()
+
+
+
+def main():
+
+  # indir='../data/empirical/global_verts/original/'
+  indir='../data/empirical/global_inverts/original/'
+
+  nets=os.listdir(indir)
+  i=1
+  for net in [n for n in nets if '_mod.csv' in n]:
+    print net,i
+    species, linkdict=read_matrix(indir+net)
+    write_matrix(species, linkdict, net)
+    i+=1
+if __name__ == '__main__':
+  main()

### 2023-04-04 code/empirical_approx.R

@@ -1,0 +1,113 @@
+require(XRJulia) ## package allowing for Julia interfacing
+require(tidyverse) ## for efficient data manipulation and plotting
+require(igraph)
+library(dplyr)
+source("getTL.R") ## to get prey average TL
+
+juliaUsing("Random, Distributions")
+
+## Monte Carlo evaluation of Bayesian network, written in Julia
+bayes_ntw_MC_jul <- juliaEval("
+                              function bayesianNetworkMonteCarlo(A, pBaseline, a, b, reps)
+                              sA = sum(A, dims = 2)
+                              nRows = length(sA)
+                              lastBasal = maximum(findall(sA .== minimum(sA)))[1][1]
+                              marginals = zeros(nRows)
+                              for rep in 1:1:reps
+                              extant = ones(nRows)
+                              for i in 1:1:lastBasal
+                              extant[i] = rand() < pBaseline[i] ? 0 : 1
+                              end
+                              for i in (lastBasal + 1):1:size(A, 1)
+                              frac = 1 - (extant' * A[i,:])[1] / sum(A[i,:])
+                              pExt = pBaseline[i] + (1 - pBaseline[i]) * cdf(Beta(a, b), frac)
+                              extant[i] = rand() < pExt ? 0 : 1
+                              end
+                              marginals = marginals + extant
+                              end
+                              return(vec(marginals) / reps)
+                              end
+                              ")
+
+## turn the above Julia code to an R function
+bayes_ntw_MC <- JuliaFunction(bayes_ntw_MC_jul)
+
+webs=as.character(list.files(path="../data/empirical/global_verts/matrix/", full.names=FALSE, pattern="*.tsv"))
+disturbance <- seq(0,1,by=0.2)     # Simulated level of proportion of basal species disturbed, transforms in a later
+                                   # stage to a disturbance level of 0.1 to 0.5 in steps of 0.1
+for(style in c('Nonlinear','Linear')){
+    for (k in 1:length(webs)){
+        web_no <- strsplit(webs[k],'.tsv')[[1]][1]
+        print(web_no)
+        A <- read.table(paste("../data/empirical/global_verts/acyclic_matrix/",
+                              web_no,'.tsv',sep=''), sep=",", header=TRUE) # topo sorted, acyclic graph
+        S <- dim(A)[1]    # no of species in food web A
[…]

### 2023-04-04 code/fig_std.R

@@ -1,0 +1,165 @@
+require(ggpmisc)
+require(tidyverse)
+require(broom)
+require(mblm)
+require(viridis)
+require(lme4)
+
+#df <- read.table(file = 'empirical_3sp_roles_participation_nonlinear.tsv', sep = '\t', header = TRUE)
+df<-read.table(file = '../data/3sp_roles_participation.tsv', sep = '\t', header = TRUE)
+
+
+labs_bp <- c(
+    "0" = "0.1",
+    "0.2" = "0.18",
+    "0.4" = "0.26",
+    "0.6" = "0.34",
+    "0.8" = "0.42",
+    "1" = "0.5")
+
+prop.labs <- c("Apparent comp.", "Three-sp. chain", "Direct comp.", "Omnivory")
+names(prop.labs) <- c("frac_acomp", "frac_chain", "frac_dcomp", "frac_omni")
+
+## modification of df
+tab_t <- df %>% 
+    filter(in_Degree > 0, Size %in% c(50,100),
+           Connectance==0.02) %>%
+    filter(in_Degree > 0) %>% 
+    unite(NW, Network, Connectance, Size, remove=FALSE) %>%
+    select(NW, Species, Persistence, m12, m36, m38, m6, 
+           Connectance, Basal_p, Size) %>%
+    mutate(omni=m38, chain=m12, dcomp=m36, acomp=m6) %>%
+    gather("motif","counts", 4:7) %>%
+    group_by(NW, Species, Persistence, Basal_p, Connectance, Size,
+             omni, chain, dcomp, acomp) %>%
+    summarise(total=sum(counts)) %>% 
+    ungroup() %>%
+    mutate(frac_omni = omni/total, frac_chain=chain/total,
+           frac_dcomp=dcomp/total, frac_acomp=acomp/total) %>%
+    gather("norm_motif", "frac", 12:15) 
+
+
+# tab_temp%>%filter(NW=="kongsfjorden_0.024_260", Species=="sp10", Basal_p==1.0)
+#lmer_per_frac <- function(df) with(df,lmer(Persistence~frac+(1|Size:Connectance)+(1|NW)))
+lmer_per_frac <- function(df) with(df,lmer(Persistence~frac+(1|NW)))
+
+## extract stats (hopefully std...)
+extract_stats_std <- function(input) {
+    input %>% summary %>% coef %>%
+        as_tibble() %>%
+        select(2) %>%
+        rename(coefs=1) %>%
[…]

### 2023-04-04 code/motif_SC_lms.py

@@ -1,9 +1,6 @@
 import sys
 import os
 import math
-import random
-from decimal import *
-import numpy as np
 
 #Pygrace libraries
 from PyGrace.grace import Grace
@@ -55,38 +52,34 @@ def format_graph(graph,graphtype):
   return graph
 
 def total_dataset(nettype):
-  if nettype=='total':
-    I=-2.846e+03
-    s=4.662e+01
-    c=-2.122e+05
-    sc=4.716e+03
-  elif nettype=='chain':
-    I=0.2617064
-    s=0.0001775
-    c=-0.1566627
-    sc=-0.0032675
+  if nettype=='chain':
+    I=-1.038068
+    s=0.001026
+    c=-0.795159
+    sc=-0.019323
   elif nettype=='apparent':
-    I=5.409e-01
-    s=-5.058e-04
-    c=-1.001e+00
-    sc=2.590e-03
+    I=0.168403
+    s=-0.002028
+    c=-4.094017
+    sc=0.010249
   elif nettype=='direct':
-    I=1.926e-01
-    s=-3.563e-05
-    c=-5.556e-01
-    sc=5.917e-03
+    I=-1.4206660
+    s=-0.0003694
+    c=-3.9670780
+    sc=0.0423147
   elif nettype=='omnivory':
-    I=4.824e-03
-    s=3.639e-04
-    c=1.714e+00
-    sc=-5.239e-03
+    I=-2.989315
+    s=0.003146
+    c=12.874525
+    sc=-0.036079
 
   datadict={}
   for S in [50,60,70,80,90,100]:
     dataset=[]
     for C in [0.02,0.06,0.1,0.14,0.18]:
       y=I+S*s+C*c+S*C*sc
-      dataset.append((C,y))
+      logity=math.exp(y)/(1+math.exp(y))
+      dataset.append((C,logity))
     datadict[S]=dataset
 
   return datadict
@@ -129,26 +122,6 @@ def populate_graph(graph,nettype):
 #
 ###############################################################################################
 ###############################################################################################
-
-
-grace=MultiPanelGrace(colors=colors)
-# grace.add_label_scheme('dummy',['Start of flowering','Mass flowering','Inferred peak','End of flowering','Season length','F','G','H'])
-# grace.set_label_scheme('dummy')
-for lm in ['mot_total']:
-  graph=grace.add_graph(Panel)
-  graph=format_graph(graph,'total')
-  graph=populate_graph(graph,'total')
[…]

### 2023-04-04 code/stat_analysis/empirical_fig_1.R

@@ -1,0 +1,81 @@
+require(ggpmisc)
+require(tidyverse)
+require(broom)
+require(mblm)
+require(viridis)
+require(lme4)
+
+# to get correct labels in fig. legend
+labs_bp <- c(
+    "0" = "0.1",
+    "0.2" = "0.18",
+    "0.4" = "0.26",
+    "0.6" = "0.34",
+    "0.8" = "0.42",
+    "1" = "0.5")
+
+# to get nice motif labels
+prop.labs <- c("Apparent comp.", "Three-sp. chain", "Direct comp.", "Omnivory")
+names(prop.labs) <- c("frac_acomp", "frac_chain", "frac_dcomp", "frac_omni")
+
+# Using data frame "data" from empirical_statistics.R
+
+tab_t <- data %>%
+    filter(in_Degree > 0) %>%   # remove basal sp.
+    unite(NW, Network, Connectance, Size, remove=FALSE) %>%  # unite info about NW, C, S in new column NW
+    select(NW, Species, Persistence, m12, m36, m38, m6, 
+           Connectance, Basal_p, Size) %>%  # get rid of some columns not necessary
+    mutate(omni=m38, chain=m12, dcomp=m36,acomp=m6) %>%  # rename for my sake
+    gather("motif","counts",4:7) %>%   
+    group_by(NW, Species, Persistence, Basal_p, Connectance, Size,
+             omni, chain, dcomp, acomp) %>%
+    summarise(total=sum(counts)) %>% # create total of each motif (grouped by above)
+    ungroup() %>%
+    mutate(frac_omni = omni/total, frac_chain=chain/total,
+           frac_dcomp=dcomp/total, frac_acomp=acomp/total) %>%
+    gather("norm_motif", "frac", 12:15) # create two columns of motif and frac
+
+# lmer model function
+lmer_per_frac <- function(data) with(data,lmer(Persistence~frac+(1|Size:Connectance)+(1|NW)))
+
+# tidyverse extraction 
+extract_stats <- function(input) {
+    input %>% summary %>% coef %>%
+        as_tibble() %>%
+        select(1) %>%
+        rename(coefs=1) %>%
+        mutate(type=c("intercept", "slope")) %>%
+        return()
+}
+
+# dark nesting of data to increase efficiency, call of lmer function
+tab_prop <- tab_t %>% 
+    group_by(norm_motif, Basal_p) %>%
+    nest() %>%
+    ungroup() %>%
+    mutate(stats = map(data, lmer_per_frac)) %>%
[…]

### 2023-04-04 code/stat_analysis/statistical_models.R

@@ -57,7 +57,7 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
         consumers$net_omni[r]=motdata$prop_omni[which(motdata$netID==net)]
     }
 
-    total_lm=with(motdata,lm(mot_total~Size*Connectance))
+    total_lm=with(motdata,glm(mot_total~Size*Connectance,family='binomial'))
     summary(total_lm)
 
     # PERMANOVA and variability of motif profiles
@@ -86,10 +86,10 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     write.table(interdata2,file='proportion_variability_SC.tsv',sep='\t')
 
     # LMs - no need for network-level random effect since it's within-network motif profiles
-    pchainlm=with(motdata,lm(prop_chain~Size*Connectance))
-    pomnilm=with(motdata,lm(prop_omni~Size*Connectance))
-    papplm=with(motdata,lm(prop_apparent~Size*Connectance))
-    pdirlm=with(motdata,lm(prop_direct~Size*Connectance))
+    pchainlm=with(motdata,glm(prop_chain~Size*Connectance,family='binomial'))
+    pomnilm=with(motdata,glm(prop_omni~Size*Connectance,family='binomial'))
+    papplm=with(motdata,glm(prop_apparent~Size*Connectance,family='binomial'))
+    pdirlm=with(motdata,glm(prop_direct~Size*Connectance,family='binomial'))
 
 
 # Section 3: How does persistence vary with network motif profile?
@@ -110,10 +110,10 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     simdata$dist=as.numeric(as.character(simdata$dist))
 
     # Mean persistence ~ chain, omni, apparent, direct with disturbance
-    netchain_mean=with(simdata,lm(per~scale(prop_chain)*scale(dist)))
-    netomni_mean=with(simdata,lm(per~scale(prop_omni)*scale(dist)))
-    netapparent_mean=with(simdata,lm(per~scale(prop_apparent)*scale(dist)))
-    netdirect_mean=with(simdata,lm(per~scale(prop_direct)*scale(dist)))
+    netchain_mean=with(simdata,glm(per~scale(prop_chain)*scale(dist)g))
[…]

### 2023-04-04 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
         We are first interested in whether participating more often in a certain motif (e.g., direct competition) might increase a species' probability of persistence.
-        To test this, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks):
+        To test this, we fit four general linear mixed-effect models with binomial error distribution (GLMMs; one per motif included in the Bayesian networks):
 
@@
         The network-level random effect accounts for the fact that species in the same network are not independent.
-        We centered and scaled the predictors before fitting the LMMs.
-        All models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
+        We centered and scaled the predictors before fitting the GLMMs.
+        All models were fit using the R~\citep{R} function `glmer' from the package \emph{lmerTest}~\citep{lmerTest}.
         We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.
@@
 
-    We compared the outcomes from linear mixed-effect models (LMMs) relating persistence to motif roles across all networks (Fig.
-3 and Table 1, \emph{Main Text}) with similar linear regressions (LRs) made on each single network separately.
+    We compared the outcomes from linear mixed-effect models (GLMMs) relating persistence to motif roles across all networks (Fig.
+3 and Table 1, \emph{Main Text}) with similar general linear regressions (GLRs) made on each single network separately.
     The main trends are similar in both analyses, but different-sized networks showed slightly different results for some motifs when considered individually (Figure~\ref{fig:density_prop_S}).
@@
 3 and Table 1, \emph{Main Text}).
-    The same pattern holds overall for LRs fit within a network  (Figure~\ref{fig:density_prop_S}): the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance.
+    The same pattern holds overall for GLRs fit within a network  (Figure~\ref{fig:density_prop_S}): the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance.
     This trend is most pronounced in large networks. (bottom panel, Figure~\ref{fig:density_prop_S}). 
@@
 
-\section{Motif profiles and global properties}
+\section{Motif profiles and global properties} [[updated]]
     % No update needed
@@
     These differences in variability can cause false positives in the PERMANOVA test. Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).
+    A series of general linear regressions support the relationship between network motif profiles and global structural properties.
+    Specifically, the frequency of the omnivory motif increased in more-connected networks (Table~\ref{network_prop_lms}) while the frequencies of other motifs were not significantly related to network size or connectance.
 
@@
 
-    The frequencies of the four possible motifs were related to global network properties.
-
      \begin{table}[hb!]
         \centering
-        \caption{Coefficients for linear regressions relating the proportion of each motif in the total motif profile of a network to network size, connectance, and their interaction.
-Coefficients in \textbf{bold} are significant (\textless0.05).}
+        \caption{Coefficients for general linear regressions relating the proportion of each motif in the total motif profile of a network to network size, connectance, and their interaction.
+Coefficients in \textbf{bold} are significant (\textless0.05).
+The regression was fit using a binomial error distribution and logit link function.}
        \label{network_prop_lms}
@@
             \hline
-            Three-species Chain & \textbf{0.262} & \textbf{1.78$\times10^{-4}$} & \textbf{-0.157} & \textbf{-3.27$\times10^{-3}$} \\
-            Apparent Competition & \textbf{0.541} & \textbf{-5.06$\times10^{-4}$} & \textbf{-1.00} & \textbf{2.59$\times10^{-3}$} \\
-            Direct Competition & \textbf{0.193} & -3.56$\times10^{-5}$ & \textbf{-0.556} & \textbf{5.92$\times10^{-3}$} \\
-            Omnivory & 4.82$\times10^{-3}$ & \textbf{3.64$\times10^{-4}$} & \textbf{1.71} & \textbf{-5.24$\times10^{-3}$}\\
+            Three-species Chain & \textbf{-1.04} & 0.001 & -0.795 & -0.019 \\
+            Apparent Competition & 0.168 & -0.002 & -4.09 & 0.010 \\
+            Direct Competition & \textbf{-1.42} & -3.69$\times10^{-4}$ & -3.97 & 0.042 \\
+            Omnivory & \textbf{-2.99} & 0.003 & \textbf{12.9} & -0.036\\
             \hline

### 2023-04-04 manuscript/JAE_reply.tex
git roundup and editing figure 6
Updates from Overleaf

@@
 
-            \textbf{R:}
+            \textbf{R:} Yes, a binomial error distribution would be more appropriate.
+We have corrected the error and updated our results accordingly and thank the Reviewer for pointing this out.
+[[something about what changed, if anything]]
 
@@
 
-            \textbf{R:}
+            \textbf{R:} [[This one might be have to be Anna's task - I can't get the ggplot2 code to run.]]
 
@@
 
-                \textbf{R:}
+                \textbf{R:} Probably the Reviewer is confused... I am confused by the comment.
 

### 2023-04-04 code/stat_analysis/persistence_vs_profiles.py
updated Fig. 5, change imperceptible

@@ -1,9 +1,6 @@
 import sys
 import os
-# import math
-# import random
-# from decimal import *
-# import numpy as np
+import math
 
 #Pygrace libraries
 from PyGrace.grace import Grace
@@ -47,10 +44,10 @@ scales={ # center, scale
 }
 
 lms={ # Intercept, motif, disturbance, interaction
-  'omnivory':(0.5743430145,-0.0140077658,-0.1385163991,-0.0005586564),
-  'apparent':(0.574343015,0.011206293,-0.138516399,0.000589999),
-  'direct':(0.5743430145,0.0033850297,-0.1385163991,0.0001809503),
-  'chain':(5.743430e-01,5.801241e-03,-1.385164e-01,-3.067919e-05)  
+  'omnivory':(0.325913,-0.062579,-0.595877,0.003218),
+  'apparent':(0.325782,0.050071,-0.595667,-0.001998),
+  'direct':(0.3255830,0.0150797,-0.5953395,-0.0005337),
+  'chain':(0.325632,0.025985,-0.595410,-0.002531)  
 }
 
 motif_names={'omnivory':'Omnivory','apparent':'Apparent competition','direct':'Direct competition','chain':'Three-species chain'}
@@ -137,8 +134,9 @@ def populate_persgraph(graph,motif):
     for x in [minx,maxx]:
       scalx=(x-scales[motif][0])/scales[motif][1]
       y=lms[motif][0]+scalx*lms[motif][1]+scaldist*lms[motif][2]+scalx*scaldist*lms[motif][3]
+      logity=math.exp(y)/(1+math.exp(y))
 
-      dats.append((x,y))
+      dats.append((x,logity))
 
     data=graph.add_dataset(dats)
     data.symbol.shape=0

### 2023-04-04 manuscript/reorder_main_text.tex

@@
 
-            To test if participating more often in a certain motif affected a species' probability of persistence, we fit four linear mixed-effect models (LMMs; one per motif included in the Bayesian networks):
+            To test if participating more often in a certain motif affected a species' probability of persistence, we fit four general linear mixed-effect models with binomial error distribution (GLMMs; one per motif included in the Bayesian networks):
                 \begin{equation}
@@
             To identify correlations between global network structure (i.e., the whole network) and participation in particular motifs (a meso-scale description of species roles), we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
-            We fit a similar set of  LMMs relating motif participation to in-degree or trophic level (i.e., local measures of a species' role; Appendix \emph{S3.2}).
+            We fit a similar set of  GLMMs relating motif participation to in-degree or trophic level (i.e., local measures of a species' role; Appendix \emph{S3.2}).
             Taken together, these supplemental models demonstrate how motif participation relates to other measures of network structure.
@@
             Persistence is also known to vary with local-scale measures of network structure. 
-            To identify the relationships between local-scale properties and species' probabilities of persistence, we fit two final LMMs relating a consumer's persistence to disturbance, in-degree or STL, and the interaction between disturbance and the local-scale measure (Eq.
+            To identify the relationships between local-scale properties and species' probabilities of persistence, we fit two final GLMMs relating a consumer's persistence to disturbance, in-degree or STL, and the interaction between disturbance and the local-scale measure (Eq.
 S10-S11).
-            These LMMs also included random effects to account for the effects of global network structure and network ID on persistence.
+            These GLMMs also included random effects to account for the effects of global network structure and network ID on persistence.
 
@@
 
-        Although the PERMANOVA test relating overall motif profiles to network properties was not reliable (Fig. S5, \emph{Appendix S9}), 
-        connectance was significantly related to the frequencies of different motifs in a network when considering motifs individually (Fig.~\ref{motif_proportion_lms}).
-        Network size and the interaction between network size and connectance did not strongly affect the frequencies of most motifs.
-        Considering each motif separately, the proportions of apparent competition and three-species chains in a network's motif profile tended to decrease with increasing connectance while the proportion of omnivory increased with increasing connectance (Table S5, \emph{Appendix S9}).
+        Considering each motif separately, the proportion of the omnivory motif in a network's motif profile increased significantly with increasing connectance (Fig.~\ref{motif_proportion_lms}, Table S5, \emph{Appendix S9}).
+        The proportions of other motifs in a network's profile were not related to network size, connectance, or their interaction.
+        Consistent with the results of these GLMMs, a PERMANOVA test suggested that a network's motif profile as a whole was related to network size and connectance.
+        However, the assumption of equal variance across groups was violated for this test and the PERMANOVA results should therefore be considered indicative only.
 

### 2023-04-04 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
         \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.
-Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportions of the other three motifs.
-Interactions between motif profiles and disturbance were significant but small.
+Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportion of the apparent competition motifs.
+Interactions between motif profiles and disturbance were not significant.
 Here we show the relationships of probabilities of extinction of basal resources of $\pi=0.1$ (top set of lines) and $\pi=0.5$ (bottom set of lines).
@@
         \includegraphics[width=\textwidth]{figures/motif_proportion_lms.eps}
-        \caption{The proportions of the four motifs in a network's motif profile were related to network size, connectance, and their interaction.
-The effect of connectance was much stronger than the main effect or interaction with network size.
-The proportion of apparent competition in a network's profile decreased with increasing connectance, while the proportion of omnivory increased.
-In small networks, the proportion of direct competition also decreased with increasing connectance.}
+        \caption{The proportion of the omnivory in a network's motif role increased significantly as connectance increased, while the proportions of the other three motifs did not vary significantly with network size, connectance, or their interaction.}
         \label{motif_proportion_lms}

### 2023-04-04 manuscript/supplement_motifs.tex

@@
 
-\section{Persistence and motif profiles}
+\section{Persistence and motif profiles} [[updated]]
 
@@
         \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile made up by a focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
+Coefficients in \textbf{bold} are significant ($\alpha$\textless0.005).
 Predictors were centered and scaled before model fitting; coefficients refer to a unit increase in predictors.
-We also provide the marginal $R^2$ value for each regression.
+We also provide McFadden's pseudo-$R^2$ value for each regression.
 }
@@
         \footnotesize
-        \begin{tabular}{c|c c | c c | c c | c c | c }
-        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \\
-        Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ \\
+        \begin{tabular}{c|c c c c c | c }
+        Motif & Intercept & Proportion & Disturbance & Interaction &  pseudo-$R^2$ \\
             \hline
-            Omnivory & 0.574 & \textless0.001 & 5.81$\times10^{-3}$ & \textless0.001 & -0.0189 & \textless0.001 & -1.775$\times10^{-7}$ & 0.885 & 0.960 \\
-            Apparent competition & 0.574 & \textless0.001 & 7.22$\times10^{-4}$ & \textless0.001 & -0.0189 & \textless0.001 & 5.20$\times10^{-6}$ & 0.003 & 0.964 \\
-            Direct competition & 0.574 & \textless0.001 &  1.38$\times10^{-4}$ & \textless0.001 & -0.0189 & \textless0.001 & 1.01$\times10^{-6}$ & 0.400 & 0.964 \\
-            Three-species chain & 0.574 & \textless0.001 & 7.22$\times10^{-4}$ & \textless0.001 & -0.0189 & \textless0.001 & 5.20$\times10^{-6}$ & 0.003 & 0.964 \\ 
+            Three-species chain & \textbf{0.326} &  0.0260 & \textbf{-0.595} & -2.53$\times10^{-3}$ & 0.*958 \\
+            Apparent competition & \textbf{0.326} & \textbf{0.0501} & \textbf{-0.596} & -2.00$\times10^{-3}$ & 0.962 \\
+            Direct competition & \textbf{0.326} & 0.0151 & \textbf{-0.595} & -5.34$\times10^{-4}$ & 0.956 \\
+            Omnivory & \textbf{0.326} & \textbf{-0.0626} & \textbf{-0.596} & 3.22$\times10^{-3}$ & 0.966 \\
         \end{tabular}
@@
         \caption{Coefficients for general linear regressions relating the proportion of each motif in the total motif profile of a network to network size, connectance, and their interaction.
-Coefficients in \textbf{bold} are significant (\textless0.05).
+Coefficients in \textbf{bold} are significant ($\alpha$\textless0.05).
 The regression was fit using a binomial error distribution and logit link function.}

### 2023-04-06 code/stat_analysis/persistence_vs_SC_lm.py
updaed motifs vs. persistence fig with logistic regressions

@@ -1,9 +1,6 @@
 import sys
 import os
 import math
-import random
-from decimal import *
-import numpy as np
 
 #Pygrace libraries
 from PyGrace.grace import Grace
@@ -33,14 +30,14 @@ scales={
 }
 
 lms={
-  'Intercept': 0.5743430,
-  'Size':-0.0004818,
-  'Connectance':-1.293e-02,
-  'Disturbance':-0.0136768,
-  'Size:Connectance':-0.0009770,
-  'Size:Disturbance':-0.0034566,
-  'Connectance:Disturbance':-0.0002251,
-  'Size:Connectance:Disturbance':-0.0003969 }
+  'Intercept': 0.3259301,
+  'Size':-0.0007211,
+  'Connectance':-0.0611595,
+  'Disturbance':-0.5959284,
+  'Size:Connectance':-0.0044764,
+  'Size:Disturbance':-0.0162922,
+  'Connectance:Disturbance':0.0045203,
+  'Size:Connectance:Disturbance':-0.0013456 }
 
 Blms={'Intercept':5.777e-01,
       'Basal':2.841e-02,
@@ -135,8 +132,8 @@ def populate_persgraph(graph,simple):
         scalmain=(level-scales[simple][0])/scales[simple][1]
         base3=base2+scalmain*lms[simple]+scalbp*scalmain*lms[simple+':Disturbance']
         final=base3+scalmain*scalalt*lms['Size:Connectance']+scalmain*scalalt*scalbp*lms['Size:Connectance:Disturbance']
-
-        dats.append((level,final))
+        logy=math.exp(final)/(1+math.exp(final))
+        dats.append((level,logy))
 
       data=graph.add_dataset(dats)
       if al==altlevels[0]:
@@ -278,29 +275,29 @@ grace.hide_redundant_labels()
 
 grace.write_file('../../manuscript/figures/persistence_vs_SC_lm.eps')
 
+## I don't think we use this figure
+# grace=MultiPanelGrace(colors=colors)
 
-grace=MultiPanelGrace(colors=colors)
-
-for simple in ['Size','Connectance','Basal']:
-  for dist in [0.1, 0.5]:
-    graph2=grace.add_graph(Panel)
-    graph2=format_graph(graph2,simple)
-    if simple=='Size':
-      graph2.add_drawing_object(DrawText,text='Disturbance: '+str(dist),char_size=0.75,x=75,y=1.1,loctype='world',just=2)
[…]

### 2023-04-06 code/stat_analysis/persistence_vs_participation.py

@@ -1,0 +1,224 @@
+import sys
+import os
+import math
+
+#Pygrace libraries
+from PyGrace.grace import Grace
+from PyGrace.colors import ColorBrewerScheme
+from PyGrace.dataset import SYMBOLS
+from PyGrace.Extensions.panel import Panel,MultiPanelGrace
+from PyGrace.Extensions.distribution import CDFGraph, PDFGraph
+from PyGrace.Extensions.latex_string import LatexString, CONVERT
+from PyGrace.drawing_objects import DrawText, DrawLine,DrawBox
+from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
+from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
+from PyGrace.Styles.el import ElGraph, ElLogColorBar
+
+colors=ColorBrewerScheme('Greys',n=10)  # The blue is very beautiful but maybe harder to see.
+# Match Anna's ggplot colours
+colors.add_color(158,101,184,'Rpurple')
+colors.add_color(145,148,215,'Rblue')
+colors.add_color(42,120,142,'Rgreen1')
+colors.add_color(34,168,132,'Rgreen2')
+colors.add_color(122,209,81,'Rgreen3')
+colors.add_color(253,231,37,'Ryellow')
+colors.add_color(254,224,139,'Ryellow2')
+colors.add_color(253,174,97,'Rorange1')
+colors.add_color(244,109,67,'Rorange2')
+colors.add_color(213,62,79,'Rred1')
+colors.add_color(158,1,66,'Rred2')
+
+mot_ranges={
+  'omnivory':(0,0.5491713),
+  'apparent':(0,1),
+  'direct':(0,1),
+  'chain':(0,1),
+}
+
+scales={ # center, scale
+  'chain':(0.2489432,0.1272357),
+  'apparent':(0.4051135,0.1461447),
+  'direct':(0.1923225,0.1143021),
+  'omnivory':(0.1536209,0.1007478),
+  'Disturbance':(0.3,0.1366261)
+}
+
+lms={ # Intercept, motif, disturbance, interaction
+  'chain':(0.316878,-0.009774,-0.598433,-0.015098),  
+  'apparent':(0.316934,-0.002628,-0.598478,0.019535),
+  'direct':(0.316977,0.030266,-0.598517,-0.001665),
+  'omnivory':(0.316886,-0.018064,-0.598426,-0.007762)
+}
[…]

### 2023-04-06 code/stat_analysis/persistence_vs_participation.py
now with 95% CI

@@ -160,10 +160,10 @@ def populate_persgraph(graph,motif):
       # 5SE is a silly error bar, just using to test plotting.
       for x in range(minx,maxx):
         scalx=(float(x)/100-scales[motif][0])/scales[motif][1]
-        y=lms[motif][0]+2*offset[motif][0]
-        y+=scalx*(lms[motif][1]+2*offset[motif][1])
-        y+=scaldist*(lms[motif][2]+2*offset[motif][2])
-        y+=scalx*scaldist*(lms[motif][3]+2*offset[motif][3])
+        y=lms[motif][0]+1.96*offset[motif][0]
+        y+=scalx*(lms[motif][1]+1.96*offset[motif][1])
+        y+=scaldist*(lms[motif][2]+1.96*offset[motif][2])
+        y+=scalx*scaldist*(lms[motif][3]+1.96*offset[motif][3])
 
         logity=math.exp(y)/(1+math.exp(y))
 

### 2023-04-06 code/stat_analysis/statistical_models.R

@@ -57,7 +57,7 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
         consumers$net_omni[r]=motdata$prop_omni[which(motdata$netID==net)]
     }
 
-    total_lm=with(motdata,glm(mot_total~Size*Connectance,family='binomial'))
+    total_lm=with(motdata,lm(mot_total~Size*Connectance))
     summary(total_lm)
 
     # PERMANOVA and variability of motif profiles
@@ -86,6 +86,7 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     write.table(interdata2,file='proportion_variability_SC.tsv',sep='\t')
 
     # LMs - no need for network-level random effect since it's within-network motif profiles
+    # Network motif profiles vs. global properties
     pchainlm=with(motdata,glm(prop_chain~Size*Connectance,family='binomial'))
     pomnilm=with(motdata,glm(prop_omni~Size*Connectance,family='binomial'))
     papplm=with(motdata,glm(prop_apparent~Size*Connectance,family='binomial'))
@@ -110,11 +111,14 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     simdata$dist=as.numeric(as.character(simdata$dist))
 
     # Mean persistence ~ chain, omni, apparent, direct with disturbance
-    netchain_mean=with(simdata,glm(per~scale(prop_chain)*scale(dist)g))
-    netomni_mean=with(simdata,glm(per~scale(prop_omni)*scale(dist)g))
-    netapparent_mean=with(simdata,glm(per~scale(prop_apparent)*scale(dist)g))
-    netdirect_mean=with(simdata,glm(per~scale(prop_direct)*scale(dist)),family='binomial')
-
+    netchain_mean=with(simdata,glm(per~scale(prop_chain)*scale(dist),family='binomial'))
+    with(summary(netchain_mean),1-deviance/null.deviance) # McFadden's pseudo r-squared
+    netomni_mean=with(simdata,glm(per~scale(prop_omni)*scale(dist),family='binomial'))
+    with(summary(netomni_mean),1-deviance/null.deviance) # McFadden's pseudo r-squared    
+    netapparent_mean=with(simdata,glm(per~scale(prop_apparent)*scale(dist),family='binomial'))
[…]

### 2023-04-14 code/plot_roles_vs_TL.py
ok fixed the figure, except maybe direct comp vs degree

@@ -16,12 +16,12 @@ from PyGrace.Styles.el import ElGraph, ElLogColorBar
 
 colors=ColorBrewerScheme('Greys',n=10)  # The blue is very beautiful but maybe harder to see.
 # Match Anna's ggplot colours
-colors.add_color(94,79,162,'Rpurple')
-colors.add_color(50,136,189,'Rblue')
-colors.add_color(102,194,165,'Rgreen1')
-colors.add_color(171,221,164,'Rgreen2')
-colors.add_color(230,245,152,'Rgreen3')
-colors.add_color(255,255,191,'Ryellow1')
+colors.add_color(158,101,184,'Rpurple')
+colors.add_color(145,148,215,'Rblue')
+colors.add_color(42,120,142,'Rgreen1')
+colors.add_color(34,168,132,'Rgreen2')
+colors.add_color(122,209,81,'Rgreen3')
+colors.add_color(253,231,37,'Ryellow')
 colors.add_color(254,224,139,'Ryellow2')
 colors.add_color(253,174,97,'Rorange1')
 colors.add_color(244,109,67,'Rorange2')
@@ -113,16 +113,16 @@ def format_graph(graph,simple,roletype):
 def populate_graph(graph,minidict,simple,roletype):
   for motif in ['Apparent','Chain','Omnivory','Direct']:
     if motif=='Apparent':
-      j=13
+      j=12
       sty=2
     elif motif=='Chain':
       j=14
       sty=3
     elif motif=='Direct':
-      j=18
+      j=15
       sty=5
     else:
-      j=20
+      j=18
       sty=1
     dats=[]
     lower=[]
@@ -171,8 +171,8 @@ def populate_persgraph(graph,persdict,simple):
 
   print persdict.keys()
 
-  j=12
-  for basal_p in [0,0.18,0.26,0.34,0.42,0.5]:
+  j=13
+  for basal_p in [0.1,0.18,0.26,0.34,0.42,0.5]:
     p=(basal_p-0.3)/0.1366261
 
     dats=[]
@@ -191,9 +191,9 @@ def populate_persgraph(graph,persdict,simple):
     data.symbol.shape=0
     data.line.configure(linestyle=1,linewidth=3.5,color=j)
 
-    j+=1
     if simple=='TL':
       data.legend=str(basal_p)
+    j+=1
 
   if simple=='TL':
     graph.add_drawing_object(DrawText,text='Basal species',x=5.25,y=0.71,loctype='world',char_size=.75) 

### 2023-04-14 code/stat_analysis/statistical_models.R
re-ran everything

@@ -34,6 +34,8 @@ consumers$prop_direct=consumers$m36/consumers$total_motifs
 consumers$prop_omni=consumers$m38/consumers$total_motifs
 consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep=':')
 
+save.image('all_tests.Rdata')
+
 
 # Section 2: How do motif profiles vary with S and C?
     motdata=read.table('../../data/3sp_motif_profiles.tsv',header=TRUE)
@@ -92,6 +94,8 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     papplm=with(motdata,glm(prop_apparent~Size*Connectance,family='binomial'))
     pdirlm=with(motdata,glm(prop_direct~Size*Connectance,family='binomial'))
 
+save.image('all_tests.Rdata')
+
 
 # Section 3: How does persistence vary with network motif profile?
     # Needs a separate data frame for profiles repeated per disturbance, mean persistence per network/disturbance
@@ -135,11 +139,14 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     persist_anova=anova(persist_disp)
     # LM dispersion ~ mean persistence
     persist_lm=lm(persist_disp$distances~round(motdata$persistence,3))
+save.image('all_tests.Rdata')
+
 
 # Section 1: How does persistence vary with global network properties?
 
     # SCper=with(consumers,lm(Persistence~scale(Size)*scale(Connectance)*scale(Disturbance)))
     SCper_mean=with(simdata,glm(per~scale(Size)*scale(Connectance)*scale(dist),family='binomial'))
+save.image('all_tests.Rdata')
 
 # Section 4: How does persistence vary with proportions?
     # # lmer with random intercepts for each level of S:C
@@ -156,39 +163,63 @@ consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep
     # Random intercepts for S:C and network ID. Conclusions same as above.
     # Singular, removing network effect
[…]

### 2023-04-14 manuscript/reorder_main_text.tex
now with 95% CI
Updates from Overleaf

@@
         The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=645, $p$\textless0.001), although this result must be considered tentative due to non-homogeneous distribution of roles (\emph{Appendix S8}).
-        Average persistence was higher in networks with higher proportions of the three-species chain, direct competition, and apparent competition motifs (Table S4, \emph{Appendix S8}). 
-        Conversely, average persistence was lower in networks with higher proportions of the omnivory motif. 
+        Average persistence was higher in networks with higher proportions of the apparent competition motif and lower in networks with higher proportions of the omnivory motif (Table S4, \emph{Appendix S8}).
         These trends were consistent across levels of disturbance (Fig.~\ref{fig:motif_profile_persistence}, Table S4), \emph{Appendix S8}).

### 2023-04-14 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
         We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.
+        Due to model singularity, we removed the network-level random effect and re-fit the models.
+        After re-fitting, the models converged but the estimated variance explained by the random intercept was 0.
+        We therefore removed this random effect as well and re-fit the model as a general linear model including only the fixed effects described above.
 
@@
 
-\section{Persistence and motif participation}
+\section{Persistence and motif participation} [updated]
 
-    Persistence varied with motif participation.
+    Persistence was significantly related to the proportion of each motif in the species' role (Table~\ref{tab:proportion}).
+Persistence tended to increase with increasing proportions of the direct competition and, at high levels of disturbance, increasing frequencies of the apparent competition motif.
+Persistence tended to decrease with increasing proportions of the omnivory motif and, at high levels of disturbance, the three-species chain.
 
@@
         \centering
-        \caption{Coefficients ($\beta$) and $p$-values ($p$) for models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction, as well as random effects to account for differences in persistence across levels of network size, connectance, and network ID  (Equation~\ref{propreq}).
-Predictors were centered and scaled before model fitting; coefficients refer to a unit increase in each parameter.
-We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
+        \caption{Coefficients ($\beta$) for general linear models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction  (Equation~\ref{propreq}).
+Coefficients in \textbf{bold} are significant ($\alpha$\textless0.005).
+Models were fit using a binomial error distribution and logit link function.
+Predictors were centered and scaled before model fitting.
+We also provide McFadden's pseudo-$R^2$ for each model.}
         \label{tab:proportion}                \footnotesize
-        \begin{tabular}{c|c c | c c | c c | c c | c c }
-        & \multicolumn{2}{c|}{Intercept} & \multicolumn{2}{c|}{Proportion} & \multicolumn{2}{c|}{Disturbance} & \multicolumn{2}{c|}{Interaction} & \multicolumn{2}{c}{$R^2$} \\
-        Motif & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ & $R^2_M$ & $R^2_C$ \\
+        \begin{tabular}{c|c c c c | c }
+        Motif & Intercept & Proportion & Disturbance & Interaction & pseudo-$R^2$  \\
         \hline
-        3-sp Chain & 0.574 & \textless0.001 & -6.25$\times10^{-4}$ & \textless0.001 & -0.0190 & \textless0.001 & -5.86$\times10^{-5}$& \textless0.001 & 0.900 & 0.912 \\
-        App. Com. & 0.574 & \textless0.001 & -5.71$\times10^{-4}$ & \textless0.001 & -0.0190 & \textless0.001 & 8.19$\times10^{-5}$ & \textless0.001 & 0.900 & 0.911 \\
-        Dir. Com. & 0.574 & \textless0.001 & 6.95$\times10^{-4}$ & \textless0.001 & -0.0190 & \textless0.001 & 4.01$\times10^{-6}$ & \textless0.001 & 0.901 & 0.911 \\
-        Omnivory & 0.574 & \textless0.001 & 8.12$\times10^{-4}$ & \textless0.001 & -0.0190 & \textless0.001 & -2.73$\times10^{-5}$ & \textless0.001 & 0.894 & 0.912 \\
+        3-sp Chain & \textbf{2.90} & \textbf{-0.204} & \textbf{-5.42} & \textbf{0.138} & 0.895 \\
+        App. Com. & \textbf{0.317} & -2.63$\times10^{-3}$ & \textbf{-5.98478} & \textbf{1.95$\times10^{-2}$} & 0.895 \\
+        Dir. Com. & \textbf{0.317} & \textbf{3.03$\times10^{-2}$} & \textbf{-0.599} & -1.67$\times10^{-3}$ &  0.897 \\
+        Omnivory & \textbf{0.317} & \textbf{-1.81$\times10^{-2}$} & \textbf{-0.598} & \textbf{-7.76$\times10^{-3}$} & 0.895 \\
         \end{tabular}
@@
 
-\section{Participation and network properties}
+\section{Participation and network properties} [not updated yet]
 
@@
         \caption{Motif participation varies with size and connectance. Here we give the range and mean ($\pm$ standard deviation) proportion of each motif in species' motif participation vectors. 
-        We also show the effects ($\beta$) and $p$-values for network size, connectance, and their interaction in linear mixed-effect models predicting the proportion of each motif in a participation vector.
-        These models also included a random effect of network ID, as species within the same network are not independent.
+        We also show the coefficients ($\beta$) for network size, connectance, and their interaction in general linear mixed-effect models predicting the proportion of each motif in a participation vector.
+        These models also included a random effect of network ID, as species within the same network are not independent and were fit with a binomial error distribution and logit link function.
         Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
@@
     \end{figure}
+
+    Motif participation also varied with a species' degree and trophic level.
+    \begin{table}[hb!]
+        \centering
+        \caption{Motif participation varies with in-degree and trophic level.
+Here we show the coefficients for models relating the frequency of each motif in a species' participation vector to in-degree or trophic level (8 models in total).
+[[Random effect?]]
+        Coefficients in \textbf{bold} are significant ($\alpha$=0.05).
+        All models were fit with binomial error distribution and logit link function.
+        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
+        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
+        \label{tab:partic_vs_SC}   
+        \footnotesize
+        \begin{tabular}{c|c c }
+            Motif & in-degree & TL \\
+            \hline
+            3-sp Chain & \\
+            App. Comp & \\
+            Dir. Comp & \\
+            Omnivory & \\
+            \hline
+            \end{tabular}
+            \end{table}
+
 \clearpage
 
-\section{Persistence, degree, and trophic level}
+\section{Persistence, degree, and trophic level} [updated]
 
-    Persistence varies with in-degree, trophic level, and their interaction.
+    Persistence varies with in-degree, trophic level, and their interactions with disturbance.
 
@@
     \begin{table}[hb!]
-        \caption{Coeffcients ($\beta$) for linear mixed-effect models of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction, as well as a random effect of network ID.
+        \caption{Coeffcients ($\beta$) for general linear mixed-effect models of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction, as well as a random effect of global network properties (size and connectance).
+The model for STL was singular when including the random effect, so we removed this term and re-fit the model as a general linear model with fixed effects only.
+All models had binomial error distribution.
 Predictors were centered and scaled before model fitting.
@@
 The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.
-We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for each model.}
+We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for the in-degree model and the pseudo-$R^2$ for the STL model.}
         \label{tab:per_vs_TLdeg}
@@
             \hline
-            STL & -0.0101 & -0.0190 & 3.34$\times10^{4}$ & 0.922 & 0.95 \\
-            in-degree & -0.0236 & -0.0190 & -0.0155 & 0.903 & 0.936 \\
+            STL & -0.0977 & -0.600 & 0.0302 & 0.920 & NA \\
+            in-degree & -1.30$\times10^{-3}$ & -0.139 & -9.08$\times10^{-3}$ & 0.903 & 0.914 \\
         \end{tabular}
@@
 
-\section{Persistence and global network properties}
+\section{Persistence and global network properties} [updated]
 
-    Connectance and species richness had smaller effects on consumer persistence than disturbance (Figs.~\ref{fig:lm_CS}, \ref{fig:heatmap_CS}, Table~\ref{tab:per_vs_SC}).
+    Connectance had a smaller, but significant, effect on consumer persistence than disturbance (Figs.~\ref{fig:lm_CS}, \ref{fig:heatmap_CS}, Table~\ref{tab:per_vs_SC}).
+    Network size had no effect on consumer persistence in these simulated networks.
 
@@
         \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
-        \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources, while network size, connectance, and their interaction had smaller effects.
-\textbf{A)} Persistence decreased slightly with increasing network size at high levels of disturbance, increased slightly with increasing network size at low levels of disturbance, and did not vary noticeably with network size at moderate levels of disturbance.
-Networks with lower connectance (dotted line) had slightly higher persistence than networks with moderate (solid line) or high (dashed line) connectance.
-        \textbf{B)} Persistence decreased with increasing connectance for all levels of disturbance and network size. Network size (line types) did not strongly affect these relationships.}
+        \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.
+The effects of network size and interactions between predictors were not significant.
+}
         \label{fig:lm_CS}
@@
     \begin{table}[hb!]
-        \caption{Mean persistence of consumers within a network depends on network size (S) and connectance (C), the level of disturbance to basal resources (D), and interactions between them.
-Here we show the effect sizes ($\beta$) and $p$-values for the regression described above.
-All predictors were centered and scaled before fitting the model; coefficients refer to a unit increase in predictors.
+        \caption{Mean persistence of consumers within a network depends on connectance (C) and the level of disturbance to basal resources (D), but not network size (S) or interactions between global properties.
+Here we show the effect sizes ($\beta$) and $p$-values for the general linear model described above.
+The model was fit using a binomial error distribution and logic link function.
+All predictors were centered and scaled before fitting the model;.
 The level of disturbance corresponds to basal resources having a probability of extinction ranging between $0.1 - 0.5$ (see \emph{Methods}).
-The regression had $R^2$=
+McFadden's pseudo-$R^2$=0.966.
         }
@@
             \hline
-            Intercept & 0.574 & \textless0.001 \\
-            Size & -8.22$\times10^{-3}$  & 0.011 \\
-            Connectance & -7.74 & \textless0.001 \\
-            Disturbance & -0.0189 & \textless0.001 \\
-            S:C & -9.43$\times10^{-4}$ & \textless0.001 \\
-            S:D & -8.07$\times10^{-3}$ & \textless0.001 \\
-            C:D & -1.74$\times10^{-6}$ & 0.233 \\
-            S:C:D: & -5.24$\times10^{-5}$ & 0.036 \\
+            Intercept & 0.326 & \textless0.001 \\
+            Size & -7.21$\times10^{-4}$  & 0.964 \\
+            Connectance & -6.12$\times10^{-2}$& \textless0.001 \\
+            Disturbance & -0.596 & \textless0.001 \\
+            S:C & -4.48$\times10^{-3}$ & 0.777 \\
+            S:D & -1.63$\times10^{-2}$ & 0.316 \\
+            C:D & -4.52$\times10^{-3}$ & 0.781 \\
+            S:C:D: & -1.35$\times10^{-3}$ & 0.934 \\
         \end{tabular}

### 2023-04-14 manuscript/JAE_reply.tex

@@
 We have corrected the error and updated our results accordingly and thank the Reviewer for pointing this out.
-[[something about what changed, if anything]]
+Our results were not strongly affected by this change in statistical methodology, except that the relationship between network size and average consumer persistence was no longer significant.
 
@@
 This would be clearer if the authors clarify whether they are studying the effects of disturbance at the primary producer level on consumer persistence right away.
-It wouldalso be helpful to define exactly what definition of disturbance they are referring to in the introduction.
+It would also be helpful to define exactly what definition of disturbance they are referring to in the introduction.
 

### 2023-04-14 code/stat_analysis/statistical_models.R
i was missing a model. fixed now.

@@ -293,31 +293,44 @@ save.image('all_tests.Rdata')
     # apparent_deg_prop_net=with(consumers[which(consumers$Disturbance==0.1),],glmer(prop_apparent~in_Degree+(1|netID),family='binomial'))
     apparent_deg_prop=with(consumers[which(consumers$Disturbance==0.1),],glm(prop_apparent~in_Degree,family='binomial'))
 
+    direct_deg_prop_randnet=with(consumers[which(consumers$Disturbance==0.1),],glmer(prop_direct~in_Degree+(1|Global)+(1|netID),family='binomial'))
+    direct_deg_prop_global=with(consumers[which(consumers$Disturbance==0.1),],glmer(prop_direct~in_Degree+(1|Global),family='binomial'))
+    direct_deg_prop_net=with(consumers[which(consumers$Disturbance==0.1),],glmer(prop_direct~in_Degree+(1|netID),family='binomial'))
+    direct_deg_prop=with(consumers[which(consumers$Disturbance==0.1),],glm(prop_direct~in_Degree,family='binomial'))
+
+
 
     # TL, prop
-    chain_TL_prop_randnet=with(consumers[which(consumers$Disturbance==0.1),],glmer(prop_chain~STL+(1|Global)+(1|netID),family='binomial'))
-    chain_TL_prop_global=with(consumers[which(consumers$Disturbance==0.1),],glmer(prop_chain~STL+(1|Global),family='binomial'))
-    chain_TL_prop_net=with(consumers[which(consumers$Disturbance==0.1),],glmer(prop_chain~STL+(1|netID),family='binomial'))
+    # chain_TL_prop_randnet=with(consumers[which(consumers$Disturbance==0.1),],glmer(prop_chain~STL+(1|Global)+(1|netID),family='binomial'))
+    # chain_TL_prop_global=with(consumers[which(consumers$Disturbance==0.1),],glmer(prop_chain~STL+(1|Global),family='binomial'))
+    # chain_TL_prop_net=with(consumers[which(consumers$Disturbance==0.1),],glmer(prop_chain~STL+(1|netID),family='binomial'))
     chain_TL_prop=with(consumers[which(consumers$Disturbance==0.1),],glm(prop_chain~STL,family='binomial'))
 
[…]

### 2023-04-14 manuscript/reorder_main_text.tex

@@
         Species' motif participation also varied with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
-        Species with higher in-degree tended to have higher frequencies of participation in the omnivory motif. Proportions of the other three motifs tended to decline with increasing degree.
-        Species with higher trophic levels tended to have higher proportions of apparent competition and three-species chains. 
+        Species with higher in-degree tended to have higher frequencies of participation in the omnivory motif and lower proportions of the other three motifs.
+        Species with higher trophic levels tended to have higher proportions of apparent competition and three-species chains and lower proportions of omnivory and direct competition. 
         A species' persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix S6}). 
@@
         \centering
-        \includegraphics[width=0.85\textwidth]{figures/persistence_vs_motifs.eps}
+        \includegraphics[width=0.85\textwidth]{figures/persistence_motif_participation.eps}
         \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).

### 2023-04-17 code/stat_analysis/participation_vs_SC.py
updating fig. S2|

@@ -1,9 +1,6 @@
 import sys
 import os
 import math
-import random
-from decimal import *
-import numpy as np
 
 #Pygrace libraries
 from PyGrace.grace import Grace
@@ -19,8 +16,8 @@ from PyGrace.Styles.el import ElGraph, ElLogColorBar
 
 colors=ColorBrewerScheme('Greys',n=10)  # The blue is very beautiful but maybe harder to see.
 # Match Anna's ggplot colours
-colors.add_color(68,1,84,'Rpurple')
-colors.add_color(65,68,135,'Rblue')
+colors.add_color(158,101,184,'Rpurple')
+colors.add_color(145,148,215,'Rblue')
 colors.add_color(42,120,142,'Rgreen1')
 colors.add_color(34,168,132,'Rgreen2')
 colors.add_color(122,209,81,'Rgreen3')
@@ -87,26 +84,38 @@ def format_graph(graph,simple):
 
 def populate_persgraph(graph,xprop,val,netprops):
   print xprop, val
-  j=12
   for motif in netprops:
+    if motif=='Apparent':
+      j=12
+      sty=2
+    elif motif=='Chain':
+      j=14
+      sty=3
+    elif motif=='Direct':
+      j=15
+      sty=5
+    else:
+      j=18
+      sty=1
     betas=netprops[motif]
     dats=[]
     if xprop=='S':
       for x in [50,60,70,80,90,100]:
         y=betas[0]+x*betas[1]+val*betas[2]+x*val*betas[3]
-        dats.append((x,y))
+        logity=math.exp(y)/(1+math.exp(y))
+        dats.append((x,logity))
     elif xprop=='C':
       for x in [0.02,0.06,0.1,0.14,0.18]:
         y=betas[0]+x*betas[2]+val*betas[1]+x*val*betas[3]      
-        dats.append((x,y))
+        logity=math.exp(y)/(1+math.exp(y))
+        dats.append((x,logity))
 
     data=graph.add_dataset(dats)
     data.symbol.shape=0
-    data.line.configure(linestyle=1,linewdith=1.5,color=j)
+    data.line.configure(linestyle=sty,linewdith=1.5,color=j)
 
-    if xprop=='C' and val==50:
+    if xprop=='C' and val==70:
       data.legend=motif_names[motif]
-    j+=3
 
[…]

### 2023-04-17 code/stat_analysis/participation_vs_SC.py
made legend nicer and colours consistent w main text

@@ -75,7 +75,7 @@ def format_graph(graph,simple):
     graph.xaxis.ticklabel.configure(format='decimal',prec=0,char_size=.75)
 
   graph.world.ymin=0
-  graph.world.ymax=.6
+  graph.world.ymax=.75
   graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
   graph.yaxis.label.configure(text='Mean persistence',char_size=1,just=2,place='normal')
   graph.yaxis.tick.configure(major=.2,onoff='on',minor_ticks=0,major_size=.4,minor_size=.5,place='both',major_linewidth=.5,minor_linewidth=1)
@@ -117,7 +117,7 @@ def populate_persgraph(graph,xprop,val,netprops):
     if xprop=='C' and val==70:
       data.legend=motif_names[motif]
 
-  graph.legend.configure(box_linestyle=0,box_fill_pattern=0,char_size=.75,loc=(0.21,0.5),loctype='world')
+  graph.legend.configure(box_linestyle=0,box_fill_pattern=0,char_size=.75,loc=(0.1,0.74),loctype='world')
   return graph
 
 ###############################################################################################
@@ -137,7 +137,7 @@ grace.add_label_scheme('dummy',['C=0.1','S=70'])
 grace.set_label_scheme('dummy')
 
 
-for xprop in ['S','C']:
+for xprop in ['C','S']:
   if xprop=='S':
     val=0.1
   elif xprop=='C':

### 2023-04-17 manuscript/reorder_main_text.tex

@@
 
-    %\subsection*{1 Why are bottom-up effects important for understanding stability?}
-     The plant community is the foundation upon which the myriad species in food webs depend for their survival.
-     Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-     As such, the structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
-     For example, the loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
+    A species' position within a food web (a network of who eats whom) is known to affect its ability to persist following the extinction of  another species in the network (i.e., secondary extinction)~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{binzer2011susceptibility, Eklof2006}.
+    Species with low in-degree are more vulnerable to extinction because they have few alternative resources if a prey species is lost, while those with high trophic levels are vulnerable to the loss of any species along the path from basal resource(s) to the focal species.
+    Based on these trends, it has also been proposed that a species' participation in \emph{meso-scale structures} (which take into account the species' direct and indirect interactions) are also related to its persistence.
+
+    One way to describe a species' participation in such meso-scale structures is using its participation in the set of \emph{motifs} of a given size~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
+    Motifs --unique patterns of $n$ interacting species-- correspond to different sets of direct and indirect interactions and provide a holistic description of a species' place in a network~\citep{Stouffer2007,Stouffer2012}.
+    Like degree and trophic level, a species' \emph{motif participation} (the vector of frequencies with which it appears in each motif) can easily be calculated once the structure of the network is known.
+    Unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the focal species' interaction partners.
+    Two species might both have three prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
+    Importantly, since specialists are generally more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than a predator of generalists.
+
+    While a relationship between motif participation and persistence is broadly intuitive, so far there has been little research demonstrating which motifs are associated with greater probabilities of persistence.
+    Here we focus on persistence after the loss of basal resources (i.e., bottom-up disturbances).
+    The plant community is the foundation upon which the myriad species in food webs depend for their survival.Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
+    The loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc. \citep{scherber2010bottom}.
+    Species' motif participation may indicate \emph{which} consumers are more or less likely to go extinct following such a disturbance.
+
+    One potential reason for the difficulty in assessing the relationship between species' motif roles and persistence are the large computational resources required to simulate population dynamics.
+    Our solution is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks~\citep{Eklof2013,Haussler2020}.
+    Although Bayesian networks currently only simulate bottom-up disturbances (our focal interest), they nevertheless capture the vast majority of secondary extinctions obtained from dynamical models including top-down and bottom-up effects~\citep{Eklof2013}.
+    The greater computational efficiency and similar [[results?
+performance?
+extinctions?]] of Bayesian networks allow us to simulate the effects of a gradient of disturbance strength and thereby explore whether participation in some motifs is associated with greater probability of extinction under strong or weak disturbances.
+
+    Although there are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, we focus on just four: the three-species chain, apparent competition, direct competition, and omnivory.
+    These four motifs are by far the most common in empirical networks~\citep{Stouffer2007, Borrelli2015a, giling2019plant}.
+    They are also the only motifs which do not include cycles (where species A and B eat each other or where A eats B eats C eats A).
+    As modelling bottom-up extinctions in a Bayesian framework requires the breaking of cycles, these are the only motifs that will appear in the resulting \emph{acyclic} networks.
+    Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited.
+
+    Moreover, these four motifs have clear ecological interpretations, which are lacking for the cyclical motifs.
+    The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
+    The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers (e.g., by depletion of the shared resource). 
+    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey (e.g., by boosting predator populations) in a top-down context, but in a bottom-up context only includes direct interactions.
+    Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator.
+
+%     %\subsection*{1 Why are bottom-up effects important for understanding stability?}
+%      The plant community is the foundation upon which the myriad species in food webs depend for their survival.
+%      Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
+%      As such, the structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
+%      For example, the loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
 \citep{scherber2010bottom}.
 
-    %\subsection*{2 Global and local network properties are known to affect stability}
-    How disturbances such as population decline or extinction lead to the loss of species via direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
-    Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances.
+%     %\subsection*{2 Global and local network properties are known to affect stability}
+%     How disturbances such as population decline or extinction lead to the loss of species via direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+%     Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances.
 Two well-known, influential global properties are network size, the number of species present in the network, and network connectance, the number of realized feeding interactions.
-    Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
-    The indisputable presence of large, persistent empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
-    Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2015}.
+%     Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
+%     The indisputable presence of large, persistent empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
+%     Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2015}.
 However, a well-connected network also includes more pathways through which a disturbance can spread~\citep{Dunne2004, Vieira2015}.
 Depending on the strength and arrangements of interactions, this means that species may be impacted through several paths, amplifying the original disturbance and causing more secondary extinctions.
-    The net effect of connectance on extinction therefore likely depends on finer-scale (species-level) network structure, as well as the type and severity of disturbance.
+%     The net effect of connectance on extinction therefore likely depends on finer-scale (species-level) network structure, as well as the type and severity of disturbance.
 
-    Species-level properties known to affect species' responses to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
-    These can also be referred to as local properties of the network.
-    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and disturbances to plant populations cause particularly many secondary extinctions \citep{Curtsdotter2011}.
-    However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
-    A more detailed perspective is desirable since, as well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
-    For example, a generalist consuming many specialist (and therefore high-risk) prey may be at greater risk of extinction than a specialist which consumes a highly generalist resource that is less likely to go extinct (Fig.~\ref{fig:concept}).
+%     Species-level properties known to affect species' responses to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
+%     These can also be referred to as local properties of the network.
+%     Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and disturbances to plant populations cause particularly many secondary extinctions \citep{Curtsdotter2011}.
+%     However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
+%     A more detailed perspective is desirable since, as well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
+%     For example, a generalist consuming many specialist (and therefore high-risk) prey may be at greater risk of extinction than a specialist which consumes a highly generalist resource that is less likely to go extinct (Fig.~\ref{fig:concept}).
 
-    %\subsection*{3 Motif participation may similarly affect species' particular extinction risk}
-    It is possible to capture information about how a consumer and its prey fit together within a community by defining participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
-    Such motif profiles of species are easily calculated once we know the network structure and require no more information than is used to calculate degree or trophic level.
-    These `motif participation' roles capture information about species' direct as well as indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
-    We refer to these motifs as meso-scale structures to emphasise their role as structures between global (network-level) and local (species-level) properties.  
-    A species' motif participation profile shows the second-step indirect interactions the species are affected by, thereby capturing important structures that are missed by global and local properties.
-    Therefore, by allowing researchers to distinguish between, for example, consumers of specialists and generalists, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
+%     %\subsection*{3 Motif participation may similarly affect species' particular extinction risk}
+%     It is possible to capture information about how a consumer and its prey fit together within a community by defining participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
+%     Such motif profiles of species are easily calculated once we know the network structure and require no more information than is used to calculate degree or trophic level.
+%     These `motif participation' roles capture information about species' direct as well as indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
+%     We refer to these motifs as meso-scale structures to emphasise their role as structures between global (network-level) and local (species-level) properties.  
+%     A species' motif participation profile shows the second-step indirect interactions the species are affected by, thereby capturing important structures that are missed by global and local properties.
+%     Therefore, by allowing researchers to distinguish between, for example, consumers of specialists and generalists, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
 
-    While the potential relationships between motif participation and species-level extinction risk have not yet been tested to our knowledge, it has been shown that the frequencies of different motifs within a whole network are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
-    These relationships between network-level motif frequencies (`motif profiles') and overall community stability mean that a network's motif profile could affect the average extinction risk for consumers within the network.
-    Moreover, if network motif profiles are related to average extinction risk, it is worth investigating whether a species' participation in different motifs also affects its extinction risk.
+%     While the potential relationships between motif participation and species-level extinction risk have not yet been tested to our knowledge, it has been shown that the frequencies of different motifs within a whole network are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
+%     These relationships between network-level motif frequencies (`motif profiles') and overall community stability mean that a network's motif profile could affect the average extinction risk for consumers within the network.
+%     Moreover, if network motif profiles are related to average extinction risk, it is worth investigating whether a species' participation in different motifs also affects its extinction risk.
 
-% Severity of disturbance
-    As mentioned above, the severity of a disturbance may interact with network structure to affect secondary extinctions.
-    A mild disturbance (e.g., leading to a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
-    Empirical studies have shown that disturbances to plant species can affect some consumers more strongly than others \citep{byrnes2011climate}, that the strength of disturbance affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}, and that the resilience and return time of the ecological community can be strongly influenced by disturbance severity \citep{rydgren2004disturbance}.
-    The severity of disturbance therefore clearly has an effect on how the ecological community will respond.
+% % Severity of disturbance
+%     As mentioned above, the severity of a disturbance may interact with network structure to affect secondary extinctions.
+%     A mild disturbance (e.g., leading to a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
+%     Empirical studies have shown that disturbances to plant species can affect some consumers more strongly than others \citep{byrnes2011climate}, that the strength of disturbance affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}, and that the resilience and return time of the ecological community can be strongly influenced by disturbance severity \citep{rydgren2004disturbance}.
+%     The severity of disturbance therefore clearly has an effect on how the ecological community will respond.
 
 %\subsection*{Here we test...}
-    Here, we are interested in whether a consumer's extinction risk after disturbance to the plant community (i.e., its risk of secondary extinction) depends on 1) the motifs in which the species participates, 2) the motif profile of the full food web and 3) the severity of the disturbance to the basal level.
-    In addition, we want to understand how the information provided by motifs is related to other global and local food web properties in order to better synthesize information provided by different scales of network structure.
-    We use synthetic networks and model secondary extinctions using Bayesian networks \citep{Eklof2013, Haussler2020}. 
-    As Bayesian networks are computationally efficient (see Methods), we are able to simulate a wide range of disturbance intensities and test whether meso-scale network structure also interacts with severity of the disturbance.
-    We focus our analysis on the four network motifs that are over-represented in empirical networks and additionally have been shown to contribute to network stability: the three-species chain, omnivory, apparent competition, and direct competition \citep{Stouffer2007, Borrelli2015a, giling2019plant}.
+    Here, we are interested in whether a consumer's extinction risk after disturbance to the plant community (i.e., its risk of secondary extinction) depends on the species' motif participation and how this relationship varies with the severity of disturbance.
+    To provide context for this result, we also consider how extinction risk varies with the overall structure of the network.
+    Specifically, we consider the frequency of motifs across the whole network, network size, and connectance (the proportion of possible links which actually occur).
+    Finally, we relate species' motif participation to overall network structure and to their in-degrees and trophic levels in order to clarify what unique information motif participation can provide.
 
@@
 
-       The relationship between frequency of participation in a motif and a species' probability of persistence depended on the level of disturbance. 
-       When disturbance ($\pi$) was low, species' persistence increased with increasing proportions of the omnivory and direct competition motifs. 
-       Conversely, persistence decreased with an increasing proportion of the apparent competition motif. 
-       Persistence was not significantly related to the proportion of the three-species chain motif (Figure~\ref{fig:prop_lmer_all}; upper lines in each panel).
+        A species' probability of persistence was significantly related to its motif participation (Table S1, \emph{Appendix S4}).
+        The probability of persistence tended to increase with increasing frequency of the direct competition motif and decrease with increasing frequency of the omnivory motif, across all levels of disturbance.
+        The relationships between participation in the apparent competition and three-species chain motifs, however, varied with the strength of disturbance (Fig.~\ref{fig:prop_lmer_all}).
+        Higher frequency of participation in the apparent competition motif was associated with lower propbability of persistence at low levels of disturbance and higher probability of persistence at high levels of disturbance.
+        Participation in the three-species chain motif showed the opposite trends.
 
-        At high levels of disturbance, a species' persistence decreased as the proportion of the three-species chain in its motif profile increased.
-        Persistence increased as the proportion of any other motif increased (Figure~\ref{fig:prop_lmer_all}; lower lines in each panel). 
-        This means that increased disturbance strengthened the effects of the three-species chain motif and changed the direction of the effect of the apparent competition motif  (Figure~\ref{fig:prop_lmer_all}, Table S1, \emph{Appendix S4}).
-        The relationship between omnivory and persistence was still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.
+        Considering each network separately ....
 
@@
     % Species-level results summary
-    A species' frequent participation in the direct competition and omnivory motifs was associated with a higher probability of persistence while participation in the three-species chain motif was associated with lower persistence, regardless of the strength of disturbance.
-    The strength of disturbance strongly affected the relationship between persistence and participation in the apparent competition motif.
-    These trends can be partially explained by underlying relationships between motif participation and local-scale network properties.
+    As with degree and trophic level, a consumer's probability of persistence varied with its participation in different three-species motifs.
+    Specifically, increasing participation in the direct competition motif was always associated with higher probability of persistence while increasing participation in the omnivory motif was always associated with lower probability of persistence.
+    The relationships between persistence and apparent competition and the three-species chain depended on the strength of the disturbance, with apparent competition associated with greater probability of persistence when basal species were very likely to go extinct and the three-species chain associated with greater probability of persistence at lower levels of disturbance.
+
+    These trends can be partially explained by underlying relationships between motif participation and other measures of a species' place in the network.
     For example, species with higher trophic levels tended to participate less often in the direct competition motif.
@@
     Accordingly, high participation in the direct competition motif, by indicating a low trophic level, is associated with high probability of persistence.
+    Similarly, participation in apparent competition decreases with increasing in-degree, and probability of persistence shows the opposite trends with respect to in-degree and participation in apparent competition.
 
-    % Degree and TL are both important for omnivory
-    The relationship between participation in the omnivory motif and persistence connects motif participation with both in-degree and trophic level (STL).
-    Species participation in the omnivory motif was associated with greater persistence, especially at low levels of disturbance.
-    This is consistent with a strong negative correlation between omnivory and trophic level (leading to the overall positive effect of omnivory) and a strong positive correlation with in-degree (leading to the reduced benefit of omnivory at high levels of disturbance).
-    Motifs can thus act as a tool for synthesizing the information provided by other measures of a species' place in its community and for understanding the interactions among these measures.
+    The effects of other motifs cannot, however, be attributed so straightforwardly to simpler measures of a species' place in the network.
+    Like apparent competition, participation in the three-species chain motif decreases with increasing in-degree, yet persistence varied with the frequency of participation in the three-species chain in the same way as in-degree.
+    While participation in the three-species chain motif was positively correlated with trophic level, this cannot explain the relationship between trophic level and probability of persistence as trophic level was negatively correlated with persistence across all levels of disturbance.
+    Similarly, higher frequencies of participation in the omnivory motif were associated with lower probabilities of persistence, yet the frequency of the omnivory motif increased with in-degree and decreased with trophic level.
+    Both high in-degree and low trophic level are associated with higher probability of persistence, except at high levels of disturbance, which would naively suggest that higher proportions of omnivory should also be associated with greater persistence.
+    Considering a species' in-degree, trophic level, \emph{and} its motif role may therefore provide extra information about its extinction risk.
 
-    % Degree and TL don't explain everything
-    The relationships between persistence and participation in the apparent competition and three-species chain motifs likewise demonstrate the potential for motif participation to synthesize different aspects of network structure.
-    In these cases, in-degree and trophic level are not sufficient to explain the trends we observe in the motifs.
-    For apparent competition, there were strong and opposing correlations with in-degree and trophic level but the relationship between apparent competition and persistence only matches the trends for in-degree.
-    This suggests that some other factor is over-riding the relationship between apparent competition and trophic level.
+    % % Degree and TL are both important for omnivory
+    % The relationship between participation in the omnivory motif and persistence connects motif participation with both in-degree and trophic level (STL).
+    % Species participation in the omnivory motif was associated with greater persistence, especially at low levels of disturbance.
+    % This is consistent with a strong negative correlation between omnivory and trophic level (leading to the overall positive effect of omnivory) and a strong positive correlation with in-degree (leading to the reduced benefit of omnivory at high levels of disturbance).
+    % Motifs can thus act as a tool for synthesizing the information provided by other measures of a species' place in its community and for understanding the interactions among these measures.
 
-    Similarly, participation in the three-species chain motif was significantly correlated with both in-degree and trophic level, but the relationship between motif participation and persistence did not reflect the disturbance-dependent relationship between in-degree and persistence.
-    This suggests that the relationship between persistence and participation in the chain motif synthesizes the effects of in-degree, trophic level, and some additional effect which counterbalances the effect of in-degree.
-    This additional effect may come from indirect interactions, which are described by motifs but not in-degree~\citep{Cirtwill2018FoodWebs}.
+    % % Degree and TL don't explain everything
+    % The relationships between persistence and participation in the apparent competition and three-species chain motifs likewise demonstrate the potential for motif participation to synthesize different aspects of network structure.
+    % In these cases, in-degree and trophic level are not sufficient to explain the trends we observe in the motifs.
+    % For apparent competition, there were strong and opposing correlations with in-degree and trophic level but the relationship between apparent competition and persistence only matches the trends for in-degree.
+    % This suggests that some other factor is over-riding the relationship between apparent competition and trophic level.
+
+    % Similarly, participation in the three-species chain motif was significantly correlated with both in-degree and trophic level, but the relationship between motif participation and persistence did not reflect the disturbance-dependent relationship between in-degree and persistence.
+    % This suggests that the relationship between persistence and participation in the chain motif synthesizes the effects of in-degree, trophic level, and some additional effect which counterbalances the effect of in-degree.
+    % This additional effect may come from indirect interactions, which are described by motifs but not in-degree~\citep{Cirtwill2018FoodWebs}.
 
     % Moving to network context
+    Relationships between a species' probability of persistence and its' position within the network should be considered within the context of overall network structure.
+    In our networks, the average probability of persistence of all consumers increased with increasing proportions of the apparent competition, three-species chain, and direct competition motifs and decreased with increasing proportions of the omnivory motif.
+    The omnivory motif was also generally less common that the other motifs in these networks, which were selected to be stable before disturbance.
+
     As well as clarifying the relationships between network structure and the effect of bottom-up disturbance on particular species, motifs were also related to the mean probability of persistence across all species in a network.

### 2023-04-17 manuscript/supplement_motifs.tex

@@
         \includegraphics[height=.4\textheight]{figures/participation_vs_SC.eps}
-        \caption{The proportion of each motif in a consumer's participation vector varied with network size, connectance, and their interaction.
+        \caption{The proportion of each motif in a consumer's participation vector varied sigificantly with connectance but not with network size or the interaction between size and connectance.
+Because the interaction was not significant, we show relationships for intermediate values of size and connectance when plotting against the other variable.
 These relationships were steepest for apparent competition and omnivory.}
@@
 Here we show the coefficients for models relating the frequency of each motif in a species' participation vector to in-degree or trophic level (8 models in total).
-[[Random effect?]]
-        Coefficients in \textbf{bold} are significant ($\alpha$=0.05).
+All relationships were significant ($\alpha$=0.05).
         All models were fit with binomial error distribution and logit link function.
@@
             \hline
-            3-sp Chain & \\
-            App. Comp & \\
-            Dir. Comp & \\
-            Omnivory & \\
+                3-sp Chain & \textbf{-6.99$\times10^{-3}$} & \textbf{9.84$\times10^{-2}$} \\
+            Omnivory & \textbf{2.29$\times10^{-2}$} & \textbf{-2.27$\times10^{-1}$}\\
+            App. Comp & \textbf{-8.76$\times10^{-3}$} & \textbf{2.40$\times10^{-1}$}\\
+            Dir. Comp & \textbf{-1.43$\times10^{-3}$} & \textbf{-2.52$\times10^{-1}$}\\
             \hline
@@
 
-    \begin{figure}[hb!]
-     \centering
-     \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP0.eps}
-     \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
+    % \begin{figure}[hb!]
+    %  \centering
+    %  \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP0.eps}
+    %  \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
 The panels show final persistence for consumers with different shortest trophic level, for increasing connectance (y-axis) and network size (x-axis).
 Persistence decreases with darker color in the heat map.}
-     \label{fig:heatmap_stl_BP0}
-    \end{figure}
+    %  \label{fig:heatmap_stl_BP0}
+    % \end{figure}
 
-    \begin{figure}[hb!]
-     \centering
-     \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP1.eps}
-     \caption{Basal resources experience a disturbance increasing their probability of extinction to $\pi = 0.5$.
+    % \begin{figure}[hb!]
+    %  \centering
+    %  \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP1.eps}
+    %  \caption{Basal resources experience a disturbance increasing their probability of extinction to $\pi = 0.5$.
 All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
@@
 Persistence decreases with darker color in the heat map.}
-     \label{fig:heatmap_stl_BP1}
-    \end{figure}
+    %  \label{fig:heatmap_stl_BP1}
+    % \end{figure}
 
@@
 
-\section{Persistence and motif profiles} [[updated]]
+\section{Persistence and whole-network motif profiles} [[updated]]
 
     % Updated and expanded to include permanova tests
-    Persistence varied with the frequencies of different motifs in a network profile ($F_{1,2998}$=645, $p$\textless0.001 for a PERMANOVA of dissimilarity in motif profiles against mean persistence).
+    Mean persistence of all consumers varied with the frequencies of different motifs in a network profile ($F_{1,2998}$=645, $p$\textless0.001 for a PERMANOVA of dissimilarity in motif profiles against mean persistence).
     However, motif profiles were not evenly distributed across mean persistences ($F_{134,2865}$=1.77, $p$\textless0.001 for an ANOVA of dissimilarity in motif profiles against mean persistence, rounded to 3 decimal places).
@@
 
-    Connectance had a smaller, but significant, effect on consumer persistence than disturbance (Figs.~\ref{fig:lm_CS}, \ref{fig:heatmap_CS}, Table~\ref{tab:per_vs_SC}).
+    Connectance had a smaller, but significant, effect on consumer persistence than disturbance (Figs.~\ref{fig:lm_CS}, Table~\ref{tab:per_vs_SC}).
     Network size had no effect on consumer persistence in these simulated networks.
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
+        \includegraphics[width=0.9\textwidth]{figures/persistence_vs_SC_lm.eps}
         \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.
@@
 
-    \begin{figure}[hb!]
-        \centering
-      \includegraphics[width=\textwidth]{figures/heatmap_CS_BPcompare.eps}
-       \caption{Heat maps of average consumer persistence in networks with varying connectance (y-axis) and network size (x-axis) at two levels of disturbance to basal resources.
+    % \begin{figure}[hb!]
+    %     \centering
+    %   \includegraphics[width=\textwidth]{figures/heatmap_CS_BPcompare.eps}
+    %    \caption{Heat maps of average consumer persistence in networks with varying connectance (y-axis) and network size (x-axis) at two levels of disturbance to basal resources.
 Persistence decreases with darker color in the heat map.
@@
 When basal species are highly disturbed, their baseline extinction risk rises to $\pi = 0.5$ while the baseline risk of consumers is not changed.}
-       \label{fig:heatmap_CS}
-    \end{figure}
+    %    \label{fig:heatmap_CS}
+    % \end{figure}
 
@@
     \begin{table}[hb!]
-        \caption{Mean persistence of consumers within a network depends on connectance (C) and the level of disturbance to basal resources (D), but not network size (S) or interactions between global properties.
-Here we show the effect sizes ($\beta$) and $p$-values for the general linear model described above.
-The model was fit using a binomial error distribution and logic link function.
-All predictors were centered and scaled before fitting the model;.
+        \caption{Mean persistence of consumers within a network depends on connectance (C) and the level of disturbance to basal resources (D), but not network size (S) or interactions between properties.
+We show effect sizes ($\beta$) and $p$-values for each coefficient in a general linear model (fit using a binomial error distribution and logic link function).
+All predictors were centered and scaled before fitting the model.
 The level of disturbance corresponds to basal resources having a probability of extinction ranging between $0.1 - 0.5$ (see \emph{Methods}).
@@
             \hline
-            Intercept & 0.326 & \textless0.001 \\
+            Intercept & \textbf{0.326} & \textless0.001 \\
             Size & -7.21$\times10^{-4}$  & 0.964 \\
-            Connectance & -6.12$\times10^{-2}$& \textless0.001 \\
-            Disturbance & -0.596 & \textless0.001 \\
+            Connectance & \textbf{-6.12$\times10^{-2}$}& \textless0.001 \\
+            Disturbance & \textbf{-0.596} & \textless0.001 \\
             S:C & -4.48$\times10^{-3}$ & 0.777 \\
@@
             C:D & -4.52$\times10^{-3}$ & 0.781 \\
-            S:C:D: & -1.35$\times10^{-3}$ & 0.934 \\
+            S:C:D & -1.35$\times10^{-3}$ & 0.934 \\
         \end{tabular}

### 2023-04-18 code/stat_analysis/per_network_regressions.R
per-network regressions

@@ -1,0 +1,61 @@
+#### TEST MOTIFS AND PERSISTENCE 
+require(ggpmisc)
+require(tidyverse)
+
+# Libraries to fit the non-plotted lms
+library(lme4)
+library(lmerTest)
+# For permanovas and betadisper
+library(vegan)
+# For R2 GLMM
+library(MuMIn)
+
+
+data<-read.table(file = '../../data/3sp_roles_participation.tsv', sep = '\t', header = TRUE)
+data$Disturbance<-0.1+0.4*data$Basal_p
+# ONLY FIT MODELS ON CONSUMERS!!!
+consumers<-data[which(data$in_Degree>0),]
+# Check that all predictors are numeric in case of R shenanigans
+# All should be numeric except for species ID
+for(c in c((1:4),(6:25))){
+    if(!is.numeric(consumers[,c])){
+        consumers[,c]=as.numeric(as.character(consumers[,c]))
+        print(paste0(colnames(consumers)[c],' made numeric'))
+    }
+}
+
+consumers$total_motifs=consumers$m12+consumers$m6+consumers$m36+consumers$m38
+consumers$prop_chain=consumers$m12/consumers$total_motifs
+consumers$prop_apparent=consumers$m6/consumers$total_motifs
+consumers$prop_direct=consumers$m36/consumers$total_motifs
+consumers$prop_omni=consumers$m38/consumers$total_motifs
+consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep=':')
+
+per_network_results=matrix(nrow=0,ncol=6)
+colnames(per_network_results)=c("Network","S","C","Disturbance","Motif","Motif_slope")
+
+for(n in levels(as.factor(consumers$Network))){
+    print(n)
+    netdat=consumers[which(consumers$Network==n),]
+    S=netdat$Size[1]
+    C=netdat$Connectance[1]
+    for (d in c(0,0.2,0.4,0.6,0.8,1)){
+        Ddat=netdat[which(netdat$Basal_p==d),]
+        omnimod=with(Ddat,glm(Persistence~prop_omni,family='binomial'))
+        chainmod=with(Ddat,glm(Persistence~prop_chain,family='binomial'))
+        ACmod=with(Ddat,glm(Persistence~prop_apparent,family='binomial'))
+        DCmod=with(Ddat,glm(Persistence~prop_direct,family='binomial'))
+        per_network_results=rbind(per_network_results,
[…]

### 2023-04-18 manuscript/JAE_reply.tex
Updates from Overleaf

@@
                 \textbf{R:}
+            The sentence the reviewer refer to is "How disturbances such as population decline or extinction lead to the loss of species via  direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research (Curtsdotter et al., 2011a; Dunne & Williams, 2009; Ekl ̈of & Ebenman, 2006 Santos et al., 2021) " Here
 
@@
 
-                \textbf{R:} Probably the Reviewer is confused... I am confused by the comment.
+                \textbf{R:} Can the Reviewer provide a reference for their claim that apparent competition is destabilising?
+Is this a very general trend, or something that only happens in rare cases?
+Note that we do not claim that apparent competition is \emph{never} destabilising, only that more frequent participation in the apparent competition motif is associated with greater persistence on average.
+Moreover, we cannot think of a reason why a bottom-up disturbance (i.e., loss of some resources) would be destabilising in an apparent competition motif - surely it is more likely that a predator with two prey will persist after losing one than than two predator sharing one prey will persist if that prey is lost.
+We can see how it is possible for strong \emph{top-down} effects to destabilise an apparent competition motif (e.g., if a predator suddenly exerts much more pressure on one of the prey than the other), but this paper is concerned with bottom-up effects only.
+Given the earlier work showing that Bayesian networks capture the majority of secondary extinctions that occur in topological models that include both top-down and bottom-up effects, we can conclude that top-down effects are much less common as a cause of extinction than bottom-up effects.
+Of course they can still be equally interesting and may be strong in particular focal systems, but as our simulations do not reflect a specific community we have no reason to expect top-down effects to have a stronger role than in the networks used to validate the Bayesian network approach.
 

### 2023-04-18 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 
-\section{Participation and network properties} [not updated yet]
+\section{Participation and network properties} [updated]
 
@@
         \centering
-        \caption{Motif participation varies with size and connectance. Here we give the range and mean ($\pm$ standard deviation) proportion of each motif in species' motif participation vectors. 
-        We also show the coefficients ($\beta$) for network size, connectance, and their interaction in general linear mixed-effect models predicting the proportion of each motif in a participation vector.
-        These models also included a random effect of network ID, as species within the same network are not independent and were fit with a binomial error distribution and logit link function.
+        \caption{Motif participation varies with size and connectance. Here we give the range and mean proportion of each motif in species' motif participation vectors. 
+        We also show the coefficients ($\beta$) for network size (S), connectance (C), and their interaction in general linear models predicting the proportion of each motif in a participation vector.
+        Coefficients in \textbf{bold} are significant ($\alpha$=0.05).
+        We also show the coefficients for similar models relating the proportion of each motif to in-degree or trophic level.
+        All models were fit with a binomial error distribution and logit link function.
         Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
@@
         \footnotesize
-        \begin{tabular}{c|c c c | c c | c c | c c}
-             &  &  &  & \multicolumn{2}{c|}{Size} & \multicolumn{2}{c|}{Connectance} & \multicolumn{2}{c}{Interaction} \\
-            Motif & Min & Max & Mean (SD) & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
+        \begin{tabular}{c|c c c | c c c | c | c}
+            Motif & Min & Max & Mean & S & C & Int. & in-Deg & TL \\
             \hline
-            3-sp Chain & 0 & 1 & 0.249 (0.127) & 2.83$\times10^{-4}$ & \textless0.001 & -0.0982 & 0.053 & -3.66$\times10^{-3}$ & \textless0.001 \\
-            App. Comp & 0.250 & 0.765 & 0.422 (0.0645) & -5.85$\times10^{-4}$ & \textless0.001 & -0.840 & \textless0.001 & 2.69$\times10^{-3}$ & 0.003 \\
-            Dir. Comp & 0.0501 & 0.359 & 0.179 (0.0407) & -6.99$\times10^{-5}$ & 0.450 & -0.636 & \textless0.001 & 6.46$\times10^{-3}$ & \textless0.001 \\
-            Omnivory & 0.006 & 0.346 & 0.164 (0.0790) & 3.70$\times10^{-4}$ & \textless0.001 & 1.57 & \textless0.001 & -5.48$\times10^{-3}$ & \textless0.001\\   
+            3-sp Chain & 0 & 1 & 0.249 & \textbf{1.20$\times10^{-3}$} &  -0.652 & \textbf{-1.77$\times10^{-2}$} & \textbf{-6.99$\times10^{-3}$} & \textbf{9.84$\times10^{-2}$} \\
+            App. Comp & 0.250 & 0.765 & 0.422 & \textbf{-1.82$\times10^{-3}$} & \textbf{-3.18} & 6.92$\times10^{-3}$ & \textbf{-8.76$\times10^{-3}$} & \textbf{2.40$\times10^{-1}$}\\
+            Dir. Comp & 0.0501 & 0.359 & 0.179 & -1.18$\times10^{-3}$ & \textbf{-4.50} & \textbf{4.79$\times10^{-2}$} & \textbf{-1.43$\times10^{-3}$} & \textbf{-2.52$\times10^{-1}$} \\
+            Omnivory & 0.006 & 0.346 & 0.164 & \textbf{3.64$\times10^{-3}$} & \textbf{12.5} & \textbf{-4.19$\times10^{-2}$} & \textbf{2.29$\times10^{-2}$} & \textbf{-2.27$\times10^{-1}$}\\   
             \hline
@@
 
-    % Updated with network-level random effect
     \begin{figure}[ht!]
         \centering
-        \includegraphics[height=.4\textheight]{figures/participation_vs_SC.eps}
+        \includegraphics[width=.9\textwidth]{figures/participation_vs_SC.eps}
         \caption{The proportion of each motif in a consumer's participation vector varied sigificantly with connectance but not with network size or the interaction between size and connectance.
@@
     \end{figure}
-
-    Motif participation also varied with a species' degree and trophic level.
-    \begin{table}[hb!]
-        \centering
-        \caption{Motif participation varies with in-degree and trophic level.
-Here we show the coefficients for models relating the frequency of each motif in a species' participation vector to in-degree or trophic level (8 models in total).
-All relationships were significant ($\alpha$=0.05).
-        All models were fit with binomial error distribution and logit link function.
-        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
-        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
-        \label{tab:partic_vs_SC}   
-        \footnotesize
-        \begin{tabular}{c|c c }
-            Motif & in-degree & TL \\
-            \hline
-                3-sp Chain & \textbf{-6.99$\times10^{-3}$} & \textbf{9.84$\times10^{-2}$} \\
-            Omnivory & \textbf{2.29$\times10^{-2}$} & \textbf{-2.27$\times10^{-1}$}\\
-            App. Comp & \textbf{-8.76$\times10^{-3}$} & \textbf{2.40$\times10^{-1}$}\\
-            Dir. Comp & \textbf{-1.43$\times10^{-3}$} & \textbf{-2.52$\times10^{-1}$}\\
-            \hline
-            \end{tabular}
-            \end{table}
 

### 2023-04-18 manuscript/JAE_reply.tex
Updates from Overleaf

@@
                 \textbf{R:}
-            The sentence the reviewer refer to is "How disturbances such as population decline or extinction lead to the loss of species via  direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research (Curtsdotter et al., 2011a; Dunne & Williams, 2009; Ekl ̈of & Ebenman, 2006 Santos et al., 2021) " Here
+            The sentence the reviewer refer to is "How disturbances such as population decline or extinction lead to the loss of species via  direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research (Curtsdotter et al., 2011a; Dunne & Williams, 2009; Eklöf & Ebenman, 2006 Santos et al., 2021)." We can understand the reviewer´s confusion, but in fact species extinctions are both the disturbance and the response; an initial extinction of a species (the primary e
 

### 2023-04-18 manuscript/JAE_reply.tex

--- 2023-04-18 manuscript/JAE_reply.tex
@@
                 \textbf{R:}
-            The sentence the reviewer refer to is "How disturbances such as population decline or extinction lead to the loss of species via  direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research (Curtsdotter et al., 2011a; Dunne & Williams, 2009; Ekl ̈of & Ebenman, 2006 Santos et al., 2021) " Here
+            The sentence the reviewer refer to is "How disturbances such as population decline or extinction lead to the loss of species via  direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research (Curtsdotter et al., 2011a; Dunne & Williams, 2009; Eklöf \& Ebenman, 2006 Santos et al., 2021)." We can understand the reviewer´s confusion, but in fact species extinctions are both the disturbance and the response; an initial extinction of a species (i.e.
+the primary extinction) can, via species interactions, cause other species to extinct (i.e.
+secondary extinctions).
 

### 2023-04-18 manuscript/supplement_motifs.tex
fixed one more SI fig

@@
 
-    Motif participation varies with global-scale network properies.
-    % Updated with network-level random effect
     \begin{table}[hb!]
@@
         \includegraphics[width=.9\textwidth]{figures/participation_vs_SC.eps}
-        \caption{The proportion of each motif in a consumer's participation vector varied sigificantly with connectance but not with network size or the interaction between size and connectance.
-Because the interaction was not significant, we show relationships for intermediate values of size and connectance when plotting against the other variable.
-These relationships were steepest for apparent competition and omnivory.}
+        \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).
+We show relationships for the smallest and largest values of network size (S) and connectance (C).}
         \label{fig:roles_vs_SC}

### 2023-04-18 code/stat_analysis/participation_vs_SC.py

@@ -114,7 +114,7 @@ def populate_persgraph(graph,xprop,val,netprops):
     data.symbol.shape=0
     data.line.configure(linestyle=sty,linewdith=1.5,color=j)
 
-    if xprop=='C' and val==70:
+    if xprop=='C' and val==50:
       data.legend=motif_names[motif]
 
   graph.legend.configure(box_linestyle=0,box_fill_pattern=0,char_size=.75,loc=(0.1,0.74),loctype='world')
@@ -133,22 +133,28 @@ datafile='roles_vs_SC.tsv'
 netprops=read_datafile(datafile)
 
 grace=MultiPanelGrace(colors=colors)
-grace.add_label_scheme('dummy',['C=0.1','S=70'])
+grace.add_label_scheme('dummy',['S=50','C=0.02','S=100','C=0.20'])
 grace.set_label_scheme('dummy')
 
+for sel in ['lo','high']:
+  for xprop in ['C','S']:
+    if xprop=='S':
+      if sel=='lo':
+        val=0.02
+      else:
+        val=0.2
+    elif xprop=='C':
+      if sel=='lo':
+        val=50
+      else:
+        val=100
+    graph2=grace.add_graph(Panel)
+    graph2=format_graph(graph2,xprop)
+    graph2=populate_persgraph(graph2,xprop,val,netprops)
+    graph2.panel_label.configure(placement='iul',char_size=.75,dx=.02,dy=.02)
 
-for xprop in ['C','S']:
-  if xprop=='S':
-    val=0.1
-  elif xprop=='C':
-    val=70
-  graph2=grace.add_graph(Panel)
-  graph2=format_graph(graph2,xprop)
-  graph2=populate_persgraph(graph2,xprop,val,netprops)
-  graph2.panel_label.configure(placement='iul',char_size=.75,dx=.02,dy=.02)
 
-
-grace.multi(rows=1,cols=2,vgap=.06,hgap=.06)
+grace.multi(rows=2,cols=2,vgap=.06,hgap=.06)
 grace.hide_redundant_labels()
 grace.set_col_yaxislabel(label='Proportion of motif participation',col=0,rowspan=(None,None),char_size=1,perpendicular_offset=.07)
 

### 2023-04-18 manuscript/JAE_reply.tex
Updates from Overleaf

@@
 This would be clearer if the authors clarify whether they are studying the effects of disturbance at the primary producer level on consumer persistence right away.
-It would also be helpful to define exactly what definition of disturbance they are referring to in the introduction.
 
@@
             The sentence the reviewer refer to is "How disturbances such as population decline or extinction lead to the loss of species via  direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research (Curtsdotter et al., 2011a; Dunne & Williams, 2009; Eklöf \& Ebenman, 2006 Santos et al., 2021)." We can understand the reviewer´s confusion, but in fact species extinctions are both the disturbance and the response; an initial extinction of a species (i.e.
-the primary extinction) can, via species interactions, cause other species to extinct (i.e.
+the primary extinction) can, since species are connected to each other via for example consumer -- resource interactions, cause other species to extinct (i.e.
 secondary extinctions).
+We have now clarified this on line XX:
+            "The loss of a species in a food web can, due to the mutual dependencies among species, trigger a cascade of additional extinctions, so called secondary extinctions.
+A species' position within a food web is known to affect its risk of going secondarily extinct ~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}."
+
+            \item It would also be helpful to define exactly what definition of disturbance they are referring to in the introduction.
+
+             \textbf{R:}
+             We have now clarified how we here define a disturbance.
+This is done on line XX.
+"The loss of a species in a food web can, due to the mutual dependencies among species, trigger a cascade of additional extinctions, so called secondary extinctions.
+A species' position within a food web is known to affect its risk of going secondarily extinct ~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}."
 

### 2023-04-18 manuscript/supplement_motifs.tex
Updates from Overleaf

--- 2023-04-18 manuscript/supplement_motifs.tex
@@
             \hline
-            Three-species chain & \textbf{0.326} &  0.0260 & \textbf{-0.595} & -2.53$\times10^{-3}$ & 0.*958 \\
+            Three-species chain & \textbf{0.326} &  0.0260 & \textbf{-0.595} & -2.53$\times10^{-3}$ & 0.958 \\
             Apparent competition & \textbf{0.326} & \textbf{0.0501} & \textbf{-0.596} & -2.00$\times10^{-3}$ & 0.962 \\

### 2023-04-18 manuscript/reorder_main_text.tex

@@
 
-    A species' position within a food web (a network of who eats whom) is known to affect its ability to persist following the extinction of  another species in the network (i.e., secondary extinction)~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    The loss of a species in a food web can, due to the mutual dependencies among species,  trigger a cascade of additional extinctions, so called secondary extinctions.
+A species' position within a food web is known to affect its risk of going secondarily extinct ~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
     For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{binzer2011susceptibility, Eklof2006}.
@@
     Like degree and trophic level, a species' \emph{motif participation} (the vector of frequencies with which it appears in each motif) can easily be calculated once the structure of the network is known.
-    Unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the focal species' interaction partners.
+    However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the focal species' interaction partners.
     Two species might both have three prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
@@
 
+    The plant community is the foundation upon which the myriad species in food webs depend for their survival.
+Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
+    The loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
+\citep{scherber2010bottom}.
+Given the severity of disturbances, here loss of species, at the lowest trophic levels, we here choose to focus on persistence after the loss of basal resources (i.e., effects of bottom-up disturbances).
+    Species' motif participation may indicate \emph{which} consumers are more or less likely to go extinct following loss of primary producers.
+
     While a relationship between motif participation and persistence is broadly intuitive, so far there has been little research demonstrating which motifs are associated with greater probabilities of persistence.
-    Here we focus on persistence after the loss of basal resources (i.e., bottom-up disturbances).
-    The plant community is the foundation upon which the myriad species in food webs depend for their survival.Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-    The loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc. \citep{scherber2010bottom}.
-    Species' motif participation may indicate \emph{which} consumers are more or less likely to go extinct following such a disturbance.
-
     One potential reason for the difficulty in assessing the relationship between species' motif roles and persistence are the large computational resources required to simulate population dynamics.
     Our solution is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks~\citep{Eklof2013,Haussler2020}.
-    Although Bayesian networks currently only simulate bottom-up disturbances (our focal interest), they nevertheless capture the vast majority of secondary extinctions obtained from dynamical models including top-down and bottom-up effects~\citep{Eklof2013}.
-    The greater computational efficiency and similar [[results?
+    One limitation with Bayesian networks is that they only take into account bottom-up effects (see Methods).
+Nevertheless, Bayesian networks do capture the vast majority of secondary extinctions obtained from fully-fledged dynamical models including top-down and bottom-up effects~\citep{Eklof2013}, and as such are a good proxy for analysing species persistence following extinctions of primary producers in complex food webs.
+    %The greater computational efficiency and similar [[results?
 performance?
@@
 
-    Although there are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, we focus on just four: the three-species chain, apparent competition, direct competition, and omnivory.
-    These four motifs are by far the most common in empirical networks~\citep{Stouffer2007, Borrelli2015a, giling2019plant}.
-    They are also the only motifs which do not include cycles (where species A and B eat each other or where A eats B eats C eats A).
-    As modelling bottom-up extinctions in a Bayesian framework requires the breaking of cycles, these are the only motifs that will appear in the resulting \emph{acyclic} networks.
-    Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited.
-
-    Moreover, these four motifs have clear ecological interpretations, which are lacking for the cyclical motifs.
+    %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
+    We will here focus three-species motifs, which are the commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}.
+We particularly analyze four of the three-species motifs, namenly the three-species chain, apparent competition, direct competition, and omnivory.
+    We choose these four motifs as they are by far the most common in empirical networks~\citep{Stouffer2007, Borrelli2015a, giling2019plant}.
+    These four motifs also have clear ecological interpretations:
     The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
@@
     Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator. 
+    Furthermore, these four motifs do not include cycles (where species A and B eat each other or where A eats B eats C eats A) and as modelling secondary extinctions in a Bayesian network requires acyclic networks (see Methods), these are the only motifs that will appear in the resulting \emph{acyclic} networks.
+
+    %\subsection*{Here we test...}
+    Here, we are interested in whether a consumer's extinction risk after disturbance to the plant community (i.e., its risk of secondary extinction) depends on the species' motif participation and how this relationship varies with the severity of disturbance.
+    To provide context for this result, we also consider how extinction risk varies with the overall structure of the network.
+    Specifically, we consider the frequency of motifs across the whole network, network size, and connectance (the proportion of possible links which actually occur).
+    Finally, we relate species' motif participation to overall network structure and to their in-degrees and trophic levels in order to clarify what unique information motif participation can provide.
+
+     %They are also the only motifs which do not include cycles (where species A and B eat each other or where A eats B eats C eats A).
+    %As modelling bottom-up extinctions in a Bayesian framework requires the breaking of cycles, these are the only motifs that will appear in the resulting \emph{acyclic} networks.
+    %Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited.
 
@@
 %     The severity of disturbance therefore clearly has an effect on how the ecological community will respond.
-
-%\subsection*{Here we test...}
-    Here, we are interested in whether a consumer's extinction risk after disturbance to the plant community (i.e., its risk of secondary extinction) depends on the species' motif participation and how this relationship varies with the severity of disturbance.
-    To provide context for this result, we also consider how extinction risk varies with the overall structure of the network.
-    Specifically, we consider the frequency of motifs across the whole network, network size, and connectance (the proportion of possible links which actually occur).
-    Finally, we relate species' motif participation to overall network structure and to their in-degrees and trophic levels in order to clarify what unique information motif participation can provide.
 
@@
 
+    \subsection*{Species persistence and other network properties}
+
+        % The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=645, $p$\textless0.001), although this result must be considered tentative due to non-homogeneous distribution of roles (\emph{Appendix S8}).
+        As well as the frequency with which each species appears in each motif, we can use the total frequencies of each motif in the whole network (i.e., the network's motif profile) to describe its structure.
+        The average probability of persistence of all consumers in the network varied with the frequencies of different motifs in this profile.
+        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:motif_profile_persistence}, Table S4, \emph{Appendix S8}).
+        There was no significant interaction between the proportion of any motif in the network profile and disturbance (Table S4), \emph{Appendix S8}).
+
+        To provide context for the relationships between consumer persistence and motif participation or motif profile, we also tested whether persistence was related to more-familiar network properties.
+        Average consumer persistence was also lower in networks with higher connectance but was not significantly related to network size (Table S6, \emph{Appendix S10}).
+        As expected, a consumer's persistence tended to decrease with increasing STL (Fig.~\ref{fig:motifs_vs_TL_and_deg}, Table S3, \emph{Appendix S6}).
+        The relationship between in-degree and persistence depended on the strength of disturbance: higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.
+
     \subsection*{Motif participation and other network properties}
 
-        Species' motif participation varied depending on the global properties of the network: size, connectance, and their interaction (Fig. S1, Table S2; \emph{Appendix S5}).
-        Species appeared most often in the apparent competition motif in all types of networks, but the proportion of apparent competition decreased with increasing network size and connectance.
-        The proportion of the omnivory motif in a species' participation role increased sharply with increasing connectance, while the proportion of the direct competition motif increased slightly with increasing network size.
+        Both the motif profile of a network and species' motif participation were related to other measures of network structure.
+        Most simply, the proportion of the omnivory motif in the network's motif role increased with increasing connectance (Fig.~\ref{motif_proportion_lms}; Table S5, \emph{Appendix S9}).
+        The proportions of other motifs did not vary significantly with connectance and none were significantly related to network size.
+
+        A species' motif participation varied with network size, connectance, and their interaction. 
+        The proportion of the omnivory motif generally increased with increasing connectance and network size, but decreased with increasing network size in the most-connected networks (Fig.
+S2 and Table S2, \emph{Appendix S5}).
+        The proportion of apparent competition decreased strongly with increasing connectance and more weakly (but still significantly) with increasing network size.
+        The proportion of the three-species chain showed similar trends to apparent competition, while the proportion of direct competition showed very little change over the range of network size and connectance we consider (though the effects of connectance and the interaction between connectance and network size were significant).
 
@@
         A species' persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix S6}).
-        Increased trophic level is generally associated with decreasing species persistence, although the rate of decline is slower at high disturbance (Table S3, \emph{Appendix S6}).
-        In-degree shows an even stronger interaction with disturbance; having more prey species is associated with higher persistence without disturbance, but lower persistence at high levels of disturbance.
 
-        \subsubsection*{Consistency in effects of motif participation on species persistence}
+        % \subsubsection*{Consistency in effects of motif participation on species persistence}
 
-            When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally showed more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
-            However, this trend varied at different levels of disturbance, especially for the direct competition and omnivory motifs.
-            At low and medium connectances there was little change in the distribution of slopes for direct competition across all levels of disturbance. 
-            At high connectances, however, the proportion of positive slopes was both lower overall and decreased with increasing disturbance. 
-            For the omnivory motif these trends were reversed. 
-            Moreover, the omnivory motif has a particularly broad distribution of slopes (i.e., an inconsistent relationship to persistence) when network connectance was low. 
-            Network size had a much smaller effect than connectance (Fig. S4; \emph{Appendix S7}).
-
-    \subsection*{Network mean persistence and motif profiles}
-
-        The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=645, $p$\textless0.001), although this result must be considered tentative due to non-homogeneous distribution of roles (\emph{Appendix S8}).
-        Average persistence was higher in networks with higher proportions of the apparent competition motif and lower in networks with higher proportions of the omnivory motif (Table S4, \emph{Appendix S8}).
-        These trends were consistent across levels of disturbance (Fig.~\ref{fig:motif_profile_persistence}, Table S4), \emph{Appendix S8}).
-
-    \subsection*{Motif profiles and network properties}
-
-        Considering each motif separately, the proportion of the omnivory motif in a network's motif profile increased significantly with increasing connectance (Fig.~\ref{motif_proportion_lms}, Table S5, \emph{Appendix S9}).
-        The proportions of other motifs in a network's profile were not related to network size, connectance, or their interaction.
-        Consistent with the results of these GLMMs, a PERMANOVA test suggested that a network's motif profile as a whole was related to network size and connectance.
-        However, the assumption of equal variance across groups was violated for this test and the PERMANOVA results should therefore be considered indicative only.
+        %     When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally showed more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
+        %     However, this trend varied at different levels of disturbance, especially for the direct competition and omnivory motifs.
+        %     At low and medium connectances there was little change in the distribution of slopes for direct competition across all levels of disturbance. 
+        %     At high connectances, however, the proportion of positive slopes was both lower overall and decreased with increasing disturbance. 
+        %     For the omnivory motif these trends were reversed. 
+        %     Moreover, the omnivory motif has a particularly broad distribution of slopes (i.e., an inconsistent relationship to persistence) when network connectance was low. 
+        %     Network size had a much smaller effect than connectance (Fig. S4; \emph{Appendix S7}).
 

### 2023-04-19 manuscript/JAE_reply.tex
Updates from Overleaf

@@
              \textbf{R:}
-             We have now clarified how we here define a disturbance.
+             We have now clarified how we in this manuscript define a disturbance.
 This is done on line XX.
-"The loss of a species in a food web can, due to the mutual dependencies among species, trigger a cascade of additional extinctions, so called secondary extinctions.
-A species' position within a food web is known to affect its risk of going secondarily extinct ~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}."
+" The loss of  basal level diversity typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
+\citep{scherber2010bottom}.
+Given these far-reaching effects we here choose to analyse persistence after loss of basal resources, i.e., effects of bottom-up disturbances.
+We here refer to disturbance as loss of basal species.
+"
 
@@
                 \textbf{R:}
+            We agree with the reviewer that this was not so clearly expressed.
+We have now re-formulated this section and clarified with an example.
+This is now located at line XX and reads: "However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the focal species' interaction partners.
+Two species might both have three prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
+Importantly, since specialists are generally more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than a predator of generalists."
 
@@
                 \textbf{R:}
+            We have now rephrased this part, now located on line XX: "Therefore, a species' motif participation profile captures the indirect interactions the species are affected by, thereby capturing important structures that are missed by global (network level) and local (species level) properties.
+As such, motif participation may complement degree and trophic level to more fully predict species' extinction risk."
 
@@
                 \textbf{R:}
+            See comment above.
 
@@
                 \textbf{R:}
+                TO DO!
 

### 2023-04-19 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
+\clearpage
 \begin{spacing}{2.0}
@@
 
-    The loss of a species in a food web can, due to the mutual dependencies among species,  trigger a cascade of additional extinctions, so called secondary extinctions.
+    The loss of a species in a food web can, due to the mutual dependencies among species,  trigger a cascade of additional (secondary) extinctions. 
     A species' position within a food web is known to affect its risk of going secondarily extinct ~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
@@
     However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the focal species' interaction partners.
-    Two species might both have three prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
-    Importantly, since specialists are generally more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than a predator of generalists.
+    Two species might both have two prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
+    Importantly, since specialists are generally more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
+    Therefore, a species' motif participation profile captures the indirect interactions that affect the species, thereby capturing important structures that are missed by larger- or smaller-scale descriptions of the network.
+    As such, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
 
     The plant community is the foundation upon which the myriad species in food webs depend for their survival. 
-Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-    The loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
-\citep{scherber2010bottom}.
-Given the severity of disturbances, here loss of species, at the lowest trophic levels, we here choose to focus on persistence after the loss of basal resources (i.e., effects of bottom-up disturbances).
-    Species' motif participation may indicate \emph{which} consumers are more or less likely to go extinct following loss of primary producers.
+    Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016}.
+    The loss of  basal level diversity typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc. \citep{scherber2010bottom}. 
+    Given these far-reaching effects we here choose to analyse persistence after loss of basal resources, i.e., effects of bottom-up disturbances. 
+    We are particularly interested in whether the strength of this disturbance affects which consumers are most likely to go extinct. 
+    For example, participation in one motif might be associated with greater probability of extinction after a mild disturbance but be associated with less risk of extinction when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances but also provide the species with alternatives when disturbances are strong.
 
@@
     %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
-    We will here focus three-species motifs, which are the commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}.
-We particularly analyze four of the three-species motifs, namenly the three-species chain, apparent competition, direct competition, and omnivory.
-    We choose these four motifs as they are by far the most common in empirical networks~\citep{Stouffer2007, Borrelli2015a, giling2019plant}.
-    These four motifs also have clear ecological interpretations:
+    We define motif participation using three-species motifs, which are the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
+    Four of these motifs, namely the three-species chain, apparent competition, direct competition, and omnivory, are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, Borrelli2015a, giling2019plant}.
+    These four motifs also have clear ecological interpretations.
     The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
@@
     Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator. 
-    Furthermore, these four motifs do not include cycles (where species A and B eat each other or where A eats B eats C eats A) and as modelling secondary extinctions in a Bayesian network requires acyclic networks (see Methods), these are the only motifs that will appear in the resulting \emph{acyclic} networks.
+    Furthermore, these four motifs do not include cycles (where species A and B eat each other or where A eats B eats C eats A).
+    As modelling secondary extinctions using Bayesian networks requires acyclic networks (see Methods), these are the only motifs possible.
 
     %\subsection*{Here we test...}
-    Here, we are interested in whether a consumer's extinction risk after disturbance to the plant community (i.e., its risk of secondary extinction) depends on the species' motif participation and how this relationship varies with the severity of disturbance.
-    To provide context for this result, we also consider how extinction risk varies with the overall structure of the network.
-    Specifically, we consider the frequency of motifs across the whole network, network size, and connectance (the proportion of possible links which actually occur).
+    Here, we analyse whether a consumer's risk of secondary extinction after removal of basal resources depends on the consumer's motif participation and how this relationship varies with the severity of disturbance.
+    To provide context for these results, we also consider how a consumer's risk of secondary extinction varies with the overall structure of the network, described by the frequency of motifs across the whole network, network size, and network connectance.
     Finally, we relate species' motif participation to overall network structure and to their in-degrees and trophic levels in order to clarify what unique information motif participation can provide.
@@
 
-    	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and trophic level, STL), 3) how network mean persistence (the average persistence of all consumers in the network) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, in-degree, and trophic level).
+    	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and trophic level, STL), 3) how network mean persistence (the average persistence of all consumers in the network) varied with network motif profiles, 4) how network motif profiles varied with large-scale structural properties (S and C), and 5) how species persistence and network mean persistence varied with other measures of network structure and species' place within it (S, C, in-degree, and trophic level).
         Where applicable, we also consider how interactions between network structure and disturbance level might affect extinction risk.
@@
             Properties which refer to different scales of network structure, including different amounts of direct and indirect interactions with the focal species, are not independent. 
-            To identify correlations between global network structure (i.e., the whole network) and participation in particular motifs (a meso-scale description of species roles), we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
+            To identify correlations between large-scale network structure (i.e., the whole network) and participation in particular motifs (a meso-scale description of species roles), we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
             We fit a similar set of  GLMMs relating motif participation to in-degree or trophic level (i.e., local measures of a species' role; Appendix \emph{S3.2}).
@@
 
-            Persistence is known to vary with global-scale and local-scale network properties.
+            Persistence is known to vary with large- and small-scale network properties.
             To correctly interpret any relationships between motifs and persistence, we must consider the context of these other relationships (Appendix \emph{S3.5}).
-            To identify the relationship between global network structure and mean probability of persistence within a network, we fit a linear regression relating mean persistence to network size, connectance, disturbance, and their interactions (Eq.
+            To identify the relationship between large-scale network structure and mean probability of persistence within a network, we fit a linear regression relating mean persistence to network size, connectance, disturbance, and their interactions (Eq.
 S9).
@@
 S10-S11).
-            These GLMMs also included random effects to account for the effects of global network structure and network ID on persistence.
+            These GLMMs also included random effects to account for the effects of large-scale network structure and network ID on persistence.
 
@@
         Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:motif_profile_persistence}, Table S4, \emph{Appendix S8}).
-        There was no significant interaction between the proportion of any motif in the network profile and disturbance (Table S4), \emph{Appendix S8}).
+        There was no significant interaction between the proportion of any motif in the network profile and level of disturbance (Table S4), \emph{Appendix S8}).
 
-        To provide context for the relationships between consumer persistence and motif participation or motif profile, we also tested whether persistence was related to more-familiar network properties.
-        Average consumer persistence was also lower in networks with higher connectance but was not significantly related to network size (Table S6, \emph{Appendix S10}).
+        To provide context for the relationships between consumer persistence and motif participation or motif profile, we also tested whether persistence was related to more commonly used network properties.
+        Average consumer persistence was lower in networks with higher connectance, but was not significantly related to network size (Table S6, \emph{Appendix S10}).
         As expected, a consumer's persistence tended to decrease with increasing STL (Fig.~\ref{fig:motifs_vs_TL_and_deg}, Table S3, \emph{Appendix S6}).
@@
         Species with higher trophic levels tended to have higher proportions of apparent competition and three-species chains and lower proportions of omnivory and direct competition. 
-        A species' persistence varies with both species in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix S6}).
+        A species' persistence varied with both species in-degree and trophic level, where the most clear result is decreased persistence with increased trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix S6}).
 
@@
 
-    The effects of bottom-up disturbances on consumers depend on the global structure of the affected food web~\citep{Dunne2002, Eklof2006, PascualDunne2006}.
+    The effects of bottom-up disturbances on consumers depend on the large-scale structure of the affected food web~\citep{Dunne2002, Eklof2006, PascualDunne2006}.
     Likewise, the distribution of meso-scale structures within a network affects the network's stability~\citep{prill2005dynamic, bascompte2005simple}.
@@
     Identifying these species is a high priority for conservation efforts~\citep{Bottrilletal2008}.
-    To advance this goal, we systematically explore the relationships between meso-scale structures and species' extinction risk, as well as how these trends overlap with the effects of global- and local-scale network properties.
+    To advance this goal, we systematically explore the relationships between meso-scale structures and species' extinction risk, as well as how these trends overlap with the effects of large- and small-scale network properties.
 
@@
     In fact, a network's motif profile had a stronger relationship to mean persistence compared to both network size and connectance. 
-    This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.
+    This suggests that network motif profiles may be more useful than large-scale network properties when predicting mean extinction risk of species in a network.
     Connectance did, however, affect which motifs were most common within a network and therefore cannot be entirely ignored. 
     Highly-connected networks tended to contain higher frequencies of the omnivory motif, which includes more links than the other three-species motifs considered here.
-    Global-scale properties may therefore indirectly affect network persistence by changing meso-scale properties, such as the distribution of motifs.
+    Large-scale properties may therefore indirectly affect network persistence by changing meso-scale properties, such as the distribution of motifs.
 

### 2023-04-19 manuscript/JAE_reply.tex
clarified a few methods points

@@
 
-Proposed intro structure: 
-1. What a motif is and why we care. 2. How do motifs relate to persistence? 3. Disturbance to plant communities influences consumers differently dependeing on network structure
-
-- but that introduces persistence before we talk about disturbance ... bit circular. 
-- main thrust is that we need to introduce fewer concepts, in more depth.
-
     \textbf{R:}
@@
                 \textbf{R:}
+            [[we could also just rephrase these as small-scale and large-scale properties. Then it's maybe more obvious why meso-scale is in the middle.]]
             We have now rephrased this part, now located on line XX: "Therefore, a species' motif participation profile captures the indirect interactions the species are affected by, thereby capturing important structures that are missed by global (network level) and local (species level) properties.
@@
 For example, networks with a high degree of omnivory may be more stable than simple food chains, but this depends on the strength of the omnivorous interactions among trophic guilds.
-It would be helpful to provide more insight into the gaps in knowledge surrounding the prevalence of these motifs in food webs more generally, and why this would help food web researchers understand consumer persistence and how it is related to resilience/stability since this is where I see the greatest amount of novelty from this article.
+[[Did they take this from our discussion, the point they found contradictory?]] It would be helpful to provide more insight into the gaps in knowledge surrounding the prevalence of these motifs in food webs more generally, and why this would help food web researchers understand consumer persistence and how it is related to resilience/stability since this is where I see the greatest amount of novelty from this article.
 
@@
 As we disturb primary producers and do not simulate top-down effects, only consumers can go secondarily extinction in our framework.
-We now define secondary extinction parenthetically in the introduction (line XX) and methods (line XX) and hope that this is clearer.
+We now define secondary extinction parenthetically in the introduction and at the start of the methods (quoted below) and hope that the meaning of secondary extinction is now clear.
+
+                \begin{quotation} 
+                Traditionally, there are two main approaches for studying secondary extinctions (i.e., extinctions of species that were not directly disturbed). 
+                \end{quotation}
 
@@
 
-                \textbf{R:} [[Alyssa doesn't know how to respond to this one]]
+                \textbf{R:} Please note that we are using a Bayesian network modelling framework and not Bayesian statistics.
+All statistical models are quite standard linear models or general linear models.
+The Bayesian network method has been used and validated in previous papers (see references in the main text) and has not been substantially modified for this paper.
+The novelty in our work lies in the relationships between motifs and persistence, not in the network modelling \emph{per se}.
 
@@
                 \textbf{R:} A focal species is simply that for which the motif role is currently being calculated ...
-all species are focal species in turn.
+all consumers are taken as the focal species in turn.
 Note that as these networks are simulated the species (nodes) in them do not correspond to any particular focal species or trophic guild.
@@
 reptiles incorporated in the networks.
-[[Another really basic misunderstanding ]][[Seems to be this line: A species’ ``motif participation role'' is the frequency with which a focal species appears in each of the motifs present in a network (Stouffer et al., 2012).
-Is it clearer if we say "A species’ ``motif participation role'' is the frequency with which it appears in each of the motifs present in a network"?]]
+
+                We have revised this line (quoted below) and hope it is now clearer that a species' motif participation role refers to that same species' appearance in different motifs, and that this applies to any species one selects.
+                \begin{quotation}
+                A species' ``motif participation role'' is the frequency with which that species appears in each of the motifs present in a network~\citep{Stouffer2012}.
+                \end{quotation}
 
@@
 
-                \textbf{R:} We only considered three-species motifs in this study, as stated in lines XX, XX, and XX.
-The fact that only four \emph{unique} three-species motifs are possible in acyclic networks is a general feature of all acyclic networks; the other 9 possible three-species motifs contain cycles of two or three species (see~\citep{Stouffer2007} for a full list of these motifs).
+                \textbf{R:} The size of a motif participation vector is always related to the number of motifs considered, not the number of species (there is one vector per species).
+We only considered three-species motifs in this study, as stated in lines XX, XX, and XX.
+Since there are only four \emph{unique} three-species motifs in these acyclic networks (a general feature of all acyclic networks), motif participation vectors are four-dimensional for all species.
+This will be true for any motif participation roles calculated using three-species motifs and acyclic networks, unless one included placeholders of 0 for the 9 three-species motifs containing cycles (see~\citep{Stouffer2007} for a full list of these motifs).
+In that case all motif participation vectors would be 13-dimensional, but still only 4 dimensions would be informative.
+
+                We have rephrased this line (quoted below) to remove the phrase `in our case' and make it clearer that four-dimensional motif participation vectors are a general feature of acyclic networks.
+                \begin{quotation}
+                    As only four unique three-species motifs can appear in acyclic networks, these are four-dimensional vectors. 
+                \end{quotation}
+
                 Note that each motif can appear within the network many times with a different set of three interacting species (e.g., a network might contain the three-species-chains `hawk-mouse-oat',  `hawk-mouse-grasshopper', and `mouse-grasshopper-oat') so that the \emph{total occurrences} of all three-species motifs in a network will be much greater than four.
-                [[I think it's this line: In our case, these are four-dimensional vectors since only four three-species motifs can appear in acyclic networks]]
-                [[How can we clarify this line?]]
 

### 2023-04-20 manuscript/JAE_reply.tex
more reply upates

@@
 
-                \textbf{R:} [[Refers to: To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
-]] [[Based on many collective years of experience working with empirical food webs...
-is that acceptable to say?]]
+                \textbf{R:} The niche model has been previously demonstrated to generate plausible network architectures, including similar motif profiles (see citations in main text).
+We chose size and connectance in order to meet computational limitations (very large networks are difficult to model) and to overlap with the ranges of size and connectance in empirical food webs used to develop much of the current network theory.
+
+                We have lightly rephrased the start of this paragraph (quoted below) and added some references to papers describing empirical networks with similar size and connectance to those we simulate.
+                Also note that we compare our simlated results with results for a set of more recent, larger, and especially well-resolved empirical networks in the SI.
+Although some of the whole-network structural properties differed, the range of species motif participations in our simulated networks spanned the whole range of observed motif participation in these empirical networks.
+
+                \begin{quotation}
+                We generated simulated networks using the niche model, which has been shown to recreate the structure of empirical networks well~\citep{Williams2000,Stouffer2007}.
+                To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002a}, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
+                All networks were generated using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.   
+                For full details and a comparison to empirical networks, see \emph{Appendix S2}.
+                \end{quotation}
 

### 2023-04-20 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
     The plant community is the foundation upon which the myriad species in food webs depend for their survival. 
-    Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-    The loss of  basal level diversity typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc. \citep{scherber2010bottom}. 
+    Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
+    The loss of  basal level diversity typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
+\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
     Given these far-reaching effects we here choose to analyse persistence after loss of basal resources, i.e., effects of bottom-up disturbances. 
@@
 
-    While a relationship between motif participation and persistence is broadly intuitive, so far there has been little research demonstrating which motifs are associated with greater probabilities of persistence.
-    One potential reason for the difficulty in assessing the relationship between species' motif roles and persistence are the large computational resources required to simulate population dynamics.
-    Our solution is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks~\citep{Eklof2013,Haussler2020}.
-    One limitation with Bayesian networks is that they only take into account bottom-up effects (see Methods). 
-    Nevertheless, Bayesian networks do capture the vast majority of secondary extinctions obtained from fully-fledged dynamical models including top-down and bottom-up effects~\citep{Eklof2013}, and as such are a good proxy for analysing species persistence following extinctions of primary producers in complex food webs.
-    %The greater computational efficiency and similar [[results?
+    % While a relationship between motif participation and persistence is broadly intuitive, so far there has been little research demonstrating which motifs are associated with greater probabilities of persistence.
+    % One potential reason for the difficulty in assessing the relationship between species' motif roles and persistence are the large computational resources required to simulate population dynamics.
+    % Our solution is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks~\citep{Eklof2013,Haussler2020}.
+    % One limitation with Bayesian networks is that they only take into account bottom-up effects (see Methods). 
+    % Nevertheless, Bayesian networks do capture the vast majority of secondary extinctions obtained from fully-fledged dynamical models including top-down and bottom-up effects~\citep{Eklof2013}, and as such are a good proxy for analysing species persistence following extinctions of primary producers in complex food webs.
+    % %The greater computational efficiency and similar [[results?
 performance?
@@
     Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator. 
-    Furthermore, these four motifs do not include cycles (where species A and B eat each other or where A eats B eats C eats A).
-    As modelling secondary extinctions using Bayesian networks requires acyclic networks (see Methods), these are the only motifs possible.
+    % Furthermore, these four motifs do not include cycles (where species A and B eat each other or where A eats B eats C eats A).
+    % As modelling secondary extinctions using Bayesian networks requires acyclic networks (see Methods), these are the only motifs possible.
 
@@
 
+    \subsection*{Network construction and describing species roles}
+
+        We generated simulated networks using the niche model, which has been shown to recreate the structure of empirical networks well~\citep{Williams2000,Stouffer2007}.
+        To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002a}, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
+        All networks were generated using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.   
+        For full details and a comparison to empirical networks, see \emph{Appendix S2}.
+        We removed any biologically unlikely networks (i.e., those with extremely long paths between any consumer and basal resources; \emph{Appendix S2}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
+
+        % To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and species topologically sorted from lowest to highest trophic level (\emph{Appendix S2}).
+        %  These processing steps allow the strictly bottom-up calculation of probability of persistence.
+        We then described each consumer species' motif participation role in each of these simulated networks before applying any disturbance. 
+        A species' ``motif participation role'' is the frequency with which the species appears in each of the motifs present in a network~\citep{Stouffer2012}.
+        As only four unique three-species motifs can appear in acyclic networks, these are four-dimensional vectors. 
+        Here we are interested in the relative frequencies of each motif rather than the total number of motifs and therefore normalized participation vectors by dividing each count by the total number of motifs the species appears in such that the vector for each species sums to 1.
+        Note, however, that this normalization does not necessarily control for differences due to degree or connectance~\citep{Cirtwill2022Oikos}.
+
+        We also calculated simpler measures of a species' role within its community.
+        These were in-degree (number of prey) and trophic level (shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}).
+        Both simple role measures were calculated in R~\citep{R} using the same algorithm as~\citet{Eklof2013}.
+
+        Finally, to provide context for species' motif participation, we also calculated network ``motif profiles'': four-dimensional vectors of the number of each three-species motif in each network~\citep{Stouffer2012}.
+        Together with network size and connectance, these motif profiles describe the structure of the whole network rather than how a particular species participates in the community.
+        To separate differences in motif structure from differences in network size or connectance (larger and more-connected networks contain more motifs), we normalized motif profiles by dividing the count of each motif by the total count of all motifs.
+        We calculated species' motif participation using the Python package \emph{pymfinder}~\citep{pymfinder}.
+
     \subsection*{Modelling secondary extinction using Bayesian networks}
@@
 
-        A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks \citep{Eklof2013,Haussler2020}. 
+        A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks \citep{Eklof2013,Haussler2020}.
         In this framework, a consumer's probability of extinction depends on the fraction of resources lost ($f = k/n$ for a species with $n$ resources, of which $k$ have gone extinct) and on a baseline probability of extinction ($\pi$) which captures factors affecting the focal species' extinction probability that are not related to the network structure (e.g., disease, stochastic extinction of small populations).
-        Modelling Bayesian networks is fast and computationally efficient, allowing analysis of networks with high species richness \citep{Haussler2020}. 
-        This increased speed allows us to simulate a range of disturbance severities rather than single species removals as in many previous studies (e.g.,~\citealp[]{Memmott2004,Staniczenko2010,Dunne2004}).
+        Importantly, the greater speed of the Bayesian network framework allows us to simulate a range of disturbance severities rather than single species removals as in many previous studies (e.g.,~\citealp[]{Memmott2004,Staniczenko2010,Dunne2004,Cirtwill2022Oikos}).
         Moreover, Bayesian network simulations are less parameter-intensive than dynamical models, meaning that they are less sensitive to assumptions made in the modelling process~\citep{Eklof2013}.
+
+        % Mention acyclic resstriction here.
+        One limitation to Bayesian networks is that consumers do not affect the extinction probabilities of their resources (i.e., only bottom-up effects are included).
+        Nevertheless, Bayesian networks do capture the vast majority of secondary extinctions obtained from fully-fledged dynamical models including top-down and bottom-up effects~\citep{Eklof2013}, and as such are a good proxy for analysing species persistence following extinctions of primary producers in complex food webs.
+        As our focus in this study is on bottom-up effects from disturbance of basal resources, Bayesian networks will capture the secondary extinctions we are interested in.
+        Bayesian networks also require the creation of acyclic networks so that a strict bottom-up sequence can be followed. 
+        This means that cycles where species A eats species B and vice versa or where A eats B eats C eats A must be broken (see \emph{Appendix SX} for details).
+        Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited.
+        Thus, making the networks acyclic in order to use the Bayesian network approach will not strongly affect extinction scenarios.        
+        The removal of cycles also affects which motifs can appear in the network. 
+        In particular, any motif which contains a cycle is impossible in an acyclic network.
+        This means that only the four motifs which are most common in empirical networks (\citealp[]{Stouffer2007}: apparent competition, direct competition, omnivory, and the three-species chain) can appear.
+        Fortunately, they are also the motifs that have been most studied in isolation (e.g.,~\citep{}) and have the clearest ecological interpretation.
+
+        \subsubsection*{Modelling consumers' risk of secondary extinction}
 
@@
 
-		% BN limitations
-		Modelling species extinctions in food webs using Bayesian networks has two main limitations. 
-		First, in Bayesian networks consumers do not affect the extinction probabilities of their resources and as such we can here only capture bottom-up effects of disturbances. 
-		However, Bayesian networks has been shown to capture the majority of secondary extinctions produced by fully-fledged dynamical models \citep{Eklof2013}.
-		\cite{Eklof2013} analyzed which functional form of consumers' response to loss of resources most accurately predicted the secondary extinctions realized in the dynamical model. 
-		We here use that functional form in our analyzes.  
-		Moreover, bottom-up effects are often the main force of secondary extinctions in empirical systems. 
-		For example, bottom-up effects regulate zooplankton communities in freshwater lakes, regulating zooplankton communities \citep{li2020bottom} and control abundances and biomass of vertebrate species in several terrestrial systems \citep{Dobson2009food, Mduma1999food, Georgiadis2007}.
-		Second, due to the strict bottom-up approach of Bayesian networks, the networks must be acyclic.
-		Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited.
-		Thus, making the networks acyclic in order to use the Bayesian network approach will not strongly affect extinction scenarios.
-		Moreover, the cyclic motifs which are excluded from an acyclic network are rare in empirical networks~\citep{Stouffer2007} and therefore unlikely to strongly affect extinction risk.
-
-    \subsection*{Disturbance scenarios}
+        \subsubsection*{Disturbance scenarios}
 
@@
             Consumer species retained $\pi=0.1$ in all cases.
-
-	\subsection*{Initial network construction}
-
-		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.
-		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
-		For full details and a comparison to empirical networks, see \emph{Appendix S2}.
-		We removed any biologically unlikely networks (\emph{Appendix S2}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
-		To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and species topologically sorted from lowest to highest trophic level (\emph{Appendix S2}).
-		 These processing steps allow the strictly bottom-up calculation of probability of persistence.
-
-	\subsection*{Defining species roles}
-
-        A species' ``motif participation role'' is the frequency with which a focal species appears in each of the motifs present in a network~\citep{Stouffer2012}.
-        In our case, these are four-dimensional vectors since only four three-species motifs can appear in acyclic networks.
-        As we are interested in relationships between the type of motif participated in rather than the total number of motifs, we normalized participation vectors by dividing each count by the total number of motifs the species appears in.
-        Note, however, that this normalization does not necessarily control for differences due to degree or connectance as high- and low-degree species (or species in more- and less-connected networks) may also tend to participate in different types of motifs.
-        We address the potential associations between motif participation, degree, and connectance explicitly below.
-
-        To provide context for species' motif participation, we also calculated network ``motif profiles'': four-dimensional vectors of the number of each three-species motif in each network~\citep{Stouffer2012}.
-        To separate differences in motif structure from differences in network size (larger networks contain more motifs), we normalized motif profiles by dividing the count of each motif by the total count of all motifs.
-		We calculated species' motif participation using the Python package \emph{pymfinder}~\citep{pymfinder}.
-
-        Finally, we calculated simpler measures of a species' role within its community.
-        These were in-degree (number of prey) and trophic level (shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}).
-        Both simple role measures were calculated in R~\citep{R} using the same algorithm as~\citet{Eklof2013}.
 

### 2023-04-20 code/stat_analysis/per_network_regressions.R
added refs

@@ -31,8 +31,8 @@ consumers$prop_direct=consumers$m36/consumers$total_motifs
 consumers$prop_omni=consumers$m38/consumers$total_motifs
 consumers$netID=paste(consumers$Size,consumers$Connectance,consumers$Network,sep=':')
 
-per_network_results=matrix(nrow=0,ncol=6)
-colnames(per_network_results)=c("Network","S","C","Disturbance","Motif","Motif_slope")
+per_network_results=matrix(nrow=0,ncol=7)
+colnames(per_network_results)=c("Network","S","C","Disturbance","P_basal_extinct","Motif","Motif_slope")
 
 for(n in levels(as.factor(consumers$Network))){
     print(n)
@@ -40,19 +40,20 @@ for(n in levels(as.factor(consumers$Network))){
     S=netdat$Size[1]
     C=netdat$Connectance[1]
     for (d in c(0,0.2,0.4,0.6,0.8,1)){
+        pBas_total=0.1+0.4*d
         Ddat=netdat[which(netdat$Basal_p==d),]
         omnimod=with(Ddat,glm(Persistence~prop_omni,family='binomial'))
         chainmod=with(Ddat,glm(Persistence~prop_chain,family='binomial'))
         ACmod=with(Ddat,glm(Persistence~prop_apparent,family='binomial'))
         DCmod=with(Ddat,glm(Persistence~prop_direct,family='binomial'))
         per_network_results=rbind(per_network_results,
-            c(n,S,C,d,'Omnivory',summary(omnimod)$coefficients[2,1]))
+            c(n,S,C,d,pBas_total,'Omnivory',summary(omnimod)$coefficients[2,1]))
         per_network_results=rbind(per_network_results,
-            c(n,S,C,d,'Chain',summary(chainmod)$coefficients[2,1]))
+            c(n,S,C,d,pBas_total,'Chain',summary(chainmod)$coefficients[2,1]))
         per_network_results=rbind(per_network_results,
-            c(n,S,C,d,'Apparent',summary(ACmod)$coefficients[2,1]))
+            c(n,S,C,d,pBas_total,'Apparent',summary(ACmod)$coefficients[2,1]))
         per_network_results=rbind(per_network_results,
-            c(n,S,C,d,'Direct',summary(DCmod)$coefficients[2,1]))
+            c(n,S,C,d,pBas_total,'Direct',summary(DCmod)$coefficients[2,1]))
     }
 }
 

### 2023-04-20 manuscript/reorder_main_text.tex

@@
         We generated simulated networks using the niche model, which has been shown to recreate the structure of empirical networks well~\citep{Williams2000,Stouffer2007}.
-        To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002a}, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
+        To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002e}, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
         All networks were generated using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.   
@@
         As only four unique three-species motifs can appear in acyclic networks, these are four-dimensional vectors. 
-        Here we are interested in the relative frequencies of each motif rather than the total number of motifs and therefore normalized participation vectors by dividing each count by the total number of motifs the species appears in such that the vector for each species sums to 1.
+        Here we are interested in the relative frequencies of each motif rather than the total number of motifs and therefore normalized participation vectors by dividing each count by the total number of motifs the species appears in (such that the vector for each species sums to 1).
         Note, however, that this normalization does not necessarily control for differences due to degree or connectance~\citep{Cirtwill2022Oikos}.
 
-        We also calculated simpler measures of a species' role within its community.
-        These were in-degree (number of prey) and trophic level (shortest trophic level (STL): the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}).
+        We also calculated two simpler measures of a species' role within its community.
+        In-degree is a species' number of prey.
+        Trophic level (shortest trophic level; STL) is the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}).
         Both simple role measures were calculated in R~\citep{R} using the same algorithm as~\citet{Eklof2013}.
@@
 
-        A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks \citep{Eklof2013,Haussler2020}.
+        A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks than is practical for dynamical models \citep{Eklof2013,Haussler2020}.
         In this framework, a consumer's probability of extinction depends on the fraction of resources lost ($f = k/n$ for a species with $n$ resources, of which $k$ have gone extinct) and on a baseline probability of extinction ($\pi$) which captures factors affecting the focal species' extinction probability that are not related to the network structure (e.g., disease, stochastic extinction of small populations).
@@
         As our focus in this study is on bottom-up effects from disturbance of basal resources, Bayesian networks will capture the secondary extinctions we are interested in.
+
         Bayesian networks also require the creation of acyclic networks so that a strict bottom-up sequence can be followed. 
         This means that cycles where species A eats species B and vice versa or where A eats B eats C eats A must be broken (see \emph{Appendix SX} for details).
-        Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited.
-        Thus, making the networks acyclic in order to use the Bayesian network approach will not strongly affect extinction scenarios.        
+        Since cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited~\citep{Allesina2009functional}, this change will not strongly affect consumers' extinction risk in our simulations.
+
         The removal of cycles also affects which motifs can appear in the network. 
-        In particular, any motif which contains a cycle is impossible in an acyclic network.
+        Any motif which contains a cycle is impossible in an acyclic network.
         This means that only the four motifs which are most common in empirical networks (\citealp[]{Stouffer2007}: apparent competition, direct competition, omnivory, and the three-species chain) can appear.
-        Fortunately, they are also the motifs that have been most studied in isolation (e.g.,~\citep{}) and have the clearest ecological interpretation.
+        Fortunately, they are also the motifs that have been most studied in isolation (e.g.,~\citep{}) and have the clearest ecological interpretation and, as such, are of greatest interest to us.
 
@@
 
-        The response of the consumer to changes in $f$ can take any form (\emph{Appendix S1}). Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$).
-		This type of response most accurately captures the secondary extinctions produced by dynamical models (\citealp[]{Eklof2013}).
+        The response of the consumer to changes in $f$ can take any form (\emph{Appendix S1}).
+Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$), the type of response that most accurately captures the secondary extinctions produced by dynamical models (\citealp[]{Eklof2013}).
 		The probability of extinction for each species $i$ can therefore be represented using the cumulative density function of a beta distribution:
@@
         For each species, we then define probability of persistence as the fraction of simulations in which the species persists. 
-There are methods for solving Bayesian networks exactly \citep{Eklof2013}.
-	    However, for larger networks the described numerical evaluation is highly efficient and produces close to identical results \citep{Haussler2020}.
+        While there are methods for solving Bayesian networks exactly \citep{Eklof2013}, for larger networks the numerical evaluation above is highly efficient and produces close to identical results \citep{Haussler2020}.
 
@@
 
-            In our baseline scenario, all species had an extinction probability of $\pi = 0.1$. 
+            In our un-disturbed scenario, all species had an extinction probability of $\pi = 0.1$. 
             To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase all basal species' probability of extinction from $0.10$ to $0.5$, in steps of $0.08$.
@@
 
-    	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and trophic level, STL), 3) how network mean persistence (the average persistence of all consumers in the network) varied with network motif profiles, 4) how network motif profiles varied with large-scale structural properties (S and C), and 5) how species persistence and network mean persistence varied with other measures of network structure and species' place within it (S, C, in-degree, and trophic level).
-        Where applicable, we also consider how interactions between network structure and disturbance level might affect extinction risk.
+        We first tested 1) how consumer species persistence varied with motif participation and the severity of disturbance.
+        For context, we also test 2) how a consumer's persistence varies with other measures of network structure and a species' place within it.
+        As different measures of network structure are not independent since all describe the same network, we also 3) relate species' motif participation to large-scale network structure and to species in-degree and trophic level.
+        All analyses are briefly described below and were performed using the R~\citep{R} package \emph{lmerTest}~\citep{lmerTest}.
+        % , \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.
         See Appendix \emph{3.1 - 3.5} for details.
-    	All analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.
 
-        \subsubsection*{1) Species persistence and motif participation}
+        \subsubsection*{1) Consumer persistence and motif participation}
 
-            To test if participating more often in a certain motif affected a species' probability of persistence, we fit four general linear mixed-effect models with binomial error distribution (GLMMs; one per motif included in the Bayesian networks):
+            To test whether a consumer's risk of extinction was associated with the frequency of each motif in its participation vector and whether this relationship varied with the strength of disturbance, we fit four general linear models with binomial error distribution (GLMs; one per motif included in the Bayesian networks):
             \begin{equation}
-                    \Psi_{ink} \approx \rho_{i} + \pi_{k} + \rho_{i}\pi_{k} +
-                    S_{n}C_{n} + N_n,
+            \Psi_{ink} \approx \rho_{i} + \pi_{k} + \rho_{i}\pi_{k} ,
+            % S_{n}C_{n} + N_n,
             \label{propreq}
             \end{equation}
-            \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$,  $S_{n}C_{n}$ is a random intercept for the species richness and connectance of network $n$, and $N_n$ is a random effect of network $n$, intended to account for the non-independence of motif participations for species in the same network.
+            \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, and $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$.
             All predictors were centered and scaled before fitting the models.
+
+            Initially, we fit general linear mixed-effect models (GLMMs) also including random  intercepts for the species richness and connectance of network $n$ and for network identity.
+            As these models were singular, we removed each random intercept in turn and re-fit the model until it converged.
+            In all four cases, the convergent model did not retain any random effects.
+
             To test whether these general trends were consistent across networks, we visually examined the distribution of slopes of simplified regressions fit to a single network and level of disturbance (Appendix \emph{3.1}).
 
-        \subsubsection*{2) Motif participation and other network properties}
+        \subsubsection*{2) Consumer persistence and other network structure metrics}
 
-            Properties which refer to different scales of network structure, including different amounts of direct and indirect interactions with the focal species, are not independent. 
-            To identify correlations between large-scale network structure (i.e., the whole network) and participation in particular motifs (a meso-scale description of species roles), we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
-            We fit a similar set of  GLMMs relating motif participation to in-degree or trophic level (i.e., local measures of a species' role; Appendix \emph{S3.2}).
-            Taken together, these supplemental models demonstrate how motif participation relates to other measures of network structure.
-
-        \subsubsection*{3) Network mean persistence and motif profiles}
-
-            As well as species-specific probabilities of persistence, the overall frequency of motifs in a network may provide information about the overall probability of any species going extinct (since they incorporate information on indirect effects).
-            We tested whether networks with similar motif profiles tend to have similar mean persistences among consumers using a PERMANOVA (Appendix \emph{S3.3}).
-            To identify the relationships between particular motifs and network mean persistence in more detail, we fit a set of four LMs (one per motif):
+            The overall structure of a network could also affect consumers' ability to persist following a disturbance.
+            Here, we describe the overall network structure using the frequency of each motif in the whole network (i.e., the network's motif profile) as well as network size and connectance (which were determined prior to network simulation).
+            To test whether the frequency of particular motifs in the network profile were related to the average persistence of all consumers, we fit a set of four GLMs with binomial error distribution (one per motif):
                 \begin{equation}
@@
             \noindent where $\overline{\Psi_{nk}}$ is the mean persistence of all consumers in network $n$ at disturbance level $k$, $\Bar{\rho}_{n}$ is the proportion of the focal motif in the network's motif profile, and $\pi_k$ is the level of disturbance (probability of extinction for basal resources) as in equation~\ref{propreq}.
+            As with the GLMs relating each consumer's persistence to its motif participation, we initially attempted to fit GLMMs including random effects of network size and connectance and network identity but removed these random effects in order to obtain non-singular models.
 
-        \subsubsection*{4) Motif profiles and network properties}
-
-            To test whether networks with the the same size and connectance have more similar motif profiles, we used a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
-            In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMs:
-            \begin{equation}
-                \Bar{\rho_{n}} \approx \Sigma_{n} + \zeta_{n} + \Sigma_{n}\zeta_{n}
-            \label{profile_global}
-            \end{equation}
-            \noindent where $\Bar{\rho}_{n}$ is the proportion of the focal motif in the motif profile of network $n$ and $\Sigma_n$ and $\zeta_{n}$ are the size and connectance of network $n$.
-
-        \subsubsection*{5) Persistence and network properties}
-
-            Persistence is known to vary with large- and small-scale network properties.
-            To correctly interpret any relationships between motifs and persistence, we must consider the context of these other relationships (Appendix \emph{S3.5}).
-            To identify the relationship between large-scale network structure and mean probability of persistence within a network, we fit a linear regression relating mean persistence to network size, connectance, disturbance, and their interactions (Eq.
+            We also confirmed previously-known relationships between persistence and other measures of network structure or a species' participation in the network.
+            To identify the relationship between large-scale network structure and mean probability of persistence within a network, and how the strength of disturbance might affect this relationship, we fit a GLM relating mean persistence to network size, connectance, disturbance, and their interactions (Eq.
 S9).
@@
 
-            Persistence is also known to vary with local-scale measures of network structure. 
-            To identify the relationships between local-scale properties and species' probabilities of persistence, we fit two final GLMMs relating a consumer's persistence to disturbance, in-degree or STL, and the interaction between disturbance and the local-scale measure (Eq.
+            A consumer's probability of persistence is also known to vary with its in-degree and trophic level.
+            To confirm these relationships in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to disturbance, in-degree or STL, and the interaction between disturbance and the measure of network participation (Eq.
 S10-S11).
-            These GLMMs also included random effects to account for the effects of large-scale network structure and network ID on persistence.
+            As with the GLMs relating persistence to motif participation and motif profiles, we initially fit GLMMs including random effects of network size and connectance and network identity but removed these terms due to model singularity.
+
+        \subsubsection*{3) Motif participation and other network structure metrics}
+
+            Since a network's motif profile and a species' motif participation reflect the same network as other measures (size, connectance, degree, and trophic level), it is likely that these measures are related.
+            We tested whether the frequency of motifs in the network's motif profile were related to network size, connectance, and their interaction using a set of four GLMs with binomial error distribution.
+            We tested whether the frequency of motifs in a consumer's motif participation vector were related to network size, connectance, and their interaction using a similar set of four GLMs with binomial error distribution.
+            Finally, we fit two additional sets of GLMs (eight total) relating the frequency of each motif in a consumer's motif participation vector to the consumer's in-degree or trophic level.
+
+        % \subsubsection*{2) Motif participation and other network properties}
+
+        %     Properties which refer to different scales of network structure, including different amounts of direct and indirect interactions with the focal species, are not independent. 
+        %     To identify correlations between large-scale network structure (i.e., the whole network) and participation in particular motifs (a meso-scale description of species roles), we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
+        %     We fit a similar set of  GLMMs relating motif participation to in-degree or trophic level (i.e., local measures of a species' role; Appendix \emph{S3.2}).
+        %     Taken together, these supplemental models demonstrate how motif participation relates to other measures of network structure.
+
+        % \subsubsection*{4) Motif profiles and network properties}
+
+        %     To test whether networks with the the same size and connectance have more similar motif profiles, we used a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
+        %     In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMs:
+        %     \begin{equation}
+        %         \Bar{\rho_{n}} \approx \Sigma_{n} + \zeta_{n} + \Sigma_{n}\zeta_{n}
+        %     \label{profile_global}
+        %     \end{equation}
+        %     \noindent where $\Bar{\rho}_{n}$ is the proportion of the focal motif in the motif profile of network $n$ and $\Sigma_n$ and $\zeta_{n}$ are the size and connectance of network $n$.
 
@@
         The relationships between participation in the apparent competition and three-species chain motifs, however, varied with the strength of disturbance (Fig.~\ref{fig:prop_lmer_all}).
-        Higher frequency of participation in the apparent competition motif was associated with lower propbability of persistence at low levels of disturbance and higher probability of persistence at high levels of disturbance.
+        Higher frequency of participation in the apparent competition motif was associated with lower probability of persistence at low levels of disturbance and higher probability of persistence at high levels of disturbance.
         Participation in the three-species chain motif showed the opposite trends.

### 2023-04-20 manuscript/reorder_main_text.tex

@@
         This means that only the four motifs which are most common in empirical networks (\citealp[]{Stouffer2007}: apparent competition, direct competition, omnivory, and the three-species chain) can appear.
-        Fortunately, they are also the motifs that have been most studied in isolation (e.g.,~\citep{}) and have the clearest ecological interpretation and, as such, are of greatest interest to us.
+        Fortunately, they are also the motifs that have been most studied in isolation (e.g.,~\citep{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}) and have the clearest ecological interpretation and, as such, are of greatest interest to us.
 

### 2023-04-22 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
+    \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=\textwidth]{figures/persistence_omnivory_detailpers.eps}
+        \label{omnivory_plus_SC}
+        \end{figure}
+
+    \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=\textwidth]{figures/persistence_chain_detailpers.eps}
+        \label{chain_plus_SC}
+        \end{figure}
+
+    \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=\textwidth]{figures/persistence_apparent_detailpers.eps}
+        \label{apparent_plus_SC}
+        \end{figure}
+
+    \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=\textwidth]{figures/persistence_direct_detailpers.eps}
+        \label{direct_plus_SC}
+        \end{figure}
+
 \clearpage

### 2023-04-24 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-            To test whether these general trends were consistent across networks, we visually examined the distribution of slopes of simplified regressions fit to a single network and level of disturbance (Appendix \emph{3.1}).
+            NEW VERSION:
+
+            To test whether a consumer's risk of extinction was associated with the frequency of each motif in its participation vector and whether this relationship varied with the strength of disturbance, network size, and/or network connectance, we fit four general linear models with binomial error distribution (GLMs; one per motif which can occur in acyclic networks):
+             \begin{equation}
+             \begin{aligned}
+            \Psi_{ink} & \approx \rho_{i} + \pi_{k} + \Sigma_{n} + \Zeta_{n} \\
+            & + \rho_{i}\pi_{k} + \rho_{i}\Sigma_{n} + \rho_{i}\Zeta_{n} + \pi_{k}\Sigma_{n} + \pi_{k}\Zeta_{n} + \Sigma_{n}\Zeta_{n} \\
+            & + \rho_{i}\pi_{k}\Sigma_{n} + \rho_{i}\pi_{k}\Zeta_{n} + \pi_{k}\Sigma_{n}\Zeta_{n} + \rho_{i}\pi_{k}\Sigma_{n}\Zeta_{n} ,
+            % S_{n}C_{n} + N_n,
+            \label{propreq}
+            \end{aligned}
+            \end{equation}
+            \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$, $\Sigma_{n}$ is the number of species in the network, and $\Zeta_{n}$ is the connectance of the network.
+            All predictors were centered and scaled before fitting the models.    
+            After fitting the full models described above, we removed any non-significant terms in a stepwise fashion, beginning with the four-way interaction and proceeding to three-way and two-way interactions.
+            If a smaller term was included in a significant larger term (e.g., a non-significant two-way interaction involving terms in a significant three-way interaction) then the smaller term was retained.
+            We attempted to fit a PERMANOVA test relating a species' motif participation vector as a whole to its persistence, but the assumption of equal variance among groups was violated and we therefore do not report the results of this test.
+
+            To test whether these general trends were consistent across networks, we visually examined the distribution of slopes of simplified regressions fit to a single network and level of disturbance.
+            These models included only the frequency of the focal motif ($\rho_{i}$ above) and again used a binomial error distrubution.
+            [[Statement about what this tells us or something else?]]
 
@@
             Since a network's motif profile and a species' motif participation reflect the same network as other measures (size, connectance, degree, and trophic level), it is likely that these measures are related.
-            We tested whether the frequency of motifs in the network's motif profile were related to network size, connectance, and their interaction using a set of four GLMs with binomial error distribution.
+            We tested whether the frequency of motifs in the network's motif profile were related to network size, connectance, and their interaction using a set of four GLMs with binomial error distribution (Appendix \emph{S3.1}, eq.
+S2).
             We tested whether the frequency of motifs in a consumer's motif participation vector were related to network size, connectance, and their interaction using a similar set of four GLMs with binomial error distribution.
-            Finally, we fit two additional sets of GLMs (eight total) relating the frequency of each motif in a consumer's motif participation vector to the consumer's in-degree or trophic level.
-
-        % \subsubsection*{2) Motif participation and other network properties}
-
-        %     Properties which refer to different scales of network structure, including different amounts of direct and indirect interactions with the focal species, are not independent. 
-        %     To identify correlations between large-scale network structure (i.e., the whole network) and participation in particular motifs (a meso-scale description of species roles), we fit a series of LMs relating motif participation to network size, connectance, and their interaction (Appendix \emph{S3.2}).
-        %     We fit a similar set of  GLMMs relating motif participation to in-degree or trophic level (i.e., local measures of a species' role; Appendix \emph{S3.2}).
-        %     Taken together, these supplemental models demonstrate how motif participation relates to other measures of network structure.
-
-        % \subsubsection*{4) Motif profiles and network properties}
-
-        %     To test whether networks with the the same size and connectance have more similar motif profiles, we used a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction (Appendix \emph{S3.4}).
-        %     In addition, we identify correlations between size, connectance, and particular motifs using a set of four LMs:
-        %     \begin{equation}
-        %         \Bar{\rho_{n}} \approx \Sigma_{n} + \zeta_{n} + \Sigma_{n}\zeta_{n}
-        %     \label{profile_global}
-        %     \end{equation}
-        %     \noindent where $\Bar{\rho}_{n}$ is the proportion of the focal motif in the motif profile of network $n$ and $\Sigma_n$ and $\zeta_{n}$ are the size and connectance of network $n$.
+            Finally, we fit two additional sets of GLMs (eight total) relating the frequency of each motif in a consumer's motif participation vector to the consumer's in-degree or trophic level (Appendix \emph{S3.1}, eq.
+S3, S4).
 

### 2023-04-24 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 \usepackage[dvips]{graphicx}
-\usepackage{multirow} 
+\usepackage{multirow,multicol} 
 \usepackage{geometry} 
@@
             \begin{itemize}
-                \item Species persistence and motif participation (Eqn. S2-3)
-                \item Motif participation and network properties (Eqn. S4-6)
-                \item Network mean persistence and motif profiles (Eqn. S7)
-                \item Network motif profiles and global network structure (Eqn. S8)
-                \item Persistence and other network properties (Eqn. S9-11)
+                \item Persistence and other network properties (Eqn. S2-3)
+                \item Motif participation, profiles and network properties (Eqn. S4-6)
+                \item Network motif profiles and global network structure (Eqn. S7)
             \end{itemize}
@@
 
-    \subsection{How does species persistence vary with motif participation?}
+    \textbf{Persistence, degree, and trophic level}
 
-        We are first interested in whether participating more often in a certain motif (e.g., direct competition) might increase a species' probability of persistence.
-        To test this, we fit four general linear mixed-effect models with binomial error distribution (GLMMs; one per motif included in the Bayesian networks):
+            In-degree and trophic level are known to influence persistence in many cases and these effects, in combination with differences in mean persistence across networks with different global properties, should also be taken into account when considering relationships between motif participation and persistence.
+            To establish this necessary context, we fit GLMs with binomial error distribution:
+
+            \begin{spacing}{1.0}
+            \begin{multicols}{2}
+            \begin{equation}
+                    \Psi_{ink} \approx \delta_{i} + \pi_{k} + \delta_{i}\pi_{k},
+                    \label{degeq}
+                \end{equation}
 
             \begin{equation}
-            \Psi_{ink} \approx \rho_{i} + \pi_{k} + \rho_{i}\pi_{k} +
-            S_{n}:C_{n} + N_n,
-            \label{propreq}
+                    \Psi_{ink} \approx \tau_{i} + \pi_{k} + \tau_{i}\pi_{k},
+                    \label{TLeq}
                 \end{equation}
+            \end{multicols}
+            \end{spacing}
 
-        \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_{k}$ is the probability of extinction for a basal resource in disturbance level $k$, and $S_{n}C_{n}$ is a random intercept for the species richness and connectance of the network containing species $i$, and $N_n$ is a random effect of network ID.
-        The network-level random effect accounts for the fact that species in the same network are not independent.
-        We centered and scaled the predictors before fitting the GLMMs.
-        All models were fit using the R~\citep{R} function `glmer' from the package \emph{lmerTest}~\citep{lmerTest}.
-        We calculated marginal (fixed effects only) and conditional (fixed and random effects) $R^2$ using the R~\citep{R} function `r.squaredGLMM' from the package \emph{MuMIn}~\citep{MuMIn}.
-        Due to model singularity, we removed the network-level random effect and re-fit the models.
-        After re-fitting, the models converged but the estimated variance explained by the random intercept was 0.
-        We therefore removed this random effect as well and re-fit the model as a general linear model including only the fixed effects described above.
+            \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$,
+            $\delta_{i}$ and $\tau_i$ are the in-degree and trophic level of species $i$, respectively.
+            We initially included random effects of network size and connectance and network identity but removed these terms in order to obtain a non-singular model.
+            We centered and scaled the predictors before fitting the models.
 
-        We were also interested in whether these general trends were consistent across networks with different global properties and different levels of disturbance. 
-        To test this, we performed a linear regression of persistence against motif participation within each simulated network, for each level of disturbance to basal resources:
-
-        \begin{equation}
-            \Psi_{ik} \approx \rho_{i} ,
-            \label{mineq}
-        \end{equation}
-
-        \noindent where all symbols are as in equation~\ref{propreq}.
-        These regressions were fit using the R~\citep{R} base function `lm'.
-        We then visually examine the distribution of slopes for the effect of motif participation across levels of disturbance, network size, and connectance.
-
-    \subsection{How does motif participation vary with network properties?}
+    \textbf{Motif participation, profiles and other network properties}
 
         Species' motif participation may co-vary with other measures of network structure and species' roles within the network.
-        To determine the relationships between motif participation and global network structure, we fit a set of four linear mixed-effect models (one per motif):
+        To determine the relationships between motif participation and global network structure, we fit a set of four general linear models with binomial error distribution (one per motif):
 
+        \begin{spacing}{1.0}
         \begin{equation}
-            \rho_{in} \approx \Sigma_{n} + \zeta_{n} + \Sigma_{n}\zeta_{n} + N_n ,
+            \rho_{in} \approx \Sigma_{n} + \zeta_{n} + \Sigma_{n}\zeta_{n},
             \label{partic_SC}
         \end{equation}
+        \end{spacing}
 
-        \noindent where $\rho_{in}$ is the proportion of the role of species $i$, belonging to network $n$, that is made up by the focal motif,
-        $\Sigma_{n}$ is the size of network $n$, $\zeta_{n}$ is the connectance of the network $n$, and $N_n$ is a random effect of network ID.
+        \noindent where $\rho_{in}$ is the frequency with which species $i$, belonging to network $n$, participates in the focal motif, $\Sigma_{n}$ is the size of network $n$, and $\zeta_{n}$ is the connectance of the network $n$.
+        As above, we initially included a random effect of network identity but removed this term in order to obtain a non-singular model.
         Predictors were not centered or scaled before fitting.
-        All models were fit using the R~\citep{R} base function `lm'.
 
-        We fit a similar set of eight linear mixed-effect models relating motif participation to in-degree or trophic level:
+        We fit a similar set of eight general linear models with binomial error distribution relating motif participation to in-degree or trophic level:
 
+        \begin{spacing}{1.0}
+        \begin{multicols}{2}
         \begin{equation}
-            \rho_{in} \approx \delta_{i} + S_{n}C_{n} + N_n,
+            \rho_{in} \approx \delta_{i},
             \label{partic_deg}
@@
         \begin{equation}
-            \rho_{in} \approx \tau_{i} + S_{n}C_{n} + N_n,
+            \rho_{in} \approx \tau_{i},
             \label{partic_tl}
         \end{equation}
+        \end{multicols}
+        \end{spacing}
 
-        \noindent where $\rho_{in}$ is the frequency with which species $i$, belonging to network $n$, participates in the focal motif, $\delta_{i}$ is the in-degree of species $i$, $\tau_{i}$ is the shortest trophic level of species $i$, $S_{n}C_{n}$ is a random effect of the combination of size and connectance for network $n$, and $N_n$ is a random effect of network ID.
+        \noindent where all symbols are as above. 
+        As above, we initially included random effects of network size and connectance and network identity but removed these terms in order to obtain a non-singular model.
         These models were fit using the R~\citep{R} function `lmer' from the package \emph{lmerTest}~\citep{lmerTest}.
 
-    \subsection{How does mean persistence vary with motif profiles?}
-
-        To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
-        As with the PERMANOVA relating motif participation to network properties, we fit the PERMANOVA using the R~\citep{R} function `adonis' from the package \emph{vegan}~\citep{vegan} and calculated significance based on 9999 permutations.
-
-        We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
-        Dispersion was calculated using the R~\citep{R} function `betadisper' from the package \emph{vegan}~\citep{vegan}.
-        In case high or low mean persistence is associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable).
-
-        To supplement these overall tests, we also fit four linear models relating mean persistence to network motif profiles, disturbance, and the interaction between them:
-
-            \begin{equation}
-                \overline{\Psi_{nk}} \approx \Bar{\rho}_{n} + \pi_{k} + \Bar{\rho}_{n}\pi_{k} ,
-                \label{netpropeq}
-            \end{equation}
-
-        \noindent where $\overline{\Psi_{nk}}$ is the mean persistence of all consumers in web $n$ at disturbance level $k$, $\Bar{\rho}_{n}$ is the proportion of the focal motif in the network's motif profile, and $\pi_k$ is the level of disturbance (probability of extinction for basal resources) as in equation~\ref{propreq}.
-        All regressions were fit using the R~\citep{R} base function `lm'.
-
-    \subsection{How do network motif profiles vary with global structure?}
+    \textbf{Network motif profiles, size, and connectance}
 
@@
         This is especially likely since we include only networks which can retain all initial species after 1000 rounds of simulated Lotka-Volterra population dynamics~\citep{Cirtwill2021_inprep}:  some meso-scale structures might be more likely to allow all species to persist than others in networks with different size and connectance constraints.
-        We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
-        Significance was calculated based on 9999 permutations.
-        PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
-        To test whether this applies in our case, we fit an ANOVA of the dispersion of motif profiles about the centroid for each combination of network size and connectance. 
-        We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
-
-        To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif in a network's motif profile to network size, connectance, and their interaction:
-
+        To test for this possibility, we fit four general linear models with binomial error distribution (one per motif) relating the proportion of a motif in a network's motif profile to network size, connectance, and their interaction:
         \begin{equation}
@@
         \end{equation}
-        where all symbols are as in equations~\ref{propreq} and~\ref{netpropeq}.
-        We fit each regression using the R~\citep{R} base function `lm'.
-
-    \subsection{How does persistence vary with other properties?}
-
-        \subsubsection{Network mean persistence and network properties}
-
-            Both species richness (S) and connectance (C) are known to influence persistence in both simulation studies and empirical food webs.
-            In addition, increasing the probability of extinction for basal species ($\pi$) is intended to decrease persistence of species at higher trophic levels.
-            Although these effects are not the main focus of our manuscript, we must establish these background effects before measuring the effects of meso-scale structure on persistence.
-            The persistence of basal species in our simulations is determined only by their baseline probability, being either $\pi = 0.1$ or higher, dependent on $\pi$.
-            We therefore considered only non-basal species (i.e., those with at least one prey) throughout this manuscript.
-
-            To test for effects of global network properties, we fit a linear regression relating the mean of consumer species' probabilities of persistence in a network to network size, connectance, the level of disturbance to basal resources, and all interactions between them.
-            All predictors were centered and scaled before fitting the model. 
-            As the three-way interaction was significant, we did not simplify the model.
-            All models were fit using the R~\citep{R} base function `lm'.
-
-            \begin{equation}
-                \overline{\Psi_{nk}} \approx \Sigma_{n} + \zeta_{n} + \pi_k + \Sigma_{n}:\zeta_{n} + \Sigma_{n}\pi_k + \zeta_{n}\pi_k + \Sigma_{n}\zeta_{n}\pi_k,
-                \label{SCeq}
-            \end{equation}
-            where all symbols are as in equation~\ref{propreq} and~\ref{partic_SC}.
-
-        \subsubsection{Species persistence and species properties}
-
-            Likewise, in-degree and trophic level are known to influence persistence in many cases.
-            These effects, in combination with differences in mean persistence across networks with different global properties, should also be taken into account when considering relationships between motif participation and persistence.
-            To establish this necessary context, we therefore fit two linear mixed-effects regressions:
-
-            \begin{equation}
-                    \Psi_{ink} \approx \delta_{i} + \pi_{k} + \delta_{i}\pi_{k} +
-                    S_{n}C_{n} + N_n,
-                    \label{degeq}
-                \end{equation}
-
-            \begin{equation}
-                    \Psi_{ink} \approx \tau_{i} + \pi_{k} + \tau_{i}\pi_{k} +
-                    S_{n}C_{n} N_n,
-                    \label{TLeq}
-                \end{equation}
-
-            where $\delta_{i}$ and $\tau_i$ are the in-degree and trophic level of species $i$, respectively, and all other terms are as in equation~\ref{propreq}.
-            We centered and scaled the predictors before fitting the models.
+        where all symbols are as in equation~\ref{partic_SC}.
 
@@
 
-    % Updated and expanded to include permanova tests
-    Mean persistence of all consumers varied with the frequencies of different motifs in a network profile ($F_{1,2998}$=645, $p$\textless0.001 for a PERMANOVA of dissimilarity in motif profiles against mean persistence).
-    However, motif profiles were not evenly distributed across mean persistences ($F_{134,2865}$=1.77, $p$\textless0.001 for an ANOVA of dissimilarity in motif profiles against mean persistence, rounded to 3 decimal places).
-    Specifically, motif profiles were more variable (higher dissimilarities) in networks with higher mean persistence.
-    As non-homogeneous variability can cause a PERMANOVA to return a false positive~\citep{Anderson2001}, this means that we must treat our PERMANOVA result as tentative and place more weight on the single-motif regressions.
-
     % Updated to use mean instead of all species
@@
 
-    A network's normalized motif profile (i.e., proportions of the total made up by each motif) was related to the network's size, connectance, and the interaction between them ($F_{1,2996}$=74.3, $p$\textless0.001; $F_{1,2996}$=4203, $p$\textless0.001; and $F_{1,2996}$=27.8, $p$\textless0.001, respectively in a PERMANOVA of dissimilarity in normalized network motif profiles against network size, connectance, and their interaction).
-    However, normalized motif profiles were not homogeneously variable across levels of network size, connectance, or their interaction  ($F_{5,2994}$=16.8, $p$\textless0.001; $F_{4,2995}$=31.7, $p$\textless0.001; and $F_{29,2970}$=11.9, $p$\textless0.001 for ANOVAs of dispersion of motif profiles against size, connectance, and combinations of size and connectance, respectively).
-    These differences in variability can cause false positives in the PERMANOVA test. Variability was highest in small and low-connectance networks (Fig.~\ref{dispersion_normmotifs}).
     A series of general linear regressions support the relationship between network motif profiles and global structural properties.

### 2023-04-25 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
             We also confirmed previously-known relationships between persistence and other measures of network structure or a species' participation in the network.
-            To identify the relationship between large-scale network structure and mean probability of persistence within a network, and how the strength of disturbance might affect this relationship, we fit a GLM relating mean persistence to network size, connectance, disturbance, and their interactions (Eq.
-S9).
+            To identify the relationship between large-scale network structure and mean probability of persistence within a network, and how the strength of disturbance might affect this relationship, we fit a GLM relating mean persistence to network size, connectance, disturbance, and their interactions.
             As the three-way interaction was significant, we did not simplify the model.
@@
             A consumer's probability of persistence is also known to vary with its in-degree and trophic level.
-            To confirm these relationships in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to disturbance, in-degree or STL, and the interaction between disturbance and the measure of network participation (Eq.
-S10-S11).
+            To confirm these relationships in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to disturbance, in-degree or STL, and the interaction between disturbance and the measure of network participation (\emph{Appendix S3}, Eq.
+S2-3).
             As with the GLMs relating persistence to motif participation and motif profiles, we initially fit GLMMs including random effects of network size and connectance and network identity but removed these terms due to model singularity.
@@
             Since a network's motif profile and a species' motif participation reflect the same network as other measures (size, connectance, degree, and trophic level), it is likely that these measures are related.
-            We tested whether the frequency of motifs in the network's motif profile were related to network size, connectance, and their interaction using a set of four GLMs with binomial error distribution (Appendix \emph{S3.1}, eq.
-S2).
+            We tested whether the frequency of motifs in the network's motif profile were related to network size, connectance, and their interaction using a set of four GLMs with binomial error distribution (Appendix \emph{S3}, eq.
+S4).
             We tested whether the frequency of motifs in a consumer's motif participation vector were related to network size, connectance, and their interaction using a similar set of four GLMs with binomial error distribution.
-            Finally, we fit two additional sets of GLMs (eight total) relating the frequency of each motif in a consumer's motif participation vector to the consumer's in-degree or trophic level (Appendix \emph{S3.1}, eq.
-S3, S4).
+            Finally, we fit two additional sets of GLMs (eight total) relating the frequency of each motif in a consumer's motif participation vector to the consumer's in-degree or trophic level (Appendix \emph{S3}, Eq.
+S5-6).
 

### 2023-04-25 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 
-    Appendices S1-S3 provide supplemental methods, while appendices S4-S8 provide supplemental results.
+    Appendices S1-S3 provide supplemental methods, while appendices S4-S6 provide supplemental results.
 
@@
             \item Comparison of empirical and simulated networks
+            \item Contains Fig. S1
             \end{itemize}
 
-        \item Statistical analysis
+        \item Supplemental statistical analysis
 
@@
                 \item Motif participation, profiles and network properties (Eqn. S4-6)
-                \item Network motif profiles and global network structure (Eqn. S7)
             \end{itemize}
 
-        \item Persistence and motif participation
+        \item Results: species motif participation 
             \begin{itemize}
-                \item Contains Table S1
+                \item Supplemental results relating species motif participation to persistence and to other network properties (S, C, in-degree, STL)
+                \item Contains Table S1-2
+                \item Contains Fig. S2
             \end{itemize}
 
-        \item Participation and network properties
+        \item Results: persistence and other network properties
 
             \begin{itemize}
-                \item Contains Table S2
-                \item Contains Fig. S1
-            \end{itemize}
-
-        \item Persistence, degree, and trophic level
-
-            \begin{itemize}
-                \item Contains Table S3
-                \item Contains Fig. S2-3
+                \item Identifying relationships between persistence and network size, connectance, in-degree, and STL to provide context for motif relationships.
+                \item Contains Tables S3-4
+                \item Contains Fig. S3
             \end{itemize}
@@
 
-        \item Persistence and network motif profiles
+        \item Results: network motif profiles
 
             \begin{itemize}
-                \item Contains Table S4
-            \end{itemize}
-
-        \item Motif profiles and global properties
-
-            \begin{itemize}
-                \item Contains Figure S5
-                \item Contains Table S5
-            \end{itemize}
-
-        \item Persistence and global network properties
-            \begin{itemize}
-                \item Contains Figures S6, S7
-                \item Contains Table S6
+                \item Relating motif profiles to persistence (see main text for equation, figure) and to network size and connectance
+                \item Contains Tables S5-6
             \end{itemize}
@@
 
-\section{Statistical analysis}
+\section{Supplemental statistical analysis}
 
@@
 
-    \textbf{Network motif profiles, size, and connectance}
-
-        As well as affecting persistence directly, global network structure (i.e., size and connectance) may affect the meso-scale structure of the network.
-        This is especially likely since we include only networks which can retain all initial species after 1000 rounds of simulated Lotka-Volterra population dynamics~\citep{Cirtwill2021_inprep}:  some meso-scale structures might be more likely to allow all species to persist than others in networks with different size and connectance constraints.
-        To test for this possibility, we fit four general linear models with binomial error distribution (one per motif) relating the proportion of a motif in a network's motif profile to network size, connectance, and their interaction:
-        \begin{equation}
-            \Bar{\rho_{n}} \approx \Sigma_{n} + \zeta_{n} + \Sigma_{n}\zeta_{n}
-        \end{equation}
-        where all symbols are as in equation~\ref{partic_SC}.
-
 \clearpage
 
-\section{Persistence and motif participation} [updated]
+\section{Results: motif participation}
 
-    Persistence was significantly related to the proportion of each motif in the species' role (Table~\ref{tab:proportion}).
-Persistence tended to increase with increasing proportions of the direct competition and, at high levels of disturbance, increasing frequencies of the apparent competition motif.
-Persistence tended to decrease with increasing proportions of the omnivory motif and, at high levels of disturbance, the three-species chain.
+    \subsection{Participation and persistence}
 
@@
         \centering
-        \caption{Coefficients ($\beta$) for general linear models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction  (Equation~\ref{propreq}).
+        \caption{Coefficients ($\beta$) for general linear models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction  (Eq.
+1, main text).
 Coefficients in \textbf{bold} are significant ($\alpha$\textless0.005).
@@
     \end{table}
-\clearpage
 
-\section{Participation and network properties} [updated]
+\subsection{Participation and network properties}
 
@@
 
-    \begin{figure}[ht!]
+    \begin{figure}[hb!]
         \centering
@@
 
-\section{Persistence, degree, and trophic level} [updated]
+\section{Results: persistence and other network properties}
 
-    Persistence varies with in-degree, trophic level, and their interactions with disturbance.
+    \begin{figure}[ht!]
+        \centering
+        \includegraphics[width=0.9\textwidth]{figures/persistence_vs_SC_lm.eps}
+        \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.}
+        \label{fig:lm_CS}
+    \end{figure}
+
+    % Updated to use mean rather than all species.
+    \begin{table}[hb!]
+        \caption{Mean persistence of consumers within a network depends on connectance (C) and the level of disturbance to basal resources (D), but not network size (S) or interactions between properties.
+We show effect sizes ($\beta$) and $p$-values for each coefficient, except the intercept (see Appendix S3 for details).
+McFadden's pseudo-$R^2$=0.966.
+}
+        \label{tab:per_vs_SC}
+        \centering
+        \footnotesize
+        \begin{tabular}{c| c c c c c c c}
+              & S & C & D & S:C & S:D & C:D & S:C:D \\
+            \hline
+             $\beta$ &	 -7.21$\times10^{-4}$  	& \textbf{-6.12$\times10^{-2}$}	& \textbf{-0.596} &	 -4.48$\times10^{-3}$ &	 -1.63$\times10^{-2}$ 	& -4.52$\times10^{-3}$ &	 -1.35$\times10^{-3}$ \\
+             $p$-value &		 0.964 &	 \textless0.001 &	 \textless0.001 &	 0.777 &	 0.316 &	 0.781 &	 0.934 \\
+                    \end{tabular}
+                \end{table}
 
@@
             \hline
-            STL & -0.0977 & -0.600 & 0.0302 & 0.920 & NA \\
-            in-degree & -1.30$\times10^{-3}$ & -0.139 & -9.08$\times10^{-3}$ & 0.903 & 0.914 \\
+            STL & \textbf{-0.0977} & \textbf{-0.600} & \textbf{0.0302} & \textbf{0.920} & NA \\
+            in-degree & \textbf{-1.30$\times10^{-3}$} & \textbf{-0.139} & \textbf{-9.08$\times10^{-3}$} & 0.903 & 0.914 \\
         \end{tabular}
@@
 
-\section{Persistence and whole-network motif profiles} [[updated]]
+\section{Supplemental results: network motif profiles}
+
+    \subsection{Persistence and whole-network motif profiles}
 
     % Updated to use mean instead of all species
-    \begin{table}[hb!]
-        \caption{Coefficients ($\beta$) for models relating persistence to the proportion of a network profile made up by a focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
+    \begin{table}[ht!]
+        \caption{These results correspond to equation 2, main text.
+Coefficients ($\beta$) for models relating persistence to the proportion of a network profile made up by a focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction.
 Coefficients in \textbf{bold} are significant ($\alpha$\textless0.005).
@@
     \end{table}
-\clearpage
 
-\section{Motif profiles and global properties} [[updated]]
+    \subsection{Motif profiles and network size, connectance}
     % No update needed
@@
 
-   \begin{figure}[hb!]
-       \centering
-       \includegraphics[width=.75\textwidth]{figures/proportion_profile_dispersion.eps}
-       \caption{Dispersion of normalized motif profiles (i.e., proportions of each motif) within a network about the centroid for that combination of network size and connectance.
+       % \begin{figure}[hb!]
+       %     \centering
+       %     \includegraphics[width=.75\textwidth]{figures/proportion_profile_dispersion.eps}
+       %     \caption{Dispersion of normalized motif profiles (i.e., proportions of each motif) within a network about the centroid for that combination of network size and connectance.
 Each circle represents a single network; circle colour indicates network size.
 Scatter has been introduced about each level of connectance to increase visual clarity; the levels of connectance we consider are indicated by vertical black lines.}
-       \label{dispersion_normmotifs}
-    \end{figure}
+       %     \label{dispersion_normmotifs}
+       %  \end{figure}
 
@@
 \clearpage
-
-\section{Persistence and global network properties} [updated]
-
-    Connectance had a smaller, but significant, effect on consumer persistence than disturbance (Figs.~\ref{fig:lm_CS}, Table~\ref{tab:per_vs_SC}).
-    Network size had no effect on consumer persistence in these simulated networks.
-
-    % updated to use network means rather than all species
-    \begin{figure}[hb!]
-        \centering
-        \includegraphics[width=0.9\textwidth]{figures/persistence_vs_SC_lm.eps}
-        \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.
-The effects of network size and interactions between predictors were not significant.
-}
-        \label{fig:lm_CS}
-    \end{figure}
-
-    % \begin{figure}[hb!]
-    %     \centering
-    %   \includegraphics[width=\textwidth]{figures/heatmap_CS_BPcompare.eps}
-    %    \caption{Heat maps of average consumer persistence in networks with varying connectance (y-axis) and network size (x-axis) at two levels of disturbance to basal resources.
-Persistence decreases with darker color in the heat map.
-When basal species are not disturbed, all species experience a baseline extinction risk of $\pi = 0.1$.
-When basal species are highly disturbed, their baseline extinction risk rises to $\pi = 0.5$ while the baseline risk of consumers is not changed.}
-    %    \label{fig:heatmap_CS}
-    % \end{figure}
-
-    % Updated to use mean rather than all species.
-    \begin{table}[hb!]
-        \caption{Mean persistence of consumers within a network depends on connectance (C) and the level of disturbance to basal resources (D), but not network size (S) or interactions between properties.
-We show effect sizes ($\beta$) and $p$-values for each coefficient in a general linear model (fit using a binomial error distribution and logic link function).
-All predictors were centered and scaled before fitting the model.
-The level of disturbance corresponds to basal resources having a probability of extinction ranging between $0.1 - 0.5$ (see \emph{Methods}).
-McFadden's pseudo-$R^2$=0.966.
-        }
-        \label{tab:per_vs_SC}
-        \centering
-        \begin{tabular}{c|c c |}
-            Predictor & $\beta$ & $p$-value \\
-            \hline
-            Intercept & \textbf{0.326} & \textless0.001 \\
-            Size & -7.21$\times10^{-4}$  & 0.964 \\
-            Connectance & \textbf{-6.12$\times10^{-2}$}& \textless0.001 \\
-            Disturbance & \textbf{-0.596} & \textless0.001 \\
-            S:C & -4.48$\times10^{-3}$ & 0.777 \\
-            S:D & -1.63$\times10^{-2}$ & 0.316 \\
-            C:D & -4.52$\times10^{-3}$ & 0.781 \\
-            S:C:D & -1.35$\times10^{-3}$ & 0.934 \\
-        \end{tabular}
-    \end{table}
 

### 2023-04-26 code/stat_analysis/persistence_vs_SC_lm.py
combined all context LMs to a single fig

@@ -14,12 +14,19 @@ from PyGrace.Extensions.colorbar import SolidRectangle, ColorBar
 from PyGrace.axis import LINEAR_SCALE, LOGARITHMIC_SCALE
 from PyGrace.Styles.el import ElGraph, ElLogColorBar
 
-colors=ColorBrewerScheme('Spectral',n=13)  # The blue is very beautiful but maybe harder to see.
-colors.add_color(173,132,191,'Purple')
-colors.add_color(34,90,34,'Green')
-# colors.add_color(120,120,120,'grey')
-# colors.add_color(255,125,125,'lightish_red')
-# colors.add_color(200,200,200,'lightgrey')
+colors=ColorBrewerScheme('Greys',n=10)  
+# Match Anna's ggplot colours
+colors.add_color(158,101,184,'Rpurple')
+colors.add_color(145,148,215,'Rblue')
+colors.add_color(42,120,142,'Rgreen1')
+colors.add_color(34,168,132,'Rgreen2')
+colors.add_color(122,209,81,'Rgreen3')
+colors.add_color(253,231,37,'Ryellow')
+colors.add_color(254,224,139,'Ryellow2')
+colors.add_color(253,174,97,'Rorange1')
+colors.add_color(244,109,67,'Rorange2')
+colors.add_color(213,62,79,'Rred1')
+colors.add_color(158,1,66,'Rred2')
 
 
 scales={
@@ -57,55 +64,57 @@ Blms={'Intercept':5.777e-01,
       'Basal:Size:Connectance:Disturbance':-1.445e-03
 }
 
+def read_persfiles(Degfile,TLfile):
+  persdict={'Deg':{},'TL':{}}
+  f=open(Degfile,'r')
+  for line in f:
+    if line.split()[0]!='"Estimate"':
+      persdict['Deg'][line.split()[0][1:-1]]=((float(line.split()[1]),float(line.split()[2])))
+  f.close()
+
+  f=open(TLfile,'r')
+  for line in f:
+    if line.split()[0]!='"Estimate"':
+      persdict['TL'][line.split()[0][1:-1]]=((float(line.split()[1]),float(line.split()[2])))
+  f.close()
+
+  return persdict
 
 def format_graph(graph,simple):
   graph.yaxis.bar.linewidth=1
   graph.xaxis.bar.linewidth=1
   graph.frame.linewidth=1
 
-
   graph.world.ymin=0
   graph.world.ymax=1
   graph.yaxis.tick.major=.2
[…]

### 2023-04-26 code/stat_analysis/persistence_vs_SC_lm.py
better order

@@ -225,7 +225,7 @@ persdict=read_persfiles('persistence_vs_Deg_norandom.tsv','persistence_vs_TL.tsv
 
 grace=MultiPanelGrace(colors=colors)
 
-for simple in ['Size','Deg','Connectance','TL']:
+for simple in ['Connectance','Deg','Size','TL']:
   graph2=grace.add_graph(Panel)
   graph2=format_graph(graph2,simple)
   if simple in ['Size','Connectance']:

### 2023-04-26 manuscript/reorder_main_text.tex

@@
              \begin{aligned}
-            \Psi_{ink} & \approx \rho_{i} + \pi_{k} + \Sigma_{n} + \Zeta_{n} \\
-            & + \rho_{i}\pi_{k} + \rho_{i}\Sigma_{n} + \rho_{i}\Zeta_{n} + \pi_{k}\Sigma_{n} + \pi_{k}\Zeta_{n} + \Sigma_{n}\Zeta_{n} \\
-            & + \rho_{i}\pi_{k}\Sigma_{n} + \rho_{i}\pi_{k}\Zeta_{n} + \pi_{k}\Sigma_{n}\Zeta_{n} + \rho_{i}\pi_{k}\Sigma_{n}\Zeta_{n} ,
+            \Psi_{ink} & \approx \rho_{i} + \pi_{k} + \Sigma_{n} + \zeta_{n} \\
+            & + \rho_{i}\pi_{k} + \rho_{i}\Sigma_{n} + \rho_{i}\zeta_{n} + \pi_{k}\Sigma_{n} + \pi_{k}\zeta_{n} + \Sigma_{n}\zeta_{n} \\
+            & + \rho_{i}\pi_{k}\Sigma_{n} + \rho_{i}\pi_{k}\zeta_{n} + \pi_{k}\Sigma_{n}\zeta_{n} + \rho_{i}\pi_{k}\Sigma_{n}\zeta_{n} ,
             % S_{n}C_{n} + N_n,
@@
             \end{equation}
-            \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$, $\Sigma_{n}$ is the number of species in the network, and $\Zeta_{n}$ is the connectance of the network.
+            \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$, $\Sigma_{n}$ is the number of species in the network, and $\zeta_{n}$ is the connectance of the network.
             All predictors were centered and scaled before fitting the models.    
@@
 
+    \begin{figure}[ht!]
+        \centering
+        \includegraphics[width=0.9\textwidth]{figures/persistence_vs_SC_lm.eps}
+        \caption{\textbf{A,C} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.
+There was no significant relationship between mean persistence and network size.
+        \textbf{B} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
+        \textbf{D} Persistence of consumers decreased with increasing STL at all levels of disturbance.}
+        \label{fig:lm_CS}
+    \end{figure}
+
+    \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=.9\textwidth]{figures/participation_vs_SC.eps}
+        \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).
+We show relationships for the smallest and largest values of network size (S) and connectance (C).}
+        \label{fig:roles_vs_SC}
+    \end{figure}
+
     \begin{figure}[hb!]

### 2023-04-27 code/participation_vs_SC.py
Now motif participation vs. other junk is in one figure

@@ -59,25 +59,25 @@ def format_graph(graph,simple):
   graph.xaxis.bar.linewidth=1
   graph.frame.linewidth=1
 
-  graph.xaxis.label.configure(char_size=1,just=2,place='normal')
+  graph.xaxis.label.configure(char_size=.75,just=2,place='normal')
   graph.xaxis.tick.configure(onoff='on',minor_ticks=0,major_size=.4,place='both',major_linewidth=.5,minor_linewidth=1)
   graph.world.xmin=0
   if simple=='C':
     graph.world.xmax=.2
-    graph.xaxis.tick.major=.04
+    graph.xaxis.tick.major=.1
     graph.xaxis.label.text='Connectance'
-    graph.xaxis.ticklabel.configure(format='decimal',prec=2,char_size=.75)
+    graph.xaxis.ticklabel.configure(format='decimal',prec=1,char_size=.5)
   else:
-    graph.world.xmin=45
-    graph.world.xmax=105
-    graph.xaxis.tick.major=10
+    graph.world.xmin=50
+    graph.world.xmax=100
+    graph.xaxis.tick.major=25
     graph.xaxis.label.text='Network size'
-    graph.xaxis.ticklabel.configure(format='decimal',prec=0,char_size=.75)
+    graph.xaxis.ticklabel.configure(format='decimal',prec=0,char_size=.5)
 
   graph.world.ymin=0
   graph.world.ymax=.75
-  graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
-  graph.yaxis.label.configure(text='Mean persistence',char_size=1,just=2,place='normal')
+  graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.5)
+  graph.yaxis.label.configure(text='Mean persistence',char_size=.75,just=2,place='normal')
   graph.yaxis.tick.configure(major=.2,onoff='on',minor_ticks=0,major_size=.4,minor_size=.5,place='both',major_linewidth=.5,minor_linewidth=1)
 
   return graph
@@ -114,10 +114,10 @@ def populate_persgraph(graph,xprop,val,netprops):
     data.symbol.shape=0
     data.line.configure(linestyle=sty,linewdith=1.5,color=j)
 
-    if xprop=='C' and val==50:
-      data.legend=motif_names[motif]
+    # if xprop=='C' and val==50:
+      # data.legend=motif_names[motif]
 
[…]

### 2023-04-27 code/plot_roles_vs_TL.py

@@ -1,6 +1,7 @@
 import sys
 import os
 import math
+import participation_vs_SC as SC
 
 #Pygrace libraries
 from PyGrace.grace import Grace
@@ -77,34 +78,24 @@ def format_graph(graph,simple,roletype):
 
   if simple=='Deg':
     graph.world.xmax=100
-    graph.xaxis.tick.major=25
+    graph.xaxis.tick.major=50
     graph.xaxis.label.text='In-degree (number of prey)'
   elif simple=='TL':
     graph.world.xmax=5
-    graph.xaxis.tick.major=1
+    graph.xaxis.tick.major=2
     graph.xaxis.label.text='Trophic level (STL)'
 
   if roletype=='Prop':
     graph.world.ymax=.600000001
     graph.yaxis.tick.major=.2
-    graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
+    graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.5)
     graph.yaxis.label.text='Proportion of motif in role'
-  elif roletype=='Count':
-    graph.world.ymax=600
-    graph.yaxis.tick.major=200
-    graph.yaxis.ticklabel.configure(format='decimal',prec=0,char_size=.75)
-    graph.yaxis.label.text='Count of motif'
-  elif roletype=='persistence':
-    graph.world.ymax=1
-    graph.yaxis.tick.major=.2
-    graph.yaxis.ticklabel.configure(format='decimal',prec=1,char_size=.75)
-    graph.yaxis.label.text='Persistence'
 
 
-  graph.xaxis.ticklabel.configure(format='decimal',prec=0,char_size=.75)
-  graph.xaxis.label.configure(char_size=1,just=2,place='normal')
+  graph.xaxis.ticklabel.configure(format='decimal',prec=0,char_size=.5)
+  graph.xaxis.label.configure(char_size=.75,just=2,place='normal')
 
-  graph.yaxis.label.configure(char_size=1,just=2,place='normal')
+  graph.yaxis.label.configure(char_size=.75,just=2,place='normal')
   graph.xaxis.tick.configure(onoff='on',minor_ticks=0,major_size=.4,place='both',major_linewidth=.5,minor_linewidth=1)
[…]

### 2023-04-27 manuscript/reorder_main_text.tex

@@
         \centering
-        \includegraphics[width=0.9\textwidth]{figures/persistence_vs_SC_lm.eps}
-        \caption{\textbf{A,C} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.
+        \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
+        \caption{
+        \textbf{A} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
+        \textbf{B,E} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.
 There was no significant relationship between mean persistence and network size.
-        \textbf{B} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
+        \textbf{C} Mean persistence of consumers decreased with increasing proportions of omnivory in the network's motif profile and increased with increasing proportions of other motifs.
+Persistence also decreased significantly with increasing probability of basal species extinction but there was no significant interaction; we therefore show trends only for $\pi$=0.1.
         \textbf{D} Persistence of consumers decreased with increasing STL at all levels of disturbance.}
@@
 
-    \begin{figure}[hb!]
-        \centering
-        \includegraphics[width=\textwidth]{figures/persistence_motif_profiles.eps}
-        \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.
+    % \begin{figure}[hb!]
+    %     \centering
+    %     \includegraphics[width=\textwidth]{figures/persistence_motif_profiles.eps}
+    %     \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.
 Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportion of the apparent competition motifs.
@@
 Relationships are shown for the observed range of proportions for each motif.}
-        \label{fig:motif_profile_persistence}
-    \end{figure}
+    %     \label{fig:motif_profile_persistence}
+    % \end{figure}
 

### 2023-04-27 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
             In all four cases, the convergent model did not retain any random effects.
-
-            NEW VERSION:
-
-            To test whether a consumer's risk of extinction was associated with the frequency of each motif in its participation vector and whether this relationship varied with the strength of disturbance, network size, and/or network connectance, we fit four general linear models with binomial error distribution (GLMs; one per motif which can occur in acyclic networks):
-             \begin{equation}
-             \begin{aligned}
-            \Psi_{ink} & \approx \rho_{i} + \pi_{k} + \Sigma_{n} + \zeta_{n} \\
-            & + \rho_{i}\pi_{k} + \rho_{i}\Sigma_{n} + \rho_{i}\zeta_{n} + \pi_{k}\Sigma_{n} + \pi_{k}\zeta_{n} + \Sigma_{n}\zeta_{n} \\
-            & + \rho_{i}\pi_{k}\Sigma_{n} + \rho_{i}\pi_{k}\zeta_{n} + \pi_{k}\Sigma_{n}\zeta_{n} + \rho_{i}\pi_{k}\Sigma_{n}\zeta_{n} ,
-            % S_{n}C_{n} + N_n,
-            \label{propreq}
-            \end{aligned}
-            \end{equation}
-            \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$, $\Sigma_{n}$ is the number of species in the network, and $\zeta_{n}$ is the connectance of the network.
-            All predictors were centered and scaled before fitting the models.    
-            After fitting the full models described above, we removed any non-significant terms in a stepwise fashion, beginning with the four-way interaction and proceeding to three-way and two-way interactions.
-            If a smaller term was included in a significant larger term (e.g., a non-significant two-way interaction involving terms in a significant three-way interaction) then the smaller term was retained.
             We attempted to fit a PERMANOVA test relating a species' motif participation vector as a whole to its persistence, but the assumption of equal variance among groups was violated and we therefore do not report the results of this test.
@@
             These models included only the frequency of the focal motif ($\rho_{i}$ above) and again used a binomial error distrubution.
-            [[Statement about what this tells us or something else?]]
+            To aid in interpretation of these single-network models with respect to network size and connectance, we fit an additional set of GLMs including proportion of motif, strength of disturbance, network size, and connectance across all networks (\emph{Appendix S6}).
 
@@
 
+        Digging deeper and exploring how persistence varied with the proportion of a motif in the species' participation vector, disturbance, and overall network structure (size and connectance), we found that the relationship between the proportion of omnivory in the participation vector and probability of persistence depended strongly on the connectance of the network.
+        Higher proportions of omnivory were associated with lower probabilities of persistence when connectance was low, regardless of network size or strength of disturbance.
+        At moderate connectance, however, higher proportions of omnivory were associated with greater probabilities of persistence when disturbance was weak.
+        In the most-connected networks, higher proportions of omnivory were associated with greater probabilities of persistence at all levels of disturbance.
+
+        The relationships between other motifs and persistence were more consistent over different levels of connectance and network size.
+        Slopes of the motif proportion-persistence relationships for the apparent compeition and direct competition motifs were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{Fig 2}.
+        The relationship between proportion of the three-species chain in the participation vector and persistence was very similar across all levels of size and connectance.
+
         Considering each network separately ....
@@
 
-        % The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=645, $p$\textless0.001), although this result must be considered tentative due to non-homogeneous distribution of roles (\emph{Appendix S8}).
         As well as the frequency with which each species appears in each motif, we can use the total frequencies of each motif in the whole network (i.e., the network's motif profile) to describe its structure.
@@
         \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence.
-\textbf{A-B)} The proportion of omnivory increased with increasing degree while all other proportions decreased.
-The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.
-\textbf{C-D)} Persistence increased with increasing in-degree for low levels of disturbance to basal resources but decreased with increasing degree for high levels of disturbance to basal resources.
-        Persistence declined with inceasing STL for all levels of disturbance to basal resources.}
+\textbf{A)} The proportion of omnivory increased with increasing degree while all other proportions decreased.
+\textbf{B)} The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.
+\textbf{C-F)} The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).
+We show relationships for the smallest and largest values of network size (S) and connectance (C).
+        }
         \label{fig:motifs_vs_TL_and_deg}
@@
 
-    \begin{figure}[hb!]
-        \centering
-        \includegraphics[width=.9\textwidth]{figures/participation_vs_SC.eps}
-        \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).
+    % \begin{figure}[hb!]
+    %     \centering
+    %     \includegraphics[width=.9\textwidth]{figures/participation_vs_SC.eps}
+    %     \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).
 We show relationships for the smallest and largest values of network size (S) and connectance (C).}
-        \label{fig:roles_vs_SC}
-    \end{figure}
+    %     \label{fig:roles_vs_SC}
+    % \end{figure}
 
@@
 
-    \begin{figure}[hb!]
-        \centering
-        \includegraphics[width=\textwidth]{figures/motif_proportion_lms.eps}
-        \caption{The proportion of the omnivory in a network's motif role increased significantly as connectance increased, while the proportions of the other three motifs did not vary significantly with network size, connectance, or their interaction.}
-        \label{motif_proportion_lms}
-    \end{figure}
-
-    \begin{figure}[hb!]
-        \centering
-        \includegraphics[width=\textwidth]{figures/persistence_omnivory_detailpers.eps}
-        \label{omnivory_plus_SC}
-        \end{figure}
-
-    \begin{figure}[hb!]
-        \centering
-        \includegraphics[width=\textwidth]{figures/persistence_chain_detailpers.eps}
-        \label{chain_plus_SC}
-        \end{figure}
-
-    \begin{figure}[hb!]
-        \centering
-        \includegraphics[width=\textwidth]{figures/persistence_apparent_detailpers.eps}
-        \label{apparent_plus_SC}
-        \end{figure}
-
-    \begin{figure}[hb!]
-        \centering
-        \includegraphics[width=\textwidth]{figures/persistence_direct_detailpers.eps}
-        \label{direct_plus_SC}
-        \end{figure}
+    % \begin{figure}[hb!]
+    %     \centering
+    %     \includegraphics[width=\textwidth]{figures/motif_proportion_lms.eps}
+    %     \caption{The proportion of the omnivory in a network's motif role increased significantly as connectance increased, while the proportions of the other three motifs did not vary significantly with network size, connectance, or their interaction.}
+    %     \label{motif_proportion_lms}
+    % \end{figure}
 

### 2023-04-27 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 
-        \item Within-network motif-persistence relationships
+        \item Motif-persistence relationships vary with S,C
 
             \begin{itemize}
+                \item Describes GLMs summarising how the relationship between motif participation and persistence varies over S and C, to complement the within-netowrk results presented in the main text
                 \item Contains Fig. S4
@@
 
-    \begin{figure}[ht!]
-        \centering
-        \includegraphics[width=0.9\textwidth]{figures/persistence_vs_SC_lm.eps}
-        \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.}
-        \label{fig:lm_CS}
-    \end{figure}
+    % \begin{figure}[ht!]
+    %     \centering
+    %     \includegraphics[width=0.9\textwidth]{figures/persistence_vs_SC_lm.eps}
+    %     \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.}
+    %     \label{fig:lm_CS}
+    % \end{figure}
 
@@
 
-\section{Within-network motif-persistence relationships}
+\section{Motif-persistence relationships vary with S,C}
 
-    We compared the outcomes from linear mixed-effect models (GLMMs) relating persistence to motif roles across all networks (Fig.
-3 and Table 1, \emph{Main Text}) with similar general linear regressions (GLRs) made on each single network separately.
-    The main trends are similar in both analyses, but different-sized networks showed slightly different results for some motifs when considered individually (Figure~\ref{fig:density_prop_S}).
-    For the three-species chain and omnivory motifs, a significant interaction between disturbance and the effect of the focal motif means that having a larger proportion of the role be made up of the chain or omnivory motif is associated with lower persistence at high levels of disturbance (Fig.
-3 and Table 1, \emph{Main Text}).
-    The same pattern holds overall for GLRs fit within a network  (Figure~\ref{fig:density_prop_S}): the fraction of negative slopes for the proportion of three-species chain and omnivory motifs increases with increasing disturbance.
-    This trend is most pronounced in large networks. (bottom panel, Figure~\ref{fig:density_prop_S}). 
-    % Besides the decrease in fraction of positive slopes with increasing disturbance, the omnivory motif showed a higher overall fraction of positive slopes in comparison to the three-species chain motif, corresponding to the positive main effect of omnivory (Fig.
-3 and Table 1, \emph{Main Text}).
+    In case the within-network trends with respect to network size and connectance are easier to understand after summarising across all networks, we fit four general linear models with binomial error distribution (GLMs; one per motif which can occur in acyclic networks):
+             \begin{equation}
+             \begin{aligned}
+            \Psi_{ink} & \approx \rho_{i} + \pi_{k} + \Sigma_{n} + \zeta_{n} \\
+            & + \rho_{i}\pi_{k} + \rho_{i}\Sigma_{n} + \rho_{i}\zeta_{n} + \pi_{k}\Sigma_{n} + \pi_{k}\zeta_{n} + \Sigma_{n}\zeta_{n} \\
+            & + \rho_{i}\pi_{k}\Sigma_{n} + \rho_{i}\pi_{k}\zeta_{n} + \pi_{k}\Sigma_{n}\zeta_{n} + \rho_{i}\pi_{k}\Sigma_{n}\zeta_{n} ,
+            % S_{n}C_{n} + N_n,
+            \label{propreq}
+            \end{aligned}
+            \end{equation}
+            \noindent where $\Psi_{ink}$ is the persistence of species $i$, belonging to network $n$, during disturbance level $k$, $\rho_{i}$ is the proportion of the role of species $i$ that is made up by the focal motif, $\pi_k$ is the probability of extinction for a basal resource in disturbance level $k$, $\Sigma_{n}$ is the number of species in the network, and $\zeta_{n}$ is the connectance of the network.
+            All predictors were centered and scaled before fitting the models.    
+            After fitting the full models described above, we removed any non-significant terms in a stepwise fashion, beginning with the four-way interaction and proceeding to three-way and two-way interactions.
+            If a smaller term was included in a significant larger term (e.g., a non-significant two-way interaction involving terms in a significant three-way interaction) then the smaller term was retained.
 
@@
         \centering
-            \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_S_allC.eps}
-            \caption{Here we show the density (y-axis) of slopes (x-axis) for regressions of probability of persistence against proportion of different motifs for all simulated webs of all connectances - a visualization of how an increased proportion of each motif (different colored lines) affects the probability of persistence for consumer species.
-Columns show the result for various disturbances on the basal level, from $\pi = 0.1$ (left) to $\pi = 0.5$ (right).
-Rows show various network sizes.
-The dotted, vertical lines indicate zero on the x-axis.
-A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increase persistence.
-The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot.
-Line color corresponds to the motif.
-}
-            \label{fig:density_prop_S}
+        \includegraphics[width=\textwidth]{figures/persistence_omnivory_detailpers.eps}
+        \caption{}
+        \label{omnivory_plus_SC}
         \end{figure}
 
-    For the apparent and direct competition motifs, the LMMs include a positive interaction between disturbance and the proportion of the motif in a species' role (Fig.
-3 and Table 1; \emph{Main Text}).
-    This means that having more of the role made up of these two motifs is increasingly beneficial at higher levels of disturbance, a trend which is also evident from the increasing fraction of positive slopes with increasing disturbance based on LRs of each network at each level of disturbance.
-% (Figure~\ref{fig:density_prop_S}).
-    As with the trends for omnivory and three-species chains, these increases are strongest in large networks. % (bottom row, Figure~\ref{fig:density_prop_S}).
-    % For direct competition, the fraction of slopes are similar across disturbance levels.
-Although this interaction is not positive when accounting for network size, the fraction of positive slopes are throughout high.
-    The increase in the proportion of positive slopes was much stronger for apparent competition than direct competition, consistent with the smaller (but still significant) interaction term for between direct competition in the LMMs.
-    A high proportion of direct competition in a species' role tended to be associated with greater probability of persistence across levels of disturbance and network size.
-    % In conclusion, species richness does not considerably alter the trends visible in the LMMs.
-High species richness more clearly display an interaction between motif proportion, persistences and disturbance levels, while this effect is smaller with low species richness.
+    \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=\textwidth]{figures/persistence_chain_detailpers.eps}
+        \caption{}
+        \label{chain_plus_SC}
+        \end{figure}
+
+    \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=\textwidth]{figures/persistence_apparent_detailpers.eps}
+        \caption{}
+        \label{apparent_plus_SC}
+        \end{figure}
+
+    \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=\textwidth]{figures/persistence_direct_detailpers.eps}
+        \caption{}
+        \label{direct_plus_SC}
+        \end{figure}
+
 \clearpage
@@
         Specifically, the frequency of the omnivory motif increased in more-connected networks (Table~\ref{network_prop_lms}) while the frequencies of other motifs were not significantly related to network size or connectance.
+
+    \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=\textwidth]{figures/motif_proportion_lms.eps}
+        \caption{The proportion of the omnivory in a network's motif role increased significantly as connectance increased, while the proportions of the other three motifs did not vary significantly with network size, connectance, or their interaction.}
+        \label{motif_proportion_lms}
+    \end{figure}
 

### 2023-04-27 manuscript/reorder_main_text.tex
one spelling error

@@
         The relationships between other motifs and persistence were more consistent over different levels of connectance and network size.
-        Slopes of the motif proportion-persistence relationships for the apparent compeition and direct competition motifs were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{Fig 2}.
+        Slopes of the motif proportion-persistence relationships for the apparent competition and direct competition motifs were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{Fig 2}.
         The relationship between proportion of the three-species chain in the participation vector and persistence was very similar across all levels of size and connectance.

### 2023-04-27 code/stat_analysis/persistence_omnivory_SCDist.py

@@ -112,6 +112,7 @@ def populate_persgraph(graph,coefs,S,C):
 
   j=13
   for dist in [0.1,0.18,0.26,0.34,0.42,0.5]:
+    print(dist,j)
   # for dist in [0.5,0.42,0.34,0.26,0.18,0.1]:
     sty=1
     scalDist=(dist-scales['Disturbance'][0])/scales['Disturbance'][1]

### 2023-05-01 manuscript/reorder_main_text.tex
Updates from Overleaf

--- 2023-04-27 manuscript/reorder_main_text.tex
@@
 
-        Digging deeper and exploring how persistence varied with the proportion of a motif in the species' participation vector, disturbance, and overall network structure (size and connectance), we found that the relationship between the proportion of omnivory in the participation vector and probability of persistence depended strongly on the connectance of the network.
+        In depth exploration of how persistence varied with the proportion of a motif in the species' participation vector, disturbance, and overall network structure (size and connectance), showed that the relationship between the proportion of omnivory in the participation vector and probability of persistence depended strongly on the connectance of the network Fig.~\ref{Fig 2}.
         Higher proportions of omnivory were associated with lower probabilities of persistence when connectance was low, regardless of network size or strength of disturbance.
@@
 
-        The relationships between other motifs and persistence were more consistent over different levels of connectance and network size.
-        Slopes of the motif proportion-persistence relationships for the apparent compeition and direct competition motifs were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{Fig 2}.
+        The relationships between the proportion of other motifs and persistence were more consistent over different levels of connectance and network size.
+        Slopes of the motif proportion-persistence relationships for the apparent competition and direct competition motifs were steepest in large and less-connected networks, but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{Fig 2}.
         The relationship between proportion of the three-species chain in the participation vector and persistence was very similar across all levels of size and connectance.
+
+         The effects of network size were overall limited Fig.~\ref{Fig 2}.
 
@@
 
-    \subsection*{Species persistence and other network properties}
+% Total frequencies of motifs in the whole network
+    \subsection*{Species persistence and network motif profile}
 
@@
 
+        % FIGURE 1
         \begin{figure}[hb!]
         \centering
-        \includegraphics[width=1.0\textwidth]{figures/concept_fig_ver2.eps}
+        \includegraphics[width=1.0\textwidth]{figures/concept_fig_ver2_edit.eps}
         \caption{Conceptual figure of a toy food web.
@@
 
+        % FIGURE 2
         \begin{figure}[hb!]
@@
 
+        % FIGURE 3
     \begin{figure}[hb!]
@@
 
+        % FIGURE 4
     \begin{figure}[ht!]
@@
 
+    % FIGURE 5
     \begin{figure}[hb!]
     \centering
-        \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.eps}
+        \includegraphics[width=\textwidth]{manuscript/figures/Fig4.pdf}
         \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.

### 2023-05-01 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 
-    \begin{figure}[hb!]
+    \begin{figure}[ht!]
         \centering
         \includegraphics[width=\textwidth]{figures/persistence_omnivory_detailpers.eps}
-        \caption{}
+        \caption{The relationship between the proportion of omnivory in the participation vector and probability of persistence depended strongly on the connectance of the network.
+        Higher proportions of omnivory were associated with lower probabilities of persistence when connectance was low, regardless of network size or strength of disturbance.
+        At moderate connectance, however, higher proportions of omnivory were associated with greater probabilities of persistence when disturbance was weak.
+        In the most-connected networks, higher proportions of omnivory were associated with greater probabilities of persistence at all levels of disturbance.}
         \label{omnivory_plus_SC}
@@
         \includegraphics[width=\textwidth]{figures/persistence_chain_detailpers.eps}
-        \caption{}
+        \caption{The relationship between proportion of the three-species chain in the participation vector and persistence was very similar across all levels of size and connectance.}
         \label{chain_plus_SC}
@@
         \includegraphics[width=\textwidth]{figures/persistence_apparent_detailpers.eps}
-        \caption{}
+        \caption{Slopes of the motif proportion-persistence relationships for the apparent competition motif were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance.}
         \label{apparent_plus_SC}
@@
         \includegraphics[width=\textwidth]{figures/persistence_direct_detailpers.eps}
-        \caption{}
+        \caption{Slopes of the motif proportion-persistence relationships for the direct competition motif were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance.}
         \label{direct_plus_SC}

### 2023-05-01 manuscript/JAE_reply.tex
old reply changes

@@
 
+    \textbf{R:}
+
     \subsection*{Comments}
@@
 
-            \textbf{R:} [[This one might be have to be Anna's task - I can't get the ggplot2 code to run.]]
+            \textbf{R:} We have reorganised the figure as suggested and, ourselves, now find it much clearer.
+We thank the Reviewer for the suggestion and hope that they also now find it easier to interpret.
+As an additional aid to understanding these results, we have added additional figures in the SI showing trends extracted from a model including proportion of motif, disturbance strength, network size, and connectance (across all networks within a size and connectance class).
+These can be compared with the peaks of the distributions of the within-network slopes to more easily understand how to these values relate to persistence.
 

### 2023-05-01 manuscript/reorder_main_text.tex

--- 2023-04-27 manuscript/reorder_main_text.tex
@@
         The relationships between other motifs and persistence were more consistent over different levels of connectance and network size.
-        Slopes of the motif proportion-persistence relationships for the apparent compeition and direct competition motifs were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{Fig 2}.
+        Slopes of the motif proportion-persistence relationships for the apparent competition and direct competition motifs were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{fig:prop_lmer_all}.
         The relationship between proportion of the three-species chain in the participation vector and persistence was very similar across all levels of size and connectance.
@@
         The average probability of persistence of all consumers in the network varied with the frequencies of different motifs in this profile.
-        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:motif_profile_persistence}, Table S4, \emph{Appendix S8}).
+        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.
+S~\ref{fig:lm_CS}, Table S4, \emph{Appendix S8}).
         There was no significant interaction between the proportion of any motif in the network profile and level of disturbance (Table S4), \emph{Appendix S8}).
@@
         Both the motif profile of a network and species' motif participation were related to other measures of network structure.
-        Most simply, the proportion of the omnivory motif in the network's motif role increased with increasing connectance (Fig.~\ref{motif_proportion_lms}; Table S5, \emph{Appendix S9}).
+        Most simply, the proportion of the omnivory motif in the network's motif role increased with increasing connectance (Fig. S6, Table S5; \emph{Appendix S9}).
         The proportions of other motifs did not vary significantly with connectance and none were significantly related to network size.
@@
         \centering
-        \includegraphics[width=1.0\textwidth]{figures/concept_fig_ver2.eps}
+        \includegraphics[width=1.0\textwidth]{figures/concept_fig_ver2_edit.eps}
         \caption{Conceptual figure of a toy food web.
@@
 For example, a disturbance to the bottom species in a three-species chain will have a direct effect on the middle species but only indirectly affect the top species while a disturbance to the bottom species in an omnivory motif will have a direct effect on the middle species and direct and indirect effects on the top species.
-        Note that the nodes participate in more motifs than the ones highlighted.
-\textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than three times more frequently (see inserted table for counts of participation in all four motifs).
+        Note that the nodes participate in more motifs than the ones highlighted and that a species' motif participation vector is the count of all motifs of each type in which the species appears.
+\textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than three times more frequently (see inserted table for counts of participation in all four motifs, giving the motif participation vector in parentheses).
         Despite having equal degree and trophic level, species A may be at greater risk of going extinct than species B because of its dependence on specialist prey species.

### 2023-05-02 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-        The relationships between other motifs and persistence were more consistent over different levels of connectance and network size.
-        Slopes of the motif proportion-persistence relationships for the apparent competition and direct competition motifs were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{fig:prop_lmer_all}.
+        The relationships between the proportion of other motifs and persistence were more consistent over different levels of connectance and network size.
+        Slopes of the motif proportion-persistence relationships for the apparent competition and direct competition motifs were steepest in large and less-connected networks, but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{fig:prop_lmer_all}.
         The relationship between proportion of the three-species chain in the participation vector and persistence was very similar across all levels of size and connectance.
+        The effects of network size were overall limited Fig.~\ref{fig:prop_lmer_all}.
 
@@
 
-    \subsection*{Species persistence and other network properties}
+    \subsection*{Species persistence and network motif profile}
 
@@
     \centering
-        \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.eps}
+        \includegraphics[width=\textwidth]{manuscript/figures/Fig4.pdf}
         \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.

### 2023-05-03 code/stat_analysis/persistence_vs_SC_lm.py
reordered to match text, updated motif colors to be distinct from disturbance

@@ -28,7 +28,15 @@ colors.add_color(244,109,67,'Rorange2')
 colors.add_color(213,62,79,'Rred1')
 colors.add_color(158,1,66,'Rred2')
 
-
+colors.add_color(228,26,28,'omnivory')
+colors.add_color(55,126,184,'direct') 
+colors.add_color(166,86,40,'apparent') 
+colors.add_color(152,78,163,'chain') 
+        # 5 E 255 127 0 
+        # 6 F 255 255 51  
+        # 7 G 166 86  40  
+        # 8 H 247 129 191 
+        # 9 I 153 153 153 
 scales={
   'Size':(79.1882,16.62225),
   'Connectance':(0.1044983,0.05559927),
@@ -213,15 +221,15 @@ def populate_specgraph(graph,persdict,simple):
     data.symbol.shape=0
     data.line.configure(linestyle=1,linewidth=3.5,color=j)
 
-    if simple=='TL':
+    if simple=='Deg':
       data.legend=str(basal_p)
     j+=1
 
-  if simple=='TL':
-    graph.add_drawing_object(DrawText,text='Basal species',x=13.25,y=0.9,loctype='world',char_size=.5) 
-    graph.add_drawing_object(DrawText,text='extinction',x=13.25,y=0.82,loctype='world',char_size=.5) 
-    graph.add_drawing_object(DrawText,text='probability',x=13.25,y=0.74,loctype='world',char_size=.5) 
-    graph.legend.configure(char_size=.5,loc=(13.25,0.72),loctype='world',box_linestyle=0,fill_pattern=0)
+  if simple=='Deg':
+    graph.add_drawing_object(DrawText,text='Basal species',x=250,y=0.9,loctype='world',char_size=.5) 
+    graph.add_drawing_object(DrawText,text='extinction',x=250,y=0.82,loctype='world',char_size=.5) 
+    graph.add_drawing_object(DrawText,text='probability',x=250,y=0.74,loctype='world',char_size=.5) 
+    graph.legend.configure(char_size=.5,loc=(250,0.72),loctype='world',box_linestyle=0,fill_pattern=0)
 
 
   return graph
@@ -255,16 +263,16 @@ def populate_motifgraph(graph,motif):
 
   for dist in [0.1]:
     if motif=='apparent':
-      j='Rpurple'
+      # j='Rpurple'
       sty=2
     elif motif=='chain':
-      j='Rblue'
+      # j='Rblue'
[…]

### 2023-05-03 manuscript/reorder_main_text.tex

@@
         A species' probability of persistence was significantly related to its motif participation (Table S1, \emph{Appendix S4}).
-        The probability of persistence tended to increase with increasing frequency of the direct competition motif and decrease with increasing frequency of the omnivory motif, across all levels of disturbance.
-        The relationships between participation in the apparent competition and three-species chain motifs, however, varied with the strength of disturbance (Fig.~\ref{fig:prop_lmer_all}).
-        Higher frequency of participation in the apparent competition motif was associated with lower probability of persistence at low levels of disturbance and higher probability of persistence at high levels of disturbance.
-        Participation in the three-species chain motif showed the opposite trends.
+        Across all levels of disturbance, species were more likely to persist if they participated more frequently in the direct competition motif and less frequently in the omnivory motif (Fig.~\ref{fig:prop_lmer_all}).
+        The relationships between probability of persistence and participation in the other two motifs varied with the strength of disturbance.
+        At low levels of disturbance, species were more likely to persist if they participated more frequently in the three-species chain motif and less frequently in the apparent competition motif.
+        These trends were reversed at high levels of disturbance.
 
-        In depth exploration of how persistence varied with the proportion of a motif in the species' participation vector, disturbance, and overall network structure (size and connectance), showed that the relationship between the proportion of omnivory in the participation vector and probability of persistence depended strongly on the connectance of the network Fig.~\ref{Fig 2}.
+        Within these overall trends, there was substantial variation in the relationships between persistence and motif participation between networks with different overall network structures (Fig.~\ref{fig:density_prop}).
+        Networks with high and low connectances had different distributions of persistence-motif participation relationships, while network size had a much smaller effect (\emph{Appendix SX}).
+        The relationship between participation in the omnivory motif and probability of persistence was especially variable.
         Higher proportions of omnivory were associated with lower probabilities of persistence when connectance was low, regardless of network size or strength of disturbance.
@@
 
-        The relationships between the proportion of other motifs and persistence were more consistent over different levels of connectance and network size.
-        Slopes of the motif proportion-persistence relationships for the apparent competition and direct competition motifs were steepest in large and less-connected networks, but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{fig:prop_lmer_all}.
-        The relationship between proportion of the three-species chain in the participation vector and persistence was very similar across all levels of size and connectance.
-        The effects of network size were overall limited Fig.~\ref{fig:prop_lmer_all}.
+        The relationships between persistence and participation in the other motifs were more consistent over different levels of connectance and network size.
+        The persistence-motif participation relationships were strongest in networks with low connectance (Fig.~\ref{fig:density_prop}).
+        For the apparent competition and three-species chain motifs, the direction of these relationships depended much more strongly on the strength of disturbance than on connectance.
+        For the direct competition motif, persistence increased with increasing participation in the motif across all disturbance levels in networks with low and medium connectance.
+        In networks with high connectance, however, the distribution of slopes for this relationship was centered around zero at all levels of disturbance.
 
-        Considering each network separately ....
+    \subsection*{Persistence and other network properties}
 
-    \subsection*{Species persistence and network motif profile}
+        To understand the relationships between motif participation and consumers' probabilities of persistence, it is helpful to understand how persistence varies with other aspects of network structure.
+        For example, consumers with higher STL were less likely to persist (Fig.~\ref{fig:motifs_vs_TL_and_deg}, Table S3, \emph{Appendix S6}).
+        The relationship between in-degree and persistence depended on the strength of disturbance: higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.
 
-        As well as the frequency with which each species appears in each motif, we can use the total frequencies of each motif in the whole network (i.e., the network's motif profile) to describe its structure.
+        The average persistence of all consumers was lower in networks with higher connectance, but did not vary significantly with network size (Table S6, \emph{Appendix S10}).
         The average probability of persistence of all consumers in the network varied with the frequencies of different motifs in this profile.
@@
         There was no significant interaction between the proportion of any motif in the network profile and level of disturbance (Table S4), \emph{Appendix S8}).
-
-        To provide context for the relationships between consumer persistence and motif participation or motif profile, we also tested whether persistence was related to more commonly used network properties.
-        Average consumer persistence was lower in networks with higher connectance, but was not significantly related to network size (Table S6, \emph{Appendix S10}).
-        As expected, a consumer's persistence tended to decrease with increasing STL (Fig.~\ref{fig:motifs_vs_TL_and_deg}, Table S3, \emph{Appendix S6}).
-        The relationship between in-degree and persistence depended on the strength of disturbance: higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.
 
@@
     \centering
-        \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
-        \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence.
-\textbf{A)} The proportion of omnivory increased with increasing degree while all other proportions decreased.
-\textbf{B)} The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.
-\textbf{C-F)} The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).
-We show relationships for the smallest and largest values of network size (S) and connectance (C).
+        \includegraphics[width=\textwidth]{manuscript/figures/Fig4.pdf}
+        \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
+Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
+Rows show various levels of connectance.
+The dotted, vertical lines indicate zero on the x-axis.
+A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increases persistence.
+The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend).
 }
-        \label{fig:motifs_vs_TL_and_deg}
+        \label{fig:density_prop}
     \end{figure}
@@
 
+    \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
+        \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence.
+\textbf{A)} The proportion of omnivory increased with increasing degree while all other proportions decreased.
+\textbf{B)} The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.
+\textbf{C-F)} The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).
+We show relationships for the smallest and largest values of network size (S) and connectance (C).
+        }
+        \label{fig:motifs_vs_TL_and_deg}
+    \end{figure}
+
     % \begin{figure}[hb!]
@@
     % \end{figure}
-
-    \begin{figure}[hb!]
-    \centering
-        \includegraphics[width=\textwidth]{manuscript/figures/Fig4.pdf}
-        \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
-Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
-Rows show various levels of connectance.
-The dotted, vertical lines indicate zero on the x-axis.
-A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increases persistence.
-The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend).
-}
-        \label{fig:density_prop_C}
-    \end{figure}
 

### 2023-05-03 code/participation_vs_SC.py
some figs updated

@@ -28,6 +28,11 @@ colors.add_color(244,109,67,'Rorange2')
 colors.add_color(213,62,79,'Rred1')
 colors.add_color(158,1,66,'Rred2')
 
+colors.add_color(228,26,28,'omnivory')
+colors.add_color(55,126,184,'direct') 
+colors.add_color(255,127,0,'apparent') 
+colors.add_color(152,78,163,'chain') 
+
 mot_ranges={
   'omnivory':(0,0.54917),
   'apparent':(0,1),
@@ -42,6 +47,7 @@ def read_datafile(datafile):
   netprops={}
   f=open(datafile,'r')
   for line in f:
+    print(line.split())
     if line.split()[0]!='"V1"':
       motif=line.split()[1][1:-1]
       intercept=float(line.split()[2][1:-1])
@@ -82,7 +88,7 @@ def format_graph(graph,simple):
 
   return graph
 
-def populate_persgraph(graph,xprop,val,netprops):
+def populate_persgraph(graph,xprop,val,netprops,netSE):
   print xprop, val
   for motif in netprops:
     if motif=='Apparent':
@@ -98,21 +104,42 @@ def populate_persgraph(graph,xprop,val,netprops):
       j=18
       sty=1
     betas=netprops[motif]
+    SEs=netSE[motif]
     dats=[]
+    lower=[]
+    upper=[]
+
     if xprop=='S':
       for x in [50,60,70,80,90,100]:
         y=betas[0]+x*betas[1]+val*betas[2]+x*val*betas[3]
         logity=math.exp(y)/(1+math.exp(y))
         dats.append((x,logity))
+        lowy=betas[0]-SEs[0]+x*(betas[1]-SEs[1])+val*(betas[2]-SEs[2])+x*val*(betas[3]-SEs[3])
+
+        logitlow=math.exp(lowy)/(1+math.exp(lowy))
+        lower.append((x,logitlow))
+
+        upy=betas[0]+SEs[0]+x*(betas[1]+SEs[1])+val*(betas[2]+SEs[2])+x*val*(betas[3]+SEs[3])
+        logitup=math.exp(upy)/(1+math.exp(upy))
+        upper.append((x,logitup))
+
     elif xprop=='C':
       for x in [0.02,0.06,0.1,0.14,0.18]:
         y=betas[0]+x*betas[2]+val*betas[1]+x*val*betas[3]      
         logity=math.exp(y)/(1+math.exp(y))
         dats.append((x,logity))
 
[…]

### 2023-05-03 code/plot_roles_vs_TL.py

@@ -29,9 +29,10 @@ colors.add_color(244,109,67,'Rorange2')
 colors.add_color(213,62,79,'Rred1')
 colors.add_color(158,1,66,'Rred2')
 
-# colors.add_color(120,120,120,'grey')
-# colors.add_color(255,125,125,'lightish_red')
-# colors.add_color(200,200,200,'lightgrey')
+colors.add_color(228,26,28,'omnivory')
+colors.add_color(55,126,184,'direct') 
+colors.add_color(255,127,0,'apparent') 
+colors.add_color(152,78,163,'chain') 
 
 def read_lmfile(lmfile):
   lmdict={'Deg':{'Count':{},'Prop':{}},'TL':{'Count':{},'Prop':{}}}
@@ -104,16 +105,12 @@ def format_graph(graph,simple,roletype):
 def populate_graph(graph,minidict,simple,roletype):
   for motif in ['Apparent','Chain','Omnivory','Direct']:
     if motif=='Apparent':
-      j=12
       sty=2
     elif motif=='Chain':
-      j=14
       sty=3
     elif motif=='Direct':
-      j=15
       sty=5
     else:
-      j=18
       sty=1
     dats=[]
     lower=[]
@@ -129,17 +126,17 @@ def populate_graph(graph,minidict,simple,roletype):
       logitup=math.exp(upy)/(1+math.exp(upy))
       upper.append((x,logitup))
 
-    upply=graph.add_dataset(upper)
-    upply.symbol.shape=0
-    upply.line.configure(linestyle=sty,linewidth=1,color=j)
+    # upply=graph.add_dataset(upper)
+    # upply.symbol.shape=0
+    # upply.line.configure(linestyle=sty,linewidth=1,color=motif.lower())
 
-    lowly=graph.add_dataset(lower)
-    lowly.symbol.shape=0
-    lowly.line.configure(linestyle=sty,linewidth=1,color=j)
+    # lowly=graph.add_dataset(lower)
+    # lowly.symbol.shape=0
+    # lowly.line.configure(linestyle=sty,linewidth=1,color=motif.lower())
 
     data=graph.add_dataset(dats)
     data.symbol.shape=0
-    data.line.configure(linestyle=sty,linewidth=2.5,color=j)
+    data.line.configure(linestyle=sty,linewidth=3.5,color=motif.lower())
 
[…]

### 2023-05-03 code/stat_analysis/persistence_vs_SC_lm.py
orange instead of brown

@@ -30,7 +30,7 @@ colors.add_color(158,1,66,'Rred2')
 
 colors.add_color(228,26,28,'omnivory')
 colors.add_color(55,126,184,'direct') 
-colors.add_color(166,86,40,'apparent') 
+colors.add_color(255,127,0,'apparent') 
 colors.add_color(152,78,163,'chain') 
         # 5 E 255 127 0 
         # 6 F 255 255 51  

### 2023-05-03 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
         To understand the relationships between motif participation and consumers' probabilities of persistence, it is helpful to understand how persistence varies with other aspects of network structure.
-        For example, consumers with higher STL were less likely to persist (Fig.~\ref{fig:motifs_vs_TL_and_deg}, Table S3, \emph{Appendix S6}).
+        Consumers with higher STL were less likely to persist (Fig.~\ref{fig:motifs_vs_TL_and_deg}, Table S3, \emph{Appendix S6}).
         The relationship between in-degree and persistence depended on the strength of disturbance: higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.
@@
 
-        Both the motif profile of a network and species' motif participation were related to other measures of network structure.
-        Most simply, the proportion of the omnivory motif in the network's motif role increased with increasing connectance (Fig. S6, Table S5; \emph{Appendix S9}).
-        The proportions of other motifs did not vary significantly with connectance and none were significantly related to network size.
-
-        A species' motif participation varied with network size, connectance, and their interaction. 
-        The proportion of the omnivory motif generally increased with increasing connectance and network size, but decreased with increasing network size in the most-connected networks (Fig.
-S2 and Table S2, \emph{Appendix S5}).
-        The proportion of apparent competition decreased strongly with increasing connectance and more weakly (but still significantly) with increasing network size.
-        The proportion of the three-species chain showed similar trends to apparent competition, while the proportion of direct competition showed very little change over the range of network size and connectance we consider (though the effects of connectance and the interaction between connectance and network size were significant).
-
-        Species' motif participation also varied with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
+        A species' motif participation varied with its position in the network and with overall network structure (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
         Species with higher in-degree tended to have higher frequencies of participation in the omnivory motif and lower proportions of the other three motifs.
         Species with higher trophic levels tended to have higher proportions of apparent competition and three-species chains and lower proportions of omnivory and direct competition.
-        A species' persistence varied with both species in-degree and trophic level, where the most clear result is decreased persistence with increased trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix S6}).
 
-        % \subsubsection*{Consistency in effects of motif participation on species persistence}
+        The frequency of participation in the omnivory motif generally increased with increasing connectance and network size, but decreased with increasing network size in the most-connected networks (Fig.
+S2 and Table S2, \emph{Appendix S5}).
+        The frequencies of participation in the apparent competition and three-species chain motifs decreased strongly with increasing connectance but showed little change with changing network size.
+        The frequency of participation in the direct competition motif decreased with increasing connectance in small networks and increased with increasing network size in highly-connected networks, but showed little change with connectance in large networks or with network size in less-connected networks.
 
-        %     When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally showed more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
-        %     However, this trend varied at different levels of disturbance, especially for the direct competition and omnivory motifs.
-        %     At low and medium connectances there was little change in the distribution of slopes for direct competition across all levels of disturbance. 
-        %     At high connectances, however, the proportion of positive slopes was both lower overall and decreased with increasing disturbance. 
-        %     For the omnivory motif these trends were reversed. 
-        %     Moreover, the omnivory motif has a particularly broad distribution of slopes (i.e., an inconsistent relationship to persistence) when network connectance was low. 
-        %     Network size had a much smaller effect than connectance (Fig. S4; \emph{Appendix S7}).
+        The network's motif profile also varied with network connectance.
+        Specifically, the proportion of the profile made up by the omnivory motif increased significantly  with increasing connectance while the proportions of the other three motifs decreased (though not significantly).
+        No proportions of motifs in the profile were significantly related to network size.
 
@@
         \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
-        \caption{
-        \textbf{A} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
-        \textbf{B,E} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.
+        \caption{The probabilities of persistence for a particular consumer (\textbf{A, B}) and the average probability of persistence across all consumers (\textbf{C-E}) varied with other elements of network structure as well as with motif participation.
+        \textbf{A} Persistence of consumers decreased with increasing STL at all levels of disturbance.
+        \textbf{B} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
+        \textbf{C,D} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.
 There was no significant relationship between mean persistence and network size.
-        \textbf{C} Mean persistence of consumers decreased with increasing proportions of omnivory in the network's motif profile and increased with increasing proportions of other motifs.
-Persistence also decreased significantly with increasing probability of basal species extinction but there was no significant interaction; we therefore show trends only for $\pi$=0.1.
-        \textbf{D} Persistence of consumers decreased with increasing STL at all levels of disturbance.}
+        \textbf{E} Mean persistence of consumers decreased with increasing proportions of omnivory in the network's motif profile and increased with increasing proportions of other motifs.
+Persistence also decreased significantly with increasing probability of basal species extinction but there was no significant interaction; we therefore show trends only for $\pi$=0.1.}
         \label{fig:lm_CS}

### 2023-05-03 code/participation_vs_SC.py
fixed line weightss

@@ -139,7 +139,7 @@ def populate_persgraph(graph,xprop,val,netprops,netSE):
 
     data=graph.add_dataset(dats)
     data.symbol.shape=0
-    data.line.configure(linestyle=sty,linewdith=3.5,color=motif.lower())
+    data.line.configure(linestyle=sty,linewidth=3.5,color=motif.lower())
 
     # if xprop=='C' and val==50:
       # data.legend=motif_names[motif]

### 2023-05-03 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 \textbf{B)} The proportions of omnivory and direct competition decreased with increasing STL while the proportions of apparent competition and three-species chains increased.
-\textbf{C-F)} The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).
-We show relationships for the smallest and largest values of network size (S) and connectance (C).
+\textbf{C-D)} The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition, where the interaction was not significant).
+To cover the largest range of slopes, we show relationships over connectance for the smallest networks (\textbf{C}) and over network size for the most-connected networks (\textbf{D}).
+See also Fig.
+SX.
 }

### 2023-05-03 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
         To understand the relationships between motif participation and consumers' probabilities of persistence, it is helpful to understand how persistence varies with other aspects of network structure.
-        Consumers with higher STL were less likely to persist (Fig.~\ref{fig:motifs_vs_TL_and_deg}, Table S3, \emph{Appendix S6}).
+        Consumers with higher STL were less likely to persist (Fig.~\ref{fig:lm_CS}, Table S3, \emph{Appendix S6}).
         The relationship between in-degree and persistence depended on the strength of disturbance: higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.
@@
         The average probability of persistence of all consumers in the network varied with the frequencies of different motifs in this profile.
-        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.
-S~\ref{fig:lm_CS}, Table S4, \emph{Appendix S8}).
+        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:lm_CS}, Table S4, \emph{Appendix S8}).
         There was no significant interaction between the proportion of any motif in the network profile and level of disturbance (Table S4), \emph{Appendix S8}).
@@
 
-    The effects of bottom-up disturbances on consumers depend on the large-scale structure of the affected food web~\citep{Dunne2002, Eklof2006, PascualDunne2006}.
-    Likewise, the distribution of meso-scale structures within a network affects the network's stability~\citep{prill2005dynamic, bascompte2005simple}.
-    In addition, both direct and indirect interactions (i.e., the local and meso-scale structure of a food web  [\citealp[]{Cirtwill2018FoodWebs}]) can affect which species are most strongly affected~\citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
-    Identifying these species is a high priority for conservation efforts~\citep{Bottrilletal2008}.
-    To advance this goal, we systematically explore the relationships between meso-scale structures and species' extinction risk, as well as how these trends overlap with the effects of large- and small-scale network properties.
+    Identifying the species that are most and least likely to persist after a disturbance is a high priority for conservation efforts~\citep{Bottrilletal2008}.
+    In a large set of simulated food webs, a consumer species' probability of persistence after a disturbance affecting basal resources depends on its frequency of participation in the four possible meso-scale motifs.
+    Species participating more frequently in the direct competition motif and less frequently in the omnivory motif are more likely to persist, while the effects of participating in the apparent competition or three-species chain motifs depend on the strength of disturbance.
+    Together with relationships between persistence and the overall structure of the network~\citep{Dunne2002, Eklof2006, PascualDunne2006,prill2005dynamic, bascompte2005simple} and simpler measures of species' positions within it~\citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}, these trends can help us to identify vulnerable species within a network.
 
     % Species-level results summary
-    As with degree and trophic level, a consumer's probability of persistence varied with its participation in different three-species motifs.
-    Specifically, increasing participation in the direct competition motif was always associated with higher probability of persistence while increasing participation in the omnivory motif was always associated with lower probability of persistence.
-    The relationships between persistence and apparent competition and the three-species chain depended on the strength of the disturbance, with apparent competition associated with greater probability of persistence when basal species were very likely to go extinct and the three-species chain associated with greater probability of persistence at lower levels of disturbance.
-
-    These trends can be partially explained by underlying relationships between motif participation and other measures of a species' place in the network.
+    Some of the relationships between motif participation and probability of persistence can be understood by examining relationships between motif participation and other measures of a species' place in the network.
     For example, species with higher trophic levels tended to participate less often in the direct competition motif.
@@
 
-    % % Degree and TL are both important for omnivory
-    % The relationship between participation in the omnivory motif and persistence connects motif participation with both in-degree and trophic level (STL).
-    % Species participation in the omnivory motif was associated with greater persistence, especially at low levels of disturbance.
-    % This is consistent with a strong negative correlation between omnivory and trophic level (leading to the overall positive effect of omnivory) and a strong positive correlation with in-degree (leading to the reduced benefit of omnivory at high levels of disturbance).
-    % Motifs can thus act as a tool for synthesizing the information provided by other measures of a species' place in its community and for understanding the interactions among these measures.
+    % Moving to network context
+    Relationships between motif participation and persistence can also be placed within the context of relationships between overall network structure, persistence, and the overall frequency of motifs in the network's motif profile.
+    For example, since networks with higher connectance contain on average more links per species, it logically follows that they will tend to include a greater proportion of the omnivory motif in the network's motif profile (Fig.
+SX) and species will tend to participate in the omnivory motif (Fig.
+5C).
+    As both higher connectance (Fig.
+4C) and more-frequent participation in the omnivory motif (Fig.
+2) are generally associated with decreased probability of persistence, it is possible that the negative effect of connectance may be because of the increased frequency of omnivory.
 
-    % % Degree and TL don't explain everything
-    % The relationships between persistence and participation in the apparent competition and three-species chain motifs likewise demonstrate the potential for motif participation to synthesize different aspects of network structure.
-    % In these cases, in-degree and trophic level are not sufficient to explain the trends we observe in the motifs.
-    % For apparent competition, there were strong and opposing correlations with in-degree and trophic level but the relationship between apparent competition and persistence only matches the trends for in-degree.
-    % This suggests that some other factor is over-riding the relationship between apparent competition and trophic level.
-
-    % Similarly, participation in the three-species chain motif was significantly correlated with both in-degree and trophic level, but the relationship between motif participation and persistence did not reflect the disturbance-dependent relationship between in-degree and persistence.
-    % This suggests that the relationship between persistence and participation in the chain motif synthesizes the effects of in-degree, trophic level, and some additional effect which counterbalances the effect of in-degree.
-    % This additional effect may come from indirect interactions, which are described by motifs but not in-degree~\citep{Cirtwill2018FoodWebs}.
-
-    % Moving to network context
-    Relationships between a species' probability of persistence and its' position within the network should be considered within the context of overall network structure.
-    In our networks, the average probability of persistence of all consumers increased with increasing proportions of the apparent competition, three-species chain, and direct competition motifs and decreased with increasing proportions of the omnivory motif.
-    The omnivory motif was also generally less common that the other motifs in these networks, which were selected to be stable before disturbance.
-
-    As well as clarifying the relationships between network structure and the effect of bottom-up disturbance on particular species, motifs were also related to the mean probability of persistence across all species in a network.
-    In fact, a network's motif profile had a stronger relationship to mean persistence compared to both network size and connectance. 
-    This suggests that network motif profiles may be more useful than large-scale network properties when predicting mean extinction risk of species in a network.
-    Connectance did, however, affect which motifs were most common within a network and therefore cannot be entirely ignored. 
-    Highly-connected networks tended to contain higher frequencies of the omnivory motif, which includes more links than the other three-species motifs considered here.
-    Large-scale properties may therefore indirectly affect network persistence by changing meso-scale properties, such as the distribution of motifs.
-
-    In general, average persistence was highest in networks with low proportions of the omnivory motif and high proportions of the apparent competition motif. 
-    These results contrast with earlier work showing that higher numbers of omnivory motifs and lower numbers of apparent competition motifs are associated with greater food-web persistence~\citep{Stouffer2010b}.
-    By using counts of motifs rather than proportions,~\citet{Stouffer2010b} may be mixing effects of total numbers of motifs with effects of motif proportions.
-    Moreover, the effect of omnivory on stability is known to be highly context-dependent \citep{bascompte2005simple, Monteiro2016,McLeod2021}. 
-    The stability of omnivory motifs in isolation depends on the strength of the omnivore-resource interaction~\citep{McLeod2021}, but at the whole-network level interactions with other species can stabilize even intrinsically unstable omnivory \citep{Kratina2012}.
-    This might be why we find different patterns between a network's motif profile (where a high proportion of omnivory indicates lower average network persistence) versus a single species motif participation (where a high proportion of the omnivory motif indicates a higher focal species persistence).
-    Thus, the ideal situation appears to be being a highly omnivorous species in a low-omnivory setting.
-
-    % High C webs and omnivory
-    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks and the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}) as in~\citet{McLeod2021}.
-    These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
+    However, as in other studies, the effect of omnivory on persistence in our networks was highly context-dependent~\citep{bascompte2005simple, Monteiro2016,McLeod2021}.
+    Under weak disturbance scenarios, the effect of increased participation in the omnivory motif on consumer persistence within a single network was likely to be neutral or positive (Fig. 3). 
+    At high levels of disturbance, however, participating in the omnivory motif was more likely to reduce the probability of persistence.
+    These trends also varied with connectance, with participation in the omnivory motif being associated with lower probability of persistence in less-connected networks and higher probability of persistence in more-connected networks as in~\citet{McLeod2021}.
+    This is also consistent with earlier work showing greater food-web persistence in networks with higher counts of the omnivory motif~\citep{Stouffer2010b}.
+    Highly-connected networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
     In such a network, it may be especially beneficial for a species to feed on multiple trophic levels and, in a sense, spread its risk of losing prey to bottom-up effects.
@@
 
+    [[Still seems relevant but don't know where to fit this in]]
+    The stability of omnivory motifs in isolation depends on the strength of the omnivore-resource interaction~\citep{McLeod2021}, but at the whole-network level interactions with other species can stabilize even intrinsically unstable omnivory \citep{Kratina2012}.
+
+    % Caveats - expand on top-down effects a bit to make R2 happy
     There are limitations to the approach presented here.
@@
         \centering
-        \includegraphics[width=1.0\textwidth]{figures/concept_fig_ver2_edit.eps}
+        \includegraphics[width=.95\textwidth]{figures/concept_fig_ver2_edit.eps}
         \caption{Conceptual figure of a toy food web.

### 2023-05-05 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-    % While a relationship between motif participation and persistence is broadly intuitive, so far there has been little research demonstrating which motifs are associated with greater probabilities of persistence.
-    % One potential reason for the difficulty in assessing the relationship between species' motif roles and persistence are the large computational resources required to simulate population dynamics.
-    % Our solution is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks~\citep{Eklof2013,Haussler2020}.
-    % One limitation with Bayesian networks is that they only take into account bottom-up effects (see Methods). 
-    % Nevertheless, Bayesian networks do capture the vast majority of secondary extinctions obtained from fully-fledged dynamical models including top-down and bottom-up effects~\citep{Eklof2013}, and as such are a good proxy for analysing species persistence following extinctions of primary producers in complex food webs.
-    % %The greater computational efficiency and similar [[results?
-performance?
-extinctions?]] of Bayesian networks allow us to simulate the effects of a gradient of disturbance strength and thereby explore whether participation in some motifs is associated with greater probability of extinction under strong or weak disturbances.
-
     %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
@@
     Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator.
-    % Furthermore, these four motifs do not include cycles (where species A and B eat each other or where A eats B eats C eats A).
-    % As modelling secondary extinctions using Bayesian networks requires acyclic networks (see Methods), these are the only motifs possible.
 
@@
         % , \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.
-        See Appendix \emph{3.1 - 3.5} for details.
+        See Appendix \emph{S3} for details.
 
@@
 
-        To understand the relationships between motif participation and consumers' probabilities of persistence, it is helpful to understand how persistence varies with other aspects of network structure.
-        Consumers with higher STL were less likely to persist (Fig.~\ref{fig:lm_CS}, Table S3, \emph{Appendix S6}).
+        To understand the relationships between motif participation and consumers' probabilities of persistence, it is helpful to understand how persistence varies with other aspects of both consumer properties and network structure.
+
+        Consumers with higher trophic level (STL) were less likely to persist (Fig.~\ref{fig:lm_CS}, Table S3, \emph{Appendix S6}).
         The relationship between in-degree and persistence depended on the strength of disturbance: higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.
@@
     Identifying the species that are most and least likely to persist after a disturbance is a high priority for conservation efforts~\citep{Bottrilletal2008}. 
-    In a large set of simulated food webs, a consumer species' probability of persistence after a disturbance affecting basal resources depends on its frequency of participation in the four possible meso-scale motifs.
+    Network structural properties such as connectance and measures of a species' position in the network such as degree and trophic level are known to affect species' vulnerability~\citep{Dunne2002, Eklof2006, PascualDunne2006,prill2005dynamic, bascompte2005simple,curtsdotter2011robustness, dunne2009cascading, Eklof2006}, but are relatively coarse metrics.
+    Defining a species' network role based on its participation in meso-scale motifs provides more nuanced information and may also predict extinction risk~\citep{}.
+    Here we tested how the persistence of a consumer in a food web depends on its participation in four different motifs with different levels of disturbance to basal resources.  
+    We found that motif participation was related to a consumer's probability of persistence in ways that can't be fully explained by simpler measures of network structure. 
     Species participating more frequently in the direct competition motif and less frequently in the omnivory motif are more likely to persist, while the effects of participating in the apparent competition or three-species chain motifs depend on the strength of disturbance.
-    Together with relationships between persistence and the overall structure of the network~\citep{Dunne2002, Eklof2006, PascualDunne2006,prill2005dynamic, bascompte2005simple} and simpler measures of species' positions within it~\citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}, these trends can help us to identify vulnerable species within a network.
 
-    % Species-level results summary
-    Some of the relationships between motif participation and probability of persistence can be understood by examining relationships between motif participation and other measures of a species' place in the network.
-    For example, species with higher trophic levels tended to participate less often in the direct competition motif.
-    A high trophic level is always associated with lower persistence since predators necessarily have lower probabilities of persistence than their prey~\citep{Eklof2013}.
-    Accordingly, high participation in the direct competition motif, by indicating a low trophic level, is associated with high probability of persistence.
-    Similarly, participation in apparent competition decreases with increasing in-degree, and probability of persistence shows the opposite trends with respect to in-degree and participation in apparent competition.
+    Disturbance strength is often neglected when predicting risk of extinction, but proved to be extremely important in our results.
+    Several of the relationships we demonstrated between probability of persistence and species roles changed direction when moving from weak to strong disturbance.
+    These simulated results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
+    For example, at low levels of disturbance a high degree may be protective since species can compensate if one or a few prey are lost. 
+    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and have lower persistence.
+    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{}.
+    These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.
 
-    The effects of other motifs cannot, however, be attributed so straightforwardly to simpler measures of a species' place in the network.
-    Like apparent competition, participation in the three-species chain motif decreases with increasing in-degree, yet persistence varied with the frequency of participation in the three-species chain in the same way as in-degree.
-    While participation in the three-species chain motif was positively correlated with trophic level, this cannot explain the relationship between trophic level and probability of persistence as trophic level was negatively correlated with persistence across all levels of disturbance.
-    Similarly, higher frequencies of participation in the omnivory motif were associated with lower probabilities of persistence, yet the frequency of the omnivory motif increased with in-degree and decreased with trophic level.
+    % Apparent competition as a case study that motifs are useful
+    The relationship between probability of persistence and participation in the apparent competition motif provides a good example of how motifs provide complementary information to that given by simpler measures of a species' network role.
+    Species with higher trophic levels tended to participate more frequently in the apparent competition motif.
+    As species with higher trophic levels generally have increased risk of going extinct~\citep{Eklof2013,others}, we would therefore expect high frequencies of participation in the apparent competition motif to also be associated with lower probability of persistence.
+    However, this is only true for low disturbance levels. 
+    For high disturbance, we see the opposite relationship – higher probability of persistence with more-frequent participation in apparent competition.
+    This change in the persistence-apparent competition relationship with changing disturbance is what we might expect based on the negative relationship between frequency of participating in apparent competition and in-degree (which showed the opposite trends with respect to persistence).
+    However, it is not obvious when trophic level should be the most important predictor of extinction risk and when in-degree should be most important.
+    Simpler measures of species' positions in a network could therefore lead us astray either because they provide conflicting predictions or because our expectations are only reasonable for specific types of disturbance.
+
+    % Try to condense the next few bits into one section on omnivory
+    Frequency of participation in the omnivory motif likewise provides information about extinction risk that is not easily obtained from in-degree and trophic level.
+    Higher frequencies of participation in the omnivory motif were associated with lower probabilities of persistence, yet the frequency of the omnivory motif increased with in-degree and decreased with trophic level.
     Both high in-degree and low trophic level are associated with higher probability of persistence, except at high levels of disturbance, which would naively suggest that higher proportions of omnivory should also be associated with greater persistence.
@@
 
-    % Moving to network context
-    Relationships between motif participation and persistence can also be placed within the context of relationships between overall network structure, persistence, and the overall frequency of motifs in the network's motif profile.
-    For example, since networks with higher connectance contain on average more links per species, it logically follows that they will tend to include a greater proportion of the omnivory motif in the network's motif profile (Fig.
-SX) and species will tend to participate in the omnivory motif (Fig.
-5C).
-    As both higher connectance (Fig.
-4C) and more-frequent participation in the omnivory motif (Fig.
-2) are generally associated with decreased probability of persistence, it is possible that the negative effect of connectance may be because of the increased frequency of omnivory.
+    While greater frequency of participation in the omnivory motif was generally associated with reduced probability of persistence across all levels of disturbance, this trend varied strongly with overall network structure (especially connectance).
+    Under weak disturbance scenarios, the effect of increased participation in the omnivory motif on consumer persistence within a single network was likely to be neutral or positive (Fig. 3). 
+    At high levels of disturbance, however, participating in the omnivory motif was more likely to reduce the probability of persistence as in~\citet{McLeod2021}.
+    This is consistent with other studies which have found the effect of omnivory to be highly context-dependent~\citep{bascompte2005simple, Monteiro2016}.
 
-    However, as in other studies, the effect of omnivory on persistence in our networks was highly context-dependent~\citep{bascompte2005simple, Monteiro2016,McLeod2021}.
-    Under weak disturbance scenarios, the effect of increased participation in the omnivory motif on consumer persistence within a single network was likely to be neutral or positive (Fig. 3). 
-    At high levels of disturbance, however, participating in the omnivory motif was more likely to reduce the probability of persistence.
-    These trends also varied with connectance, with participation in the omnivory motif being associated with lower probability of persistence in less-connected networks and higher probability of persistence in more-connected networks as in~\citet{McLeod2021}.
-    This is also consistent with earlier work showing greater food-web persistence in networks with higher counts of the omnivory motif~\citep{Stouffer2010b}.
     Highly-connected networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
@@
 
-    [[Still seems relevant but don't know where to fit this in]]
-    The stability of omnivory motifs in isolation depends on the strength of the omnivore-resource interaction~\citep{McLeod2021}, but at the whole-network level interactions with other species can stabilize even intrinsically unstable omnivory \citep{Kratina2012}.
+    In our simulated networks, increased persistence with greater participation in the omnivory motif could have emerged from the requirement that all networks permit the persistence of all species before disturbance was applied.
+    In highly-connected networks, which will tend to include more omnivory motifs simply because this motif contains one more link than the others (higher connectance implies more links per species), the abundant omnivory motifs must be generally more stable (or stabilizing) than is necessary for rarer omnivory motifs in less-connected networks.
+    Although testing the stability of individual motifs is outside the scope of this study, previous work on suggests that omnivory tends to be more stabilising/stable in isolation when the interactions involved are weak~\citep{}.
+    Although omnivory motifs tend to be less stable in isolation the the other motifs we consider here~\citep{Borrelli2015a}, when embedded in a network interactions with other species can stabilise even intrinsically unstable omnivory~\citep{Kratina2012}.
 
@@
     Most importantly, the Bayesian network framework operates on a strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
-    Here our focus was to analyze the importance of disturbances to the basal resources, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
+    Here our focus was to analyze bottom-up affects following disturbances to the basal resources, in which case the Bayesian network approach have been shown to capture a majority of the secondary extinctions obtained in a fully-fledged dynamical model but doing so with much greater efficiency~\citep{Eklof2013}.
     If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
-    For example, one recent study using a dynamical model shows that time to extinction for species following disturbance is tightly coupled to the motifs a species participates in, and that adding top-down effects can lead omnivory to be negatively correlated with persistence \citep{Cirtwill2021_inprep}.
-    If top-down effects are not crucial, the Bayesian network approach could be extended by explicitly incorporating information on species traits.
-    In empirical webs, species traits are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000},
-    adding an extra dimension of information which we do not consider here.
+    For example, although a recent study using a dynamical model to relate motif participation to persistence after a disturbance found similar trends for omnivory, participation in the direct competition motif was not universally positive~\citep{Cirtwill2021_inprep}.
+    In particular, species which frequently appeared in the prey position of the direct competition motif tended to have shorter persistence times after disturbance in that study - possibly due to top-down effects from the pair of predators involved.
+    However, as these two studies applied different types of disturbances, further study would be required to disentangle the effects of top-down and bottom-up effects on species with different motif participation, and how each relates to the type and strength of disturbance.
+    For this preliminary study, we have also neglected species traits.
+    As traits such as body size are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000} in empirical webs, this is also a potentially useful addition to future work.
 
-    In summary, our results show that participating in different three-species motifs is related to a species' risk of extinction following a bottom-up disturbance, and that these relationships synthesize information provided by other, simpler measures of species roles.
-    Critically, our results also underscore the importance of considering the level of disturbance when identifying which species are most at risk of extinction.
-    There were strong interactions between the level of disturbance and the relationships between persistence and multiple measures of species roles (motifs and in-degree).
-    These simulated results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
-    For example, at low levels of disturbance a high degree may be protective since species can compensate if one or a few prey are lost. 
-    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and have lower persistence.
-    These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.
+    [[Some kind of bridge between caveats and conclusion]]
+    Many ecologists, including ourselves, have argued that simple measures of species' roles within food webs are important for understanding species' extinction risk~\citep{}. 
+    However, here we show that relationships between extinction risk and measures of a species' place in its food web are likely more complex, varying with the strength of disturbance and the overall network structure.
+    This context-dependence means that even limited changes in network structure or the type of disturbance could change which species are most sensitive to disturbance. 
+    This argues for further caution in targeted conservation actions and increased acknowledgment of the importance and sensitivity of network structure in ecosystems.
 
@@
         \centering
-        \includegraphics[width=.95\textwidth]{figures/concept_fig_ver2_edit.eps}
+        \includegraphics[width=.9\textwidth]{figures/concept_fig_ver2_edit.eps}
         \caption{Conceptual figure of a toy food web.

### 2023-05-09 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.} \\
-    2. Coarse measures of a species' structural role, like in-degree and trophic level, are known to affect a consumer's vulnerability to bottom-up disturbance. \\
+    2.
+Given known relationships between coarse measures of a consumer's position in a food web and its ability to persist following disturbance, we expect that a species' participation in three-species motifs --meso-scale structures which provide information on a species' direct and indirect interactions-- will also be related to its probability of persistence.\\
     3.
-We are interested in whether a richer set of species roles --based on participation in three-species motifs-- is also related to persistence and how these motif-based roles relate to simpler measures.
-\\
-    4. We show that a consumer species' participation in different motifs is related to its probability of persistence. \\ 
-    5. Similarly, the overall motif profile of a network is related to the average persistence of consumer species. \\
-    6.
-Notably, we found that networks containing a higher proportion of the `omnivory' motif tend to have lower mean persistence following disturbance, while more frequent participation in the omnivory motif was associated with higher species-level probabilities of persistence.
-\\
-    7. Additionally, we find that relationships between motif participation and persistence can synthesize the information provided by in-degree and trophic level. \\
+We show that a consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.
+    4.
+However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.
+    5.
+
     8. Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities.
@@
     The loss of a species in a food web can, due to the mutual dependencies among species,  trigger a cascade of additional (secondary) extinctions. 
-    A species' position within a food web is known to affect its risk of going secondarily extinct ~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    A species' risk of going secondarily extinct is known to depend on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
     For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{binzer2011susceptibility, Eklof2006}.
     Species with low in-degree are more vulnerable to extinction because they have few alternative resources if a prey species is lost, while those with high trophic levels are vulnerable to the loss of any species along the path from basal resource(s) to the focal species.
-    Based on these trends, it has also been proposed that a species' participation in \emph{meso-scale structures} (which take into account the species' direct and indirect interactions) are also related to its persistence.
+    As well as these simple measures of food-web position, it has also been proposed that a species' participation in \emph{meso-scale structures} (which take into account the species' direct and indirect interactions) are related to its persistence.
 
@@
     Like degree and trophic level, a species' \emph{motif participation} (the vector of frequencies with which it appears in each motif) can easily be calculated once the structure of the network is known.
-    However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the focal species' interaction partners.
+    However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the positions of focal species' interaction partners.
     Two species might both have two prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
@@
 
-    The plant community is the foundation upon which the myriad species in food webs depend for their survival. 
-    Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
-    The loss of  basal level diversity typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
-\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
-    Given these far-reaching effects we here choose to analyse persistence after loss of basal resources, i.e., effects of bottom-up disturbances. 
-    We are particularly interested in whether the strength of this disturbance affects which consumers are most likely to go extinct. 
+    Here, we focus on persistence after the loss of plants or other basal resources, i.e., the effects of bottom-up disturbances.
+    The plant community is the foundation upon which the myriad species in food webs depend for their survival, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
+    Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
+    It is therefore critically important to understand how bottom-up effects of disturbance affect the rest of the food web.
+    As well as identifying identifying patterns of motif participation that are associated with greater or lesser ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
     For example, participation in one motif might be associated with greater probability of extinction after a mild disturbance but be associated with less risk of extinction when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances but also provide the species with alternatives when disturbances are strong.
@@
     To provide context for these results, we also consider how a consumer's risk of secondary extinction varies with the overall structure of the network, described by the frequency of motifs across the whole network, network size, and network connectance.
-    Finally, we relate species' motif participation to overall network structure and to their in-degrees and trophic levels in order to clarify what unique information motif participation can provide.
+    Finally, because measures of food-web position are not independent, we identify how species' motif participation varies with overall network structure and with in-degree and trophic level.
 
@@
         A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks than is practical for dynamical models \citep{Eklof2013,Haussler2020}.
-        In this framework, a consumer's probability of extinction depends on the fraction of resources lost ($f = k/n$ for a species with $n$ resources, of which $k$ have gone extinct) and on a baseline probability of extinction ($\pi$) which captures factors affecting the focal species' extinction probability that are not related to the network structure (e.g., disease, stochastic extinction of small populations).
+        In this framework, a consumer's probability of extinction depends on the fraction of resources lost ($f = k/n$ for a species with $n$ resources, of which $k$ have gone extinct) and on a baseline probability of extinction ($\pi$) which captures factors affecting the focal species' extinction probability that are not related to the network structure (e.g., disease or stochastic extinction of small populations).
         Importantly, the greater speed of the Bayesian network framework allows us to simulate a range of disturbance severities rather than single species removals as in many previous studies (e.g.,~\citealp[]{Memmott2004,Staniczenko2010,Dunne2004,Cirtwill2022Oikos}).
@@
         Any motif which contains a cycle is impossible in an acyclic network.
-        This means that only the four motifs which are most common in empirical networks (\citealp[]{Stouffer2007}: apparent competition, direct competition, omnivory, and the three-species chain) can appear.
-        Fortunately, they are also the motifs that have been most studied in isolation (e.g.,~\citep{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}) and have the clearest ecological interpretation and, as such, are of greatest interest to us.
+        This means that only four motifs can appear: apparent competition, direct competition, omnivory, and the three-species chain).
+        Fortunately, they are also the motifs that are most common in empirical networks (\citealp[]{Stouffer2007}, have been most studied in isolation (e.g.,~\citep{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.
 
@@
 		Similarly, if all resources are extinct ($f = 1$), the extinction probability $P(\lnot C|f)$ will be equal to 1 and the consumer will go extinct.
-		If the fraction $f$ is neither 1 nor 0, the non-linear sigmoid curve will determine the response of the consumer, which will be more severe if the fraction resources lost is high and less severe if the fraction is small.
+		If the fraction $f$ is neither 1 nor 0, the non-linear sigmoid curve will determine the response of the consumer, which will be more severe if the fraction of resources lost is high.
 This will in turn affect the extinction probability $P(\lnot C|f)$.
-		For basal resources, we assume that abiotic resources are always available and $f=0$. 
+		For basal resources, we assume that the abiotic resources they require are always available and therefore $f=0$. 
 		Following Equation \ref{betafunc}, the extinction probability of a basal resource will only depend on $\pi$.
@@
         As different measures of network structure are not independent since all describe the same network, we also 3) relate species' motif participation to large-scale network structure and to species in-degree and trophic level.
-        All analyses are briefly described below and were performed using the R~\citep{R} package \emph{lmerTest}~\citep{lmerTest}.
+        All analyses are briefly described below and were performed using  R~\citep{R} package.% \emph{lmerTest}~\citep{lmerTest}.
         % , \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.
@@
             As the three-way interaction was significant, we did not simplify the model. 
-
-            A consumer's probability of persistence is also known to vary with its in-degree and trophic level.
-            To confirm these relationships in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to disturbance, in-degree or STL, and the interaction between disturbance and the measure of network participation (\emph{Appendix S3}, Eq.
+            To confirm known relationships between persistence and in-degree or trophic level in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to disturbance, in-degree or STL, and the interaction between disturbance and the measure of network participation (\emph{Appendix S3}, Eq.
 S2-3).
@@
 
-            Since a network's motif profile and a species' motif participation reflect the same network as other measures (size, connectance, degree, and trophic level), it is likely that these measures are related.
-            We tested whether the frequency of motifs in the network's motif profile were related to network size, connectance, and their interaction using a set of four GLMs with binomial error distribution (Appendix \emph{S3}, eq.
+            Since a network's motif profile and a species' motif participation reflect the same network as other measures (size, connectance, degree, and trophic level), these measures are not independent.
+            We identified how the frequencies of motifs in the network's motif profile were related to network size, connectance, and their interaction using a set of four GLMs with binomial error distribution (Appendix \emph{S3}, eq.
 S4).
-            We tested whether the frequency of motifs in a consumer's motif participation vector were related to network size, connectance, and their interaction using a similar set of four GLMs with binomial error distribution.
+            We identified how the frequencies of motifs in a consumer's motif participation vector were related to network size, connectance, and their interaction using a similar set of four GLMs with binomial error distribution.
             Finally, we fit two additional sets of GLMs (eight total) relating the frequency of each motif in a consumer's motif participation vector to the consumer's in-degree or trophic level (Appendix \emph{S3}, Eq.
@@
         \centering
-        \includegraphics[width=.9\textwidth]{figures/concept_fig_ver2_edit.eps}
+        \includegraphics[width=.9\textwidth]{manuscript/figures/Figure_1.png}
         \caption{Conceptual figure of a toy food web.

### 2023-05-09 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
+% Limit is 8500 words.
+% Currently c. 9000
 \begin{document} 
@@
         For full details and a comparison to empirical networks, see \emph{Appendix S2}.
-        We removed any biologically unlikely networks (i.e., those with extremely long paths between any consumer and basal resources; \emph{Appendix S2}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
+        We removed any biologically unlikely networks (i.e., those with extremely long paths between any consumer and basal resources~\citep{Borrelli2014}; \emph{Appendix S2}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
 
@@
     Identifying the species that are most and least likely to persist after a disturbance is a high priority for conservation efforts~\citep{Bottrilletal2008}. 
-    Network structural properties such as connectance and measures of a species' position in the network such as degree and trophic level are known to affect species' vulnerability~\citep{Dunne2002, Eklof2006, PascualDunne2006,prill2005dynamic, bascompte2005simple,curtsdotter2011robustness, dunne2009cascading, Eklof2006}, but are relatively coarse metrics.
-    Defining a species' network role based on its participation in meso-scale motifs provides more nuanced information and may also predict extinction risk~\citep{}.
+    Network structural properties such as connectance and measures of a species' position in the network such as degree and trophic level are known to affect species' vulnerability~\citep{Dunne2002, Eklof2006, PascualDunne2006,prill2005dynamic, bascompte2005simple,curtsdotter2011robustness, dunne2009cascading}, but are relatively coarse metrics.
+    Defining a species' network role based on its participation in meso-scale motifs provides more nuanced information and may also predict extinction risk~\citep{Stouffer2007,Stouffer2010b,Simmons2019}.
     Here we tested how the persistence of a consumer in a food web depends on its participation in four different motifs with different levels of disturbance to basal resources.  
@@
     At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and have lower persistence.
-    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{}.
+    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{binzer2011susceptibility}.
     These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.
@@
     Species with higher trophic levels tended to participate more frequently in the apparent competition motif.
-    As species with higher trophic levels generally have increased risk of going extinct~\citep{Eklof2013,others}, we would therefore expect high frequencies of participation in the apparent competition motif to also be associated with lower probability of persistence.
+    As species with higher trophic levels generally have increased risk of going extinct~\citep{Eklof2006,Eklof2013}, we would therefore expect high frequencies of participation in the apparent competition motif to also be associated with lower probability of persistence.
     However, this is only true for low disturbance levels. 
@@
     In highly-connected networks, which will tend to include more omnivory motifs simply because this motif contains one more link than the others (higher connectance implies more links per species), the abundant omnivory motifs must be generally more stable (or stabilizing) than is necessary for rarer omnivory motifs in less-connected networks.
-    Although testing the stability of individual motifs is outside the scope of this study, previous work on suggests that omnivory tends to be more stabilising/stable in isolation when the interactions involved are weak~\citep{}.
+    Although testing the stability of individual motifs is outside the scope of this study, previous work on suggests that omnivory tends to be more stabilising/stable in isolation when the interactions involved are weak~\citep{Emmerson2004} or when the omnivore preys upon a pair of non-basal species~\citep{McLeod2021}.
     Although omnivory motifs tend to be less stable in isolation the the other motifs we consider here~\citep{Borrelli2015a}, when embedded in a network interactions with other species can stabilise even intrinsically unstable omnivory~\citep{Kratina2012}.
@@
     [[Some kind of bridge between caveats and conclusion]]
-    Many ecologists, including ourselves, have argued that simple measures of species' roles within food webs are important for understanding species' extinction risk~\citep{}. 
+    Many ecologists, including ourselves, have argued that simple measures of species' roles within food webs are important for understanding species' extinction risk~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006,binzer2011susceptibility}.
     However, here we show that relationships between extinction risk and measures of a species' place in its food web are likely more complex, varying with the strength of disturbance and the overall network structure.
@@
         \centering
-        \includegraphics[width=.9\textwidth]{manuscript/figures/Figure_1.png}
+        \includegraphics[width=.9\textwidth]{figures/Figure_1.png}
         \caption{Conceptual figure of a toy food web.
@@
     \label{fig:concept}
+    \end{figure}
+
+        \begin{figure}[hb!]
+        \centering
+        \includegraphics[width=.9\textwidth]{figures/concept_fig_ver2_edit.eps}
+        \caption{Note motif participation vectors in the right-hand column of the box. I still think it's nice to show - can you add back into current Fig. 1?}
+    \label{fig:concept_wrongcolour}
     \end{figure}

### 2023-05-09 manuscript/JAE_reply.tex

@@
 
-    \textbf{R:}
+    \textbf{R:} We thank the Editor for the option to revise and resubmit and for the helpful feedback from all concerned.
+We have taken careful note of the Associate Editor's and Reviewers' comments, particularly with regard to clarity and our statistical methods, and have extensively revised the manuscript.
+As a result, we are confident that our aims and results are now clearer and that that introduction and discussion more clearly place our work within the context of the literature.
+We hope that the Editor will agree that our work now meets the high standards of the Journal of Animal Ecology.
+
+    [[Some nice closing?]]
 
@@
 
-    \textbf{R:}
+    \textbf{R:} We thank the Associate Editor for his feedback and appreciate the points raised by the Reviewers.
+We have extensively revised our manuscript with regard to clarity and believe that the current version makes our main aim and conclusions clearer.
+We have also redone our analyses in line with the concern raised by Reviewer 1.
+These changes have, to our reading, greatly improved the manuscript and we are pleased to present this revision for new consideration.
 
@@
 
-\section*{Reply to Reviewer \#1}
+\section*{Reply to Reviewer \#1} [[Done?]]
 
@@
 
-    \textbf{R:}
+    \textbf{R:} We thank the Reviewer for taking the time to read the manuscript and for their thoughtful feedback.
+We have redone our analyses with binomial error distribution as suggested (though with glm rather than glmer as the random effects proved singular).
+Our main results did not change much with this improved approach, though some of our supplemental results (e.g., those relating to overall network structure) were simplified.
+We hope that the revised manuscript is now clearer and easier to read and that our statistical approach is more reasonable.
+Below, we respond to each of the Reviewer's comments (preceded by \textbf{R:}) in turn.
 
@@
 
-    \textbf{R:}
+    \textbf{R:} We thank the Reviewer for taking the time to read our manuscript and for the thoughtful feedback.
+We particularly appreciate the pointers where the structure of our manuscript was unclear an where it was difficult to extract our main aims and results.
+In the present revision we have substantially reworked the introduction and discussion and reordered the methods and results in order to present our work in a more streamlined and, hopefully, understandable manner.
+We believe that the manuscript has been greatly improved as a result.
+Note that, in response to a comment from the other Reviewer, we have also redone our analyses and some results have slightly changed from the previous submission.
+Below, we respond to each of the Reviewer's comments (preceded by \textbf{R:}) in turn.
 
@@
 
-                \textbf{R:}
+                \textbf{R:} We have thoroughly revised the summary to make it clearer that our focus is on species participation in motifs rather than the presence of motifs within a network.
+We hope that it now more-closely parallels the rest of the manuscript and is easier to follow.
 
@@
 
-                \textbf{R:}
-            The sentence the reviewer refer to is "How disturbances such as population decline or extinction lead to the loss of species via  direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research (Curtsdotter et al., 2011a; Dunne & Williams, 2009; Eklöf \& Ebenman, 2006 Santos et al., 2021)." We can understand the reviewer´s confusion, but in fact species extinctions are both the disturbance and the response; an initial extinction of a species (i.e.
-the primary extinction) can, since species are connected to each other via for example consumer -- resource interactions, cause other species to extinct (i.e.
+                \textbf{R:} The sentence the reviewer refers to is ``How disturbances such as population decline or extinction lead to the loss of species via  direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research (Curtsdotter et al., 2011a; Dunne & Williams, 2009; Eklöf \& Ebenman, 2006 Santos et al., 2021).'' We can understand the reviewer´s confusion, but in fact species extinctions are both the disturbance and the response.
+                An initial extinction of a species (i.e.
+the primary extinction) can, since species are connected to each other via  consumer--resource interactions, cause other species to go extinct (i.e.
 secondary extinctions).
 We have now clarified this on line XX:
-            "The loss of a species in a food web can, due to the mutual dependencies among species, trigger a cascade of additional extinctions, so called secondary extinctions.
-A species' position within a food web is known to affect its risk of going secondarily extinct ~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}."
+
+                \begin{quotation}
+                The loss of a species in a food web can, due to the mutual dependencies among species, trigger a cascade of additional extinctions, so called secondary extinctions.
+A species' risk of going secondarily extinct is known to depend on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+                \end{quotation}
 
@@
              \textbf{R:}
-             We have now clarified how we in this manuscript define a disturbance.
-This is done on line XX.
-" The loss of  basal level diversity typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
-\citep{scherber2010bottom}.
-Given these far-reaching effects we here choose to analyse persistence after loss of basal resources, i.e., effects of bottom-up disturbances.
-We here refer to disturbance as loss of basal species.
-"
+             We have now clarified how we in this manuscript define a disturbance. This is done on line XX.
+
+             \begin{quotation}
+             Here, we focus on persistence after the loss of plants or other basal resources, i.e., the effects of bottom-up disturbances.
+            \end{quotation}
 
@@
 
-                \textbf{R:}
-            We agree with the reviewer that this was not so clearly expressed.
-We have now re-formulated this section and clarified with an example.
-This is now located at line XX and reads: "However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the focal species' interaction partners.
+                \textbf{R:} We agree with the reviewer that this was not so clearly expressed. We have now re-formulated this section and clarified with an example, beginning at line XX.
+
+                \begin{quotation}
+                However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the focal species' interaction partners.
 Two species might both have three prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
-Importantly, since specialists are generally more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than a predator of generalists."
+Importantly, since specialists are generally more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than a predator of generalists.
+                \end{quotation}
 
@@
 
-                \textbf{R:}
-            [[we could also just rephrase these as small-scale and large-scale properties. Then it's maybe more obvious why meso-scale is in the middle.]]
-            We have now rephrased this part, now located on line XX: "Therefore, a species' motif participation profile captures the indirect interactions the species are affected by, thereby capturing important structures that are missed by global (network level) and local (species level) properties.
-As such, motif participation may complement degree and trophic level to more fully predict species' extinction risk."
+                \textbf{R:} To avoid introducing extra terminology, we have replaced all instances of `global' and `local' with `large-scale' and `small-scale' respectively.
+We hope it is now easier to understand motifs as a meso-scale descriptor of network structure.
+See line XX, quoted below, for a short description.
+We have also, as described above, explained how motif participation captures information about a focal species' interaction partners, and thereby the potential for indirect interactions.
+We do not refer to second-step indirect interactions in order to avoid confusion with that term.
+
+                \begin{quotation}
+                Therefore, a species' motif participation profile captures the indirect interactions the species are affected by, thereby capturing important structures that are missed by global (network level) and local (species level) properties.
+As such, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
+                \end{quotation}
 
@@
 
-                \textbf{R:}
-            See comment above.
+                \textbf{R:} See comment above.
 
@@
                 We have lightly rephrased the start of this paragraph (quoted below) and added some references to papers describing empirical networks with similar size and connectance to those we simulate.
-                Also note that we compare our simlated results with results for a set of more recent, larger, and especially well-resolved empirical networks in the SI.
+                Also note that we compare our simulated results with results for a set of more recent, larger, and especially well-resolved empirical networks in the SI.
 Although some of the whole-network structural properties differed, the range of species motif participations in our simulated networks spanned the whole range of observed motif participation in these empirical networks.
@@
 2, 3, and 5 for consistency.
+We have also thoroughly revised the results section to emphasize our main findings and place supplemental results in the proper context.
 
@@
 
-                \textbf{R:} Can the Reviewer provide a reference for their claim that apparent competition is destabilising?
-Is this a very general trend, or something that only happens in rare cases?
-Note that we do not claim that apparent competition is \emph{never} destabilising, only that more frequent participation in the apparent competition motif is associated with greater persistence on average.
-Moreover, we cannot think of a reason why a bottom-up disturbance (i.e., loss of some resources) would be destabilising in an apparent competition motif - surely it is more likely that a predator with two prey will persist after losing one than than two predator sharing one prey will persist if that prey is lost.
-We can see how it is possible for strong \emph{top-down} effects to destabilise an apparent competition motif (e.g., if a predator suddenly exerts much more pressure on one of the prey than the other), but this paper is concerned with bottom-up effects only.
-Given the earlier work showing that Bayesian networks capture the majority of secondary extinctions that occur in topological models that include both top-down and bottom-up effects, we can conclude that top-down effects are much less common as a cause of extinction than bottom-up effects.
-Of course they can still be equally interesting and may be strong in particular focal systems, but as our simulations do not reflect a specific community we have no reason to expect top-down effects to have a stronger role than in the networks used to validate the Bayesian network approach.
+                \textbf{R:} Previous research has shown that omnivory can either be stabilising or destabilising, depending on the precise disturbance, strengths of interactions, etc.
+Our results, which deal specifically with persistence after bottom-up disturbance, add to that picture of complexity without contradicting earlier findings.
+Note that we do not say that omnivory is \emph{never} beneficial - only that in our networks, in response to our focal disturbance, is is generally detrimental.
+We also note that our results show apparent competition as associated with less persistence when disturbance is weak, which is compatible with the possibility that the Reviewer mentions above (though we do not specifically model prey switching, which seems to be what the Reviewer refers to).
+In the revision we have expanded our discussion on how the strength and direction of motif participation-persistence relationships vary network-to-network (especially in networks with different connectance).
+We hope that this addition provides some of the nuance the Reviewer missed from the previous revision.
 
@@
                 None of this is to discount the reality of top-down extinctions or their interest from a scientific standpoint.
-However, given that our interest is in bottom-up extinctions in particular and that we are up front about the absence of top-down extinctions in our model [[]]
-
-                [[Add a comment about how strong top-down effects could change our interpretation to discussion?]]
+However, since the Bayesian network framework does not include top-down effects any comment we might make about how top-down effects might affect our results would necessarily be pure speculation.
+Instead, we have added a reference to another paper exploring the relationship between motif roles and persistence in a dynamical context.
+That paper, due to the computational limits inherent in the dynamical approach, uses only a single-species disturbance.
+It does, however, include top-down as well as bottom-up effects and may provide the information the Reviewer is seeking.
 
@@
 
-                \textbf{R:}
-                [[refocus on motif participation is definitely a good idea. Though there were interaction effects with disturbance strength, so we can't leave that out.]]
+                \textbf{R:} We have extensively revised the discussion to focus more clearly on the relationships between motif participation and persistence, although we note that some of these relationships depend on the strength of disturbance.
+As such, disturbance strength remains a strong secondary theme in the present version.
+We hope that this revision is easier to follow.
 

### 2023-05-11 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 \medskip
-\small$^3$ BioFrontiers Institute\\
-University of Colorado, Boulder, Boulder, U.S.A.\\
+\small$^3$ School of Biological Sciences\\
+University of Canterbury, Christchurch, New Zealand\\
 \medskip
@@
     3.
-We show that a consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.
+We show that consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.\\
     4.
-However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.
+However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.\\
     5.
@@
     A species' risk of going secondarily extinct is known to depend on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
-    For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{binzer2011susceptibility, Eklof2006}.
+    For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{Allesina2012,binzer2011susceptibility, Eklof2006}.
     Species with low in-degree are more vulnerable to extinction because they have few alternative resources if a prey species is lost, while those with high trophic levels are vulnerable to the loss of any species along the path from basal resource(s) to the focal species.
@@
     Two species might both have two prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
-    Importantly, since specialists are generally more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
+    Importantly, since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
     Therefore, a species' motif participation profile captures the indirect interactions that affect the species, thereby capturing important structures that are missed by larger- or smaller-scale descriptions of the network.
@@
     It is therefore critically important to understand how bottom-up effects of disturbance affect the rest of the food web.
-    As well as identifying identifying patterns of motif participation that are associated with greater or lesser ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
+    As well as identifying patterns of motif participation that are associated with greater or lesser ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
     For example, participation in one motif might be associated with greater probability of extinction after a mild disturbance but be associated with less risk of extinction when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances but also provide the species with alternatives when disturbances are strong.
@@
     %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
-    We define motif participation using three-species motifs, which are the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
+    We define motif participation using three-species motifs, which are the most commonly-used size for food web analyses~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
     Four of these motifs, namely the three-species chain, apparent competition, direct competition, and omnivory, are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, Borrelli2015a, giling2019plant}.
@@
     The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers (e.g., by depletion of the shared resource). 
-    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey (e.g., by boosting predator populations) in a top-down context, but in a bottom-up context only includes direct interactions.
+    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey (e.g., by boosting predator populations) in a top-down context, but in a bottom-up context, such as here, only includes direct interactions.
     Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator.
@@
         In-degree is a species' number of prey.
-        Trophic level (shortest trophic level; STL) is the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}).
+        Trophic level (shortest trophic level; STL) is the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}.
         Both simple role measures were calculated in R~\citep{R} using the same algorithm as~\citet{Eklof2013}.
 
-        Finally, to provide context for species' motif participation, we also calculated network ``motif profiles'': four-dimensional vectors of the number of each three-species motif in each network~\citep{Stouffer2012}.
+        Finally, to provide context for species' motif participation, we calculated network ``motif profiles'': four-dimensional vectors of the number of each three-species motif in each network~\citep{Stouffer2012}.
         Together with network size and connectance, these motif profiles describe the structure of the whole network rather than how a particular species participates in the community.
@@
         This means that cycles where species A eats species B and vice versa or where A eats B eats C eats A must be broken (see \emph{Appendix SX} for details).
-        Since cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited~\citep{Allesina2009functional}, this change will not strongly affect consumers' extinction risk in our simulations.
+        Since cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effects on network robustness are very limited~\citep{Allesina2009functional}, this change will not strongly affect consumers' extinction risk in our simulations.
 
@@
         Any motif which contains a cycle is impossible in an acyclic network.
-        This means that only four motifs can appear: apparent competition, direct competition, omnivory, and the three-species chain).
+        This means that only four motifs can appear: apparent competition, direct competition, omnivory, and the three-species chain.
         Fortunately, they are also the motifs that are most common in empirical networks (\citealp[]{Stouffer2007}, have been most studied in isolation (e.g.,~\citep{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.
@@
         As different measures of network structure are not independent since all describe the same network, we also 3) relate species' motif participation to large-scale network structure and to species in-degree and trophic level.
-        All analyses are briefly described below and were performed using  R~\citep{R} package.% \emph{lmerTest}~\citep{lmerTest}.
+        All analyses are briefly described below and were performed using  R~\citep{R} package. % \emph{lmerTest}~\citep{lmerTest}.
         % , \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.
@@
     Network structural properties such as connectance and measures of a species' position in the network such as degree and trophic level are known to affect species' vulnerability~\citep{Dunne2002, Eklof2006, PascualDunne2006,prill2005dynamic, bascompte2005simple,curtsdotter2011robustness, dunne2009cascading}, but are relatively coarse metrics.
-    Defining a species' network role based on its participation in meso-scale motifs provides more nuanced information and may also predict extinction risk~\citep{Stouffer2007,Stouffer2010b,Simmons2019}.
-    Here we tested how the persistence of a consumer in a food web depends on its participation in four different motifs with different levels of disturbance to basal resources.  
-    We found that motif participation was related to a consumer's probability of persistence in ways that can't be fully explained by simpler measures of network structure. 
+    Defining a species' network role based on its participation in meso-scale motifs provides more nuanced information and may also predict extinction risk~\citep{Stouffer2007,Stouffer2010b,Simmons2019,Cirtwill2018FoodWebs}, but this potential is only beginning to be explored~\citep{Cirtwill2022Oikos}.
+
+    Here we partially fill this knowledge gap by testing how the persistence of a consumer in a food web depends on its participation in four different motifs, and how this relationship varies with different levels of disturbance to basal resources.
+    We found that motif participation was related to a consumer's probability of persistence in ways not fully explained by simpler measures of network structure. 
     Species participating more frequently in the direct competition motif and less frequently in the omnivory motif are more likely to persist, while the effects of participating in the apparent competition or three-species chain motifs depend on the strength of disturbance.
 
-    Disturbance strength is often neglected when predicting risk of extinction, but proved to be extremely important in our results.
+    Disturbance strength is often neglected when predicting risk of extinction, but proved important in our results.
     Several of the relationships we demonstrated between probability of persistence and species roles changed direction when moving from weak to strong disturbance.
-    These simulated results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
+    These simulation results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
     For example, at low levels of disturbance a high degree may be protective since species can compensate if one or a few prey are lost. 
-    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and have lower persistence.
-    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{binzer2011susceptibility}.
+    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and therefore have lower persistence.
+    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{Allesina2012,binzer2011susceptibility}.
     These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.
@@
     % Apparent competition as a case study that motifs are useful
-    The relationship between probability of persistence and participation in the apparent competition motif provides a good example of how motifs provide complementary information to that given by simpler measures of a species' network role.
-    Species with higher trophic levels tended to participate more frequently in the apparent competition motif.
-    As species with higher trophic levels generally have increased risk of going extinct~\citep{Eklof2006,Eklof2013}, we would therefore expect high frequencies of participation in the apparent competition motif to also be associated with lower probability of persistence.
-    However, this is only true for low disturbance levels. 
+    The relationship between probability of persistence and participation in the apparent competition motif provides a good example of how motifs provide complementary information to simpler measures of a species' network role.
+    Species with higher trophic levels tend to participate more frequently in the apparent competition motif and
+    species with higher trophic levels generally have increased extinction risk~\citep{Eklof2006,Eklof2013}.
+We would therefore expect high frequencies of participation in the apparent competition motif to also be associated with lower persistence probability.
+    However, this is only true for low disturbance. 
     For high disturbance, we see the opposite relationship – higher probability of persistence with more-frequent participation in apparent competition.
-    This change in the persistence-apparent competition relationship with changing disturbance is what we might expect based on the negative relationship between frequency of participating in apparent competition and in-degree (which showed the opposite trends with respect to persistence).
+    This change in the persistence-apparent competition relationship across disturbance levels may be driven by the negative relationship between apparent competition and in-degree, and in-degree's negative relationship with persistence.
     However, it is not obvious when trophic level should be the most important predictor of extinction risk and when in-degree should be most important.
-    Simpler measures of species' positions in a network could therefore lead us astray either because they provide conflicting predictions or because our expectations are only reasonable for specific types of disturbance.
+    Such simpler measures of species' positions in a network could therefore lead us astray either because they provide conflicting predictions or because our expectations are only reasonable for specific types or levels of disturbance.
 
     % Try to condense the next few bits into one section on omnivory
-    Frequency of participation in the omnivory motif likewise provides information about extinction risk that is not easily obtained from in-degree and trophic level.
-    Higher frequencies of participation in the omnivory motif were associated with lower probabilities of persistence, yet the frequency of the omnivory motif increased with in-degree and decreased with trophic level.
-    Both high in-degree and low trophic level are associated with higher probability of persistence, except at high levels of disturbance, which would naively suggest that higher proportions of omnivory should also be associated with greater persistence.
-    Considering a species' in-degree, trophic level, \emph{and} its motif role may therefore provide extra information about its extinction risk.
-
-    While greater frequency of participation in the omnivory motif was generally associated with reduced probability of persistence across all levels of disturbance, this trend varied strongly with overall network structure (especially connectance).
-    Under weak disturbance scenarios, the effect of increased participation in the omnivory motif on consumer persistence within a single network was likely to be neutral or positive (Fig. 3). 
-    At high levels of disturbance, however, participating in the omnivory motif was more likely to reduce the probability of persistence as in~\citet{McLeod2021}.
+    Omnivory motif participation likewise provides information about extinction risk that is not easily obtained from in-degree and trophic level.
+    Omnivory motif participation decreased with trophic level (as does persistence) and increased with in-degree (as does persistence, except at high disturbance).
+This would naively suggest that higher omnivory participation should also be associated with greater persistence, but instead we (usually) find the opposite.
+The exception to this pattern is found at high connectance, where increased omnivory participation was more likely to have a positive or neutral relationship with consumer persistence, especially at low disturbance levels (Fig.
+3).
+    %While greater frequency of participation in the omnivory motif was generally associated with reduced probability of persistence across all levels of disturbance, this trend varied strongly with overall network structure (especially connectance).
+    %Under weak disturbance scenarios, the effect of increased participation in the omnivory motif on consumer persistence within a single network was likely to be neutral or positive (Fig. 3). 
+    %At high levels of disturbance, however, participating in the omnivory motif was more likely to reduce the probability of persistence as in~\citet{McLeod2021}.
     This is consistent with other studies which have found the effect of omnivory to be highly context-dependent~\citep{bascompte2005simple, Monteiro2016}.
+    %Considering a species' in-degree, trophic level, \emph{and} its motif role may therefore provide extra information about its extinction risk.
 
@@
     Most importantly, the Bayesian network framework operates on a strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
-    Here our focus was to analyze bottom-up affects following disturbances to the basal resources, in which case the Bayesian network approach have been shown to capture a majority of the secondary extinctions obtained in a fully-fledged dynamical model but doing so with much greater efficiency~\citep{Eklof2013}.
+    Here our focus was to analyze bottom-up affects following disturbances to the basal resources, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
     If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 

### 2023-05-11 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.} \\
-    2.
-Given known relationships between coarse measures of a consumer's position in a food web and its ability to persist following disturbance, we expect that a species' participation in three-species motifs --meso-scale structures which provide information on a species' direct and indirect interactions-- will also be related to its probability of persistence.\\
+    2. A species' position in a food web (e.g., its trophic level or number of prey) is known to affect its ability to persist following disturbance.
     3.
+We therefore expect that a species' participation in three-species motifs --meso-scale structures which provide information on a species' direct and indirect interactions-- will also be related to its probability of persistence.\\
+    4.
 We show that consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.\\
-    4.
+    5.
 However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.\\
-    5.
-
-    8. Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities.
+    6.
+Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities, but must be considered in the context of network structure as a whole and of the specific disturbance applied.
 
@@
     Species with low in-degree are more vulnerable to extinction because they have few alternative resources if a prey species is lost, while those with high trophic levels are vulnerable to the loss of any species along the path from basal resource(s) to the focal species.
-    As well as these simple measures of food-web position, it has also been proposed that a species' participation in \emph{meso-scale structures} (which take into account the species' direct and indirect interactions) are related to its persistence.
+    As well as these simple measures of food-web position, it has also been proposed that a species' participation in \emph{meso-scale structures} (which take into account the species' direct and indirect interactions) are related to its persistence~\citep{Cirtwill2022Oikos}.
 
@@
     The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers (e.g., by depletion of the shared resource). 
-    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey (e.g., by boosting predator populations) in a top-down context, but in a bottom-up context, such as here, only includes direct interactions.
+    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey (e.g., by boosting predator populations) in a top-down context, but only includes direct bottom-up interactions.
     Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator.
@@
     Here, we analyse whether a consumer's risk of secondary extinction after removal of basal resources depends on the consumer's motif participation and how this relationship varies with the severity of disturbance.
-    To provide context for these results, we also consider how a consumer's risk of secondary extinction varies with the overall structure of the network, described by the frequency of motifs across the whole network, network size, and network connectance.
+    To provide context for these results, we also consider how a consumer's risk of secondary extinction varies with the overall structure of the network, described by the frequency of motifs, network size, and network connectance.
     Finally, because measures of food-web position are not independent, we identify how species' motif participation varies with overall network structure and with in-degree and trophic level.
-
-     %They are also the only motifs which do not include cycles (where species A and B eat each other or where A eats B eats C eats A).
-    %As modelling bottom-up extinctions in a Bayesian framework requires the breaking of cycles, these are the only motifs that will appear in the resulting \emph{acyclic} networks.
-    %Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited.
-
-%     %\subsection*{1 Why are bottom-up effects important for understanding stability?}
-%      The plant community is the foundation upon which the myriad species in food webs depend for their survival.
-%      Disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant} and consumer diversity \citep{scherber2010bottom, Baiser2016}.
-%      As such, the structure of the plant community affects stability across all levels of ecosystem organization, which is crucial for sustaining a healthy energy flow in the ecosystem as a whole~\citep{proulx2010diversity, scherber2010bottom, Rosenblatt2016}.
-%      For example, the loss of diversity at the basal level typically causes declines in both abundance and richness of all types of consumers: herbivores, predators, parasitoids, etc.
-\citep{scherber2010bottom}.
-
-%     %\subsection*{2 Global and local network properties are known to affect stability}
-%     How disturbances such as population decline or extinction lead to the loss of species via direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
-%     Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances.
-Two well-known, influential global properties are network size, the number of species present in the network, and network connectance, the number of realized feeding interactions.
-%     Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
-%     The indisputable presence of large, persistent empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
-%     Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2015}.
-However, a well-connected network also includes more pathways through which a disturbance can spread~\citep{Dunne2004, Vieira2015}.
-Depending on the strength and arrangements of interactions, this means that species may be impacted through several paths, amplifying the original disturbance and causing more secondary extinctions.
-%     The net effect of connectance on extinction therefore likely depends on finer-scale (species-level) network structure, as well as the type and severity of disturbance.
-
-%     Species-level properties known to affect species' responses to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator).
-%     These can also be referred to as local properties of the network.
-%     Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and disturbances to plant populations cause particularly many secondary extinctions \citep{Curtsdotter2011}.
-%     However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
-%     A more detailed perspective is desirable since, as well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
-%     For example, a generalist consuming many specialist (and therefore high-risk) prey may be at greater risk of extinction than a specialist which consumes a highly generalist resource that is less likely to go extinct (Fig.~\ref{fig:concept}).
-
-%     %\subsection*{3 Motif participation may similarly affect species' particular extinction risk}
-%     It is possible to capture information about how a consumer and its prey fit together within a community by defining participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}.
-%     Such motif profiles of species are easily calculated once we know the network structure and require no more information than is used to calculate degree or trophic level.
-%     These `motif participation' roles capture information about species' direct as well as indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
-%     We refer to these motifs as meso-scale structures to emphasise their role as structures between global (network-level) and local (species-level) properties.  
-%     A species' motif participation profile shows the second-step indirect interactions the species are affected by, thereby capturing important structures that are missed by global and local properties.
-%     Therefore, by allowing researchers to distinguish between, for example, consumers of specialists and generalists, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
-
-%     While the potential relationships between motif participation and species-level extinction risk have not yet been tested to our knowledge, it has been shown that the frequencies of different motifs within a whole network are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}.
-%     These relationships between network-level motif frequencies (`motif profiles') and overall community stability mean that a network's motif profile could affect the average extinction risk for consumers within the network.
-%     Moreover, if network motif profiles are related to average extinction risk, it is worth investigating whether a species' participation in different motifs also affects its extinction risk.
-
-% % Severity of disturbance
-%     As mentioned above, the severity of a disturbance may interact with network structure to affect secondary extinctions.
-%     A mild disturbance (e.g., leading to a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
-%     Empirical studies have shown that disturbances to plant species can affect some consumers more strongly than others \citep{byrnes2011climate}, that the strength of disturbance affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}, and that the resilience and return time of the ecological community can be strongly influenced by disturbance severity \citep{rydgren2004disturbance}.
-%     The severity of disturbance therefore clearly has an effect on how the ecological community will respond.
 
@@
         We then described each consumer species' motif participation role in each of these simulated networks before applying any disturbance. 
-        A species' ``motif participation role'' is the frequency with which the species appears in each of the motifs present in a network~\citep{Stouffer2012}.
+        A species' ``motif participation role'' (Fig.~\ref{fig:concept}) is the frequency with which the species appears in each of the motifs present in a network~\citep{Stouffer2012}.
         As only four unique three-species motifs can appear in acyclic networks, these are four-dimensional vectors. 
@@
 As this method is probabilistic, we repeat these calculations 100 000 times per network, with unique random draws.
-        For each species, we then define probability of persistence as the fraction of simulations in which the species persists. 
+        For each species, we then define probability of persistence (hereafter `persistence') as the fraction of simulations in which the species persists. 
         While there are methods for solving Bayesian networks exactly \citep{Eklof2013}, for larger networks the numerical evaluation above is highly efficient and produces close to identical results \citep{Haussler2020}.
@@
 
-            To test whether a consumer's risk of extinction was associated with the frequency of each motif in its participation vector and whether this relationship varied with the strength of disturbance, we fit four general linear models with binomial error distribution (GLMs; one per motif included in the Bayesian networks):
+            To test whether a consumer's risk of extinction was associated with the frequency of each motif in its participation vector (hereafter simply `participation in motif X') and whether this relationship varied with the strength of disturbance, we fit four general linear models with binomial error distribution (GLMs; one per motif included in the Bayesian networks):
             \begin{equation}
@@
             To test whether these general trends were consistent across networks, we visually examined the distribution of slopes of simplified regressions fit to a single network and level of disturbance.
-            These models included only the frequency of the focal motif ($\rho_{i}$ above) and again used a binomial error distrubution.
+            These models included only participation in the focal motif ($\rho_{i}$ above) and again used a binomial error distrubution.
             To aid in interpretation of these single-network models with respect to network size and connectance, we fit an additional set of GLMs including proportion of motif, strength of disturbance, network size, and connectance across all networks (\emph{Appendix S6}).
@@
             We also confirmed previously-known relationships between persistence and other measures of network structure or a species' participation in the network.
-            To identify the relationship between large-scale network structure and mean probability of persistence within a network, and how the strength of disturbance might affect this relationship, we fit a GLM relating mean persistence to network size, connectance, disturbance, and their interactions.
+            To identify the relationship between large-scale network structure and mean persistence within a network, and how the strength of disturbance might affect this relationship, we fit a GLM relating mean persistence to network size, connectance, disturbance, and their interactions.
             As the three-way interaction was significant, we did not simplify the model. 
-            To confirm known relationships between persistence and in-degree or trophic level in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to disturbance, in-degree or STL, and the interaction between disturbance and the measure of network participation (\emph{Appendix S3}, Eq.
+            To confirm known relationships between persistence and in-degree or trophic level in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to in-degree or STL, disturbance, and the interaction between them (\emph{Appendix S3}, Eq.
 S2-3).
@@
             We identified how the frequencies of motifs in a consumer's motif participation vector were related to network size, connectance, and their interaction using a similar set of four GLMs with binomial error distribution.
-            Finally, we fit two additional sets of GLMs (eight total) relating the frequency of each motif in a consumer's motif participation vector to the consumer's in-degree or trophic level (Appendix \emph{S3}, Eq.
-S5-6).
+            Finally, we fit two additional sets of GLMs (eight total) relating participation in each motif to the consumer's in-degree or trophic level (Appendix \emph{S3}, Eq. S5-6).
 
@@
 
-        A species' probability of persistence was significantly related to its motif participation (Table S1, \emph{Appendix S4}).
+        A species' probability of persistence (hereafter `persistence') was significantly related to its motif participation (Table S1, \emph{Appendix S4}).
         Across all levels of disturbance, species were more likely to persist if they participated more frequently in the direct competition motif and less frequently in the omnivory motif (Fig.~\ref{fig:prop_lmer_all}).
-        The relationships between probability of persistence and participation in the other two motifs varied with the strength of disturbance.
+        The relationships between persistence and participation in the other two motifs varied with the strength of disturbance.
         At low levels of disturbance, species were more likely to persist if they participated more frequently in the three-species chain motif and less frequently in the apparent competition motif.
@@
         Networks with high and low connectances had different distributions of persistence-motif participation relationships, while network size had a much smaller effect (\emph{Appendix SX}).
-        The relationship between participation in the omnivory motif and probability of persistence was especially variable.
+        The relationship between participation in the omnivory motif and persistence was especially variable.
         Higher proportions of omnivory were associated with lower probabilities of persistence when connectance was low, regardless of network size or strength of disturbance.
@@
 
-        The average persistence of all consumers was lower in networks with higher connectance, but did not vary significantly with network size (Table S6, \emph{Appendix S10}).
-        The average probability of persistence of all consumers in the network varied with the frequencies of different motifs in this profile.
+        The average persistence of all consumers was lower in networks with higher connectance, did not vary significantly with network size (Table S6, \emph{Appendix S10}), and  varied with the frequencies of different motifs in this profile.
         Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:lm_CS}, Table S4, \emph{Appendix S8}).
@@
 
-        The frequency of participation in the omnivory motif generally increased with increasing connectance and network size, but decreased with increasing network size in the most-connected networks (Fig.
+        Omnivory motif participation generally increased with increasing connectance and network size, but decreased with increasing network size in the most-connected networks (Fig.
 S2 and Table S2, \emph{Appendix S5}).
-        The frequencies of participation in the apparent competition and three-species chain motifs decreased strongly with increasing connectance but showed little change with changing network size.
-        The frequency of participation in the direct competition motif decreased with increasing connectance in small networks and increased with increasing network size in highly-connected networks, but showed little change with connectance in large networks or with network size in less-connected networks.
+        Participation in the apparent competition and three-species chain motifs decreased strongly with increasing connectance but showed little change with changing network size.
+        Direct competition motif participation decreased with increasing connectance in small networks and increased with increasing network size in highly-connected networks, but showed little change with connectance in large networks or with network size in less-connected networks.
 
@@
     Here we partially fill this knowledge gap by testing how the persistence of a consumer in a food web depends on its participation in four different motifs, and how this relationship varies with different levels of disturbance to basal resources.
-    We found that motif participation was related to a consumer's probability of persistence in ways not fully explained by simpler measures of network structure. 
+    We found that motif participation was related to a consumer's persistence in ways not fully explained by simpler measures of network structure. 
     Species participating more frequently in the direct competition motif and less frequently in the omnivory motif are more likely to persist, while the effects of participating in the apparent competition or three-species chain motifs depend on the strength of disturbance.
@@
     % Apparent competition as a case study that motifs are useful
-    The relationship between probability of persistence and participation in the apparent competition motif provides a good example of how motifs provide complementary information to simpler measures of a species' network role.
+    The relationship between persistence and participation in the apparent competition motif provides a good example of how motifs provide complementary information to simpler measures of a species' network role.
     Species with higher trophic levels tend to participate more frequently in the apparent competition motif and
-    species with higher trophic levels generally have increased extinction risk~\citep{Eklof2006,Eklof2013}.
-We would therefore expect high frequencies of participation in the apparent competition motif to also be associated with lower persistence probability.
+    species with higher trophic levels generally have lower persistence~\citep{Eklof2006,Eklof2013}.
+We would therefore expect high participation in the apparent competition motif to also be associated with lower persistence.
     However, this is only true for low disturbance. 
-    For high disturbance, we see the opposite relationship – higher probability of persistence with more-frequent participation in apparent competition.
+    For high disturbance, we see the opposite relationship.
     This change in the persistence-apparent competition relationship across disturbance levels may be driven by the negative relationship between apparent competition and in-degree, and in-degree's negative relationship with persistence.
-    However, it is not obvious when trophic level should be the most important predictor of extinction risk and when in-degree should be most important.
+    However, it is not obvious when trophic level should be the most important predictor of persistence and when in-degree should be most important.
     Such simpler measures of species' positions in a network could therefore lead us astray either because they provide conflicting predictions or because our expectations are only reasonable for specific types or levels of disturbance.
 
-    % Try to condense the next few bits into one section on omnivory
-    Omnivory motif participation likewise provides information about extinction risk that is not easily obtained from in-degree and trophic level.
-    Omnivory motif participation decreased with trophic level (as does persistence) and increased with in-degree (as does persistence, except at high disturbance).
-This would naively suggest that higher omnivory participation should also be associated with greater persistence, but instead we (usually) find the opposite.
+    Omnivory motif participation likewise provides information about persistence that is not easily obtained from in-degree and trophic level.
+    Although omnivory participation decreased with trophic level (as does persistence) and increased with in-degree (as does persistence, except at high disturbance), we found that greater omnivory participation was usually associated with \emph{lower} persistence-- contrary to what we would expect based on trophic level and in-degree.
     The exception to this pattern is found at high connectance, where increased omnivory participation was more likely to have a positive or neutral relationship with consumer persistence, especially at low disturbance levels (Fig.
 3).
-    %While greater frequency of participation in the omnivory motif was generally associated with reduced probability of persistence across all levels of disturbance, this trend varied strongly with overall network structure (especially connectance).
-    %Under weak disturbance scenarios, the effect of increased participation in the omnivory motif on consumer persistence within a single network was likely to be neutral or positive (Fig. 3). 
-    %At high levels of disturbance, however, participating in the omnivory motif was more likely to reduce the probability of persistence as in~\citet{McLeod2021}.
-    This is consistent with other studies which have found the effect of omnivory to be highly context-dependent~\citep{bascompte2005simple, Monteiro2016}.
-    %Considering a species' in-degree, trophic level, \emph{and} its motif role may therefore provide extra information about its extinction risk.
-
-    Highly-connected networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
-    In such a network, it may be especially beneficial for a species to feed on multiple trophic levels and, in a sense, spread its risk of losing prey to bottom-up effects.
-    Although species in our model cannot swap prey, the reduced extinction risk for omnivorous species resembles empirical studies showing that adaptive omnivory can increase the permanence of three-species chains~\citep{Fagan1997, Kvrivan2005, AbramsFung2010}.
-
-    In our simulated networks, increased persistence with greater participation in the omnivory motif could have emerged from the requirement that all networks permit the persistence of all species before disturbance was applied.
-    In highly-connected networks, which will tend to include more omnivory motifs simply because this motif contains one more link than the others (higher connectance implies more links per species), the abundant omnivory motifs must be generally more stable (or stabilizing) than is necessary for rarer omnivory motifs in less-connected networks.
-    Although testing the stability of individual motifs is outside the scope of this study, previous work on suggests that omnivory tends to be more stabilising/stable in isolation when the interactions involved are weak~\citep{Emmerson2004} or when the omnivore preys upon a pair of non-basal species~\citep{McLeod2021}.
+    This is consistent with other studies which have found the effect of omnivory to be highly context-dependent~\citep{bascompte2005simple, Monteiro2016} and particularly beneficial when interaction strengths are weak~\citep{Emmerson2004}.
     Although omnivory motifs tend to be less stable in isolation the the other motifs we consider here~\citep{Borrelli2015a}, when embedded in a network interactions with other species can stabilise even intrinsically unstable omnivory~\citep{Kratina2012}.
+    Though we do not explicitly model interaction strengths, we do note that the greater number of interactions per species in high-connectance networks should lead to weaker interactions in general, as well as allowing many opportunities for additional interactions that could stabilise omnivory.
 
@@
     However, as these two studies applied different types of disturbances, further study would be required to disentangle the effects of top-down and bottom-up effects on species with different motif participation, and how each relates to the type and strength of disturbance.
-    For this preliminary study, we have also neglected species traits.
-    As traits such as body size are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000} in empirical webs, this is also a potentially useful addition to future work.
+    Another potentially useful addition to future work would be to include species traits such as body size, which are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000} in empirical webs.
 
-    [[Some kind of bridge between caveats and conclusion]]
+    \subsection*{Conclusions}
+
     Many ecologists, including ourselves, have argued that simple measures of species' roles within food webs are important for understanding species' extinction risk~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006,binzer2011susceptibility}.
@@
     This context-dependence means that even limited changes in network structure or the type of disturbance could change which species are most sensitive to disturbance. 
-    This argues for further caution in targeted conservation actions and increased acknowledgment of the importance and sensitivity of network structure in ecosystems.
+    As such, caution and increased attention to both network structure and the specifics of disturbance are needed for targetted conservation actions to be most effective.
 
@@
     \centering
-        \includegraphics[width=\textwidth]{manuscript/figures/Fig4.pdf}
+        \includegraphics[width=\textwidth]{figures/Fig4.pdf}
         \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.

### 2023-05-11 manuscript/reorder_main_text.tex
better length calc

--- 2023-05-11 manuscript/reorder_main_text.tex
@@
 \begin{spacing}{2.0}
-\linenumbers
+% \linenumbers
 
@@
         \centering
-        \includegraphics[width=.9\textwidth]{figures/Figure_1.png}
+        % \includegraphics[width=.9\textwidth]{figures/Figure_1.png}
         \caption{Conceptual figure of a toy food web.
@@
 
-        \begin{figure}[hb!]
-        \centering
-        \includegraphics[width=.9\textwidth]{figures/concept_fig_ver2_edit.eps}
-        \caption{Note motif participation vectors in the right-hand column of the box. I still think it's nice to show - can you add back into current Fig. 1?}
-    \label{fig:concept_wrongcolour}
-    \end{figure}
+    %     \begin{figure}[hb!]
+    %     \centering
+    %     \includegraphics[width=.9\textwidth]{figures/concept_fig_ver2_edit.eps}
+    %     \caption{Note motif participation vectors in the right-hand column of the box. I still think it's nice to show - can you add back into current Fig. 1?}
+    % \label{fig:concept_wrongcolour}
+    % \end{figure}
 
@@
         \centering
-        \includegraphics[width=0.85\textwidth]{figures/persistence_motif_participation.eps}
+        % \includegraphics[width=0.85\textwidth]{figures/persistence_motif_participation.eps}
         \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).
@@
     \centering
-        \includegraphics[width=\textwidth]{manuscript/figures/Fig4.pdf}
+        % \includegraphics[width=\textwidth]{figures/Fig4.pdf}
         \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
+        % \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
         \caption{The probabilities of persistence for a particular consumer (\textbf{A, B}) and the average probability of persistence across all consumers (\textbf{C-E}) varied with other elements of network structure as well as with motif participation.
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
+        % \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
         \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence.

### 2023-05-11 manuscript/reorder_main_text.tex

--- 2023-05-11 manuscript/reorder_main_text.tex
@@
     \centering
-        \includegraphics[width=\textwidth]{figures/Fig4.pdf}
+        % \includegraphics[width=\textwidth]{figures/Fig4.pdf}
         \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.

### 2023-05-11 manuscript/reorder_main_text.tex
Updates from Overleaf

--- 2023-05-11 manuscript/reorder_main_text.tex
@@
 
-% Limit is 8500 words.
+% Limit is 8200 words.
 % Currently c. 9000
@@
     The loss of a species in a food web can, due to the mutual dependencies among species,  trigger a cascade of additional (secondary) extinctions. 
-    A species' risk of going secondarily extinct is known to depend on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+    A species' risk of secondary extinction depends on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
     For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{Allesina2012,binzer2011susceptibility, Eklof2006}.
-    Species with low in-degree are more vulnerable to extinction because they have few alternative resources if a prey species is lost, while those with high trophic levels are vulnerable to the loss of any species along the path from basal resource(s) to the focal species.
-    As well as these simple measures of food-web position, it has also been proposed that a species' participation in \emph{meso-scale structures} (which take into account the species' direct and indirect interactions) are related to its persistence~\citep{Cirtwill2022Oikos}.
+    These species are more vulnerable because they are impacted by the loss of any species along a long path from basal resources or because they have few alternative resources if a prey species is lost.
+    Similarly, a species' participation in \emph{meso-scale motifs} (which provide a more nuanced description of a species' place in the food web) are related to its persistence~\citep{Cirtwill2015a,Cirtwill2018FoodWebs,Cirtwill2022Oikos}.
 
-    One way to describe a species' participation in such meso-scale structures is using its participation in the set of \emph{motifs} of a given size~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
-    Motifs --unique patterns of $n$ interacting species-- correspond to different sets of direct and indirect interactions and provide a holistic description of a species' place in a network~\citep{Stouffer2007,Stouffer2012}.
-    Like degree and trophic level, a species' \emph{motif participation} (the vector of frequencies with which it appears in each motif) can easily be calculated once the structure of the network is known.
-    However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the positions of focal species' interaction partners.
+    Motifs are unique patterns of $n$ interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
+    The vector of frequencies with which a species appears in each motif --its \emph{motif participation}-- describes the species' position in the meso-scale structure of a food web and, like degree and trophic level, is easily calculated once the structure of the network is known.
+    Importantly, motif participation incorporates information on the positions of focal species' interaction partners, as well as potential indirect interactions between species.
     Two species might both have two prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
     Importantly, since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
-    Therefore, a species' motif participation profile captures the indirect interactions that affect the species, thereby capturing important structures that are missed by larger- or smaller-scale descriptions of the network.
-    As such, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
+    This extra information on the focal species' interaction partners may thus provide extra insight when predicting extinction risk.
 
-    Here, we focus on persistence after the loss of plants or other basal resources, i.e., the effects of bottom-up disturbances.
-    The plant community is the foundation upon which the myriad species in food webs depend for their survival, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
+    Here, we focus on extinction risk after the loss of plants or other basal resources, i.e., the effects of bottom-up disturbances.
+    The basal resource community is the foundation upon which all other species depend, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
     Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
-    It is therefore critically important to understand how bottom-up effects of disturbance affect the rest of the food web.
-    As well as identifying patterns of motif participation that are associated with greater or lesser ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
-    For example, participation in one motif might be associated with greater probability of extinction after a mild disturbance but be associated with less risk of extinction when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances but also provide the species with alternatives when disturbances are strong.
+    It is therefore crucial to understand how bottom-up effects of disturbance to basal resources affect the rest of the food web.
+    As well as identifying patterns of motif participation that are associated with ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
+    For example, participation in one motif might be associated with greater extinction risk after a mild disturbance but be associated with less risk when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances while providing `back-up' resources when many prey are removed by a strong disturbance.
 
     %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
-    We define motif participation using three-species motifs, which are the most commonly-used size for food web analyses~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
-    Four of these motifs, namely the three-species chain, apparent competition, direct competition, and omnivory, are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, Borrelli2015a, giling2019plant}.
+    We define motif participation using the well-studied three-species motifs~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
+    Four of these motifs --the three-species chain, apparent competition, direct competition, and omnivory-- are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, giling2019plant} and more stable than other motifs when modelled in isolation~\citep{Borelli2015a}.
     These four motifs also have clear ecological interpretations.

### 2023-05-11 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
     We define motif participation using the well-studied three-species motifs~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
-    Four of these motifs --the three-species chain, apparent competition, direct competition, and omnivory-- are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, giling2019plant} and more stable than other motifs when modelled in isolation~\citep{Borelli2015a}.
+    Four of these motifs --the three-species chain, apparent competition, direct competition, and omnivory-- are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, giling2019plant} and more stable than other motifs when modelled in isolation~\citep{Borrelli2015a}.
     These four motifs also have clear ecological interpretations.
@@
 
-        We generated simulated networks using the niche model, which has been shown to recreate the structure of empirical networks well~\citep{Williams2000,Stouffer2007}.
+        We generated simulated networks using the niche model, which recreates the structure of empirical networks well~\citep{Williams2000,Stouffer2007}.
         To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002e}, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
+        We removed any biologically unlikely networks (i.e., those with extremely long paths between any consumer and basal resources~\citep{Borrelli2014} and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
         All networks were generated using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.
 For full details and a comparison to empirical networks, see \emph{Appendix S2}.
-        We removed any biologically unlikely networks (i.e., those with extremely long paths between any consumer and basal resources~\citep{Borrelli2014}; \emph{Appendix S2}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
 
@@
         %  These processing steps allow the strictly bottom-up calculation of probability of persistence.
-        We then described each consumer species' motif participation role in each of these simulated networks before applying any disturbance. 
-        A species' ``motif participation role'' (Fig.~\ref{fig:concept}) is the frequency with which the species appears in each of the motifs present in a network~\citep{Stouffer2012}.
+        We described the motif participation of each consumer species in each of these simulated networks before applying any disturbance. 
+        A species' motif participation (Fig.~\ref{fig:concept}) is the vector of frequencies with which it appears in each type of motif present in a network~\citep{Stouffer2012}.
         As only four unique three-species motifs can appear in acyclic networks, these are four-dimensional vectors. 
         Here we are interested in the relative frequencies of each motif rather than the total number of motifs and therefore normalized participation vectors by dividing each count by the total number of motifs the species appears in (such that the vector for each species sums to 1).
-        Note, however, that this normalization does not necessarily control for differences due to degree or connectance~\citep{Cirtwill2022Oikos}.
+        Note, however, that this normalization does not fully control for differences due to degree or connectance~\citep{Cirtwill2022Oikos}.
 
-        We also calculated two simpler measures of a species' role within its community.
+        We also calculated two simpler measures of a species' position within its community.
         In-degree is a species' number of prey.
@@
         Finally, to provide context for species' motif participation, we calculated network ``motif profiles'': four-dimensional vectors of the number of each three-species motif in each network~\citep{Stouffer2012}.
-        Together with network size and connectance, these motif profiles describe the structure of the whole network rather than how a particular species participates in the community.
+        Together with network size and connectance, these motif profiles describe the structure of the whole network.
         To separate differences in motif structure from differences in network size or connectance (larger and more-connected networks contain more motifs), we normalized motif profiles by dividing the count of each motif by the total count of all motifs.
-        We calculated species' motif participation using the Python package \emph{pymfinder}~\citep{pymfinder}.
+        We calculated species' motif participation and network motif profiles using the Python package \emph{pymfinder}~\citep{pymfinder}.
 
@@
         Traditionally, there are two main approaches for studying secondary extinctions (i.e., extinctions of species that were not directly disturbed). 
-        First, there are topological models, based only on food web structure \citep{dunne2009cascading}. 
-        Here, extinctions only affect other species in the network once a consumer has lost all of its prey and therefore must go extinct. 
-        The second approach uses dynamical models which explicitly simulate population dynamics using a system of differential equations \citep{binzer2011susceptibility}. 
-        Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network. 
+        First, there are topological models where secondary extinctions only occur once a consumer has lost all of its prey~\citep{dunne2009cascading}. 
+        Alternatively, dynamical models explicitly simulate population dynamics using a system of differential equations and take changes in prey or predator populations into account when calculating populations of other species in the network~\citep{binzer2011susceptibility}.
         This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are highly parameter intensive and simulations are much more time-consuming than topological models.

### 2023-05-11 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-        A middle‐ground approach for simulating secondary extinctions is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks than is practical for dynamical models \citep{Eklof2013,Haussler2020}.
-        In this framework, a consumer's probability of extinction depends on the fraction of resources lost ($f = k/n$ for a species with $n$ resources, of which $k$ have gone extinct) and on a baseline probability of extinction ($\pi$) which captures factors affecting the focal species' extinction probability that are not related to the network structure (e.g., disease or stochastic extinction of small populations).
+        A middle‐ground approach is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks than is practical for dynamical models \citep{Eklof2013,Haussler2020}.
+        In this framework, a consumer's probability of extinction depends on the fraction of resources lost ($f = k/n$ for a species with $n$ resources, of which $k$ have gone extinct) and on a baseline probability of extinction ($\pi$) which capturing factors that are not related to the network structure (e.g., disease or stochastic extinction of small populations).
         Importantly, the greater speed of the Bayesian network framework allows us to simulate a range of disturbance severities rather than single species removals as in many previous studies (e.g.,~\citealp[]{Memmott2004,Staniczenko2010,Dunne2004,Cirtwill2022Oikos}).
-        Moreover, Bayesian network simulations are less parameter-intensive than dynamical models, meaning that they are less sensitive to assumptions made in the modelling process~\citep{Eklof2013}.
+        Moreover, Bayesian network simulations are less parameter-intensive than dynamical models and are therefore less sensitive to assumptions made in the modelling process~\citep{Eklof2013}.
 
@@
 
-        \subsubsection*{Modelling consumers' risk of secondary extinction}
+        \subsubsection*{Modelling consumers' risk of secondary extinction} [[trimmed down to here]]
 

### 2023-05-12 manuscript/JAE_reply.tex
Updates from Overleaf

@@
 
-                \textbf{R:}
-                TO DO!
+                \textbf{R:} Every study we are aware of that relates motifs to stability either models one or more motifs in isolation or examines the effect of motif profiles (or elements thereof, e.g., frequency of omnivory) on the whole network.
+Neither one is exactly applicable to the effect of motif participation on per-species extinction risk.
+The exception is an earlier study by some authors of the present work, which we mention but do not wish to over-promote by making this paper a keystone of the introduction.
+For these reasons, together with constraints of length, we have been judicious in our selection of additional material on previous motif-stability relationships.
+Particularly for omnivory, the existing literature paints a picture of context-dependence and variable effects rather than a clear general trend.
+Reviewing this material would be more suited to a general review (perhaps with supplemental modelling to fill gaps between tests of different types of stability, disturbance, etc.) than to a paragraph in our introduction.
+We hope that the refocusing and revision of the introduction provides enough clarity to compensate for this necessarily cursory overview.
 

### 2023-05-12 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-% Limit is 8200 words.
-% Currently c. 9000
+% Limit is 8500 words.
+% Currently c. 8066 - we're good!
 \begin{document} 
@@
     We define motif participation using the well-studied three-species motifs~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
-    Four of these motifs --the three-species chain, apparent competition, direct competition, and omnivory-- are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, giling2019plant} and more stable than other motifs when modelled in isolation~\citep{Borrelli2015a}.
+    Four of these motifs --the three-species chain, apparent competition, direct competition, and omnivory-- are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, giling2019plant} and more stable than other motifs (i.e., are more likely to retain all three species in the motif) when modelled in isolation~\citep{Borrelli2015a}.
     These four motifs also have clear ecological interpretations.
     The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
-    The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers (e.g., by depletion of the shared resource). 
-    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey (e.g., by boosting predator populations) in a top-down context, but only includes direct bottom-up interactions.
+    The dynamics of three-species chains in isolation are well-understood from Lotka-Volterra models, and stable equilibria with all three species persisting can often be identified~\citep{Borrelli2015a,Hsu2015}.
+    In general, food chains are more likely to be stable when predator-prey body-mass ratios are smaller~\citep{Jonsson1998}.
+    The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers by depletion of the shared resource in competitive exclusion~\citep{Hsu2015}.
+    Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.
+    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey indicates a generalist predator (which may be able persist following reduction or loss of one prey) but can capture indirect effects between prey that mae the predator more likely to go extinct~\citep{Melian2002}.
     Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator. 
+    Omnivory is less stable in isolation than the other three motifs~\citep{Borelli2015a}, though there is a region of stable equilibrium for Lotka-Volterra model of omnivory~\citep{Hsu2015}.
+    In whole food webs, frequent omnivory has been proposed to both increase and decrease overall stability in different conditions~\citep{McCann1997,Emmerson2004,Monteiro2016}.
+    As most of the above studies either model motifs in isolation or examine overall effects on the whole food web, it remains unclear how a particular species' motif participation will affect its specific extinction risk.
 

### 2023-05-12 manuscript/reorder_main_text.tex
refs

@@
     Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator. 
-    Omnivory is less stable in isolation than the other three motifs~\citep{Borelli2015a}, though there is a region of stable equilibrium for Lotka-Volterra model of omnivory~\citep{Hsu2015}.
+    Omnivory is less stable in isolation than the other three motifs~\citep{Borrelli2015a}, though there is a region of stable equilibrium for Lotka-Volterra model of omnivory~\citep{Hsu2015}.
     In whole food webs, frequent omnivory has been proposed to both increase and decrease overall stability in different conditions~\citep{McCann1997,Emmerson2004,Monteiro2016}.

### 2023-05-12 manuscript/reorder_main_text.tex
fixed SI refs

@@
         All networks were generated using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.
-For full details and a comparison to empirical networks, see \emph{Appendix S2}.
+For full details and a comparison to empirical networks, see \emph{Appendix S1}.
 
-        % To make the simulated networks compatible with the Bayesian simulation framework, all networks were rendered acyclic following~\citet{Allesina2009functional} and species topologically sorted from lowest to highest trophic level (\emph{Appendix S2}).
-        %  These processing steps allow the strictly bottom-up calculation of probability of persistence.
         We described the motif participation of each consumer species in each of these simulated networks before applying any disturbance. 
@@
         Bayesian networks also require the creation of acyclic networks so that a strict bottom-up sequence can be followed. 
-        This means that cycles where species A eats species B and vice versa or where A eats B eats C eats A must be broken (see \emph{Appendix SX} for details).
+        This means that cycles where species A eats species B and vice versa or where A eats B eats C eats A must be broken (see \emph{Appendix S1.3} for details).
         Since cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effects on network robustness are very limited~\citep{Allesina2009functional}, this change will not strongly affect consumers' extinction risk in our simulations.
@@
 
-        The response of the consumer to changes in $f$ can take any form (\emph{Appendix S1}).
+        The response of the consumer to changes in $f$ can take any form (\emph{Appendix S2.1}).
 Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$), the type of response that most accurately captures the secondary extinctions produced by dynamical models (\citealp[]{Eklof2013}).
@@
 	\subsection*{Statistical analysis}
-	% Appendix 3, with numbered subheadings.
 
@@
             These models included only participation in the focal motif ($\rho_{i}$ above) and again used a binomial error distrubution.
-            To aid in interpretation of these single-network models with respect to network size and connectance, we fit an additional set of GLMs including proportion of motif, strength of disturbance, network size, and connectance across all networks (\emph{Appendix S6}).
+            To aid in interpretation of these single-network models with respect to network size and connectance, we fit an additional set of GLMs including proportion of motif, strength of disturbance, network size, and connectance across all networks (\emph{Appendix S4.1}).
 
@@
             As the three-way interaction was significant, we did not simplify the model. 
-            To confirm known relationships between persistence and in-degree or trophic level in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to in-degree or STL, disturbance, and the interaction between them (\emph{Appendix S3}, Eq.
+            To confirm known relationships between persistence and in-degree or trophic level in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to in-degree or STL, disturbance, and the interaction between them (\emph{Appendix S3.1}, Eq.
 S2-3).
@@
             Since a network's motif profile and a species' motif participation reflect the same network as other measures (size, connectance, degree, and trophic level), these measures are not independent.
-            We identified how the frequencies of motifs in the network's motif profile were related to network size, connectance, and their interaction using a set of four GLMs with binomial error distribution (Appendix \emph{S3}, eq.
+            We identified how the frequencies of motifs in the network's motif profile were related to network size, connectance, and their interaction using a set of four GLMs with binomial error distribution (Appendix \emph{S3.2}, eq.
 S4).
             We identified how the frequencies of motifs in a consumer's motif participation vector were related to network size, connectance, and their interaction using a similar set of four GLMs with binomial error distribution.
-            Finally, we fit two additional sets of GLMs (eight total) relating participation in each motif to the consumer's in-degree or trophic level (Appendix \emph{S3}, Eq. S5-6).
+            Finally, we fit two additional sets of GLMs (eight total) relating participation in each motif to the consumer's in-degree or trophic level (Appendix \emph{S3.2}, Eq. S5-6).
 
@@
 
-        A species' probability of persistence (hereafter `persistence') was significantly related to its motif participation (Table S1, \emph{Appendix S4}).
+        A species' probability of persistence (hereafter `persistence') was significantly related to its motif participation (Table S1, \emph{Appendix S5}).
         Across all levels of disturbance, species were more likely to persist if they participated more frequently in the direct competition motif and less frequently in the omnivory motif (Fig.~\ref{fig:prop_lmer_all}).
@@
         Within these overall trends, there was substantial variation in the relationships between persistence and motif participation between networks with different overall network structures (Fig.~\ref{fig:density_prop}).
-        Networks with high and low connectances had different distributions of persistence-motif participation relationships, while network size had a much smaller effect (\emph{Appendix SX}).
+        Networks with high and low connectances had different distributions of persistence-motif participation relationships, while network size had a much smaller effect (\emph{Appendix S4.2}).
         The relationship between participation in the omnivory motif and persistence was especially variable.
@@
 
-        Consumers with higher trophic level (STL) were less likely to persist (Fig.~\ref{fig:lm_CS}, Table S3, \emph{Appendix S6}).
+        Consumers with higher trophic level (STL) were less likely to persist (Fig.~\ref{fig:lm_CS}, Table S3, \emph{Appendix S6.1}).
         The relationship between in-degree and persistence depended on the strength of disturbance: higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.
 
-        The average persistence of all consumers was lower in networks with higher connectance, did not vary significantly with network size (Table S6, \emph{Appendix S10}), and  varied with the frequencies of different motifs in this profile.
-        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:lm_CS}, Table S4, \emph{Appendix S8}).
-        There was no significant interaction between the proportion of any motif in the network profile and level of disturbance (Table S4), \emph{Appendix S8}).
+        The average persistence of all consumers was lower in networks with higher connectance, did not vary significantly with network size (Table S4, \emph{Appendix S6.1}), and  varied with the frequencies of different motifs in this profile.
+        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:lm_CS}, Table S5, \emph{Appendix S6.2}).
+        There was no significant interaction between the proportion of any motif in the network profile and level of disturbance).
 
@@
         Omnivory motif participation generally increased with increasing connectance and network size, but decreased with increasing network size in the most-connected networks (Fig.
-S2 and Table S2, \emph{Appendix S5}).
+S7 and Table S6, \emph{Appendix S7}).
         Participation in the apparent competition and three-species chain motifs decreased strongly with increasing connectance but showed little change with changing network size.

### 2023-05-12 manuscript/supplement_motifs.tex
Updates from Overleaf

@@
 
-    Appendices S1-S3 provide supplemental methods, while appendices S4-S6 provide supplemental results.
+    Appendices S1-S4 provide supplemental methods, while appendices S4-S7 provide supplemental results.
 
     \begin{enumerate}
-
-        \item Calculating persistence
-
-            \begin{itemize}
-                \item Details on functional response of consumers
-                \item Contains Eqn. S1
-            \end{itemize}
 
@@
             \item Conversion of networks to acyclic form
-            \item Comparison of empirical and simulated networks
-            \item Contains Fig. S1
+            \item Comparison of empirical and simulated networks (Fig. S1)
+            \end{itemize}
+
+        \item Calculating persistence
+
+            \begin{itemize}
+                \item Details on functional response of consumers (Eqn. S1)
             \end{itemize}
@@
             \begin{itemize}
-                \item Persistence and other network properties (Eqn. S2-3)
-                \item Motif participation, profiles and network properties (Eqn. S4-6)
+                \item Describes GLMs relating persistence and other network properties (Eqn. S2-3) and GLMs relating motif participation, profiles and network properties (Eqn. S4-6)
+            \end{itemize}
+
+        \item Motif-persistence relationships vary with S,C
+
+            \begin{itemize}
+                \item Describes GLMs summarising how the relationship between motif participation and persistence varies over S and C, to complement the within-network results presented in the main text (Eqn.
+S7)
+                \item Plots of these GLMs for each motif (Figs. S2-5)
             \end{itemize}
@@
                 \item Supplemental results relating species motif participation to persistence and to other network properties (S, C, in-degree, STL)
-                \item Contains Table S1-2
-                \item Contains Fig. S2
+                \item Table S1-2l; Fig. S6
             \end{itemize}
 
-        \item Results: persistence and other network properties
+        \item Supplemental results: persistence and other network properties
 
             \begin{itemize}
-                \item Identifying relationships between persistence and network size, connectance, in-degree, and STL to provide context for motif relationships.
-                \item Contains Tables S3-4
-                \item Contains Fig. S3
+                \item Supplemental results identifying relationships between persistence and network size, connectance, in-degree, and STL to provide context for motif relationships (Tables S3-4).
+                \item Also relating average persistence to network motif profiles (Table S5)
             \end{itemize}
 
-        \item Motif-persistence relationships vary with S,C
+        \item Supplemental results: network motif profiles
 
             \begin{itemize}
-                \item Describes GLMs summarising how the relationship between motif participation and persistence varies over S and C, to complement the within-netowrk results presented in the main text
-                \item Contains Fig. S4
-            \end{itemize}
-
-        \item Results: network motif profiles
-
-            \begin{itemize}
-                \item Relating motif profiles to persistence (see main text for equation, figure) and to network size and connectance
-                \item Contains Tables S5-6
+                \item Relating motif profiles  to network size and connectance (Table S6)
             \end{itemize}
@@
             \begin{itemize}
-                \item Contains Table S7
+                \item Definitions of network terminology used in main text (Table S7)
             \end{itemize}
@@
 \linenumbers
+\section{Network creation}
+
+    \subsection{Generation of simulated networks}
+
+        We simulated a suite of food webs based on the probabilistic niche model, which assigns predator-prey links based on the body-mass ratios between individuals of different species~\citep{Williams2000,Delmas2017}.
+        The meso-scale structure of niche-model networks closely mimics that of empirical food webs~\citep{Stouffer2007}. 
+        To ensure that we captured a variety of realistic community sizes and structures, we generated networks ranging between 50 and 100 species (in steps of 10) with connectance values between 0.02 and 0.2 (in steps of 0.02).
+        The range of network sizes was chosen to reflect moderately well-sampled empirical webs while working within our computational limits, while the range of connectance values was chosen to cover that observed in most empirical food webs~\citep{Dunne2002e}.
+        We generated a total of 100 networks with each combination of parameters, for a total of 6000 networks. 
+        All networks were generated using the function ``nichemodel'' within the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}. 
+        If a simulated network contained any disconnected species (species without predators or prey) or disconnected components (a group of species connected among themselves but not to the rest of the network), the network was rejected and a new network simulated.
+        Finally, networks where the path lengths between each species and a basal resource could not be resolved (i.e., trophic levels were undefined) were rejected and new networks simulated.
+
+        As a further check that our simulated networks are realistic (e.g., could plausibly retain all species), we simulated community dynamics for 500 timesteps and discarded any networks which did not retain all species.
+        We simulated dynamics using the function ``simulate'' from the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
+        This function uses Lotka-Volterra predator-prey models including density dependence and type 2 functional responses for all species (please see~\citet{Delmas2017} for full details).
+        All non-basal species were designated as vertebrates to ensure a good match between metabolic and predator-prey body-mass ratio values.
+Metabolic rates in the Lotka-Volterra model are based on each species' body mass (i.e.
+mass of a single individual).
+        We assigned relative body masses based on each species' trophic level, which was, in turn, calculated based on the food-web structure provided by the niche model. 
+        After basal species were assigned a body mass of 1, we used a predator-prey body-mass ratio of 3.065 to calculate the relative body masses of higher trophic levels. 
+        We selected this ratio based on the estimate for vertebrates (averaged across ecosystem and metabolic types) in~\citet{Brose2006}. 
+        We excluded reported body-mass ratios for invertebrates as these could include parasites and parasitoids, which are generally smaller than their prey, and because interactions among vertebrates are better represented in the food-web literature than interactions involving invertebrates.
+
+        The persistence of each species in our simulated networks also depends on its population biomass. 
+        We randomly assigned initial population biomasses (i.e. cumulative biomass across all individuals of a species) for each species from a uniform distribution [0,1]. 
+        Note that population biomasses and individual body masses are not calculated on the same scale. 
+        We then simulated community dynamics for 1000 time steps to obtain an `equilibrium' community. 
+        To ensure that species did not `recover' from unrealistically low biomasses during the simulation, we considered a species extinct if it dropped below an arbitrary threshold biomass of 1$\times10^{-5}$.
+        We rejected any network where one or more species dropped below this biomass threshold. 
+        Consumers were assumed to have no preferences such that the consumption rate $w_{ij}$ of predator $i$ eating prey $j$ is equal to $\frac{1}{n}$, where $n$ is the number of prey for predator $i$.
+        If the network did not retain all species for 1000 time steps, a new set of initial population biomasses was applied and the simulation repeated.
+        If a set of biomasses where all species persist still had not been reached after 100 sets of randomly-assigned initial biomasses, we discarded the network and simulated another to replace it.
+
+    \subsection{Filtering of biologically unrealistic networks}
+
+        After simulating our initial niche-model networks, we filtered networks for biological plausibility.
+        Specifically, we removed any network containing disconnected components (species or groups of species not connected to the rest of the network) or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
+        Any removed network was replaced with a fresh simulation and re-checked until 100 realistic networks for each size-connectance combination had been obtained.
+
+    \subsection{Conversion of networks to acyclic form}
+
+        A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}.
+        All persistence calculations therefore begin by determining the status of the basal resources (who do not depend on other species).
+        Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on basal resources), then herbivores, and so on up the network.
+
+        While niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
+        We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009functional}.
+        This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009functional}.
+        These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
+        Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citet{Tarjan1972} and \citet{Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
+        After these steps, calculations of persistence probabilities can begin.
+
+    \subsection{Comparison of empirical and simulated networks}
+
+        Although previous work has shown that the niche model replicates the motif structure of empirical networks~\citep{Stouffer2005a,Stouffer2006}, we additionally provide a comparison of the motif profiles and consumer motif participation for our simulated networks and the highest-quality empirical networks we could find within the same size range.
+        For this comparison, we obtained all networks containing 50-100 species from \citet{globalwebdb}.
+        We then removed networks where sets of predators and prey were completely different (i.e., bipartite networks) or links were not clear (i.e., +/- effects rather than predation specifically).
+        This yielded five vertebrate-dominated food webs describing the California coast in different years, four invertebrate-dominated food webs describing mainland US streams, 23 invertebrate-dominated food webs describing New Zealand streams, and five other invertebrate-dominated food webs.
+        This final set of webs had 50-77 species and connectances (after rendering the networks acyclic) of 0.095-0.227 - covering the upper half of the connectance range we simulate, as well as slightly higher connectances.
+        This may be partly due to relatively poor resolution of basal resources in these empirical webs: aggregation of many taxa into nodes such as `phytoplankton' will tend to decrease the number of species but leave the number of links unchanged, raising connectance.
+        We note that most of these webs were compiled by just three research groups.
+        Moreover, many of the webs are repeated samples of the same site  or closely-grouped sites.
+        This means that these empirical networks cannot be considered fully independent samples.
+
+        \begin{figure}
+            \centering
+            \includegraphics[width=\textwidth]{figures/motif_profiles_participation_vs_empirical.eps}
+            \caption{\textbf{A)} Proportions of acyclic motifs in the motif profiles of empirical and simulated networks.
+Mean proportions ($\pm$SD) in the simulated networks are indicated in black, with the range of proportions in the simulated networks indicated by black boxes.
+\textbf{B)} Proportions of acyclic motifs in the motif participation vectors of consumers in simulated and empirical networks.}
+            \label{empirical_compare}
+        \end{figure}
+
+    \subsubsection{Comparing motif profiles}
+
+      The proportions of omnivory and apparent competition in the motif profiles of simulated networks matched those of the empirical webs fairly well, but many of the empirical webs had higher proportions of direct competition and lower proportions of the three-species chain than any simulated web.
+      The empirical webs with high proportions of direct competition tend to be invertebrate-dominated webs representing small streams.
+      High proportions of direct competition in these system suggest large numbers of fairly generalist herbivore/detritivores, with relatively few secondary consumers present to create chains.
+
+      A few webs had lower proportions of apparent competition than any simulated web but similar proportions of the other motifs.
+      Most of these webs represented the California coast in different years (2003-2007).
+      Basal resources in these webs were highly aggregated into only two size classes of phytoplankton and three types of detritus.
+      This aggregation at the basal level leads to artificial ``specialisation'' in consumers with low trophic levels who may consume, for example, many species of phytoplankton that are grouped into a single size class.
+      As the basal resources in our simulated networks are not aggregated, our webs feature higher proportions of apparent competition as seen in the other empirical webs.
+
+    \subsubsection{Comparing motif participation}
+
+      Despite the differences in network motif profiles between some of the empirical food webs and the simulated webs, we note that the motif participation of consumers in the simulated webs entirely spans the range of motif participation for consumers in the empirical webs.
+      Consumers in the NZ stream food webs had higher proportions of direct competition and lower proportions of apparent competition than the mean of the simulated webs, consistent with the network motif profiles.
+      Nonetheless, the greater differences in network profiles than consumer motif participation suggests that the largest differences between the simulated and empirical networks are in the roles of basal resources.
+      In empirical webs, basal resources are often highly aggregated. 
+      Improved resolution at the basal level should improve the correspondence between simulated and empirical webs and allow better modelling of empirical communities.
+
+\clearpage
+
 \section{Calculating persistence}
@@
 
-\section{Network creation}
-
-    \subsection*{Generation of simulated networks}
-
-        We simulated a suite of food webs based on the probabilistic niche model, which assigns predator-prey links based on the body-mass ratios between individuals of different species~\citep{Williams2000,Delmas2017}.
-        The meso-scale structure of niche-model networks closely mimics that of empirical food webs~\citep{Stouffer2007}. 
-        To ensure that we captured a variety of realistic community sizes and structures, we generated networks ranging between 50 and 100 species (in steps of 10) with connectance values between 0.02 and 0.2 (in steps of 0.02).
-        The range of network sizes was chosen to reflect moderately well-sampled empirical webs while working within our computational limits, while the range of connectance values was chosen to cover that observed in most empirical food webs~\citep{Dunne2002e}.
-        We generated a total of 100 networks with each combination of parameters, for a total of 6000 networks. 
-        All networks were generated using the function ``nichemodel'' within the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}. 
-        If a simulated network contained any disconnected species (species without predators or prey) or disconnected components (a group of species connected among themselves but not to the rest of the network), the network was rejected and a new network simulated.
-        Finally, networks where the path lengths between each species and a basal resource could not be resolved (i.e., trophic levels were undefined) were rejected and new networks simulated.
-
-        As a further check that our simulated networks are realistic (e.g., could plausibly retain all species), we simulated community dynamics for 500 timesteps and discarded any networks which did not retain all species.
-        We simulated dynamics using the function ``simulate'' from the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}.
-        This function uses Lotka-Volterra predator-prey models including density dependence and type 2 functional responses for all species (please see~\citet{Delmas2017} for full details).
-        All non-basal species were designated as vertebrates to ensure a good match between metabolic and predator-prey body-mass ratio values.
-Metabolic rates in the Lotka-Volterra model are based on each species' body mass (i.e.
-mass of a single individual).
-        We assigned relative body masses based on each species' trophic level, which was, in turn, calculated based on the food-web structure provided by the niche model. 
-        After basal species were assigned a body mass of 1, we used a predator-prey body-mass ratio of 3.065 to calculate the relative body masses of higher trophic levels. 
-        We selected this ratio based on the estimate for vertebrates (averaged across ecosystem and metabolic types) in~\citet{Brose2006}. 
-        We excluded reported body-mass ratios for invertebrates as these could include parasites and parasitoids, which are generally smaller than their prey, and because interactions among vertebrates are better represented in the food-web literature than interactions involving invertebrates.
-
-        The persistence of each species in our simulated networks also depends on its population biomass. 
-        We randomly assigned initial population biomasses (i.e. cumulative biomass across all individuals of a species) for each species from a uniform distribution [0,1]. 
-        Note that population biomasses and individual body masses are not calculated on the same scale. 
-        We then simulated community dynamics for 1000 time steps to obtain an `equilibrium' community. 
-        To ensure that species did not `recover' from unrealistically low biomasses during the simulation, we considered a species extinct if it dropped below an arbitrary threshold biomass of 1$\times10^{-5}$.
-        We rejected any network where one or more species dropped below this biomass threshold. 
-        Consumers were assumed to have no preferences such that the consumption rate $w_{ij}$ of predator $i$ eating prey $j$ is equal to $\frac{1}{n}$, where $n$ is the number of prey for predator $i$.
-        If the network did not retain all species for 1000 time steps, a new set of initial population biomasses was applied and the simulation repeated.
-        If a set of biomasses where all species persist still had not been reached after 100 sets of randomly-assigned initial biomasses, we discarded the network and simulated another to replace it.
-
-    \subsection*{Filtering of biologically unrealistic networks}
-
-        After simulating our initial niche-model networks, we filtered networks for biological plausibility.
-        Specifically, we removed any network containing disconnected components (species or groups of species not connected to the rest of the network) or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
-        Any removed network was replaced with a fresh simulation and re-checked until 100 realistic networks for each size-connectance combination had been obtained.
-
-    \subsection*{Conversion of networks to acyclic form}
-
-        A Bayesian food web depicts probabilistic relationships among a set of species where each species' probability of persistence depends upon the probabilities of its resource species persisting~\citep{Jensen_Nielsen,Eklof2013}.
-        All persistence calculations therefore begin by determining the status of the basal resources (who do not depend on other species).
-        Calculations continue in a strictly bottom-up manner with primary consumers (who depend only on basal resources), then herbivores, and so on up the network.
-
-        While niche-model food webs can contain cycles (e.g., species A eats B, B eats C, and C eats A), such cycles make it impossible to calculate persistence in the Bayesian network framework~\citep{Tarjan1972}.
-        We therefore removed any cycles within each network by removing the link in each cycle which contributed least to the robustness of the food web, following~\citet{Allesina2009functional}.
-        This is achieved by first finding the set of resources for each consumer and then removing the consumer-resource connections which have the lowest eigenvalue centrality~\citep{Allesina2009functional}.
-        These links have the least effect on the overall stability of the network; removing them creates an acyclic network with very similar properties to the original network.
-        Next, we ordered the species from lowest to highest trophic level using a topological sorting routine following \citet{Tarjan1972} and \citet{Allesinaetal2005}, ensuring that probability calculations follow a strict bottom-up order.
-        After these steps, calculations of persistence probabilities can begin.
-
-    \subsection*{Comparison of empirical and simulated networks}
-
-        Although previous work has shown that the niche model replicates the motif structure of empirical networks~\citep{Stouffer2005a,Stouffer2006}, we additionally provide a comparison of the motif profiles and consumer motif participation for our simulated networks and the highest-quality empirical networks we could find within the same size range.
-        For this comparison, we obtained all networks containing 50-100 species from \citet{globalwebdb}.
-        We then removed networks where sets of predators and prey were completely different (i.e., bipartite networks) or links were not clear (i.e., +/- effects rather than predation specifically).
-        This yielded five vertebrate-dominated food webs describing the California coast in different years, four invertebrate-dominated food webs describing mainland US streams, 23 invertebrate-dominated food webs describing New Zealand streams, and five other invertebrate-dominated food webs.
-        This final set of webs had 50-77 species and connectances (after rendering the networks acyclic) of 0.095-0.227 - covering the upper half of the connectance range we simulate, as well as slightly higher connectances.
-        This may be partly due to relatively poor resolution of basal resources in these empirical webs: aggregation of many taxa into nodes such as `phytoplankton' will tend to decrease the number of species but leave the number of links unchanged, raising connectance.
-        We note that most of these webs were compiled by just three research groups.
-        Moreover, many of the webs are repeated samples of the same site  or closely-grouped sites.
-        This means that these empirical networks cannot be considered fully independent samples.
-
-        \begin{figure}
-            \centering
-            \includegraphics[width=\textwidth]{figures/motif_profiles_participation_vs_empirical.eps}
-            \caption{\textbf{A)} Proportions of acyclic motifs in the motif profiles of empirical and simulated networks.
-Mean proportions ($\pm$SD) in the simulated networks are indicated in black, with the range of proportions in the simulated networks indicated by black boxes.
-\textbf{B)} Proportions of acyclic motifs in the motif participation vectors of consumers in simulated and empirical networks.}
-            \label{empirical_compare}
-        \end{figure}
-
-    \subsubsection*{Comparing motif profiles}
-
-      The proportions of omnivory and apparent competition in the motif profiles of simulated networks matched those of the empirical webs fairly well, but many of the empirical webs had higher proportions of direct competition and lower proportions of the three-species chain than any simulated web.
-      The empirical webs with high proportions of direct competition tend to be invertebrate-dominated webs representing small streams.
-      High proportions of direct competition in these system suggest large numbers of fairly generalist herbivore/detritivores, with relatively few secondary consumers present to create chains.
-
-      A few webs had lower proportions of apparent competition than any simulated web but similar proportions of the other motifs.
-      Most of these webs represented the California coast in different years (2003-2007).
-      Basal resources in these webs were highly aggregated into only two size classes of phytoplankton and three types of detritus.
-      This aggregation at the basal level leads to artificial ``specialisation'' in consumers with low trophic levels who may consume, for example, many species of phytoplankton that are grouped into a single size class.
-      As the basal resources in our simulated networks are not aggregated, our webs feature higher proportions of apparent competition as seen in the other empirical webs.
-
-    \subsubsection*{Comparing motif participation}
-
-      Despite the differences in network motif profiles between some of the empirical food webs and the simulated webs, we note that the motif participation of consumers in the simulated webs entirely spans the range of motif participation for consumers in the empirical webs.
-      Consumers in the NZ stream food webs had higher proportions of direct competition and lower proportions of apparent competition than the mean of the simulated webs, consistent with the network motif profiles.
-      Nonetheless, the greater differences in network profiles than consumer motif participation suggests that the largest differences between the simulated and empirical networks are in the roles of basal resources.
-      In empirical webs, basal resources are often highly aggregated. 
-      Improved resolution at the basal level should improve the correspondence between simulated and empirical webs and allow better modelling of empirical communities.
-
-\clearpage
-
 \section{Supplemental statistical analysis}
 
-    \textbf{Persistence, degree, and trophic level}
+    \subsection{Persistence, degree, and trophic level}
 
@@
 
-    \textbf{Motif participation, profiles and other network properties}
+    \subsection{Motif participation, profiles and other network properties}
 
@@
 
-\section{Results: motif participation}
+\section{Motif-persistence relationships vary with S,C}
 
-    \subsection{Participation and persistence}
-
-    % Updated with new random effect, converted to data-scale
-    \begin{table}[ht!]
-        \centering
-        \caption{Coefficients ($\beta$) for general linear models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction  (Eq.
-1, main text).
-Coefficients in \textbf{bold} are significant ($\alpha$\textless0.005).
-Models were fit using a binomial error distribution and logit link function.
-Predictors were centered and scaled before model fitting.
-We also provide McFadden's pseudo-$R^2$ for each model.}
-        \label{tab:proportion}                \footnotesize
-        \begin{tabular}{c|c c c c | c }
-        Motif & Intercept & Proportion & Disturbance & Interaction & pseudo-$R^2$  \\
-        \hline
-        3-sp Chain & \textbf{2.90} & \textbf{-0.204} & \textbf{-5.42} & \textbf{0.138} & 0.895 \\
-        App. Com. & \textbf{0.317} & -2.63$\times10^{-3}$ & \textbf{-5.98478} & \textbf{1.95$\times10^{-2}$} & 0.895 \\
-        Dir. Com. & \textbf{0.317} & \textbf{3.03$\times10^{-2}$} & \textbf{-0.599} & -1.67$\times10^{-3}$ &  0.897 \\
-        Omnivory & \textbf{0.317} & \textbf{-1.81$\times10^{-2}$} & \textbf{-0.598} & \textbf{-7.76$\times10^{-3}$} & 0.895 \\
-        \end{tabular}
-    \end{table}
-
-\subsection{Participation and network properties}
-
-    \begin{table}[hb!]
-        \centering
-        \caption{Motif participation varies with size and connectance. Here we give the range and mean proportion of each motif in species' motif participation vectors. 
-        We also show the coefficients ($\beta$) for network size (S), connectance (C), and their interaction in general linear models predicting the proportion of each motif in a participation vector.
-        Coefficients in \textbf{bold} are significant ($\alpha$=0.05).
-        We also show the coefficients for similar models relating the proportion of each motif to in-degree or trophic level.
-        All models were fit with a binomial error distribution and logit link function.
-        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
-        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
-        \label{tab:partic_vs_SC}   
-        \footnotesize
-        \begin{tabular}{c|c c c | c c c | c | c}
-            Motif & Min & Max & Mean & S & C & Int. & in-Deg & TL \\
-            \hline
-            3-sp Chain & 0 & 1 & 0.249 & \textbf{1.20$\times10^{-3}$} &  -0.652 & \textbf{-1.77$\times10^{-2}$} & \textbf{-6.99$\times10^{-3}$} & \textbf{9.84$\times10^{-2}$} \\
-            App. Comp & 0.250 & 0.765 & 0.422 & \textbf{-1.82$\times10^{-3}$} & \textbf{-3.18} & 6.92$\times10^{-3}$ & \textbf{-8.76$\times10^{-3}$} & \textbf{2.40$\times10^{-1}$}\\
-            Dir. Comp & 0.0501 & 0.359 & 0.179 & -1.18$\times10^{-3}$ & \textbf{-4.50} & \textbf{4.79$\times10^{-2}$} & \textbf{-1.43$\times10^{-3}$} & \textbf{-2.52$\times10^{-1}$} \\
-            Omnivory & 0.006 & 0.346 & 0.164 & \textbf{3.64$\times10^{-3}$} & \textbf{12.5} & \textbf{-4.19$\times10^{-2}$} & \textbf{2.29$\times10^{-2}$} & \textbf{-2.27$\times10^{-1}$}\\   
-            \hline
-            \end{tabular}
-            \end{table}
-
-    % \begin{figure}[hb!]
-    %     \centering
-    %     \includegraphics[width=.9\textwidth]{figures/participation_vs_SC.eps}
-    %     \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).
-We show relationships for the smallest and largest values of network size (S) and connectance (C).}
-    %     \label{fig:roles_vs_SC}
-    % \end{figure}
-
-\clearpage
-
-\section{Results: persistence and other network properties}
-
-    % \begin{figure}[ht!]
-    %     \centering
-    %     \includegraphics[width=0.9\textwidth]{figures/persistence_vs_SC_lm.eps}
-    %     \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.}
-    %     \label{fig:lm_CS}
-    % \end{figure}
-
-    % Updated to use mean rather than all species.
-    \begin{table}[hb!]
-        \caption{Mean persistence of consumers within a network depends on connectance (C) and the level of disturbance to basal resources (D), but not network size (S) or interactions between properties.
-We show effect sizes ($\beta$) and $p$-values for each coefficient, except the intercept (see Appendix S3 for details).
-McFadden's pseudo-$R^2$=0.966.
-}
-        \label{tab:per_vs_SC}
-        \centering
-        \footnotesize
-        \begin{tabular}{c| c c c c c c c}
-              & S & C & D & S:C & S:D & C:D & S:C:D \\
-            \hline
-             $\beta$ &	 -7.21$\times10^{-4}$  	& \textbf{-6.12$\times10^{-2}$}	& \textbf{-0.596} &	 -4.48$\times10^{-3}$ &	 -1.63$\times10^{-2}$ 	& -4.52$\times10^{-3}$ &	 -1.35$\times10^{-3}$ \\
-             $p$-value &		 0.964 &	 \textless0.001 &	 \textless0.001 &	 0.777 &	 0.316 &	 0.781 &	 0.934 \\
-                    \end{tabular}
-                \end{table}
-
-    % Updated with network-level random effect
-    \begin{table}[hb!]
-        \caption{Coeffcients ($\beta$) for general linear mixed-effect models of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction, as well as a random effect of global network properties (size and connectance).
-The model for STL was singular when including the random effect, so we removed this term and re-fit the model as a general linear model with fixed effects only.
-All models had binomial error distribution.
-Predictors were centered and scaled before model fitting.
-All $p$-values were \textless0.001.
-The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.
-We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for the in-degree model and the pseudo-$R^2$ for the STL model.}
-        \label{tab:per_vs_TLdeg}
-        \centering
-        \begin{tabular}{l|c  c  c | c c  }
-            Simple role measure & Role & Disturbance & Interaction & R$^2_M$ & R$^2_C$ \\
-            \hline
-            STL & \textbf{-0.0977} & \textbf{-0.600} & \textbf{0.0302} & \textbf{0.920} & NA \\
-            in-degree & \textbf{-1.30$\times10^{-3}$} & \textbf{-0.139} & \textbf{-9.08$\times10^{-3}$} & 0.903 & 0.914 \\
-        \end{tabular}
-    \end{table}
-
-    % \begin{figure}[hb!]
-    %  \centering
-    %  \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP0.eps}
-    %  \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
-The panels show final persistence for consumers with different shortest trophic level, for increasing connectance (y-axis) and network size (x-axis).
-Persistence decreases with darker color in the heat map.}
-    %  \label{fig:heatmap_stl_BP0}
-    % \end{figure}
-
-    % \begin{figure}[hb!]
-    %  \centering
-    %  \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP1.eps}
-    %  \caption{Basal resources experience a disturbance increasing their probability of extinction to $\pi = 0.5$.
-All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
-The panels show final persistence for consumers with different shortest trophic level, for increasing connectance (y-axis) and network size (x-axis).
-Persistence decreases with darker color in the heat map.}
-    %  \label{fig:heatmap_stl_BP1}
-    % \end{figure}
-
-\clearpage
-
-\section{Motif-persistence relationships vary with S,C}
+    \subsection{Supplemental methods}
 
@@
             If a smaller term was included in a significant larger term (e.g., a non-significant two-way interaction involving terms in a significant three-way interaction) then the smaller term was retained.
+
+    \subsection{Supplemental results}
 
@@
 
-\section{Supplemental results: network motif profiles}
+\section{Supplemental results: motif participation}
+
+    \subsection{Participation and persistence}
+
+    % Updated with new random effect, converted to data-scale
+    \begin{table}[ht!]
+        \centering
+        \caption{Coefficients ($\beta$) for general linear models relating persistence to the proportion of a species' role made up by the focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction  (Eq.
+1, main text).
+Coefficients in \textbf{bold} are significant ($\alpha$\textless0.005).
+Models were fit using a binomial error distribution and logit link function.
+Predictors were centered and scaled before model fitting.
+We also provide McFadden's pseudo-$R^2$ for each model.}
+        \label{tab:proportion}                \footnotesize
+        \begin{tabular}{c|c c c c | c }
+        Motif & Intercept & Proportion & Disturbance & Interaction & pseudo-$R^2$  \\
+        \hline
+        3-sp Chain & \textbf{2.90} & \textbf{-0.204} & \textbf{-5.42} & \textbf{0.138} & 0.895 \\
+        App. Com. & \textbf{0.317} & -2.63$\times10^{-3}$ & \textbf{-5.98478} & \textbf{1.95$\times10^{-2}$} & 0.895 \\
+        Dir. Com. & \textbf{0.317} & \textbf{3.03$\times10^{-2}$} & \textbf{-0.599} & -1.67$\times10^{-3}$ &  0.897 \\
+        Omnivory & \textbf{0.317} & \textbf{-1.81$\times10^{-2}$} & \textbf{-0.598} & \textbf{-7.76$\times10^{-3}$} & 0.895 \\
+        \end{tabular}
+    \end{table}
+
+\subsection{Participation and network properties}
+
+    \begin{figure}[ht!]
+        \centering
+        \includegraphics[width=\textwidth]{figures/roles_vs_SC_all.eps}
+        \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).}
+        \label{fig:motifs_vs_TL_and_deg}
+    \end{figure}
+
+    \begin{table}[hb!]
+        \centering
+        \caption{Motif participation varies with size and connectance. Here we give the range and mean proportion of each motif in species' motif participation vectors. 
+        We also show the coefficients ($\beta$) for network size (S), connectance (C), and their interaction in general linear models predicting the proportion of each motif in a participation vector.
+        Coefficients in \textbf{bold} are significant ($\alpha$=0.05).
+        We also show the coefficients for similar models relating the proportion of each motif to in-degree or trophic level.
+        All models were fit with a binomial error distribution and logit link function.
+        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
+        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
+        \label{tab:partic_vs_SC}   
+        \footnotesize
+        \begin{tabular}{c|c c c | c c c | c | c}
+            Motif & Min & Max & Mean & S & C & Int. & in-Deg & TL \\
+            \hline
+            3-sp Chain & 0 & 1 & 0.249 & \textbf{1.20$\times10^{-3}$} &  -0.652 & \textbf{-1.77$\times10^{-2}$} & \textbf{-6.99$\times10^{-3}$} & \textbf{9.84$\times10^{-2}$} \\
+            App. Comp & 0.250 & 0.765 & 0.422 & \textbf{-1.82$\times10^{-3}$} & \textbf{-3.18} & 6.92$\times10^{-3}$ & \textbf{-8.76$\times10^{-3}$} & \textbf{2.40$\times10^{-1}$}\\
+            Dir. Comp & 0.0501 & 0.359 & 0.179 & -1.18$\times10^{-3}$ & \textbf{-4.50} & \textbf{4.79$\times10^{-2}$} & \textbf{-1.43$\times10^{-3}$} & \textbf{-2.52$\times10^{-1}$} \\
+            Omnivory & 0.006 & 0.346 & 0.164 & \textbf{3.64$\times10^{-3}$} & \textbf{12.5} & \textbf{-4.19$\times10^{-2}$} & \textbf{2.29$\times10^{-2}$} & \textbf{-2.27$\times10^{-1}$}\\   
+            \hline
+            \end{tabular}
+            \end{table}
+
+\clearpage
+
+\section{Supplemental results: persistence and other network properties}
+
+    \subsection{Persistence and non-motif measures}
+    % \begin{figure}[ht!]
+    %     \centering
+    %     \includegraphics[width=0.9\textwidth]{figures/persistence_vs_SC_lm.eps}
+    %     \caption{Persistence decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance.}
+    %     \label{fig:lm_CS}
+    % \end{figure}
+
+    % Updated with network-level random effect
+    \begin{table}[hb!]
+        \caption{Coeffcients ($\beta$) for general linear mixed-effect models of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction, as well as a random effect of global network properties (size and connectance).
+The model for STL was singular when including the random effect, so we removed this term and re-fit the model as a general linear model with fixed effects only.
+All models had binomial error distribution.
+Predictors were centered and scaled before model fitting.
+All $p$-values were \textless0.001.
+The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.
+We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for the in-degree model and the pseudo-$R^2$ for the STL model.}
+        \label{tab:per_vs_TLdeg}
+        \centering
+        \begin{tabular}{l|c  c  c | c c  }
+            Simple role measure & Role & Disturbance & Interaction & R$^2_M$ & R$^2_C$ \\
+            \hline
+            STL & \textbf{-0.0977} & \textbf{-0.600} & \textbf{0.0302} & \textbf{0.920} & NA \\
+            in-degree & \textbf{-1.30$\times10^{-3}$} & \textbf{-0.139} & \textbf{-9.08$\times10^{-3}$} & 0.903 & 0.914 \\
+        \end{tabular}
+    \end{table}
+
+    % Updated to use mean rather than all species.
+    \begin{table}[hb!]
+        \caption{Mean persistence of consumers within a network depends on connectance (C) and the level of disturbance to basal resources (D), but not network size (S) or interactions between properties.
+We show effect sizes ($\beta$) and $p$-values for each coefficient, except the intercept (see Appendix S3 for details).
+McFadden's pseudo-$R^2$=0.966.
+}
+        \label{tab:per_vs_SC}
+        \centering
+        \footnotesize
+        \begin{tabular}{c| c c c c c c c}
+              & S & C & D & S:C & S:D & C:D & S:C:D \\
+            \hline
+             $\beta$ &	 -7.21$\times10^{-4}$  	& \textbf{-6.12$\times10^{-2}$}	& \textbf{-0.596} &	 -4.48$\times10^{-3}$ &	 -1.63$\times10^{-2}$ 	& -4.52$\times10^{-3}$ &	 -1.35$\times10^{-3}$ \\
+             $p$-value &		 0.964 &	 \textless0.001 &	 \textless0.001 &	 0.777 &	 0.316 &	 0.781 &	 0.934 \\
+                    \end{tabular}
+                \end{table}
+
+    % \begin{figure}[hb!]
+    %  \centering
+    %  \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP0.eps}
+    %  \caption{All species across all levels experience a risk of going extinct due to causes not related to the web of $\pi = 0.1$.
+The panels show final persistence for consumers with different shortest trophic level, for increasing connectance (y-axis) and network size (x-axis).
+Persistence decreases with darker color in the heat map.}
+    %  \label{fig:heatmap_stl_BP0}
+    % \end{figure}
+
+    % \begin{figure}[hb!]
+    %  \centering
+    %  \includegraphics[width=.9\textwidth]{figures/heatmap_STL_allCS_BP1.eps}
+    %  \caption{Basal resources experience a disturbance increasing their probability of extinction to $\pi = 0.5$.
+All consumer species go extinct due to causes not related to the web itself with probability $\pi = 0.1$.
+The panels show final persistence for consumers with different shortest trophic level, for increasing connectance (y-axis) and network size (x-axis).
+Persistence decreases with darker color in the heat map.}
+    %  \label{fig:heatmap_stl_BP1}
+    % \end{figure}
 
@@
     \end{table}
+
+\clearpage
+
+\section{Supplemental results: network motif profiles}
 

### 2023-05-19 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
     Motifs are unique patterns of $n$ interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
+    Some motifs are over-represented in stable food webs, suggesting that participation in these motifs provides stability (refs).
+    Indeed, dynamic simulations suggest that participation in these motifs can provide greater persistence after a disturbance~\citep{Cirtwill2022Oikos}.
     The vector of frequencies with which a species appears in each motif --its \emph{motif participation}-- describes the species' position in the meso-scale structure of a food web and, like degree and trophic level, is easily calculated once the structure of the network is known.
@@
     Two species might both have two prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
-    Importantly, since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
+    Since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
     This extra information on the focal species' interaction partners may thus provide extra insight when predicting extinction risk.
@@
     Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.
-    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey indicates a generalist predator (which may be able persist following reduction or loss of one prey) but can capture indirect effects between prey that mae the predator more likely to go extinct~\citep{Melian2002}.
+    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey.
+The motif indicates a generalist predator (which may be able persist following reduction or loss of one prey) but can capture indirect effects between prey that make the predator more likely to go extinct~\citep{Melian2002}.
     Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator. 

### 2023-06-04 manuscript/JAE_reply.tex
SI ready to submit

@@
 
-    [[Some nice closing?]]
+    Thank you for your time and consideration of our manuscript.
 
@@
 
-            \textbf{R:} Yes, a binomial error distribution would be more appropriate.
+            \textbf{R:} Yes, we agree with the Reviewer that a binomial error distribution would be more appropriate.
 We have corrected the error and updated our results accordingly and thank the Reviewer for pointing this out.
@@
 
-                \textbf{R:} If the Reviewer is thinking of EcoPath with EcoSim, that type of model only works with quite a small food web with very well-understood interaction strengths and initial population sizes.
-We are not aware of any dynamical model that is fast over a large food web - certainly this was not the case when some of the authors conducted dynamical simulations of webs with 50-100 species.
+                \textbf{R:} We do think that our statement here is valid.
+Topological models for simulating species extinctions are only taking the food web structure in to account, and a secondary extinction will occur when a consumer lost all its resources.
+So, in comparison to topological models simulating the biomass (or similar) dynamics between many species will require both more parameters and simulation time.
+If we compare to for example the ATN-model requires parameters describing each species body mass and several parameters for defining for example attack rates and handling times.
+Similarly, models like EcoPath with EcoSim, is limited to smaller food webs with well-understood interaction strengths and initial population sizes and as such require more parameters which even in simple cases are more complex than topological models.
+However, these more complex models can model more complex mechanisms than what topological models could.
+
+                %We are not aware of any dynamical model that is fast over a large food web - certainly this was not the case when some of the authors conducted dynamical simulations of webs with 50-100 species.
 [[Anyone have a better answer?
@@
 Our results, which deal specifically with persistence after bottom-up disturbance, add to that picture of complexity without contradicting earlier findings.
-Note that we do not say that omnivory is \emph{never} beneficial - only that in our networks, in response to our focal disturbance, is is generally detrimental.
+Note that we do not say that omnivory is never beneficial - only that in our networks, in response to our focal disturbance, is is generally detrimental.
 We also note that our results show apparent competition as associated with less persistence when disturbance is weak, which is compatible with the possibility that the Reviewer mentions above (though we do not specifically model prey switching, which seems to be what the Reviewer refers to).
-In the revision we have expanded our discussion on how the strength and direction of motif participation-persistence relationships vary network-to-network (especially in networks with different connectance).
+In the revision, we have expanded our discussion on how the strength and direction of motif participation-persistence relationships vary network-to-network (especially in networks with different connectance).
 We hope that this addition provides some of the nuance the Reviewer missed from the previous revision.
@@
 However, since the Bayesian network framework does not include top-down effects any comment we might make about how top-down effects might affect our results would necessarily be pure speculation.
-Instead, we have added a reference to another paper exploring the relationship between motif roles and persistence in a dynamical context.
+Instead, we have added a reference to another paper (quoted below) exploring the relationship between motif roles and persistence in a dynamical context.
 That paper, due to the computational limits inherent in the dynamical approach, uses only a single-species disturbance.
 It does, however, include top-down as well as bottom-up effects and may provide the information the Reviewer is seeking.
+
+                \begin{quotation}
+
+                    There are limitations to the approach presented here.
+                    Most importantly, the Bayesian network framework operates on a strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}.
+                    Here our focus was to analyze bottom-up affects following disturbances to the basal resources, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
+                    If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
+                    For example, although a recent study using a dynamical model to relate motif participation to persistence after a disturbance found similar trends for omnivory, participation in the direct competition motif was not universally positive~\citep{Cirtwill2021_inprep}.
+                    In particular, species which frequently appeared in the prey position of the direct competition motif tended to have shorter persistence times after disturbance in that study - possibly due to top-down effects from the pair of predators involved.
+                    However, as these two studies applied different types of disturbances, further study would be required to disentangle the effects of top-down and bottom-up effects on species with different motif participation, and how each relates to the type and strength of disturbance.
+
+                \end{quotation}
 

### 2023-06-04 manuscript/reorder_main_text.tex
SI ready to submit

@@
 
+% Now 286 words. Max 350. 
     \indent{1.
-Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.} \\
-    2. A species' position in a food web (e.g., its trophic level or number of prey) is known to affect its ability to persist following disturbance.
+Loss of species in food webs can set in motion a cascade of additional, secondary,  extinctions.A species' position in a food web (e.g., its trophic level or number of interactions) is known to affect its ability to persist following disturbance.
+Theses simple measures, however, offer only a coarse description of how species fit into their community.
+Therefore one would expect that more detailed structure measures, such as a species' participation in three-species motifs (meso-scale structures which provide information on a species' direct and indirect interactions) will also be related to its probability of persistence.\\
+    2.
+Disturbances, such as species extinctions, affecting the basal resources have particularly strong effects on the rest of the food web.
+However, how the disturbance branch out and affect consumer persistence is depending on the structural pattern of species interactions in several steps.
+The magnitude, e.g.
+proportion of basal resources lost, will likely also affect the outcome.\\
     3.
-We therefore expect that a species' participation in three-species motifs --meso-scale structures which provide information on a species' direct and indirect interactions-- will also be related to its probability of persistence.\\
+Here, we analyse whether a consumer's risk of going secondary extinct after removal of basal resources depends on the consumer's motif participation and how this relationship varies with the severity of disturbance.\\
     4.
@@
     6.
-Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities, but must be considered in the context of network structure as a whole and of the specific disturbance applied.
+Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities, but must be considered in the context of network structure as a whole and of the specific disturbance applied}
 
@@
     These species are more vulnerable because they are impacted by the loss of any species along a long path from basal resources or because they have few alternative resources if a prey species is lost.
-    Similarly, a species' participation in \emph{meso-scale motifs} (which provide a more nuanced description of a species' place in the food web) are related to its persistence~\citep{Cirtwill2015a,Cirtwill2018FoodWebs,Cirtwill2022Oikos}.
+    The relationships between these relatively simple measures of a species' position in a food web suggest that more detailed measures, such as a species' participation in \emph{meso-scale motifs} will also be related to its persistence~\citep{Cirtwill2015a,Cirtwill2018FoodWebs,Cirtwill2022Oikos}.
 
     Motifs are unique patterns of $n$ interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
-    Some motifs are over-represented in stable food webs, suggesting that participation in these motifs provides stability (refs).
-    Indeed, dynamic simulations suggest that participation in these motifs can provide greater persistence after a disturbance~\citep{Cirtwill2022Oikos}.
+    Some motifs are over-represented in stable food webs, suggesting that these motifs provides stability~\citep{Borrelli2015a}.
+This further suggest that the more frequently species are participating in these motifs, the more stable the food web would be.
+    % Indeed, dynamic simulations show that participation in some motifs can provide greater persistence after the loss of a single species~\citep{Cirtwill2022Oikos}, suggesting that motif participation may be related to persistence after other types of disturbance.
     The vector of frequencies with which a species appears in each motif --its \emph{motif participation}-- describes the species' position in the meso-scale structure of a food web and, like degree and trophic level, is easily calculated once the structure of the network is known.
     Importantly, motif participation incorporates information on the positions of focal species' interaction partners, as well as potential indirect interactions between species.
-    Two species might both have two prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
+    Two species might have the same trophic level and the same number of prey, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists (for an example see Fig.~\ref{fig:concept}).
     Since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
     This extra information on the focal species' interaction partners may thus provide extra insight when predicting extinction risk.
-
-    Here, we focus on extinction risk after the loss of plants or other basal resources, i.e., the effects of bottom-up disturbances.
-    The basal resource community is the foundation upon which all other species depend, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
-    Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
-    It is therefore crucial to understand how bottom-up effects of disturbance to basal resources affect the rest of the food web.
-    As well as identifying patterns of motif participation that are associated with ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
-    For example, participation in one motif might be associated with greater extinction risk after a mild disturbance but be associated with less risk when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances while providing `back-up' resources when many prey are removed by a strong disturbance.
 
@@
     The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers by depletion of the shared resource in competitive exclusion~\citep{Hsu2015}.
-    Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.
-    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey.
-The motif indicates a generalist predator (which may be able persist following reduction or loss of one prey) but can capture indirect effects between prey that make the predator more likely to go extinct~\citep{Melian2002}.
-    Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator. 
-    Omnivory is less stable in isolation than the other three motifs~\citep{Borrelli2015a}, though there is a region of stable equilibrium for Lotka-Volterra model of omnivory~\citep{Hsu2015}.
+   ( Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.)
+    The apparent competition motif indicates a predator which consumes two resources and, by being somewhat generalist, may be able to persist following the loss or reduction of one prey.
+    However, there may also be indirect effects between the linked prey that make the predator more likely to go extinct~\citep{Melian2002}.
+    Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes  the potential for a plant to have both direct and indirect effects on the predator. 
+    Omnivory is less stable in isolation than the other three motifs~\citep{Borrelli2015a}, though there is a region of stable equilibrium for Lotka-Volterra models of omnivory~\citep{Hsu2015}.
     In whole food webs, frequent omnivory has been proposed to both increase and decrease overall stability in different conditions~\citep{McCann1997,Emmerson2004,Monteiro2016}.
     As most of the above studies either model motifs in isolation or examine overall effects on the whole food web, it remains unclear how a particular species' motif participation will affect its specific extinction risk.
+
+    Previous work using dynamical simulations has shown that participation in some motifs is associated with longer persistence after the loss of a single species in the food web~\citep{Cirtwill2022Oikos}, but the relationship between motif participation and persistence in the face of other disturbances remains unclear.
+   As the basal resource community is the foundation upon which all other species depend, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
+    Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
+    It is therefore crucial to understand how bottom-up effects from disturbance to basal resources affect the rest of the food web.
+Here, we are interested in understanding if motif participation can provide additional information on how extinction risk of consumer species are affected by the loss of basal resources, i.e., the effects of bottom-up disturbances.
+    As well as identifying patterns of motif participation that are associated with ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
+    For example, participation in one motif might be associated with greater extinction risk after a mild disturbance but be associated with less risk when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances while providing `back-up' resources when many prey are removed by a strong disturbance.
 
@@
         \centering
-        % \includegraphics[width=.9\textwidth]{figures/Figure_1.png}
+        \includegraphics[width=.9\textwidth]{figures/Figure_1.eps}
         \caption{Conceptual figure of a toy food web.
-\textbf{a)} This food web contains examples of the four motifs: apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow).
+\textbf{a)} This food web contains examples of the four motifs: apparent competition (orange), three-species chain (purple), omnivory (red), and direct competition (blue).
 These motifs capture different ways in which disturbances to basal resources can propagate through the network.
@@
         Note that the nodes participate in more motifs than the ones highlighted and that a species' motif participation vector is the count of all motifs of each type in which the species appears.
-\textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than three times more frequently (see inserted table for counts of participation in all four motifs, giving the motif participation vector in parentheses).
+\textbf{b)} In the same food web, species with identical degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only three chains while species B (dark grey node with its direct consumption links colored dark grey) participates in seven (see inserted table for counts of participation in all four motifs; species A has the motif participation vector [3, 3, 0, 1] while species B has the motif participation vector [3, 7, 0, 1]).
         Despite having equal degree and trophic level, species A may be at greater risk of going extinct than species B because of its dependence on specialist prey species.
@@
         \centering
-        % \includegraphics[width=0.85\textwidth]{figures/persistence_motif_participation.eps}
-        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis).
-The effect of participating in each motif is based on the fixed effects in Equation~\ref{propreq}.
+        \includegraphics[width=\textwidth]{figures/persistence_motif_participation.eps}
+        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on probability of persistence following disturbance to basal resources (y-axis).
+The effect of participating in each motif is based on the Equation~\ref{propreq}.
 The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance).
-95\% confidence intervals for each line are shown in grey.
+95\% confidence intervals for each line are shown in lighter shading.
 Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
@@
     \centering
-        % \includegraphics[width=\textwidth]{figures/Fig4.pdf}
+        \includegraphics[width=\textwidth]{figures/Fig4.pdf}
         \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.
-Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right).
-Rows show various levels of connectance.
+Columns show the result for different motifs, rows show results for low, medium, and high connectance, and colour indicates the strength of disturbance to basal resources ($\pi_{disturbed}$; probability of extinction for a basal resource).
 The dotted, vertical lines indicate zero on the x-axis.
-A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increases persistence.
-The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend).
-}
+A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increases persistence.}
         \label{fig:density_prop}
@@
         \centering
-        % \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
+        \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
         \caption{The probabilities of persistence for a particular consumer (\textbf{A, B}) and the average probability of persistence across all consumers (\textbf{C-E}) varied with other elements of network structure as well as with motif participation.
@@
         \centering
-        % \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
+        \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
         \caption{Motif participation correlated with in-degree and trophic level, and both of these simpler role measures were also correlated with persistence.
@@
 See also Fig.
-SX.
-}
+S6.}
         \label{fig:motifs_vs_TL_and_deg}

### 2023-06-04 manuscript/supplement_motifs.tex
SI ready to submit

@@
 
-\title{Species motif participation provides unique information about species risk of extinction}
+\title{Species motif participation provides unique information about species risk of extinction: supplemental information}
 
@@
             \begin{itemize}
-                \item Describes GLMs relating persistence and other network properties (Eqn. S2-3) and GLMs relating motif participation, profiles and network properties (Eqn. S4-6)
+                \item GLMs relating persistence and other network properties (Eqn. S2-S3)
+                \item GLMs relating motif participation, profiles and network properties (Eqn. S4-S6)
             \end{itemize}
 
-        \item Motif-persistence relationships vary with S,C
+        \item Motif-persistence relationships vary with S, C
 
             \begin{itemize}
-                \item Describes GLMs summarising how the relationship between motif participation and persistence varies over S and C, to complement the within-network results presented in the main text (Eqn.
+                \item GLMs summarising how the relationship between motif participation and persistence varies over S and C, to complement the within-network results presented in the main text (Eqn.
 S7)
-                \item Plots of these GLMs for each motif (Figs. S2-5)
+                \item Plots of these GLMs for each motif (Figs. S2-S5)
             \end{itemize}
 
-        \item Results: species motif participation 
+        \item Supplemental results: motif participation 
             \begin{itemize}
                 \item Supplemental results relating species motif participation to persistence and to other network properties (S, C, in-degree, STL)
-                \item Table S1-2l; Fig. S6
+                \item Table S1-S2; Fig. S6
             \end{itemize}
 
-        \item Supplemental results: persistence and other network properties
+        \item Supplemental results: persistence
 
             \begin{itemize}
-                \item Supplemental results identifying relationships between persistence and network size, connectance, in-degree, and STL to provide context for motif relationships (Tables S3-4).
-                \item Also relating average persistence to network motif profiles (Table S5)
+                \item Supplemental results identifying relationships between persistence and network size, connectance, in-degree, and STL to provide context for motif relationships (Tables S3-S4).
+                \item Supplemental results relating average persistence to network motif profiles (Table S5)
             \end{itemize}
@@
             \begin{itemize}
-                \item Relating motif profiles  to network size and connectance (Table S6)
+                \item Supplemental results relating motif profiles  to network size and connectance (Table S6; FIg. S7)
             \end{itemize}
@@
         If a simulated network contained any disconnected species (species without predators or prey) or disconnected components (a group of species connected among themselves but not to the rest of the network), the network was rejected and a new network simulated.
-        Finally, networks where the path lengths between each species and a basal resource could not be resolved (i.e., trophic levels were undefined) were rejected and new networks simulated.
+        Finally, networks where the path lengths between each species and a basal resource could not be resolved (i.e., trophic levels were undefined) were rejected, as was any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
+        Rejected networks were replaced with fresh simulations.
 
@@
         If a set of biomasses where all species persist still had not been reached after 100 sets of randomly-assigned initial biomasses, we discarded the network and simulated another to replace it.
-
-    \subsection{Filtering of biologically unrealistic networks}
-
-        After simulating our initial niche-model networks, we filtered networks for biological plausibility.
-        Specifically, we removed any network containing disconnected components (species or groups of species not connected to the rest of the network) or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
-        Any removed network was replaced with a fresh simulation and re-checked until 100 realistic networks for each size-connectance combination had been obtained.
 
@@
 
-\section{Motif-persistence relationships vary with S,C}
+\section{Motif-persistence relationships vary with S, C}
 
@@
         \includegraphics[width=\textwidth]{figures/persistence_apparent_detailpers.eps}
-        \caption{Slopes of the motif proportion-persistence relationships for the apparent competition motif were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance.}
+        \caption{Slopes of the motif proportion-persistence relationships for the apparent competition motif were steepest in large and less-connected networks or small and highly-connected networks but the directions of slopes varied only with the strength of disturbance.}
         \label{apparent_plus_SC}
@@
         \includegraphics[width=\textwidth]{figures/persistence_direct_detailpers.eps}
-        \caption{Slopes of the motif proportion-persistence relationships for the direct competition motif were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance.}
+        \caption{Slopes of the motif proportion-persistence relationships for the direct competition motif were steepest in less-connected networks but the directions of slopes varied only with the strength of disturbance.}
         \label{direct_plus_SC}
@@
         \includegraphics[width=\textwidth]{figures/roles_vs_SC_all.eps}
-        \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).}
+        \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition, where there was no significant interaction).}
         \label{fig:motifs_vs_TL_and_deg}
@@
 
-\section{Supplemental results: persistence and other network properties}
+\section{Supplemental results: persistence}
 
@@
 All models had binomial error distribution.
-Predictors were centered and scaled before model fitting.
 All $p$-values were \textless0.001.
@@
 Coefficients in \textbf{bold} are significant ($\alpha$\textless0.005).
-Predictors were centered and scaled before model fitting; coefficients refer to a unit increase in predictors.
 We also provide McFadden's pseudo-$R^2$ value for each regression.

### 2023-06-06 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-% Now 286 words. Max 350. 
-    \indent{1.
-Loss of species in food webs can set in motion a cascade of additional, secondary,  extinctions.A species' position in a food web (e.g., its trophic level or number of interactions) is known to affect its ability to persist following disturbance.
-Theses simple measures, however, offer only a coarse description of how species fit into their community.
-Therefore one would expect that more detailed structure measures, such as a species' participation in three-species motifs (meso-scale structures which provide information on a species' direct and indirect interactions) will also be related to its probability of persistence.\\
-    2.
-Disturbances, such as species extinctions, affecting the basal resources have particularly strong effects on the rest of the food web.
-However, how the disturbance branch out and affect consumer persistence is depending on the structural pattern of species interactions in several steps.
-The magnitude, e.g.
-proportion of basal resources lost, will likely also affect the outcome.\\
-    3.
-Here, we analyse whether a consumer's risk of going secondary extinct after removal of basal resources depends on the consumer's motif participation and how this relationship varies with the severity of disturbance.\\
+% Now 281 words. Max 350. 
+    1. Loss of species in food webs can set in motion a cascade of additional (secondary)  extinctions.
+    2. A species' position in a food web (e.g., its trophic level or number of interactions) is known to affect its ability to persist following disturbance. 
+    3. These simple measures, however, offer only a coarse description of how species fit into their community. 
     4.
-We show that consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.\\
-    5.
-However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.\\
-    6.
-Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities, but must be considered in the context of network structure as a whole and of the specific disturbance applied}
+One would therefore expect that more detailed structural measures such as  participation in three-species motifs (meso-scale structures which provide information on a species' direct and indirect interactions) will also be related to probability of persistence.
+    5. Disturbances affecting the basal resources have particularly strong effects on the rest of the food web. 
+    6. However, how disturbances branch out and affect consumer persistence depends on the structural pattern of species interactions in several steps. 
+    7. The magnitude, e.g. proportion of basal resources lost, will likely also affect the outcome.     
+    8.
+Here, we analyse whether a consumer's risk of secondary extinction after removal of basal resources depends on the consumer's motif participation and how this relationship varies with the severity of disturbance.
+    9.
+We show that consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.
+    10.
+However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.
+    11.
+Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities, but must be considered in the context of network structure as a whole and of the specific disturbance applied.
 
@@
     Motifs are unique patterns of $n$ interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
-    Some motifs are over-represented in stable food webs, suggesting that these motifs provides stability~\citep{Borrelli2015a}.
-This further suggest that the more frequently species are participating in these motifs, the more stable the food web would be.
+    Some motifs are over-represented in stable food webs, suggesting that these motifs provide stability~\citep{Borrelli2015a}. 
+    This further suggests that the more frequently species participate in these motifs, the more stable the food web would be.
     % Indeed, dynamic simulations show that participation in some motifs can provide greater persistence after the loss of a single species~\citep{Cirtwill2022Oikos}, suggesting that motif participation may be related to persistence after other types of disturbance.
@@
     The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
-    The dynamics of three-species chains in isolation are well-understood from Lotka-Volterra models, and stable equilibria with all three species persisting can often be identified~\citep{Borrelli2015a,Hsu2015}.
+    The dynamics of three-species chains in isolation are well understood from Lotka-Volterra models, and stable equilibria with all three species persisting can often be identified~\citep{Borrelli2015a,Hsu2015}.
     In general, food chains are more likely to be stable when predator-prey body-mass ratios are smaller~\citep{Jonsson1998}.
     The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers by depletion of the shared resource in competitive exclusion~\citep{Hsu2015}.
-   ( Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.)
+   Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.
     The apparent competition motif indicates a predator which consumes two resources and, by being somewhat generalist, may be able to persist following the loss or reduction of one prey.
@@
     It is therefore crucial to understand how bottom-up effects from disturbance to basal resources affect the rest of the food web.
-Here, we are interested in understanding if motif participation can provide additional information on how extinction risk of consumer species are affected by the loss of basal resources, i.e., the effects of bottom-up disturbances.
+Here, we are interested in understanding if motif participation can provide additional information on how extinction risks of consumer species are affected by the loss of basal resources, i.e., the effects of bottom-up disturbances.
     As well as identifying patterns of motif participation that are associated with ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
@@
         This means that only four motifs can appear: apparent competition, direct competition, omnivory, and the three-species chain.
-        Fortunately, they are also the motifs that are most common in empirical networks (\citealp[]{Stouffer2007}, have been most studied in isolation (e.g.,~\citep{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.
+        Fortunately, they are also the motifs that are most common in empirical networks~\citep{Stouffer2007}, have been most studied in isolation (e.g.,~\citealp[]{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.
 
-        \subsubsection*{Modelling consumers' risk of secondary extinction} [[trimmed down to here]]
+        \subsubsection*{Modelling consumers' risk of secondary extinction} %[[trimmed down to here]]
 
@@
 S2-3).
-            As with the GLMs relating persistence to motif participation and motif profiles, we initially fit GLMMs including random effects of network size and connectance and network identity but removed these terms due to model singularity.
+            When relating persistence to STL, we initially included a random effect of network size and connectance but removed these terms due to model singularity as above.
+            The model for in-degree was non-singular when including these random effects and so they were retained in this case.
 
@@
 
-        The average persistence of all consumers was lower in networks with higher connectance, did not vary significantly with network size (Table S4, \emph{Appendix S6.1}), and  varied with the frequencies of different motifs in this profile.
+        The average persistence of all consumers was lower in networks with higher connectance, did not vary significantly with network size (Table S4, \emph{Appendix S6.1}), and  varied with the frequencies of different motifs in the network's motif profile.
         Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:lm_CS}, Table S5, \emph{Appendix S6.2}).
-        There was no significant interaction between the proportion of any motif in the network profile and level of disturbance).
+        These relationships are broadly similar to, but weaker than, the relationships between species' probabilities of persistence and motif participation.
+        There was no significant interaction between the proportion of any motif in the network motif profile and level of disturbance.
 
@@
     At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and therefore have lower persistence.
-    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{Allesina2012,binzer2011susceptibility}.
+    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{Allesina2012,binzer2011susceptibility}, though high-degree species may be less vulnerable than we expect if they are able to switch to alternative prey.
+    Our model does not include interaction rewiring, but it is possible that predators with many observed prey may also be more able to attack new prey species after a disturbance than specialists.
     These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.
@@
 ARC is supported by a Finnish Academy postdoctoral research grant (\#332999).
+ARC, A\r{A}, and AE conceived the study.
+ARC and KLW simulated networks.
+A\r{A} and ARC conducted analyses.
+All authors contributed to writing the manuscript.
 
@@
         \centering
-        \includegraphics[width=.9\textwidth]{figures/Figure_1.eps}
+        % \includegraphics[width=.9\textwidth]{figures/Figure_1.eps}
         \caption{Conceptual figure of a toy food web.
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/persistence_motif_participation.eps}
+        % \includegraphics[width=0.95\textwidth]{figures/persistence_motif_participation.eps}
         \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on probability of persistence following disturbance to basal resources (y-axis).

### 2023-06-06 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 \clearpage
-\begin{spacing}{2.0}
+\begin{spacing}{1.0}
 % \linenumbers
@@
 
-    The loss of a species in a food web can, due to the mutual dependencies among species,  trigger a cascade of additional (secondary) extinctions. 
-    A species' risk of secondary extinction depends on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
-    For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{Allesina2012,binzer2011susceptibility, Eklof2006}.
-    These species are more vulnerable because they are impacted by the loss of any species along a long path from basal resources or because they have few alternative resources if a prey species is lost.
-    The relationships between these relatively simple measures of a species' position in a food web suggest that more detailed measures, such as a species' participation in \emph{meso-scale motifs} will also be related to its persistence~\citep{Cirtwill2015a,Cirtwill2018FoodWebs,Cirtwill2022Oikos}.
+    A single extinction can, due to the mutual dependencies among species in a food web,  trigger a cascade of additional (secondary) extinctions. 
+    A species' risk of secondary extinction depends on its position within a food web~\citep{Santos2021,dunne2009cascading, Eklof2006}.
+    For example, higher trophic levels (long paths to basal resources) and lower in-degree (few prey) increase extinction risk~\citep{Allesina2012,binzer2011susceptibility,Eklof2006}.
+    These species are more vulnerable because they are impacted by the loss of any species along a long path from basal resources or because they have few alternative resources to replace lost prey.
+    The relationships between these relatively simple measures of a species' position in a food web suggest that more detailed measures, such as a species' participation in \emph{meso-scale motifs} are also related to its persistence~\citep{Cirtwill2022Oikos}.
 
-    Motifs are unique patterns of $n$ interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
-    Some motifs are over-represented in stable food webs, suggesting that these motifs provide stability~\citep{Borrelli2015a}. 
-    This further suggests that the more frequently species participate in these motifs, the more stable the food web would be.
-    % Indeed, dynamic simulations show that participation in some motifs can provide greater persistence after the loss of a single species~\citep{Cirtwill2022Oikos}, suggesting that motif participation may be related to persistence after other types of disturbance.
+    Motifs are unique patterns of interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
+    Some motifs are over-represented in stable food webs, suggesting that the more frequently species participate in these motifs, the more stable the network will be~\citep{Borrelli2015a}. 
     The vector of frequencies with which a species appears in each motif --its \emph{motif participation}-- describes the species' position in the meso-scale structure of a food web and, like degree and trophic level, is easily calculated once the structure of the network is known.
-    Importantly, motif participation incorporates information on the positions of focal species' interaction partners, as well as potential indirect interactions between species.
-    Two species might have the same trophic level and the same number of prey, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists (for an example see Fig.~\ref{fig:concept}).
+    Importantly, motif participation incorporates information on the positions of focal species' interaction partners.
+    Two species might have the same trophic level and in-degree but quite different motif participation if one species's prey are strict specialists and the others' prey are generalists (Fig.~\ref{fig:concept}).
     Since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
@@
     %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
-    We define motif participation using the well-studied three-species motifs~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
+    We define motif participation using the well-studied three-species motifs~\citep{Milo2004,Stouffer2007,Stouffer2012}. 
     Four of these motifs --the three-species chain, apparent competition, direct competition, and omnivory-- are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, giling2019plant} and more stable than other motifs (i.e., are more likely to retain all three species in the motif) when modelled in isolation~\citep{Borrelli2015a}.
-    These four motifs also have clear ecological interpretations.
+    These four motifs also have clear ecological interpretations and their dynamics explored using Lotka-Volterra models~\citep{Hsu2015}.
     The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
-    The dynamics of three-species chains in isolation are well understood from Lotka-Volterra models, and stable equilibria with all three species persisting can often be identified~\citep{Borrelli2015a,Hsu2015}.
     In general, food chains are more likely to be stable when predator-prey body-mass ratios are smaller~\citep{Jonsson1998}.
     The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers by depletion of the shared resource in competitive exclusion~\citep{Hsu2015}.
-   Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.
-    The apparent competition motif indicates a predator which consumes two resources and, by being somewhat generalist, may be able to persist following the loss or reduction of one prey.
-    However, there may also be indirect effects between the linked prey that make the predator more likely to go extinct~\citep{Melian2002}.
+   % Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.
+    The apparent competition motif indicates a predator which consumes two resources and, by being somewhat generalist, may be able to persist following the loss or reduction of one prey but can also be affected by indirect effects between the linked prey~\citep{Melian2002}.
     Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes  the potential for a plant to have both direct and indirect effects on the predator. 
@@
 
-    Previous work using dynamical simulations has shown that participation in some motifs is associated with longer persistence after the loss of a single species in the food web~\citep{Cirtwill2022Oikos}, but the relationship between motif participation and persistence in the face of other disturbances remains unclear.
-   As the basal resource community is the foundation upon which all other species depend, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
-    Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
+    Previous work using dynamical simulations has shown that motif participation is related to persistence after the loss of a single species in the food web~\citep{Cirtwill2022Oikos}, but the relationship between motif participation and persistence in the face of other disturbances remains unclear.
+   As the basal resource community is the foundation upon which all other species depend, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food,Georgiadis2007}.
+    Moreover, disturbances to the plant community have been shown to affect both the biomass and diversity of consumers~\citep{chen2019plant,scherber2010bottom,li2020bottom}.
     It is therefore crucial to understand how bottom-up effects from disturbance to basal resources affect the rest of the food web.  
-Here, we are interested in understanding if motif participation can provide additional information on how extinction risks of consumer species are affected by the loss of basal resources, i.e., the effects of bottom-up disturbances.
     As well as identifying patterns of motif participation that are associated with ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
@@
         First, there are topological models where secondary extinctions only occur once a consumer has lost all of its prey~\citep{dunne2009cascading}. 
-        Alternatively, dynamical models explicitly simulate population dynamics using a system of differential equations and take changes in prey or predator populations into account when calculating populations of other species in the network~\citep{binzer2011susceptibility}.
-        This additional detail means that realistic processes such as indirect interactions can be included, but also means that dynamical models are highly parameter intensive and simulations are much more time-consuming than topological models.
+        Alternatively, dynamical models explicitly simulate population dynamics using a system of differential equations~\citep{binzer2011susceptibility}. 
+        Dynamical models include realistic processes such as indirect interactions but are also highly parameter intensive and require much more computation time than topological models.
 
-        A middle‐ground approach is to use Bayesian networks, which are much more computationally efficient and facilitate the analysis of larger networks than is practical for dynamical models \citep{Eklof2013,Haussler2020}.
+        Bayesian networks, which are much more computationally efficient and less-parameter-intensive (and therefore less sensative to modelling assumptions) than dynamical models~\citep{Eklof2013} but are more realistic than topological models, offer a middle-ground approach \citep{Eklof2013,Haussler2020}.
         In this framework, a consumer's probability of extinction depends on the fraction of resources lost ($f = k/n$ for a species with $n$ resources, of which $k$ have gone extinct) and on a baseline probability of extinction ($\pi$) which capturing factors that are not related to the network structure (e.g., disease or stochastic extinction of small populations).
-        Importantly, the greater speed of the Bayesian network framework allows us to simulate a range of disturbance severities rather than single species removals as in many previous studies (e.g.,~\citealp[]{Memmott2004,Staniczenko2010,Dunne2004,Cirtwill2022Oikos}).
-        Moreover, Bayesian network simulations are less parameter-intensive than dynamical models and are therefore less sensitive to assumptions made in the modelling process~\citep{Eklof2013}.
+        The efficiency of the Bayesian network framework allows us to simulate a range of disturbance severities rather than single species removals as in many previous studies (e.g.,~\citealp[]{Memmott2004,Staniczenko2010,Cirtwill2022Oikos}).
 
@@
         This means that only four motifs can appear: apparent competition, direct competition, omnivory, and the three-species chain.
-        Fortunately, they are also the motifs that are most common in empirical networks~\citep{Stouffer2007}, have been most studied in isolation (e.g.,~\citealp[]{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.
+        Fortunately, they are also the motifs that are most common in empirical networks~\citep{Stouffer2007}, have been most studied in isolation (e.g.,~\citealp[]{Bascompte2005,Polis1989,Zabalo2012,Kondoh2008,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.
 
@@
         Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:lm_CS}, Table S5, \emph{Appendix S6.2}).
-        These relationships are broadly similar to, but weaker than, the relationships between species' probabilities of persistence and motif participation.
         There was no significant interaction between the proportion of any motif in the network motif profile and level of disturbance.
@@
     Identifying the species that are most and least likely to persist after a disturbance is a high priority for conservation efforts~\citep{Bottrilletal2008}. 
-    Network structural properties such as connectance and measures of a species' position in the network such as degree and trophic level are known to affect species' vulnerability~\citep{Dunne2002, Eklof2006, PascualDunne2006,prill2005dynamic, bascompte2005simple,curtsdotter2011robustness, dunne2009cascading}, but are relatively coarse metrics.
-    Defining a species' network role based on its participation in meso-scale motifs provides more nuanced information and may also predict extinction risk~\citep{Stouffer2007,Stouffer2010b,Simmons2019,Cirtwill2018FoodWebs}, but this potential is only beginning to be explored~\citep{Cirtwill2022Oikos}.
+    Network structural properties such as connectance and measures of a species' position in the network such as degree and trophic level are known to affect species' vulnerability~\citep{Dunne2002, Eklof2006,bascompte2005simple,dunne2009cascading}, but are relatively coarse metrics.
+    Defining a species' network role based on its participation in meso-scale motifs provides more nuanced information and may also predict extinction risk~\citep{Stouffer2012,Simmons2019,Cirtwill2018FoodWebs}, but this potential is only beginning to be explored~\citep{Cirtwill2022Oikos}.
 
@@
     However, as these two studies applied different types of disturbances, further study would be required to disentangle the effects of top-down and bottom-up effects on species with different motif participation, and how each relates to the type and strength of disturbance.
-    Another potentially useful addition to future work would be to include species traits such as body size, which are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000} in empirical webs.
+    Another potentially useful addition to future work would be to include species traits such as body size, which are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017,Cardillo2005, Purvis2000} in empirical webs.
 
@@
 
-    Many ecologists, including ourselves, have argued that simple measures of species' roles within food webs are important for understanding species' extinction risk~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006,binzer2011susceptibility}.
+    Many ecologists, including ourselves, have argued that simple measures of species' roles within food webs are important for understanding species' extinction risk~\citep{Santos2021,dunne2009cascading, Eklof2006,binzer2011susceptibility}.
     However, here we show that relationships between extinction risk and measures of a species' place in its food web are likely more complex, varying with the strength of disturbance and the overall network structure.

### 2023-06-06 manuscript/reorder_main_text.tex
Updates from Overleaf

@@
 
-% Limit is 8500 words.
-% Currently c. 8066 - we're good!
 \begin{document} 
@@
 
-% Now 281 words. Max 350. 
+% Now 282 words. Max 350. 
     1. Loss of species in food webs can set in motion a cascade of additional (secondary)  extinctions.
@@
 
-\section*{Introduction} %New
+\section*{Introduction}
 
@@
 
-    %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
     We define motif participation using the well-studied three-species motifs~\citep{Milo2004,Stouffer2007,Stouffer2012}. 
@@
     The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers by depletion of the shared resource in competitive exclusion~\citep{Hsu2015}.
-   % Despite the potential for exclusion where one predator is competitively superior, this motif is highly robust to changes in interaction strength and therefore has high quasi-sign stability~\citep{Borrelli2015a}.
     The apparent competition motif indicates a predator which consumes two resources and, by being somewhat generalist, may be able to persist following the loss or reduction of one prey but can also be affected by indirect effects between the linked prey~\citep{Melian2002}.
@@
 
-    %\subsection*{Here we test...}
     Here, we analyse whether a consumer's risk of secondary extinction after removal of basal resources depends on the consumer's motif participation and how this relationship varies with the severity of disturbance.
@@
 
-\section*{Methods}
+\section*{Materials and methods}
 
@@
 
-        We generated simulated networks using the niche model, which recreates the structure of empirical networks well~\citep{Williams2000,Stouffer2007}.
-        To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002e}, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
-        We removed any biologically unlikely networks (i.e., those with extremely long paths between any consumer and basal resources~\citep{Borrelli2014} and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.
+        Simulated networks were generated using the niche model, which recreates the structure of empirical networks well~\citep{Williams2000,Stouffer2007}.
+        To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002e}, networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04) were simulated.
+        Biologically unlikely networks (i.e., those with extremely long paths between any consumer and basal resources~\citep{Borrelli2014} were removed and replaced with new simulated networks until 100 suitable networks in each combination of size and connectance were obtained.
         All networks were generated using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.     
@@
 
-        We described the motif participation of each consumer species in each of these simulated networks before applying any disturbance. 
+        The motif participation of each consumer species was described in each of these simulated networks before applying any disturbance. 
         A species' motif participation (Fig.~\ref{fig:concept}) is the vector of frequencies with which it appears in each type of motif present in a network~\citep{Stouffer2012}.
         As only four unique three-species motifs can appear in acyclic networks, these are four-dimensional vectors. 
-        Here we are interested in the relative frequencies of each motif rather than the total number of motifs and therefore normalized participation vectors by dividing each count by the total number of motifs the species appears in (such that the vector for each species sums to 1).
+        To focus on the relative frequencies of each motif rather than the total number of motifs, participation vectors were normalized by dividing each count by the total number of motifs the species appears in (such that the vector for each species sums to 1).
         Note, however, that this normalization does not fully control for differences due to degree or connectance~\citep{Cirtwill2022Oikos}.
 
-        We also calculated two simpler measures of a species' position within its community.
+        Two simpler measures of a species' position within its community were also calculated.
         In-degree is a species' number of prey.
@@
 
-        Finally, to provide context for species' motif participation, we calculated network ``motif profiles'': four-dimensional vectors of the number of each three-species motif in each network~\citep{Stouffer2012}.
+        Finally, network ``motif profiles'' --four-dimensional vectors of the number of each three-species motif in each network~\citep{Stouffer2012}-- were calculated to provide context for species' motif participation..
         Together with network size and connectance, these motif profiles describe the structure of the whole network.
-        To separate differences in motif structure from differences in network size or connectance (larger and more-connected networks contain more motifs), we normalized motif profiles by dividing the count of each motif by the total count of all motifs.
-        We calculated species' motif participation and network motif profiles using the Python package \emph{pymfinder}~\citep{pymfinder}.
+        To separate differences in motif structure from differences in network size or connectance (larger and more-connected networks contain more motifs), motif profiles were normalised by dividing the count of each motif by the total count of all motifs.
+        Motif participation and profiles were calculated using the Python package \emph{pymfinder}~\citep{pymfinder}.
 
@@
 
-        % Mention acyclic resstriction here.
         One limitation to Bayesian networks is that consumers do not affect the extinction probabilities of their resources (i.e., only bottom-up effects are included).
         Nevertheless, Bayesian networks do capture the vast majority of secondary extinctions obtained from fully-fledged dynamical models including top-down and bottom-up effects~\citep{Eklof2013}, and as such are a good proxy for analysing species persistence following extinctions of primary producers in complex food webs.
-        As our focus in this study is on bottom-up effects from disturbance of basal resources, Bayesian networks will capture the secondary extinctions we are interested in.
+        As our focus in this study is on bottom-up effects from disturbance of basal resources, Bayesian networks will capture the secondary extinctions of interest.
 
@@
         This means that only four motifs can appear: apparent competition, direct competition, omnivory, and the three-species chain.
-        Fortunately, they are also the motifs that are most common in empirical networks~\citep{Stouffer2007}, have been most studied in isolation (e.g.,~\citealp[]{Bascompte2005,Polis1989,Zabalo2012,Kondoh2008,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.
+        Fortunately, they are also the motifs that are most common in empirical networks~\citep{Stouffer2007}, have been most studied in isolation (e.g.,~\citealp[]{Bascompte2005,Kondoh2008,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.
 
-        \subsubsection*{Modelling consumers' risk of secondary extinction} %[[trimmed down to here]]
+        \subsubsection*{Modelling consumers' risk of secondary extinction}
 
-        The response of the consumer to changes in $f$ can take any form (\emph{Appendix S2.1}).
-Here we use a sigmoid non-linear function ($\alpha > 1, \beta > 1$), the type of response that most accurately captures the secondary extinctions produced by dynamical models (\citealp[]{Eklof2013}).
+           Response of the consumer to changes in $f$ were modelled using a sigmoid non-linear function ($\alpha > 1, \beta > 1$) which most accurately captures the secondary extinctions produced by dynamical models (\citealp[]{Eklof2013}; \emph{Appendix S2.1}).
             The probability of extinction for each species $i$ can therefore be represented using the cumulative density function of a beta distribution:
@@
 
-		If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability will be $P(\lnot C|f) = \pi$. 
-		Similarly, if all resources are extinct ($f = 1$), the extinction probability $P(\lnot C|f)$ will be equal to 1 and the consumer will go extinct.
-		If the fraction $f$ is neither 1 nor 0, the non-linear sigmoid curve will determine the response of the consumer, which will be more severe if the fraction of resources lost is high.
-This will in turn affect the extinction probability $P(\lnot C|f)$.
-		For basal resources, we assume that the abiotic resources they require are always available and therefore $f=0$. 
-		Following Equation \ref{betafunc}, the extinction probability of a basal resource will only depend on $\pi$.
-		Note that we here assume that only the fraction, and not the identity, of the resources lost is important. 
+            If all resource species of a consumer \textit{C} are present ($f = 0$), the extinction probability is $P(\lnot C|f) = \pi$. 
+            Similarly, if all resources are extinct ($f = 1$), $P(\lnot C|f)=1$  and the consumer will go extinct.
+            If the fraction $f$ is neither 1 nor 0, the non-linear sigmoid curve will determine the response of the consumer, and $P(\lnot C|f)$ is greater if the fraction of resources lost is high. 
+            For basal resources, the abiotic resources they require are assumed to be always available.
+            Therefore $f=0$ and, following Equation~\ref{betafunc}, $P(\lnot C|f) = \pi$.
+            It is assumed that only the fraction, and not the identity, of the resources lost is important. 
             However, the model can easily be modified to incorporate weighted resource interactions \citep[see][]{Eklof2013}.
 
-		% how we calculate persistences 
-        Beginning with basal resources and working systematically up through the food web, we use Equation \ref{betafunc} to calculate each species $i$'s probability of extinction $P(\lnot i|f)$.
-        Using this probability, we then perform Bernoulli trials to determine whether the species goes extinct or not. 
+            After species $i$'s probability of extinction $P(\lnot i|f)$ is calculated using Equation~\ref{betafunc}, Bernoulli trials are used to determine whether the species goes extinct or not. 
             Species are considered extinct if a random number drawn from a uniform distribution $[0-1]$ is less than $P(\lnot i|f)$.
-        These simulated extinctions are used when calculating the fraction of resources lost $f$ for each consumer.
-        This allows us to update Equation~\ref{betafunc} with information on extinctions of a focal consumer's resources and simulate bottom-up effects.
-As this method is probabilistic, we repeat these calculations 100 000 times per network, with unique random draws.
-        For each species, we then define probability of persistence (hereafter `persistence') as the fraction of simulations in which the species persists. 
+            Moving up through the food web, the simulated extinctions of prey species are used to calculate the fraction of resources lost $f$ for each consumer and Equation~\ref{betafunc} is updated accordingly.
+            This probabilistic method was repeated 100 000 times per network, with unique random draws.
+            For each species, probability of persistence (hereafter `persistence') is defined as the fraction of simulations in which the species persists. 
             While there are methods for solving Bayesian networks exactly \citep{Eklof2013}, for larger networks the numerical evaluation above is highly efficient and produces close to identical results \citep{Haussler2020}.
@@
             In our un-disturbed scenario, all species had an extinction probability of $\pi = 0.1$. 
-            To simulate various threats to basal resources, making some species more vulnerable to extinction, we increase all basal species' probability of extinction from $0.10$ to $0.5$, in steps of $0.08$.
+            To simulate various threats to basal resources, making some species more vulnerable to extinction, probability of extinction for all basal species was increased from $0.10$ to $0.5$, in steps of $0.08$.
             The highest disturbance level, $\pi = 0.5$, corresponds to basal species having a 50\% risk of going extinct. 
@@
 
-        We first tested 1) how consumer species persistence varied with motif participation and the severity of disturbance.
-        For context, we also test 2) how a consumer's persistence varies with other measures of network structure and a species' place within it.
-        As different measures of network structure are not independent since all describe the same network, we also 3) relate species' motif participation to large-scale network structure and to species in-degree and trophic level.
-        All analyses are briefly described below and were performed using  R~\citep{R} package. % \emph{lmerTest}~\citep{lmerTest}.
-        % , \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.
-        See Appendix \emph{S3} for details.
-
         \subsubsection*{1) Consumer persistence and motif participation}
 
-            To test whether a consumer's risk of extinction was associated with the frequency of each motif in its participation vector (hereafter simply `participation in motif X') and whether this relationship varied with the strength of disturbance, we fit four general linear models with binomial error distribution (GLMs; one per motif included in the Bayesian networks):
+            To test whether a consumer's probability of persistence was associated with the frequency of each motif in its participation vector (hereafter simply `participation in motif X') and whether this relationship varied with the strength of disturbance, four general linear models were fit with binomial error distribution (GLMs; one per motif included in the Bayesian networks):
             \begin{equation}
@@
 
-            Initially, we fit general linear mixed-effect models (GLMMs) also including random  intercepts for the species richness and connectance of network $n$ and for network identity.
-            As these models were singular, we removed each random intercept in turn and re-fit the model until it converged.
-            In all four cases, the convergent model did not retain any random effects.
-            We attempted to fit a PERMANOVA test relating a species' motif participation vector as a whole to its persistence, but the assumption of equal variance among groups was violated and we therefore do not report the results of this test.
+            Initially, general linear mixed-effect models (GLMMs) were fit including random  intercepts for the species richness and connectance of network $n$ and for network identity.
+            As these models were singular, each random intercept was removed in turn and the model re-fit until it converged (no random effects remaining).
+            All models here and below were fit using the R~\citep{R} functions `glm' from the \emph{stats} package and `glmer' from the package \emph{lmerTest}~\citep{lmerTest}.
 
-            To test whether these general trends were consistent across networks, we visually examined the distribution of slopes of simplified regressions fit to a single network and level of disturbance.
-            These models included only participation in the focal motif ($\rho_{i}$ above) and again used a binomial error distrubution.
-            To aid in interpretation of these single-network models with respect to network size and connectance, we fit an additional set of GLMs including proportion of motif, strength of disturbance, network size, and connectance across all networks (\emph{Appendix S4.1}).
+            To test whether these general trends were consistent across networks, the distribution of slopes of simplified regressions fit to a single network and level of disturbance were visually examined.
+            These models included only participation in the focal motif ($\rho_{i}$ above) and again used a binomial error distribution.
+            To aid in interpretation of these single-network models with respect to network size and connectance, an additional set of GLMs was fit including proportion of motif, strength of disturbance, network size, and connectance across all networks (\emph{Appendix S4.1}).
 
@@
             The overall structure of a network could also affect consumers' ability to persist following a disturbance.
-            Here, we describe the overall network structure using the frequency of each motif in the whole network (i.e., the network's motif profile) as well as network size and connectance (which were determined prior to network simulation).
-            To test whether the frequency of particular motifs in the network profile were related to the average persistence of all consumers, we fit a set of four GLMs with binomial error distribution (one per motif):
+            Here, the overall network structure was described using the frequency of each motif in the network's motif profile as well as initial network size and connectance.
+            To test whether the frequencies of motifs in the network profile were related to the average persistence of all consumers, a set of four GLMs  (one per motif) were fit with binomial error distribution:
                 \begin{equation}
@@
                 \end{equation}
-            \noindent where $\overline{\Psi_{nk}}$ is the mean persistence of all consumers in network $n$ at disturbance level $k$, $\Bar{\rho}_{n}$ is the proportion of the focal motif in the network's motif profile, and $\pi_k$ is the level of disturbance (probability of extinction for basal resources) as in equation~\ref{propreq}.
-            As with the GLMs relating each consumer's persistence to its motif participation, we initially attempted to fit GLMMs including random effects of network size and connectance and network identity but removed these random effects in order to obtain non-singular models.
+            \noindent where $\overline{\Psi_{nk}}$ is the mean persistence of consumers in network $n$ at disturbance level $k$, $\Bar{\rho}_{n}$ is the frequency of the focal motif in the network's motif profile, and $\pi_k$ is the level of disturbance (probability of extinction for basal resources).
+            As with the GLMs relating each consumer's persistence to its motif participation, GLMMs including random effects of network size and connectance and network identity were initially fit but the random effects were removed in order to obtain non-singular models.
 
-            We also confirmed previously-known relationships between persistence and other measures of network structure or a species' participation in the network.
-            To identify the relationship between large-scale network structure and mean persistence within a network, and how the strength of disturbance might affect this relationship, we fit a GLM relating mean persistence to network size, connectance, disturbance, and their interactions.
-            As the three-way interaction was significant, we did not simplify the model. 
-            To confirm known relationships between persistence and in-degree or trophic level in our simulated networks and identify how they might vary with different strengths of disturbance, we fit two GLMs with binomial error distribution relating a consumer's persistence to in-degree or STL, disturbance, and the interaction between them (\emph{Appendix S3.1}, Eq.
+            To identify the relationships between large-scale network structure,  mean persistence within a network, and strength of disturbance, a GLM was fit relating mean persistence to network size, connectance, disturbance, and their interactions.
+            To confirm known relationships between persistence and in-degree or trophic level and identify how these relationships might vary with disturbance strength, two GLMs were fit with binomial error distribution relating a consumer's persistence to in-degree or STL, disturbance, and the interaction between them (\emph{Appendix S3.1}, Eq.
 S2-3).
-            When relating persistence to STL, we initially included a random effect of network size and connectance but removed these terms due to model singularity as above.
+            When relating persistence to STL, random effects of network size and connectance were initially included but removed due to model singularity as above.
             The model for in-degree was non-singular when including these random effects and so they were retained in this case.
@@
             Since a network's motif profile and a species' motif participation reflect the same network as other measures (size, connectance, degree, and trophic level), these measures are not independent.
-            We identified how the frequencies of motifs in the network's motif profile were related to network size, connectance, and their interaction using a set of four GLMs with binomial error distribution (Appendix \emph{S3.2}, eq.
+            The relationships between frequencies of motifs in the network's motif profile and network size, connectance, and their interaction were described using a set of four GLMs with binomial error distribution (Appendix \emph{S3.2}, eq.
 S4).
-            We identified how the frequencies of motifs in a consumer's motif participation vector were related to network size, connectance, and their interaction using a similar set of four GLMs with binomial error distribution.
-            Finally, we fit two additional sets of GLMs (eight total) relating participation in each motif to the consumer's in-degree or trophic level (Appendix \emph{S3.2}, Eq. S5-6).
+            The relationships between frequencies of motifs in a consumer's motif participation vector and network size, connectance, and their interaction were described using a similar set of four GLMs with binomial error distribution.
+            Finally, two additional sets of GLMs (eight total) were fit relating participation in each motif to the consumer's in-degree or trophic level (Appendix \emph{S3.2}, Eq. S5-6).
 
@@
         A species' probability of persistence (hereafter `persistence') was significantly related to its motif participation (Table S1, \emph{Appendix S5}).
-        Across all levels of disturbance, species were more likely to persist if they participated more frequently in the direct competition motif and less frequently in the omnivory motif (Fig.~\ref{fig:prop_lmer_all}).
+        Across all levels of disturbance, persistence increased with increasing participation in the direct competition motif and decreasing participation in the omnivory motif (Fig.~\ref{fig:prop_lmer_all}).
         The relationships between persistence and participation in the other two motifs varied with the strength of disturbance.
-        At low levels of disturbance, species were more likely to persist if they participated more frequently in the three-species chain motif and less frequently in the apparent competition motif.
+        At low levels of disturbance, persistence increased with increasing participation in the three-species chain motif and decreasing participation in the apparent competition motif.
         These trends were reversed at high levels of disturbance.
 
-        Within these overall trends, there was substantial variation in the relationships between persistence and motif participation between networks with different overall network structures (Fig.~\ref{fig:density_prop}).
-        Networks with high and low connectances had different distributions of persistence-motif participation relationships, while network size had a much smaller effect (\emph{Appendix S4.2}).
+        Within these overall trends, there was substantial variation between networks (Fig.~\ref{fig:density_prop}).
+        Distributions of persistence-motif participation relationships varied strongly with connectance, while network size had a much smaller effect (\emph{Appendix S4.2}).
         The relationship between participation in the omnivory motif and persistence was especially variable.
-        Higher proportions of omnivory were associated with lower probabilities of persistence when connectance was low, regardless of network size or strength of disturbance.
-        At moderate connectance, however, higher proportions of omnivory were associated with greater probabilities of persistence when disturbance was weak.
-        In the most-connected networks, higher proportions of omnivory were associated with greater probabilities of persistence at all levels of disturbance.
+        Increased participation in the omnivory motif were associated with lower  persistence when connectance was low, regardless of network size or strength of disturbance.
+        At moderate connectance, however, increased participation in omnivory was associated with greater persistence when disturbance was weak.
+        In the most-connected networks, increased participation in omnivory were associated with greater persistence at all levels of disturbance.
 
-        The relationships between persistence and participation in the other motifs were more consistent over different levels of connectance and network size.
-        The persistence-motif participation relationships were strongest in networks with low connectance (Fig.~\ref{fig:density_prop}).
+        The relationships between persistence and participation in the other motifs were more consistent over different levels of connectance and network size but tended to be strongest in less-connected networks (Fig.~\ref{fig:density_prop}).
         For the apparent competition and three-species chain motifs, the direction of these relationships depended much more strongly on the strength of disturbance than on connectance.
@@
 
-        To understand the relationships between motif participation and consumers' probabilities of persistence, it is helpful to understand how persistence varies with other aspects of both consumer properties and network structure.
-
        Consumers with higher trophic level (STL) were less likely to persist (Fig.~\ref{fig:lm_CS}, Table S3, \emph{Appendix S6.1}).
-        The relationship between in-degree and persistence depended on the strength of disturbance: higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.
+        The relationship between in-degree and persistence depended on the strength of disturbance.
+        Higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.
 
         The average persistence of all consumers was lower in networks with higher connectance, did not vary significantly with network size (Table S4, \emph{Appendix S6.1}), and  varied with the frequencies of different motifs in the network's motif profile.
-        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and was not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:lm_CS}, Table S5, \emph{Appendix S6.2}).
+        Specifically, average consumer persistence was higher in networks with higher proportions of the apparent competition motif, lower in networks with higher proportions of the omnivory motif, and not significantly related to the frequencies of the three-species chain or direct competition motifs (Fig.~\ref{fig:lm_CS}, Table S5, \emph{Appendix S6.2}).
         There was no significant interaction between the proportion of any motif in the network motif profile and level of disturbance.
@@
 
-    Identifying the species that are most and least likely to persist after a disturbance is a high priority for conservation efforts~\citep{Bottrilletal2008}. 
-    Network structural properties such as connectance and measures of a species' position in the network such as degree and trophic level are known to affect species' vulnerability~\citep{Dunne2002, Eklof2006,bascompte2005simple,dunne2009cascading}, but are relatively coarse metrics.
-    Defining a species' network role based on its participation in meso-scale motifs provides more nuanced information and may also predict extinction risk~\citep{Stouffer2012,Simmons2019,Cirtwill2018FoodWebs}, but this potential is only beginning to be explored~\citep{Cirtwill2022Oikos}.
-
-    Here we partially fill this knowledge gap by testing how the persistence of a consumer in a food web depends on its participation in four different motifs, and how this relationship varies with different levels of disturbance to basal resources.
+    Here we partially fill the knowledge gap surrounding relationships between a species' position in its food web and its ability to persist following disturbance by testing how the persistence of a consumer in a food web depends on its participation in four different motifs, and how this relationship varies with different levels of disturbance to basal resources.
     We found that motif participation was related to a consumer's persistence in ways not fully explained by simpler measures of network structure. 
@@
 
-    % Apparent competition as a case study that motifs are useful
     The relationship between persistence and participation in the apparent competition motif provides a good example of how motifs provide complementary information to simpler measures of a species' network role.
@@
     We would therefore expect high participation in the apparent competition motif to also be associated with lower persistence.
-    However, this is only true for low disturbance. 
-    For high disturbance, we see the opposite relationship.
-    This change in the persistence-apparent competition relationship across disturbance levels may be driven by the negative relationship between apparent competition and in-degree, and in-degree's negative relationship with persistence.
+    However, this is only true for low disturbance; at high disturbance the opposite relationship holds.
+    This change across disturbance levels may be driven by the negative relationship between apparent competition and in-degree and in-degree's negative relationship with persistence.
     However, it is not obvious when trophic level should be the most important predictor of persistence and when in-degree should be most important.
-    Such simpler measures of species' positions in a network could therefore lead us astray either because they provide conflicting predictions or because our expectations are only reasonable for specific types or levels of disturbance.
+    Such simpler measures of species' positions in a network could therefore be misleading either because they provide conflicting predictions or because our expectations are only reasonable for specific types or levels of disturbance.
 
-    Omnivory motif participation likewise provides information about persistence that is not easily obtained from in-degree and trophic level.
+    Participation in the omnivory motif likewise provides information about persistence that is not easily obtained from in-degree and trophic level.
     Although omnivory participation decreased with trophic level (as does persistence) and increased with in-degree (as does persistence, except at high disturbance), we found that greater omnivory participation was usually associated with \emph{lower} persistence-- contrary to what we would expect based on trophic level and in-degree.
-    The exception to this pattern is found at high connectance, where increased omnivory participation was more likely to have a positive or neutral relationship with consumer persistence, especially at low disturbance levels (Fig.
-3).
-    This is consistent with other studies which have found the effect of omnivory to be highly context-dependent~\citep{bascompte2005simple, Monteiro2016} and particularly beneficial when interaction strengths are weak~\citep{Emmerson2004}.
-    Although omnivory motifs tend to be less stable in isolation the the other motifs we consider here~\citep{Borrelli2015a}, when embedded in a network interactions with other species can stabilise even intrinsically unstable omnivory~\citep{Kratina2012}.
-    Though we do not explicitly model interaction strengths, we do note that the greater number of interactions per species in high-connectance networks should lead to weaker interactions in general, as well as allowing many opportunities for additional interactions that could stabilise omnivory.
+    The exception to this pattern is found at high connectance, where increased omnivory participation was more likely to have a positive or neutral relationship with consumer persistence, especially at low disturbance levels (Fig.~\ref{fig:density_prop}).
+    This is consistent with other studies showing context-dependent effects of omnivory~\citep{bascompte2005simple, Monteiro2016} and particularly beneficial effects when interaction strengths are weak (as will generally be the case at high connectance)~\citep{Emmerson2004}.
+    Although omnivory motifs tend to be less stable in isolation the the other motifs we consider here~\citep{Borrelli2015a}, interactions with other species can stabilise even intrinsically unstable omnivory when embedded in a network~\citep{Kratina2012}.
+    As well as leading to weaker average interaction strengths, a higher connectance offers many opportunities for additional interactions that could stabilise unstable omnivory.
 
-    % Caveats - expand on top-down effects a bit to make R2 happy
     There are limitations to the approach presented here.
@@
     If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
-    For example, although a recent study using a dynamical model to relate motif participation to persistence after a disturbance found similar trends for omnivory, participation in the direct competition motif was not universally positive~\citep{Cirtwill2021_inprep}.
+    For example, although a recent study using a dynamical model to relate motif participation to persistence after a disturbance found similar trends for omnivory, participation in the direct competition motif was not universally positive~\citep{Cirtwill2022Oikos}.
     In particular, species which frequently appeared in the prey position of the direct competition motif tended to have shorter persistence times after disturbance in that study - possibly due to top-down effects from the pair of predators involved.
     However, as these two studies applied different types of disturbances, further study would be required to disentangle the effects of top-down and bottom-up effects on species with different motif participation, and how each relates to the type and strength of disturbance.
-    Another potentially useful addition to future work would be to include species traits such as body size, which are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017,Cardillo2005, Purvis2000} in empirical webs.
+    Another potentially useful addition to future work would be to include species traits such as body size, which are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017,Cardillo2005,Purvis2000} in empirical webs.
 
@@
     This context-dependence means that even limited changes in network structure or the type of disturbance could change which species are most sensitive to disturbance. 
-    As such, caution and increased attention to both network structure and the specifics of disturbance are needed for targetted conservation actions to be most effective.
+    As such, caution and increased attention to both network structure and the specifics of disturbance are needed for targeted conservation actions to be most effective.
 
@@
 
+% \clearpage
+
+\bibliographystyle{jae} 
+\bibliography{anna_bib_new}
+
 \clearpage
 
-\section*{Figures}
+\section*{Figure legends}
 
@@
     \end{figure}
-
-    %     \begin{figure}[hb!]
-    %     \centering
-    %     \includegraphics[width=.9\textwidth]{figures/concept_fig_ver2_edit.eps}
-    %     \caption{Note motif participation vectors in the right-hand column of the box. I still think it's nice to show - can you add back into current Fig. 1?}
-    % \label{fig:concept_wrongcolour}
-    % \end{figure}
 
@@
 
-    % \begin{figure}[hb!]
-    %     \centering
-    %     \includegraphics[width=.9\textwidth]{figures/participation_vs_SC.eps}
-    %     \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).
-We show relationships for the smallest and largest values of network size (S) and connectance (C).}
-    %     \label{fig:roles_vs_SC}
-    % \end{figure}
-
-    % \begin{figure}[hb!]
-    %     \centering
-    %     \includegraphics[width=\textwidth]{figures/persistence_motif_profiles.eps}
-    %     \caption{The proportions of the four motifs in a network's motif profile are related to the mean persistence of species in the network.
-Specifically, persistence decreases as the proportion of omnivory in the network's motif profile increases while persistence increases with the proportion of the apparent competition motifs.
-Interactions between motif profiles and disturbance were not significant.
-Here we show the relationships of probabilities of extinction of basal resources of $\pi=0.1$ (top set of lines) and $\pi=0.5$ (bottom set of lines).
-Relationships are shown for the observed range of proportions for each motif.}
-    %     \label{fig:motif_profile_persistence}
-    % \end{figure}
-
-    % \begin{figure}[hb!]
-    %     \centering
-    %     \includegraphics[width=\textwidth]{figures/motif_proportion_lms.eps}
-    %     \caption{The proportion of the omnivory in a network's motif role increased significantly as connectance increased, while the proportions of the other three motifs did not vary significantly with network size, connectance, or their interaction.}
-    %     \label{motif_proportion_lms}
-    % \end{figure}
-
-\clearpage
-
-% \section*{Glossary}
-% \begin{table}[h!]
-% \label{glossary}
-% \caption{Glossary of terms relating to motifs and Bayesian networks}
-%     \footnotesize{
-% \begin{tabular}{l|l}
-%     Term & Definition \\
-%     \hline
-%     Motif & Set of $n$ interacting species. In this case, $n=3$ \\
-%     Network motif profile & Vector describing the frequency of each motif in the network. \\
-%     & Normalised by dividing counts of each motif by the total across all motifs.\\
-%     Motif participation role & Vector describing the frequency with which a focal species appears in each motif.\\
-%     & Normalised by dividing counts for each motif by the total across all motifs. \\
-%     Bayesian network & A directed acyclic graph, used to predict species' likelihood of persistence. \\
-%     Network persistence & The mean likelihood of consumers in a network not going extinct.\\
-%     Species persistence & The individual likelihood of a species not going extinct.\\
-%     In-degree & Number of prey species to a consumer species.\\
-%     Trophic level (STL) & The shortest food chain between the focal species and any basal species.\\
-%     Disturbance ($\pi_{disturbed}$) & Probability of extinction of a basal resource when extra disturbance is added. \\
-%     Baseline extinction &  When no disturbance is applied, $\pi_{base}$=0.1.\\
-% \end{tabular}}
-% \end{table}
-
-% Notes on what expressions and words to use
-% "probability of extinction" or "level of disturbance" for the different disturbace scenarios?
-
-\bibliographystyle{jae} 
-\bibliography{anna_bib_new} % Abbreviate journal titles.
-
 \end{spacing}

### 2023-06-06 manuscript/JAE_reply.tex
Updates from Overleaf

@@
 
-\section*{Reply to Reviewer \#1} [[Done?]]
+\section*{Reply to Reviewer \#1}
 
@@
 
-                \textbf{R:} The sentence the reviewer refers to is ``How disturbances such as population decline or extinction lead to the loss of species via  direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research (Curtsdotter et al., 2011a; Dunne & Williams, 2009; Eklöf \& Ebenman, 2006 Santos et al., 2021).'' We can understand the reviewer´s confusion, but in fact species extinctions are both the disturbance and the response.
-                An initial extinction of a species (i.e.
+                \textbf{R:} An initial extinction of a species (i.e.
 the primary extinction) can, since species are connected to each other via  consumer--resource interactions, cause other species to go extinct (i.e.
@@
 We hope it is now easier to understand motifs as a meso-scale descriptor of network structure.
-See line XX, quoted below, for a short description.
+See line XX, quoted below, for a short description of motifs and definition of meso-scale structure.
 We have also, as described above, explained how motif participation captures information about a focal species' interaction partners, and thereby the potential for indirect interactions.
@@
                 \begin{quotation}
-                Therefore, a species' motif participation profile captures the indirect interactions the species are affected by, thereby capturing important structures that are missed by global (network level) and local (species level) properties.
-As such, motif participation may complement degree and trophic level to more fully predict species' extinction risk.
+                    Motifs are unique patterns of interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale (i.e., intermediate between overall and fine-scale) network structure~\citep{Stouffer2007,Stouffer2012}.
                 \end{quotation}
@@
 For example, networks with a high degree of omnivory may be more stable than simple food chains, but this depends on the strength of the omnivorous interactions among trophic guilds.
-[[Did they take this from our discussion, the point they found contradictory?]] It would be helpful to provide more insight into the gaps in knowledge surrounding the prevalence of these motifs in food webs more generally, and why this would help food web researchers understand consumer persistence and how it is related to resilience/stability since this is where I see the greatest amount of novelty from this article.
+It would be helpful to provide more insight into the gaps in knowledge surrounding the prevalence of these motifs in food webs more generally, and why this would help food web researchers understand consumer persistence and how it is related to resilience/stability since this is where I see the greatest amount of novelty from this article.
 
-                \textbf{R:} Every study we are aware of that relates motifs to stability either models one or more motifs in isolation or examines the effect of motif profiles (or elements thereof, e.g., frequency of omnivory) on the whole network.
+                \textbf{R:} We have expanded our description of existing knowledge about motifs but have kept this section relatively brief for several reasons.
+First, nearly every study we are aware of that relates motifs to stability either models one or more motifs in isolation or examines the effect of motif profiles (or elements thereof, e.g., frequency of omnivory) on the whole network.
 Neither one is exactly applicable to the effect of motif participation on per-species extinction risk.
-The exception is an earlier study by some authors of the present work, which we mention but do not wish to over-promote by making this paper a keystone of the introduction.
-For these reasons, together with constraints of length, we have been judicious in our selection of additional material on previous motif-stability relationships.
-Particularly for omnivory, the existing literature paints a picture of context-dependence and variable effects rather than a clear general trend.
-Reviewing this material would be more suited to a general review (perhaps with supplemental modelling to fill gaps between tests of different types of stability, disturbance, etc.) than to a paragraph in our introduction.
+One exception is an earlier study by some authors of the present work, which we mention but do not wish to over-promote by making this paper a keystone of the introduction.
+Second, in order to place more of the methods in the main text while complying with length restrictions we must keep the introduction brief.
                 We hope that the refocusing and revision of the introduction provides enough clarity to compensate for this necessarily cursory overview.
@@
                 \textbf{R:} We do think that our statement here is valid.
-Topological models for simulating species extinctions are only taking the food web structure in to account, and a secondary extinction will occur when a consumer lost all its resources.
-So, in comparison to topological models simulating the biomass (or similar) dynamics between many species will require both more parameters and simulation time.
-If we compare to for example the ATN-model requires parameters describing each species body mass and several parameters for defining for example attack rates and handling times.
-Similarly, models like EcoPath with EcoSim, is limited to smaller food webs with well-understood interaction strengths and initial population sizes and as such require more parameters which even in simple cases are more complex than topological models.
-However, these more complex models can model more complex mechanisms than what topological models could.
-
-                %We are not aware of any dynamical model that is fast over a large food web - certainly this was not the case when some of the authors conducted dynamical simulations of webs with 50-100 species.
-[[Anyone have a better answer?
-From the focus on major energy flows I think it's EcoSim they're thinking of...]]
+In large food webs, there are very many parameters affecting the population dynamics of interacting species.
+While in empirical systems it may be possible to ignore many of these parameters and focus on a small set of `key' parameters as you suggest, as in models like EcoPath with EcoSim, these are limited to smaller food webs with well-understood interaction strengths and initial population sizes.
+                In simulated networks, selection of such key parameters would either be highly arbitrary (and potentially strongly affecting model results) or require an extensive (and computationally costly) sensitivity analysis.
+                On the other hand, the ATN-model requires parameters describing each species' body mass and several parameters for defining for example attack rates and handling times. 
+                This is many more than either topological or Bayesian network models and leads to a much longer computation time.
+                Note that we do not say that dynamical models are not useful (indeed, we have used them ourselves in different contexts), only that the computation time required for large networks prevents us from exploring multiple disturbance scenarios.
 
@@
                 \textbf{R:} Please note that we are using a Bayesian network modelling framework and not Bayesian statistics.
-All statistical models are quite standard linear models or general linear models.
+All statistical models are  standard linear models or general linear models.
 The Bayesian network method has been used and validated in previous papers (see references in the main text) and has not been substantially modified for this paper.
@@
                 We generated simulated networks using the niche model, which has been shown to recreate the structure of empirical networks well~\citep{Williams2000,Stouffer2007}.
-                To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002a}, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
+                To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002e}, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
                 All networks were generated using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.   
@@
 These frequencies are equivalent to the proportion of a species' total motif participation made up by each motif.
-We also now use the perminology ``proportion of motif in role'' and ``proportion of motif in network profile'' in the axis labels of Figs.
+We also now use the terminology ``proportion of motif in role'' and ``proportion of motif in network profile'' in the axis labels of Figs.
 2, 3, and 5 for consistency.
@@
 However, in a parallel study (started around the same time but by chance published earlier) some of the authors of the present study did simulate both top-down and bottom-up effects in a dynamical model.
-The dynamical model has its own caveats (in general, dynamical models are sensative to parameter choice and, due to the time and computational resources required, a full sensitivity analysis could not be performed).
-However, [[something comparing the results between papers - somewhat similar?]]
+The dynamical model has its own caveats (in general, dynamical models are sensitive to parameter choice and, due to the time and computational resources required, a full sensitivity analysis could not be performed).
+However, we note that a dynamical analysis on the same set of networks (albeit with a different disturbance and persistence metric) found substantial consistency in motif-persistence relationships.
 
-                It may also reassure the Reviewer to know that in the initial studies introducing the Bayesian network framework~\citep{}, Bayesian networks captured XX\% of the extinctions identified by dynamical models run on the same set of networks.
+                It may also reassure the Reviewer to know that in the initial studies introducing the Bayesian network framework~\citep{Eklof2013}, Bayesian networks captured the majority of the extinctions identified by dynamical models run on the same set of networks.
 As we use the same modelling framework as in those studies, we can reasonably expect similar results.
@@
                     If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
-                    For example, although a recent study using a dynamical model to relate motif participation to persistence after a disturbance found similar trends for omnivory, participation in the direct competition motif was not universally positive~\citep{Cirtwill2021_inprep}.
+                    For example, although a recent study using a dynamical model to relate motif participation to persistence after a disturbance found similar trends for omnivory, participation in the direct competition motif was not universally positive~\citep{Cirtwill2022Oikos}.
                     In particular, species which frequently appeared in the prey position of the direct competition motif tended to have shorter persistence times after disturbance in that study - possibly due to top-down effects from the pair of predators involved.
@@
 
-    \bibliographystyle{ecollett} 
-    \bibliography{manual} % Do not abbreviate journal titles, papers with >
+    \bibliographystyle{jae} 
+    \bibliography{anna_bib_new} % Do not abbreviate journal titles, papers with >
 

### 2023-06-06 manuscript/JAE_reply.tex
syncing quotes

@@
         I feel that this study includes a lot of work and analysis that is meaningful.
-For example, at line 195-196, the authors state “This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.”.
+For example, at line 195-196, the authors state ``This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.''.
 To me, this sentence highlights how their results fit into an up-and-coming area of ecological research: monitoring the likelihood of consumer persistence in efficient ways.
@@
 It seems like it is both from the results presented, but the authors highlight the former extensively and leave the latter as more of an after-thought.
-This may be addressed by simply taking out words like “additionally” from the start of the last few summary points.
+This may be addressed by simply taking out words like ``additionally'' from the start of the last few summary points.
 
@@
 secondary extinctions).
-We have now clarified this on line XX:
+We have now clarified this on lines 2-5:
 
                 \begin{quotation}
-                The loss of a species in a food web can, due to the mutual dependencies among species, trigger a cascade of additional extinctions, so called secondary extinctions.
-A species' risk of going secondarily extinct is known to depend on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
+                A single extinction can, due to the mutual dependencies among species in a food web,  trigger a cascade of additional (secondary) extinctions. 
+                A species' risk of secondary extinction depends on its position within a food web~\citep{Santos2021,dunne2009cascading, Eklof2006}.
                 \end{quotation}
@@
              \textbf{R:}
-             We have now clarified how we in this manuscript define a disturbance. This is done on line XX.
+             We have now clarified how we in this manuscript define a disturbance. This is done on line 71 (quoted below) and in the summary.
 
              \begin{quotation}
-             Here, we focus on persistence after the loss of plants or other basal resources, i.e., the effects of bottom-up disturbances.
+             Here, we analyse whether a consumer's risk of secondary extinction after removal of basal resources depends on the consumer's motif participation and how this relationship varies with the severity of disturbance.
             \end{quotation}
@@
 
-                \textbf{R:} We agree with the reviewer that this was not so clearly expressed. We have now re-formulated this section and clarified with an example, beginning at line XX.
+                \textbf{R:} We agree with the reviewer that this was not so clearly expressed. We have now re-formulated this section and clarified with an example, beginning at line 21.
 
                 \begin{quotation}
-                However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the focal species' interaction partners.
-Two species might both have three prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
-Importantly, since specialists are generally more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than a predator of generalists.
+                    Importantly, motif participation incorporates information on the positions of focal species' interaction partners.
+                    Two species might have the same trophic level and in-degree but quite different motif participation if one species's prey are strict specialists and the others' prey are generalists (Fig.
+1).
+                    Since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
+                    This extra information on the focal species' interaction partners may thus provide extra insight when predicting extinction risk.
                 \end{quotation}
@@
             \item Line 48: This paragraph contains important information to understanding the concept of motifs; however, I found that the description of motif participation is vague and difficult to interpret in the context of this study.
-For example, the “second-step indirect interactions the species are affected by, thereby capturing important structures that are missed by global and local properties” does not provide any details that would help the reader understand why motif participation is a relevant metric.
+For example, the ``second-step indirect interactions the species are affected by, thereby capturing important structures that are missed by global and local properties'' does not provide any details that would help the reader understand why motif participation is a relevant metric.
 Specifically, what important structures are the authors referring to here, and how are they captured by the second-step indirect interactions?
@@
 We hope it is now easier to understand motifs as a meso-scale descriptor of network structure.
-See line XX, quoted below, for a short description of motifs and definition of meso-scale structure.
+See line 13, quoted below, for a short description of motifs and definition of meso-scale structure.
 We have also, as described above, explained how motif participation captures information about a focal species' interaction partners, and thereby the potential for indirect interactions.
@@
                 \begin{quotation}
-                We generated simulated networks using the niche model, which has been shown to recreate the structure of empirical networks well~\citep{Williams2000,Stouffer2007}.
-                To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002e}, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04).
+                Simulated networks were generated using the niche model, which recreates the structure of empirical networks well~\citep{Williams2000,Stouffer2007}.
+                To capture a range of network architectures similar to those in well-studied empirical networks~\citep{Dunne2002,Dunne2002e}, networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04) were simulated.
+                Biologically unlikely networks (i.e., those with extremely long paths between any consumer and basal resources~\citep{Borrelli2014} were removed and replaced with new simulated networks until 100 suitable networks in each combination of size and connectance were obtained.
                 All networks were generated using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}.     
-                For full details and a comparison to empirical networks, see \emph{Appendix S2}.
+                For full details and a comparison to empirical networks, see \emph{Appendix S1}.
                 \end{quotation}
@@
                 \begin{quotation}
-                A species' ``motif participation role'' is the frequency with which that species appears in each of the motifs present in a network~\citep{Stouffer2012}.
+                A species' motif participation (Fig. 1) is the vector of frequencies with which it appears in each type of motif present in a network~\citep{Stouffer2012}.
                 \end{quotation}
@@
                 \textbf{R:} The size of a motif participation vector is always related to the number of motifs considered, not the number of species (there is one vector per species).
-We only considered three-species motifs in this study, as stated in lines XX, XX, and XX.
+We only considered three-species motifs in this study, as stated in the summary and lines 28 and 95.
 Since there are only four \emph{unique} three-species motifs in these acyclic networks (a general feature of all acyclic networks), motif participation vectors are four-dimensional for all species.
@@
 
-                We have rephrased this line (quoted below) to remove the phrase `in our case' and make it clearer that four-dimensional motif participation vectors are a general feature of acyclic networks.
+                We have rephrased line 95 (quoted below) to remove the phrase `in our case' and make it clearer that four-dimensional motif participation vectors are a general feature of acyclic networks.
                 \begin{quotation}
@@
                 \begin{quotation}
-
     There are limitations to the approach presented here.
@@
     However, as these two studies applied different types of disturbances, further study would be required to disentangle the effects of top-down and bottom-up effects on species with different motif participation, and how each relates to the type and strength of disturbance.
+    Another potentially useful addition to future work would be to include species traits such as body size, which are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017,Cardillo2005,Purvis2000} in empirical webs.
 

### 2023-06-06 code/plot_roles_vs_TL.py

@@ -142,8 +142,8 @@ def populate_graph(graph,minidict,simple,roletype):
       data.legend=motif
 
   if simple=='TL':
-    graph.add_drawing_object(DrawText,text='Motif',x=5.25,y=0.47,loctype='world',char_size=.5) 
-    graph.legend.configure(char_size=.5,loc=(5.25,0.450),loctype='world',box_linestyle=0,fill_pattern=0)
+    graph.add_drawing_object(DrawText,text='Motif',x=3.15,y=-.21,loctype='world',char_size=.5) 
+    graph.legend.configure(char_size=.5,loc=(3,-.22),loctype='world',box_linestyle=0,fill_pattern=0)
 
   return graph
 

### 2023-06-06 code/stat_analysis/persistence_vs_SC_lm.py

@@ -329,12 +329,12 @@ for simple in ['TL','Connectance','Motifs','Deg','Size','dummy']:
 
 grace.multi(rows=2,cols=3,vgap=.09,hgap=.09)
 grace.hide_redundant_labels()
-grace.graphs[0].set_view(0.14, 0.7367058823529411, 0.38, 0.95)
-grace.graphs[1].set_view(0.46, 0.7367058823529411, 0.7, 0.95)
-grace.graphs[2].set_view(0.74, 0.7367058823529411, 0.98, 0.95)
-grace.graphs[3].set_view(0.14, 0.4334117647058823, 0.38, 0.6467058823529411)
-grace.graphs[4].set_view(0.46, 0.4334117647058823, 0.7, 0.6467058823529411)
-grace.graphs[5].set_view(0.74, 0.4334117647058823, .98, 0.6467058823529411)
+grace.graphs[0].set_view(0.14, 0.7367058823529411, 0.37, 0.95)
+grace.graphs[1].set_view(0.46, 0.7367058823529411, 0.69, 0.95)
+grace.graphs[2].set_view(0.72, 0.7367058823529411, 0.95, 0.95)
+grace.graphs[3].set_view(0.14, 0.4334117647058823, 0.37, 0.6467058823529411)
+grace.graphs[4].set_view(0.46, 0.4334117647058823, 0.69, 0.6467058823529411)
+grace.graphs[5].set_view(0.72, 0.4334117647058823, .95, 0.6467058823529411)
 
 
 grace.set_col_yaxislabel(label='Mean consumer probability of persistence',col=1,rowspan=(None,None),char_size=1,perpendicular_offset=.03)

### 2023-06-06 manuscript/JAE_reply.tex
Reply ready for JAE

@@
     \textbf{R:} We thank the Reviewer for taking the time to read the manuscript and for their thoughtful feedback.
-We have redone our analyses with binomial error distribution as suggested (though with glm rather than glmer as the random effects proved singular).
+We have redone our analyses with binomial error distribution as suggested (though usually with glm rather than glmer as the random effects proved singular).
 Our main results did not change much with this improved approach, though some of our supplemental results (e.g., those relating to overall network structure) were simplified.
@@
     \textbf{R:} We thank the Reviewer for taking the time to read our manuscript and for the thoughtful feedback.
-We particularly appreciate the pointers where the structure of our manuscript was unclear an where it was difficult to extract our main aims and results.
+We particularly appreciate the pointers where the structure of our manuscript was unclear and where it was difficult to extract our main aims and results.
 In the present revision we have substantially reworked the introduction and discussion and reordered the methods and results in order to present our work in a more streamlined and, hopefully, understandable manner.
@@
 
-                \textbf{R:} A focal species is simply that for which the motif role is currently being calculated ...
-all consumers are taken as the focal species in turn.
-Note that as these networks are simulated the species (nodes) in them do not correspond to any particular focal species or trophic guild.
-The vagueness is intentional as there are no characteristics of, say, invertebrates vs.
+                \textbf{R:} A focal species is simply that for which the motif role is currently being calculated.
+All consumers are taken as the focal species in turn.
+Note that as these networks are simulated the species (nodes) in them do not correspond to any particular real species or trophic guilds.
+The vagueness about species identity is intentional as there are no characteristics of, say, invertebrates vs.
 vertebrates or mammals vs.
@@
 Our results, which deal specifically with persistence after bottom-up disturbance, add to that picture of complexity without contradicting earlier findings.
-Note that we do not say that omnivory is never beneficial - only that in our networks, in response to our focal disturbance, is is generally detrimental.
-We also note that our results show apparent competition as associated with less persistence when disturbance is weak, which is compatible with the possibility that the Reviewer mentions above (though we do not specifically model prey switching, which seems to be what the Reviewer refers to).
+Note that we do not say that omnivory is never beneficial (indeed, it is generally associated with higher persistence in some highly-connected webs, as shown in Fig.
+3).
+However, in our networks overall, in response to our focal disturbance, omnivory is generally detrimental.
+We also note that our results show apparent competition as associated with less persistence when disturbance is weak, which is compatible with the possibility that the Reviewer mentions above (though we do not specifically model prey switching, which seems to be what the Reviewer refers to, this is included in the ability of a consumer to persist on a subset of its prey).
 In the revision, we have expanded our discussion on how the strength and direction of motif participation-persistence relationships vary network-to-network (especially in networks with different connectance).
@@
 However, in a parallel study (started around the same time but by chance published earlier) some of the authors of the present study did simulate both top-down and bottom-up effects in a dynamical model.
-The dynamical model has its own caveats (in general, dynamical models are sensitive to parameter choice and, due to the time and computational resources required, a full sensitivity analysis could not be performed).
-However, we note that a dynamical analysis on the same set of networks (albeit with a different disturbance and persistence metric) found substantial consistency in motif-persistence relationships.
-
+Although that study has its own caveats (in general, dynamical models are sensitive to parameter choice and, due to the time and computational resources required, a full sensitivity analysis could not be performed), we refer readers interested in combined top-down and bottom-up effects there.
                 It may also reassure the Reviewer to know that in the initial studies introducing the Bayesian network framework~\citep{Eklof2013}, Bayesian networks captured the majority of the extinctions identified by dynamical models run on the same set of networks.
@@
 While this obviously does occur in nature, logically this is a smaller number of extinctions than the set of herbivore and predator extinctions caused by bottom-up effects.
-This is also reflected in the similar results from Bayesian networks and dynammical models in the studies cited above - if top-down effects were responsible for the majority of secondary extinctions we would see quite different performance.
+This is also reflected in the similar results from Bayesian networks and dynamical models in the studies cited above - if top-down effects were responsible for the majority of secondary extinctions we would see quite different performance.
 
@@
 However, since the Bayesian network framework does not include top-down effects any comment we might make about how top-down effects might affect our results would necessarily be pure speculation.
-Instead, we have added a reference to another paper (quoted below) exploring the relationship between motif roles and persistence in a dynamical context.
-That paper, due to the computational limits inherent in the dynamical approach, uses only a single-species disturbance.
+Instead, we have added a reference to our dynamical paper (quoted below).
+That paper, due to the computational limits inherent in the dynamical approach, uses only a single-species disturbance and a different estimate of persistence.
 It does, however, include top-down as well as bottom-up effects and may provide the information the Reviewer is seeking.
@@
                     Another potentially useful addition to future work would be to include species traits such as body size, which are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017,Cardillo2005,Purvis2000} in empirical webs.
-
                 \end{quotation}

### 2024-02-16 manuscript/JAE_reply_2.tex
tiny revisions

\documentclass[12pt]{article} 
\usepackage{amsmath} 
\usepackage[dvips]{graphicx}
\usepackage{multirow} 
\usepackage{geometry} 
\usepackage{pdflscape}
\usepackage[labelfont=bf]{caption} 
\usepackage{setspace}
\usepackage[running]{lineno} 
% \usepackage[numbers,sort]{natbib}
\usepackage[round]{natbib} 
\usepackage{array}
\usepackage[table]{xcolor}

\topmargin -1.5cm % 0.0cm 
\oddsidemargin 0.0cm % 0.2cm 
\textwidth 6.5in
\textheight 9.0in % 21cmhttps://www.overleaf.com/project/6048001a3ac22e3aff2ced4a
\footskip 1.0cm % 1.0cm

\newenvironment{refquote}{\bigskip \begin{it}}{\end{it}\medskip}

\usepackage{authblk}

\begin{document}

% Due June 18.

\section*{Reply to Editor}

    \begin{refquote}

    Dear Dr Alyssa Cirtwill

    Re: Species motif participation provides unique information about species risk of extinction.  Cirtwill, Alyssa; Åkesson, Anna; Wootton, Katherine; Eklöf, Anna.

    Thank you for submitting your manuscript to Journal of Animal Ecology.

    I have now received reviewers' reports and the Associate Editor's comments on your manuscript and looked at it myself. As you can see from their comments below, a few final revisions are required.

    I suspect you can deal with these minor revisions fairly easily, and I look forward to seeing your revised manuscript by 17-Mar-2024. If you foresee any problems with meeting this deadline, please contact the editorial office on: admin@journalofanimalecology.org

    * Please respond to all the comments of the reviewers and Associate Editor, which are set out below, by changes in the manuscript and point by point in your response.  Please ensure you insert your response in the correct 'Response to decision letter' section when you submit your revision online. Please ensure that you complete this thoroughly so that it is clear to the editors and reviewers how you have improved your manuscript. Please insert your point by point response in the correct section of the submission process in the Authors Response to Comments text box.  Please bear in mind that in the Author Response to Comments text box has limited formatting (no italics/ bold), so please distinguish your answers clearly.

    *Please upload a Word document for use by our Publisher. PDF files are not acceptable.

    Yours sincerely
    Professor Darren Evans
    Editor, Journal of Animal Ecology

    \end{refquote}

    \textbf{R:}

\clearpage

\section*{Reply to Associate Editor}

    \begin{refquote}
    Dear Alyssa Cirtwill et al,

    The revised manuscript has been seen by a reviewer and myself and we agree that all the comments have been addressed correctly. However, there a few minor points left to address.

    Best wishes,
    Daniel
    \end{refquote}

    - The Supplementary material is not uploaded, maybe due to some problems or errors during the submission process?
    \textbf{R:} We have now uploaded a corrected Supplementary material file and apologise for the error in the previous submission. We are not sure how that occurred.
    - Line 68: ‘when’ is written twice
    \textbf{R:} Corrected.
    - Line 358-360: This sentence is not clear. Please rewrite
    \textbf{R:} We have revised the sentence (as well as the preceding and following sentences) and hope that the end of this paragraph is now clearer.

    Revised sentence:
    \begin{quotation}
    This is consistent with other studies showing context-dependent effects of omnivory~\citep{Bascompte2005,Monteiro2016}.
    In highly-connected networks, species in omnivory motifs are very likely to interact with other species in the network (outside the omnivory motif).
    These `external' interactions can stabilise omnivory motifs that would be unstable in isolation~\citep{Kratina2012}, perhaps mitigating the risk of participating in omnivory rather than motifs that are more stable in isolation~\citep{Borelli2015}.
    Moreover, omnivory tends to be particularly beneficial when interaction strengths are weak, which is generally the case in highly-connected networks~\citep{Emmerson2004}.
    These earlier findings suggest that high connectance reduces the risk and increases the benefit of participating in omnivory motifs, consistent with our simulation results.

    \end{quotation}

    Original lines 354-362:

    \begin{quotation}
    This is consistent with other studies showing context-dependent effects of omnivory~\citep{Bascompte2005,Monteiro2016} and particularly beneficial effects when interaction strengths are weak (as will generally be the case at high connectance)~\citep{Emmerson2004}. Although omnivory motifs tend to be less stable in isolation the the other motifs we consider here~\citep{Borrelli2015}, interactions with other species can stabilise even intrinsically unstable omnivory when embedded in a network~\citep{Kratina2012}. As well as leading to weaker average interaction strengths, a higher connectance offers many opportunities for additional interactions that could stabilise unstable omnivory.
    \end{quotation}

\clearpage

\section*{Reply to reviewer 1}

    \begin{refquote}

        This revision is very well done, and the authors responses to reviewer comments are comprehensive. The changes made by the authors have greatly improved the clarity of the manuscript. The new Fig 4 (now Fig 3) offers a much better means of comparing the distribution of slopes across disturbance and connectance levels within each motif type.

        In the uploaded documents it seems like the supplement "motif_participation_SI.pdf" contains  another version of the paper rather than the Appendix, and I did not see the supplementary materials in any of the files.

    \end{refquote}

    \textbf{R:} We thank the Reviewer for their careful comments on the previous submission which greatly improved the manuscript. We have double-checked the Appendix and uploaded the correct version and apologise for the error.

\clearpage

\clearpage

    \bibliographystyle{jae} 
    \bibliography{anna_bib_new} % Do not abbreviate journal titles, papers with >

\end{document}

### 2024-02-16 manuscript/JAE_resub1.tex
tiny revisions

--- 2023-06-06 manuscript/reorder_main_text.tex
@@
 \clearpage
-\begin{spacing}{1.0}
-% \linenumbers
+\begin{spacing}{2.0}
+\linenumbers
 
@@
     These species are more vulnerable because they are impacted by the loss of any species along a long path from basal resources or because they have few alternative resources to replace lost prey.
-    The relationships between these relatively simple measures of a species' position in a food web suggest that more detailed measures, such as a species' participation in \emph{meso-scale motifs} are also related to its persistence~\citep{Cirtwill2022Oikos}.
+    The relationships between these relatively simple measures of a species' position in a food web suggest that more detailed measures, such as a species' participation in \emph{motifs} are also related to its persistence~\citep{Cirtwill2022Oikos}.
 
-    Motifs are unique patterns of interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale network structure~\citep{Stouffer2007,Stouffer2012}.
+    Motifs are unique patterns of interacting species that correspond to different sets of direct and indirect interactions, providing a holistic description of meso-scale (i.e., intermediate between overall and fine-scale) network structure~\citep{Stouffer2007,Stouffer2012}.
     Some motifs are over-represented in stable food webs, suggesting that the more frequently species participate in these motifs, the more stable the network will be~\citep{Borrelli2015a}. 

### 2024-02-16 manuscript/diff.tex
tiny revisions

diff --git a/reorder_main_text.tex b/../../../manuscript/reorder_main_text.tex
index 858fe9b..1cb25e3 100644
--- a/reorder_main_text.tex
+++ b/../../../manuscript/reorder_main_text.tex
@@ -379,21 +379,22 @@ Motif participation describes extinction risk
         Participation in the three-species chain motif showed the opposite trends.

-        Digging deeper and exploring how persistence varied with the proportion of a motif in the species' participation vector, disturbance, and overall network structure (size and connectance), we found that the relationship between the proportion of omnivory in the participation vector and probability of persistence depended strongly on the connectance of the network.
+        In depth exploration of how persistence varied with the proportion of a motif in the species' participation vector, disturbance, and overall network structure (size and connectance), showed that the relationship between the proportion of omnivory in the participation vector and probability of persistence depended strongly on the connectance of the network Fig.~\ref{Fig 2}.
         Higher proportions of omnivory were associated with lower probabilities of persistence when connectance was low, regardless of network size or strength of disturbance.
         At moderate connectance, however, higher proportions of omnivory were associated with greater probabilities of persistence when disturbance was weak.
-        In the most-connected networks, higher proportions of omnivory were associated with greater probabilities of persistence at all levels of disturbance.
+        In the most-connected networks, higher proportions of omnivory were associated with greater probabilities of persistence at all levels of disturbance.

-        The relationships between other motifs and persistence were more consistent over different levels of connectance and network size.
-        Slopes of the motif proportion-persistence relationships for the apparent competition and direct competition motifs were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{Fig 2}.
+        The relationships between the proportion of other motifs and persistence were more consistent over different levels of connectance and network size.
+        Slopes of the motif proportion-persistence relationships for the apparent competition and direct competition motifs were steepest in large and less-connected networks, but the directions of slopes varied only with the strength of disturbance, as in Fig.~\ref{Fig 2}.
         The relationship between proportion of the three-species chain in the participation vector and persistence was very similar across all levels of size and connectance.

+         The effects of network size were overall limited Fig.~\ref{Fig 2}.

         Considering each network separately ....

-
-    \subsection*{Species persistence and other network properties}
+% Total frequencies of motifs in the whole network
+    \subsection*{Species persistence and network motif profile}

         As well as the frequency with which each species appears in each motif, we can use the total frequencies of each motif in the whole network (i.e., the network's motif profile) to describe its structure.
         The average probability of persistence of all consumers in the network varied with the frequencies of different motifs in this profile.
@@ -544,7 +545,7 @@ Motif participation describes extinction risk

 \section*{Figures}

-
+        % FIGURE 1
         \begin{figure}[hb!]
         \centering
         \includegraphics[width=1.0\textwidth]{figures/concept_fig_ver2.eps}
@@ -554,7 +555,7 @@ Motif participation describes extinction risk
     \label{fig:concept}
     \end{figure}

-    
+        % FIGURE 2
         \begin{figure}[hb!]
         \centering
         \includegraphics[width=0.85\textwidth]{figures/persistence_motif_participation.eps}
@@ -562,7 +563,7 @@ Motif participation describes extinction risk
     \label{fig:prop_lmer_all}
     \end{figure}

-
+        % FIGURE 3
     \begin{figure}[hb!]
         \centering
         \includegraphics[width=\textwidth]{figures/roles_vs_TL.eps}
@@ -571,6 +572,7 @@ Motif participation describes extinction risk
         \label{fig:motifs_vs_TL_and_deg}
     \end{figure}

+        % FIGURE 4
     \begin{figure}[ht!]
         \centering
         \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
@@ -590,10 +592,10 @@ Motif participation describes extinction risk
     %     \label{fig:roles_vs_SC}
     % \end{figure}

-
+    % FIGURE 5
     \begin{figure}[hb!]
     \centering
-        \includegraphics[width=\textwidth]{figures/prop_dens_bp_vs_C_allS.eps}
+        \includegraphics[width=\textwidth]{manuscript/figures/Fig4.pdf}
         \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species. Columns show the result for various disturbances on the basal level, from $\pi_{disturbed} = 0.1$ (left) to $\pi_{disturbed} = 0.5$ (right). Rows show various levels of connectance. The dotted, vertical lines indicate zero on the x-axis. A negative slope value reflects a negative relationship between increased participation in a motif and persistence, while a positive slope value reflects a positive relationship - an increased proportion of a specific motif increases persistence. The fraction of replicates with a slope greater than zero are stated in numbers in each sub-plot, the color corresponding to each motif (legend). }
         \label{fig:density_prop_C}
     \end{figure}

### 2024-02-21 manuscript/JAE_reply_2.tex
tiny revision for JAE

@@
 
-    \textbf{R:}
+    \textbf{R:} Dear Professor Evans,\\
+
+    Thank you for your optimism about our manuscript and for your efforts coordinating the review process.
+In the current resubmission we have rephrased the sentence flagged by the Associate Editor as unclear, corrected our supporting information file, and converted the main text to .docx.
+Please note that the text was originally prepared using LaTeX and we do not have a paid subscription to Microsoft Office.
+In case of any formatting discrepancies (e.g., in equations), please refer to the self-compiled PDF file for the correct versions.
+Once again, thank you for your time and efforts.
+
+    Sincerely,
+    Alyssa Cirtwill (on behalf of all authors)
 
@@
 
+    \textbf{R:} Dear Daniel,
+
+    Thank you for your continued efforts to help us improve the manuscript.
+We agree that the sentence you flagged was not clear and have revised the second half of the relevant paragraph to rectify the situation.
+As line numbers have been changed during the required conversion to .docx, we quote both the original and revised sections below.
+We have also uploaded a correct supplementary material file.
+
+    Best regards,
+    Alyssa Cirtwill (on behalf of all authors)
+
     - The Supplementary material is not uploaded, maybe due to some problems or errors during the submission process?
@@
     In highly-connected networks, species in omnivory motifs are very likely to interact with other species in the network (outside the omnivory motif).
-    These `external' interactions can stabilise omnivory motifs that would be unstable in isolation~\citep{Kratina2012}, perhaps mitigating the risk of participating in omnivory rather than motifs that are more stable in isolation~\citep{Borelli2015}.
+    These `external' interactions can stabilise omnivory motifs that would be unstable in isolation~\citep{Kratina2012}, perhaps mitigating the risk of participating in omnivory rather than motifs that are more stable in isolation~\citep{Borrelli2015a}.
     Moreover, omnivory tends to be particularly beneficial when interaction strengths are weak, which is generally the case in highly-connected networks~\citep{Emmerson2004}.
@@
     This is consistent with other studies showing context-dependent effects of omnivory~\citep{Bascompte2005,Monteiro2016} and particularly beneficial effects when interaction strengths are weak (as will generally be the case at high connectance)~\citep{Emmerson2004}.
-Although omnivory motifs tend to be less stable in isolation the the other motifs we consider here~\citep{Borrelli2015}, interactions with other species can stabilise even intrinsically unstable omnivory when embedded in a network~\citep{Kratina2012}.
+Although omnivory motifs tend to be less stable in isolation the the other motifs we consider here~\citep{Borrelli2015a}, interactions with other species can stabilise even intrinsically unstable omnivory when embedded in a network~\citep{Kratina2012}.
 As well as leading to weaker average interaction strengths, a higher connectance offers many opportunities for additional interactions that could stabilise unstable omnivory.
@@
 
-        In the uploaded documents it seems like the supplement "motif_participation_SI.pdf" contains  another version of the paper rather than the Appendix, and I did not see the supplementary materials in any of the files.
+        In the uploaded documents it seems like the supplement "motif\_participation\_SI.pdf" contains  another version of the paper rather than the Appendix, and I did not see the supplementary materials in any of the files.
 

### 2024-02-21 manuscript/reorder_main_text.tex
tiny revision for JAE

@@
 \begin{document} 
-\maketitle 
+% \maketitle 
 \raggedright
@@
     As well as identifying patterns of motif participation that are associated with ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
-    For example, participation in one motif might be associated with greater extinction risk after a mild disturbance but be associated with less risk when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances while providing `back-up' resources when many prey are removed by a strong disturbance.
+    For example, participation in one motif might be associated with greater extinction risk after a mild disturbance but be associated with less risk when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances while providing `back-up' resources when many prey are removed by a strong disturbance.
 
@@
     The exception to this pattern is found at high connectance, where increased omnivory participation was more likely to have a positive or neutral relationship with consumer persistence, especially at low disturbance levels (Fig.~\ref{fig:density_prop}).
-    This is consistent with other studies showing context-dependent effects of omnivory~\citep{bascompte2005simple, Monteiro2016} and particularly beneficial effects when interaction strengths are weak (as will generally be the case at high connectance)~\citep{Emmerson2004}.
-    Although omnivory motifs tend to be less stable in isolation the the other motifs we consider here~\citep{Borrelli2015a}, interactions with other species can stabilise even intrinsically unstable omnivory when embedded in a network~\citep{Kratina2012}.
-    As well as leading to weaker average interaction strengths, a higher connectance offers many opportunities for additional interactions that could stabilise unstable omnivory.
+    This is consistent with other studies showing context-dependent effects of omnivory~\citep{Bascompte2005,Monteiro2016}.
+    In highly-connected networks, species in omnivory motifs are very likely to interact with other species in the network (outside the omnivory motif).
+    These `external' interactions can stabilise omnivory motifs that would be unstable in isolation~\citep{Kratina2012}, perhaps mitigating the risk of participating in omnivory rather than motifs that are more stable in isolation~\citep{Borrelli2015a}.
+    Moreover, omnivory tends to be particularly beneficial when interaction strengths are weak, which is generally the case in highly-connected networks~\citep{Emmerson2004}.
+    These earlier findings suggest that high connectance reduces the risk and increases the benefit of participating in omnivory motifs, consistent with our simulation results.
 
@@
 
-\section*{Acknowledgements}
+% \section*{Acknowledgements}
 
-    The authors thank Gy\"{o}rgy Barab\'{a}s for excellent discussion of the ideas in this manuscript.
+%     The authors thank Gy\"{o}rgy Barab\'{a}s for excellent discussion of the ideas in this manuscript.
 A{\AA} is supported by the Swedish Research Council for Sustainable Development Formas (\#2015-01262) granted to AE.
@@
         \centering
-        % \includegraphics[width=.9\textwidth]{figures/Figure_1.eps}
+        \includegraphics[width=.9\textwidth]{figures/Figure_1.eps}
         \caption{Conceptual figure of a toy food web.
@@
         \centering
-        % \includegraphics[width=0.95\textwidth]{figures/persistence_motif_participation.eps}
+        \includegraphics[width=0.95\textwidth]{figures/persistence_motif_participation.eps}
         \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on probability of persistence following disturbance to basal resources (y-axis).
@@
     \centering
-        % \includegraphics[width=\textwidth]{figures/Fig4.pdf}
+        \includegraphics[width=\textwidth]{figures/Fig4.pdf}
         \caption{Here we show the density (y-axis) of slopes (x-axis) of persistence against proportion of different motifs for all simulated webs of all sizes - a visualization of how an increased proportion of each motif (different colored lines) affects persistence of consumer species.

### 2024-02-21 code/stat_analysis/statistical_models.R

@@ -328,6 +328,13 @@ save.image('all_tests.Rdata')
     results[4,]=c("Direct",summary(direct_prop_norandom)$coefficients[,1])
     write.table(results,file='roles_vs_SC.tsv',sep='\t')
 
+    SEs=matrix(nrow=4,ncol=5)
+    SEs[1,]=c("Omnivory",summary(omni_prop_norandom)$coefficients[,2])
+    SEs[2,]=c("Chain",summary(chain_prop_norandom)$coefficients[,2])
+    SEs[3,]=c("Apparent",summary(apparent_prop_norandom)$coefficients[,2])
+    SEs[4,]=c("Direct",summary(direct_prop_norandom)$coefficients[,2])
+    write.table(SEs,file='roles_vs_SC_SE.tsv',sep='\t')
+
     # # Deg and TL
     # # relating counts and proportions to degree, TL (repeats analysis in Cirwill & Wootton, in prep.)
     # # # All singular.
